### 2018-04-09 Manuscript/manuscript.tex
Initial commit with all current topologies and cleaned Data and scripts folders. Manuscript template adopted from ECC-POPC work.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% This is a (brief) model paper using the achemso class
%% The document class accepts keyval options, which should include
%% the target journal and optionally the manuscript type.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[journal=jpcbfk,manuscript=article]{achemso}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Place any additional packages needed here.  Only include packages
%% which are essential, to avoid problems later. Do NOT use any
%% packages which require e-TeX (for example etoolbox): the e-TeX
%% extensions are not currently available on the ACS conversion
%% servers.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{rotating} 
\usepackage{times}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{amsmath}
\usepackage{epstopdf}
\usepackage{csquotes}
\usepackage{mhchem}
\usepackage{chemfig}

%\usepackage{markdown}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% If issues arise when submitting your manuscript, you may want to
%% un-comment the next line.  This provides information on the
%% version of every file you have used.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%\listfiles

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Place any additional macros here.  Please use \newcommand* where
%% possible, and avoid layout-changing macros (which are not used
%% when typesetting).
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand*\mycommand[1]{\texttt{\emph{#1}}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Meta-data block
%% ---------------
%% Each author should be given as a separate \author command.
%%
%% Corresponding authors should have an e-mail given after the author
%% name as an \email command. Phone and fax numbers can be given
%% using \phone and \fax, respectively; this information is optional.
%%
%% The affiliation of authors is given after the authors; each
%% \affiliation command applies to all preceding authors not already
%% assigned an affiliation.
%%
%% The affiliation takes an option argument for the short name.  This
%% will typically be something like "University of Somewhere".
%%
%% The \altaffiliation macro should be used for new address, etc.
%% On the other hand, \alsoaffiliation is used on a per author basis
%% when authors are associated with multiple institutions.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\author{Josef Melcr}
\author{Hector Martinez-Seara}
\affiliation{Institute of Organic Chemistry and Biochemistry,
Academy of Sciences of the Czech Republic, 
Prague 6, Czech Republic}
\author{Ji{\v r}{\' i} Kolafa}
\affiliation{Department of Physical Chemistry, Institute of Chemical Technology, Prague 6, Czech Republic}
\author{Pavel Jungwirth}
\affiliation{Institute of Organic Chemistry and Biochemistry,
Academy of Sciences of the Czech Republic, 
Prague 6, Czech Republic}
\alsoaffiliation{Department of Physics, Tampere University of Technology, P.O. Box 692, FI-33101
Tampere, Finland}

\author{O. H. Samuli Ollila}
\email{samuli.ollila@helsinki.fi}
%\homepage[]{Your web page}
\affiliation{Institute of Organic Chemistry and Biochemistry,
Academy of Sciences of the Czech Republic, 
Prague 6, Czech Republic}
\alsoaffiliation{Institute of Biotechnology, University of Helsinki}

%\author{Andrew N. Other}
%\altaffiliation{A shared footnote}
%\author{Fred T. Secondauthor}
%\altaffiliation{Current address: Some other place, Othert\"own,
%Germany}
%\author{I. Ken Groupleader}
%\altaffiliation{A shared footnote}
%\email{i.k.groupleader@unknown.uu}
%\phone{+123 (0)123 4445556}
%\fax{+123 (0)123 4445557}
%\affiliation[Unknown University]
%{Department of Chemistry, Unknown University, Unknown Town}
%\alsoaffiliation[Second University]
%{Department of Chemistry, Second University, Nearby Town}
%\author{Susanne K. Laborator}
%\email{s.k.laborator@bigpharma.co}
%\affiliation[BigPharma]
%{Lead Discovery, BigPharma, Big Town, USA}
%\author{Kay T. Finally}
%\affiliation[Unknown University]
%{Department of Chemistry, Unknown University, Unknown Town}
%\alsoaffiliation[Second University]
%{Department of Chemistry, Second University, Nearby Town}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The document title should be given as usual. Some journals require
%% a running title from the author: this should be supplied as an
%% optional argument to \title.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title[An \textsf{achemso} demo]
  {Binding of cations to Neutral and Negatively Charged Phospholipid Bilayers 
   made accurate through the use of 
   classical fixed-charge MD models with implicit polarizability }

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Some journals require a list of abbreviations or keywords to be
%% supplied. These should be set up here, and will be printed after
%% the title and author information, if needed.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\abbreviations{MD,IR,NMR,UV}
\keywords{American Chemical Society, \LaTeX}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The manuscript does not need to include \maketitle, which is
%% executed automatically.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The "tocentry" environment can be used to create an entry for the
%% graphical table of contents. It is given here as some journals
%% require that it is printed as part of the abstract page. It will
%% be automatically moved as appropriate.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{tocentry}

Some journals require a graphical entry for the Table of Contents.
This should be laid out ``print ready'' so that the sizing of the
text is correct.

Inside the \texttt{tocentry} environment, the font used is Helvetica
8\,pt, as required by \emph{Journal of the American Chemical
Society}.

The surrounding frame is 9\,cm by 3.5\,cm, which is the maximum
permitted for  \emph{Journal of the American Chemical Society}
graphical table of content entries. The box will not resize if the
content is too big: instead it will overflow the edge of the box.

This box and the associated title will always be printed on a
separate page at the end of the document.

\end{tocentry}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The abstract environment will automatically gobble the contents
%% if an abstract is not used by the target journal.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{abstract}
Binding affinities and stoichiometries of Na$^+$ and Ca$^{2+}$ ions to phospholipid
bilayers are of paramount significance in the properties and functionality of cellular
membranes. Current estimates of binding affinities and
stoichiometries of cations are, however, inconsistent due to limitations in the available experimental and
computational methods. In this work, we improve the description of the binding details of Na$^+$ and Ca$^{2+}$ ions to the
1-Palmitoyl-2-oleoyl-phosphatidylcholine (POPC) bilayer by implicitly including electronic
polarization as a mean field correction, known as the electronic continuum correction (ECC).
This is applied by scaling the partial charges of a selected state-of-the-art POPC lipid model for molecular dynamics simulations.
Our improved ECC-POPC model reproduces not only the
experimentally measured structural parameters for the ion-free membrane, but also the
response of lipid head group to a strongly bound cationic amphiphile, as well as the binding affinities of 
Na$^+$ and Ca$^{2+}$ ions. With our new model we observe on the one side negligible binding of Na$^+$ ions to POPC
bilayer, while on the other side stronger interactions of Ca$^{2+}$ primarily with phosphate oxygens, which is
in agreement with the previous interpretations of the experimental spectroscopic data.
The present model results in Ca$^{2+}$ ions forming complexes with one to three POPC molecules with almost
equal probabilities, suggesting more complex binding stoichiometries than those from simple models
used to interpret the NMR data previously.
The results of this work pave the way to quantitative molecular simulations with realistic electrostatic interactions 
of complex biochemical systems at cellular membranes.
\end{abstract}

%\begin{abstract}
%  This is an example document for the \textsf{achemso} document
%  class, intended for submissions to the American Chemical Society
%  for publication. The class is based on the standard \LaTeXe\
%  \textsf{report} file, and does not seek to reproduce the appearance
%  of a published paper.

%  This is an abstract for the \textsf{achemso} document class
%  demonstration document.  An abstract is only allowed for certain
%  manuscript types.  The selection of \texttt{journal} and
%  \texttt{manuscript} will determine if an abstract is valid.  If
%  not, the class will issue an appropriate error.
%\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Start the main part of the manuscript here.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Introduction}
Interactions of ions with cellular membranes play a key role in many important biological processes~\cite{seelig90,cevc90}. Ions, especially multivalent cations, modify general properties of the membrane which also modulate the embedded transmembrane proteins~\cite{cevc90,tocanne90, pabst07}. For example, Ca$^{2+}$  is crucial in propagation of neural signals. It promotes membrane fusion by bridging lipids in the vesicles carrying neurotransmitters and the neuron synapsis membrane~\cite{papahadjopoulos90,brouwer15}. Ca$^{2+}$ also participates in the T-cell receptor activation. It induces the detachment of the positively charged cytosolic tails of the CD3 protein complex from the negatively charged intracellular membrane making it accessible for the Lck protein~\cite{Shi2013Ca2}. Another example involving sugar-lipid interactions mediated by ions concerns Ca$^{2+}$ modulating the presentation of the sugars present in the PI(4,5)P$_2$ lipid which ultimately modulates the phospholipase C delta 1 pleckstrin homology domain (PLC $\delta$1-H)~\cite{Bilkova2017Calcium}. Interestingly, this effect is not present for Mg$^{2+}$, which illustrates the selectivity of these processes with respect to particular ions. Despite our increasing understanding of the role of ions in cell membrane-related processes the molecular details of the mechanisms behind such processes remain elusive in many cases.

Direct measurements of ion-membrane interactions in complex biological systems are difficult. 
Hence, simplified lipid bilayers are often used as entry-level models to shed light on the role of ions in complex biological membranes~\cite{scherer87,seelig90,cevc90}. 
For this reason, interactions of biologically relevant cations, in particular Na$^+$ and Ca$^{2+}$, with zwitterionic phosphocholine (PC) bilayers have been widely studied experimentally~\cite{akutsu81, altenbach84, seelig90, cevc90, tocanne90, binder02, pabst07, uhrikova08} and using molecular dynamics (MD) simulations~\cite{bockmann03, bockmann04, Berkowitz12, melcrova16, javanainen17}. The details of ion binding are, however, not fully consistent in the literature. Non-invasive spectroscopic methods, like nuclear magnetic resonance (NMR), scattering, and infrared spectroscopies mostly suggest that Na$^+$ ions exhibit negligible binding to PC lipid bilayers. In contrast, Ca$^{2+}$ is observed to specifically bind to a couple of PC molecules via their phosphate groups~\cite{hauser76, hauser78, herbette84, akutsu81, altenbach84, binder02, pabst07, uhrikova08}. Most atomistic resolution MD simulation models predict stronger membrane bindings of the cations than observed in experiments~\cite{catte16}. Namely, simulations report various degrees of Na$^+$ accumulation at the lipid interface~\cite{bockmann03} and for Ca$^{2+}$ strong binding to up to four PC lipids simultaneously, involving not only interactions with phosphates but also with carbonyl oxygens~\cite{bockmann04, melcrova16, javanainen17}.

Recent studies within the NMRlipids project (\url{nmrlipids.blogspot.fi})~\cite{catte16} made an attempt to resolve these disagreements. A direct comparison of ion binding affinities to PC bilayers between simulations and experiments has been made possible using the electrometer concept~\cite{seelig87}. Namely, the changes in NMR order parameters of the head group upon addition of ions are directly compared to the MD simulations results. 
Analyzing massive amounts of data collected within an open collaboration approach, 
it was concluded that the accuracy of the current state-of-the-art lipid models for MD simulations is not sufficient for a detailed interpretation of the interactions of cations with PC lipid bilayers~\cite{catte16}.

In this work, we improve the description of cation binding to a POPC bilayer within MD simulations 
via an implicit inclusion of electronic polarizability in the polar region of phospholipids. 
To this end, we apply the electronic continuum correction (ECC)~\cite{leontyev11}, 
within which scaling of the atomic partial charges is employed to account for the missing electronic polarizability in standard force field MD simulations. 
Such an approach has been shown previously to improve the behavior of ions in simulations of aqueous salt solutions~\cite{martinek17, Pluharova2014, kohagen14, kohagen16}. Here, we extend the ECC approach to aqueous lipid membranes taking the Lipid14 force field~\cite{dickson14} as a starting point for refining the POPC model. This choice is justified by the fact that the Lipid14 provides the least inaccurate descriptions of cation binding among the existing force fields\cite{catte16}. 
The newly developed ECC-POPC model reproduces the experimentally
measurable structural parameters of an ion-free POPC lipid bilayer with the accuracy comparable to the best state-of-the-art lipid models, while at the same time significantly improving the membrane binding affinities of sodium and calcium cations.

\section{Methods}

\subsection{Electronic continuum correction for lipid bilayers}\label{section:ecc}
The lack of electronic polarizability in standard MD force fields has been considered a serious issue since the early days of lipid bilayer simulations. In this work, we circumvent the demanding explicit inclusion of electronic polarization effects~\cite{lucas12, chowdhary13} by accounting for the electronic part of polarizability in lipid bilayer simulations implicitly via the electronic continuum correction (ECC)~\cite{leontyev11}. Technically, this is similar to the phenomenological charge-scaling applied in earlier studies of surfactants, lipids or ionic liquids \cite{jonsson86, egberts94, beichel14}. However, the present concept of ECC is physically well justified and has theoretical support~\cite{leontyev09, leontyev10, leontyev11, leontyev14}.

According to ECC, the electronic polarizability can be included into non-polarizable MD in a mean-field way by 
embedding the ions in a homogeneous dielectric continuum with a dielectric constant $\epsilon_{el}$, which is the electronic part of the dielectric constant of the medium~\cite{leontyev11}. Following Coulomb's law,  ECC can be directly incorporated by scaling the charges with a constant scaling factor of $f_q = \epsilon _{el} ^{-1/2}$, yielding
\begin{equation}
Q^{ECC} = f_q \cdot Q
\end{equation}
for the ECC corrected charges. 
Given that the  high frequency dielectric constant of water is $\epsilon _{el} = 1.78$ (\textit{i.e.,} the square of the refraction index), the scaling factor for ions in water is roughly $f_q \approx 0.75$. This scaling factor significantly improves the accuracy of simulations of solvated ions, when quantitatively compared with neutron scattering data \cite{kohagen14,kohagen16, Pluharova2014, martinek17}.
It is important to note that the value of the
high frequency dielectric constant 
is around 2 for almost any biologically relevant environment \cite{leontyev11}.
The dielectric discontinuity in a lipid bilayer thus arises only
from the orientational polarization of the molecules, which is accounted for explicitly in standard MD simulations. 
Therefore, the same correction for the electronic polarizability can be 
applied throughout the lipid bilayer/aqueous solution interface.

While using the scaling factor of $f_q = 0.75$ for ions in water is well justified in the ECC theory \cite{leontyev11}, it is not clear whether the same factor should be applied to partial charges used to describe molecules in MD models, e.g., lipids in our case \cite{leontyev14}. Unlike the total charge of an ion, atomic partial charges within molecules are not physical observables. Several computational schemes exist for the assignment of partial charges for biomolecules~\cite{Hu2007}, with the restrained electrostatic potential method (RESP) being commonly used~\cite{RESP_paper, Singh1984}. Considering that water is often included in RESP calculations or charges are refined to improve certain experimental observables, the electronic polarizability effects of the solvent may to some extent be included in standard force fields~\cite{RESP_paper, Singh1984, jorgensen96, ipolq2013, benavides17}. Thus, our application of the scaling factor, $f_q$, to existing partial charges in molecules does not necessarily follow $f_q = \epsilon _{el} ^{-1/2}$. Instead, a consistent scaling factor should lie between the value of 0.75  (\textit{i.e.,} no electronic polarizability included in the original partial charges) and 1 (\textit{i.e.,} electronic polarizability fully included in the original partial charges).

Here we develop a ECC-POPC lipid model that accurately describes binding 
of sodium and calcium ions to the POPC  lipid bilayer. 
The Lipid14~\cite{dickson14} force field 
(available in a Gromacs format from Ref. \citenum{lipid14files}) was used as a starting 
point since 
it provides the most accurate response of the head groups to ions among the available 
lipid models (see Figs.~2~and~5 in Ref.~\citenum{catte16}). Additionally, the Lipid14 model 
provides relatively realistic head group, glycerol backbone, and acyl chain structures~\cite{dickson14,botan15}.
We applied the ECC correction 
to the Lipid14 model of POPC by scaling 
partial charges of the head group, glycerol 
backbone, and carbonyl regions. 
These are the polar parts of phospholipids which can 
contribute to the cation binding.

To reproduce the experimental ion binding affinities,
we scanned possible values of the scaling factor from the interval $f_q~\in~(0.75, 1.0)$.
The ion binding affinity was benchmarked against experiments
using the head group order parameters and the electrometer concept \cite{seelig87,catte16},
as discussed more detail in the next section.
Scaling down the partial charges reduced the ion binding affinity.
We found the most accurate ion binding affinities with a scaling factor of $f_q = 0.8$,
which is only slightly higher than the ECC one ($f_q=0.75$).
Note that common empirical scaling factors for monovalent ions in water are 0.8 or
even closer to unity \cite{benavides17,skinner14,nacleps}.  In contrast, modern force fields
for ionic liquids suggest values of 0.6--0.65, which are lower than $\epsilon^{-1/2}_{el}$\cite{holm14}.
Directly applying
the 0.8 scaling to the partial charges of the head group, the glycerol backbone, and
the carbonyls reduced the area per lipid to 60~\AA$^2$. This area is smaller than in the
original Lipid14 model ($65.6 \pm 0.5$~\AA$^2$)~\cite{dickson14} and in experiments
(64.3~\AA$^2$) \cite{kucerka11}. The decrease of the area per lipid arises from a
reduced hydration of the lipid head group region after scaling of the partial charges, which effectively
reduces the head group polarity. We solve this problem by slightly reducing the effective radii of
the modified head group atoms via lowering the $\sigma$ parameters in the Lennard-Jones potential by a
factor of $f_\sigma = 0.89$. The same approach was successfully adopted for the ECC ions in aqueous
solutions previously \cite{kohagen14, kohagen16, Pluharova2014, martinek17}. After reducing the head group atoms $\sigma$ parameters, the area per molecule is restored to the experimental value (Table~\ref{tab:apls}).

\subsection{Electrometer concept} \label{section:electrometer}
Comparing MD simulation to NMR experiments, we can validate the ion
binding affinity in lipid bilayer simulations using the 'electrometer concept'~ \cite{seelig87, catte16}.
This method is based on the experimental observation that the C-H bond order parameters of $\alpha$ and $\beta$ carbons in a PC lipid head group (Fig.~\ref{simVSexpNOions}) are proportional to the amount of charge bound per lipid~\cite{seelig87}.
The order parameters for all C-H bonds in lipid molecules can be accurately measured using $^2$H NMR or $^{13}$C NMR techniques \cite{ollila16}. From MD simulations the order parameters can be calculated using the relation
\begin{equation}\label{OP}
S_{\rm CH}=\frac{1}{2}\langle 3\cos^2\theta -1 \rangle,
\end{equation}
where $\theta$ is the angle between the C-H bond and the membrane normal. Angular brackets point to the average over all sampled configurations.

The relation between the amount of the bound charge per lipid, $X^\pm$, and the head group order parameter change, $\Delta S_{\rm{CH}}^{i}$, is empirically quantified as~\cite{seelig87,ferreira16}
\begin{equation}\label{OPchangeEQ}
\Delta S_{\rm{CH}}^{i}= S_{\rm{CH}}^{i}(X^\pm)-S_{\rm{CH}}^{i}(0) \approx m_i \frac{4 }{3\chi}X^\pm,
\end{equation}
where $i$ refers to either the $\alpha$ or $\beta$ carbons, $S_{\rm{CH}}^{i}(0)$ denotes the order parameter in the absence of bound charge, $\chi$ is the quadrupole coupling constant ($\chi \approx$\,167\,kHz), and $m_i$ is an empirical constant depending on the valency and location of the bound charge.

The measured change of the order parameter depends on the head group response to the bound charge and on the amount of the bound charge (\textit{i.e.,} $m_i$ and $X^\pm$ in Eq.~\ref{OPchangeEQ}, respectively).  The empirical factor $m_i$ has to be adequately quantified before the electrometer concept can be used to analyze the binding affinities. This calibration has been done experimentally for a wide range of systems~\cite{seelig87, beschiasvili91}. To calibrate the response of the head group order parameters to the bound charge in simulations, we use experimental data for a strong cationic surfactant dihexadecyldimethylammonium bromide  (DHAB) mixed with a POPC bilayer~\cite{scherer89}. DHAB\\[0.5cm]
%\vspace{0.5cm}
\chemfig{
  -[:0,3.5,,,draw=none]\chemabove{N}{\scriptstyle\oplus} (-[:150]H_3C)(-[:210]H_3C)(-[:330]{(}CH_2{)}_{15}-CH_3)(-[:30]{(}CH_2{)}_{15}-CH_3)
}  
\vspace{0.5cm} \\
is a cationic surfactant having two acyl chains and bearing a unit charge at the hydrophilic end. Due to its structure it is expected to locate in the bilayer similarly to the phospholipids and its molar ratio then gives directly the amount of bound unit charge per lipid $X^\pm$ in these systems~\cite{scherer89}.

\subsection{Salt concentrations and binding affinities}
When measuring head group order parameters, the NMR experiments report the employed salt concentrations in two different ways.
Some studies provide the salt concentrations in water before solvating the lipids~\cite{akutsu81}, $C_{ion}'$,
while others use atomic absorption spectroscopy and report the salt concentration in
the supernatant after the solvation of lipids~\cite{altenbach84}, $C_{ion}$. In this work,
we stick to the latter definition by estimating the salt concentration in the aqueous
bulk region using the farthest point from the lipid bilayer in the aqueous phase.
Note that the former definition was used by some of us previously.~\cite{catte16}.
Although these two definitions provide somewhat different values when applied
to Ca$^{2+}$ concentrations, their particular choice does not affect the conclusions of this work in any significant way.

To quantify the ion binding affinities to a lipid bilayer, we calculate the relative surface excess of ions with respect to water, $\Gamma_{ion}^{water}$~\cite{chattorajBOOK}. Such a quantity compares the adsorption of ions to the adsorption of water molecules at the interface without the necessity of defining a Gibbs dividing surface between the membrane interior and the water bulk region. In our simulations, we only assume that the interface is located between the ion-free hydrophobic interior of the lipid bilayer and the aqueous region far from the membrane.  Such a setup and the above definition of bulk ion concentrations provides a simple relation for the relative surface excess $\Gamma_{ion}^{water}$ for simulations of lipid bilayers,
 \begin{equation}\label{surfexcess}
  \Gamma_{ion}^{water}=\frac{1}{2A_b} \left ( n_{ion} - n_{water} \frac{C_{ion}}{C_{water}} \right ).
\end{equation}
Here, $n_{water}$ and $n_{ion}$ are the total numbers of water molecules and ions in the system,
$C_{water}$ and $C_{ion}$ are their respective bulk concentrations in the aqueous phase
(estimated from the solution density of farthest point between periodic bilayers),
and $A_b$ is the area of the unit cell in the membrane plane.
The total area of the interface is then twice the area of the membrane, i.e., $2A_b$,
because the bilayer has an interface at each of the two leaflets.

\subsection{Validation of lipid bilayer structure against NMR and scattering experiments}
The structures of lipid bilayers in simulations without ions were validated against NMR by calculating the order parameters for the C-H bonds and against \mbox{X-ray} scattering experiments by evaluating the scattering form factors. NMR order parameters validate the structures sampled by the individual lipid molecules with atomic resolution. The simulated order parameters were calculated for all C-H bonds in lipid molecules from Eq. \ref{OP}. Scattering form factors validate the dimensions of the lipid bilayer (i.e., the bilayer thickness and area per molecule). Form factors were calculated using a relation

\begin{equation}
  F(q) = \int _{-D/2} ^{D/2} \left ( \rho_{el}(z) - \rho_{el}^s \right ) \cos (zq_z) \mathrm{d}z,
  %F(q) = \int _{-D/2} ^{D/2} \left ( \sum _\alpha f_\alpha (q_z) n_\alpha (z) - \rho _s \right ) \exp (izq_z) \mathrm{d}z,
\end{equation}

\noindent where $\rho_{el} (z)$ is the total electron density, $\rho_{el}^s$ is the electron density of the solvent far in the aqueous bulk, and $z$ is the distance from the membrane center along its normal with $D/2$ being half of the unit cell size.

\subsection{Simulation details}

\subsubsection{Simulations of POPC bilayers with aqueous ions}
Simulations of a POPC bilayer in pure water or at varying salt concentrations consisted of 128 POPC molecules
and approximately 50~water molecules per lipid in an orthorhombic simulation box with periodic boundary conditions.
The SPC/E~\cite{Berendsen1987} water model was used in all ECC-POPC model simulations reported
in the main text. This model was also used in the previous parametrization
of ECC-ions \cite{martinek17, kohagen16, Pluharova2014} because its lowered dielectric constant is consistent with
the ECC concept~\cite{leontyev11, leontyev14}.
The robustness of our approach with other water models (OPC \cite{Izadi14},
OPC3~\cite{Izadi16}, TIP3P~\cite{jorgensen83}, TIP3P-FB and TIP4P-FB~\cite{Wang2014}, and TIP4P/2005 \cite{Abascal2005})
was verified and is presented in the Supporting Information (SI).
Sodium, calcium, and chloride ions were modeled as ECC-ions
with parameters from Refs. \citenum{martinek17, Pluharova2014}. Scaled charges and Lennard-Jones radii for atoms
forming the POPC lipid were derived in this work starting from the Lipid14 force field~\cite{dickson14}.
For comparison, simulations with the Lipid14 model~\cite{dickson14} and
ion models by Dang and coworkers~\cite{smith94,chang1999,dang2006} or ECC-ions \cite{martinek17, kohagen16, Pluharova2014}
were also performed. The TIP3P water model~\cite{jorgensen83} was used in all simulations with the original Lipid14 model.
Simulation data for the Lipid14 model with \AA{}qvist ions~\cite{aqvist90} and the TIP3P~\cite{jorgensen83} water model were taken directly from
Refs.~\citenum{catte16,lipid14POPC0mMNaClfiles,lipid14POPC350mMCaClfiles,lipid14POPC350mMCaClfilesNC,lipid14POPC1000mMNaClfiles}.

MD simulations were performed using the GROMACS \cite{Abraham15} simulation package (version 5.1.4). The simulation parameters used
in this work are summarized in Table~S1 in the SI. 
Compatibility with the openMM simulation package \cite{openmm7} was also
tested and is presented in the SI. 
For deposited trajectories and parameter files see 
Refs.~\citenum{lipid14_POPC_Dang_and_ECC-ions_files, ECC-POPC_nacl_cacl2_files, ECC-POPC_nacl_cacl2_wat-models_files}.

\subsubsection{Simulations of POPC bilayers with cationic surfactants}
The topology of the dihexadecyldimethylammonium cation was created with the automated topology builder \cite{malde11}. The AmberTools program~\cite{amber} was used to generate the Amber-type force field parameters. These parameters were then converted to the Gromacs format with the acpype tool \cite{acpype}. The partial charges were manually modified in the non-scaled version to approximately match similar segments in Lipid14 \cite{dickson14}. For simulation with ECC-POPC, the partial charges of the head group of the cation were scaled by a factor of $f_q=0.8$, and the $\sigma$ of the scale atoms reduced by a factor of 0.89 similarly as for POPC-ECC.

The cationic surfactants were randomly mixed among the phospholipids to form bilayer structures with mole fractions of 10\%, 20\%, 30\%, 42\%, or 50\% of surfactant in the POPC bilayer. All these systems contained 50 POPC molecules per leaflet, 6340 water molecules, and 6, 14, 21, 35, or 50 surfactants per leaflet.  Chloride counter ions were used in all the simulations, while bromide was the counter ion in the experiment~\cite{scherer89}. For POPC, either Lipid14 with TIP3P~\cite{jorgensen83} or ECC-POPC with SPC/E~\cite{Berendsen1987} parameters  were used. The first 20~ns of the total simulation time of 200~ns was considered as an equilibration period and was thus omitted from the analysis. We checked that a sufficient lipid  exchange occurred during the simulations.
For deposited trajectories and parameter files of scaled and unscaled simulations see
Refs.~\citenum{ECC-POPC_cationic-surf_files} and~\citenum{POPClipid14T313K,POPClipid1410perCATsurfT313K,POPClipid1420perCATsurfT313K,POPClipid1430perCATsurfT313K,POPClipid1442perCATsurfT313K,POPClipid1450perCATsurfT313K}, respectively.

\section{Results and Discussion}

\subsection{Structural parameters of a pure ECC-POPC model membrane: Agreement with experiments}

\begin{figure}[tb!]
  \centering
  \includegraphics[width=8.2cm]{../Fig/ipython_nb/Order-parameters_form-factors_exp-L14-ECCL17_q80_sig89_POPC-struct.pdf}
  \caption{\label{simVSexpNOions}
    Top: X-ray scattering form factors from simulations with the Lipid14 \cite{dickson14} and
    the ECC-POPC models compared with experiments~\cite{kucerka11} at 303~K. \\
    Middle: Order parameters of POPC head group, glycerol backbone and acyl chains 
    from simulations with the Lipid14 \cite{dickson14} and the ECC-POPC models
    compared with experiments \cite{ferreira13} at 300~K.
    The size of the markers for the head group order parameters correspond to
    the error estimate $\pm 0.02$ for experiments \cite{botan15,ollila16},
    while the error estimate for simulations is $\pm 0.005$
    (Bayesian estimate of 95\% confidence interval \cite{scipy}).
    The size of the points for acyl chains are decreased by a factor of 3 to improve the clarity of the plot.
    Open/closed symbols are used for palmitoyl/oleoyl chains of POPC. \\
    Bottom: The chemical structure of POPC and the labeling of the carbon segments.
  } 
\end{figure}

\begin{table}[tb!]
  \caption{Values of the area per lipid (APL) of POPC bilayers without ions. \label{tab:apls}
  }
  \begin{tabular}{l|c c}
    model          & APL (\AA$^2$)   & Temperature [K] \\
    \hline
    Lipid14                   & 65.1$\pm$ 0.6  &  300 \\
    Lipid14 \cite{dickson14}  & 65.6$\pm$ 0.5  &  303 \\
    \hline
    ECC-POPC                & 63.2$\pm$ 0.6  &  300       \\
    \hline
    experiment \cite{kucerka11} & 64.3  &  303    \\
    \hline
  \end{tabular}
\end{table}

First, we present results for bilayers in pure water.
The ECC-POPC and Lipid14 models both reproduce the experimental X-ray scattering form factors
of a POPC bilayer with a comparable accuracy (see Fig.~\ref{simVSexpNOions}).
The area per lipid from the Lipid14 model is by $\approx$1\AA{} larger than the
experimental value in Table~\ref{tab:apls}, while the value from the ECC-POPC model
is by $\approx$1\AA{} smaller than the experimental one.
The values of the area per lipid of the ECC-POPC model vary slightly
when simulated with different water models (i.e., within the interval of 62.2--66.8 \AA{}, see Table~S2 in SI),
while still being close to the experimentally reported values.
We can thus conclude that the ECC-POPC model reproduces the experimental dimensions of the POPC
lipid bilayer with a comparable accuracy to other state-of-the-art lipid models~\cite{ollila16}.

Similarly, the acyl chain order parameters of the ECC-POPC model, as well as those of the Lipid14 model~\cite{dickson14}, agree with the experimental values within the error bars, as presented in Fig.~\ref{simVSexpNOions}. Notably, the experimentally measured forking and small order parameter values of the $C_2$ segment in {\it sn}-2 chain are well reproduced by both models. This feature has been suggested to indicate that the carbonyl of the {\it sn}-2 chain is directed towards the water phase, in contrast to the carbonyl in the {\it sn}-1 chain, which orients more along the bilayer plane~\cite{seelig75,schindler75,gawrisch92}.
This arrangement, which is not fully reproduced by other available lipid models~\cite{ollila16}, may be a relevant feature for the ion binding details.

The order parameters of the $\alpha$ and $\beta$ carbons in the head group are slightly larger in the ECC-POPC model than in the Lipid14 model, which is apparently related to the P-N vector orienting by about 7$^{\circ}$ more toward the water phase in the former model, see Fig.~\ref{OrderParameterCHANGESsurf}. While both models perform relatively well, considering the available experimental evidence, it is not possible to decide which of the two models provides more realistic head group orientations. The ECC-POPC model gives the $\beta$ carbon order parameter value closer to experiments than the Lipid14 model, while the opposite is true for the $\alpha$ carbon. The accuracy of both models in the glycerol backbone region is comparable to other state-of-the-art lipid model available in literature \cite{botan15}, see Fig.~\ref{simVSexpNOions}.

\subsection{Calibration of head group response to membrane-bound charge using cationic surfactant}\label{section:boundCHARGE}

\begin{figure}[tb!]
  \centering
  \includegraphics[width=8.0cm]{../Fig/ipython_nb/PN_angle_OrdPars-A-B_L14-ECCL17_q80_sig89_surf.pdf}
  \caption{\label{OrderParameterCHANGESsurf}
    The changes of head group order parameters and P-N vector orientation as a function of
    a molar fraction of the cationic surfactant dihexadecyldimethylammonium in a POPC bilayer
    from simulations and experiments \cite{scherer89} at 313 K.
  }
\end{figure}

Before studying the sodium and calcium ion binding affinities, we quantify the response of the head group order parameters to the amount of bound charge by using mixtures of monovalent cationic surfactants (dihexadecyldimethylammonium) and POPC~\cite{scherer89}. These mixtures have a well-defined amount of bound charge per PC, namely, the molar fraction of cationic surfactants. This is due to the ability of dihexadecyldimethylammonium to directly insert in the lipid bilayer due to its two hydrophobic acyl chains. Furthermore, available experimental data for these systems can be used to validate the sensitivity of lipid head group order parameters to the amount of bound charge in simulations \cite{scherer89}.

The changes of the head group order parameters with increasing amount of the cationic surfactant are compared between simulations and experiments~\cite{scherer89} in Fig.~\ref{OrderParameterCHANGESsurf}. An approximately linear decrease of the order parameters, as expected from Eq. \ref{OPchangeEQ}, is observed in both simulations and experiments at least for mole fractions below $\sim$30\%. The slope is, however, too steep in the Lipid14 model indicating that the response of head group order parameters to the bound positive charge is too strong. In contrast, the slope of the ECC-POPC model is in a very good agreement with experiments for the $\alpha$ segment, while being only slightly underestimated for the $\beta$ segment.

In Fig.~\ref{OrderParameterCHANGESsurf}, we show the headgroup P-N vector angle as a function of the mole fraction of the cationic surfactant. As suggested previously~\cite{seelig87}, the headgroup orients more towards the water phase with the increasing amount of positive charge in the PC lipid bilayer. The effect is more pronounced in the Lipid14 model than in the ECC-POPC model.   For example, the addition of 50\% mole fraction of the cationic surfactant leads to a decrease of 20$^{\circ}$ of the P-N vector angle for the Lipid14 model while only of 11$^{\circ}$ in the ECC-POPC model. The difference is in line with the smaller order parameter changes and the reduced charge--dipole interactions in the latter model. The weaker sensitivity of the P-N vector angle response in the ECC-POPC model is in better agreement with experiments.

\subsection{Validation of ECC-POPC model using binding affinities to Na$^+$ and Ca$^{2+}$ cations: the electrometer concept}

\begin{figure}[htb!]
  \centering
  \includegraphics[width=8.0cm]{../Fig/ipython_nb/OrdPars-A-B_L14-ECCL17_q80_sig89_NaCl.pdf}
  \caption{\label{fig:delta_ordPar_NaCl}
    Changes of the head group order parameters of a POPC bilayer as a function of NaCl concentration
    in bulk ($C_{ion}$) from simulations with different force fields at 313 K together with 
    experimental data for DPPC (323\,K) \cite{akutsu81} and POPC (313\,K) \cite{altenbach84}.
    Simulation data with Lipid14 and \AA{}qvist ion parameters at 298 K are taken directly from
    Refs.~\cite{lipid14POPC0mMNaClfiles,lipid14POPC1000mMNaClfiles}.
  }
\end{figure}

\begin{figure}[htb!]
  \centering
  \includegraphics[width=8.0cm]{../Fig/ipython_nb/PN_angle_OrdPars-A-B-g3_L14-ECCL17_q80_sig89_CaCl.pdf}
  \caption{\label{fig:delta_ordPar_CaCl}
    Changes of the head group order parameters and P-N vector orientation of a POPC bilayer 
    as a function of the CaCl$_2$ concentration in bulk ($C_{ion}$)
    from simulations at 313 K together with experimental data 
    (DPPC (323\,K) \cite{akutsu81} and POPC (313\,K) \cite{altenbach84}). 
    The error estimate for bulk concentrations is approximately 10\,mM.
    The order of magnitude larger error in the
    simulation with Lipid14 and ECC-ions is due to unconverged bulk densities  (shown if Fig.~\ref{fig:cacl-dens}) limited by
    the simulation box. 
    Simulation data with Lipid14 and \AA{}qvist ion parameters at 298 K are taken directly from
    Refs.~\citenum{lipid14POPC0mMNaClfiles,lipid14POPC350mMCaClfiles,lipid14POPC350mMCaClfilesNC}.
  }
\end{figure}

Changes of the lipid bilayer head group order parameters extracted from simulations and
experiments \cite{akutsu81, altenbach84} are shown in Figs.~\ref{fig:delta_ordPar_NaCl}
and~\ref{fig:delta_ordPar_CaCl} as functions of NaCl or CaCl$_2$ concentrations.
As seen in Fig.~\ref{OrderParameterCHANGESsurf}, the order parameters decrease
proportionally to the amount of the bound positive charge.
These results can be thus used to compare the ion binding affinities to lipid bilayers between
simulations and experiments using the electrometer concept~\cite{seelig87, catte16}.

The experimentally measured small order parameter
changes with NaCl (Fig.~\ref{fig:delta_ordPar_NaCl}) 
are reproduced by the Lipid14 model simulated with \AA{}qvist ions.
However, the same combination of models overestimates the order parameter changes with CaCl$_2$ (Fig.~\ref{fig:delta_ordPar_CaCl}).
Replacing \AA{}qvist ions with ion parameters by Dang et al.~\cite{smith94, chang1999, dang2006}
or ECC-ions~\cite{martinek17, kohagen16, Pluharova2014} did not improve
the results~(Figs.~\ref{fig:delta_ordPar_NaCl} and \ref{fig:delta_ordPar_CaCl}).
In line with the previous work \cite{catte16}, the results suggest that improvements
in the lipid parameters are required to correctly describe the binding of cations to phospholipid bilayers.

The results from simulations combining the ECC-POPC with the ECC-ion models \cite{martinek17, kohagen16, Pluharova2014} exhibit a significantly improved behavior of the POPC head group order parameters as a function of NaCl or CaCl$_2$ concentrations, see Fig.~\ref{fig:delta_ordPar_NaCl} and Fig.~\ref{fig:delta_ordPar_CaCl}. Considering that we are also able to reproduce the experimental response in systems with known charge density (see above section \ref{section:boundCHARGE}), we conclude that our ECC model correctly reproduces the binding affinities of Na$^{+}$ and Ca$^{2+}$ ions to the POPC lipid bilayer. Furthermore, while the response of the glycerol backbone $g_3$ order parameter to CaCl$_2$ was significantly overestimated in the original Lipid14 model, the ECC-POPC model provides an improved agreement with experiment, as seen in Fig.~\ref{fig:delta_ordPar_CaCl}.
Also the changes of the P-N vector angle are too pronounced for the Lipid14 model,
for which the largest tilting toward water phase induced by a $780\,\mathrm{mM}$
CaCl$_2$ concentration is approximately 17$^{\circ}$. The corresponding value
for the ECC-POPC simulation is only 6$^{\circ}$ ($820\,\mathrm{mM}$ CaCl$_2$).

Within the Lipid14 model, the overestimated changes in the lipid headgroup order parameter of POPC  as functions of the CaCl$_2$ concentration arise both from the overestimated binding affinity and the excessive sensitivity of the headgroup tilt to the bound positive charge. It is plausible to assume that the same applies to the other lipid models tested in a previous study~\cite{catte16}, which underlines the importance of validation of the lipid headgroup order parameter response to the bound charge.

Finally, the ion binding affinities for the ECC-POPC model with different water models are compared in SI. In general, the performance of ECC-POPC with any of the tested water models is better than that of the original Lipid14 model, with the order parameter changes being slightly overestimated with the four-site water models and with TIP3P model.

\subsection{Binding affinities of Na$^+$ and Ca$^{2+}$ cations to the POPC membrane}
\label{sec:affinity}

\begin{table}[tb!]
  \caption{Bulk concentrations of Ca$^{2+}$ (C$_b$), relative surface excess of calcium with respect to water ($\Gamma_{Ca}^{\rm water}$),
    and the percetages of Ca$^{2+}$ bound to phosphate or carbonyl oxyges ($r^\mathrm{Ca^{2+}} _\mathrm{PO_4} $ and $r^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}}$)
    in different POPC bilayer models. All systems have the same molar concentration of Ca$^{2+}$ with respect to water ($C_{ion}'$=350mM).
  \label{tab:binding}}
  \begin{tabular}{l|c c | c | c c}
    model                  & $C_{ion}'$ & $C_{ion}\,/\,\mathrm{mM}$ & $\Gamma_{Ca}^{\rm water} \,/\, \mathrm{nm}^{-2}$  & $r^\mathrm{Ca^{2+}} _\mathrm{PO_4} $ & $r^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}} $ \\
    \hline
    ECC-POPC             &  350  &  $280\pm 10 $  &  $0.06 \pm 0.01 $                           &  99\%  &    25\%    \\
    Lipid14/\AA{}qvist     &  350  &  $210\pm 10 $  &  $0.13 \pm 0.01 $                          & 100\%  &    37\%     \\
    Lipid14/Dang           &  350  &  $160\pm 10 $  &  $0.23 \pm 0.03 $                            & 100\%  &    14\%    \\
    Lipid14/ECC-ions       &  350  &  $120\pm 100$  &  $0.35 \pm 0.11 $                         & 100\%  &    23\%    \\
  \end{tabular}
\end{table}

Binding affinities of Ca$^{2+}$ ions to a POPC bilayer in different simulation models were quantified by calculating the relative surface excess of calcium with respect to water molecules, $\Gamma_{\rm ion}^{\rm water}$, from Eq.~\ref{surfexcess}.
The values of $\Gamma_{\rm ion}^{\rm water}$
from different simulations with the same molar concetration of cations with respect
to water ($C_{ion}'$=350mM) are shown in Table~\ref{tab:binding}.
As expected from the changes of the lipid headgroup order parameters in Fig.~ \ref{fig:delta_ordPar_CaCl}, the relative surface excess of calcium, $\Gamma_{\rm Ca}^{\rm water}$ = 0.06~nm$^{-2}$, is significantly smaller for the ECC-POPC model than for the other models, 0.13--0.35~nm$^{-2}$.
Interestingly, the calculated relative surface excess of \ce{NaCl} at $1\,\mathrm{M}$ concentration (ECC-ions~\cite{Pluharova2014}) using our ECC-POPC model is not only quantitatively but also qualitatively different from \ce{CaCl2} having actually a negative value of $\Gamma_{Na}^{water} = -0.11 \pm 0.01) \rm{nm}^{-2}$ (Fig.~\ref{fig:cacl-dens}. This 
means that on average water molecules are preferred to sodium and chloride ions at the membrane-water interface.  
This is in contradiction with most of the available lipid force fields, which predict a positive surface excess of sodium at PC lipid bilayers \cite{catte16}.

\begin{figure}[htbp!]
  \centering
  \includegraphics[height=19.9cm]{../Fig/ipython_nb/density_profiles_ca_cl_wat_phos_models-compar.pdf}
  \caption{\label{fig:cacl-dens}
    Number density profiles of \ce{Ca^{2+}}, \ce{Na^{+}} and \ce{Cl^-} along membrane normal axis
    for different force fields.
    In order to visualize the density profiles with a scale comparable to the profile of \ce{Ca^{2+}}, 
    the density profiles of~\ce{Cl^-} and \ce{Na^{+}} ions are divided by 2, and
    the density profiles of phosphate groups and water are divided by 5 and 200, respectively. 
    All simulations with \ce{CaCl2} shown here have the same molar concentration of ions in water ($C_{ion}'$=350~mM).
    The simulation with \ce{NaCl} has $C_{ion}'$=1000~mM.
    }
\end{figure}

\subsection{Molecular interactions between Na$^+$ or Ca$^{2+}$ cations and POPC oxygens}
We analyzed the ratio of the number of calcium cations bound to either phosphate or carbonyl moieties and the total number of bound cations in our POPC bilayers as done previously in Ref.~\citenum{javanainen17}. A maximum distance of 0.3~nm from any lipid oxygen is used to define a bound calcium. The results from ECC-POPC simulation in Table~\ref{tab:binding} show that almost all (99\%) of the bound Ca$^{2+}$ ions are in direct contact with phosphate oxygens. From these ions, only one third (32\%) also interacts with the carbonyl oxygens, while the interaction of calcium ions with carbonyl oxygens only is rare (1\%). The most abundand interaction scenarios between Ca$^{2+}$ ions and phosphate oxygens are visualized using the probability density isocontours in Fig.~\ref{fig:volmaps}. While higher concentrations of \ce{CaCl2} increase the number of contacts per lipid, the distribution of contacts between phosphate and carbonyl oxygens is not affected.

\begin{figure}[tb!]
  \centering
  \includegraphics[width=8.0cm]{../Fig/isocontours_r37_ca_O-carb.png} %\\
  \caption{\label{fig:volmaps}
    Isocontours of spatial number density of \ce{Ca^{2+}} (dark blue, 0.001~\AA$^{-3}$) 
    and POPC carbonyl oxygen atoms (light semi-transparent red, 0.008~\AA$^{-3}$, all POPC lipids contribute). 
    Calcium cations localize mostly around phosphate oxygens (oxygens red, phosphorus bronze).
    Interactions with carbonyl oxygens is less likely than with phosphate oxygens, 
    and it is contributed more by other neighbouring phospholipids than by the same lipid. 
    Transparent structures are shown to depict the variability of choline configurations 
    (colour warps from red to blue along the simulation time). 
    The number density was evaluated for each lipid, 
    after its structural alignment using only phosphate group.
    MDAnalysis \cite{mdanalysis2011} library was used for 
    the calculations of the structural alignment and the spatial number density. 
    VMD \cite{hump96} was used for visualisation. 
    Carbon atoms are depicted in cyan, hydrogen atoms in white, oxygen atoms in red, nitrogen in blue.
  }
\end{figure}

Even though \ce{Na+} ions do not bind strongly to a POPC bilayer, they still interact mostly with its oxygen moieties. The results from a simulation at a $1\,$M \ce{NaCl} concentration show that 55\% of \ce{Na+} ions at the bilayer interact with phosphate oxygens of POPC only and 20\% with carbonyl oxygens only, with the remaining 25\%, is interacting with both negatively charged groups.

In conclusion, the results suggest that calcium ions bind specifically to the phosphate oxygens, occasionally interacting also with the carbonyls of the PC lipids. This is in a qualitative agreement with previous conclusions from several experimental studies~\cite{hauser76, hauser78, herbette84, cevc90, binder02}. However, the present results suggest, in
agreement with experiments, an overally weaker binding to the bilayer, in particular with a lower relative binding affinity to the carbonyls than inferred from previous MD simulation studies~\cite{bockmann03, bockmann04, melcrova16, javanainen17}. Sodium ions, which do not exhibit any appreciable affinity for the bilayer, also interact primarily with phosphate oxygens of the POPC, but in contrast to calcium, the interactions purely with carbonyls are also significant.

\subsection{Binding stoichiometry of \ce{Na+} and \ce{Ca^{2+}} cations to POPC membrane}
Simple binding models have been used previously to interpret the same experimental data \cite{altenbach84,macdonald87} as employed in this work to validate the simulation models (Fig.~\ref{fig:delta_ordPar_CaCl}). In particular, NMR data concerning the PC headgroup order parameters response and atomic absorption spectra were explained best using a ternary complex binding model with a binding stoichiometry of one \ce{Ca^{2+}} per two POPC lipids~\cite{altenbach84}. Nevertheless, a Langmuir adsorption model assuming a \ce{Ca^{2+}}:POPC stoichiometry of 1:1 also provided a good fit to the experimental data when considering \ce{CaCl2} at low concentrations only~\cite{macdonald87}.

In this work, we reproduce the same experimental data used to infer binding stoichiometries employing our ECC-POPC model. Thanks to our simulations, we have a direct access to atomistic details of the binding stoichiometry without a need for any binding model as employed for interpreting in experiments~\cite{altenbach84, macdonald87}.
To evaluate the relative propensities for each of the stoichiometric complexes (i.e.,~1~Ca$^{2+}$:~n~POPC),
we calculated for each bound Ca$^{2+}$ the number of POPC molecules having oxygen atoms within a distance of 0.3~nm.
Results from the POPC bilayer simulation with a 285~mM bulk concentration of CaCl$_2$ are shown in Fig.~\ref{fig:cacl_complexes}.
We found the largest propensity for the 1:2 complex (41\%), with probabilities of complexes with the stoichiometries of 1:1~(25\%)~and 1:3~(34\%)~being only slightly lower. This suggests a more complex binding model than considered in a simple 1:2 ternary complex model previously. Nevertheless, with a broad brushstroke, the simulation data can be viewed such that one calcium binds to two lipids on average, because the probabilities of the complexes with 1 or 3 lipids are almost equal to each other  (and complexes with more than three lipids per one calcium ion were not observed). This probably explains why the simple the ternary complex model fits adequately the experimental data, as well as the ECC-POPC simulation results (see Fig.~S3 in SI).

\begin{figure}[tb!]
  \centering
  \includegraphics[width=8.0cm]{../Fig/ipython_nb/stoichiometry_NaCl-CaCl2_comparison_Ecc-lipids.pdf} \\
  \caption{\label{fig:cacl_complexes}
      Relative probabilities of existence of \ce{Na^{+}} or \ce{Ca^{2+}} complexes
      with a certain number of POPC lipids. 
      \ce{Na^{+}} complexes were evaluated from the simulation with 1~M concentration;
      and \ce{Ca^{2+}} complexes were evaluated from the simulation with 287~mM concentration.
  }
\end{figure}

The probabilities of different complexes formed by \ce{Na+} ions and POPC
analyzed from the simulation with the ECC-POPC model at $1\,$M concentration of NaCl are also 
shown in Fig.~\ref{fig:cacl_complexes}. In contrast to calcium, the
probability is largest (67\%) for 1:1 complex, significantly smaller (29\%)
for 1:2 complexes and very small (4\%) for 1:3 of \ce{Na+}:POPC complexes.

\subsection{Residence times of \ce{Na+} and \ce{Ca^{2+}} cations in the POPC membrane}

Equilibration of \ce{Ca^{2+}} ions at a POPC bilayer in MD simulations is a microsecond time scale process with current force fields, such  as CHARMM36 and Slipids force fields~\cite{javanainen17}. This suggests that at least several microseconds are required to reach the ion binding/unbinding equilibrium.
To quantify the exchange of ions between the membrane and aqueous solution in simulations, we evaluated residence times of ions bound to the membrane. Within our analysis, an ion is considered to be bound when it is within 0.3~nm from any oxygen atom belonging to a POPC molecule.

The histograms of residence times of \ce{Ca^{2+}} in a POPC bilayer ($C_{ion}'$ = 450~mM) from simulations with 
ECC-POPC and CHARMM36 (simulation from Refs.~\citenum{javanainen17,zenodo.259376}) are shown in Fig.~S4 in SI.
In the CHARMM36 simulation, a significant number of the calcium ions is bound to the membrane for the whole length of the trajectory (800~ns).
In contrast, at least an order of magnitude faster bound/unbound calcium exchange is observed within the ECC-POPC model,
where 90\% of the \ce{Ca^{2+}} residence times to a POPC membrane are shorter than $60\,\mathrm{ns}$. The longest observed
residence time is around 150~ns, which is below the total length of the simulation used for analysis, i.e., 200~ns.
Note that these results are in line with the experimental estimate that the residence time of \ce{Ca^{2+}} at each PC
headgroup is of the order of $10\,\mu\mathrm{s}$~\cite{altenbach84}. Note that the exchange of \ce{Na+} ions at the POPC membrane
is yet another order of magnitude faster, with 90\% of the residence times smaller than~1~ns and the longest residence time being~6~ns.

In summary, the results from the ECC-POPC model suggest that the exchange of calcium between the POPC bilayer and the solvent occurs within the $\sim$100~ns timeframe, which is significantly faster than observed in simulations emloying most of the presently available lipid models~\cite{javanainen17}. Sodium cations exhibit an even more rapid exchange between the membrane and the aqueous solution. Our results suggest that simulations with a length of several hundreds of nanoseconds are sufficient to simulate alkali and alkali earth ion binding to phospholipid bilayers in equilibrium when realistic force fields are used. This has not been the case with previous lipid force fields, which overestimate the binding strength of the sodium and, in particular,  calcium cations \cite{javanainen17, catte16}.

\section{Conclusions}

In this study, we employed the electrometer concept to demonstrate that the binding of \ce{Na^+} and \ce{Ca^2+} ions to a POPC lipid bilayer can be accurately described within a force field MD simulation, provided that electronic polarization is implicitly included via the electronic continuum correction~\cite{leontyev11}. This ECC-POPC model is built upon the Lipid14 POPC model~\cite{dickson14} by scaling the partial charges by a factor of 0.8 and reducing the Lennard-Jones radii by a factor of 0.89 for the headgroup, glycerol backbone, and carbonyl atoms. While the structural details of a POPC lipid bilayer in pure water simulated with the newly developed ECC-POPC model agree with experiments with an accuracy comparable to the other state of the art lipid models, the new model in addition reproduces the experimental lipid head group order parameter responses to a cationic surfactant, NaCl, and CaCl$_2$ at varying concentrations. It thus represents a significant improvement over currently available lipid models, which tend to overestimate cation binding affinities~\cite{catte16}.

The good agreement with experiments enables us to interpret NMR experiments with atomistic details using MD simulations with the ECC-POPC model. In line with previous interpretations of the experimental data \cite{hauser76,hauser78,herbette84,binder02}, Ca$^{2+}$ ions interact mainly with phosphate oxygens, with interactions to the carbonyls being of a secondary importance. However, the stoichiometry of calcium binding is significantly more complicated than the simple ternary complex model, used to interpret NMR data previously, within which one calcium binds to two POPC molecules~\cite{altenbach84}. While complexes with one calcium ion bound to two lipids are the most probable also in the ECC-POPC model, complexes of one or three lipids per one calcium were observed to be relatively abundant and also almost equally likely to occur. While the success of the simple ternary complex model in fitting NMR data is understandable based on the simulation results, it cannot capture the detailed nature of calcium binding to phospholipid bilayers observed in the present simulation.

Accurate description of cation binding to POPC bilayer paves the way for simulations of complex biochemical systems at cellular membranes with realistically described electrostatic interactions, including a mean-field account for electronic polarization effects. To this end, the compatibility of the ECC-POPC model with existing models for proteins and other biological molecules needs to be verified and, if necessary, further adjustments following the ECC concept to the force fields of the biomolecules interacting with the lipids should be made.

This work can be reached as a repository containing all data and scripts for analysis at Ref. \citenum{github}.

% If you have acknowledgments, this puts in the proper section head.
\begin{acknowledgement}
% Put your acknowledgments here.
P.J. acknowledges support from the Czech Science Foundation (grant no. 16-01074S) 
and from the Academy of Finland via the FiDiPro award.
J.K. acknowledges support from the Czech Science Foundation, Project No. 15-12386S.
Computational resources were supplied by the Ministry of Education, Youth and Sports
of the Czech Republic under the Projects CESNET (Project No. LM2015042) and CERIT-Scientific
Cloud (Project No. LM2015085) provided within the program Projects of Large Research,
Development and Innovations Infrastructures.
O.H.S.O. acknowledges financial support from
Integrated Structural Biology Research Infrastructure of
Helsinki Institute of Life Science (Instruct-HiLIFE).
\end{acknowledgement}

\begin{suppinfo}

%A listing of the contents of each file supplied as Supporting Information
%should be included. For instructions on what should be included in the
%Supporting Information as well as how to prepare this material for
%publications, refer to the journal's Instructions for Authors.

The following files are available free of charge.
\begin{itemize}
  \item SI.pdf: Simulation details and further analysis.
%  \item Filename: brief description
\end{itemize}

\end{suppinfo}

\bibliography{refs.bib}

\end{document}

### 2018-04-09 scripts/atomic_partial_charges_compar_plot_l14_slipids_C36/molecule_mapper_plots_charges.py

@@ -1,0 +1,257 @@
+# -*- coding: utf-8 -*-
+"""
+reorders atoms in molecules
+according to the names in a reference molecule
+
+and prints out their partial charges in plots.
+
+IN: itp files (hard-coded in)
+
+Originally developed to be used with NMRlipids project 6
+by
+J. Melcr
+"""
+
+import pmx
+import pmx.forcefield
+import numpy as np
+import matplotlib.pyplot as plt
+#import math
+
+#%%
+
+def fetch_atom(mol, at_name):
+    """
+    return the first occurence of an atom with a given name.
+    return the 1st atom if searched atom is not found.
+    """
+    try:
+        found = False
+        for otherat in mol.atoms:
+            if otherat.name == at_name:
+                found = True
+                #print "Found atom {atname}.".format(atname=at_name)
+                return otherat
+    except:
+        print "Something went wrong during atom searching (function fetch_atom). \
+        \nIs mol a pmx.forcefield.ITPfile object?"
+    finally:
+        if not found:
+            print "Atom {atname} not found! -- will substitute it with atom no.1.".format(atname=at_name)
+            # so that there's no missing space in the sequence.
+            return mol.atoms[0]
+
+
+def autolabel(rects, i_max=None, font_size=2.5, q_label_offset=0.03):
+    """
+    Attach a text label above/below each bar displaying its height
+    This implementation works with a GLOBAL variable ax -- QUICK&DIRTY HACK!!
+    """
+    if i_max == None:
+        i_max = len(rects)
+    for i,rect in enumerate(rects):
+        if i<i_max :
+            height = rect.get_height()
+            if rect.get_y() >= 0:
+                ax.text(rect.get_x() + rect.get_width()/2., rect.get_y() + height + q_label_offset,
+                    '%2.2f' % height,
+                    ha='center', va='bottom',
+                    fontsize=font_size,
+                    rotation=90)
+            else:
+                ax.text(rect.get_x() + rect.get_width()/2., rect.get_y() - q_label_offset,
+                    '-%2.2f' % height,
+                    ha='center', va='top',
[…]

### 2018-04-19 Manuscript/manuscript.tex
First pilot content added to the manuscript + figures for both Lipid14/17 and ECC-lipids + some results and references + Makefile for an easy manuscript generation

@@
 \newcommand*\mycommand[1]{\texttt{\emph{#1}}} 
+\newlength{\figwidth}
+\setlength{\figwidth}{8 cm}
 
@@
 \author{Josef Melcr} 
-\author{Hector Martinez-Seara}
+\email{melcr@uochb.cas.cz} 
+%%\homepage[]{https://jmelcr.github.io/}
 \affiliation{Institute of Organic Chemistry and Biochemistry, 
@@
 Prague 6, Czech Republic} 
-\author{Ji{\v r}{\' i} Kolafa}
-\affiliation{Department of Physical Chemistry, Institute of Chemical Technology, Prague 6, Czech Republic}
 \author{Pavel Jungwirth} 
+%%\homepage[]{http://jungwirth.uochb.cas.cz/}
 \affiliation{Institute of Organic Chemistry and Biochemistry, 
@@
 \email{samuli.ollila@helsinki.fi} 
-%\homepage[]{Your web page}
+%%\homepage[]{Your web page} 
 \affiliation{Institute of Organic Chemistry and Biochemistry, 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
-\title[An \textsf{achemso} demo]
-  {Binding of cations to Neutral and Negatively Charged Phospholipid Bilayers 
-   made accurate through the use of 
-   classical fixed-charge MD models with implicit polarizability }
+\title[] 
+  {Accurate interactions of cations with 
+   neutral and negatively charged membranes 
+   via combination of experiments and molecular simulation}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
-\abbreviations{MD,IR,NMR,UV}
-\keywords{American Chemical Society, \LaTeX}
+\abbreviations{IR,NMR,UV,MD,ECC,PC,PS,POPS,POPS} 
+\keywords{MD simulation, molecular modeling, 
+          polarizability, polarization,
+          phospholipids, phosphatidylserine}
 
@@
 \begin{abstract} 
-Binding affinities and stoichiometries of Na$^+$ and Ca$^{2+}$ ions to phospholipid
-bilayers are of paramount significance in the properties and functionality of cellular
-membranes. Current estimates of binding affinities and
-stoichiometries of cations are, however, inconsistent due to limitations in the available experimental and
-computational methods. In this work, we improve the description of the binding details of Na$^+$ and Ca$^{2+}$ ions to the
-1-Palmitoyl-2-oleoyl-phosphatidylcholine (POPC) bilayer by implicitly including electronic
-polarization as a mean field correction, known as the electronic continuum correction (ECC).
-This is applied by scaling the partial charges of a selected state-of-the-art POPC lipid model for molecular dynamics simulations.
-Our improved ECC-POPC model reproduces not only the
-experimentally measured structural parameters for the ion-free membrane, but also the
-response of lipid head group to a strongly bound cationic amphiphile, as well as the binding affinities of 
-Na$^+$ and Ca$^{2+}$ ions. With our new model we observe on the one side negligible binding of Na$^+$ ions to POPC
-bilayer, while on the other side stronger interactions of Ca$^{2+}$ primarily with phosphate oxygens, which is
-in agreement with the previous interpretations of the experimental spectroscopic data.
-The present model results in Ca$^{2+}$ ions forming complexes with one to three POPC molecules with almost
-equal probabilities, suggesting more complex binding stoichiometries than those from simple models
-used to interpret the NMR data previously.
-The results of this work pave the way to quantitative molecular simulations with realistic electrostatic interactions 
-of complex biochemical systems at cellular membranes.
+Binding affinities and stoichiometries of sodium and calcium ions to phospholipid bilayers are of paramount significance in the properties and functionality of cellular membranes.
+Current estimates of binding affinities and stoichiometries of cations are, however, controversial due to limitations in the available experimental and theoretical methods.
+Experimentally one can assess these parameters by titrating several membrane properties as a function of the ion concentration.
+However, the interpretation of the experiments relies on theoretical models, as direct molecular information is not available.
+Classical molecular dynamics (MD) simulations provide details of the ion binding process with atomistic resolution, therefore offering all the necessary information to interpret experimental data without the need to resort to over-simplified models.
+However, the accuracy of the currently available lipid models when interacting with ions is not sufficient for such interpretation \cite{catte16}.
+Recently, we have improved the binding details of sodium and calcium ions to the 1-Palmitoyl-2-oleoyl-phosphatidylcholine (POPC) bilayer by implicitly including electronic polarization as a mean field correction, known as the electronic continuum correction (ECC).
+Our improved ECC-POPC model reproduces not only the experimentally measured structural parameters for the ion-free membrane, but also improves significantly the response of lipid head group to a bound positive charge, and the binding affinities of sodium and calcium ions.
+\cite{melcr18}
+We are now extending our work to other lipid membranes composed of larger varieties of different biologically relevant lipids.
+In particular, we focus our attention on membranes with phosphatidylethanolamine (PE) and negatively charged phosphatidylserine (PS).
+We compare available NMR and x-ray experiments with MD simulations to obtain atomistic-resolution picture of the interactions between cations and phospholipid membranes.
+Our preliminary results indicate that electronic polarization is a non-negligable interaction, which dramatically improves the description of cation binding to the membranes.
 \end{abstract}
@@
 \section{Introduction}
-Interactions of ions with cellular membranes play a key role in many important biological processes~\cite{seelig90,cevc90}.
-Ions, especially multivalent cations, modify general properties of the membrane which also modulate the embedded transmembrane proteins~\cite{cevc90,tocanne90, pabst07}.
-For example, Ca$^{2+}$  is crucial in propagation of neural signals.
-It promotes membrane fusion by bridging lipids in the vesicles carrying neurotransmitters and the neuron synapsis membrane~\cite{papahadjopoulos90,brouwer15}.
-Ca$^{2+}$ also participates in the T-cell receptor activation.
-It induces the detachment of the positively charged cytosolic tails of the CD3 protein complex from the negatively charged intracellular membrane making it accessible for the Lck protein~\cite{Shi2013Ca2}.
-Another example involving sugar-lipid interactions mediated by ions concerns Ca$^{2+}$ modulating the presentation of the sugars present in the PI(4,5)P$_2$ lipid which ultimately modulates the phospholipase C delta 1 pleckstrin homology domain (PLC $\delta$1-H)~\cite{Bilkova2017Calcium}.
-Interestingly, this effect is not present for Mg$^{2+}$, which illustrates the selectivity of these processes with respect to particular ions.
-Despite our increasing understanding of the role of ions in cell membrane-related processes the molecular details of the mechanisms behind such processes remain elusive in many cases.
-
-Direct measurements of ion-membrane interactions in complex biological systems are difficult. 
-Hence, simplified lipid bilayers are often used as entry-level models to shed light on the role of ions in complex biological membranes~\cite{scherer87,seelig90,cevc90}. 
-For this reason, interactions of biologically relevant cations, in particular Na$^+$ and Ca$^{2+}$, with zwitterionic phosphocholine (PC) bilayers have been widely studied experimentally~\cite{akutsu81, altenbach84, seelig90, cevc90, tocanne90, binder02, pabst07, uhrikova08} and using molecular dynamics (MD) simulations~\cite{bockmann03, bockmann04, Berkowitz12, melcrova16, javanainen17}.
-The details of ion binding are, however, not fully consistent in the literature.
-Non-invasive spectroscopic methods, like nuclear magnetic resonance (NMR), scattering, and infrared spectroscopies mostly suggest that Na$^+$ ions exhibit negligible binding to PC lipid bilayers.
-In contrast, Ca$^{2+}$ is observed to specifically bind to a couple of PC molecules via their phosphate groups~\cite{hauser76, hauser78, herbette84, akutsu81, altenbach84, binder02, pabst07, uhrikova08}.
-Most atomistic resolution MD simulation models predict stronger membrane bindings of the cations than observed in experiments~\cite{catte16}.
-Namely, simulations report various degrees of Na$^+$ accumulation at the lipid interface~\cite{bockmann03} and for Ca$^{2+}$ strong binding to up to four PC lipids simultaneously, involving not only interactions with phosphates but also with carbonyl oxygens~\cite{bockmann04, melcrova16, javanainen17}.
-
-Recent studies within the NMRlipids project (\url{nmrlipids.blogspot.fi})~\cite{catte16} made an attempt to resolve these disagreements.
-A direct comparison of ion binding affinities to PC bilayers between simulations and experiments has been made possible using the electrometer concept~\cite{seelig87}.
-Namely, the changes in NMR order parameters of the head group upon addition of ions are directly compared to the MD simulations results.
-Analyzing massive amounts of data collected within an open collaboration approach, 
-it was concluded that the accuracy of the current state-of-the-art lipid models for MD simulations is not sufficient for a detailed interpretation of the interactions of cations with PC lipid bilayers~\cite{catte16}.
-
-In this work, we improve the description of cation binding to a POPC bilayer within MD simulations 
-via an implicit inclusion of electronic polarizability in the polar region of phospholipids. 
-To this end, we apply the electronic continuum correction (ECC)~\cite{leontyev11}, 
-within which scaling of the atomic partial charges is employed to account for the missing electronic polarizability in standard force field MD simulations. 
-Such an approach has been shown previously to improve the behavior of ions in simulations of aqueous salt solutions~\cite{martinek17, Pluharova2014, kohagen14, kohagen16}.
-Here, we extend the ECC approach to aqueous lipid membranes taking the Lipid14 force field~\cite{dickson14} as a starting point for refining the POPC model.
-This choice is justified by the fact that the Lipid14 provides the least inaccurate descriptions of cation binding among the existing force fields\cite{catte16}.
-The newly developed ECC-POPC model reproduces the experimentally
-measurable structural parameters of an ion-free POPC lipid bilayer with the accuracy comparable to the best state-of-the-art lipid models, while at the same time significantly improving the membrane binding affinities of sodium and calcium cations.
 
@@
 \subsection{Electronic continuum correction for lipid bilayers}\label{section:ecc} 
-The lack of electronic polarizability in standard MD force fields has been considered a serious issue since the early days of lipid bilayer simulations.
-In this work, we circumvent the demanding explicit inclusion of electronic polarization effects~\cite{lucas12, chowdhary13} by accounting for the electronic part of polarizability in lipid bilayer simulations implicitly via the electronic continuum correction (ECC)~\cite{leontyev11}.
-Technically, this is similar to the phenomenological charge-scaling applied in earlier studies of surfactants, lipids or ionic liquids \cite{jonsson86, egberts94, beichel14}.
-However, the present concept of ECC is physically well justified and has theoretical support~\cite{leontyev09, leontyev10, leontyev11, leontyev14}.
+This section may not be required, we may just cite ECC-POPS 
+and put a short paragraph describing what we did on POPS.
 
-According to ECC, the electronic polarizability can be included into non-polarizable MD in a mean-field way by 
-embedding the ions in a homogeneous dielectric continuum with a dielectric constant $\epsilon_{el}$, which is the electronic part of the dielectric constant of the medium~\cite{leontyev11}.
-Following Coulomb's law,  ECC can be directly incorporated by scaling the charges with a constant scaling factor of $f_q = \epsilon _{el} ^{-1/2}$, yielding
-\begin{equation}
-Q^{ECC} = f_q \cdot Q
-\end{equation}
-for the ECC corrected charges. 
-Given that the  high frequency dielectric constant of water is $\epsilon _{el} = 1.78$ (\textit{i.e.,} the square of the refraction index), the scaling factor for ions in water is roughly $f_q \approx 0.75$.
-This scaling factor significantly improves the accuracy of simulations of solvated ions, when quantitatively compared with neutron scattering data \cite{kohagen14,kohagen16, Pluharova2014, martinek17}.
-It is important to note that the value of the
-high frequency dielectric constant 
-is around 2 for almost any biologically relevant environment \cite{leontyev11}.
-The dielectric discontinuity in a lipid bilayer thus arises only
-from the orientational polarization of the molecules, which is accounted for explicitly in standard MD simulations. 
-Therefore, the same correction for the electronic polarizability can be 
-applied throughout the lipid bilayer/aqueous solution interface.
-
-While using the scaling factor of $f_q = 0.75$ for ions in water is well justified in the ECC theory \cite{leontyev11}, it is not clear whether the same factor should be applied to partial charges used to describe molecules in MD models, e.g., lipids in our case \cite{leontyev14}.
-Unlike the total charge of an ion, atomic partial charges within molecules are not physical observables.
-Several computational schemes exist for the assignment of partial charges for biomolecules~\cite{Hu2007}, with the restrained electrostatic potential method (RESP) being commonly used~\cite{RESP_paper, Singh1984}.
-Considering that water is often included in RESP calculations or charges are refined to improve certain experimental observables, the electronic polarizability effects of the solvent may to some extent be included in standard force fields~\cite{RESP_paper, Singh1984, jorgensen96, ipolq2013, benavides17}.
-Thus, our application of the scaling factor, $f_q$, to existing partial charges in molecules does not necessarily follow $f_q = \epsilon _{el} ^{-1/2}$.
-Instead, a consistent scaling factor should lie between the value of 0.75  (\textit{i.e.,} no electronic polarizability included in the original partial charges) and 1 (\textit{i.e.,} electronic polarizability fully included in the original partial charges).
-
-Here we develop a ECC-POPC lipid model that accurately describes binding 
-of sodium and calcium ions to the POPC  lipid bilayer. 
-The Lipid14~\cite{dickson14} force field 
-(available in a Gromacs format from Ref. \citenum{lipid14files}) was used as a starting 
-point since 
-it provides the most accurate response of the head groups to ions among the available 
-lipid models (see Figs.~2~and~5 in Ref.~\citenum{catte16}). Additionally, the Lipid14 model 
-provides relatively realistic head group, glycerol backbone, and acyl chain structures~\cite{dickson14,botan15}.
-We applied the ECC correction 
-to the Lipid14 model of POPC by scaling 
-partial charges of the head group, glycerol 
-backbone, and carbonyl regions. 
-These are the polar parts of phospholipids which can 
-contribute to the cation binding.
-
-To reproduce the experimental ion binding affinities,
-we scanned possible values of the scaling factor from the interval $f_q~\in~(0.75, 1.0)$.
-The ion binding affinity was benchmarked against experiments
-using the head group order parameters and the electrometer concept \cite{seelig87,catte16},
-as discussed more detail in the next section.
-Scaling down the partial charges reduced the ion binding affinity.
-We found the most accurate ion binding affinities with a scaling factor of $f_q = 0.8$,
-which is only slightly higher than the ECC one ($f_q=0.75$).
-Note that common empirical scaling factors for monovalent ions in water are 0.8 or
-even closer to unity \cite{benavides17,skinner14,nacleps}.  In contrast, modern force fields
-for ionic liquids suggest values of 0.6--0.65, which are lower than $\epsilon^{-1/2}_{el}$\cite{holm14}.
-Directly applying
-the 0.8 scaling to the partial charges of the head group, the glycerol backbone, and
-the carbonyls reduced the area per lipid to 60~\AA$^2$. This area is smaller than in the
-original Lipid14 model ($65.6 \pm 0.5$~\AA$^2$)~\cite{dickson14} and in experiments
-(64.3~\AA$^2$) \cite{kucerka11}. The decrease of the area per lipid arises from a
-reduced hydration of the lipid head group region after scaling of the partial charges, which effectively
-reduces the head group polarity. We solve this problem by slightly reducing the effective radii of
-the modified head group atoms via lowering the $\sigma$ parameters in the Lennard-Jones potential by a
-factor of $f_\sigma = 0.89$. The same approach was successfully adopted for the ECC ions in aqueous
-solutions previously \cite{kohagen14, kohagen16, Pluharova2014, martinek17}.
-After reducing the head group atoms $\sigma$ parameters, the area per molecule is restored to the experimental value (Table~\ref{tab:apls}).
+Compared to ECC-POPC model \cite{melcr18}, 
+which used used a scaling factor $f_q = 0.8$ for all charges but lipid tails,
+this model, ECC-POPS, has to adopt a scaling factor $f_q = 0.75$ 
+as it has non-zero total charge in line with ECC theory. \cite{leontyev09}
+Hence, electronic continuum correction of POPS 
+has the same characteristics as for cations and anions, 
+denoted as ECC-ions \cite{martinek17, Pluharova2014, kohagen14, kohagen16}.
 
 \subsection{Electrometer concept} \label{section:electrometer} 
-Comparing MD simulation to NMR experiments, we can validate the ion
-binding affinity in lipid bilayer simulations using the 'electrometer concept'~ \cite{seelig87, catte16}.
-This method is based on the experimental observation that the C-H bond order parameters of $\alpha$ and $\beta$ carbons in a PC lipid head group (Fig.~\ref{simVSexpNOions}) are proportional to the amount of charge bound per lipid~\cite{seelig87}.
-The order parameters for all C-H bonds in lipid molecules can be accurately measured using $^2$H NMR or $^{13}$C NMR techniques \cite{ollila16}.
-From MD simulations the order parameters can be calculated using the relation
-\begin{equation}\label{OP}
-S_{\rm CH}=\frac{1}{2}\langle 3\cos^2\theta -1 \rangle,
-\end{equation}
-where $\theta$ is the angle between the C-H bond and the membrane normal. Angular brackets point to the average over all sampled configurations.
-
-The relation between the amount of the bound charge per lipid, $X^\pm$, and the head group order parameter change, $\Delta S_{\rm{CH}}^{i}$, is empirically quantified as~\cite{seelig87,ferreira16}
-\begin{equation}\label{OPchangeEQ}
-\Delta S_{\rm{CH}}^{i}= S_{\rm{CH}}^{i}(X^\pm)-S_{\rm{CH}}^{i}(0) \approx m_i \frac{4 }{3\chi}X^\pm,
-\end{equation}
-where $i$ refers to either the $\alpha$ or $\beta$ carbons, $S_{\rm{CH}}^{i}(0)$ denotes the order parameter in the absence of bound charge, $\chi$ is the quadrupole coupling constant ($\chi \approx$\,167\,kHz), and $m_i$ is an empirical constant depending on the valency and location of the bound charge.
-
-The measured change of the order parameter depends on the head group response to the bound charge and on the amount of the bound charge (\textit{i.e.,} $m_i$ and $X^\pm$ in Eq.~\ref{OPchangeEQ}, respectively).
-The empirical factor $m_i$ has to be adequately quantified before the electrometer concept can be used to analyze the binding affinities.
-This calibration has been done experimentally for a wide range of systems~\cite{seelig87, beschiasvili91}.
-To calibrate the response of the head group order parameters to the bound charge in simulations, we use experimental data for a strong cationic surfactant dihexadecyldimethylammonium bromide  (DHAB) mixed with a POPC bilayer~\cite{scherer89}.
-DHAB\\[0.5cm]
-%\vspace{0.5cm}
-\chemfig{
-  -[:0,3.5,,,draw=none]\chemabove{N}{\scriptstyle\oplus} (-[:150]H_3C)(-[:210]H_3C)(-[:330]{(}CH_2{)}_{15}-CH_3)(-[:30]{(}CH_2{)}_{15}-CH_3)
-}  
-\vspace{0.5cm} \\
-is a cationic surfactant having two acyl chains and bearing a unit charge at the hydrophilic end.
-Due to its structure it is expected to locate in the bilayer similarly to the phospholipids and its molar ratio then gives directly the amount of bound unit charge per lipid $X^\pm$ in these systems~\cite{scherer89}.
-
-\subsection{Salt concentrations and binding affinities}
-When measuring head group order parameters, the NMR experiments report the employed salt concentrations in two different ways.
-Some studies provide the salt concentrations in water before solvating the lipids~\cite{akutsu81}, $C_{ion}'$,
-while others use atomic absorption spectroscopy and report the salt concentration in
-the supernatant after the solvation of lipids~\cite{altenbach84}, $C_{ion}$. In this work,
-we stick to the latter definition by estimating the salt concentration in the aqueous
-bulk region using the farthest point from the lipid bilayer in the aqueous phase.
-Note that the former definition was used by some of us previously.~\cite{catte16}.
-Although these two definitions provide somewhat different values when applied
-to Ca$^{2+}$ concentrations, their particular choice does not affect the conclusions of this work in any significant way.
-
-To quantify the ion binding affinities to a lipid bilayer, we calculate the relative surface excess of ions with respect to water, $\Gamma_{ion}^{water}$~\cite{chattorajBOOK}.
-Such a quantity compares the adsorption of ions to the adsorption of water molecules at the interface without the necessity of defining a Gibbs dividing surface between the membrane interior and the water bulk region.
-In our simulations, we only assume that the interface is located between the ion-free hydrophobic interior of the lipid bilayer and the aqueous region far from the membrane.
-Such a setup and the above definition of bulk ion concentrations provides a simple relation for the relative surface excess $\Gamma_{ion}^{water}$ for simulations of lipid bilayers,
- \begin{equation}\label{surfexcess}
-  \Gamma_{ion}^{water}=\frac{1}{2A_b} \left ( n_{ion} - n_{water} \frac{C_{ion}}{C_{water}} \right ).
-\end{equation}
-Here, $n_{water}$ and $n_{ion}$ are the total numbers of water molecules and ions in the system,
-$C_{water}$ and $C_{ion}$ are their respective bulk concentrations in the aqueous phase
-(estimated from the solution density of farthest point between periodic bilayers),
-and $A_b$ is the area of the unit cell in the membrane plane.
-The total area of the interface is then twice the area of the membrane, i.e., $2A_b$,
-because the bilayer has an interface at each of the two leaflets.
+This section may not be required, 
+we may just cite ECC-POPS and NMRlipids projects papers (and ongoing works)
 
 \subsection{Validation of lipid bilayer structure against NMR and scattering experiments} 
+\emph{This section probably has to stay. Following text to be modified. Taken from Ref.~\citenum{melcr18}}
+
 The structures of lipid bilayers in simulations without ions were validated against NMR by calculating the order parameters for the C-H bonds and against \mbox{X-ray} scattering experiments by evaluating the scattering form factors.
 NMR order parameters validate the structures sampled by the individual lipid molecules with atomic resolution. 
-The simulated order parameters were calculated for all C-H bonds in lipid molecules from Eq.
-\ref{OP}.
-Scattering form factors validate the dimensions of the lipid bilayer (i.e., the bilayer thickness and area per molecule).
-Form factors were calculated using a relation
+The simulated order parameters were calculated for all C-H bonds in lipid molecules from Eq. \ref{OP}.
+Scattering form factors validate the dimensions of the lipid bilayer (i.e., the bilayer thickness and area per molecule). Form factors were calculated using a relation
 
@@
 
-\subsubsection{Simulations of POPC bilayers with aqueous ions}
-Simulations of a POPC bilayer in pure water or at varying salt concentrations consisted of 128 POPC molecules
-and approximately 50~water molecules per lipid in an orthorhombic simulation box with periodic boundary conditions.
-The SPC/E~\cite{Berendsen1987} water model was used in all ECC-POPC model simulations reported
-in the main text. This model was also used in the previous parametrization
-of ECC-ions \cite{martinek17, kohagen16, Pluharova2014} because its lowered dielectric constant is consistent with
-the ECC concept~\cite{leontyev11, leontyev14}.
-The robustness of our approach with other water models (OPC \cite{Izadi14},
-OPC3~\cite{Izadi16}, TIP3P~\cite{jorgensen83}, TIP3P-FB and TIP4P-FB~\cite{Wang2014}, and TIP4P/2005 \cite{Abascal2005})
-was verified and is presented in the Supporting Information (SI).
-Sodium, calcium, and chloride ions were modeled as ECC-ions
-with parameters from Refs. \citenum{martinek17, Pluharova2014}. Scaled charges and Lennard-Jones radii for atoms
-forming the POPC lipid were derived in this work starting from the Lipid14 force field~\cite{dickson14}.
-For comparison, simulations with the Lipid14 model~\cite{dickson14} and
-ion models by Dang and coworkers~\cite{smith94,chang1999,dang2006} or ECC-ions \cite{martinek17, kohagen16, Pluharova2014}
-were also performed. The TIP3P water model~\cite{jorgensen83} was used in all simulations with the original Lipid14 model.
-Simulation data for the Lipid14 model with \AA{}qvist ions~\cite{aqvist90} and the TIP3P~\cite{jorgensen83} water model were taken directly from
-Refs.~\citenum{catte16,lipid14POPC0mMNaClfiles,lipid14POPC350mMCaClfiles,lipid14POPC350mMCaClfilesNC,lipid14POPC1000mMNaClfiles}.
+General simulation setup.
 
-MD simulations were performed using the GROMACS \cite{Abraham15} simulation package (version 5.1.4). The simulation parameters used
-in this work are summarized in Table~S1 in the SI. 
-Compatibility with the openMM simulation package \cite{openmm7} was also
-tested and is presented in the SI. 
-For deposited trajectories and parameter files see 
-Refs.~\citenum{lipid14_POPC_Dang_and_ECC-ions_files, ECC-POPC_nacl_cacl2_files, ECC-POPC_nacl_cacl2_wat-models_files}.
+\subsubsection{Simulations of POPS:POPS bilayers with aqueous ions}
 
-\subsubsection{Simulations of POPC bilayers with cationic surfactants}
-The topology of the dihexadecyldimethylammonium cation was created with the automated topology builder \cite{malde11}.
-The AmberTools program~\cite{amber} was used to generate the Amber-type force field parameters.
-These parameters were then converted to the Gromacs format with the acpype tool \cite{acpype}.
-The partial charges were manually modified in the non-scaled version to approximately match similar segments in Lipid14 \cite{dickson14}.
-For simulation with ECC-POPC, the partial charges of the head group of the cation were scaled by a factor of $f_q=0.8$, and the $\sigma$ of the scale atoms reduced by a factor of 0.89 similarly as for POPC-ECC.
+Any specifics for simulations with cations.
 
-The cationic surfactants were randomly mixed among the phospholipids to form bilayer structures with mole fractions of 10\%, 20\%, 30\%, 42\%, or 50\% of surfactant in the POPC bilayer.
-All these systems contained 50 POPC molecules per leaflet, 6340 water molecules, and 6, 14, 21, 35, or 50 surfactants per leaflet.
-Chloride counter ions were used in all the simulations, while bromide was the counter ion in the experiment~\cite{scherer89}.
-For POPC, either Lipid14 with TIP3P~\cite{jorgensen83} or ECC-POPC with SPC/E~\cite{Berendsen1987} parameters  were used.
-The first 20~ns of the total simulation time of 200~ns was considered as an equilibration period and was thus omitted from the analysis.
-We checked that a sufficient lipid  exchange occurred during the simulations.
-For deposited trajectories and parameter files of scaled and unscaled simulations see
-Refs.~\citenum{ECC-POPC_cationic-surf_files} and~\citenum{POPClipid14T313K,POPClipid1410perCATsurfT313K,POPClipid1420perCATsurfT313K,POPClipid1430perCATsurfT313K,POPClipid1442perCATsurfT313K,POPClipid1450perCATsurfT313K}, respectively.
+\subsubsection{Simulations of POPS:POPS bilayers with anionic/cationic surfactants/peptides}
+
+Any specifics for simulations with anionic/cationinc surfactants/peptides. 
+e.g. where the parameters for them came from.
+
+% ++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
+% ++++++++  RESULTS AND DISCUSSION section +++++++++++++++++++++++++++++++++++++++++++
+% ++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
 
@@
 
-\subsection{Structural parameters of a pure ECC-POPC model membrane: Agreement with experiments}
+\subsection{Structural parameters of a pure ECC-POPS model membrane: Agreement with experiments}
 
@@
   \centering 
-  \includegraphics[width=8.2cm]{../Fig/ipython_nb/Order-parameters_form-factors_exp-L14-ECCL17_q80_sig89_POPC-struct.pdf}
+  \includegraphics[width=\figwidth]{../Fig/order_parameters_actual_pure-POPS.pdf} \\
+  \includegraphics[width=0.7\figwidth]{../Fig/form-f_l17-ecc-pops-exp_compar.pdf} 
+  \includegraphics[width=0.55 \figwidth]{../Fig/lipids_chemfig.pdf} 
   \caption{\label{simVSexpNOions} 
-    Top: X-ray scattering form factors from simulations with the Lipid14 \cite{dickson14} and
-    the ECC-POPC models compared with experiments~\cite{kucerka11} at 303~K. \\
-    Middle: Order parameters of POPC head group, glycerol backbone and acyl chains 
-    from simulations with the Lipid14 \cite{dickson14} and the ECC-POPC models
-    compared with experiments \cite{ferreira13} at 300~K.
-    The size of the markers for the head group order parameters correspond to
-    the error estimate $\pm 0.02$ for experiments \cite{botan15,ollila16},
-    while the error estimate for simulations is $\pm 0.005$
-    (Bayesian estimate of 95\% confidence interval \cite{scipy}).
-    The size of the points for acyl chains are decreased by a factor of 3 to improve the clarity of the plot.
-    Open/closed symbols are used for palmitoyl/oleoyl chains of POPC. \\
-    Bottom: The chemical structure of POPC and the labeling of the carbon segments.
+    X-ray scattering form factors from simulations with the Lipid17 \cite{lipid17-future} and 
+    the ECC-POPS models compared with experiments~\cite{SDP-CHARMM36_comparison_paper_Samuli-knows} at 298~K. 
+    The disagreements between ECC-POPS model and experiment 
+    may come from an intrinsincally incorrect structure of the head group (and, hence, electron density in the region), 
+    however, lipid packing and overall membrane properties 
+    may be still caputured well enough -- after considering areas per lipid reported in Table~\ref{tab:apls}.  \\
+    Order parameters of POPS head group, glycerol backbone and acyl chains  
+    from simulations with the Lipid17 \cite{lipid17-future} and the ECC-POPS models 
+    compared with \emph{experiments at 300~K (check and change)}. 
+    Open/closed symbols are used for palmitoyl/oleoyl chains of POPS. \\
+    The chemical structure of POPS and the labeling of the carbon segments. 
   }  
@@
 \begin{table}[tb!] 
-  \caption{Values of the area per lipid (APL) of POPC bilayers without ions. \label{tab:apls}
+  \caption{Values of the area per lipid (APL) of POPS bilayers with only \ce{Na^+} counterions. \label{tab:apls} 
   } 
@@
     \hline 
-    Lipid14                   & 65.1$\pm$ 0.6  &  300 \\
-    Lipid14 \cite{dickson14}  & 65.6$\pm$ 0.5  &  303 \\
+    Lipid14                   & 53.5$\pm$ 0.8  &  298 \\ 
     \hline 
-    ECC-POPC                & 63.2$\pm$ 0.6  &  300       \\
+    ECC-POPS                & 60.3$\pm$ 0.6  &  298       \\ 
     \hline 
-    experiment \cite{kucerka11} & 64.3  &  303    \\
+    experiment \cite{SDP-CHARMM36_comparison_paper_Samuli-knows} & 62.3  &  298    \\ 
     \hline 
@@
 
-First, we present results for bilayers in pure water.
-The ECC-POPC and Lipid14 models both reproduce the experimental X-ray scattering form factors
-of a POPC bilayer with a comparable accuracy (see Fig.~\ref{simVSexpNOions}).
-The area per lipid from the Lipid14 model is by $\approx$1\AA{} larger than the
-experimental value in Table~\ref{tab:apls}, while the value from the ECC-POPC model
-is by $\approx$1\AA{} smaller than the experimental one.
-The values of the area per lipid of the ECC-POPC model vary slightly
-when simulated with different water models (i.e., within the interval of 62.2--66.8 \AA{}, see Table~S2 in SI),
-while still being close to the experimentally reported values.
-We can thus conclude that the ECC-POPC model reproduces the experimental dimensions of the POPC
-lipid bilayer with a comparable accuracy to other state-of-the-art lipid models~\cite{ollila16}.
+\subsection{Electrometer concept in negatively charged membranes with various cations}
 
-Similarly, the acyl chain order parameters of the ECC-POPC model, as well as those of the Lipid14 model~\cite{dickson14}, agree with the experimental values within the error bars, as presented in Fig.~\ref{simVSexpNOions}.
-Notably, the experimentally measured forking and small order parameter values of the $C_2$ segment in {\it sn}-2 chain are well reproduced by both models.
-This feature has been suggested to indicate that the carbonyl of the {\it sn}-2 chain is directed towards the water phase, in contrast to the carbonyl in the {\it sn}-1 chain, which orients more along the bilayer plane~\cite{seelig75,schindler75,gawrisch92}.
-This arrangement, which is not fully reproduced by other available lipid models~\cite{ollila16}, may be a relevant feature for the ion binding details.
+Preliminary results are plotted in the figures in this section. 
+Beware! -- the y-axis scales are variable!
+Colours:
+\begin{itemize}
+  \item Red -- ECC-lipids, ECC-ions
+  \item Blue -- Lipid14/17, Dang ions
+\end{itemize}
 
-The order parameters of the $\alpha$ and $\beta$ carbons in the head group are slightly larger in the ECC-POPC model than in the Lipid14 model, which is apparently related to the P-N vector orienting by about 7$^{\circ}$ more toward the water phase in the former model, see Fig.~\ref{OrderParameterCHANGESsurf}.
-While both models perform relatively well, considering the available experimental evidence, it is not possible to decide which of the two models provides more realistic head group orientations.
-The ECC-POPC model gives the $\beta$ carbon order parameter value closer to experiments than the Lipid14 model, while the opposite is true for the $\alpha$ carbon.
-The accuracy of both models in the glycerol backbone region is comparable to other state-of-the-art lipid model available in literature \cite{botan15}, see Fig.~\ref{simVSexpNOions}.
+Changes of the lipid bilayer head group order parameters extracted from simulations and 
+experiments \cite{roux90} are shown in Figs.~\ref{fig:delta_ordPar_KCl}, \ref{fig:delta_ordPar_NaCl} 
+and~\ref{fig:delta_ordPar_CaCl} as functions of KCl, NaCl or CaCl$_2$ concentrations,
+and in Figs.~\ref{fig:delta_ordPar_NaCl_PC-PS_mix} as a function of PC and PS content.
 
-\subsection{Calibration of head group response to membrane-bound charge using cationic surfactant}\label{section:boundCHARGE}
-
-\begin{figure}[tb!]
-  \centering
-  \includegraphics[width=8.0cm]{../Fig/ipython_nb/PN_angle_OrdPars-A-B_L14-ECCL17_q80_sig89_surf.pdf}
-  \caption{\label{OrderParameterCHANGESsurf}
-    The changes of head group order parameters and P-N vector orientation as a function of
-    a molar fraction of the cationic surfactant dihexadecyldimethylammonium in a POPC bilayer
-    from simulations and experiments \cite{scherer89} at 313 K.
-  }
-\end{figure}
-
-Before studying the sodium and calcium ion binding affinities, we quantify the response of the head group order parameters to the amount of bound charge by using mixtures of monovalent cationic surfactants (dihexadecyldimethylammonium) and POPC~\cite{scherer89}.
-These mixtures have a well-defined amount of bound charge per PC, namely, the molar fraction of cationic surfactants.
-This is due to the ability of dihexadecyldimethylammonium to directly insert in the lipid bilayer due to its two hydrophobic acyl chains.
-Furthermore, available experimental data for these systems can be used to validate the sensitivity of lipid head group order parameters to the amount of bound charge in simulations \cite{scherer89}.
-
-The changes of the head group order parameters with increasing amount of the cationic surfactant are compared between simulations and experiments~\cite{scherer89} in Fig.~\ref{OrderParameterCHANGESsurf}.
-An approximately linear decrease of the order parameters, as expected from Eq.
-\ref{OPchangeEQ}, is observed in both simulations and experiments at least for mole fractions below $\sim$30\%.
-The slope is, however, too steep in the Lipid14 model indicating that the response of head group order parameters to the bound positive charge is too strong.
-In contrast, the slope of the ECC-POPC model is in a very good agreement with experiments for the $\alpha$ segment, while being only slightly underestimated for the $\beta$ segment.
-
-In Fig.~\ref{OrderParameterCHANGESsurf}, we show the headgroup P-N vector angle as a function of the mole fraction of the cationic surfactant.
-As suggested previously~\cite{seelig87}, the headgroup orients more towards the water phase with the increasing amount of positive charge in the PC lipid bilayer.
-The effect is more pronounced in the Lipid14 model than in the ECC-POPC model.
-For example, the addition of 50\% mole fraction of the cationic surfactant leads to a decrease of 20$^{\circ}$ of the P-N vector angle for the Lipid14 model while only of 11$^{\circ}$ in the ECC-POPC model.
-The difference is in line with the smaller order parameter changes and the reduced charge--dipole interactions in the latter model.
-The weaker sensitivity of the P-N vector angle response in the ECC-POPC model is in better agreement with experiments.
-
-\subsection{Validation of ECC-POPC model using binding affinities to Na$^+$ and Ca$^{2+}$ cations: the electrometer concept}
-
+% ============================== 
+% ==  ECC-lipids plots ========= 
+% ==     and           ========= 
+% ==  Lipid14/17 plots ========= 
+% ============================== 
 \begin{figure}[htb!] 
   \centering 
-  \includegraphics[width=8.0cm]{../Fig/ipython_nb/OrdPars-A-B_L14-ECCL17_q80_sig89_NaCl.pdf}
-  \caption{\label{fig:delta_ordPar_NaCl}
-    Changes of the head group order parameters of a POPC bilayer as a function of NaCl concentration
-    in bulk ($C_{ion}$) from simulations with different force fields at 313 K together with 
-    experimental data for DPPC (323\,K) \cite{akutsu81} and POPC (313\,K) \cite{altenbach84}.
-    Simulation data with Lipid14 and \AA{}qvist ion parameters at 298 K are taken directly from
-    Refs.~\cite{lipid14POPC0mMNaClfiles,lipid14POPC1000mMNaClfiles}.
+  \includegraphics[width=\figwidth]{../Fig/l17/order_parameters_changes_A-B_POPC_kcl.pdf} 
+  \includegraphics[width=\figwidth]{../Fig/l17/order_parameters_changes_A-B_POPS_kcl.pdf} 
+  \caption{\label{fig:delta_ordPar_KCl_l17} 
+    Changes of the head group order parameters of a POPC:POPS 5:1 bilayer as a function of KCl concentration 
+    in bulk ($C_{ion}$) from simulations with different force fields at 298 K together with  
   } 
@@
   \centering 
-  \includegraphics[width=8.0cm]{../Fig/ipython_nb/PN_angle_OrdPars-A-B-g3_L14-ECCL17_q80_sig89_CaCl.pdf}
-  \caption{\label{fig:delta_ordPar_CaCl}
-    Changes of the head group order parameters and P-N vector orientation of a POPC bilayer 
-    as a function of the CaCl$_2$ concentration in bulk ($C_{ion}$)
-    from simulations at 313 K together with experimental data 
-    (DPPC (323\,K) \cite{akutsu81} and POPC (313\,K) \cite{altenbach84}). 
-    The error estimate for bulk concentrations is approximately 10\,mM.
-    The order of magnitude larger error in the
-    simulation with Lipid14 and ECC-ions is due to unconverged bulk densities  (shown if Fig.~\ref{fig:cacl-dens}) limited by
-    the simulation box. 
-    Simulation data with Lipid14 and \AA{}qvist ion parameters at 298 K are taken directly from
-    Refs.~\citenum{lipid14POPC0mMNaClfiles,lipid14POPC350mMCaClfiles,lipid14POPC350mMCaClfilesNC}.
+  \includegraphics[width=\figwidth]{../Fig/order_parameters_changes_A-B_POPC_kcl.pdf} 
+  \includegraphics[width=\figwidth]{../Fig/order_parameters_changes_A-B_POPS_kcl.pdf} 
+  \caption{\label{fig:delta_ordPar_KCl} 
+    Changes of the head group order parameters of a POPC:POPS 5:1 bilayer as a function of KCl concentration 
+    in bulk ($C_{ion}$) from simulations with different force fields at 298 K together with  
   } 
@@
 
-Changes of the lipid bilayer head group order parameters extracted from simulations and
-experiments \cite{akutsu81, altenbach84} are shown in Figs.~\ref{fig:delta_ordPar_NaCl}
-and~\ref{fig:delta_ordPar_CaCl} as functions of NaCl or CaCl$_2$ concentrations.
-As seen in Fig.~\ref{OrderParameterCHANGESsurf}, the order parameters decrease
-proportionally to the amount of the bound positive charge.
-These results can be thus used to compare the ion binding affinities to lipid bilayers between
-simulations and experiments using the electrometer concept~\cite{seelig87, catte16}.
+\begin{figure}[htb!] 
+  \centering 
+  \includegraphics[width=\figwidth]{../Fig/l17/order_parameters_changes_A-B_POPC_nacl.pdf} 
+  \includegraphics[width=\figwidth]{../Fig/l17/order_parameters_changes_A-B_POPS_nacl.pdf} 
+  \caption{\label{fig:delta_ordPar_NaCl_l17} 
+    Changes of the head group order parameters of a POPC:POPS 5:1 bilayer as a function of NaCl concentration 
+    in bulk ($C_{ion}$) from simulations with different force fields at 298 K together with  
+  } 
+\end{figure}
 
-The experimentally measured small order parameter
-changes with NaCl (Fig.~\ref{fig:delta_ordPar_NaCl}) 
-are reproduced by the Lipid14 model simulated with \AA{}qvist ions.
-However, the same combination of models overestimates the order parameter changes with CaCl$_2$ (Fig.~\ref{fig:delta_ordPar_CaCl}).
-Replacing \AA{}qvist ions with ion parameters by Dang et al.~\cite{smith94, chang1999, dang2006}
-or ECC-ions~\cite{martinek17, kohagen16, Pluharova2014} did not improve
-the results~(Figs.~\ref{fig:delta_ordPar_NaCl} and \ref{fig:delta_ordPar_CaCl}).
-In line with the previous work \cite{catte16}, the results suggest that improvements
-in the lipid parameters are required to correctly describe the binding of cations to phospholipid bilayers.
+\begin{figure}[htb!] 
+  \centering 
+  \includegraphics[width=\figwidth]{../Fig/order_parameters_changes_A-B_POPC_nacl.pdf} 
+  \includegraphics[width=\figwidth]{../Fig/order_parameters_changes_A-B_POPS_nacl.pdf} 
+  \caption{\label{fig:delta_ordPar_NaCl} 
+    Changes of the head group order parameters of a POPC:POPS 5:1 bilayer as a function of NaCl concentration 
+    in bulk ($C_{ion}$) from simulations with different force fields at 298 K together with  
+  } 
+\end{figure}
 
-The results from simulations combining the ECC-POPC with the ECC-ion models \cite{martinek17, kohagen16, Pluharova2014} exhibit a significantly improved behavior of the POPC head group order parameters as a function of NaCl or CaCl$_2$ concentrations, see Fig.~\ref{fig:delta_ordPar_NaCl} and Fig.~\ref{fig:delta_ordPar_CaCl}.
-Considering that we are also able to reproduce the experimental response in systems with known charge density (see above section \ref{section:boundCHARGE}), we conclude that our ECC model correctly reproduces the binding affinities of Na$^{+}$ and Ca$^{2+}$ ions to the POPC lipid bilayer.
-Furthermore, while the response of the glycerol backbone $g_3$ order parameter to CaCl$_2$ was significantly overestimated in the original Lipid14 model, the ECC-POPC model provides an improved agreement with experiment, as seen in Fig.~\ref{fig:delta_ordPar_CaCl}.
-Also the changes of the P-N vector angle are too pronounced for the Lipid14 model,
-for which the largest tilting toward water phase induced by a $780\,\mathrm{mM}$
-CaCl$_2$ concentration is approximately 17$^{\circ}$. The corresponding value
-for the ECC-POPC simulation is only 6$^{\circ}$ ($820\,\mathrm{mM}$ CaCl$_2$).
+\begin{figure}[htb!] 
+  \centering 
+  \includegraphics[width=\figwidth]{../Fig/l17/order_parameters_changes_A-B_POPC_cacl.pdf} 
+  \includegraphics[width=\figwidth]{../Fig/l17/order_parameters_changes_A-B_POPS_cacl.pdf} 
+  \caption{\label{fig:delta_ordPar_CaCl_l17} 
+    Changes of the head group order parameters of a POPC:POPS 5:1 bilayer as a function of CaCl concentration 
+    in bulk ($C_{ion}$) from simulations with different force fields at 298 K together with  
+  } 
+\end{figure}
 
-Within the Lipid14 model, the overestimated changes in the lipid headgroup order parameter of POPC  as functions of the CaCl$_2$ concentration arise both from the overestimated binding affinity and the excessive sensitivity of the headgroup tilt to the bound positive charge.
-It is plausible to assume that the same applies to the other lipid models tested in a previous study~\cite{catte16}, which underlines the importance of validation of the lipid headgroup order parameter response to the bound charge.
+\begin{figure}[htb!] 
+  \centering 
+  \includegraphics[width=\figwidth]{../Fig/order_parameters_changes_A-B_POPC_cacl.pdf} 
+  \includegraphics[width=\figwidth]{../Fig/order_parameters_changes_A-B_POPS_cacl.pdf} 
+  \caption{\label{fig:delta_ordPar_CaCl} 
+    Changes of the head group order parameters of a POPC:POPS 5:1 bilayer as a function of CaCl concentration 
+    in bulk ($C_{ion}$) from simulations with different force fields at 298 K together with  
+  } 
+\end{figure}
 
-Finally, the ion binding affinities for the ECC-POPC model with different water models are compared in SI.
-In general, the performance of ECC-POPC with any of the tested water models is better than that of the original Lipid14 model, with the order parameter changes being slightly overestimated with the four-site water models and with TIP3P model.
+\begin{figure}[htb!] 
+  \centering 
+  \includegraphics[width=\figwidth]{../Fig/l17/order_parameters_changes_A-B_PC-PS_mix_POPC_nacl.pdf} 
+  \includegraphics[width=\figwidth]{../Fig/l17/order_parameters_changes_A-B_PC-PS_mix_POPS_nacl.pdf} 
+  \caption{\label{fig:delta_ordPar_NaCl_PC-PS_mix_l17} 
+    Changes of the head group order parameters of a POPC:POPS 5:1 bilayer as a function of PS content.
+    Experimental points have inherent systematic error, as they come from different experiments.
+    A complete series of points can be found (?) in \cite{roux90} -- which is to be added to the plot.
+  } 
+\end{figure}
 
-\subsection{Binding affinities of Na$^+$ and Ca$^{2+}$ cations to the POPC membrane}
+\begin{figure}[htb!] 
+  \centering 
+  \includegraphics[width=\figwidth]{../Fig/order_parameters_changes_A-B_PC-PS_mix_POPC_nacl.pdf} 
+  \includegraphics[width=\figwidth]{../Fig/order_parameters_changes_A-B_PC-PS_mix_POPS_nacl.pdf} 
+  \caption{\label{fig:delta_ordPar_NaCl_PC-PS_mix} 
+    Changes of the head group order parameters of a POPC:POPS 5:1 bilayer as a function of PS content
+    Experimental points have inherent systematic error, as they come from different experiments.
+    A complete series of points can be found (?) in \cite{roux90} -- which is to be added to the plot.
+  } 
+\end{figure}
+
+\subsection{Molecular interaction and binding affinities of K$^+$, Na$^+$ or Ca$^{2+}$  cations to the mixed POPC:POPS (5:1) membrane} 
 \label{sec:affinity}
+
+Provide density plots and show, where and how the cations bind to the membrane.
+
+Where do the cations bind? Preferentially to POPS. Discuss further \dots
+
+After validation above, it can be an option to show results also for systems, for which we have no experiments.
 
@@
     and the percetages of Ca$^{2+}$ bound to phosphate or carbonyl oxyges ($r^\mathrm{Ca^{2+}} _\mathrm{PO_4} $ and $r^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}}$) 
-    in different POPC bilayer models. All systems have the same molar concentration of Ca$^{2+}$ with respect to water ($C_{ion}'$=350mM).
+    in different mixed POPC:POPS (5:1) bilayer models. 
   \label{tab:binding}} 
@@
     \hline 
-    ECC-POPC             &  350  &  $280\pm 10 $  &  $0.06 \pm 0.01 $                           &  99\%  &    25\%    \\
-    Lipid14/\AA{}qvist     &  350  &  $210\pm 10 $  &  $0.13 \pm 0.01 $                          & 100\%  &    37\%     \\
-    Lipid14/Dang           &  350  &  $160\pm 10 $  &  $0.23 \pm 0.03 $                            & 100\%  &    14\%    \\
-    Lipid14/ECC-ions       &  350  &  $120\pm 100$  &  $0.35 \pm 0.11 $                         & 100\%  &    23\%    \\
+    ECC-POPS               &  tba  &  $tba\pm 10 $  &  $0.?? \pm 0.01 $   &  ??\%  &    ??\%    \\ 
+    Lipid14/Dang           &  tba  &  $tba\pm 10 $  &  $0.?? \pm 0.03 $   &  ??\%  &    ??\%    \\ 
   \end{tabular} 
@@
 
-Binding affinities of Ca$^{2+}$ ions to a POPC bilayer in different simulation models were quantified by calculating the relative surface excess of calcium with respect to water molecules, $\Gamma_{\rm ion}^{\rm water}$, from Eq.~\ref{surfexcess}.
-The values of $\Gamma_{\rm ion}^{\rm water}$
-from different simulations with the same molar concetration of cations with respect
-to water ($C_{ion}'$=350mM) are shown in Table~\ref{tab:binding}.
-As expected from the changes of the lipid headgroup order parameters in Fig.~ \ref{fig:delta_ordPar_CaCl}, the relative surface excess of calcium, $\Gamma_{\rm Ca}^{\rm water}$ = 0.06~nm$^{-2}$, is significantly smaller for the ECC-POPC model than for the other models, 0.13--0.35~nm$^{-2}$.
-Interestingly, the calculated relative surface excess of \ce{NaCl} at $1\,\mathrm{M}$ concentration (ECC-ions~\cite{Pluharova2014}) using our ECC-POPC model is not only quantitatively but also qualitatively different from \ce{CaCl2} having actually a negative value of $\Gamma_{Na}^{water} = -0.11 \pm 0.01) \rm{nm}^{-2}$ (Fig.~\ref{fig:cacl-dens}.
-This
-means that on average water molecules are preferred to sodium and chloride ions at the membrane-water interface.  
-This is in contradiction with most of the available lipid force fields, which predict a positive surface excess of sodium at PC lipid bilayers \cite{catte16}.
+%\begin{figure}[htbp!] 
+%  \centering 
+%  \includegraphics[height=19.9cm]{../Fig/}
+%  \caption{\label{fig:cacl-dens} 
+%    Number density profiles of \ce{Ca^{2+}}, \ce{Na^{+}} and \ce{Cl^-} along membrane normal axis 
+%    for different force fields. 
+%    } 
+%\end{figure}
 
-\begin{figure}[htbp!]
-  \centering
-  \includegraphics[height=19.9cm]{../Fig/ipython_nb/density_profiles_ca_cl_wat_phos_models-compar.pdf}
-  \caption{\label{fig:cacl-dens}
-    Number density profiles of \ce{Ca^{2+}}, \ce{Na^{+}} and \ce{Cl^-} along membrane normal axis
-    for different force fields.
-    In order to visualize the density profiles with a scale comparable to the profile of \ce{Ca^{2+}}, 
-    the density profiles of~\ce{Cl^-} and \ce{Na^{+}} ions are divided by 2, and
-    the density profiles of phosphate groups and water are divided by 5 and 200, respectively. 
-    All simulations with \ce{CaCl2} shown here have the same molar concentration of ions in water ($C_{ion}'$=350~mM).
-    The simulation with \ce{NaCl} has $C_{ion}'$=1000~mM.
-    }
-\end{figure}
+%\subsection{Molecular interactions between K$^+$, Na$^+$ or Ca$^{2+}$ cations and mixed POPC:POPS membrane}
 
-\subsection{Molecular interactions between Na$^+$ or Ca$^{2+}$ cations and POPC oxygens}
-We analyzed the ratio of the number of calcium cations bound to either phosphate or carbonyl moieties and the total number of bound cations in our POPC bilayers as done previously in Ref.~\citenum{javanainen17}.
-A maximum distance of 0.3~nm from any lipid oxygen is used to define a bound calcium.
-The results from ECC-POPC simulation in Table~\ref{tab:binding} show that almost all (99\%) of the bound Ca$^{2+}$ ions are in direct contact with phosphate oxygens.
-From these ions, only one third (32\%) also interacts with the carbonyl oxygens, while the interaction of calcium ions with carbonyl oxygens only is rare (1\%).
-The most abundand interaction scenarios between Ca$^{2+}$ ions and phosphate oxygens are visualized using the probability density isocontours in Fig.~\ref{fig:volmaps}.
-While higher concentrations of \ce{CaCl2} increase the number of contacts per lipid, the distribution of contacts between phosphate and carbonyl oxygens is not affected.
+%\subsection{Binding stoichiometry of cations to mixed POPC:POPS membrane}
 
-\begin{figure}[tb!]
-  \centering
-  \includegraphics[width=8.0cm]{../Fig/isocontours_r37_ca_O-carb.png} %\\
-  \caption{\label{fig:volmaps}
-    Isocontours of spatial number density of \ce{Ca^{2+}} (dark blue, 0.001~\AA$^{-3}$) 
-    and POPC carbonyl oxygen atoms (light semi-transparent red, 0.008~\AA$^{-3}$, all POPC lipids contribute). 
-    Calcium cations localize mostly around phosphate oxygens (oxygens red, phosphorus bronze).
-    Interactions with carbonyl oxygens is less likely than with phosphate oxygens, 
-    and it is contributed more by other neighbouring phospholipids than by the same lipid. 
-    Transparent structures are shown to depict the variability of choline configurations 
-    (colour warps from red to blue along the simulation time). 
-    The number density was evaluated for each lipid, 
-    after its structural alignment using only phosphate group.
-    MDAnalysis \cite{mdanalysis2011} library was used for 
-    the calculations of the structural alignment and the spatial number density. 
-    VMD \cite{hump96} was used for visualisation. 
-    Carbon atoms are depicted in cyan, hydrogen atoms in white, oxygen atoms in red, nitrogen in blue.
-  }
-\end{figure}
+%\subsection{Residence times of cations in the mixed POPC:POPS membrane}
 
-Even though \ce{Na+} ions do not bind strongly to a POPC bilayer, they still interact mostly with its oxygen moieties.
-The results from a simulation at a $1\,$M \ce{NaCl} concentration show that 55\% of \ce{Na+} ions at the bilayer interact with phosphate oxygens of POPC only and 20\% with carbonyl oxygens only, with the remaining 25\%, is interacting with both negatively charged groups.
+%\begin{figure}[tb!] 
+%  \centering 
+%  \includegraphics[width=8.0cm]{../Fig/ipython_nb/stoichiometry_NaCl-CaCl2_comparison_Ecc-lipids.pdf} \\ 
+%  \caption{\label{fig:cacl_complexes} 
+%      Relative probabilities of existence of \ce{Na^{+}} or \ce{Ca^{2+}} complexes 
+%      with a certain number of POPS lipids.  
+%      \ce{Na^{+}} complexes were evaluated from the simulation with 1~M concentration; 
+%      and \ce{Ca^{2+}} complexes were evaluated from the simulation with 287~mM concentration. 
+%  } 
+%\end{figure}
 
-In conclusion, the results suggest that calcium ions bind specifically to the phosphate oxygens, occasionally interacting also with the carbonyls of the PC lipids.
-This is in a qualitative agreement with previous conclusions from several experimental studies~\cite{hauser76, hauser78, herbette84, cevc90, binder02}.
-However, the present results suggest, in
-agreement with experiments, an overally weaker binding to the bilayer, in particular with a lower relative binding affinity to the carbonyls than inferred from previous MD simulation studies~\cite{bockmann03, bockmann04, melcrova16, javanainen17}.
-Sodium ions, which do not exhibit any appreciable affinity for the bilayer, also interact primarily with phosphate oxygens of the POPC, but in contrast to calcium, the interactions purely with carbonyls are also significant.
+\subsection{Electronic Continuum Correction as a generally applicable model polarizability?}
 
-\subsection{Binding stoichiometry of \ce{Na+} and \ce{Ca^{2+}} cations to POPC membrane}
-Simple binding models have been used previously to interpret the same experimental data \cite{altenbach84,macdonald87} as employed in this work to validate the simulation models (Fig.~\ref{fig:delta_ordPar_CaCl}).
-In particular, NMR data concerning the PC headgroup order parameters response and atomic absorption spectra were explained best using a ternary complex binding model with a binding stoichiometry of one \ce{Ca^{2+}} per two POPC lipids~\cite{altenbach84}.
-Nevertheless, a Langmuir adsorption model assuming a \ce{Ca^{2+}}:POPC stoichiometry of 1:1 also provided a good fit to the experimental data when considering \ce{CaCl2} at low concentrations only~\cite{macdonald87}.
-
-In this work, we reproduce the same experimental data used to infer binding stoichiometries employing our ECC-POPC model.
-Thanks to our simulations, we have a direct access to atomistic details of the binding stoichiometry without a need for any binding model as employed for interpreting in experiments~\cite{altenbach84, macdonald87}.
-To evaluate the relative propensities for each of the stoichiometric complexes (i.e.,~1~Ca$^{2+}$:~n~POPC),
-we calculated for each bound Ca$^{2+}$ the number of POPC molecules having oxygen atoms within a distance of 0.3~nm.
-Results from the POPC bilayer simulation with a 285~mM bulk concentration of CaCl$_2$ are shown in Fig.~\ref{fig:cacl_complexes}.
-We found the largest propensity for the 1:2 complex (41\%), with probabilities of complexes with the stoichiometries of 1:1~(25\%)~and 1:3~(34\%)~being only slightly lower.
-This suggests a more complex binding model than considered in a simple 1:2 ternary complex model previously.
-Nevertheless, with a broad brushstroke, the simulation data can be viewed such that one calcium binds to two lipids on average, because the probabilities of the complexes with 1 or 3 lipids are almost equal to each other  (and complexes with more than three lipids per one calcium ion were not observed).
-This probably explains why the simple the ternary complex model fits adequately the experimental data, as well as the ECC-POPC simulation results (see Fig.~S3 in SI).
-
-\begin{figure}[tb!]
-  \centering
-  \includegraphics[width=8.0cm]{../Fig/ipython_nb/stoichiometry_NaCl-CaCl2_comparison_Ecc-lipids.pdf} \\
-  \caption{\label{fig:cacl_complexes}
-      Relative probabilities of existence of \ce{Na^{+}} or \ce{Ca^{2+}} complexes
-      with a certain number of POPC lipids. 
-      \ce{Na^{+}} complexes were evaluated from the simulation with 1~M concentration;
-      and \ce{Ca^{2+}} complexes were evaluated from the simulation with 287~mM concentration.
-  }
-\end{figure}
-
-The probabilities of different complexes formed by \ce{Na+} ions and POPC
-analyzed from the simulation with the ECC-POPC model at $1\,$M concentration of NaCl are also 
-shown in Fig.~\ref{fig:cacl_complexes}. In contrast to calcium, the
-probability is largest (67\%) for 1:1 complex, significantly smaller (29\%)
-for 1:2 complexes and very small (4\%) for 1:3 of \ce{Na+}:POPC complexes.
-
-\subsection{Residence times of \ce{Na+} and \ce{Ca^{2+}} cations in the POPC membrane}
-
-Equilibration of \ce{Ca^{2+}} ions at a POPC bilayer in MD simulations is a microsecond time scale process with current force fields, such  as CHARMM36 and Slipids force fields~\cite{javanainen17}.
-This suggests that at least several microseconds are required to reach the ion binding/unbinding equilibrium.
-To quantify the exchange of ions between the membrane and aqueous solution in simulations, we evaluated residence times of ions bound to the membrane.
-Within our analysis, an ion is considered to be bound when it is within 0.3~nm from any oxygen atom belonging to a POPC molecule.
-
-The histograms of residence times of \ce{Ca^{2+}} in a POPC bilayer ($C_{ion}'$ = 450~mM) from simulations with 
-ECC-POPC and CHARMM36 (simulation from Refs.~\citenum{javanainen17,zenodo.259376}) are shown in Fig.~S4 in SI.
-In the CHARMM36 simulation, a significant number of the calcium ions is bound to the membrane for the whole length of the trajectory (800~ns).
-In contrast, at least an order of magnitude faster bound/unbound calcium exchange is observed within the ECC-POPC model,
-where 90\% of the \ce{Ca^{2+}} residence times to a POPC membrane are shorter than $60\,\mathrm{ns}$. The longest observed
-residence time is around 150~ns, which is below the total length of the simulation used for analysis, i.e., 200~ns.
-Note that these results are in line with the experimental estimate that the residence time of \ce{Ca^{2+}} at each PC
-headgroup is of the order of $10\,\mu\mathrm{s}$~\cite{altenbach84}. Note that the exchange of \ce{Na+} ions at the POPC membrane
-is yet another order of magnitude faster, with 90\% of the residence times smaller than~1~ns and the longest residence time being~6~ns.
-
-In summary, the results from the ECC-POPC model suggest that the exchange of calcium between the POPC bilayer and the solvent occurs within the $\sim$100~ns timeframe, which is significantly faster than observed in simulations emloying most of the presently available lipid models~\cite{javanainen17}.
-Sodium cations exhibit an even more rapid exchange between the membrane and the aqueous solution.
-Our results suggest that simulations with a length of several hundreds of nanoseconds are sufficient to simulate alkali and alkali earth ion binding to phospholipid bilayers in equilibrium when realistic force fields are used.
-This has not been the case with previous lipid force fields, which overestimate the binding strength of the sodium and, in particular,  calcium cations \cite{javanainen17, catte16}.
+Show the successes and possible \emph{limits} of ECC applied on top of various models \cite{melcr18, martinek17, Pluharova2014, kohagen14, kohagen16}.
 
 \section{Conclusions}
-
-In this study, we employed the electrometer concept to demonstrate that the binding of \ce{Na^+} and \ce{Ca^2+} ions to a POPC lipid bilayer can be accurately described within a force field MD simulation, provided that electronic polarization is implicitly included via the electronic continuum correction~\cite{leontyev11}.
-This ECC-POPC model is built upon the Lipid14 POPC model~\cite{dickson14} by scaling the partial charges by a factor of 0.8 and reducing the Lennard-Jones radii by a factor of 0.89 for the headgroup, glycerol backbone, and carbonyl atoms.
-While the structural details of a POPC lipid bilayer in pure water simulated with the newly developed ECC-POPC model agree with experiments with an accuracy comparable to the other state of the art lipid models, the new model in addition reproduces the experimental lipid head group order parameter responses to a cationic surfactant, NaCl, and CaCl$_2$ at varying concentrations.
-It thus represents a significant improvement over currently available lipid models, which tend to overestimate cation binding affinities~\cite{catte16}.
-
-The good agreement with experiments enables us to interpret NMR experiments with atomistic details using MD simulations with the ECC-POPC model.
-In line with previous interpretations of the experimental data \cite{hauser76,hauser78,herbette84,binder02}, Ca$^{2+}$ ions interact mainly with phosphate oxygens, with interactions to the carbonyls being of a secondary importance.
-However, the stoichiometry of calcium binding is significantly more complicated than the simple ternary complex model, used to interpret NMR data previously, within which one calcium binds to two POPC molecules~\cite{altenbach84}.
-While complexes with one calcium ion bound to two lipids are the most probable also in the ECC-POPC model, complexes of one or three lipids per one calcium were observed to be relatively abundant and also almost equally likely to occur.
-While the success of the simple ternary complex model in fitting NMR data is understandable based on the simulation results, it cannot capture the detailed nature of calcium binding to phospholipid bilayers observed in the present simulation.
-
-Accurate description of cation binding to POPC bilayer paves the way for simulations of complex biochemical systems at cellular membranes with realistically described electrostatic interactions, including a mean-field account for electronic polarization effects.
-To this end, the compatibility of the ECC-POPC model with existing models for proteins and other biological molecules needs to be verified and, if necessary, further adjustments following the ECC concept to the force fields of the biomolecules interacting with the lipids should be made.
-
-This work can be reached as a repository containing all data and scripts for analysis at Ref. \citenum{github}.
 
@@
 and from the Academy of Finland via the FiDiPro award. 
-J.K. acknowledges support from the Czech Science Foundation, Project No. 15-12386S.
 Computational resources were supplied by the Ministry of Education, Youth and Sports 
@@
 
-The following files are available free of charge.
-\begin{itemize}
-  \item SI.pdf: Simulation details and further analysis.
+%The following files are available free of charge. 
+%\begin{itemize} 
 %  \item Filename: brief description 
-\end{itemize}
+%\end{itemize}
 

### 2018-05-14 Manuscript/manuscript.tex
Written a paragraph on population analysis using number of contacts, updated figures, now all simulations with Ca2+ are at least 500ns, added TODO latex package.

@@
 \usepackage{chemfig} 
+\usepackage[obeyFinal]{easy-todo}
 
@@
 
-Where do the cations bind? Preferentially to POPS. Discuss further \dots
-
-After validation above, it can be an option to show results also for systems, for which we have no experiments.
+\textbf{Where do the cations bind?} 
+The affinity towards PC and PS lipids and some of their groups was evaluated 
+by counting contacts between the cation and the oxygen atoms of the lipids
+similarly as was done in \cite{melcr18}. 
+The threshold for counting a contact was set to be $0.3 \mathrm{nm}$, 
+which encompasses the first peak of radial distribution function between the cations and the oxygen atoms of the lipids. 
+In Table~\ref{tab:binding}, there are percentages of the membrane-bound calcium population for various membrane moieties. 
+Although the lipid ratio in the membrane is 5~PC:1~PS,
+half of the total population of calcium cations is bound to PS lipids
+with 7\% bound only to them. 
+This corroborates the intrinsincally higher affinity of PS lipids to calcium cations compared to PC lipids. 
+The population analysis also suggests 
+that the calcium cations prefer to reside in the phosphate region of the membrane. 
+Hence, the carboxylate group in PS is coordinated to the cations bound in that region. 
+The increased response of head group order parameters $\alpha$ and $\beta$ of PC 
+in mixed 5~PC:1~PS bilayer compared to pure PC
+is due to higer affinity of the membrane mediated by even a relatively small fraction (1/6) of PS lipids. 
+This is in line with the steep onset of the response of the PS head group order parameters at lower concentrations.
+The complex response of the head group order parameters of PS lipids 
+is due to the confomational change of the carboxylate group that is attracted more towards the phosphate region. 
+In order to examine whether the carboxylate and phosphate groups in PS are bound 
+to the same calcium cation or rather different cations, 
+more detailed analysis is required. 
+\todo{Detailed analysis of the interactions of cations and PC/PS head groups using MSM.}
 
 \begin{table}[tb!] 
-  \caption{Bulk concentrations of Ca$^{2+}$ (C$_b$), relative surface excess of calcium with respect to water ($\Gamma_{Ca}^{\rm water}$), 
-    and the percetages of Ca$^{2+}$ bound to phosphate or carbonyl oxyges ($r^\mathrm{Ca^{2+}} _\mathrm{PO_4} $ and $r^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}}$) 
-    in different mixed POPC:POPS (5:1) bilayer models. 
+  \caption{Bulk concentrations of Ca$^{2+}$ (C$_b$), 
+           and the percentages of Ca$^{2+}$ bound to 
+           either to PC or PS, 
+           to PC or PS without phosphate group,
+           to PC or PS without carboxylate group,
+           to PC and PS at the same time,
+           only to PC, 
+           and only to PS 
+           in a mixed POPC:POPS (5:1) bilayer. 
           \label{tab:binding}} 
-  \begin{tabular}{l|c c | c | c c} 
-    model                  & $C_{ion}'$ & $C_{ion}\,/\,\mathrm{mM}$ & $\Gamma_{Ca}^{\rm water} \,/\, \mathrm{nm}^{-2}$  & $r^\mathrm{Ca^{2+}} _\mathrm{PO_4} $ & $r^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}} $ \\
+  \begin{tabular}{l| c | c c c c c c } 
+    model                  & $C_{ion}\,/\,\mathrm{mM}$ & PC+PS  &  PC+PS no PO$_4$ & PC+PS no COO$^-$ & PC-$Ca^{2+}$-PS  & PC & PS  \\ 
     \hline 
-    ECC-POPS               &  tba  &  $tba\pm 10 $  &  $0.?? \pm 0.01 $   &  ??\%  &    ??\%    \\ 
-    Lipid14/Dang           &  tba  &  $tba\pm 10 $  &  $0.?? \pm 0.03 $   &  ??\%  &    ??\%    \\ 
+    ECC-lipids             &  $240\pm 10 $             & 100    &   98-99          & 100              & 43               & 50 &  7  \\
+    \hline 
+    Lipid14/17             &  $280\pm 10 $             & 100    &   ---            & ---              & ---              & ---& --- \\
+    \hline 
   \end{tabular} 
@@
 Show the successes and possible \emph{limits} of ECC applied on top of various models \cite{melcr18, martinek17, Pluharova2014, kohagen14, kohagen16}.
+
+After validation above, it can be an option to show results also for systems, for which we have no experiments.
 

### 2018-08-06 Manuscript/manuscript.tex
Small update, comments.
Major update to the manuscript:   copied-in contents of my thesis for further editing.   Updated packages in latex template   updated biliography   plots are updated (my doctoral thesis submittion state)

--- 2018-04-19 Manuscript/manuscript.tex
@@
 
+%% Workaround for malfunctioning textsuperscript with pdfx and T1 encoding
+\let\tmpa\textsuperscript
+\DeclareTextCommandDefault{\textsuperscript}{\tmpa}
+
+\usepackage[x11names]{xcolor}  % typesetting in color
+%% Generate PDF/A-2u
+\usepackage[a-2u]{pdfx}
+
+\usepackage[super]{natbib}         % citation style AUTHOR (YEAR), or AUTHOR [NUMBER]
+
+\usepackage{achemso}  
 \usepackage{rotating}  
-\usepackage{times} 
+%\usepackage{times} 
+%\usepackage{lmodern}    %% Prefer Latin Modern fonts
 \usepackage{graphicx} 
@@
 \usepackage{chemfig} 
+\usepackage[obeyFinal]{easy-todo}
+\usepackage[english]{babel}
+
+\usepackage{subfig}
+\usepackage{dcolumn}        % improved alignment of table columns
+\usepackage{booktabs}       % improved horizontal lines in tables
+\usepackage{paralist}       % improved enumerate and itemize
+
+%% Character encoding: usually latin2, cp1250 or utf8:
+\usepackage[utf8]{inputenc}
+\usepackage[T1]{fontenc}
 
 %\usepackage{markdown}
+
+%% The hyperref package for clickable links in PDF and also for storing
+%% metadata to PDF (including the table of contents).
+%% Most settings are pre-set by the pdfx package.
+\hypersetup{unicode}
+\hypersetup{breaklinks=true}
+\hypersetup{colorlinks,allcolors=DodgerBlue4}
 
@@
 \newcommand*\mycommand[1]{\texttt{\emph{#1}}} 
+%%% Custom variables
+% width suitable for fitting into a column in 1-column page layout
 \newlength{\figwidth}
-\setlength{\figwidth}{8 cm}
+\setlength{\figwidth}{9 cm} 
+\newlength{\figwidthsmall}
+\setlength{\figwidthsmall}{6 cm} 
+\newlength{\figwidthfull}
+\setlength{\figwidthfull}{14 cm}
 
@@
 \begin{abstract} 
+	\emph{Adapt abstract to the final message: properties of POPS-containing bilayers -or- ECC-concept applicability?}
+
 Binding affinities and stoichiometries of sodium and calcium ions to phospholipid bilayers are of paramount significance in the properties and functionality of cellular membranes.
@@
 However, the interpretation of the experiments relies on theoretical models, as direct molecular information is not available.
+
 Classical molecular dynamics (MD) simulations provide details of the ion binding process with atomistic resolution, therefore offering all the necessary information to interpret experimental data without the need to resort to over-simplified models.
@@
 \cite{melcr18}
-We are now extending our work to other lipid membranes composed of larger varieties of different biologically relevant lipids.
-In particular, we focus our attention on membranes with phosphatidylethanolamine (PE) and negatively charged phosphatidylserine (PS).
+
+We are now extending our work to other lipid membranes composed of larger varieties of different biologically relevant lipids. In particular, we focus our attention on membranes with 
+	%phosphatidylethanolamine (PE) and 
+	negatively charged phosphatidylserine (PS) lipids. 
 	We compare available NMR and x-ray experiments with MD simulations to obtain atomistic-resolution picture of the interactions between cations and phospholipid membranes.
-Our preliminary results indicate that electronic polarization is a non-negligable interaction, which dramatically improves the description of cation binding to the membranes.
+Our results indicate that electronic polarization is a non-negligable interaction, which dramatically improves the description of cation binding to phospholipid membranes.
 \end{abstract}
@@
 \subsection{Electronic continuum correction for lipid bilayers}\label{section:ecc}
-This section may not be required, we may just cite ECC-POPS 
-and put a short paragraph describing what we did on POPS.
 
-Compared to ECC-POPC model \cite{melcr18}, 
-which used used a scaling factor $f_q = 0.8$ for all charges but lipid tails,
-this model, ECC-POPS, has to adopt a scaling factor $f_q = 0.75$ 
-as it has non-zero total charge in line with ECC theory. \cite{leontyev09}
-Hence, electronic continuum correction of POPS 
-has the same characteristics as for cations and anions, 
-denoted as ECC-ions \cite{martinek17, Pluharova2014, kohagen14, kohagen16}.
+\emph{ Directly taken from my thesis.}
+
+The lack of electronic polarizability in classical MD simulations
+has been considered a serious issue since the early days of the modeling of lipid bilayers. 
+% the commented text on polarizability in this section comes from \citep{huang2017mapping}
+There exist two main approaches for the explicit representation of electronic polarizability -- atomic induced dipoles and Drude oscillators.
+%The AMOEBA FF adopts the point induced dipole model in which each atom carries an induced dipole moment determined by the electric field and its atomic polarizability. 
+%Permanent dipole and quadrupoles are also assigned to each atomic site for a more accurate description of electrostatic interactions. 
+The induced dipole approach models the polarizability through a point dipole moment on each atom.
+Such an electric dipole is induced by the acting electric field proportionally to its polarizability. 
+This model is implemented in the widely adopted force field AMOEBA \citep{amoeba06, amoeba10, amoeba13}.
+
+%The Drude FF adopts the classical Drude oscillator model where auxiliary charges (Drude oscillators or particles) are attached to non-hydrogen atoms with harmonic springs, and the displacement of the Drude particle relative to its parent atom represents the response to the external electric field.
+%Virtual particles representative of lone pairs are included in the Drude FF to improve the treatment of electrostatic interactions typically involving hydrogen bond acceptors.
+%The charge distribution between a Drude particle and its parent atom is a finite difference representation of a dipole moment. 
+%Similarly, lone pairs can be considered as real space representations of permanent multipoles. 
+%Such analogy between the classical Drude oscillator model and the multipole and induced dipole (MPID) model suggests that in theory one can interconvert polarizable FFs between the two models. 
+The classical Drude oscillator model is inspired by a classical representation of a lone pair of electrons. 
+It uses an auxiliary charge,
+also termed Drude oscillator or Drude particle, 
+which is attached to each atom through a harmonic spring.
+The displacement of the Drude particle represents the induced polarization proportionally to the acting electric field,
+as implemented for example in the Drude force field \citep{lemkul2016empirical}.
+
+%In this work, we convert the Drude polarizable force field from its original classical Drude oscillator formalism onto a permanent multipole and induced dipole formalism, which we refer to as MPID. 
+%We show that a direct mapping without any reparametrization can transfer the Drude FF into the MPID FF, yielding a similar level of agreement as the Drude FF in reproducing condensed phase properties of organic liquids, crystals, and interfaces.
+%Our results thus also serve as an additional validation of these novel simulation algorithms.
+Although the two approaches, 
+the induced dipole model and the classical Drude oscillator model, 
+are very different in how they represent polarizability in simulations,
+it was shown that the classical Drude oscillator formalism can be converted to the induced dipole formalism 
+proving the two methods to be equivalent. \citep{huang2017mapping}
+From the point of view of lipids, however, the classical models with fixed charges and no explicit polarizability that are currently available 
+provide comparable accuracy at a fraction of the computational cost of the models with explicit polarization. \citep{lucas12,chowdhary13}
+
+%The effects of polarization from rather computationally demanding simulations with explicit models 
+%can be implemented using an implicit model of polarization. 
+Instead of the computationally demanding explicit polarization approach, 
+electronic polarization can also be accounted for implicitly.
+Such an implicit mean field model of electronic polarizability 
+%was developed by \citet{leontyev09} under the name Molecular Dynamics in Electronic Continuum, MDEC. 
+termed Electronic continuum correction (ECC) 
+was first suggested by \citet{leontyev09}, 
+and was then systematically applied to solutions of ions by \citet{Pluharova2014, kohagen14, kohagen16, martinek17}. 
+Through a comparison of neutron scattering and/or \emph{ab-initio} molecular dynamics data with classical MD simulations, 
+the authors developed an array of models for cations and anions, termed here as ECC-ions,
+which provide realistic structures even for concentrated solutions of both monovalent and divalent cations. 
+%The term Electronic continuum correction, ECC, originates in the works by \citet{Pluharova2014, kohagen14, kohagen16, martinek17},
+%where the rigorously developed theory by \citet{leontyev14} was systematicaly applied to solutions of ions.
+%that use ECC to represent electronic polarizability.
+
+%Electronic continuum correction (ECC) 
+In ECC, a system of polarizable particles is represented 
+with an equivalent system of particles with fixed charges. %\citep{leontyev09, leontyev10, leontyev11, leontyev14}
+%including the effects of electronic polarization \emph{implicitly} in a mean-field way. 
+This can be done by a simple transform of the partial charges,
+which does not change the form of the empirical force field in Equation~\ref{eq:amber}, 
+providing the effects of the electronic polarization in a mean field way at no additional computational cost. 
+Although this is technically similar to the phenomenological scaling of partial charges 
+applied in earlier studies of aqueous solutions or ionic liquids \citep{jonsson86,egberts94,beichel14},
+%where also other possible effects (e.g. charge transfer) were considered,
+ECC is a physically well justified and rigorously derived model \citep{leontyev09, leontyev10, leontyev11, leontyev14}.
+Assuming that the electronic polarizability of electrons is approximately constant in biological solutions, 
+it was shown by \citet{leontyev11} that the electrostatic interactions between atoms 
+are screened like in a dielectric continuum 
+with an electronic dielectric constant, $\epsilon _{el}$, 
+also known as the high frequency dielectric constant. 
+This can be formally written as
+\begin{equation}  \label{eq:coulomb}
+   %U = \frac{1}{2} \displaystyle \sum ^N _{j\neq i} \frac{q_i q_j}{r_{ij}} + \frac{1}{2} \displaystyle \sum ^N _{i,j=1} U ^{dip} _{ij} - \displaystyle \sum ^N _{i=1} E^q(r_i) d_i  \propto  \frac{1}{2} \displaystyle \sum ^N _{j\neq i} \frac{q_i q_j}{\epsilon_{el} \, r_{ij}}  =  \frac{1}{2} \displaystyle \sum ^N _{j\neq i} \frac{q^{ECC}_i q^{ECC}_j}{r_{ij}}
+   %U = \frac{1}{2} \displaystyle \sum ^N _{j\neq i} \frac{q_i q_j}{r_{ij}} + \frac{1}{2} \displaystyle \sum ^N _{i,j=1} U ^{dip-dip} _{ij} - \displaystyle \sum ^N _{i=1} U^{q-dip}_i  \propto  \frac{1}{2} \displaystyle \sum ^N _{j\neq i} \frac{q_i q_j}{\epsilon_{el} \, r_{ij}}  =  \frac{1}{2} \displaystyle \sum ^N _{j\neq i} \frac{\frac{q_i}{\sqrt{\epsilon _{el}}} \frac{q_j}{\sqrt{\epsilon _{el}}}}{r_{ij}}  ,
+   U = \frac{1}{2} \displaystyle \sum ^N _{j\neq i} \frac{q_i q_j}{\epsilon_{el} \, r_{ij}}  =  \frac{1}{2} \displaystyle \sum ^N _{j\neq i} \frac{\frac{q_i}{\sqrt{\epsilon _{el}}} \frac{q_j}{\sqrt{\epsilon _{el}}}}{r_{ij}}  ,
+\end{equation} 
+where $q_i$ is the point charge of the particle $i$,
+and $r_{ij}$ are the distances between particles $i$ and $j$.
+%$U^{dip-dip} _{ij}$ represent the interactions between induced dipoles and their polarization energies,
+%and the terms $U^{q-dip} _i$ are the interactions between point charges and induced dipoles. 
+It naturally follows from the above equation that 
+embedding all atoms into a dielectric continuum 
+is equivalent to a formal scaling of the atomic point charges 
+\begin{equation}  \label{eq:scaling}
+ q_i ^{ECC} = \frac{q_i}{\sqrt{\epsilon _{el}}} ,
+\end{equation} 
+where $q_i$ and $q_i^{ECC}$ are respectively the original partial charges and the scaled charges, 
+which represent the effects of the electronic polarization. 
+Given that the  high frequency dielectric constant of water is 1.78 (i.e., the square of the refraction index of 1.33), 
+the scaling factor for ions in water is $\approx 0.75$.
+
+It is important to note that the value of the high frequency dielectric constant  
+is around 2 for almost any biologically relevant environment \citep{leontyev11}. 
+This means that even interfaces like biological membranes do not contain discontinuities of the electronic continuum. 
+The dielectric discontinuity in a lipid bilayer thus arises only 
+from the orientational polarization of the molecules, which is accounted for explicitly in standard MD simulations.  
+Therefore, the same correction for the electronic polarizability can be  
+applied throughout the lipid bilayer/aqueous solution interface.
 
@@
 
-\subsubsection{Simulations of POPS:POPS bilayers with aqueous ions}
-
-Any specifics for simulations with cations.
-
-\subsubsection{Simulations of POPS:POPS bilayers with anionic/cationic surfactants/peptides}
+Any specifics for simulations with aqueous cations.
 
@@
   \centering 
-  \includegraphics[width=\figwidth]{../Fig/order_parameters_actual_pure-POPS.pdf} \\
-  \includegraphics[width=0.7\figwidth]{../Fig/form-f_l17-ecc-pops-exp_compar.pdf} 
-  \includegraphics[width=0.55 \figwidth]{../Fig/lipids_chemfig.pdf} 
-  \caption{\label{simVSexpNOions} 
-    X-ray scattering form factors from simulations with the Lipid17 \cite{lipid17-future} and 
-    the ECC-POPS models compared with experiments~\cite{SDP-CHARMM36_comparison_paper_Samuli-knows} at 298~K. 
-    The disagreements between ECC-POPS model and experiment 
-    may come from an intrinsincally incorrect structure of the head group (and, hence, electron density in the region), 
-    however, lipid packing and overall membrane properties 
-    may be still caputured well enough -- after considering areas per lipid reported in Table~\ref{tab:apls}.  \\
-    Order parameters of POPS head group, glycerol backbone and acyl chains  
-    from simulations with the Lipid17 \cite{lipid17-future} and the ECC-POPS models 
-    compared with \emph{experiments at 300~K (check and change)}. 
-    Open/closed symbols are used for palmitoyl/oleoyl chains of POPS. \\
-    The chemical structure of POPS and the labeling of the carbon segments. 
+  \includegraphics[width=\figwidth]{../img/ecc_pops/Order-parameters_form-factors_exp-L17-ECC-lipids.pdf}
+  \includegraphics[width=\figwidth]{../img/ecc_pops/pops_chemfig.pdf} 
+\hfill
+  \caption{\label{simVSexpNOions_POPS} 
+    Top: X-ray scattering form factors from simulations with the Lipid17 \citep{lipid17-future} and 
+    the ECC-POPS models compared with experiments~\citep{kucerka14} at 298~K. 
+    Middle: Order parameters of POPS head group, glycerol backbone and acyl chains  
+    from simulations with the Lipid17 \citep{lipid17-future} and the ECC-POPS models 
+    compared with experiments at 298~K. \citep{nmrlipids_proj4}
+    Open/closed symbols are used for palmitoyl/oleoyl chains of POPS. 
+    Bottom: The chemical structure of POPS and the labeling of the carbon segments. 
   }  
@@
 \begin{table}[tb!] 
-  \caption{Values of the area per lipid (APL) of POPS bilayers with only \ce{Na^+} counterions. \label{tab:apls} 
-  } 
+\centering
+  \caption{Values of the area per lipid (APL) of POPS bilayers 
+	   at 298~K with only \ce{Na^+} counterions and no additional ions. \label{tab:apls} } 
   \begin{tabular}{l|c c} 
-    model          & APL (\AA$^2$)   & Temperature [K] \\ 
+%    \multicolumn{3}{c}{POPC} \\
+%    model          & APL / Å$^2$   & $T$ / K  \\ 
+%    \hline 
+%    Lipid14 POPC \citep{melcr18}    & 65.1$\pm$ 0.6  &  300 \\ 
+%    Lipid14 POPC \citep{dickson14}  & 65.6$\pm$ 0.5  &  303 \\ 
+%    \hline 
+%    ECC-POPC     \citep{melcr18}    & 63.2$\pm$ 0.6  &  300       \\ 
+%    \hline 
+%    experiment (SDP model) \citep{kucerka11} & 64.3  &  303    \\ 
+%    \hline 
+%    \multicolumn{3}{c}{} \\
+%    \multicolumn{3}{c}{POPE} \\
+%    model          & APL / Å$^2$   & $T$ / K  \\ 
+%    \hline 
+%    ECC-POPE                 & 56.7$\pm$ 0.8  &  298 \\ 
+%    \hline 
+%    experiment   \citep{parsegian89} & 56.6  &  310    \\ 
+%    experiment   \citep{rappolt03}   & 59--61 &  303--313  \\ 
+%    \hline 
+%    \multicolumn{3}{c}{} \\
+    \multicolumn{3}{c}{POPS} \\
+    model          & APL / Å$^2$   & $T$ / K  \\ 
     \hline 
-    Lipid14                   & 53.5$\pm$ 0.8  &  298 \\ 
+    Lipid17 POPS              & 53.5$\pm$ 0.8  &  298 \\ 
     \hline 
@@
     \hline 
-    experiment \cite{SDP-CHARMM36_comparison_paper_Samuli-knows} & 62.3  &  298    \\ 
+    experiment (SDP model) \citep{kucerka14} & 62.3  &  298    \\ 
     \hline 
@@
 
-\subsection{Electrometer concept in negatively charged membranes with various cations}
+We compared X-ray scattering form factors and NMR order parameters of bilayers
+in pure water without any ions (or only counter ions)
+from simulations and experiments
+as the first step in the assessment of the quality of the model. 
+The experimental X-ray scattering form factors 
+of a bilayer are well reproduced for all lipids employing the presented ECC-lipids model 
+(see Figs.~\ref{fig:simVSexpNOions}, \ref{simVSexpNOions_POPE} and \ref{simVSexpNOions_POPS}).
 
-Preliminary results are plotted in the figures in this section. 
-Beware! -- the y-axis scales are variable!
-Colours:
-\begin{itemize}
-  \item Red -- ECC-lipids, ECC-ions
-  \item Blue -- Lipid14/17, Dang ions
-\end{itemize}
+The area per lipid is often used as a relatively simple structural parameter reporting on the bilayer properties and the packing of lipids. 
+In experiments, modeling is used on top of the scattering factors to obtain it \citep{kucerka14}. 
+From simulations, this property is easily extracted. 
+We compare the values from experiments and simulations in Table~\ref{tab:apls}.
+
+While the agreement between the scattering form factors 
+from the simulation of a pure POPS bilayer and experiment 
+are excellent (Fig.~\ref{simVSexpNOions_POPS}),
+there is a non-negligible difference between the values of the area per lipid in Table~\ref{tab:apls}. 
+Since both values are derived from the scattering form factors through modeling of the electron density of the bilayer,
+we cannot decide, which of the values is more reliable. 
+In general, we can conclude that ECC-lipids
+%the presented lipid models with ECC-lipids 
+reproduce the experimental structural parameters of the lipid bilayers 
+with a comparable accuracy to existing state-of-the-art lipid models~\citep{botan15, ollila16, Pluhackova2016}.
+
+The head group and acyl chain order parameters within ECC-lipids
+are in general in a good agreement with the experimental values 
+as shown in Fig. \ref{simVSexpNOions_POPS}. 
+The acyl chain order parameters in particular are almost all within the experimental error bars.
+The order parameters of the head groups are at an accuracy comparable to 
+other currently available classical models of lipids \citep{botan15, catte16, Pluhackova2016}.
+
+The head group order parameters $\alpha$ and $\beta$ are highly relevant for this work,
+as they are being used in the electrometer concept \cite{altenbach84, catte16, melcr18}.
+For POPC in pure water, the order parameter $\beta$ agrees well with the experiment, 
+while the order parameter $\alpha$ is somewhat lower. 
+In the case of POPS, the situation is a bit more complicated
+compared to POPC as the order parameter $\alpha$ exhibits a notable forking (see Fig.~\ref{simVSexpNOions_POPS}).
+One of the order parameters of ECC-POPS, $\alpha_1$, agrees well with the experiment, 
+while the other, $\alpha_2$, adopts a higher value underestimating the experimentally reported forking. 
+There is only one order parameter $\beta$ in POPS, 
+which has a higher value closer to zero in the ECC-lipids model than in experiment. 
+Such a feature suggests that the model overestimates the orientational freedom of its head group.
+
+\subsection{Electrometer concept: cationic and anionic amphiphilic molecules}
 
 Changes of the lipid bilayer head group order parameters extracted from simulations and 
-experiments \cite{roux90} are shown in Figs.~\ref{fig:delta_ordPar_KCl}, \ref{fig:delta_ordPar_NaCl} 
-and~\ref{fig:delta_ordPar_CaCl} as functions of KCl, NaCl or CaCl$_2$ concentrations,
+experiments \cite{roux90} are shown in Figs.~\ref{fig:delta_ordPar_KCl_PCPS}, \ref{fig:delta_ordPar_NaCl_PCPS} 
+and~\ref{fig:delta_ordPar_CaC_PCPS} as functions of KCl, NaCl or CaCl$_2$ concentrations,
 and in Figs.~\ref{fig:delta_ordPar_NaCl_PC-PS_mix} as a function of PC and PS content. 
+and in Figs.~\ref{OrderParameterCHANGESsurf} as a function of the content of dihexadecylammonium bromide, a cationic amphiphile.
 
-% ============================== 
-% ==  ECC-lipids plots ========= 
-% ==     and           ========= 
-% ==  Lipid14/17 plots ========= 
-% ============================== 
-\begin{figure}[htb!] 
+\begin{figure}[tb!] 
   \centering 
-  \includegraphics[width=\figwidth]{../Fig/l17/order_parameters_changes_A-B_POPC_kcl.pdf} 
-  \includegraphics[width=\figwidth]{../Fig/l17/order_parameters_changes_A-B_POPS_kcl.pdf} 
-  \caption{\label{fig:delta_ordPar_KCl_l17} 
-    Changes of the head group order parameters of a POPC:POPS 5:1 bilayer as a function of KCl concentration 
-    in bulk ($C_{ion}$) from simulations with different force fields at 298 K together with  
+  \includegraphics[width=\figwidth]{../img/ecc_popc/PN_angle_OrdPars-A-B_L14-ECCL17_q80_sig89_surf.pdf} 
+  \caption{\label{OrderParameterCHANGESsurf} 
+    Changes of the head group order parameters $\alpha$, $\beta$ and P-N vector orientation as a function of 
+    the molar fraction of a cationic surfactant dihexadecyldimethylammonium in a POPC bilayer 
+    from simulations and experiments \citep{scherer89} at 313 K.
+    Image reproduced from \cite{melcr18}.
   } 
+\todo{Combine with Fig.~\ref{fig:delta_ordPar_NaCl_PC-PS_mix} into a single plot.}
 \end{figure}
@@
   \centering 
-  \includegraphics[width=\figwidth]{../Fig/order_parameters_changes_A-B_POPC_kcl.pdf} 
-  \includegraphics[width=\figwidth]{../Fig/order_parameters_changes_A-B_POPS_kcl.pdf} 
-  \caption{\label{fig:delta_ordPar_KCl} 
-    Changes of the head group order parameters of a POPC:POPS 5:1 bilayer as a function of KCl concentration 
-    in bulk ($C_{ion}$) from simulations with different force fields at 298 K together with  
-  } 
-\end{figure}
-
-\begin{figure}[htb!] 
-  \centering 
-  \includegraphics[width=\figwidth]{../Fig/l17/order_parameters_changes_A-B_POPC_nacl.pdf} 
-  \includegraphics[width=\figwidth]{../Fig/l17/order_parameters_changes_A-B_POPS_nacl.pdf} 
-  \caption{\label{fig:delta_ordPar_NaCl_l17} 
-    Changes of the head group order parameters of a POPC:POPS 5:1 bilayer as a function of NaCl concentration 
-    in bulk ($C_{ion}$) from simulations with different force fields at 298 K together with  
-  } 
-\end{figure}
-
-\begin{figure}[htb!] 
-  \centering 
-  \includegraphics[width=\figwidth]{../Fig/order_parameters_changes_A-B_POPC_nacl.pdf} 
-  \includegraphics[width=\figwidth]{../Fig/order_parameters_changes_A-B_POPS_nacl.pdf} 
-  \caption{\label{fig:delta_ordPar_NaCl} 
-    Changes of the head group order parameters of a POPC:POPS 5:1 bilayer as a function of NaCl concentration 
-    in bulk ($C_{ion}$) from simulations with different force fields at 298 K together with  
-  } 
-\end{figure}
-
-\begin{figure}[htb!] 
-  \centering 
-  \includegraphics[width=\figwidth]{../Fig/l17/order_parameters_changes_A-B_POPC_cacl.pdf} 
-  \includegraphics[width=\figwidth]{../Fig/l17/order_parameters_changes_A-B_POPS_cacl.pdf} 
-  \caption{\label{fig:delta_ordPar_CaCl_l17} 
-    Changes of the head group order parameters of a POPC:POPS 5:1 bilayer as a function of CaCl concentration 
-    in bulk ($C_{ion}$) from simulations with different force fields at 298 K together with  
-  } 
-\end{figure}
-
-\begin{figure}[htb!] 
-  \centering 
-  \includegraphics[width=\figwidth]{../Fig/order_parameters_changes_A-B_POPC_cacl.pdf} 
-  \includegraphics[width=\figwidth]{../Fig/order_parameters_changes_A-B_POPS_cacl.pdf} 
-  \caption{\label{fig:delta_ordPar_CaCl} 
-    Changes of the head group order parameters of a POPC:POPS 5:1 bilayer as a function of CaCl concentration 
-    in bulk ($C_{ion}$) from simulations with different force fields at 298 K together with  
-  } 
-\end{figure}
-
-\begin{figure}[htb!] 
-  \centering 
-  \includegraphics[width=\figwidth]{../Fig/l17/order_parameters_changes_A-B_PC-PS_mix_POPC_nacl.pdf} 
-  \includegraphics[width=\figwidth]{../Fig/l17/order_parameters_changes_A-B_PC-PS_mix_POPS_nacl.pdf} 
-  \caption{\label{fig:delta_ordPar_NaCl_PC-PS_mix_l17} 
-    Changes of the head group order parameters of a POPC:POPS 5:1 bilayer as a function of PS content.
-    Experimental points have inherent systematic error, as they come from different experiments.
-    A complete series of points can be found (?) in \cite{roux90} -- which is to be added to the plot.
-  } 
-\end{figure}
-
-\begin{figure}[htb!] 
-  \centering 
-  \includegraphics[width=\figwidth]{../Fig/order_parameters_changes_A-B_PC-PS_mix_POPC_nacl.pdf} 
-  \includegraphics[width=\figwidth]{../Fig/order_parameters_changes_A-B_PC-PS_mix_POPS_nacl.pdf} 
+  \includegraphics[width=\figwidthsmall]{../Fig/l17/order_parameters_changes_A-B_PC-PS_mix_POPC_nacl.pdf} 
+  \includegraphics[width=\figwidthsmall]{../Fig/l17/order_parameters_changes_A-B_PC-PS_mix_POPS_nacl.pdf} 
+  \includegraphics[width=\figwidthsmall]{../Fig/order_parameters_changes_A-B_PC-PS_mix_POPC_nacl.pdf} 
+  \includegraphics[width=\figwidthsmall]{../Fig/order_parameters_changes_A-B_PC-PS_mix_POPS_nacl.pdf} 
   \caption{\label{fig:delta_ordPar_NaCl_PC-PS_mix} 
@@
 
-\subsection{Molecular interaction and binding affinities of K$^+$, Na$^+$ or Ca$^{2+}$  cations to the mixed POPC:POPS (5:1) membrane} 
+\subsection{Molecular interaction and binding affinities of K$^+$ and Na$^+$ cations to mixed POPC:POPS (5:1) membrane} 
 \label{sec:affinity}
+
+\begin{figure}[tbp!] 
+  \centering 
+  \includegraphics[width=\figwidthsmall]{../img/ecc_pops/order_parameters_changes_ecc-lip_L14_A-B-PN-COO_POPC_nacl.pdf} 
+  \includegraphics[width=\figwidthsmall]{../img/ecc_pops/order_parameters_changes_ecc-lip_L14_A-B-PN-COO_POPS_nacl.pdf} 
+  \caption{\label{fig:delta_ordPar_NaCl_PCPS} 
+    Changes of the head group order parameters $\alpha$, $\beta$ and the orientations of the carboxylate group and the P-N vector  
+    of POPC (left) and POPS (right) phospholipids in a POPC:POPS 5:1 bilayer as a function of \ce{NaCl} concentration 
+    in bulk ($C_{ion}$) from simulations with different force fields at 298 K.
+    Because data with \ce{NaCl} are not available for POPC, 
+    we show experimental data for \ce{LiCl} (dashed line, left) 
+    as an upper bound for the magnitude of the response to \ce{NaCl}, 
+    which has a lower affinity to phospholipid bilayers compared to \ce{LiCl} \citep{roux90}. 
+    The orientation of the \ce{COO^-} group is defined as 
+    the connector from the $\beta$ carbon to the carbon in \ce{COO^-} (stars, bottom right). 
+  } 
+\end{figure}
+
+\begin{figure}[tbp!] 
+  \centering 
+  \includegraphics[width=\figwidthsmall]{../img/ecc_pops/order_parameters_changes_ecc-lip_L14_A-B-PN-COO_POPC_kcl.pdf} 
+  \includegraphics[width=\figwidthsmall]{../img/ecc_pops/order_parameters_changes_ecc-lip_L14_A-B-PN-COO_POPS_kcl.pdf} 
+  \caption{\label{fig:delta_ordPar_KCl_PCPS} 
+    Changes of the head group order parameters $\alpha$, $\beta$ and the orientations of the carboxylate group and the P-N vector  
+    of POPC (left) and POPS (right) phospholipids in a POPC:POPS 5:1 bilayer as a function of \ce{KCl} concentration 
+    in bulk ($C_{ion}$) from simulations with different force fields and experiments at 298 K. \citep{roux90}
+    The orientation of the \ce{COO^-} group is defined as 
+    the connector from the $\beta$ carbon to the carbon in \ce{COO^-} (stars, bottom right). 
+  } 
+\end{figure}
+
+% Pavel suggested to remove this as it only recaps previous sections
+%It was presented in the previous section,
+%that accounting for electronic polarization,
+%a distinct feature of ECC-lipids compared to other lipid models, 
+%is crucial for an accurate description of the response of the POPC electrometer. 
+%In this section,
+%we employ ECC-lipids to simulate neutral and negatively charged bilayers 
+%in the presence of monovalent salts, namely \ce{KCl} and \ce{NaCl}.
+
+The binding of \ce{Na^+} cations to phospholipid bilayers 
+is not generally agreed on between simulations 
+\citep{bockmann03,sachs04,berkowitz06,cordomi09}
+and experiments 
+\citep{cevc90,tocanne90,hauser76,herbette84,uhrikova08}.
+We attempted to find a model, 
+which would successfully interpret the experiments on binding of \ce{Na^+}
+to neutral and negatively charged phospholipid bilayers \citep{akutsu81, roux90} 
+in our works \citep{catte16, nmrlipids_proj4}.
+After testing the currently available force fields for phospholipids,
+we concluded
+that the interactions are generally overestimated in magnitude in almost all models 
+but Lipid14 (PC) \citep{dickson14}, resp. Lipid17 (PS) \citep{lipid17-future}. 
+This model yields a semi-quantitative agreement with the experimentally measured small changes of the order parameters 
+when used with the model of ions by \citet{aqvist90} (Fig.~\ref{fig:catte16}). 
+However, when used with a more accurate model of ions by \citet{Pluharova2014, martinek17},
+the model overestimates the binding affinity of \ce{Na^+}
+measured with lipid electrometer concept. \citep{melcr18}
+In total, these results suggest that improvements 
+in the lipid parameters are required for more accurate interactions even with monovalent cations. 
+\citep{catte16, melcr18, nmrlipids_proj4}
+
+Generally improved behaviour 
+of the POPC and POPS head group order parameters 
+with \ce{NaCl} or \ce{KCl} concentrations 
+was achieved through the combination of models 
+ECC-lipids \citep{melcr18} and ECC-ions \citep{martinek17, kohagen16, Pluharova2014}. 
+Simulations with these models 
+reveal a good agreement with the NMR experiments 
+for both neutral and negatively charged membranes.
+The results are plotted for \ce{NaCl} in Fig. \ref{fig:delta_ordPar_NaCl_PCPS}, 
+and for \ce{KCl} in Fig. \ref{fig:delta_ordPar_KCl_PCPS}.
+
+The interaction with \ce{K^+}, which binds very weakly to both neutral and negatively charged membranes, 
+renders a qualitatively different response of the order parameter $S^\beta$ in POPS in the mixed negatively charged bilayers compared to the neutral bilayers. 
+While the order parameter $S^\beta$ increases for both \ce{Na^+} and \ce{Ca^{2+}},
+it decreases in the presence of \ce{K^+}.
+%This feature is, \emph{qualitatively} captured by a few models in \citep{nmrlipids_proj4},
+%neither of them is as close to a \emph{qunatitative} agreement with the experiments 
+%as the combination of ECC-lipids with ECC-ions. 
+However, no model studied in \citep{nmrlipids_proj4} describes the behaviour of a PS head group correctly enough to reveal this effect. 
+In contrast, ECC-lipids with ECC-ions capture the different response of the order parameters $S^{\beta}$, $S^{\alpha _1}$ and $S^{\alpha _2}$ in POPS to various salts accurately. 
+Such a detailed description of the changes of the structural parameters 
+demonstrates that including electronic polarization
+improves the description of the interactions even for very weakly binding cations like \ce{K^+}.
+
+\begin{figure}[tbp!] 
+  \centering 
+  \includegraphics[width=\figwidth]{../img/ecc_pops/density_profiles_na_k_cl_wat_phos_models-compar_4-6_NaCl-and-KCl-series.pdf}
+  \caption{\label{fig:nacl-dens_PCPS} 
+    Number density profiles of \ce{K^{+}}, \ce{Na^{+}} and \ce{Cl^-} along the membrane normal axis 
+    for the negatively charged membranes with the composition of 5\,PC:1\,PS using ECC-lipids and ECC-ions.  
+    The top profile shows the simulation without any additional salt concentration, i.e., only with \ce{Na^+} counterions. 
+    The middle profile shows the simulation with an additional \ce{KCl} concentration and \ce{Na^+} counterions. 
+    The bottom profile shows the simulation with an additional \ce{NaCl} concentration and \ce{Na^+} counterions, which are not distinguished from the added salt. 
+    In order to visualize the density profiles with a scale comparable to the profile of \ce{Ca^{2+}}, 
+    the density profiles of~\ce{Cl^-}, \ce{K^+} and \ce{Na^+} ions are divided by 2, and 
+    the density profiles of phosphate groups and water are divided by 5 and 200, respectively.  
+    } 
+\end{figure}
+
+The difference between the affinities of \ce{Na^+} and \ce{K^+} to neutral and negatively charged membranes
+can be described by their relative surface excesses with respect to water, $\Gamma ^{w} _{ion}$, 
+which are shown in the plots of the density profiles of the ions in Figs.~\ref{fig:nacl-dens} (neutral bilayer) and \ref{fig:nacl-dens_PCPS} (negatively charged bilayer). 
+Such a quantity compares the adsorption of ions to the adsorption of water molecules at an interface 
+without the necessity of defining a Gibbs dividing surface. \citep{melcr18, chattorajBOOK}
+While \ce{K^+} maintains negative values of $\Gamma^{w}_{K}$ even for the negatively charged bilayer,
+the value of $\Gamma^{w}_{Na}$ for \ce{Na^+} changes from negative to positive
+in a neutral POPC bilayer versus in a negatively charged bilayer with the composition 5\,PC:1\,PS.
+%This means that at the given concentration the bilayer interface has a small preference to \ce{Na^+} cations compared to water molecules. 
+Interestingly, this value is slightly decreased in the presence of an additional \ce{NaCl} concentration adding also \ce{Cl^-} anions, 
+which are not present in the system when only counterions are used
+(bottom resp. top plot in Fig.~\ref{fig:nacl-dens_PCPS}). 
+The interaction of a neutral POPC bilayer with \ce{NaCl} is discussed in a greater detail in our publication \citep{melcr18}.
+
+\subsection{Molecular interaction and binding affinities of \ce{Ca^{2+}} cations to mixed POPC:POPS (5:1) membrane} 
+\label{section:lip-ion_ca}
+
+\begin{figure}[tbp!] 
+  \centering 
+  \includegraphics[width=\figwidthsmall]{../img/ecc_pops/order_parameters_changes_ecc-lip_L14_A-B-PN-COO_POPC_cacl.pdf} 
+  \includegraphics[width=\figwidthsmall]{../img/ecc_pops/order_parameters_changes_ecc-lip_L14_A-B-PN-COO_POPS_cacl.pdf} 
+  \caption{\label{fig:delta_ordPar_CaCl_PCPS} 
+    Changes of the head group order parameters $\alpha$, $\beta$ and the orientations of the carboxylate group and the P-N vector  
+    of POPC (left) and POPS (right) phospholipids in a POPC:POPS 5:1 bilayer as a function of \ce{CaCl2} concentration 
+    in bulk ($C_{ion}$) from simulations with different force fields and experiments at 298~K. \citep{roux90}
+    The orientation of the \ce{COO^-} group is defined as 
+    the connector from the $\beta$ carbon to the carbon in \ce{COO^-} (stars, bottom right). 
+  } 
+\end{figure}
+
+%%% Sum up what is to be presented briefly
+%The importance of treating polarizability for interactions of phospholipids with monovalent ions was demonstrated in the previous section. 
+Electronic polarization is a non-negligible contribution to the interactions of calcium even in simple aqueous solutions of \ce{CaCl2} \citep{martinek17, kohagen16, Pluharova2014}. 
+In this section,
+we will show the results from simulations of neutral and negatively charged phospholipid bilayers at varying \ce{CaCl2} concentrations
+using the recently developed models ECC-lipids and ECC-ions \citep{melcr18, martinek17}, 
+which implicitly include the effects of electronic polarization through electronic continuum correction \citep{leontyev11}. 
+Validated with the concept of a lipid electrometer in section \ref{section:electrometer_exp_sim},
+these implicitly polarizable models yield accurate description of 
+the interaction of \ce{Ca^{2+}} with both neutral and negatively charged phospholipids.
+
+The changes of the head group order parameters $S^\alpha$ and $S^\beta$ from simulations and experiments 
+are shown in Fig.~\ref{fig:delta_ordPar_CaCl} for a neutral POPC bilayer, 
+and in Fig.~\ref{fig:delta_ordPar_CaCl_PCPS} for a negatively charged bilayer with the composition 5\,PC:1\,PS. 
+For a direct comparison and a connection to our works \citep{catte16, nmrlipids_proj4},
+we show simulation results from ECC-lipids and also from Lipid17 \citep{lipid17-future}. 
+This also highlights the improvements in ECC-lipids over Lipid17 arising from the electronic polarization. 
+%Although Lipid17 already belongs to the top-performing models in terms of the responses of the head group order parameters in those studies,  
+%including electronic polarizability to form ECC-lipids improves the results even further.
+The effect is probably the most striking for POPS, 
+for which also the structure of a pure POPS bilayer with only counterions is dramatically improved with the augmentation.
+
+%%% Discuss the changes of OPs and vectors from neutral and neg. membranes
+%While the changes induced by \ce{CaCl2} are \emph{qualitatively} correct for Lipid14/17, 
+Increasing concentrations of \ce{CaCl2} induce a systematic decrease of the order parameters $S^\alpha$ and $S^\beta$ in POPC. 
+Although the total magnitude of the response of the PC head group order parameters 
+is only slightly higher in the negatively charged bilayers than in the neutral bilayers, 
+the shape of the changes in the latter shows a steeper onset at low concentrations. 
+This is apparently due to the presence of POPS, 
+which has a higher affinity to \ce{Ca^{2+}} compared to POPC. 
+ECC-lipids is the first model,
+which achieves a \emph{quantitative} agreement with the changes induced by \ce{CaCl2} in experiments.
+
+\begin{figure}[htbp!] 
+  \centering 
+  \includegraphics[width=\figwidth]{../img/ecc_pops/density_profiles_ca_na_cl_wat_phos_models-compar_1-3_CaCl2-series.pdf}
+  \caption{\label{fig:cacl-dens_PCPS} 
+    Number density profiles of \ce{Ca^{2+}}, \ce{Na^{+}} and \ce{Cl^-} along the normal of the membrane starting at the centre
+    for the negatively charged membrane with the composition 5\,PC:1\,PS
+    at various bulk concentrations of \ce{CaCl2} from simulations with ECC-lipids. 
+    All profiles contain \ce{Na^+} counterions and an additional concentration of \ce{CaCl2}. 
+    In order to visualize the density profiles with a scale comparable to the profile of \ce{Ca^{2+}},  
+    the density profiles of~\ce{Cl^-} ions are divided by 2, and 
+    the density profiles of phosphate groups and water are divided by 5 and 200, respectively.  
+    } 
+\todo{JM: Compare with pure POPC bilayer?}
+\end{figure}
+
+%%% Present the density profiles (different concentrations for both POPC and mixed)
+The increase in the amount of bound calcium cations from pure POPC \cite{melcr18}
+to the mixed negatively charged bilayer containing POPS
+is well demonstrated using the relative surface excess, $\Gamma ^w _{Ca}$,
+summarized in Table~\ref{tab:binding}. 
+Distributions of \ce{Ca^{2+}}, \ce{Na^+} counterions and also \ce{Cl^-} 
+are plotted in Fig.~\ref{fig:cacl-dens_PCPS} for the negatively charged bilayer. 
+In contrast to \ce{KCl} or added concentrations of \ce{NaCl}, 
+\ce{Na^+} counterions are substituted with \ce{Ca^{2+}} even at low concentrations of \ce{CaCl2}. 
+The increasing concentration of \ce{CaCl2} and, hence, a higher amount of bound \ce{Ca^{2+}}
+also attracts \ce{Cl^-} anions to the bilayer 
+as can be seen from its growing density at the interface.
+
+%%% discuss the table with Gammas and follow up with molecular details, where do cations bind? Support with the spatial density figure
+The density profiles of ions suggest that
+the dominant contribution to the binding of \ce{Ca^{2+}} to phospholipid bilayers
+comes from the interactions with the phosphate groups in both POPC and POPS. 
+%The population analysis also suggests 
+%that the calcium cations prefer to reside in the phosphate region of the membrane. 
+%Such a finding is further validated with a more detailed analysis using Markov state modeling (MSM) \citep{Pande_MSM_paper} \todoi{Add papers reviewing MSM, e.g. recent Pande's paper.}. 
+%The set of states of a calcium cation 
+%encompassed all possible combinations of up to three surrounding lipids. 
+%In the case of PC, we used only the phosphate moitety, which forms the dominant contribution for calcium binding.
+%For PS we also distinguished configurations in which calcium interacts with the carboxylate moiety. 
+%All possible combinations of such states were used to build a Markov model at a lag time $25\,\mathrm{ns}$ using pyEMMA code by \citet{pyemma} \todoi{add citation for pyemma}. 
+%The resulting MSM was validated using Chapman-Kolmogorov test \citep{FrankNoe_papers_MSM} \todoi{Add citation for Noe's papers on MSM, especially Chapman Kolmogorov test.}. 
+%The stationary distribution of the states reveals a strong preference of the calcium cations to reside in the phosphate region of the mixed PC-PS membrane. 
+%When interacting with PS, configurations with the phosphate moiety from either PC or PS dominate the population.
+%Moreover, states containing interactions with the carboxylate group in PS 
+%bear higher probability when interacting also with the phosphate group of the same lipid or other lipids. 
+This is also reflected in the shift of the mean orientation 
+of the \ce{COO^-} group from $62^\circ$ to $73^\circ$ (420~mM \ce{CaCl2})
+which was measured as the connector of the carbon atoms 
+forming the bond between the group and the $\beta$-carbon of the phospholipid. 
+The interactions of the carboxylate group in PS with calcium and other phosphate groups
+shed light into the qualitatively different response of the head group order parameters $\alpha$ and $\beta$ in PS compared to PC 
+(see Figs.~\ref{fig:delta_ordPar_CaCl} and~\ref{fig:delta_ordPar_CaCl_PCPS}). 
+We find that the complex response of the head group order parameters of POPS 
+is affected by the conformational changes of the carboxylate group,
+which is attracted more towards the phosphate region, 
+where the calcium cations dominantly bind. 
+In line with the experimental work by \citet{browning80},
+this is also very likely the reason, 
+why the magnitude of the P-N vector change in POPS is diminished compared to POPC, 
+which is not restrained by an additional cation binding group like \ce{COO^-} in POPS.
+
+Further details about the interactions of \ce{Ca^{2+}} with various moieties in POPC or POPS
+were obtained by counting contacts between the cations and the oxygen atoms of the lipids
+similarly as was done in \cite{melcr18}. 
+The threshold for counting a contact was set to $0.3\,\mathrm{nm}$, 
+which encompasses the first peak of the radial distribution function between the cations and the oxygen atoms of the lipids.
+
+\begin{table}[tb!] 
+\centering
+  \caption{Bulk concentrations, $C _{Ca}$, and molar fractions, $C' _{Ca}$, of Ca$^{2+}$;
+           relative surface excess of calcium with respect to water ($\Gamma_{Ca}^{\rm water}$); 
+           and percentages of the population 
+           of bound Ca$^{2+}$ to various moieties 
+           in a neutral membrane composed of POPC
+           and in a negatively charged membrane with the composition 5\,PC:1\,PS.
+           \label{tab:binding}} 
+  \begin{tabular}{ l | c c } 
+	                     &  5\,POPC:1\,POPS &  POPC   \\
+	\hline
+	$C _{Ca}\,/\,\mathrm{mM}$  &  $240\pm 10 $  &  $280\pm 10 $  \\
+	$C'_{Ca}\,/\,\mathrm{mM}$  &  $400\pm 10 $  &  $350\pm 10 $  \\
+	$\Gamma_{Ca}^{\rm water}\, / \,\mathrm{nm}^{-2}$  &  $0.24 \pm 0.01 $  &  $0.06 \pm 0.01 $  \\
+	\hline
+                             &  \multicolumn{2}{c}{ } \\
+        interacting moiety   &  \multicolumn{2}{c}{percentage of bound \ce{Ca^{2+}} } \\
+	\hline
+	     PC              &   59   &  100   \\
+	     PO$_4$    in PC &   41   &   67   \\
+	     carbonyls in PC &   <1   &   ~1   \\
+	\hline
+	     PS              &    8   &        \\ 
+	     PO$_4$  in PS   &    2   &        \\
+	     COO$^-$ in PS   &    4   &        \\
+	     carbonyls in PS &   <1   &        \\
+	\hline
+	both PC and PS       &   33   &        \\
+  \end{tabular} 
+\end{table}
+
+The percentages of the populations of membrane-bound calcium cations for various membrane moieties 
+are summarized in Table~\ref{tab:binding}.
+Even though the negatively charged membrane contains only 18\% of POPS, 
+approximately 41\% of the total population of bound calcium cations is in contact with PS lipids
+with 8\% bound only to them. 
+This corroborates the intrinsically higher affinity of PS lipids to calcium cations compared to neutral PC lipids. 
+POPC interacts with the calcium cations almost entirely through its phosphate group 
+in both neutral and negatively charged membranes. 
+Interactions of \ce{Ca^{2+}} with carbonyl groups are also present, 
+however, they are always accompanied by interactions with phosphate groups. 
+%but a very small negligable amount much below 1\%.
+
+\begin{figure}[tb!] 
+  \centering 
+  \includegraphics[width=\figwidth]{../img/stoichiometry_CaCl2_comparison_Ecc-lipids_PC-vs-PCPS.pdf} \\ 
+  \caption{\label{fig:cacl_complexes} 
+      Relative probabilities of existence of \ce{Ca^{2+}} complexes 
+      with a certain number of lipids.  
+      All lipids were taken into account with the exception of the complexes in light green and orange, 
+      for which we counted only contacts with POPC resp POPS from the mixed 5\,PC:1\,PS negatively charged bilayer. 
+      The calculated probabilities of the calcium-lipid complexes also reflect only POPC (light green) resp POPS (orange). 
+      Probabilities were taken from simulations with comparable bulk concentrations of calcium around 250~mM. 
+      Clusters of four or more lipids were not observed in either membrane. 
+      Simulation data for POPC bilayers were taken directly from \cite{melcr18}. 
+  } 
+\end{figure}
+
+Relative probabilities of \ce{Ca^{2+}} complexes with a certain number of lipids are presented in Fig.~\ref{fig:cacl_complexes}. 
+Calcium cations that are bound only to PC in the mixed bilayer with PS 
+behave similarly as in the pure PC bilayer
+maintaining comparable probabilities for clustering one, two, or even three PC lipids together. 
+In contrast, PS lipids prefer 1:1 ratio with \ce{Ca^{2+}},
+which may also be due to their low molar fraction in the the mixed bilayer. 
+In total, however, the negatively charged membrane has its stoichiometry 
+shifted towards complexes with three phospholipids and one calcium. 
+This is obviously due to the presence of POPS, 
+which also contributes to the pre-formed one- and two-membered clusters of POPC. 
+Clusters of four or more lipids were not observed in either membrane.
+
+%This is also reflected in the increased probability 
+%of a single phospholipid interacting with two \ce{Ca^{2+}} cations 
+%compared to the neutral POPC bilayer, 
+%where such a probability is almost negligable. 
+% Not shown through numbers, however, I have the plots to back it.
+
+%\todo{Stationary distribution: Make a figure documenting the populations of bound \ce{Ca^{2+}} cations (like I have in the presentation) that would accompany Table \ref{tab:Ca_binding_PCPS}. 
+%This will roughly correspond to the PC stoichiometry plot \ref{fig:cacl_complexes}. }
+
+\begin{figure}[tb!]
+  \centering
+  \includegraphics[width=\figwidth]{../img/histogram_bound_times_26CaCl2_comparison_PC-PCPS.pdf}
+  \caption{\label{fig:hist_residence_times}
+   Histograms of residence times of \ce{Ca^{2+}} 
+   in a neutral membrane composed of POPC (orange)
+   and in a negatively charged membrane with the composition 5\,PC:1\,PS (blue)
+   from simulations with ECC-lipids and ECC-ions.
+   The simulation with the neutral membrane has a bulk concentration of calcium $C_{ion} = 280\mathrm{mM}$, 
+   the simulation with the negatively charged membrane has a bulk concentration of calcium $C_{ion} = 240\mathrm{mM}$. 
+   In the simulation with the neutral membrane, 
+   90\% of the residence times of calcium cations are
+   shorter than $60\,\mathrm{ns}$, % exactly $53\,\mathrm{ns}$                                                                          
+   with the longest observed residence time being $141\,\mathrm{ns}$. 
+   In the simulation with the negatively charged membrane, 
+   90\% of the residence times of calcium cations are
+   shorter than $200\,\mathrm{ns}$, % exactly $53\,\mathrm{ns}$                                                                          
+   with the longest observed residence time being $485\,\mathrm{ns}$. 
+   Simulation data for POPC bilayers were taken directly from \cite{melcr18}. 
+   }
+\end{figure}
+
+Timescales associated with the binding of calcium cations from solution to the membrane
+are plotted for each binding event as a histogram in Fig.~\ref{fig:hist_residence_times}. 
+Using these plots, we can estimate that 90\% of the residence times of any calcium cation 
+will be lower than $60\,\mathrm{ns}$ for pure POPC neutral bilayer 
+and shorter than $200\,\mathrm{ns}$ for the mixed 5\,PC:1\,PS negatively charged bilayer. 
+The longest observed residence times in the simulations were $141\,\mathrm{ns}$ for the neutral membrane 
+and $485\,\mathrm{ns}$ for the negatively charged membrane. 
+Both estimates of the residence times come from simulations with comparable concentrations of around $250\mathrm{mM}$;
+the simulation with the neutral membrane has a bulk concentration of calcium $C_{ion} = 280\mathrm{mM}$, 
+whereas the simulation with the negatively charged membrane has a bulk concentration of calcium $C_{ion} = 240\mathrm{mM}$.
+
+In summary, the results from ECC-lipids suggest 
+that the exchange of calcium between the POPC bilayer and the solvent 
+occurs at the order of $\sim$10--100~ns, 
+which is significantly faster than observed in simulations 
+with other presently available non-polarizable models of lipids and ions~\citep{javanainen17, catte16}. 
+Our results suggest 
+that simulation trajectories with a characteristic length of several hundreds of nanoseconds 
+are necessary to capture the binding of calcium to neutral POPC bilayers 
+in equilibrium when more realistic \emph{polarizable} force fields are used. 
+Interestingly enough, almost an order of magnitude longer simulations
+are required for the negatively charged bilayers.
+
+%In addition to such estimates of the time scales, we used the Markov model on top of the simulation with the negatviely charged mixed bilayer
+%to calculate the time of the mean first passage of calcium from solution to the membrane (and in reverse) resulting in $55\,\mathrm{ns}$  ($165\,\mathrm{ns}$).
+
+%\todo{Fluxes: committor analysis, dominant fluxes (table and figure)}
+%From the spectrum of the transition matrix, we also observe that 
+%the slowest transitions are associated with the binding to two or three phosphate moieties from either PC or PS,
+%which are also among the states with the highest probabilities.
+%The net fluxes of calcium cations from solution to such states also form a large contribution to the total flux ($\approx 45\%$). 
+%\todo{Make a table and a figure of the state probabilities and fluxes to support this statement.}. 
+%Analysis of the possible binding pathways of calcium cations to the negatively charged mixed bilayer 
+%reveal that the cations mostly enter the membrane bound states directly from solution 
+%without complicated transitions at the time scales of the order of the lag time of the Markov state model, $25\,\mathrm{ns}$.
 
@@
 
-Where do the cations bind? Preferentially to POPS. Discuss further \dots
+\subsection{Electronic Continuum Correction as a generally applicable model polarizability?}
+
+\todo{Look up possible applications of AMOEBA and Drude on anions/cations and possibly lipids. 
+Does ECC provide improvment over them? 
+This is crucial -- if this section will make the core message of this paper, 
+I have to put the ECC-method into context of other polarizable models. 
+If there are no other simulations like these and no data on the accuracy with cation/anions, 
+our approach (ECC) is hence the pioneer in the field and, hence, sets the standard instead. 
+}
+
+Show the successes and possible \emph{limits} of ECC applied on top of various models \cite{melcr18, martinek17, Pluharova2014, kohagen14, kohagen16}.
 
@@
 
-\begin{table}[tb!] 
-  \caption{Bulk concentrations of Ca$^{2+}$ (C$_b$), relative surface excess of calcium with respect to water ($\Gamma_{Ca}^{\rm water}$), 
-    and the percetages of Ca$^{2+}$ bound to phosphate or carbonyl oxyges ($r^\mathrm{Ca^{2+}} _\mathrm{PO_4} $ and $r^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}}$) 
-    in different mixed POPC:POPS (5:1) bilayer models. 
-  \label{tab:binding}} 
-  \begin{tabular}{l|c c | c | c c} 
-    model                  & $C_{ion}'$ & $C_{ion}\,/\,\mathrm{mM}$ & $\Gamma_{Ca}^{\rm water} \,/\, \mathrm{nm}^{-2}$  & $r^\mathrm{Ca^{2+}} _\mathrm{PO_4} $ & $r^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}} $ \\
-    \hline 
-    ECC-POPS               &  tba  &  $tba\pm 10 $  &  $0.?? \pm 0.01 $   &  ??\%  &    ??\%    \\ 
-    Lipid14/Dang           &  tba  &  $tba\pm 10 $  &  $0.?? \pm 0.03 $   &  ??\%  &    ??\%    \\ 
-  \end{tabular} 
-\end{table}
+\section{Conclusions}
 
-%\begin{figure}[htbp!] 
-%  \centering 
-%  \includegraphics[height=19.9cm]{../Fig/}
-%  \caption{\label{fig:cacl-dens} 
-%    Number density profiles of \ce{Ca^{2+}}, \ce{Na^{+}} and \ce{Cl^-} along membrane normal axis 
-%    for different force fields. 
-%    } 
-%\end{figure}
-
-%\subsection{Molecular interactions between K$^+$, Na$^+$ or Ca$^{2+}$ cations and mixed POPC:POPS membrane}
-
-%\subsection{Binding stoichiometry of cations to mixed POPC:POPS membrane}
-
-%\subsection{Residence times of cations in the mixed POPC:POPS membrane}
-
-%\begin{figure}[tb!] 
-%  \centering 
-%  \includegraphics[width=8.0cm]{../Fig/ipython_nb/stoichiometry_NaCl-CaCl2_comparison_Ecc-lipids.pdf} \\ 
-%  \caption{\label{fig:cacl_complexes} 
-%      Relative probabilities of existence of \ce{Na^{+}} or \ce{Ca^{2+}} complexes 
-%      with a certain number of POPS lipids.  
-%      \ce{Na^{+}} complexes were evaluated from the simulation with 1~M concentration; 
-%      and \ce{Ca^{2+}} complexes were evaluated from the simulation with 287~mM concentration. 
-%  } 
-%\end{figure}
-
-\subsection{Electronic Continuum Correction as a generally applicable model polarizability?}
-
-Show the successes and possible \emph{limits} of ECC applied on top of various models \cite{melcr18, martinek17, Pluharova2014, kohagen14, kohagen16}.
-
-\section{Conclusions}
+\listoftodos
 

### 2018-08-09 Manuscript/manuscript.tex
Add some comments in the manuscript

@@
 Tampere, Finland} 
-
+%SAMULI: We need to put tiago because the experimental acyl chain order parameters
+\author{Tiago Ferreira}
+\affiliation{Halle} 
 \author{O. H. Samuli Ollila} 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
+%\title[] 
+%  {Accurate interactions of cations with 
+%   neutral and negatively charged membranes 
+%   via combination of experiments and molecular simulation}
+
+%%SAMULI: new suggestion for the preliminary title
 \title[] 
-  {Accurate interactions of cations with 
-   neutral and negatively charged membranes 
-   via combination of experiments and molecular simulation}
+  {Effective Inclusion of Electronic Polarization to a Negatively Charged Membrane}
 
@@
 
-Binding affinities and stoichiometries of sodium and calcium ions to phospholipid bilayers are of paramount significance in the properties and functionality of cellular membranes.
+%%% SAMULI: I commented this out because it way too general. We can write abstract later when the final content is clear.
+
+%Binding affinities and stoichiometries of sodium and calcium ions to phospholipid bilayers are of paramount significance in the properties and functionality of cellular membranes.
 Current estimates of binding affinities and stoichiometries of cations are, however, controversial due to limitations in the available experimental and theoretical methods.
@@
 
-Classical molecular dynamics (MD) simulations provide details of the ion binding process with atomistic resolution, therefore offering all the necessary information to interpret experimental data without the need to resort to over-simplified models.
+%Classical molecular dynamics (MD) simulations provide details of the ion binding process with atomistic resolution, therefore offering all the necessary information to interpret experimental data without the need to resort to over-simplified models.
 However, the accuracy of the currently available lipid models when interacting with ions is not sufficient for such interpretation \cite{catte16}.
@@
 
-We are now extending our work to other lipid membranes composed of larger varieties of different biologically relevant lipids. In particular, we focus our attention on membranes with 
-	%phosphatidylethanolamine (PE) and 
-	negatively charged phosphatidylserine (PS) lipids. 
-	We compare available NMR and x-ray experiments with MD simulations to obtain atomistic-resolution picture of the interactions between cations and phospholipid membranes.
+%We are now extending our work to other lipid membranes composed of larger varieties of different biologically relevant lipids. In particular, we focus our attention on membranes with 
+%	%phosphatidylethanolamine (PE) and 
+%	negatively charged phosphatidylserine (PS) lipids. 
+%	We compare available NMR and x-ray experiments with MD simulations to obtain atomistic-resolution picture of the interactions between cations and phospholipid membranes.
 Our results indicate that electronic polarization is a non-negligable interaction, which dramatically improves the description of cation binding to phospholipid membranes.
@@
 
+\todo{SAMULI: I can write preliminary version of the introduction asap.}
+
 \section{Methods}
@@
 \subsection{Electronic continuum correction for lipid bilayers}\label{section:ecc}
+
+\todo{SAMULI: I think that all the current content is too general for this paper. The focus should be
+in application of ECC to negatively charge lipids versus zwitterionic lipid in the previous paper.}
 
@@
 
-\subsection{Electrometer concept} \label{section:electrometer} 
-This section may not be required, 
-we may just cite ECC-POPS and NMRlipids projects papers (and ongoing works)
+% SAMULI:
+%\subsection{Electrometer concept} \label{section:electrometer} 
+%This section may not be required, 
+%we may just cite ECC-POPS and NMRlipids projects papers (and ongoing works)
+% 
+% 
+% 
+%\subsection{Validation of lipid bilayer structure against NMR and scattering experiments} 
+%\emph{This section probably has to stay. Following text to be modified. Taken from Ref.~\citenum{melcr18}}
+%
+%
 
-\subsection{Validation of lipid bilayer structure against NMR and scattering experiments} 
-\emph{This section probably has to stay. Following text to be modified. Taken from Ref.~\citenum{melcr18}}
+\subsection{Validation of membrane properties} 
+\todo{SAMULI: I would put under under this section.
+1) a very short explanation of the usage of electrometer concept in this paper by mainly citing NMRlipids IV.
+2) a very short explanation of the usage of form factor and order parameters by citing approriate publications}
 
@@
 
+\subsection{Measurements of acyl chain order parameters}
+\todo{SAMULI: Short description about the order parameter experiments by Tiago.}
+
 \subsection{Simulation details}
 
+\todo{SAMULI: Table of simulated systems would be useful.}
 General simulation setup.
@@
 
+\todo{SAMULI: This section should be titled something like ''PS-PC headgroup interactions in mixtures'',
+unless we actually run simulations with anionic amphiphilic molecules.}
+
 Changes of the lipid bilayer head group order parameters extracted from simulations and 
@@
   } 
-\todo{Combine with Fig.~\ref{fig:delta_ordPar_NaCl_PC-PS_mix} into a single plot.}
+\todo{Combine with Fig.~\ref{fig:delta_ordPar_NaCl_PC-PS_mix} into a single plot.} \\
+\todo{SAMULI: I do not see any reason why we should show this data in this paper at all.}
 \end{figure}
@@
   } 
+\todo{SAMULI: It would be probably better to show also here the Lipid17 and ECC results in the same 
+  plot as done in the other figures.}
 \end{figure}
@@
   } 
+  \todo{SAMULI: The forking of PS alpha carbon should not be put to zero with zero salt concentrations
+    in this or in other figures. This is confusing expecially for the results with CaCl$_2$ where the forking
+    seems to increase with CaCl$_2$, which is not correct. In NMRlipids IV, I have now put the lower order parameter
+    to zero and kept the forking correct with all concentrations of ions.}
 \end{figure}
@@
     } 
-\todo{JM: Compare with pure POPC bilayer?}
+\todo{JM: Compare with pure POPC bilayer? 
+SAMULI: Comparison of the relative surface excess values in the text and in the table would be probably sufficient.}
 \end{figure}
@@
 our approach (ECC) is hence the pioneer in the field and, hence, sets the standard instead. 
-}
+SAMULI: We should keep the focus on lipids here. We should check the recent literature, but
+I am not aware of charged lipid simulations with explicitly polarizable force fields. 
+As for PC, the charge scaling was used in the early simulations of PS lipids from Berendsen group.
+We should mention this. Otherwise, I think that one of our conclusions will be that our
+results indicate that ECC generally improves the results of lipids bilayer MD simulations.
+However, we should not write a separate section about it. It should be in the conclusions section.}
 

### 2018-08-16 Manuscript/manuscript.tex
Figure PC-PS interaction contains the correct experimental series now. Data with both Na+ and K+ counterions plotted. Few details in text updated, e.g. the main message.

@@
 \begin{abstract}
-	\emph{Adapt abstract to the final message: properties of POPS-containing bilayers -or- ECC-concept applicability?}
+
+\textbf{Main message: 
+ECC generally improves interaction and structure of both neutral and charged molecules. 
+Polarizability is necessarry for accurate simulations of cellular membranes at physiological ionic conditions.}
 
@@
 
-\subsection{Electrometer concept: cationic and anionic amphiphilic molecules}
-
-\todo{SAMULI: This section should be titled something like ''PS-PC headgroup interactions in mixtures'',
-unless we actually run simulations with anionic amphiphilic molecules.}
+\subsection{Interactions between PC and PS head groups in mixed PC:PS bilayers}
 
@@
   \centering 
-  \includegraphics[width=\figwidthsmall]{../Fig/l17/order_parameters_changes_A-B_PC-PS_mix_POPC_nacl.pdf} 
-  \includegraphics[width=\figwidthsmall]{../Fig/l17/order_parameters_changes_A-B_PC-PS_mix_POPS_nacl.pdf} 
   \includegraphics[width=\figwidthsmall]{../Fig/order_parameters_changes_A-B_PC-PS_mix_POPC_nacl.pdf} 
   \includegraphics[width=\figwidthsmall]{../Fig/order_parameters_changes_A-B_PC-PS_mix_POPS_nacl.pdf} 
+  \includegraphics[width=\figwidthsmall]{../Fig/order_parameters_changes_A-B_PC-PS_mix_POPC_kcl.pdf} 
+  \includegraphics[width=\figwidthsmall]{../Fig/order_parameters_changes_A-B_PC-PS_mix_POPS_kcl.pdf} 
   \caption{\label{fig:delta_ordPar_NaCl_PC-PS_mix} 
-    Changes of the head group order parameters of a POPC:POPS 5:1 bilayer as a function of PS content
-    Experimental points have inherent systematic error, as they come from different experiments.
-    A complete series of points can be found (?) in \cite{roux90} -- which is to be added to the plot.
+    Changes of the head group order parameters of a POPC:POPS (5:1) bilayer as a function of PS content
+    with \ce{Na^+} and \ce{K^+} counterions from simulations with ECC-lipids and Lipid17 \cite{lipid17-future} 
+    and experiments (only \ce{Na^+} counterions). \cite{roux90}. 
+    The underestimated response of ECC-lipids with \ce{Na^+} counterions 
+    is likely due to a still slightly overestimated binding affinity of \ce{Na^+} to the phospholipids,
+    which is corroborated by the series with \ce{K^+} counterions (lower affinity than \ce{Na^+}),
+    where the response is closer to the experiment (which uses only \ce{Na^+} counterions). 
+    Experimental points for POPS inherently contain larger systematic error, 
+    as they come from independent experiments.
   } 
-\todo{SAMULI: It would be probably better to show also here the Lipid17 and ECC results in the same 
-  plot as done in the other figures.}
+\todo{Josef: Simulation at 50:50 PC:PS might be useful for this plot (PC-PS interactions). 
+Now it may make an impression, that we are consciously avoiding it. }
 \end{figure}
@@
 
+\textbf{This will be exemplified on POPS throughout this paper and, hence, it does not require additional section. }
+
 \todo{Look up possible applications of AMOEBA and Drude on anions/cations and possibly lipids. 

### 2018-08-22 Manuscript/manuscript.tex
Add preliminary intro

@@
 
-\todo{SAMULI: I can write preliminary version of the introduction asap.}
+Lipids with negatively charged headgroups, like phophatidylserine
+play important role in biomembranes. Signaling domains interact with
+anionic lipid membranes due to electrostatic and specific interactions
+with negatively charged lipids \cite{??}.
+Anionic lipids and their interactions with ions play also role in membrane
+fusion \cite{??}.
+These interactions depend on
+chemical details of lipid headgroups and binding of counter and co-ions
+to anionic membranes.
+
+Various interaction schemes have been proposed for the negatively charged
+lipids with ions and other lipids. Some studies suggest that the
+binding constant to negatively charged lipids is similar to zwitterionic
+and the binding affinity is increased only due to the increased cation
+concentration in the vicinity of the membrane \cite{??}. On the other hand,
+PS headgroups can form dehydrated complexes with calcium \cite{??} and ?? are
+proposed to form complexes with ions \cite{??}. Different scenarios are
+difficult to evaluate experimentally, especially with physiological
+ion concentrations which are typically quite low.
+
+Spectroscopic experiments give detailed information about the
+interactions \cite{??}. The data is, however, often indirect and difficult
+to interpret. Classical molecular dynamics (MD) simulations give
+access to all the details, but the current force fields are not
+sufficiently accurate to interpret the interactions of cationic
+lipid headgroups with other lipids and ions \cite{NMRlipidsIV}.
+Implicit inclusion of electronic polarizability using
+electronic continuum correction (ECC) to lipid models
+substantially improved the interactions between cations and
+zwitterionic lipids \cite{melcr18}. Here, we extend the approach
+to anionic PS lipids to more accurately describe its interactions with ions
+and PC headgroups in mixed lipid bilayer. The results pave
+the way for more realistic MD simulations of negatively charged
+lipid bilayers to be used in wide range of applications.
 

### 2018-08-23 Manuscript/manuscript.tex
Add references to intor

--- 2018-08-16 Manuscript/manuscript.tex
@@
 
-\todo{SAMULI: I can write preliminary version of the introduction asap.}
+Lipids with negatively charged headgroups, like phophatidylserine
+play important role in biomembranes. Signaling domains interact with
+anionic lipid membranes due to electrostatic and specific interactions
+with negatively charged lipids \cite{lemmon08,leventis10}.
+Anionic lipids and their interactions with ions play also role in membrane
+fusion \cite{??}. \todo{Pavel can probably improve this and give a citation}
+These interactions depend on
+chemical details of lipid headgroups and binding of counter and co-ions
+to anionic membranes.
+
+Various interaction schemes have been proposed for the negatively charged
+lipids with ions and other lipids. Some studies suggest that the
+binding constant to negatively charged lipids is similar to zwitterionic
+and the binding affinity is increased only due to the increased cation
+concentration in the vicinity of the membrane \cite{seelig90,sinn06}.
+On the other hand, calcium forms dehydrated complexes with PS headgroups
+which cause phase separation \cite{hauser77,kurland79,hauser85,feigenson86,mattai89,roux90,roux91,boettcher11}.
+Different theories are difficult to evaluate experimentally,
+especially with physiological ion concentrations which are typically quite low.
+
+Spectroscopic experiments give detailed information about the
+interactions between ions and PS lipids, but the data is often indirect and difficult to
+interpret \cite{hauser77,kurland79,eisenberg79,hauser83,dluhy83,hauser85,feigenson86,mattai89,roux90,roux91,boettcher11}.
+Classical molecular dynamics (MD) simulations give
+access to all the details, but the current force fields are not
+sufficiently accurate to interpret the interactions between anionic
+lipid headgroups and other lipids or ions \cite{NMRlipidsIV}.
+Implicit inclusion of electronic polarizability using
+electronic continuum correction (ECC) to lipid models
+substantially improved the interactions between cations and
+zwitterionic lipids \cite{melcr18}. Here, we extend the approach
+to anionic PS lipids to more accurately describe their interactions with ions
+and PC headgroups in mixed lipid bilayers. The results pave
+the way for more realistic MD simulations of negatively charged
+lipid bilayers to be used in wide range of applications.
 

### 2018-09-05 Manuscript/manuscript.tex
started a table with all simulations

@@
 
-\todo{SAMULI: Table of simulated systems would be useful.}
+\begin{table*}[tbp]
+\centering
+\caption{  List of MD simulations. 
+           Given concentrations adopt the form $C_{bulk}/C_{buffer}$. 
+        }
+\todo{Joe: to be filled-in. I'll generate it automatically from my Jupyter notebooks.}  
+\label{IONsystems_conc}
+\begin{tabular}{l c c r r r r }
+  lipid/counter-ions &  KCl (mM)  &  NaCl (mM) & CaCl$_2$\,(mM) &  \footnote{Number of lipid molecules with the largest mole fraction}N$_{\rm l}$   &  \footnote{Number of water molecules}N$_{\rm w}$ \todoi{Should confirm that the amounts of water in experiments matched those in simulations.}   &  \footnote{Simulation temperature}T (K)  \\
+  \hline
+    POPC:POPS (5:1)/K$^+$  & tba & tba & 60:12 & 9999 & 298  \\
+    \hline
+    POPC:POPS (5:1)/Na$^+$ & tba & tba & 60:12 & 9999 & 298  \\
+    \hline
+\end{tabular}
+\end{table*}
+
 General simulation setup.

### 2018-09-05 Manuscript/manuscript.tex
Added basic information and sentences into section simulation details

@@
 \todo{Joe: to be filled-in. I'll generate it automatically from my Jupyter notebooks.}  
-\label{IONsystems_conc}
+\label{tbl:sim-list}
 \begin{tabular}{l c c r r r r }
@@
 
-General simulation setup.
-
-Any specifics for simulations with aqueous cations.
-
-Any specifics for simulations with anionic/cationinc surfactants/peptides. 
-e.g. where the parameters for them came from.
+MD simulations were performed using the GROMACS \cite{Abraham15} simulation package (version 2018). 
+The simulation parameters used in this work are summarized in Table~\ref{tbl:mdpar}. 
+A comprehensive list of simulaitons used in this work is in Table~\ref{tbl:sim-list}.
+
+\begin{table}[tbp]
+  \caption{Simulation parameters}
+  \label{tbl:mdpar}
+  \begin{tabular}{ll}
+    simulation property & parameter   \\
+    \hline
+    time-step           & 2~fs         \\
+    equilibration time  & tba~ns  \\
+    total simulation time     & 1.5~$\mu$s  \\
+    temperature         & 298~K       \\
+    thermostat          & v-rescale  \cite{bussi07}   \\
+    barostat            & Parrinello-Rahman, semi-isotropic \cite{parrinello81} \\
+    long-range electrostatics & PME  \cite{darden93}  \\
+    cut-off scheme      & Verlet \cite{Pall13}      \\
+    Coulomb and VdW cut-off & 1.0~nm \\
+    constraints         & LINCS, only hydrogen atoms \cite{hess97} \\
+    constraints for water & SETTLE  \cite{miyamoto92} \\
+    \hline
+  \end{tabular}
+\end{table}
+
+For deposited trajectories and parameter files see 
+Refs.~\citenum{TBA}.
+\todo{Joe: Upload trajectories and add references here.}
 

### 2018-09-06 Manuscript/manuscript.tex
Improved simulation details

@@
 
+MD simulations were performed using the GROMACS 2018 \cite{Abraham15} simulation package. 
+The simulated system consisted of lipids (POPC and POPS) forming a single lipid bilayer,
+an amount of water molecules comparable to the hydration levels in experiments, 
+\todo{Check this claim.}
+and concentrations of ions (\ce{K^+}, \ce{Na^+}, \ce{Ca^{2+}} and \ce{Cl^{-}})
+in an orthorhombic simulation box with periodic boundary conditions. 
+The simulation parameters used in this work are summarized in Table~\ref{tbl:mdpar}. 
+A comprehensive list of simulations used in this work is in Table~\ref{tbl:sim-list}.
+
+We used the SPC/E~\cite{Berendsen1987} water model 
+and ECC-ions \cite{martinek17, kohagen16, Pluharova2014} 
+for the simulations with ECC-lipids.
+Force field parameters for POPC were taken from Ref.~\citenum{melcr18}. 
+The ECC-POPS model as derived again from the Lipid14 force field~\cite{dickson14, lipid17-future}, as was described above.
+For a direct comparison, 
+we also performed simulations with the Lipid14 model~\cite{dickson14},
+the TIP3P water model~\cite{jorgensen83}, 
+and ion models by Dang and coworkers~\cite{smith94,chang1999,dang2006}.
+
+For deposited trajectories and parameter files see Refs.~
+\todo{Joe: Upload trajectories and add references here.}
+
 \begin{table*}[tbp]
@@
 \end{table*}
-
-MD simulations were performed using the GROMACS \cite{Abraham15} simulation package (version 2018). 
-The simulation parameters used in this work are summarized in Table~\ref{tbl:mdpar}. 
-A comprehensive list of simulaitons used in this work is in Table~\ref{tbl:sim-list}.
 
@@
 \end{table}
-
-For deposited trajectories and parameter files see 
-Refs.~\citenum{TBA}.
-\todo{Joe: Upload trajectories and add references here.}
 

### 2018-09-06 Manuscript/manuscript.tex
Small updates -- mostly todos.

@@
 
+\todo{ Make a literature search for any possible applications of polarizable force fields (AMOEBA and Drude) on anions/cations and lipids. 
+If found -- Does ECC provide improvment over them? 
+As for PC, the charge scaling was used in the early simulations of PS lipids from Berendsen group.  }
+
 \section{Methods}
@@
 \subsection{Electronic continuum correction for lipid bilayers}\label{section:ecc}
+
+\todo{Joe: It might be a very short section -- Shall we merge it with the simulation details section?}
 
@@
 (bottom resp. top plot in Fig.~\ref{fig:nacl-dens_PCPS}). 
-The interaction of a neutral POPC bilayer with \ce{NaCl} is discussed in a greater detail in our publication \citep{melcr18}.
+The relative surface excess of calcium, $\Gamma^{w}_{Ca}$, also reflects
+the higher affinity of the negatively charged PC:PS (5:1) membrane compared to a neutral POPC bilayer; 
+its value for the latter is       $\Gamma^{w}_{Ca} = 0.06\mathrm{nm^{-2}}$, 
+while for the PC:PS mixture it is $\Gamma^{w}_{Ca} = 0.24\mathrm{nm^{-2}}$
+at comparable concentrations $C _{Ca} = 260\pm 30 \mathrm{mM} $.
 
@@
     } 
-\todo{JM: Compare with pure POPC bilayer? 
-SAMULI: Comparison of the relative surface excess values in the text and in the table would be probably sufficient.}
 \end{figure}
@@
 
-\subsection{Electronic Continuum Correction as a generally applicable model polarizability?}
-
-\textbf{This will be exemplified on POPS throughout this paper and, hence, it does not require additional section. }
-
-\todo{Look up possible applications of AMOEBA and Drude on anions/cations and possibly lipids. 
-Does ECC provide improvment over them? 
-This is crucial -- if this section will make the core message of this paper, 
-I have to put the ECC-method into context of other polarizable models. 
-If there are no other simulations like these and no data on the accuracy with cation/anions, 
-our approach (ECC) is hence the pioneer in the field and, hence, sets the standard instead. 
-SAMULI: We should keep the focus on lipids here. We should check the recent literature, but
-I am not aware of charged lipid simulations with explicitly polarizable force fields. 
-As for PC, the charge scaling was used in the early simulations of PS lipids from Berendsen group.
-We should mention this. Otherwise, I think that one of our conclusions will be that our
-results indicate that ECC generally improves the results of lipids bilayer MD simulations.
-However, we should not write a separate section about it. It should be in the conclusions section.}
-
-Show the successes and possible \emph{limits} of ECC applied on top of various models \cite{melcr18, martinek17, Pluharova2014, kohagen14, kohagen16}.
-
-After validation above, it can be an option to show results also for systems, for which we have no experiments.
-
 \section{Conclusions}

### 2018-09-06 Manuscript/manuscript.tex
Updated table list of simulations -- now shall contain almost all simulations with ECC-lipids The format of the table is yet to be discussed

@@
 \caption{  List of MD simulations. 
-           Given concentrations adopt the form $C_{bulk}/C_{buffer}$. 
+           %Given concentrations adopt the form $C_{bulk}/C_{buffer}$. 
         }
-\todo{Joe: to be filled-in. I'll generate it automatically from my Jupyter notebooks.}  
 \label{tbl:sim-list}
-\begin{tabular}{l c c r r r r }
-  lipid/counter-ions &  KCl (mM)  &  NaCl (mM) & CaCl$_2$\,(mM) &  \footnote{Number of lipid molecules with the largest mole fraction}N$_{\rm l}$   &  \footnote{Number of water molecules}N$_{\rm w}$ \todoi{Should confirm that the amounts of water in experiments matched those in simulations.}   &  \footnote{Simulation temperature}T (K)  \\
-  \hline
-    POPC:POPS (5:1)/K$^+$  & tba & tba & 60:12 & 9999 & 298  \\
-    \hline
-    POPC:POPS (5:1)/Na$^+$ & tba & tba & 60:12 & 9999 & 298  \\
+\begin{tabular}{l c c r r r r r }
+  lipid/counter-ions &  K  &  Na & Ca & Cl &  \footnote{Number of water molecules}N$_{\rm w}$ \todoi{Should confirm that the amounts of water in experiments matched those in simulations.}   &  \footnote{Number of lipid molecules with the largest mole fraction}N$_{\rm l}$     &  \footnote{Simulation temperature}T (K)  \\
+  \hline
+PC:PS (1:1)/CIon  &  0  &  64  &  0  &  0  &  5253  &  64:64  &  298 \\ 
+PC:PS (1:1)/CIon  &  64  &  0  &  0  &  0  &  5253  &  64:64  &  298 \\ 
+PC:PS (10:1)/CIon  &  0  &  6  &  0  &  0  &  3600  &  60:6  &  298 \\ 
+PC:PS (10:1)/CIon  &  6  &  0  &  0  &  0  &  3600  &  60:6  &  298 \\ 
+PC:PS (0:1)/CIon  &  72  &  0  &  0  &  0  &  3600  &  0:72  &  298 \\ 
+PC:PS (0:1)/CIon  &  0  &  72  &  0  &  0  &  3600  &  0:72  &  298 \\ 
+PC:PS (4:1)/CIon  &  0  &  12  &  0  &  0  &  3600  &  48:12  &  298 \\ 
+PC:PS (4:1)/CIon  &  12  &  0  &  0  &  0  &  3600  &  48:12  &  298 \\ 
+PC:PS (5:1)/CIon  &  0  &  12  &  13  &  26  &  3561  &  60:12  &  298 \\ 
+PC:PS (5:1)/CIon  &  0  &  12  &  26  &  52  &  3522  &  60:12  &  298 \\ 
+PC:PS (5:1)/CIon  &  0  &  12  &  0  &  0  &  3600  &  60:12  &  298 \\ 
+PC:PS (5:1)/CIon  &  12  &  0  &  0  &  0  &  3600  &  60:12  &  298 \\ 
+PC:PS (5:1)/CIon  &  0  &  12  &  39  &  78  &  3483  &  60:12  &  298 \\ 
+PC:PS (5:1)/CIon  &  39  &  12  &  0  &  39  &  3483  &  60:12  &  298 \\ 
+PC:PS (5:1)/CIon  &  0  &  51  &  0  &  39  &  3483  &  60:12  &  298 \\ 
   \hline

### 2018-09-07 Manuscript/manuscript.tex
Improved table list of simulations

--- 2018-09-06 Manuscript/manuscript.tex
@@
 \centering
-\caption{  List of MD simulations. 
-           Given concentrations adopt the form $C_{bulk}/C_{buffer}$. 
+\caption{  List of MD simulations with
+           numbers of individual molecules for each simulation
+           at temperature $T$. 
         }
-\todo{Joe: to be filled-in. I'll generate it automatically from my Jupyter notebooks.}  
 \label{tbl:sim-list}
-\begin{tabular}{l c c r r r r }
-  lipid/counter-ions &  KCl (mM)  &  NaCl (mM) & CaCl$_2$\,(mM) &  \footnote{Number of lipid molecules with the largest mole fraction}N$_{\rm l}$   &  \footnote{Number of water molecules}N$_{\rm w}$ \todoi{Should confirm that the amounts of water in experiments matched those in simulations.}   &  \footnote{Simulation temperature}T (K)  \\
-  \hline
-    POPC:POPS (5:1)/K$^+$  & tba & tba & 60:12 & 9999 & 298  \\
-    \hline
-    POPC:POPS (5:1)/Na$^+$ & tba & tba & 60:12 & 9999 & 298  \\
+\begin{tabular}{l c c r r r r r }
+  lipids/counter-ions &  \ce{K^+}  &  \ce{Na^+} & \ce{Ca^{2+}} & \ce{Cl^-} & \ce{H2O} &  PC:PS total  &  $T$~(K)  \\
+  \hline
+PC:PS (1:1)/Na+  &  0  &  64  &  0  &  0  &  5253  &  64:64  &  298 \\ 
+PC:PS (1:1)/K+  &  64  &  0  &  0  &  0  &  5253  &  64:64  &  298 \\ 
+PC:PS (10:1)/Na+  &  0  &  6  &  0  &  0  &  3600  &  60:6  &  298 \\ 
+PC:PS (10:1)/K+  &  6  &  0  &  0  &  0  &  3600  &  60:6  &  298 \\ 
+PC:PS (4:1)/Na+  &  0  &  12  &  0  &  0  &  3600  &  48:12  &  298 \\ 
+PC:PS (4:1)/K+  &  12  &  0  &  0  &  0  &  3600  &  48:12  &  298 \\ 
+PC:PS (5:1)/Na+  &  0  &  12  &  13  &  26  &  3561  &  60:12  &  298 \\ 
+PC:PS (5:1)/Na+  &  0  &  12  &  26  &  52  &  3522  &  60:12  &  298 \\ 
+PC:PS (5:1)/Na+  &  0  &  12  &  0  &  0  &  3600  &  60:12  &  298 \\ 
+PC:PS (5:1)/K+  &  12  &  0  &  0  &  0  &  3600  &  60:12  &  298 \\ 
+PC:PS (5:1)/Na+  &  0  &  12  &  39  &  78  &  3483  &  60:12  &  298 \\ 
+PC:PS (5:1)/Na+  &  39  &  12  &  0  &  39  &  3483  &  60:12  &  298 \\ 
+PC:PS (5:1)/Na+  &  0  &  51  &  0  &  39  &  3483  &  60:12  &  298 \\ 
+  \hline
+PS/K+  &  72  &  0  &  0  &  0  &  3600  &  0:72  &  298 \\ 
+PS/Na+  &  0  &  72  &  0  &  0  &  3600  &  0:72  &  298 \\ 
   \hline

### 2018-09-07 Manuscript/manuscript.tex
Updated PC-PS interaction plots (now also contain 50:50 mix)
Removed the plot POPC w.r.t surfactant DHMDAB

@@
 and in Figs.~\ref{fig:delta_ordPar_NaCl_PC-PS_mix} as a function of PC and PS content.
-and in Figs.~\ref{OrderParameterCHANGESsurf} as a function of the content of dihexadecylammonium bromide, a cationic amphiphile.
-
-\begin{figure}[tb!] 
-  \centering 
-  \includegraphics[width=\figwidth]{../img/ecc_popc/PN_angle_OrdPars-A-B_L14-ECCL17_q80_sig89_surf.pdf} 
-  \caption{\label{OrderParameterCHANGESsurf} 
-    Changes of the head group order parameters $\alpha$, $\beta$ and P-N vector orientation as a function of 
-    the molar fraction of a cationic surfactant dihexadecyldimethylammonium in a POPC bilayer 
-    from simulations and experiments \citep{scherer89} at 313 K.
-    Image reproduced from \cite{melcr18}.
-  } 
-\todo{Combine with Fig.~\ref{fig:delta_ordPar_NaCl_PC-PS_mix} into a single plot.} \\
-\todo{SAMULI: I do not see any reason why we should show this data in this paper at all.}
-\end{figure}
 
@@
   } 
-\todo{Josef: Simulation at 50:50 PC:PS might be useful for this plot (PC-PS interactions). 
-Now it may make an impression, that we are consciously avoiding it. }
 \end{figure}

### 2018-09-07 Manuscript/manuscript.tex
Written section validation of membrane properties

@@
 \subsection{Validation of membrane properties}
-\todo{SAMULI: I would put under under this section.
-1) a very short explanation of the usage of electrometer concept in this paper by mainly citing NMRlipids IV.
-2) a very short explanation of the usage of form factor and order parameters by citing approriate publications}
-
-The structures of lipid bilayers in simulations without ions were validated against NMR by calculating the order parameters for the C-H bonds and against \mbox{X-ray} scattering experiments by evaluating the scattering form factors.
-NMR order parameters validate the structures sampled by the individual lipid molecules with atomic resolution. 
-The simulated order parameters were calculated for all C-H bonds in lipid molecules from Eq. \ref{OP}.
-Scattering form factors validate the dimensions of the lipid bilayer (i.e., the bilayer thickness and area per molecule). Form factors were calculated using a relation
-
-\begin{equation} 
-  F(q) = \int _{-D/2} ^{D/2} \left ( \rho_{el}(z) - \rho_{el}^s \right ) \cos (zq_z) \mathrm{d}z, 
-  %F(q) = \int _{-D/2} ^{D/2} \left ( \sum _\alpha f_\alpha (q_z) n_\alpha (z) - \rho _s \right ) \exp (izq_z) \mathrm{d}z, 
-\end{equation}
-
-\noindent where $\rho_{el} (z)$ is the total electron density, $\rho_{el}^s$ is the electron density of the solvent far in the aqueous bulk, and $z$ is the distance from the membrane center along its normal with $D/2$ being half of the unit cell size.
+
+The structures of lipid bilayers in simulations without ions were validated against NMR 
+by calculating the order parameters for the C-H bonds and against \mbox{X-ray} scattering experiments by evaluating the scattering form factors
+as was done in Refs.~\citenum{catte16, melcr18, nmrlipids_proj4}.
+%NMR order parameters validate the structures sampled by the individual lipid molecules with atomic resolution. 
+The simulated order parameters were calculated for all C-H bonds in lipid molecules, 
+while all order paremeters in the head group and only some in the \ce{C-H} groups in tails were resolved experimentally
+(see next section and Fig.~\ref{simVSexpNOions_POPS}).
+
+Ion binding to lipid bilayers can be experimentally quantified
+by measuring the changes of the head group order parameters in the lipids using solid state NMR.
+In the case of the head group order parameters $S^\alpha$ and $S^\beta$ in phosphatidylcholine
+(see Fig.~\ref{simVSexpNOions_POPS} for the definition of the order parameters)
+such changes are known under the term ''electrometer concept'' \citep{seelig87,catte16, ollila16, melcr18, nmrlipids_proj4}. 
+It is experimentally observed that the \ce{C-H} bond
+order parameters of $S^\alpha$ and $S^\beta$ carbons in a PC lipid head group
+are proportional to the amount of charge, positive or negative, bound per lipid~\citep{seelig87, roux90}.
+The change of the order parameters measured with varying concentrations of aqueous ions 
+can be then related to the amount of bound ions.
 
@@
 The experimental X-ray scattering form factors 
-of a bilayer are well reproduced for all lipids employing the presented ECC-lipids model 
-(see Figs.~\ref{fig:simVSexpNOions}, \ref{simVSexpNOions_POPE} and \ref{simVSexpNOions_POPS}).
+of a bilayer are well reproduced  
+(see Fig.~\ref{simVSexpNOions_POPS}).
 

### 2018-09-10 Manuscript/manuscript.tex
Added simple lit search to the methods section. Still to be rewritten!

@@
 
-\emph{ Directly taken from my thesis.}
+\todo{The following contains reference, which are not added to the bibliography yet. 
+These paragraphs are not meant as a final intro -- it is rather a summary of my recent lit search. 
+Everything is to be completely rewritten. }
+
+Polarizability in classical MD simulations is a matter of an open discussion for many years. 
+While the most of the MD community uses and tries to reassure themselves
+that the lack of polarizability is not an issue for their applications,
+there are several groups which point at this apparent problem. 
+It was already demonstrated on several case problems
+that electronic polarization is a non-negligable part of the interactions between molecules. \cite{lopes2009}
+To give a few examples, 
+the community around ionic liquids considers not only polarization, 
+but also a charge transfer between moiteties an iportant effect in their simulations. \cite{borodin2009}
+Simulations of DNA suffer from overestimated interactions with divalent cations,
+which are much improved upon introduction of polarizability
+concluding that polarizability is crucial for a correct 
+characterization of structure and dynamics of DNA even without the ions. \cite{lemkul2016empirical}
+
+Using a common scaling factor is not a new approach in classical MD simulations,
+as it can be find in the early work on lipid bilayer simulation \cite{berkowitz06},
+in force fields employing explicit polarization \cite{lemkul2016empirical},
+in ionic liquids \cite{borodin2009} 
+\todo{Check these references whether they really contain scaling, especially the ionic liquid one.}
+Great improvement in permeability of charged solutes in gramicidin channel is observed with a force field using charge equilibration to model polarizability. \cite{lucas2012}
+
+Classical Durde model polarizable force field for phospholipids is gradually being developed,
+however, only PC and PE are available at this moment. \cite{chowdhary13, chowdhary17}
+While the polarizable model shows a potential improvement from many perspectives -- for example
+dipole potential of a membrane is found to be changed dramaticly 
+when polarizability is considered \cite{harder2009} -- 
+its non-polarizable version still performes better in some cases (e.g., in head group structure). 
+Anyhow, to date there is no study that would show any results 
+about negatively charged lipids using polarizable force fields
+nor about their interactions with cations and anions. 
+From this perspective,
+this work is the first to study interactions of POPS with monovalent and even divalent cations.
+
+Even though the software packages are more and more optimized for running simulations with explicit polarization,
+its application still seems to be rather far from ready for common production use. \cite{jiang2011, lopes2009}
 
@@
 %Permanent dipole and quadrupoles are also assigned to each atomic site for a more accurate description of electrostatic interactions. 
-The induced dipole approach models the polarizability through a point dipole moment on each atom.
-Such an electric dipole is induced by the acting electric field proportionally to its polarizability. 
+%The induced dipole approach models the polarizability through a point dipole moment on each atom.
+%Such an electric dipole is induced by the acting electric field proportionally to its polarizability. 
 This model is implemented in the widely adopted force field AMOEBA \citep{amoeba06, amoeba10, amoeba13}.
@@
 %Such analogy between the classical Drude oscillator model and the multipole and induced dipole (MPID) model suggests that in theory one can interconvert polarizable FFs between the two models. 
-The classical Drude oscillator model is inspired by a classical representation of a lone pair of electrons. 
-It uses an auxiliary charge,
-also termed Drude oscillator or Drude particle, 
-which is attached to each atom through a harmonic spring.
-The displacement of the Drude particle represents the induced polarization proportionally to the acting electric field,
+%The classical Drude oscillator model is inspired by a classical representation of a lone pair of electrons. 
+%It uses an auxiliary charge,
+%also termed Drude oscillator or Drude particle, 
+%which is attached to each atom through a harmonic spring.
+%The displacement of the Drude particle represents the induced polarization proportionally to the acting electric field,
 as implemented for example in the Drude force field \citep{lemkul2016empirical}.
@@
 This can be formally written as
-\begin{equation}  \label{eq:coulomb}
-   %U = \frac{1}{2} \displaystyle \sum ^N _{j\neq i} \frac{q_i q_j}{r_{ij}} + \frac{1}{2} \displaystyle \sum ^N _{i,j=1} U ^{dip} _{ij} - \displaystyle \sum ^N _{i=1} E^q(r_i) d_i  \propto  \frac{1}{2} \displaystyle \sum ^N _{j\neq i} \frac{q_i q_j}{\epsilon_{el} \, r_{ij}}  =  \frac{1}{2} \displaystyle \sum ^N _{j\neq i} \frac{q^{ECC}_i q^{ECC}_j}{r_{ij}}
-   %U = \frac{1}{2} \displaystyle \sum ^N _{j\neq i} \frac{q_i q_j}{r_{ij}} + \frac{1}{2} \displaystyle \sum ^N _{i,j=1} U ^{dip-dip} _{ij} - \displaystyle \sum ^N _{i=1} U^{q-dip}_i  \propto  \frac{1}{2} \displaystyle \sum ^N _{j\neq i} \frac{q_i q_j}{\epsilon_{el} \, r_{ij}}  =  \frac{1}{2} \displaystyle \sum ^N _{j\neq i} \frac{\frac{q_i}{\sqrt{\epsilon _{el}}} \frac{q_j}{\sqrt{\epsilon _{el}}}}{r_{ij}}  ,
-   U = \frac{1}{2} \displaystyle \sum ^N _{j\neq i} \frac{q_i q_j}{\epsilon_{el} \, r_{ij}}  =  \frac{1}{2} \displaystyle \sum ^N _{j\neq i} \frac{\frac{q_i}{\sqrt{\epsilon _{el}}} \frac{q_j}{\sqrt{\epsilon _{el}}}}{r_{ij}}  ,
-\end{equation} 
-where $q_i$ is the point charge of the particle $i$,
-and $r_{ij}$ are the distances between particles $i$ and $j$.
-%$U^{dip-dip} _{ij}$ represent the interactions between induced dipoles and their polarization energies,
+%\begin{equation}  \label{eq:coulomb}
+%   %U = \frac{1}{2} \displaystyle \sum ^N _{j\neq i} \frac{q_i q_j}{r_{ij}} + \frac{1}{2} \displaystyle \sum ^N _{i,j=1} U ^{dip} _{ij} - \displaystyle \sum ^N _{i=1} E^q(r_i) d_i  \propto  \frac{1}{2} \displaystyle \sum ^N _{j\neq i} \frac{q_i q_j}{\epsilon_{el} \, r_{ij}}  =  \frac{1}{2} \displaystyle \sum ^N _{j\neq i} \frac{q^{ECC}_i q^{ECC}_j}{r_{ij}}
+%   %U = \frac{1}{2} \displaystyle \sum ^N _{j\neq i} \frac{q_i q_j}{r_{ij}} + \frac{1}{2} \displaystyle \sum ^N _{i,j=1} U ^{dip-dip} _{ij} - \displaystyle \sum ^N _{i=1} U^{q-dip}_i  \propto  \frac{1}{2} \displaystyle \sum ^N _{j\neq i} \frac{q_i q_j}{\epsilon_{el} \, r_{ij}}  =  \frac{1}{2} \displaystyle \sum ^N _{j\neq i} \frac{\frac{q_i}{\sqrt{\epsilon _{el}}} \frac{q_j}{\sqrt{\epsilon _{el}}}}{r_{ij}}  ,
+%   U = \frac{1}{2} \displaystyle \sum ^N _{j\neq i} \frac{q_i q_j}{\epsilon_{el} \, r_{ij}}  =  \frac{1}{2} \displaystyle \sum ^N _{j\neq i} \frac{\frac{q_i}{\sqrt{\epsilon _{el}}} \frac{q_j}{\sqrt{\epsilon _{el}}}}{r_{ij}}  ,
+%\end{equation} 
+%where $q_i$ is the point charge of the particle $i$,
+%and $r_{ij}$ are the distances between particles $i$ and $j$.
+%%$U^{dip-dip} _{ij}$ represent the interactions between induced dipoles and their polarization energies,
 %and the terms $U^{q-dip} _i$ are the interactions between point charges and induced dipoles. 
-It naturally follows from the above equation that 
-embedding all atoms into a dielectric continuum 
-is equivalent to a formal scaling of the atomic point charges 
-\begin{equation}  \label{eq:scaling}
- q_i ^{ECC} = \frac{q_i}{\sqrt{\epsilon _{el}}} ,
-\end{equation} 
-where $q_i$ and $q_i^{ECC}$ are respectively the original partial charges and the scaled charges, 
-which represent the effects of the electronic polarization. 
+%It naturally follows from the above equation that 
+%embedding all atoms into a dielectric continuum 
+%is equivalent to a formal scaling of the atomic point charges 
+%\begin{equation}  \label{eq:scaling}
+% q_i ^{ECC} = \frac{q_i}{\sqrt{\epsilon _{el}}} ,
+%\end{equation} 
+%where $q_i$ and $q_i^{ECC}$ are respectively the original partial charges and the scaled charges, 
+%which represent the effects of the electronic polarization. 
 Given that the  high frequency dielectric constant of water is 1.78 (i.e., the square of the refraction index of 1.33), 

### 2018-09-14 Manuscript/manuscript.tex
Comments added into the manuscript

@@
 
-\todo{Joe: It might be a very short section -- Shall we merge it with the simulation details section?}
+\todo{Joe: It might be a very short section -- Shall we merge it with the simulation details section?
+SAMULI: Let's write it first and then see.}
 
@@
 
+\todo{SAMULI: Short discussion about mixtures should be added by citing the NMRlipids IV.}
 The structures of lipid bilayers in simulations without ions were validated against NMR 
@@
 and ion models by Dang and coworkers~\cite{smith94,chang1999,dang2006}.
+\todo{The usage of Dang ions should be justified here.}
 
@@
          }
+         \todo{SAMULI: Temperature column is not necessary because it is always the same.}
+         \todo{SAMULI: Columns reporting the additional ion concentrations would be useful.}
+         \todo{SAMULI: Order of the PC:PS (5:1)/Na+ systems in little bit confusing.}
 \label{tbl:sim-list}
@@
 
+\todo{SAMULI: Lipid17 simulations are too compressed, which is probably due to the
+  overbinding of Dang ions. Lipid17 with Åqvist ions (default in Amber) has larger
+  area in NMRlipids IV. This is similar to ECC-POPC paper where changing from Åqvist to
+  Dang or ECC-ions increased binding affinity in Lipid14. However, the headgroup order
+  parameters are better in ECC-POPS compared to Lipid17 simulations with any ions.
+  These should be discussed in this section.
+}
+
 We compared X-ray scattering form factors and NMR order parameters of bilayers
@@
 
-\begin{figure}[htb!] 
+\begin{figure*}[!htb] 
   \centering 
@@
   }
-\end{figure}
+  \todo{SAMULI: Caption says Åqvist, but I guess that Dang is used also here?} \\
+  \todo{SAMULI: Do you have a simulation of Lipid14 pure POPC at 298K?
+    This is not in the table. This simulation would be useful also for NMRlipids IV.
+  } \\
+  \todo{SAMULI: Error bars for lipid14 missing.}
+\end{figure*}
 
@@
     seems to increase with CaCl$_2$, which is not correct. In NMRlipids IV, I have now put the lower order parameter
-    to zero and kept the forking correct with all concentrations of ions.}
+    to zero and kept the forking correct with all concentrations of ions.} \\
+  \todo{SAMULI: Ion concentrations should be updated to the concentrations in water before solvating the lipids.} \\
+  \todo{SAMULI: Error bars should be recalculated with the fixed code} \\
+  \todo{SAMULI: We should have at least one more points with higher concentration (maybe close to 2M), or maybe two more points.} \\
   \end{figure}
@@
     seems to increase with CaCl$_2$, which is not correct. In NMRlipids IV, I have now put the lower order parameter
-    to zero and kept the forking correct with all concentrations of ions.}
+    to zero and kept the forking correct with all concentrations of ions.} \\              
+  \todo{SAMULI: Ion concentrations should be updated to the concentrations in water before solvating the lipids.} \\
+  \todo{SAMULI: Error bars should be recalculated with the fixed code} \\
+  \todo{SAMULI: We should have at least one more points with higher concentration (maybe close to 2M), or maybe two more points.} \\
 \end{figure}
@@
   }
+  \todo{SAMULI: Top figure should go together with figure of only K+ counterions and figure with pure PS mentioned
+    another todo point
+  } \\
+  \todo{SAMULI: Division of ion densities with 2 in figure without calcium is not necessary and makes
+  comparison to other results more difficult.}
 \end{figure}
@@
 can be described by their relative surface excesses with respect to water, $\Gamma ^{w} _{ion}$, 
-which are shown in the plots of the density profiles of the ions in Figs.~\ref{fig:nacl-dens} (neutral bilayer) and \ref{fig:nacl-dens_PCPS} (negatively charged bilayer). 
+which are shown in the plots of the density profiles of the ions in Figs.~\ref{fig:nacl-dens} (neutral bilayer)
+\todo{SAMULI: We should have a figure showing density profiles of Na+ and K+ in pure POPS and PC:PS (5:1) mixture
+  with only counterions from ECC-POPS model and Lipid17 model.
+}
+and \ref{fig:nacl-dens_PCPS} (negatively charged bilayer). 
 Such a quantity compares the adsorption of ions to the adsorption of water molecules at an interface 
@@
     seems to increase with CaCl$_2$, which is not correct. In NMRlipids IV, I have now put the lower order parameter
-    to zero and kept the forking correct with all concentrations of ions.}
+    to zero and kept the forking correct with all concentrations of ions.} \\
+  \todo{SAMULI: Ion concentrations should be updated to the concentrations in water before solvating the lipids.} \\
+  \todo{SAMULI: Error bars should be recalculated with the fixed code} \\
+  \todo{SAMULI: Maybe we should have at least one more points with smaller concentration
+    (maybe close to 100mM in solution before solvating lipids).} \\
 \end{figure}
@@
 which achieves a \emph{quantitative} agreement with the changes induced by \ce{CaCl2} in experiments.
+
+\todo{SAMULI: The qualitative iprovement of PS headgroup order parameter response to bound Ca2+
+  should be discussed. We should, however, point out that the response PS alpha carbon to bound
+  Ca2+ is not excatly reproduced even by the ECC-POPS model.
+}
 
@@
   }
+  \todo{SAMULI: Ion concentrations should be updated to the concentrations in water before solvating the lipids.} \\
+  \todo{SAMULI: We should compare to the density profiles from Lipid17 simulations. These are also needed for NMRlipids IV.}
 \end{figure}
@@
 
+\todo{SAMULI: This paragraph in unclear. Should be clarified and the changes
+  of headgroup orientations in PC and PS clearly discussed.
+}
 %%% discuss the table with Gammas and follow up with molecular details, where do cations bind? Support with the spatial density figure
@@
 %but a very small negligable amount much below 1\%. 
+\todo{Also COO- vs. phosphate binding should be discussed and the results should be
+  discussed with respect to the literature: https://doi.org/10.1016/j.bpj.2018.08.044 and
+  Melcrova et al. Scientific Reports | 6:38035 | DOI: 10.1038/srep38035 }
 

### 2018-09-21 Manuscript/manuscript.tex

@@
   \centering 
+  \includegraphics[width=\figwidth]{../img/density_profile_K-Na_counterions_compar_pure-ECC-POPS.pdf}
+  \caption{\label{fig:POPS-counterions-dens}
+    Number density profiles of \ce{K^{+}}, \ce{Na^{+}} counterions along the membrane normal axis 
+    for the negatively charged POPS bilayers using ECC-lipids and ECC-ions.  }
+\todo{Joe: this figure requires updating and polishing, but serves the purpose of “showing the data”.}
+\end{figure}
+
+\begin{figure}[tbp!] 
+  \centering 
   \includegraphics[width=\figwidth]{../img/ecc_pops/density_profiles_na_k_cl_wat_phos_models-compar_4-6_NaCl-and-KCl-series.pdf}

### 2018-09-26 Manuscript/manuscript.tex
Improved density-profile figure with PC:PS mixed membranes and added NaCl/KCl concentrations.

@@
     for the negatively charged POPS bilayers using ECC-lipids and ECC-ions.  }
-\todo{Joe: this figure requires updating and polishing, but serves the purpose of “showing the data”.}
+\todo{Joe: TO BE UPDATED + results from PC:PS mixutres to be added.}
 \end{figure}
@@
   \centering 
-  \includegraphics[width=\figwidth]{../img/ecc_pops/density_profiles_na_k_cl_wat_phos_models-compar_4-6_NaCl-and-KCl-series.pdf}
+  \includegraphics[width=\figwidth]{../img/ecc_pops/density_profiles_na_k_cl_wat_phos_models-compar_5-6_NaCl-and-KCl-series.pdf}
   \caption{\label{fig:nacl-dens_PCPS} 
     Number density profiles of \ce{K^{+}}, \ce{Na^{+}} and \ce{Cl^-} along the membrane normal axis 
-    for the negatively charged membranes with the composition of 5\,PC:1\,PS using ECC-lipids and ECC-ions.  
-    The top profile shows the simulation without any additional salt concentration, i.e., only with \ce{Na^+} counterions. 
-    The middle profile shows the simulation with an additional \ce{KCl} concentration and \ce{Na^+} counterions. 
+    for the negatively charged membranes with the composition of 5\,PC:1\,PS using ECC-lipids and ECC-ions
+    from simulations with added concentrations of \ce{KCl} or \ce{NaCl}.  
+    The top    profile shows the simulation with an additional \ce{KCl} concentration and \ce{Na^+} counterions. 
     The bottom profile shows the simulation with an additional \ce{NaCl} concentration and \ce{Na^+} counterions, which are not distinguished from the added salt. 
-    In order to visualize the density profiles with a scale comparable to the profile of \ce{Ca^{2+}}, 
-    the density profiles of~\ce{Cl^-}, \ce{K^+} and \ce{Na^+} ions are divided by 2, and 
-    the density profiles of phosphate groups and water are divided by 5 and 200, respectively.  
-  }
-  \todo{SAMULI: Top figure should go together with figure of only K+ counterions and figure with pure PS mentioned
-    another todo point
-  } \\
-  \todo{SAMULI: Division of ion densities with 2 in figure without calcium is not necessary and makes
-  comparison to other results more difficult.}
+    The density profiles of phosphate groups and water are divided by 4 and 100, respectively.  
+  }
 \end{figure}

### 2018-10-07 Manuscript/manuscript.tex
Written a short description of the choice of scaling factor 0.75 for PS compared to 0.8 for PC.in the methods section.
The other rubbish is still left there.

@@
 Everything is to be completely rewritten. }
+
+The polarization of electrons in ECC-lipids uses the implicit method termed electronic continuum correction (ECC). \cite{leontyev09, Pluharova2014, martinek17, melcr18} 
+In this method, the effects of electronic polarization are represented through rescaling of charges. \cite{melcr18}
+The scaling factor is derived from the high-frequency dielectric constant of electrons, $\epsilon_{el}$, as
+\begin{equation}
+f_q = \frac{1}{\sqrt{\epsilon_{el}}} .
+\end{equation}
+The value of $\epsilon_{el}$ for water solutions is $\epsilon_{el} = 1.78$ \cite{leontyev09},
+hence, the scaling factor yields $f_q \approx 0.75$.
+
+In the case of POPS, which is a monovalently negatively charged lipid, 
+we used exactly this scaling factor to represent the polarizability of the electrons
+yielding a total charge of $-0.75$ in the ECC formalism. \cite{martinek17, leontyev09, melcr18}
+
+Note that we used a different scaling factor of $f_q = 0.8$ 
+to derive a ECC-POPC model, which also uses ECC. \cite{melcr18}
+Compared to POPS, POPC is a zwitterionic neutral molecule with a zero total charge; 
+scaling it by any factor does not affect it. 
+Partial charges of the atoms in POPC are not physical observabels, 
+and vary greatly between different models. \cite{botan15, catte16, nmrlipids_proj4, Pluhackova2016}
+The choice of a larger scaling factor in ECC-POPC represents the fact that 
+the set of partial charges in the initial model for ECC-POPC, the Lipid14 model \cite{dickson14}, 
+was determined as less polar compared to other models.
+
+\textbf{This explains the choice of different scaling factor. uff..}
 

### 2018-10-13 Manuscript/manuscript.tex
Improved methods (ECC implementation) and part of intro

--- 2018-09-26 Manuscript/manuscript.tex
@@
 substantially improved the interactions between cations and
-zwitterionic lipids \cite{melcr18}. Here, we extend the approach
+zwitterionic lipids \cite{melcr18}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% On the simulation part .. %%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+An empirical scaling factor, which reduces electrostatic forces between molecules, 
+is a well known concept in molecular modelling;
+such an approach was adopted in the early simulations of lipid bilayers \cite{egberts94, berendsen1996},
+in force fields employing explicit polarization \cite{lemkul2016empirical},
+or in the simulations of ionic liquids \cite{Bhargava2007}. 
+%An improvement of permeability of charged solutes in the gramicidin A channel 
+%is observed with a force field using charge equilibration to model polarizability. \cite{lucas2012}
+
+From the perspective of phospholipid bilayers,
+a force field with an explicit polarizability 
+using the Classical Durde model 
+is gradually being developed, however, 
+only parameters for PC and PE are available at this moment. \cite{chowdhary13, chowdhary17}
+While such a model shows a potential improvement from many perspectives -- 
+for example accounting for polarizability has a significant impact on the dipole potential of a membrane \cite{harder2009} -- 
+the non-polarizable version of the same force field 
+performes comparably well at a fraction of the computational cost 
+of the models with explicit polarization. \citep{lucas12,chowdhary13}
+
+%To our knowledge, there is no study focusing on
+%charged lipids using polarizable force fields. 
+%nor about their interactions with cations and anions. 
+%From this perspective,
+%this work is the first pioneering study providing atomistic resolution insight 
+%into the interactions of POPS with monovalent and even divalent cations.
+
+%Even though the software packages are more and more optimized for running simulations with explicit polarization,
+%its application still seems to be rather far from ready for common production use. \cite{jiang2011, lopes2009}
+
+Here, we extend the approach
 to anionic PS lipids to more accurately describe their interactions with ions
@@
 
-\todo{ Make a literature search for any possible applications of polarizable force fields (AMOEBA and Drude) on anions/cations and lipids. 
-If found -- Does ECC provide improvment over them? 
-As for PC, the charge scaling was used in the early simulations of PS lipids from Berendsen group.  }
-
 \section{Methods}
@@
 
-\todo{Joe: It might be a very short section -- Shall we merge it with the simulation details section?
-SAMULI: Let's write it first and then see.}
-
-\todo{SAMULI: I think that all the current content is too general for this paper. The focus should be
-in application of ECC to negatively charge lipids versus zwitterionic lipid in the previous paper.}
-
-\todo{The following contains reference, which are not added to the bibliography yet. 
-These paragraphs are not meant as a final intro -- it is rather a summary of my recent lit search. 
-Everything is to be completely rewritten. }
-
-Polarizability in classical MD simulations is a matter of an open discussion for many years. 
-While the most of the MD community uses and tries to reassure themselves
-that the lack of polarizability is not an issue for their applications,
-there are several groups which point at this apparent problem. 
-It was already demonstrated on several case problems
-that electronic polarization is a non-negligable part of the interactions between molecules. \cite{lopes2009}
-To give a few examples, 
-the community around ionic liquids considers not only polarization, 
-but also a charge transfer between moiteties an iportant effect in their simulations. \cite{borodin2009}
-Simulations of DNA suffer from overestimated interactions with divalent cations,
-which are much improved upon introduction of polarizability
-concluding that polarizability is crucial for a correct 
-characterization of structure and dynamics of DNA even without the ions. \cite{lemkul2016empirical}
-
-Using a common scaling factor is not a new approach in classical MD simulations,
-as it can be find in the early work on lipid bilayer simulation \cite{berkowitz06},
-in force fields employing explicit polarization \cite{lemkul2016empirical},
-in ionic liquids \cite{borodin2009} 
-\todo{Check these references whether they really contain scaling, especially the ionic liquid one.}
-Great improvement in permeability of charged solutes in gramicidin channel is observed with a force field using charge equilibration to model polarizability. \cite{lucas2012}
-
-Classical Durde model polarizable force field for phospholipids is gradually being developed,
-however, only PC and PE are available at this moment. \cite{chowdhary13, chowdhary17}
-While the polarizable model shows a potential improvement from many perspectives -- for example
-dipole potential of a membrane is found to be changed dramaticly 
-when polarizability is considered \cite{harder2009} -- 
-its non-polarizable version still performes better in some cases (e.g., in head group structure). 
-Anyhow, to date there is no study that would show any results 
-about negatively charged lipids using polarizable force fields
-nor about their interactions with cations and anions. 
-From this perspective,
-this work is the first to study interactions of POPS with monovalent and even divalent cations.
-
-Even though the software packages are more and more optimized for running simulations with explicit polarization,
-its application still seems to be rather far from ready for common production use. \cite{jiang2011, lopes2009}
-
-The lack of electronic polarizability in classical MD simulations
-has been considered a serious issue since the early days of the modeling of lipid bilayers. 
-% the commented text on polarizability in this section comes from \citep{huang2017mapping}
-There exist two main approaches for the explicit representation of electronic polarizability -- atomic induced dipoles and Drude oscillators.
-%The AMOEBA FF adopts the point induced dipole model in which each atom carries an induced dipole moment determined by the electric field and its atomic polarizability. 
-%Permanent dipole and quadrupoles are also assigned to each atomic site for a more accurate description of electrostatic interactions. 
-%The induced dipole approach models the polarizability through a point dipole moment on each atom.
-%Such an electric dipole is induced by the acting electric field proportionally to its polarizability. 
-This model is implemented in the widely adopted force field AMOEBA \citep{amoeba06, amoeba10, amoeba13}.
-
-%The Drude FF adopts the classical Drude oscillator model where auxiliary charges (Drude oscillators or particles) are attached to non-hydrogen atoms with harmonic springs, and the displacement of the Drude particle relative to its parent atom represents the response to the external electric field.
-%Virtual particles representative of lone pairs are included in the Drude FF to improve the treatment of electrostatic interactions typically involving hydrogen bond acceptors.
-%The charge distribution between a Drude particle and its parent atom is a finite difference representation of a dipole moment. 
-%Similarly, lone pairs can be considered as real space representations of permanent multipoles. 
-%Such analogy between the classical Drude oscillator model and the multipole and induced dipole (MPID) model suggests that in theory one can interconvert polarizable FFs between the two models. 
-%The classical Drude oscillator model is inspired by a classical representation of a lone pair of electrons. 
-%It uses an auxiliary charge,
-%also termed Drude oscillator or Drude particle, 
-%which is attached to each atom through a harmonic spring.
-%The displacement of the Drude particle represents the induced polarization proportionally to the acting electric field,
-as implemented for example in the Drude force field \citep{lemkul2016empirical}.
-
-%In this work, we convert the Drude polarizable force field from its original classical Drude oscillator formalism onto a permanent multipole and induced dipole formalism, which we refer to as MPID. 
-%We show that a direct mapping without any reparametrization can transfer the Drude FF into the MPID FF, yielding a similar level of agreement as the Drude FF in reproducing condensed phase properties of organic liquids, crystals, and interfaces.
-%Our results thus also serve as an additional validation of these novel simulation algorithms.
-Although the two approaches, 
-the induced dipole model and the classical Drude oscillator model, 
-are very different in how they represent polarizability in simulations,
-it was shown that the classical Drude oscillator formalism can be converted to the induced dipole formalism 
-proving the two methods to be equivalent. \citep{huang2017mapping}
-From the point of view of lipids, however, the classical models with fixed charges and no explicit polarizability that are currently available 
-provide comparable accuracy at a fraction of the computational cost of the models with explicit polarization. \citep{lucas12,chowdhary13}
-
-%The effects of polarization from rather computationally demanding simulations with explicit models 
-%can be implemented using an implicit model of polarization. 
-Instead of the computationally demanding explicit polarization approach, 
-electronic polarization can also be accounted for implicitly.
-Such an implicit mean field model of electronic polarizability 
-%was developed by \citet{leontyev09} under the name Molecular Dynamics in Electronic Continuum, MDEC. 
-termed Electronic continuum correction (ECC) 
-was first suggested by \citet{leontyev09}, 
-and was then systematically applied to solutions of ions by \citet{Pluharova2014, kohagen14, kohagen16, martinek17}. 
-Through a comparison of neutron scattering and/or \emph{ab-initio} molecular dynamics data with classical MD simulations, 
-the authors developed an array of models for cations and anions, termed here as ECC-ions,
-which provide realistic structures even for concentrated solutions of both monovalent and divalent cations. 
-%The term Electronic continuum correction, ECC, originates in the works by \citet{Pluharova2014, kohagen14, kohagen16, martinek17},
-%where the rigorously developed theory by \citet{leontyev14} was systematicaly applied to solutions of ions.
-%that use ECC to represent electronic polarizability.
-
-%Electronic continuum correction (ECC) 
-In ECC, a system of polarizable particles is represented 
-with an equivalent system of particles with fixed charges. %\citep{leontyev09, leontyev10, leontyev11, leontyev14}
-%including the effects of electronic polarization \emph{implicitly} in a mean-field way. 
-This can be done by a simple transform of the partial charges,
-which does not change the form of the empirical force field in Equation~\ref{eq:amber}, 
-providing the effects of the electronic polarization in a mean field way at no additional computational cost. 
-Although this is technically similar to the phenomenological scaling of partial charges 
-applied in earlier studies of aqueous solutions or ionic liquids \citep{jonsson86,egberts94,beichel14},
-%where also other possible effects (e.g. charge transfer) were considered,
-ECC is a physically well justified and rigorously derived model \citep{leontyev09, leontyev10, leontyev11, leontyev14}.
-Assuming that the electronic polarizability of electrons is approximately constant in biological solutions, 
-it was shown by \citet{leontyev11} that the electrostatic interactions between atoms 
-are screened like in a dielectric continuum 
-with an electronic dielectric constant, $\epsilon _{el}$, 
-also known as the high frequency dielectric constant. 
-This can be formally written as
-%\begin{equation}  \label{eq:coulomb}
-%   %U = \frac{1}{2} \displaystyle \sum ^N _{j\neq i} \frac{q_i q_j}{r_{ij}} + \frac{1}{2} \displaystyle \sum ^N _{i,j=1} U ^{dip} _{ij} - \displaystyle \sum ^N _{i=1} E^q(r_i) d_i  \propto  \frac{1}{2} \displaystyle \sum ^N _{j\neq i} \frac{q_i q_j}{\epsilon_{el} \, r_{ij}}  =  \frac{1}{2} \displaystyle \sum ^N _{j\neq i} \frac{q^{ECC}_i q^{ECC}_j}{r_{ij}}
-%   %U = \frac{1}{2} \displaystyle \sum ^N _{j\neq i} \frac{q_i q_j}{r_{ij}} + \frac{1}{2} \displaystyle \sum ^N _{i,j=1} U ^{dip-dip} _{ij} - \displaystyle \sum ^N _{i=1} U^{q-dip}_i  \propto  \frac{1}{2} \displaystyle \sum ^N _{j\neq i} \frac{q_i q_j}{\epsilon_{el} \, r_{ij}}  =  \frac{1}{2} \displaystyle \sum ^N _{j\neq i} \frac{\frac{q_i}{\sqrt{\epsilon _{el}}} \frac{q_j}{\sqrt{\epsilon _{el}}}}{r_{ij}}  ,
-%   U = \frac{1}{2} \displaystyle \sum ^N _{j\neq i} \frac{q_i q_j}{\epsilon_{el} \, r_{ij}}  =  \frac{1}{2} \displaystyle \sum ^N _{j\neq i} \frac{\frac{q_i}{\sqrt{\epsilon _{el}}} \frac{q_j}{\sqrt{\epsilon _{el}}}}{r_{ij}}  ,
-%\end{equation} 
-%where $q_i$ is the point charge of the particle $i$,
-%and $r_{ij}$ are the distances between particles $i$ and $j$.
-%%$U^{dip-dip} _{ij}$ represent the interactions between induced dipoles and their polarization energies,
-%and the terms $U^{q-dip} _i$ are the interactions between point charges and induced dipoles. 
-%It naturally follows from the above equation that 
-%embedding all atoms into a dielectric continuum 
-%is equivalent to a formal scaling of the atomic point charges 
-%\begin{equation}  \label{eq:scaling}
-% q_i ^{ECC} = \frac{q_i}{\sqrt{\epsilon _{el}}} ,
-%\end{equation} 
-%where $q_i$ and $q_i^{ECC}$ are respectively the original partial charges and the scaled charges, 
-%which represent the effects of the electronic polarization. 
-Given that the  high frequency dielectric constant of water is 1.78 (i.e., the square of the refraction index of 1.33), 
-the scaling factor for ions in water is $\approx 0.75$.
+The polarization of electrons in ECC-lipids uses the implicit method termed electronic continuum correction (ECC). \cite{leontyev09, Pluharova2014, martinek17, melcr18} 
+In this method, a system of polarizable particles is represented with an equivalent system of particles with fixed charges. 
+This is practically achieved rescaling the charges with a pre-determined scaling factor, $f_q$,
+which is derived from the high-frequency dielectric constant of electrons, $\epsilon_{el}$, as
+\begin{equation}
+f_q = \frac{1}{\sqrt{\epsilon_{el}}} .
+\end{equation}
+The value of $\epsilon_{el}$ for water solutions is $\epsilon_{el} = 1.78$ \cite{leontyev09},
+hence, the scaling factor for aquaeous environments yields $f_q \approx 0.75$.
+
+In the case of POPS, which is a monovalently negatively charged lipid, 
+we used exactly this scaling factor 
+to represent the effects of electronic polarization
+yielding a total charge of $-0.75$ in the ECC formalism. \cite{martinek17, leontyev09, melcr18}
 
@@
 
-% SAMULI:
-%\subsection{Electrometer concept} \label{section:electrometer} 
-%This section may not be required, 
-%we may just cite ECC-POPS and NMRlipids projects papers (and ongoing works)
-% 
-% 
-% 
-%\subsection{Validation of lipid bilayer structure against NMR and scattering experiments} 
-%\emph{This section probably has to stay. Following text to be modified. Taken from Ref.~\citenum{melcr18}}
-%
-%
+Note that we used a different scaling factor $f_q = 0.8$ 
+to for the ECC-POPC model. \cite{melcr18}
+Compared to POPS, POPC is a zwitterionic neutral molecule with a zero total charge; 
+scaling it by any factor does not affect it. 
+Partial charges of the atoms in POPC are not physical observables, 
+and vary greatly between different models. \cite{botan15, catte16, nmrlipids_proj4, Pluhackova2016}
+The choice of a larger scaling factor in ECC-POPC represents the fact that 
+the set of partial charges in the initial model for ECC-POPC, the Lipid14 model \cite{dickson14}, 
+was determined in the beginning as less polar compared to other models.
 

### 2018-10-13 Manuscript/manuscript.tex
Small update to methods - scaling of POPS - where and why

@@
 
-In the case of POPS, which is a monovalently negatively charged lipid, 
-we used exactly this scaling factor 
-to represent the effects of electronic polarization
-yielding a total charge of $-0.75$ in the ECC formalism. \cite{martinek17, leontyev09, melcr18}
-
 It is important to note that the value of the high frequency dielectric constant  
@@
 applied throughout the lipid bilayer/aqueous solution interface.
+
+In the case of POPS, which is a monovalently negatively charged lipid, 
+we used exactly the pre-determined scaling factor 
+to represent the effects of electronic polarization
+yielding a total charge of $-0.75$ in the ECC formalism. \cite{martinek17, leontyev09, melcr18}
+Similarly to the ECC-POPC model, we changed the charges only of the polar parts of the phospholipid,
+i.e., charges on all atoms except on alkyl tails, 
+which do not encounter direct contacts with ions.
 

### 2018-10-14 Manuscript/manuscript.tex
Improvments in the end of intro

@@
 Classical molecular dynamics (MD) simulations give
-access to all the details, but the current force fields are not
+access to all the details, but the current non-polarizable force fields are not
 sufficiently accurate to interpret the interactions between anionic
 lipid headgroups and other lipids or ions \cite{NMRlipidsIV}.
-Implicit inclusion of electronic polarizability using
-electronic continuum correction (ECC) to lipid models
-substantially improved the interactions between cations and
-zwitterionic lipids \cite{melcr18}.
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%From the perspective of phospholipid bilayers,
+There are force field parameters for neutral lipids PC and PE 
+with explicit polarizability using the Classical Durde model. \cite{chowdhary13, chowdhary17}
+While such a model shows a potential improvement over non-polarizable force fields from many perspectives -- 
+for example accounting for polarizability has a significant impact on the dipole potential of a membrane \cite{harder2009} -- 
+the non-polarizable version of the same force field 
+performes comparably well at a fraction of the computational cost 
+of the models with explicit polarization. \citep{lucas12,chowdhary13}
 
@@
 %is observed with a force field using charge equilibration to model polarizability. \cite{lucas2012} 
-
-From the perspective of phospholipid bilayers,
-a force field with an explicit polarizability 
-using the Classical Durde model 
-is gradually being developed, however, 
-only parameters for PC and PE are available at this moment. \cite{chowdhary13, chowdhary17}
-While such a model shows a potential improvement from many perspectives -- 
-for example accounting for polarizability has a significant impact on the dipole potential of a membrane \cite{harder2009} -- 
-the non-polarizable version of the same force field 
-performes comparably well at a fraction of the computational cost 
-of the models with explicit polarization. \citep{lucas12,chowdhary13}
-
-%To our knowledge, there is no study focusing on
-%charged lipids using polarizable force fields. 
-%nor about their interactions with cations and anions. 
-%From this perspective,
-%this work is the first pioneering study providing atomistic resolution insight 
-%into the interactions of POPS with monovalent and even divalent cations.
-
-%Even though the software packages are more and more optimized for running simulations with explicit polarization,
-%its application still seems to be rather far from ready for common production use. \cite{jiang2011, lopes2009}
-
-Here, we extend the approach
-to anionic PS lipids to more accurately describe their interactions with ions
-and PC headgroups in mixed lipid bilayers. The results pave
-the way for more realistic MD simulations of negatively charged
-lipid bilayers to be used in wide range of applications.
+At the same time, 
+the implicit inclusion of electronic polarizability to lipid models
+via charge re-scaling using electronic continuum correction (ECC) 
+substantially improved the interactions between cations and zwitterionic lipids \cite{melcr18}.
+
+Here, we extend the approach to anionic PS lipids.
+The newly developed model termed ECC-POPS 
+not only describes the interactions of PS with ions more accurately, 
+but also improves the structure of the PS head group 
+and its interactions with PC in mixed lipid bilayers. 
+The results pave the way to more realistic MD simulations 
+of both neutral and negatively charged lipid bilayers 
+for a wide range of applications.
 

### 2018-10-14 Manuscript/manuscript.tex
Added/improved a list of MD simulations.

@@
 \centering
-\caption{  List of MD simulations with
-           numbers of individual molecules for each simulation
-           at temperature $T$. 
-         }
-         \todo{SAMULI: Temperature column is not necessary because it is always the same.}
-         \todo{SAMULI: Columns reporting the additional ion concentrations would be useful.}
-         \todo{SAMULI: Order of the PC:PS (5:1)/Na+ systems in little bit confusing.}
+\caption{  List of MD simulations reporting their respective
+           added buffer concentrations,
+           amounts of individual molecules,
+           and the deployed models.
+         }
 \label{tbl:sim-list}
-\begin{tabular}{l c c r r r r r }
-  lipids/counter-ions &  \ce{K^+}  &  \ce{Na^+} & \ce{Ca^{2+}} & \ce{Cl^-} & \ce{H2O} &  PC:PS total  &  $T$~(K)  \\
-  \hline
-PC:PS (1:1)/Na+  &  0  &  64  &  0  &  0  &  5253  &  64:64  &  298 \\ 
-PC:PS (1:1)/K+  &  64  &  0  &  0  &  0  &  5253  &  64:64  &  298 \\ 
-PC:PS (10:1)/Na+  &  0  &  6  &  0  &  0  &  3600  &  60:6  &  298 \\ 
-PC:PS (10:1)/K+  &  6  &  0  &  0  &  0  &  3600  &  60:6  &  298 \\ 
-PC:PS (4:1)/Na+  &  0  &  12  &  0  &  0  &  3600  &  48:12  &  298 \\ 
-PC:PS (4:1)/K+  &  12  &  0  &  0  &  0  &  3600  &  48:12  &  298 \\ 
-PC:PS (5:1)/Na+  &  0  &  12  &  13  &  26  &  3561  &  60:12  &  298 \\ 
-PC:PS (5:1)/Na+  &  0  &  12  &  26  &  52  &  3522  &  60:12  &  298 \\ 
-PC:PS (5:1)/Na+  &  0  &  12  &  0  &  0  &  3600  &  60:12  &  298 \\ 
-PC:PS (5:1)/K+  &  12  &  0  &  0  &  0  &  3600  &  60:12  &  298 \\ 
-PC:PS (5:1)/Na+  &  0  &  12  &  39  &  78  &  3483  &  60:12  &  298 \\ 
-PC:PS (5:1)/Na+  &  39  &  12  &  0  &  39  &  3483  &  60:12  &  298 \\ 
-PC:PS (5:1)/Na+  &  0  &  51  &  0  &  39  &  3483  &  60:12  &  298 \\ 
-  \hline
-PS/K+  &  72  &  0  &  0  &  0  &  3600  &  0:72  &  298 \\ 
-PS/Na+  &  0  &  72  &  0  &  0  &  3600  &  0:72  &  298 \\ 
+\begin{tabular}{l | r r r r | r r | c c }
+                          &  \multicolumn{4}{c}{added buffer concentration / mM}    & \multicolumn{2}{c}{no. molecules} &  \multicolumn{2}{c}{simulated with model}  \\
+PC:PS (ratio)/counterions &  \ce{K^+}  &  \ce{Na^+} & \ce{Ca^{2+}} & \ce{Cl^-}      & \ce{H2O} &  PC:PS                 &  ECC-lipids  &  Lipid17    \\
+  \hline
+PC:PS (0:1)/K+   &      0  &      0  &      0  &      0  &  3600  &  0:72  &  X  &  X  \\ 
+PC:PS (0:1)/Na+  &      0  &      0  &      0  &      0  &  3600  &  0:72  &  X  &  X  \\ 
+  \hline
+PC:PS (10:1)/K+  &      0  &      0  &      0  &      0  &  3600  &  60:6  &  X  &  -  \\ 
+PC:PS (10:1)/Na+ &      0  &      0  &      0  &      0  &  3600  &  60:6  &  X  &  -  \\ 
+  \hline
+PC:PS (1:1)/K+   &      0  &      0  &      0  &      0  &  5253  &  64:64  &  X  &  X  \\ 
+PC:PS (1:1)/Na+  &      0  &      0  &      0  &      0  &  5253  &  64:64  &  X  &  X  \\ 
+  \hline
+PC:PS (4:1)/K+   &      0  &      0  &      0  &      0  &  3600  &  48:12  &  X  &  -  \\ 
+PC:PS (4:1)/Na+  &      0  &      0  &      0  &      0  &  3600  &  48:12  &  X  &  -  \\ 
+  \hline
+PC:PS (5:1)/K+   &      0  &      0  &      0  &      0  &  3600  &  60:12  &  X  &  X  \\ 
+PC:PS (5:1)/Na+  &      0  &      0  &      0  &      0  &  3600  &  60:12  &  X  &  X  \\ 
+PC:PS (5:1)/Na+  &      0  &      0  &     78  &     78  &  3561  &  60:12  &  X  &  X  \\ 
+PC:PS (5:1)/Na+  &      0  &      0  &    125  &    125  &  3561  &  60:12  &  X  &  X  \\ 
+PC:PS (5:1)/Na+  &      0  &      0  &    202  &    202  &  3561  &  60:12  &  X  &  X  \\ 
+PC:PS (5:1)/Na+  &      0  &      0  &    409  &    409  &  3522  &  60:12  &  X  &  X  \\ 
+PC:PS (5:1)/Na+  &      0  &      0  &    621  &    621  &  3483  &  60:12  &  X  &  X  \\ 
+PC:PS (5:1)/Na+  &      0  &    621  &      0  &    621  &  3483  &  60:12  &  X  &  X  \\ 
+PC:PS (5:1)/Na+  &      0  &   1510  &      0  &   1510  &  3377  &  60:12  &  X  &  X  \\ 
+PC:PS (5:1)/Na+  &      0  &   3002  &      0  &   3002  &  3213  &  60:12  &  X  &  X  \\ 
+PC:PS (5:1)/Na+  &    621  &      0  &      0  &    621  &  3483  &  60:12  &  X  &  X  \\ 
+PC:PS (5:1)/Na+  &   1510  &      0  &      0  &   1510  &  3377  &  60:12  &  X  &  X  \\ 
+PC:PS (5:1)/Na+  &   3002  &      0  &      0  &   3002  &  3213  &  60:12  &  X  &  X  \\ 
   \hline

### 2018-10-14 Manuscript/manuscript.tex
Removed TODOs relating to figures.

@@
   \caption{\label{fig:delta_ordPar_NaCl_PC-PS_mix} 
-    Changes of the head group order parameters of a POPC:POPS (5:1) bilayer as a function of PS content
+    The head group order parameters of a POPC:POPS (5:1) bilayer as a function of PS content
     with \ce{Na^+} and \ce{K^+} counterions from simulations with ECC-lipids and Lipid17 \cite{lipid17-future} 
@@
     where the response is closer to the experiment (which uses only \ce{Na^+} counterions). 
-    Experimental points for POPS inherently contain larger systematic error, 
-    as they come from independent experiments.
-  }
-  \todo{SAMULI: Caption says Åqvist, but I guess that Dang is used also here?} \\
-  \todo{SAMULI: Do you have a simulation of Lipid14 pure POPC at 298K?
-    This is not in the table. This simulation would be useful also for NMRlipids IV.
-  } \\
-  \todo{SAMULI: Error bars for lipid14 missing.}
+  }
 \end{figure*}
@@
   } 
-  \todo{SAMULI: The forking of PS alpha carbon should not be put to zero with zero salt concentrations
-    in this or in other figures. This is confusing expecially for the results with CaCl$_2$ where the forking
-    seems to increase with CaCl$_2$, which is not correct. In NMRlipids IV, I have now put the lower order parameter
-    to zero and kept the forking correct with all concentrations of ions.} \\
-  \todo{SAMULI: Ion concentrations should be updated to the concentrations in water before solvating the lipids.} \\
-  \todo{SAMULI: Error bars should be recalculated with the fixed code} \\
-  \todo{SAMULI: We should have at least one more points with higher concentration (maybe close to 2M), or maybe two more points.} \\
   \end{figure}
@@
   } 
-  \todo{SAMULI: The forking of PS alpha carbon should not be put to zero with zero salt concentrations
-    in this or in other figures. This is confusing expecially for the results with CaCl$_2$ where the forking
-    seems to increase with CaCl$_2$, which is not correct. In NMRlipids IV, I have now put the lower order parameter
-    to zero and kept the forking correct with all concentrations of ions.} \\              
-  \todo{SAMULI: Ion concentrations should be updated to the concentrations in water before solvating the lipids.} \\
-  \todo{SAMULI: Error bars should be recalculated with the fixed code} \\
-  \todo{SAMULI: We should have at least one more points with higher concentration (maybe close to 2M), or maybe two more points.} \\
 \end{figure}
@@
   } 
-  \todo{SAMULI: The forking of PS alpha carbon should not be put to zero with zero salt concentrations
-    in this or in other figures. This is confusing expecially for the results with CaCl$_2$ where the forking
-    seems to increase with CaCl$_2$, which is not correct. In NMRlipids IV, I have now put the lower order parameter
-    to zero and kept the forking correct with all concentrations of ions.} \\
-  \todo{SAMULI: Ion concentrations should be updated to the concentrations in water before solvating the lipids.} \\
-  \todo{SAMULI: Error bars should be recalculated with the fixed code} \\
-  \todo{SAMULI: Maybe we should have at least one more points with smaller concentration
-    (maybe close to 100mM in solution before solvating lipids).} \\
 \end{figure}
@@
   }
-  \todo{SAMULI: Ion concentrations should be updated to the concentrations in water before solvating the lipids.} \\
   \todo{SAMULI: We should compare to the density profiles from Lipid17 simulations. These are also needed for NMRlipids IV.}

### 2018-10-14 Manuscript/manuscript.tex
Formatting figures...

@@
 \setlength{\figwidthfull}{14 cm} 
+\newlength{\figheightsmall}
+\setlength{\figheightsmall}{8 cm} 
+\newlength{\figheight}
+\setlength{\figheight}{12 cm}
 
@@
 
-\begin{figure*}[!htb] 
-  \centering 
-  \includegraphics[width=\figwidthsmall]{../Fig/order_parameters_changes_A-B_PC-PS_mix_POPC_nacl.pdf} 
-  \includegraphics[width=\figwidthsmall]{../Fig/order_parameters_changes_A-B_PC-PS_mix_POPS_nacl.pdf} 
-  \includegraphics[width=\figwidthsmall]{../Fig/order_parameters_changes_A-B_PC-PS_mix_POPC_kcl.pdf} 
-  \includegraphics[width=\figwidthsmall]{../Fig/order_parameters_changes_A-B_PC-PS_mix_POPS_kcl.pdf} 
+\begin{figure*}[!tbp] 
+  \centering 
+  \includegraphics[height=\figheightsmall]{../Fig/order_parameters_changes_A-B_PC-PS_mix_POPC_nacl.pdf} 
+  \includegraphics[height=\figheightsmall]{../Fig/order_parameters_changes_A-B_PC-PS_mix_POPS_nacl.pdf} 
+  \includegraphics[height=\figheightsmall]{../Fig/order_parameters_changes_A-B_PC-PS_mix_POPC_kcl.pdf} 
+  \includegraphics[height=\figheightsmall]{../Fig/order_parameters_changes_A-B_PC-PS_mix_POPS_kcl.pdf} 
   \caption{\label{fig:delta_ordPar_NaCl_PC-PS_mix} 
@@
   \centering 
-  \includegraphics[width=\figwidthsmall]{../img/ecc_pops/order_parameters_changes_ecc-lip_L14_A-B-PN-COO_POPC_nacl.pdf} 
-  \includegraphics[width=\figwidthsmall]{../img/ecc_pops/order_parameters_changes_ecc-lip_L14_A-B-PN-COO_POPS_nacl.pdf} 
+  \includegraphics[height=\figheight]{../img/ecc_pops/order_parameters_changes_ecc-lip_L14_A-B-PN-COO_POPC_nacl.pdf} 
+  \includegraphics[height=\figheight]{../img/ecc_pops/order_parameters_changes_ecc-lip_L14_A-B-PN-COO_POPS_nacl.pdf} 
   \caption{\label{fig:delta_ordPar_NaCl_PCPS} 
@@
   \centering 
-  \includegraphics[width=\figwidthsmall]{../img/ecc_pops/order_parameters_changes_ecc-lip_L14_A-B-PN-COO_POPC_kcl.pdf} 
-  \includegraphics[width=\figwidthsmall]{../img/ecc_pops/order_parameters_changes_ecc-lip_L14_A-B-PN-COO_POPS_kcl.pdf} 
+  \includegraphics[height=\figheight]{../img/ecc_pops/order_parameters_changes_ecc-lip_L14_A-B-PN-COO_POPC_kcl.pdf} 
+  \includegraphics[height=\figheight]{../img/ecc_pops/order_parameters_changes_ecc-lip_L14_A-B-PN-COO_POPS_kcl.pdf} 
   \caption{\label{fig:delta_ordPar_KCl_PCPS} 
@@
   \centering 
-  \includegraphics[width=\figwidthsmall]{../img/ecc_pops/order_parameters_changes_ecc-lip_L14_A-B-PN-COO_POPC_cacl.pdf} 
-  \includegraphics[width=\figwidthsmall]{../img/ecc_pops/order_parameters_changes_ecc-lip_L14_A-B-PN-COO_POPS_cacl.pdf} 
+  \includegraphics[height=\figheight]{../img/ecc_pops/order_parameters_changes_ecc-lip_L14_A-B-PN-COO_POPC_cacl.pdf} 
+  \includegraphics[height=\figheight]{../img/ecc_pops/order_parameters_changes_ecc-lip_L14_A-B-PN-COO_POPS_cacl.pdf} 
   \caption{\label{fig:delta_ordPar_CaCl_PCPS} 

### 2018-10-25 Manuscript/manuscript.tex
Add some comments in the manuscript

@@
 
+\todo{SAMULI: Controversies in previous simulations (https://doi.org/10.1016/j.bpj.2018.08.044 and
+  Melcrova et al. Scientific Reports | 6:38035 | DOI: 10.1038/srep38035) should be mentioned also
+in the intro.}
+
 \section{Methods}
 
-\subsection{Electronic continuum correction for lipid bilayers}\label{section:ecc}
+\subsection{Electronic continuum correction for PS lipids}\label{section:ecc}
 
@@
 an amount of water molecules comparable to the hydration levels in experiments, 
-\todo{Check this claim.}
+\todo{Check this claim. SAMULI: Check discussin in the NMRlipids IV manuscript}
 and concentrations of ions (\ce{K^+}, \ce{Na^+}, \ce{Ca^{2+}} and \ce{Cl^{-}})
@@
 
-We used the SPC/E~\cite{Berendsen1987} water model 
+We used the SPC/E~\cite{Berendsen1987} water model \todo{SAMULI: The choice of water model should be justified.}
 and ECC-ions \cite{martinek17, kohagen16, Pluharova2014} 
@@
 Force field parameters for POPC were taken from Ref.~\citenum{melcr18}. 
-The ECC-POPS model as derived again from the Lipid14 force field~\cite{dickson14, lipid17-future}, as was described above.
+The ECC-POPS model as derived again from the Lipid14
+\todo{SAMULI: References to Lipid14/Lipid17 models should be consistent here and elsewhere.
+We could also consider using term Amber lipd force field.}
+force field~\cite{dickson14, lipid17-future}, as was described above.
 For a direct comparison, 
-we also performed simulations with the Lipid14 model~\cite{dickson14},
+we also performed simulations with the Lipid14
+\todo{SAMULI: Also Lipid17 is used. The origin of Gromacs suitable files (Zenodo, acpype, etc.) should be described.}
+model~\cite{dickson14},
 the TIP3P water model~\cite{jorgensen83}, 
@@
          }
+         \todo{SAMULI: All the tails are PO, right? This could be mentioned in the caption.}
+
 \label{tbl:sim-list}
@@
 
+\todo{SAMULI: For this discussion we need the counterion density profiles from pure POPS
+  simulation with Lipid17 and Dang counterions. This could be added to the figure \ref{fig:POPS-counterions-dens} and
+the figure could be moved here.}
+
 We compared X-ray scattering form factors and NMR order parameters of bilayers
@@
   }
+  \todo{SAMULI: Use the same y-axis scale in the bottom figures showing angles for POPC and POPS to ease the comparison.}
   \end{figure}
@@
   }
+  \todo{SAMULI: Use the same y-axis scale in the bottom figures showing angles for POPC and POPS to ease the comparison.}
 \end{figure}
@@
 
-\todo{SAMULI: The qualitative iprovement of PS headgroup order parameter response to bound Ca2+
-  should be discussed. We should, however, point out that the response PS alpha carbon to bound
-  Ca2+ is not excatly reproduced even by the ECC-POPS model.
-}
+\todo{SAMULI: The new results show that we cannot actually reproduce the rapid changes with low CaCl2
+  concentrations in POPS order parameters. This should be somehow discussed.}
 
@@
 
+\todo{SAMULI: I have putted here two preliminary sentences}
+Applying ECC to POPS lipids improves cation binding affinity to membranes
+mixed with POPC and also the response of PS headgroup to the bound cations.
+However, further improvement of the parameters is needed to fully reproduce the
+details of calcium binding to PS headgroups.
+
 \listoftodos

### 2018-11-01 Manuscript/manuscript.tex
Added density plot with only counterions for both PC:PS mix and pure POPS and ECC-lipids resp. Lipid17

@@
   \centering 
-  \includegraphics[width=\figwidth]{../img/density_profile_K-Na_counterions_compar_pure-ECC-POPS.pdf}
+  \includegraphics[height=\figheightsmall]{../img/ecc_pops/density_profiles_na-k-counterions_wat_phos_compar_purePOPS.pdf}
+  \includegraphics[height=\figheightsmall]{../img/ecc_pops/density_profiles_na-k-counterions_wat_phos_compar_5PC-1PS.pdf} \\
+  \includegraphics[height=\figheightsmall]{../img/ecc_pops/l17/density_profiles_na-k-counterions_wat_phos_compar_purePOPS_lipid17.pdf}
+  \includegraphics[height=\figheightsmall]{../img/ecc_pops/l17/density_profiles_na-k-counterions_wat_phos_compar_5PC-1PS_lipid17.pdf}
   \caption{\label{fig:POPS-counterions-dens}
     Number density profiles of \ce{K^{+}}, \ce{Na^{+}} counterions along the membrane normal axis 
-    for the negatively charged POPS bilayers using ECC-lipids and ECC-ions.  }
-\todo{Joe: TO BE UPDATED + results from PC:PS mixutres to be added.}
+    for the negatively charged bilayers composed of only POPS (left) or POPC:POPS (5:1, right) 
+    using ECC-lipids and ECC-ions (top) and Lipid17 (bottom).  }
 \end{figure}

### 2018-11-01 Manuscript/manuscript.tex
Update methods text

@@
 
-The polarization of electrons in ECC-lipids uses the implicit method termed electronic continuum correction (ECC). \cite{leontyev09, Pluharova2014, martinek17, melcr18} 
-In this method, a system of polarizable particles is represented with an equivalent system of particles with fixed charges. 
-This is practically achieved rescaling the charges with a pre-determined scaling factor, $f_q$,
-which is derived from the high-frequency dielectric constant of electrons, $\epsilon_{el}$, as
+The lack of electronic polarizability has been a concern since the
+early days of classical MD simulations of lipids \cite{jonsson86,egberts94, berendsen1996},
+where the atomic charges were scaled by a factor of 1/2 to effectively
+include the electrostatic screening from the missing electronic polarizability.
+The electronic continuum correction (ECC) is a more sophisticated version of the charge scaling.
+In ECC, the charges are rescaled with a pre-determined scaling factor, $f_q$,
 \begin{equation}
-f_q = \frac{1}{\sqrt{\epsilon_{el}}} .
+  f_q = \frac{1}{\sqrt{\epsilon_{el}}} ,
 \end{equation}
-The value of $\epsilon_{el}$ for water solutions is $\epsilon_{el} = 1.78$ \cite{leontyev09},
-hence, the scaling factor for aquaeous environments yields $f_q \approx 0.75$.
-
-It is important to note that the value of the high frequency dielectric constant  
-is around 2 for almost any biologically relevant environment \citep{leontyev11}. 
-This means that even interfaces like biological membranes do not contain discontinuities of the electronic continuum. 
-The dielectric discontinuity in a lipid bilayer thus arises only 
-from the orientational polarization of the molecules, which is accounted for explicitly in standard MD simulations.  
-Therefore, the same correction for the electronic polarizability can be  
-applied throughout the lipid bilayer/aqueous solution interface.
-
-In the case of POPS, which is a monovalently negatively charged lipid, 
-we used exactly the pre-determined scaling factor 
-to represent the effects of electronic polarization
-yielding a total charge of $-0.75$ in the ECC formalism. \cite{martinek17, leontyev09, melcr18}
-Similarly to the ECC-POPC model, we changed the charges only of the polar parts of the phospholipid,
-i.e., charges on all atoms except on alkyl tails, 
-which do not encounter direct contacts with ions.
-
-Note that we used a different scaling factor $f_q = 0.8$ 
-to for the ECC-POPC model. \cite{melcr18}
-Compared to POPS, POPC is a zwitterionic neutral molecule with a zero total charge; 
-scaling it by any factor does not affect it. 
-Partial charges of the atoms in POPC are not physical observables, 
-and vary greatly between different models. \cite{botan15, catte16, nmrlipids_proj4, Pluhackova2016}
-The choice of a larger scaling factor in ECC-POPC represents the fact that 
-the set of partial charges in the initial model for ECC-POPC, the Lipid14 model \cite{dickson14}, 
-was determined in the beginning as less polar compared to other models.
+where $\epsilon_{el}$ is the high-frequency dielectric constant of electrons.
+The value of $\epsilon_{el} = 1.78$ for water solutions gives the scaling
+factor of $f_q \approx 0.75$ for aquaeous environments \cite{leontyev09}.
+This has been recently applied to improve classical models for ions and
+other biomolecules \cite{Pluharova2014, martinek17, melcr18}
+\todo{Pavel can probably update this}.
+
+Recently we applied the ECC also to the Amber based lipid14 force field \cite{dickson14} to
+improve the cation binding affinity to zwitterionic POPC lipid bilayers \cite{melcr18}.
+In contrast to the total charge of ions, the partial charges in force fields of zwitterionic
+depend on the methods used to derive the force field which may already partially include the
+electronic polarizability effects. Therefore, the optimal scaling factor for charges in
+zwitterionic lipids may be larger than the theoretically derived $f_q \approx 0.75$.
+In the recent work, we found optimal results for cation binding affinity and
+bilayer structural properties when the partial charges and Lennard-Jones $\sigma$
+parameters of headgroup, glycerol backbone, and carbonyl regions of the lipid14 POPC model
+were scaled with the factors of $f_q = 0.8$ and $f_\sigma = 0.89$, respectively \cite{melcr18}.
+
+%Compared to POPS, POPC is a zwitterionic neutral molecule with a zero total charge; 
+%scaling it by any factor does not affect it. 
+%Partial charges of the atoms in POPC are not physical observables, 
+%and vary greatly between different models. \cite{botan15, catte16, nmrlipids_proj4, Pluhackova2016}
+%The choice of a larger scaling factor in ECC-POPC represents the fact that 
+%the set of partial charges in the initial model for ECC-POPC, the Lipid14 model \cite{dickson14}, 
+%was determined in the beginning as less polar compared to other models.
+
+Here, we apply the ECC to the Amber based lipid17 parameters \cite{could18} of POPS lipid with
+a total monovalent negative charge. In this case we apply the scaling factor of $f_q = 0.75$ 
+also to the lipid because it carries a physical total charge similar to the ions in solution and
+the lipid charges has to neutralize the counterion charges in simulations.
+For the Lennard-Jones $\sigma$ parameters, we used the same scaling factor of $f_\sigma = 0.89$
+as in the ECC-POPC model \cite{melcr18}.
+%we used exactly the pre-determined scaling factor 
+%to represent the effects of electronic polarization
+%yielding a total charge of $-0.75$ in the ECC formalism. \cite{martinek17, leontyev09, melcr18}
+Similarly to the ECC-POPC model \cite{melcr18}, we applied the ECC only to the
+headgroup, glycerol backbone, and carbonyl regions.
+%polar parts of the phospholipid,
+%i.e., charges on all atoms except on alkyl tails, 
+%which do not encounter direct contacts with ions.
+
+%It is important to note that the value of the high frequency dielectric constant  
+%is around 2 for almost any biologically relevant environment \citep{leontyev11}. 
+%This means that even interfaces like biological membranes do not contain discontinuities of the electronic continuum. 
+%The dielectric discontinuity in a lipid bilayer thus arises only 
+%from the orientational polarization of the molecules, which is accounted for explicitly in standard MD simulations.  
+%Therefore, the same correction for the electronic polarizability can be  
+%applied throughout the lipid bilayer/aqueous solution interface.
 

### 2018-11-01 Manuscript/manuscript.tex
Update methods text

@@
   \centering 
-  \includegraphics[height=\figheightsmall]{../img/ecc_pops/density_profiles_na-k-counterions_wat_phos_compar_purePOPS.pdf}
-  \includegraphics[height=\figheightsmall]{../img/ecc_pops/density_profiles_na-k-counterions_wat_phos_compar_5PC-1PS.pdf} \\
-  \includegraphics[height=\figheightsmall]{../img/ecc_pops/l17/density_profiles_na-k-counterions_wat_phos_compar_purePOPS_lipid17.pdf}
-  \includegraphics[height=\figheightsmall]{../img/ecc_pops/l17/density_profiles_na-k-counterions_wat_phos_compar_5PC-1PS_lipid17.pdf}
+  \includegraphics[width=\figwidth]{../img/density_profile_K-Na_counterions_compar_pure-ECC-POPS.pdf}
   \caption{\label{fig:POPS-counterions-dens}
     Number density profiles of \ce{K^{+}}, \ce{Na^{+}} counterions along the membrane normal axis 
-    for the negatively charged bilayers composed of only POPS (left) or POPC:POPS (5:1, right) 
-    using ECC-lipids and ECC-ions (top) and Lipid17 (bottom).  }
+    for the negatively charged POPS bilayers using ECC-lipids and ECC-ions.  }
+\todo{Joe: TO BE UPDATED + results from PC:PS mixutres to be added.}
 \end{figure}

### 2018-11-01 Manuscript/manuscript.tex
Update the methods text

--- 2018-11-01 Manuscript/manuscript.tex
@@
 
-\todo{SAMULI: Short discussion about mixtures should be added by citing the NMRlipids IV.}
-The structures of lipid bilayers in simulations without ions were validated against NMR 
-by calculating the order parameters for the C-H bonds and against \mbox{X-ray} scattering experiments by evaluating the scattering form factors
-as was done in Refs.~\citenum{catte16, melcr18, nmrlipids_proj4}.
+To validate the structures of lipid bilayers, the scattering form factors and C-H bond order parameters
+were calculated and compared against the experimental data as described previously~\cite{botan15,ollila16,melcr18,nmrlipids_proj4}.
+%in simulations without ions were validated against NMR 
+%by calculating the order parameters for the C-H bonds and against \mbox{X-ray} scattering experiments by evaluating 
+%as was done in Refs.~\citenum{catte16, melcr18, nmrlipids_proj4}.
 %NMR order parameters validate the structures sampled by the individual lipid molecules with atomic resolution. 
-The simulated order parameters were calculated for all C-H bonds in lipid molecules, 
-while all order paremeters in the head group and only some in the \ce{C-H} groups in tails were resolved experimentally
-(see next section and Fig.~\ref{simVSexpNOions_POPS}).
-
-Ion binding to lipid bilayers can be experimentally quantified
-by measuring the changes of the head group order parameters in the lipids using solid state NMR.
-In the case of the head group order parameters $S^\alpha$ and $S^\beta$ in phosphatidylcholine
-(see Fig.~\ref{simVSexpNOions_POPS} for the definition of the order parameters)
-such changes are known under the term ''electrometer concept'' \citep{seelig87,catte16, ollila16, melcr18, nmrlipids_proj4}. 
-It is experimentally observed that the \ce{C-H} bond
-order parameters of $S^\alpha$ and $S^\beta$ carbons in a PC lipid head group
-are proportional to the amount of charge, positive or negative, bound per lipid~\citep{seelig87, roux90}.
-The change of the order parameters measured with varying concentrations of aqueous ions 
-can be then related to the amount of bound ions.
+%The simulated order parameters were calculated for all C-H bonds in lipid molecules, 
+%while all order paremeters in the head group and only some in the \ce{C-H} groups in tails were resolved experimentally
+%(see next section and Fig.~\ref{simVSexpNOions_POPS}). 
+The order parameters for each C-H bond of individual lipids in the trajectory were first averaged
+over the time. Then the average over lipids was taken and the standard error of the mean of this
+average was used as an error bar \cite{botan15,,ollila16,nmrlipids_proj4}.
+
+Cation binding affinity to phosphatidylcholine (PC) lipid bilayers can be evaluated using 
+the order parameters of $\alpha$ and $\beta$ carbons in the lipid headgroup 
+(see Fig.~\ref{simVSexpNOions_POPS} for the labeling
+\todo{SAMULI: Currently the figure has only PS structure while we use PC also in this work.
+We should either add also PC structure or refer to somewhere else for the labeling.}) \cite{catte16}.
+According to the ''electrometer concept'' \citep{seelig87}, these order parameters decrease
+proportionally to the bound charge and can therefore be used to quantify the cation binding
+affinity to PC lipid bilayers. The situation is more complicated for lipid bilayers containing
+PS lipids because the ion free reference state does not exist due to the counterions and additional
+ions often lead to phase transitions \cite{nmrlipids_proj4,feigenson86,mattai89,roux91,roux90}
+Here, we overcome these complications by comparing simulations to the experimental data measured
+from POPC:POPS (5:1) mixtures with additional ions\cite{roux90}, as demonstrated in the
+recent work by the NMRlipids project \cite{nmrlipids_proj4}.
+
+%experimentally quantified
+%by measuring the changes of the head group order parameters in the lipids using solid state NMR.
+%In the case of 
+%such changes are known under the term 
+%It is experimentally observed that the \ce{C-H} bond
+%order parameters of $S^\alpha$ and $S^\beta$ carbons in a PC lipid head group
+%are proportional to the amount of charge, positive or negative, bound per lipid~\citep{seelig87, roux90}.
+%The change of the order parameters measured with varying concentrations of aqueous ions 
+%can be then related to the amount of bound ions.
 
@@
 
-MD simulations were performed using the GROMACS 2018 \cite{Abraham15} simulation package. 
-The simulated system consisted of lipids (POPC and POPS) forming a single lipid bilayer,
-an amount of water molecules comparable to the hydration levels in experiments, 
-\todo{Check this claim. SAMULI: Check discussin in the NMRlipids IV manuscript}
-and concentrations of ions (\ce{K^+}, \ce{Na^+}, \ce{Ca^{2+}} and \ce{Cl^{-}})
-in an orthorhombic simulation box with periodic boundary conditions. 
-The simulation parameters used in this work are summarized in Table~\ref{tbl:mdpar}. 
-A comprehensive list of simulations used in this work is in Table~\ref{tbl:sim-list}.
-
+MD simulations of POPC:POPS lipid bilayers with different molar ratios 
+and ion concentrations (\ce{K^+}, \ce{Na^+}, \ce{Ca^{2+}} and \ce{Cl^{-}})
+were performed in an orthorhombic simulation box with periodic boundary conditions
+using the GROMACS 2018 \cite{Abraham15} simulation package. 
+%The simulated system consisted of lipids (POPC and POPS) forming a single lipid bilayer,
+%an amount of water molecules comparable to the hydration levels in experiments, 
+%\todo{Check this claim. SAMULI: Check discussin in the NMRlipids IV manuscript}
+The simulated systems are listed in Table~\ref{tbl:sim-list} and the used parameters
+in Table~\ref{tbl:mdpar}. 
+\todo{SAMULI: All the simulations were ran ?? ns and the first ?? ns were omitted as
+an equilibration period. Or some other description of simulation times and equlibration.}
+For deposited trajectories and parameter files see Refs.~
+\todo{Joe: Upload trajectories and add references here.}
+
+In the reference simulations with the standard Amber force field,
+we used the Lipid14 parameters~\cite{dickson14} for POPC, the Lipid17 parameters for POPS \cite{gould18},
+the TIP3P water model~\cite{jorgensen83}, and ion models by Dang and coworkers~\cite{smith94,chang1999,dang2006}.
+The Lipid14 parameters in Gromacs format were downloaded from ?? \todo{SAMULI: Add this} and
+the Lipid17 parameters were \todo{SAMULI: Describe how lipid17 parameters in Gromacs were generated.}.
+The ion model by Dang and coworkers was used in these simulations because
+the default ion parameters in Amber exhibit clustering in solution with relatively
+low concentrations which complicates the evaluation of binding affinity against experiments \cite{nmrlipids_proj4}.
 We used the SPC/E~\cite{Berendsen1987} water model \todo{SAMULI: The choice of water model should be justified.}
@@
 force field~\cite{dickson14, lipid17-future}, as was described above.
-For a direct comparison, 
-we also performed simulations with the Lipid14
-\todo{SAMULI: Also Lipid17 is used. The origin of Gromacs suitable files (Zenodo, acpype, etc.) should be described.}
-model~\cite{dickson14},
-the TIP3P water model~\cite{jorgensen83}, 
-and ion models by Dang and coworkers~\cite{smith94,chang1999,dang2006}.
-\todo{The usage of Dang ions should be justified here.}
-
-For deposited trajectories and parameter files see Refs.~
-\todo{Joe: Upload trajectories and add references here.}
 

### 2018-11-14 Manuscript/manuscript.tex

@@
 
-Here, we apply the ECC to the Amber based lipid17 parameters \cite{could18} of POPS lipid with
+Here, we apply the ECC to the Amber based lipid17 parameters \cite{lipid17-future} of POPS lipid with
 a total monovalent negative charge. In this case we apply the scaling factor of $f_q = 0.75$ 
@@
 In the reference simulations with the standard Amber force field,
-we used the Lipid14 parameters~\cite{dickson14} for POPC, the Lipid17 parameters for POPS \cite{gould18},
+we used the Lipid14 parameters~\cite{dickson14} for POPC, the Lipid17 parameters for POPS \cite{lipid17-future},
 the TIP3P water model~\cite{jorgensen83}, and ion models by Dang and coworkers~\cite{smith94,chang1999,dang2006}.
@@
 the Lipid17 parameters were \todo{SAMULI: Describe how lipid17 parameters in Gromacs were generated.}.
-The ion model by Dang and coworkers was used in these simulations because
-the default ion parameters in Amber exhibit clustering in solution with relatively
-low concentrations which complicates the evaluation of binding affinity against experiments \cite{nmrlipids_proj4}.
-We used the SPC/E~\cite{Berendsen1987} water model \todo{SAMULI: The choice of water model should be justified.}
-and ECC-ions \cite{martinek17, kohagen16, Pluharova2014} 
-for the simulations with ECC-lipids.
-Force field parameters for POPC were taken from Ref.~\citenum{melcr18}. 
-The ECC-POPS model as derived again from the Lipid14
-\todo{SAMULI: References to Lipid14/Lipid17 models should be consistent here and elsewhere.
-We could also consider using term Amber lipd force field.}
-force field~\cite{dickson14, lipid17-future}, as was described above.
+The ion model by Dang and coworkers~\cite{smith94,chang1999,dang2006} was used with Amber lipids because
+analysis of simulations with the default Amber ion parameters is complicated by the artificial
+clustering of ions in solution \cite{nmrlipids_proj4}.
+In simulations with ECC-lipids,
+we used the ECC-POPC parameters~\cite{melcr18} (available at ??\todo{Add reference}) and ECC-POPS
+parameteres derived in this work (section \ref{section:ecc}) together with the SPC/E~\cite{Berendsen1987}
+water model and ECC-ions \cite{martinek17, kohagen16, Pluharova2014}.
+The SPC/E water model is used because it lower dielectric constant is consistent with the
+ECC concept \cite{leontyev11,leontyev14}.
 
@@
            and the deployed models.
-         }
+         }\label{tbl:sim-list}
          \todo{SAMULI: All the tails are PO, right? This could be mentioned in the caption.}
-
-\label{tbl:sim-list}
 \begin{tabular}{l | r r r r | r r | c c }

### 2018-11-15 Manuscript/manuscript.tex
Added plots with only counterions - they now combine Lipid17+ecclipids in one single plot

@@
   \centering 
-  \includegraphics[height=\figheightsmall]{../img/ecc_pops/density_profiles_na-k-counterions_wat_phos_compar_purePOPS.pdf}
-  \includegraphics[height=\figheightsmall]{../img/ecc_pops/density_profiles_na-k-counterions_wat_phos_compar_5PC-1PS.pdf} \\
-  \includegraphics[height=\figheightsmall]{../img/ecc_pops/l17/density_profiles_na-k-counterions_wat_phos_compar_purePOPS_lipid17.pdf}
-  \includegraphics[height=\figheightsmall]{../img/ecc_pops/l17/density_profiles_na-k-counterions_wat_phos_compar_5PC-1PS_lipid17.pdf}
+  \includegraphics[height=\figheight]{../img/ecc_pops/density_profiles_na-k-counterions_wat_phos_compar_purePOPS_ecclipids-lipid17.pdf}
+  \includegraphics[height=\figheight]{../img/ecc_pops/density_profiles_na-k-counterions_wat_phos_compar_5PC-1PS_ecclipids-lipid17.pdf}
   \caption{\label{fig:POPS-counterions-dens}
@@
     for the negatively charged bilayers composed of only POPS (left) or POPC:POPS (5:1, right) 
-    using ECC-lipids and ECC-ions (top) and Lipid17 (bottom).  }
+    using ECC-lipids and ECC-ions (solid lines) and Lipid17 (dashed lines).  
+    The density profiles of phosphate groups and water are divided by 4 and 100, respectively.  
+}
 \end{figure}

### 2018-11-15 Manuscript/manuscript.tex
Updated order parameter plots and added references to simulations into text.

@@
 over the time. Then the average over lipids was taken and the standard error of the mean of this
-average was used as an error bar \cite{botan15,,ollila16,nmrlipids_proj4}.
+average was used as an error bar \cite{botan15,ollila16,nmrlipids_proj4}.
 
@@
 an equilibration period. Or some other description of simulation times and equlibration.}
-For deposited trajectories and parameter files see Refs.~
-\todo{Joe: Upload trajectories and add references here.}
+For deposited trajectories and parameter files see 
+Refs.~\citenum{ecclipids_pcps_nacl_kcl_series, ecclipids_pcps_cacl2_series, ecclipids_pcps_mixtures_counterions, 
+               lipid17_nacl_kcl_series, lipid17_ff99_ions, lipid17_cacl_series}
 

### 2018-11-30 Manuscript/manuscript.tex

@@
 The electronic continuum correction (ECC) is a more sophisticated version of the charge scaling.
+
+The polarization of electrons in ECC-lipids uses the implicit method termed electronic continuum correction (ECC). \cite{leontyev09, Pluharova2014, martinek17, melcr18} 
+In this method, a system of polarizable particles is represented with an equivalent system of particles with fixed charges. 
+This is practically achieved rescaling the charges with a pre-determined scaling factor, $f_q$,
+which is derived from the high-frequency dielectric constant of electrons, $\epsilon_{el}$, as
+-OR-
 In ECC, the charges are rescaled with a pre-determined scaling factor, $f_q$,
@@
 were scaled with the factors of $f_q = 0.8$ and $f_\sigma = 0.89$, respectively \cite{melcr18}.
-
-%Compared to POPS, POPC is a zwitterionic neutral molecule with a zero total charge; 
-%scaling it by any factor does not affect it. 
-%Partial charges of the atoms in POPC are not physical observables, 
-%and vary greatly between different models. \cite{botan15, catte16, nmrlipids_proj4, Pluhackova2016}
-%The choice of a larger scaling factor in ECC-POPC represents the fact that 
-%the set of partial charges in the initial model for ECC-POPC, the Lipid14 model \cite{dickson14}, 
-%was determined in the beginning as less polar compared to other models.
 
@@
 
-%It is important to note that the value of the high frequency dielectric constant  
-%is around 2 for almost any biologically relevant environment \citep{leontyev11}. 
-%This means that even interfaces like biological membranes do not contain discontinuities of the electronic continuum. 
-%The dielectric discontinuity in a lipid bilayer thus arises only 
-%from the orientational polarization of the molecules, which is accounted for explicitly in standard MD simulations.  
-%Therefore, the same correction for the electronic polarizability can be  
-%applied throughout the lipid bilayer/aqueous solution interface.
-
 \subsection{Validation of membrane properties}
@@
 over the time. Then the average over lipids was taken and the standard error of the mean of this
-average was used as an error bar \cite{botan15,ollila16,nmrlipids_proj4}.
+average was used as an error bar \cite{botan15,,ollila16,nmrlipids_proj4}.
 
@@
 ECC concept \cite{leontyev11,leontyev14}.
+
+The simulation parameters used in this work are summarized in Table~\ref{tbl:mdpar}.
 

### 2018-12-05 Manuscript/manuscript.tex
Add details of NMR experiments written by Tiago

@@
 \subsection{Measurements of acyl chain order parameters}
-\todo{SAMULI: Short description about the order parameter experiments by Tiago.}
+A R-type Proton Detected Local Field (R-PDLF) experiment was performed to determine C--H bond order
+parameters, $S_{\rm{CH}}=1/2\langle3cos^2\theta-1\rangle$, where $\theta$ denotes the angle of the C--H bond
+with the bilayer normal and the angular brackets define a time average on a time scale of approximately 1 microsecond.
+The experiment was done using a Bruker Avance III 400 spectrometer operating at a 1H Larmor frequency of 400.03 MHz
+equipped with a standard 4 mm CP-MAS HXY probe. The set up of the R-PDLF experiment was the following
+(using the notation as in figures 1c and 2c of the original reference for the R-PDLF
+experiment [S. V. Dvinskikh, H. Zimmermann, A. Maliniak and D. Sandström, J. Mag. Res. 168, 194 (2004)]).
+The magic angle spinning (MAS) frequency used was 5.15 kHz. The recoupling pulses in the R18 blocks had
+therefore a length of 10.79 us which correspond to a nutation frequency of 46.35 kHz. Increments of 18x2 recoupling
+pulses where used giving a spectral width of 2.6 kHz in the indirect dimension and a total number of 32 points in
+the indirect dimension were recorded. These settings enable to record dipolar slices with a C--H bond order parameter
+resolution of +- 0.01. The C--H bond order parameter determined from a given dipolar
+splitting, $\Delta\nu$ (see e.g. Figure XX as SI) is equal to $\Delta\nu/(0.315\times21.5 kHz)$, where 0.315 is
+the scaling factor of the R18 recoupling sequence and 21.5 kHz is the maximum 1H-13C dipolar coupling for a C--H bond.
+The refocused-INEPT transfer [G. Morris and R. Freeman, J. Am. Chem. Soc. 101, 760 (1979); D. Burum, R. Ernst, J. Mag. Res. 39, 163 (1980)]
+was used for transferring the polarization from 1H nuclei to the covalently bond 13C nuclei with the delays $\tau_1$ = 1.94 ms and $\tau_2$ = 0.97 ms
+(set as multiples of the MAS rotation period). The pulses for the refocused-INEPT sequence had a nutation frequency equal to 63.45 kHz.
+For acquiring the spectra, a total number of 1024 transients were recorded for each point in the indirect dimension, using an acquisition time of 0.1 s under SPINAL64 1H decoupling [B.
+Fung, A.
+Khitrin, K.Ermolaev, J.
+Mag.
+Res.
+142, 97 (2000)] with a nutation frequency of 50 kHz, and with dwell time giving a spectral width of 200 ppm for the 13C chemical shift dimension.
+The set of free induction decays recorded were then processed as described e.g.
+in [T.
+M.
+Ferreira, B.
+Medronho, R.
+W.
+Martin, D.
+Topgaard, Phys.
+Chem.
+Chem.
+Phys.
+10, 6033 (2008)].
+For assigning the dipolar splittings in the crowded spectral region between 29 and 30 ppm both the assignments made for POPC in [ T.
+M.
+Ferreira, F.
+Coreta- Gomes, O.
+H.
+S.
+Ollila, M.
+J.
+Moreno, W.
+L.
+C.
+Vaz, D.
+Topgaard, Phys.
+Chem.
+Chem.
+Phys.
+15, 1976-1989 (2013)] and the POPS order parameters simulated and reported in this work were taken as a reference.
 

### 2018-12-17 Manuscript/manuscript.tex
Small improvements in Introduction

@@
 Lipids with negatively charged headgroups, like phophatidylserine
-play important role in biomembranes. Signaling domains interact with
-anionic lipid membranes due to electrostatic and specific interactions
-with negatively charged lipids \cite{lemmon08,leventis10}.
+play important role in biomembranes. 
+Signaling molecules in cells interact with lipid membranes 
+due to specific, often electrostatically driven, interactions 
+with anionic lipids \cite{lemmon08,leventis10}.
 Anionic lipids and their interactions with ions play also role in membrane
 fusion \cite{??}. \todo{Pavel can probably improve this and give a citation}
-These interactions depend on
-chemical details of lipid headgroups and binding of counter and co-ions
-to anionic membranes.
+These interactions depend both on
+the chemical details of lipid headgroups as well as 
+on the binding of counter and co-ions to anionic membranes.
 
 Various interaction schemes have been proposed for the negatively charged
-lipids with ions and other lipids. Some studies suggest that the
-binding constant to negatively charged lipids is similar to zwitterionic
+lipids with ions and other lipids. 
+Some studies suggest that the
+binding constant of ions to negatively charged lipids is similar to that of zwitterionic lipids,
 and the binding affinity is increased only due to the increased cation
@@
 
+Short paragraph describing previous MD-work on this topic (PS), i.e., \dots 
+\todo{SAMULI: Controversies in previous simulations (https://doi.org/10.1016/j.bpj.2018.08.044 and
+  Melcrova et al. Scientific Reports | 6:38035 | DOI: 10.1038/srep38035) should be mentioned also
+in the intro.}
+
 An empirical scaling factor, which reduces electrostatic forces between molecules, 
@@
 for a wide range of applications.
-
-\todo{SAMULI: Controversies in previous simulations (https://doi.org/10.1016/j.bpj.2018.08.044 and
-  Melcrova et al. Scientific Reports | 6:38035 | DOI: 10.1038/srep38035) should be mentioned also
-in the intro.}
 

### 2018-12-17 Manuscript/manuscript.tex
Small improvemnets in Methods: Electronic continuum correction for PS lipids (just this small section)

@@
 
+\todo{Joe: this paragraph requires editing.}
 The lack of electronic polarizability has been a concern since the
@@
 The polarization of electrons in ECC-lipids uses the implicit method termed electronic continuum correction (ECC). \cite{leontyev09, Pluharova2014, martinek17, melcr18} 
-In this method, a system of polarizable particles is represented with an equivalent system of particles with fixed charges. 
-This is practically achieved rescaling the charges with a pre-determined scaling factor, $f_q$,
+In this method, a system of polarizable particles is represented 
+with an equivalent system of particles with fixed charges,
+which account for the electronic polarizability \emph{implicitly}. 
+This is practically achieved by rescaling the charges with a pre-determined scaling factor, $f_q$,
 which is derived from the high-frequency dielectric constant of electrons, $\epsilon_{el}$, as
--OR-
-In ECC, the charges are rescaled with a pre-determined scaling factor, $f_q$,
 \begin{equation}
@@
 \end{equation}
-where $\epsilon_{el}$ is the high-frequency dielectric constant of electrons.
 The value of $\epsilon_{el} = 1.78$ for water solutions gives the scaling
@@
 other biomolecules \cite{Pluharova2014, martinek17, melcr18}
-\todo{Pavel can probably update this}.
+\todo{Pavel can probably update this (insulin?)}.
 
@@
 improve the cation binding affinity to zwitterionic POPC lipid bilayers \cite{melcr18}.
-In contrast to the total charge of ions, the partial charges in force fields of zwitterionic
-depend on the methods used to derive the force field which may already partially include the
+In contrast to the total charge of ions, 
+the partial charges in force fields of zwitterionic molecules
+depend on the methods used to derive the force field, 
+which may already partially include the
 electronic polarizability effects. Therefore, the optimal scaling factor for charges in
-zwitterionic lipids may be larger than the theoretically derived $f_q \approx 0.75$.
+zwitterionic lipids may vary from (be even larger) than the theoretically derived $f_q \approx 0.75$.
 In the recent work, we found optimal results for cation binding affinity and
-bilayer structural properties when the partial charges and Lennard-Jones $\sigma$
-parameters of headgroup, glycerol backbone, and carbonyl regions of the lipid14 POPC model
-were scaled with the factors of $f_q = 0.8$ and $f_\sigma = 0.89$, respectively \cite{melcr18}.
+bilayer structural properties when the partial charges and Lennard-Jones~$\sigma$
+parameters of all non-tail atoms of the lipid14 POPC model
+(i.e., headgroup, glycerol backbone, and carbonyl regions) 
+were scaled with the factors $f_q = 0.8$ and $f_\sigma = 0.89$  \cite{melcr18}.
 
 Here, we apply the ECC to the Amber based lipid17 parameters \cite{lipid17-future} of POPS lipid with
-a total monovalent negative charge. In this case we apply the scaling factor of $f_q = 0.75$ 
-also to the lipid because it carries a physical total charge similar to the ions in solution and
-the lipid charges has to neutralize the counterion charges in simulations.
-For the Lennard-Jones $\sigma$ parameters, we used the same scaling factor of $f_\sigma = 0.89$
+a total monovalent negative charge. 
+In this case we treat the POPS lipids like ions
+applying to it the scaling factor of $f_q = 0.75$,
+as it carries a physical total charge similar to the ions in solution, 
+and the lipid charges has to neutralize the counterion charges in simulations.
+For the Lennard-Jones~$\sigma$ parameters, we used the same scaling factor of $f_\sigma = 0.89$
 as in the ECC-POPC model \cite{melcr18}.
@@
 Similarly to the ECC-POPC model \cite{melcr18}, we applied the ECC only to the
-headgroup, glycerol backbone, and carbonyl regions.
+non-tail atoms, i.e., to atoms in the headgroup, glycerol backbone, and carbonyl regions. 
 %polar parts of the phospholipid,

### 2018-12-18 Manuscript/manuscript.tex
Small edits in Methods section (finished till its end)

@@
 over the time. Then the average over lipids was taken and the standard error of the mean of this
-average was used as an error bar \cite{botan15,ollila16,nmrlipids_proj4}.
+average was used as the error estimate of the mean \cite{botan15,ollila16,nmrlipids_proj4}.
 
@@
 the order parameters of $\alpha$ and $\beta$ carbons in the lipid headgroup 
-(see Fig.~\ref{simVSexpNOions_POPS} for the labeling
-\todo{SAMULI: Currently the figure has only PS structure while we use PC also in this work.
-We should either add also PC structure or refer to somewhere else for the labeling.}) \cite{catte16}.
+(see Fig.~\ref{fig:chemstruct_pc_ps} for the labeling) \cite{catte16}.
 According to the ''electrometer concept'' \citep{seelig87}, these order parameters decrease
 proportionally to the bound charge and can therefore be used to quantify the cation binding
-affinity to PC lipid bilayers. The situation is more complicated for lipid bilayers containing
-PS lipids because the ion free reference state does not exist due to the counterions and additional
-ions often lead to phase transitions \cite{nmrlipids_proj4,feigenson86,mattai89,roux91,roux90}
-Here, we overcome these complications by comparing simulations to the experimental data measured
-from POPC:POPS (5:1) mixtures with additional ions\cite{roux90}, as demonstrated in the
-recent work by the NMRlipids project \cite{nmrlipids_proj4}.
+affinity to lipid bilayers. The situation is more complicated for lipid bilayers containing
+PS lipids, for which there is no ion-free reference state due to the presence of counterions.
+Moreover, additional ions often lead to phase transitions in such bilayers \cite{nmrlipids_proj4,feigenson86,mattai89,roux91,roux90}
+Here, we overcome this complication by comparing simulations to the experimental data 
+measured from POPC:POPS (5:1) mixtures, 
+which are not as highly negatively charged, 
+and do not ondergo phase transitions with additional ions \cite{roux90} 
+as was also demonstrated in the recent work by the NMRlipids project \cite{nmrlipids_proj4}.
 
@@
 %\todo{Check this claim. SAMULI: Check discussin in the NMRlipids IV manuscript}
-The simulated systems are listed in Table~\ref{tbl:sim-list} and the used parameters
-in Table~\ref{tbl:mdpar}. 
-\todo{SAMULI: All the simulations were ran ?? ns and the first ?? ns were omitted as
-an equilibration period. Or some other description of simulation times and equlibration.}
+The complete set of simulated systems is listed in Table~\ref{tbl:sim-list},
+and the used simulation parameters are in Table~\ref{tbl:mdpar}. 
+All simulations were ran for a minimum of $1 \, \mu$s,
+and the first  $50 \,$ns were omitted as an equilibration period. 
 For deposited trajectories and parameter files see 
@@
 the TIP3P water model~\cite{jorgensen83}, and ion models by Dang and coworkers~\cite{smith94,chang1999,dang2006}.
-The Lipid14 parameters in Gromacs format were downloaded from ?? \todo{SAMULI: Add this} and
-the Lipid17 parameters were \todo{SAMULI: Describe how lipid17 parameters in Gromacs were generated.}.
+The Lipid14 parameters in Gromacs format were downloaded from Ref.~\citenum{lipid14files}. 
+The Lipid17 parameters were obtained from AmberTools18 \cite{amber18} 
+and they were converted to Gromacs format using acpype \cite{acpype}.  
 The ion model by Dang and coworkers~\cite{smith94,chang1999,dang2006} was used with Amber lipids because
@@
 clustering of ions in solution \cite{nmrlipids_proj4}.
-In simulations with ECC-lipids,
-we used the ECC-POPC parameters~\cite{melcr18} (available at ??\todo{Add reference}) and ECC-POPS
-parameteres derived in this work (section \ref{section:ecc}) together with the SPC/E~\cite{Berendsen1987}
-water model and ECC-ions \cite{martinek17, kohagen16, Pluharova2014}.
-The SPC/E water model is used because it lower dielectric constant is consistent with the
+In the simulations with ECC-lipids,
+we used the ECC-POPC parameters~\cite{melcr18} (available at \cite{ECC-POPC_nacl_cacl2_files}) and 
+ECC-POPS parameteres derived in this work (section \ref{section:ecc}) 
+together with the SPC/E~\cite{Berendsen1987} water model and 
+ECC-ions \cite{martinek17, kohagen16, Pluharova2014}.
+We used the SPC/E water model as its lower dielectric constant is consistent with the
 ECC concept \cite{leontyev11,leontyev14}.
-
-The simulation parameters used in this work are summarized in Table~\ref{tbl:mdpar}.
 
@@
 with a comparable accuracy to existing state-of-the-art lipid models~\citep{botan15, ollila16, Pluhackova2016}.
+
+\begin{figure}[tb!] 
+  \centering 
+  \includegraphics[width=\figwidth]{../Fig/lipids_chemfig_POPC_POPS.pdf} 
+  \caption{ \label{fig:chemstruct_pc_ps} 
+            The chemical structure of POPC and POPS 
+            including the labeling of the carbon segments
+            used in the measurements of order parameteres. 
+  }  
+\end{figure}
 

### 2018-12-19 Manuscript/manuscript.tex
Small fixes and moving aroung figures and tables in Methods and Results (beginning)

@@
 
+\begin{table*}[tbp]
+\centering
+\caption{  List of MD simulations reporting their respective
+           added buffer concentrations,
+           amounts of individual molecules,
+           and the deployed models.
+           The fatty acid chains for all lipids are plamitoyl (sn-1) and oleoyl (sn-2).
+         }\label{tbl:sim-list}
+\begin{tabular}{l | r r r r | r r | c c }
+                          &  \multicolumn{4}{c}{added buffer concentration / mM}    & \multicolumn{2}{c}{no. molecules} &  \multicolumn{2}{c}{simulated with model}  \\
+PC:PS (ratio)/counterions &  \ce{K^+}  &  \ce{Na^+} & \ce{Ca^{2+}} & \ce{Cl^-}      & \ce{H2O} &  PC:PS                 &  ECC-lipids  &  Lipid17    \\
+  \hline
+PC:PS (0:1)/K+   &      0  &      0  &      0  &      0  &  3600  &  0:72  &  X  &  X  \\ 
+PC:PS (0:1)/Na+  &      0  &      0  &      0  &      0  &  3600  &  0:72  &  X  &  X  \\ 
+  \hline
+PC:PS (10:1)/K+  &      0  &      0  &      0  &      0  &  3600  &  60:6  &  X  &  -  \\ 
+PC:PS (10:1)/Na+ &      0  &      0  &      0  &      0  &  3600  &  60:6  &  X  &  -  \\ 
+  \hline
+PC:PS (1:1)/K+   &      0  &      0  &      0  &      0  &  5253  &  64:64  &  X  &  X  \\ 
+PC:PS (1:1)/Na+  &      0  &      0  &      0  &      0  &  5253  &  64:64  &  X  &  X  \\ 
+  \hline
+PC:PS (4:1)/K+   &      0  &      0  &      0  &      0  &  3600  &  48:12  &  X  &  -  \\ 
+PC:PS (4:1)/Na+  &      0  &      0  &      0  &      0  &  3600  &  48:12  &  X  &  -  \\ 
+  \hline
+PC:PS (5:1)/K+   &      0  &      0  &      0  &      0  &  3600  &  60:12  &  X  &  X  \\ 
+PC:PS (5:1)/Na+  &      0  &      0  &      0  &      0  &  3600  &  60:12  &  X  &  X  \\ 
+PC:PS (5:1)/Na+  &      0  &      0  &     78  &     78  &  3561  &  60:12  &  X  &  X  \\ 
+PC:PS (5:1)/Na+  &      0  &      0  &    125  &    125  &  3561  &  60:12  &  X  &  X  \\ 
+PC:PS (5:1)/Na+  &      0  &      0  &    202  &    202  &  3561  &  60:12  &  X  &  X  \\ 
+PC:PS (5:1)/Na+  &      0  &      0  &    409  &    409  &  3522  &  60:12  &  X  &  X  \\ 
+PC:PS (5:1)/Na+  &      0  &      0  &    621  &    621  &  3483  &  60:12  &  X  &  X  \\ 
+PC:PS (5:1)/Na+  &      0  &    621  &      0  &    621  &  3483  &  60:12  &  X  &  X  \\ 
+PC:PS (5:1)/Na+  &      0  &   1510  &      0  &   1510  &  3377  &  60:12  &  X  &  X  \\ 
+PC:PS (5:1)/Na+  &      0  &   3002  &      0  &   3002  &  3213  &  60:12  &  X  &  X  \\ 
+PC:PS (5:1)/Na+  &    621  &      0  &      0  &    621  &  3483  &  60:12  &  X  &  X  \\ 
+PC:PS (5:1)/Na+  &   1510  &      0  &      0  &   1510  &  3377  &  60:12  &  X  &  X  \\ 
+PC:PS (5:1)/Na+  &   3002  &      0  &      0  &   3002  &  3213  &  60:12  &  X  &  X  \\ 
+  \hline
+\end{tabular}
+\end{table*}
+
 MD simulations of POPC:POPS lipid bilayers with different molar ratios 
@@
 
-\begin{table*}[tbp]
-\centering
-\caption{  List of MD simulations reporting their respective
-           added buffer concentrations,
-           amounts of individual molecules,
-           and the deployed models.
-         }\label{tbl:sim-list}
-         \todo{SAMULI: All the tails are PO, right? This could be mentioned in the caption.}
-\begin{tabular}{l | r r r r | r r | c c }
-                          &  \multicolumn{4}{c}{added buffer concentration / mM}    & \multicolumn{2}{c}{no. molecules} &  \multicolumn{2}{c}{simulated with model}  \\
-PC:PS (ratio)/counterions &  \ce{K^+}  &  \ce{Na^+} & \ce{Ca^{2+}} & \ce{Cl^-}      & \ce{H2O} &  PC:PS                 &  ECC-lipids  &  Lipid17    \\
-  \hline
-PC:PS (0:1)/K+   &      0  &      0  &      0  &      0  &  3600  &  0:72  &  X  &  X  \\ 
-PC:PS (0:1)/Na+  &      0  &      0  &      0  &      0  &  3600  &  0:72  &  X  &  X  \\ 
-  \hline
-PC:PS (10:1)/K+  &      0  &      0  &      0  &      0  &  3600  &  60:6  &  X  &  -  \\ 
-PC:PS (10:1)/Na+ &      0  &      0  &      0  &      0  &  3600  &  60:6  &  X  &  -  \\ 
-  \hline
-PC:PS (1:1)/K+   &      0  &      0  &      0  &      0  &  5253  &  64:64  &  X  &  X  \\ 
-PC:PS (1:1)/Na+  &      0  &      0  &      0  &      0  &  5253  &  64:64  &  X  &  X  \\ 
-  \hline
-PC:PS (4:1)/K+   &      0  &      0  &      0  &      0  &  3600  &  48:12  &  X  &  -  \\ 
-PC:PS (4:1)/Na+  &      0  &      0  &      0  &      0  &  3600  &  48:12  &  X  &  -  \\ 
-  \hline
-PC:PS (5:1)/K+   &      0  &      0  &      0  &      0  &  3600  &  60:12  &  X  &  X  \\ 
-PC:PS (5:1)/Na+  &      0  &      0  &      0  &      0  &  3600  &  60:12  &  X  &  X  \\ 
-PC:PS (5:1)/Na+  &      0  &      0  &     78  &     78  &  3561  &  60:12  &  X  &  X  \\ 
-PC:PS (5:1)/Na+  &      0  &      0  &    125  &    125  &  3561  &  60:12  &  X  &  X  \\ 
-PC:PS (5:1)/Na+  &      0  &      0  &    202  &    202  &  3561  &  60:12  &  X  &  X  \\ 
-PC:PS (5:1)/Na+  &      0  &      0  &    409  &    409  &  3522  &  60:12  &  X  &  X  \\ 
-PC:PS (5:1)/Na+  &      0  &      0  &    621  &    621  &  3483  &  60:12  &  X  &  X  \\ 
-PC:PS (5:1)/Na+  &      0  &    621  &      0  &    621  &  3483  &  60:12  &  X  &  X  \\ 
-PC:PS (5:1)/Na+  &      0  &   1510  &      0  &   1510  &  3377  &  60:12  &  X  &  X  \\ 
-PC:PS (5:1)/Na+  &      0  &   3002  &      0  &   3002  &  3213  &  60:12  &  X  &  X  \\ 
-PC:PS (5:1)/Na+  &    621  &      0  &      0  &    621  &  3483  &  60:12  &  X  &  X  \\ 
-PC:PS (5:1)/Na+  &   1510  &      0  &      0  &   1510  &  3377  &  60:12  &  X  &  X  \\ 
-PC:PS (5:1)/Na+  &   3002  &      0  &      0  &   3002  &  3213  &  60:12  &  X  &  X  \\ 
-  \hline
-\end{tabular}
-\end{table*}
-
 \begin{table}[tbp]
@@
     time-step           & 2~fs         \\
-    equilibration time  & tba~ns  \\
-    total simulation time     & 1.5~$\mu$s  \\
+    equilibration time  & 50~ns  \\
+    total simulation time     & 1~$\mu$s  \\
     temperature         & 298~K       \\
@@
   \caption{\label{simVSexpNOions_POPS} 
-    Top: X-ray scattering form factors from simulations with the Lipid17 \citep{lipid17-future} and 
-    the ECC-POPS models compared with experiments~\citep{kucerka14} at 298~K. 
-    Middle: Order parameters of POPS head group, glycerol backbone and acyl chains  
+    \textbf{Top:} X-ray scattering form factors from simulations with Lipid17/Dang \citep{lipid17-future, dang2006} and 
+    the ECC-POPS/ECC-ions \cite{martinek17, Pluhackova2016} compared with experiments~\citep{kucerka14} at 298~K. 
+    \textbf{Middle:} Order parameters of POPS head group, glycerol backbone and acyl chains  
     from simulations with the Lipid17 \citep{lipid17-future} and the ECC-POPS models 
@@
     Open/closed symbols are used for palmitoyl/oleoyl chains of POPS. 
-    Bottom: The chemical structure of POPS and the labeling of the carbon segments. 
-  }  
+    \textbf{Bottom:} The chemical structure of POPS and the labeling of the carbon segments. 
+  }  
+\todo{J: flip the chemical structure so that it agrees with the left-to-right reading of the plot.}
 \end{figure}
@@
 	   at 298~K with only \ce{Na^+} counterions and no additional ions. \label{tab:apls} } 
-  \begin{tabular}{l|c c} 
-%    \multicolumn{3}{c}{POPC} \\
-%    model          & APL / Å$^2$   & $T$ / K  \\ 
-%    \hline 
-%    Lipid14 POPC \citep{melcr18}    & 65.1$\pm$ 0.6  &  300 \\ 
-%    Lipid14 POPC \citep{dickson14}  & 65.6$\pm$ 0.5  &  303 \\ 
-%    \hline 
-%    ECC-POPC     \citep{melcr18}    & 63.2$\pm$ 0.6  &  300       \\ 
-%    \hline 
-%    experiment (SDP model) \citep{kucerka11} & 64.3  &  303    \\ 
-%    \hline 
-%    \multicolumn{3}{c}{} \\
-%    \multicolumn{3}{c}{POPE} \\
-%    model          & APL / Å$^2$   & $T$ / K  \\ 
-%    \hline 
-%    ECC-POPE                 & 56.7$\pm$ 0.8  &  298 \\ 
-%    \hline 
-%    experiment   \citep{parsegian89} & 56.6  &  310    \\ 
-%    experiment   \citep{rappolt03}   & 59--61 &  303--313  \\ 
-%    \hline 
-%    \multicolumn{3}{c}{} \\
-    \multicolumn{3}{c}{POPS} \\
-    model          & APL / Å$^2$   & $T$ / K  \\ 
-    \hline 
-    Lipid17 POPS              & 53.5$\pm$ 0.8  &  298 \\ 
-    \hline 
-    ECC-POPS                & 60.3$\pm$ 0.6  &  298       \\ 
-    \hline 
-    experiment (SDP model) \citep{kucerka14} & 62.3  &  298    \\ 
+  \begin{tabular}{l|c } 
+    \multicolumn{2}{c}{POPS} \\
+    model          & APL / Å$^2$    \\ 
+    \hline 
+    Lipid17/Dang              & 53.5$\pm$ 0.8   \\ 
+    Lipid17/ff99 \todoi{add?} & XX.x$\pm$ 0.8   \\ 
+    \hline 
+    ECC-POPS/ECC-ions         & 60.3$\pm$ 0.6   \\ 
+    \hline 
+    experiment (SDP model) \citep{kucerka14} & 62.3  \\ 
     \hline 
@@
 \end{table}
-
-\todo{SAMULI: Lipid17 simulations are too compressed, which is probably due to the
-  overbinding of Dang ions. Lipid17 with Åqvist ions (default in Amber) has larger
-  area in NMRlipids IV. This is similar to ECC-POPC paper where changing from Åqvist to
-  Dang or ECC-ions increased binding affinity in Lipid14. However, the headgroup order
-  parameters are better in ECC-POPS compared to Lipid17 simulations with any ions.
-  These should be discussed in this section.
-}
-
-\todo{SAMULI: For this discussion we need the counterion density profiles from pure POPS
-  simulation with Lipid17 and Dang counterions. This could be added to the figure \ref{fig:POPS-counterions-dens} and
-the figure could be moved here.}
 
@@
 We compare the values from experiments and simulations in Table~\ref{tab:apls}.
+
+\todo{SAMULI: Lipid17 simulations are too compressed, which is probably due to the
+  overbinding of Dang ions. Lipid17 with Åqvist ions (default in Amber) has larger
+  area in NMRlipids IV. This is similar to ECC-POPC paper where changing from Åqvist to
+  Dang or ECC-ions increased binding affinity in Lipid14. However, the headgroup order
+  parameters are better in ECC-POPS compared to Lipid17 simulations with any ions.
+  These should be discussed in this section.
+}
+
+\begin{figure}[tbp!] 
+  \centering 
+  \includegraphics[height=\figheight]{../img/ecc_pops/density_profiles_na-k-counterions_wat_phos_compar_purePOPS_ecclipids-lipid17.pdf}
+  \includegraphics[height=\figheight]{../img/ecc_pops/density_profiles_na-k-counterions_wat_phos_compar_5PC-1PS_ecclipids-lipid17.pdf}
+  \caption{\label{fig:POPS-counterions-dens}
+    Number density profiles of \ce{K^{+}}, \ce{Na^{+}} counterions along the membrane normal axis 
+    for the negatively charged bilayers composed of only POPS (left) or POPC:POPS (5:1, right) 
+    using ECC-lipids/ECC-ions (solid lines) and Lipid17/Dang ions (dashed lines).  
+    The density profiles of phosphate groups and water are divided by 4 and 100, respectively.  
+}
+\end{figure}
 
@@
 improves the description of the interactions even for very weakly binding cations like \ce{K^+}.
-
-\begin{figure}[tbp!] 
-  \centering 
-  \includegraphics[height=\figheight]{../img/ecc_pops/density_profiles_na-k-counterions_wat_phos_compar_purePOPS_ecclipids-lipid17.pdf}
-  \includegraphics[height=\figheight]{../img/ecc_pops/density_profiles_na-k-counterions_wat_phos_compar_5PC-1PS_ecclipids-lipid17.pdf}
-  \caption{\label{fig:POPS-counterions-dens}
-    Number density profiles of \ce{K^{+}}, \ce{Na^{+}} counterions along the membrane normal axis 
-    for the negatively charged bilayers composed of only POPS (left) or POPC:POPS (5:1, right) 
-    using ECC-lipids and ECC-ions (solid lines) and Lipid17 (dashed lines).  
-    The density profiles of phosphate groups and water are divided by 4 and 100, respectively.  
-}
-\end{figure}
 

### 2018-12-19 Manuscript/manuscript.tex
Small edits in the Results-first paragraph (added refs)

@@
 in pure water without any ions (or only counter ions)
-from simulations and experiments
-as the first step in the assessment of the quality of the model. 
+from simulations and experiments. 
 The experimental X-ray scattering form factors 
-of a bilayer are well reproduced  
+of a bilayer are well reproduced by ECC-POPS, 
+while for Lipid17/Dang they are in a slight disagreement
 (see Fig.~\ref{simVSexpNOions_POPS}). 
+The agreement of Lipid17 can be improved by using different parametes for ions \cite{aqvist90},
+which, however, bring other kinds of artifacts with them 
+(e.g., incorrect pairing of ions, condensation at larger salt concentrations \cite{kohagen16, chen07}).
 

### 2019-02-03 Manuscript/manuscript.tex
Added parts of text from NMRlipids4 paper for inspiration for writing.

@@
 
+\todo{EDIT: Taken from NMRlipids4 as an interesting information, to be edited.}
+In the experimental $^2$H NMR literature data used in this work~\cite{browning80,roux90},
+the lipids were first soluted to the buffer and then centrifuged to a pellet which
+was used in the measurements. Such samples have lower lipid concentration
+(approximately 10~wt~\% of lipids~\cite{browning80,roux88,roux90}) than 
+gravimetric samples (60~wt~\%) and simulations (approximately 50-60~wt~\%) in this work.
+Larger multilamellar repeat distances are expected in the samples with lower lipid
+concentrations due to the swelling caused by electrostatic repulsion in pure PS lipid systems~\cite{millman82}.
+However, the PS headgroup order parameters measured from gravimetric sample in this work
+are in good agreement with the results from centrifuged sample in the literature \cite{browning80} (Fig. \ref{HGorderParameters}).
+Furthermore, the equlibrium repeat distance rapidly decreases with the addtion of monovalent
+salts and is close to the simulation box sizes already above 500~mM concentrations~\cite{millman82,rand89}.
+Therefore, the hydration levels of multilamellae are expected to be sufficiently similar
+in the used simulations and reference experiments.
+
 In the reference simulations with the standard Amber force field,
@@
 and in Figs.~\ref{fig:delta_ordPar_NaCl_PC-PS_mix} as a function of PC and PS content.
+
+\todo{EDIT: Taken from NMRlipids4 as an interesting information, to be edited.}
+The headgroup order parameters
+of negatively charged PS and PG lipids exhibit systemic, but less
+characterized dependence on the bound charge~\cite{borle85,macdonald87,roux86,roux90}.
+Therefore, the ion binding affinity to negatively charged bilayers
+can be better characterized by measuring the PC headgroup order parameters from 
+mixed bilayers~\cite{roux86,roux90,roux91},
+
+\todo{EDIT: Taken from NMRlipids4 as an interesting information, to be edited.}
+Previous experimental studies have concluded that
+PS headgroups are more rigid than phophocholines (PC)
+due to the hydrogen bonding network or
+electrostatic interactions \cite{browning80,buldt81}.
+%
+Dilution of bilayers with PC lipids makes PS headgroups
+less rigid and reduces propensity for the formation of
+strong complexes with multivalent ions \cite{browning80,buldt81,roux90,roux91}.
+The molecular level interpretation of these observations is,
+however, not available.
 
@@
 \label{section:lip-ion_ca}
+
+\todo{EDIT: Taken from NMRlipids4 as an interesting information, to be edited.}
+The dehydrated complexes of PS headgroup and calcium ions can also lead to the
+phase separation \cite{hauser77,kurland79,hauser85,feigenson86,mattai89,roux90,roux91}.
+%
+CHARMM36 force field \cite{klauda10,venable13} combined with 2D infrared spectroscopy
+suggests that calcium ions interacts only with the carboxylate group of PS lipids \cite{valentine18}, while
+the same force field without the NBfix parameters together with the NMR chemical shifts and
+REDOR experiments suggests a significant binding affinity also to the phosphate region \cite{hallock18}.
+On the other hand, simulations with the Berger force field \cite{berger97,mukhopadhyay04}
+combined with fluorescent and vibrational sum frequency spectroscopy suggested a significant
+calcium binding also to the carbonyls in the acyl chains \cite{melcrova16}.
 

### 2019-02-05 Manuscript/manuscript.tex
Inputs from NMRlipid4 and small touches to the MS in methods and results-structure of POPS bilayer

@@
   \caption{Values of the area per lipid (APL) of POPS bilayers 
-	   at 298~K with only \ce{Na^+} counterions and no additional ions. \label{tab:apls} } 
+	   at 298~K with only \ce{Na^+} counterions and no additional ions. 
+           Standard deviations are given. \label{tab:apls} } 
   \begin{tabular}{l|c } 
@@
     \hline 
-    Lipid17/Dang              & 53.5$\pm$ 0.8   \\ 
-    Lipid17/ff99 \todoi{add?} & XX.x$\pm$ 0.8   \\ 
-    \hline 
-    ECC-POPS/ECC-ions         & 60.3$\pm$ 0.6   \\ 
+    Lipid17/Dang              & 53.5$\pm$ 1.7   \\ 
+    Lipid17/ff99              & 57.9$\pm$ 1.7   \\ 
+    \hline 
+    ECC-POPS/ECC-ions         & 59.8$\pm$ 1.8   \\ 
     \hline 
@@
 which, however, bring other kinds of artifacts with them 
-(e.g., incorrect pairing of ions, condensation at larger salt concentrations \cite{kohagen16, chen07}).
-
-The area per lipid is often used as a relatively simple structural parameter reporting on the bilayer properties and the packing of lipids. 
-In experiments, modeling is used on top of the scattering factors to obtain it \citep{kucerka14}. 
-From simulations, this property is easily extracted. 
+(e.g., incorrect pairing of ions, condensation at larger salt concentrations \cite{kohagen16, chen07, nmrlipids_proj4}).
+
+%The area per lipid is often used as a relatively simple structural parameter reporting on the bilayer properties and the packing of lipids. 
+In order to obtain the area per lipid from experiments, modeling is used on top of the scattering factors \citep{kucerka14}. 
+In ocntrast, this property is easily extracted directly from simulations with flat phospholipid bilayers. 
 We compare the values from experiments and simulations in Table~\ref{tab:apls}.
@@
 as shown in Fig. \ref{simVSexpNOions_POPS}. 
-The acyl chain order parameters in particular are almost all within the experimental error bars.
+The acyl chain order parameters in particular agree with the experimental data within error bars.
 The order parameters of the head groups are at an accuracy comparable to 
-other currently available classical models of lipids \citep{botan15, catte16, Pluhackova2016}.
+other currently available classical models of lipids \citep{botan15, catte16, Pluhackova2016, nmrlipids_proj4}.
 
@@
 as they are being used in the electrometer concept \cite{altenbach84, catte16, melcr18}.
-For POPC in pure water, the order parameter $\beta$ agrees well with the experiment, 
-while the order parameter $\alpha$ is somewhat lower. 
-In the case of POPS, the situation is a bit more complicated
-compared to POPC as the order parameter $\alpha$ exhibits a notable forking (see Fig.~\ref{simVSexpNOions_POPS}).
+%For POPC in pure water, the order parameter $\beta$ agrees well with the experiment, while the order parameter $\alpha$ is somewhat lower. 
+The head group order parameters of PS are a bit more complicated 
+compared to PC as the order parameter $\alpha$ exhibits a notable forking 
+(i.e., the hydrogen atoms in the group give rise to different order parameters, see Fig.~\ref{simVSexpNOions_POPS}).
 One of the order parameters of ECC-POPS, $\alpha_1$, agrees well with the experiment, 
@@
 which has a higher value closer to zero in the ECC-lipids model than in experiment. 
-Such a feature suggests that the model overestimates the orientational freedom of its head group.
+Such a feature suggests that the model likely overestimates the orientational freedom of the PS head group.
 

### 2019-02-05 Manuscript/manuscript.tex
Written PC-PS interaction section

@@
 
-\todo{SAMULI: Lipid17 simulations are too compressed, which is probably due to the
-  overbinding of Dang ions. Lipid17 with Åqvist ions (default in Amber) has larger
-  area in NMRlipids IV. This is similar to ECC-POPC paper where changing from Åqvist to
-  Dang or ECC-ions increased binding affinity in Lipid14. However, the headgroup order
-  parameters are better in ECC-POPS compared to Lipid17 simulations with any ions.
-  These should be discussed in this section.
-}
-
 \begin{figure}[tbp!] 
@@
 other currently available classical models of lipids \citep{botan15, catte16, Pluhackova2016, nmrlipids_proj4}. 
-
-The head group order parameters $\alpha$ and $\beta$ are highly relevant for this work,
-as they are being used in the electrometer concept \cite{altenbach84, catte16, melcr18}.
+In particular, ECC-POPS is closer to the experimental order parameters 
+than Lipid17 with either Dang \cite{smith94,chang1999,dang2006} 
+or ff99 ions \citet{aqvist90}.
+
+%The head group order parameters $\alpha$ and $\beta$ are highly relevant for this work,
+%as they are being used in the electrometer concept \cite{altenbach84, catte16, melcr18}.
 %For POPC in pure water, the order parameter $\beta$ agrees well with the experiment, while the order parameter $\alpha$ is somewhat lower. 
@@
 
-Changes of the lipid bilayer head group order parameters extracted from simulations and 
-experiments \cite{roux90} are shown in Figs.~\ref{fig:delta_ordPar_KCl_PCPS}, \ref{fig:delta_ordPar_NaCl_PCPS} 
-and~\ref{fig:delta_ordPar_CaC_PCPS} as functions of KCl, NaCl or CaCl$_2$ concentrations,
-and in Figs.~\ref{fig:delta_ordPar_NaCl_PC-PS_mix} as a function of PC and PS content.
-
-\todo{EDIT: Taken from NMRlipids4 as an interesting information, to be edited.}
-The headgroup order parameters
-of negatively charged PS and PG lipids exhibit systemic, but less
-characterized dependence on the bound charge~\cite{borle85,macdonald87,roux86,roux90}.
-Therefore, the ion binding affinity to negatively charged bilayers
-can be better characterized by measuring the PC headgroup order parameters from 
-mixed bilayers~\cite{roux86,roux90,roux91},
-
-\todo{EDIT: Taken from NMRlipids4 as an interesting information, to be edited.}
-Previous experimental studies have concluded that
-PS headgroups are more rigid than phophocholines (PC)
-due to the hydrogen bonding network or
-electrostatic interactions \cite{browning80,buldt81}.
-%
-Dilution of bilayers with PC lipids makes PS headgroups
-less rigid and reduces propensity for the formation of
-strong complexes with multivalent ions \cite{browning80,buldt81,roux90,roux91}.
-The molecular level interpretation of these observations is,
-however, not available.
+According to the electrometer concept,
+the head group order parameters $\alpha$ and $\beta$ of PC lipids
+change proportionally to the amount of bound charge
+in a fixed ratio. \cite{seelig87,scherer87,roux90} 
+This was demonstrated for a bound positive charge in simulations with POPC and 
+cationic surfactants or aquaeous cations employing the ECC-POPC model \cite{melcr18}.
+In this work, we will use the electrometer concept also for a bound negative charge 
+using PS lipids as negatively charged surfactants. \cite{roux90}
+
+Unlinke in pure PC bilayers,
+Interpreting the changes of the head group order parameters of 
+mixed PC:PS lipid bilayers 
+from both simulations and experiments
+is complicated by the presence of counterions. \cite{nmrlipids_proj4}
+In order to see the effects of different counterions, 
+we performed simulations with \ce{Na+} and also with \ce{K+} counterions, 
+which bind more weakly to phospholipids. \cite{nmrlipids_proj4, melcr18}
+This can also be seen in the density plots of the counterions in Fig.~\ref{fig:POPS-counterions-dens},
+which show a significantly reduced concentration of \ce{K+} ions 
+at the membrane head group region compared to \ce{Na+}
+for both PS and mixed PC:PS (5:1) bilayers.
+
+The slopes of the response of the PC head group order parameters $\alpha$ and $\beta$ 
+in the mixed PC:PS (5:1) bilayer with \ce{K+} counterions
+(Fig.~\ref{fig:delta_ordPar_NaCl_PC-PS_mix})
+agree qualitatively well with experiments for both employed models,
+i.e., for both Lipid14/17 and ECC-lipids. 
+The same property from the simualtions with \ce{Na+} counterions, 
+which were used in the experimental measurement,
+are not as good, however.
+In particular,
+ECC-lipids have a weaker response of their PC head group order parameters compared to the experiments, 
+while the simulations with Lipid17 model reveal a completely opposite trend.
+This effect likely arises from an overestimated binding of the sodium cations to the phospholipids \cite{nmrlipids_proj4},
+which is likely still present in a small amount also in the simulations with ECC-lipids. 
+The large difference between the affinities of the employed lipid models to the counterions 
+is visible in the density plots in Fig~\ref{fig:POPS-counterions-dens}.
+
+The order parameters of the PS head groups including their changes with the content of PC lipids
+are challenging to capture correctly in classical MD simulation models. \cite{nmrlipids_proj4}
+From such a perspective,
+the changes of the PS head group order parameters from ECC-lipids 
+with \ce{Na+}, and also \ce{K+}, counterions 
+agree comparably well to the experimental measurements
+(Fig.~\ref{fig:delta_ordPar_NaCl_PC-PS_mix}). 
+Interestingly enough, 
+although the Lipid17 model performed well with \ce{K+} counterions 
+for the PC head group order parameter changes,
+it does not agree in the equivalent changes in PS head groups
+pointing at non-negligible effects of even relatively weakly binding cations 
+on the structure of negatively charged lipids.
+
+Despite the great improvements in the overall structure of ECC-POPS 
+and its response to PC lipids and counterions,
+there is still a large room for improving it
+to agree with the experimental measurements.
 
@@
     The head group order parameters of a POPC:POPS (5:1) bilayer as a function of PS content
-    with \ce{Na^+} and \ce{K^+} counterions from simulations with ECC-lipids and Lipid17 \cite{lipid17-future} 
-    and experiments (only \ce{Na^+} counterions). \cite{roux90}. 
+    with \ce{Na^+} (top) and \ce{K^+} (bottom) counterions from simulations with ECC-lipids and Lipid17 \cite{lipid17-future} 
+    and experiments (only \ce{Na^+} counterions, but shown also in bottom plots). \cite{roux90}. 
     The underestimated response of ECC-lipids with \ce{Na^+} counterions 
@@
     which is corroborated by the series with \ce{K^+} counterions (lower affinity than \ce{Na^+}),
-    where the response is closer to the experiment (which uses only \ce{Na^+} counterions). 
+    where the response is closer to the experiment, which uses only \ce{Na^+} counterions. 
   }
@@
 %in the presence of monovalent salts, namely \ce{KCl} and \ce{NaCl}.
+
+\todo{EDIT: Taken from NMRlipids4 as an interesting information, to be edited.}
+The headgroup order parameters
+of negatively charged PS and PG lipids exhibit systemic, but less
+characterized dependence on the bound charge~\cite{borle85,macdonald87,roux86,roux90}.
+Therefore, the ion binding affinity to negatively charged bilayers
+can be better characterized by measuring the PC headgroup order parameters from 
+mixed bilayers~\cite{roux86,roux90,roux91},
+%
+Previous experimental studies have concluded that
+PS headgroups are more rigid than phophocholines (PC)
+due to the hydrogen bonding network or
+electrostatic interactions \cite{browning80,buldt81}.
+%
+Dilution of bilayers with PC lipids makes PS headgroups
+less rigid and reduces propensity for the formation of
+strong complexes with multivalent ions \cite{browning80,buldt81,roux90,roux91}.
+The molecular level interpretation of these observations is,
+however, not available.
+
+Figs.~\ref{fig:delta_ordPar_KCl_PCPS}, \ref{fig:delta_ordPar_NaCl_PCPS} 
+and~\ref{fig:delta_ordPar_CaCl_PCPS} as functions of KCl, NaCl or CaCl$_2$ concentrations,
+and in Figs.~\ref{fig:delta_ordPar_NaCl_PC-PS_mix} as a function of PC and PS content.
 

### 2019-02-22 Manuscript/manuscript.tex
Changed the title of the PC:PS section to include counterion binding.
Edited and cleaned Section on additional KCl and NaCl concnetntrations

@@
 }
+\todo{JOE: add labels with the relative surface excess $\Gamma$, which are promised to be present in the text.}
+\todo{JOE: split this figure into two: separate for POPS (to SI) and PC:PS 5:1 (main text)?}
 \end{figure}
@@
 
-\subsection{Interactions between PC and PS head groups in mixed PC:PS bilayers}
+\subsection{Binding of counterions to negatively charged PC:PS bilayers and interactions between PC and PS head groups}
 
@@
   }
+\todo{JOE+SAMULI: split this figure into separate plots for PC and PS, 
+and change the PS-plot x-axis to reflect the dilution of POPS by PC (now shows the content of PS). 
+The experimental lines in the bottom plot shall be different to make clear that they are not with K counterions (gray dashed?).}
 \end{figure*}
@@
 \subsection{Molecular interaction and binding affinities of K$^+$ and Na$^+$ cations to mixed POPC:POPS (5:1) membrane} 
-\label{sec:affinity}
+\label{sec:K_Na_added}
 
@@
     Changes of the head group order parameters $\alpha$, $\beta$ and the orientations of the carboxylate group and the P-N vector  
-    of POPC (left) and POPS (right) phospholipids in a POPC:POPS 5:1 bilayer as a function of \ce{NaCl} concentration 
-    in bulk ($C_{ion}$) from simulations with different force fields at 298 K.
+    of POPC (left) and POPS (right) phospholipids in a POPC:POPS 5:1 bilayer 
+    as a function of \ce{NaCl} concentration in the buffer 
+    from simulations with different force fields and experiments at 298 K. \citep{roux90}
     Because data with \ce{NaCl} are not available for POPC, 
@@
     Changes of the head group order parameters $\alpha$, $\beta$ and the orientations of the carboxylate group and the P-N vector  
-    of POPC (left) and POPS (right) phospholipids in a POPC:POPS 5:1 bilayer as a function of \ce{KCl} concentration 
-    in bulk ($C_{ion}$) from simulations with different force fields and experiments at 298 K. \citep{roux90}
+    of POPC (left) and POPS (right) phospholipids in a POPC:POPS 5:1 bilayer 
+    as a function of \ce{KCl} concentration in the buffer 
+    from simulations with different force fields and experiments at 298 K. \citep{roux90}
     The orientation of the \ce{COO^-} group is defined as 
@@
 
-% Pavel suggested to remove this as it only recaps previous sections
-%It was presented in the previous section,
-%that accounting for electronic polarization,
-%a distinct feature of ECC-lipids compared to other lipid models, 
-%is crucial for an accurate description of the response of the POPC electrometer. 
-%In this section,
-%we employ ECC-lipids to simulate neutral and negatively charged bilayers 
-%in the presence of monovalent salts, namely \ce{KCl} and \ce{NaCl}.
-
-\todo{EDIT: Taken from NMRlipids4 as an interesting information, to be edited.}
-The headgroup order parameters
-of negatively charged PS and PG lipids exhibit systemic, but less
-characterized dependence on the bound charge~\cite{borle85,macdonald87,roux86,roux90}.
-Therefore, the ion binding affinity to negatively charged bilayers
-can be better characterized by measuring the PC headgroup order parameters from 
-mixed bilayers~\cite{roux86,roux90,roux91},
+% This§: 
+% NA/K binding status, 
+% how we tried finding a model (NMRlipids4), and 
+% how it was futile
 % 
-Previous experimental studies have concluded that
-PS headgroups are more rigid than phophocholines (PC)
-due to the hydrogen bonding network or
-electrostatic interactions \cite{browning80,buldt81}.
-%
-Dilution of bilayers with PC lipids makes PS headgroups
-less rigid and reduces propensity for the formation of
-strong complexes with multivalent ions \cite{browning80,buldt81,roux90,roux91}.
-The molecular level interpretation of these observations is,
-however, not available.
-
-Figs.~\ref{fig:delta_ordPar_KCl_PCPS}, \ref{fig:delta_ordPar_NaCl_PCPS} 
-and~\ref{fig:delta_ordPar_CaCl_PCPS} as functions of KCl, NaCl or CaCl$_2$ concentrations,
-and in Figs.~\ref{fig:delta_ordPar_NaCl_PC-PS_mix} as a function of PC and PS content.
-
 The binding of \ce{Na^+} cations to phospholipid bilayers 
@@
 but Lipid14 (PC) \citep{dickson14}, resp. Lipid17 (PS) \citep{lipid17-future}. 
-This model yields a semi-quantitative agreement with the experimentally measured small changes of the order parameters 
-when used with the model of ions by \citet{aqvist90} (Fig.~\ref{fig:catte16}). 
+This model yields a semi-quantitative agreement with the experimentally measured small changes of the order parameters \cite{catte16}
+when used with the model of ions by \citet{aqvist90}. 
 However, when used with a more accurate model of ions by \citet{Pluharova2014, martinek17},
@@
 
+% This§: 
+% we have developed a model that can describe the NMR exp. using ECC
+% 
 Generally improved behaviour 
@@
 was achieved through the combination of models 
-ECC-lipids \citep{melcr18} and ECC-ions \citep{martinek17, kohagen16, Pluharova2014}. 
+ECC-lipids and ECC-ions \citep{martinek17, kohagen16, Pluharova2014}. 
 Simulations with these models 
@@
 
+% This§: 
+% Other models do not yield good reponse, but
+% ECC-lipid describe the changes well
+% => Elect. polar. necessarry.
+% 
 The interaction with \ce{K^+}, which binds very weakly to both neutral and negatively charged membranes, 
-renders a qualitatively different response of the order parameter $S^\beta$ in POPS in the mixed negatively charged bilayers compared to the neutral bilayers. 
+renders a qualitatively different response of the order parameter $S^\beta$ in POPS 
+in the mixed negatively charged bilayers compared to the neutral bilayers. 
 While the order parameter $S^\beta$ increases for both \ce{Na^+} and \ce{Ca^{2+}},
-it decreases in the presence of \ce{K^+}.
-%This feature is, \emph{qualitatively} captured by a few models in \citep{nmrlipids_proj4},
-%neither of them is as close to a \emph{qunatitative} agreement with the experiments 
-%as the combination of ECC-lipids with ECC-ions. 
-However, no model studied in \citep{nmrlipids_proj4} describes the behaviour of a PS head group correctly enough to reveal this effect. 
-In contrast, ECC-lipids with ECC-ions capture the different response of the order parameters $S^{\beta}$, $S^{\alpha _1}$ and $S^{\alpha _2}$ in POPS to various salts accurately. 
-Such a detailed description of the changes of the structural parameters 
-demonstrates that including electronic polarization
-improves the description of the interactions even for very weakly binding cations like \ce{K^+}.
+it decreases in the presence of \ce{K^+}
+(\ce{KCl}:  Fig.~\ref{fig:delta_ordPar_KCl_PCPS}, 
+ \ce{NaCl}: Fig.~\ref{fig:delta_ordPar_NaCl_PCPS}, and
+\ce{CaCl2}: Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}). 
+Despite the tremendous effort in the development of classical MD models of lipids, however,
+there is no such model, 
+which would agree with the experimental trends for these salt solutions \cite{nmrlipids_proj4}. 
+In contrast, ECC-lipids with ECC-ions capture the different response 
+of the order parameters $S^{\beta}$, $S^{\alpha _1}$ and $S^{\alpha _2}$ in POPS 
+to various concentrations of \ce{NaCl} and \ce{KCl} in a good agreement with the experiments
+(\ce{KCl}:  Fig.~\ref{fig:delta_ordPar_KCl_PCPS}, 
+ \ce{NaCl}: Fig.~\ref{fig:delta_ordPar_NaCl_PCPS}). 
+Such a dramatic improvement in the structural parameters 
+demonstrates that including electronic polarization in phospholipids is crucial 
+even for interactions with soft cations like \ce{K^+}.
 
@@
 
-The difference between the affinities of \ce{Na^+} and \ce{K^+} to neutral and negatively charged membranes
-can be described by their relative surface excesses with respect to water, $\Gamma ^{w} _{ion}$, 
-which are shown in the plots of the density profiles of the ions in Figs.~\ref{fig:nacl-dens} (neutral bilayer)
-\todo{SAMULI: We should have a figure showing density profiles of Na+ and K+ in pure POPS and PC:PS (5:1) mixture
-  with only counterions from ECC-POPS model and Lipid17 model.
-}
-and \ref{fig:nacl-dens_PCPS} (negatively charged bilayer). 
+% This§: 
+% adsorption and affinities of Na and K to PC:PS memb
+% from densities and rel.surf.excess
+% 
+%in  Fig.~\ref{fig:POPS-counterions-dens} (only counterions)
+%and Fig.~\ref{fig:nacl-dens_PCPS}        (additional salt concentrations), 
+%
+The affinities of \ce{Na^+} and \ce{K^+} to phospholipid membranes
+can be described by their relative surface excesses with respect to water, $\Gamma ^{w} _{ion}$. 
 Such a quantity compares the adsorption of ions to the adsorption of water molecules at an interface 
 without the necessity of defining a Gibbs dividing surface. \citep{melcr18, chattorajBOOK}
-While \ce{K^+} maintains negative values of $\Gamma^{w}_{K}$ even for the negatively charged bilayer,
-the value of $\Gamma^{w}_{Na}$ for \ce{Na^+} changes from negative to positive
-in a neutral POPC bilayer versus in a negatively charged bilayer with the composition 5\,PC:1\,PS.
-%This means that at the given concentration the bilayer interface has a small preference to \ce{Na^+} cations compared to water molecules. 
-Interestingly, this value is slightly decreased in the presence of an additional \ce{NaCl} concentration adding also \ce{Cl^-} anions, 
-which are not present in the system when only counterions are used
-(bottom resp. top plot in Fig.~\ref{fig:nacl-dens_PCPS}). 
-The relative surface excess of calcium, $\Gamma^{w}_{Ca}$, also reflects
-the higher affinity of the negatively charged PC:PS (5:1) membrane compared to a neutral POPC bilayer; 
-its value for the latter is       $\Gamma^{w}_{Ca} = 0.06\mathrm{nm^{-2}}$, 
-while for the PC:PS mixture it is $\Gamma^{w}_{Ca} = 0.24\mathrm{nm^{-2}}$
-at comparable concentrations $C _{Ca} = 260\pm 30 \mathrm{mM} $.
+Interestingly enough, \ce{K^+} maintains negative values of $\Gamma^{w}_{K}$ even for the negatively charged PC:PS~(5:1) bilayer.
+This is also evident from the density plots in  Fig.~\ref{fig:nacl-dens_PCPS}
+showing that the density profile of \ce{K^+} along the membrane normal 
+decays clearly faster towards the membrane centre compared to the density of water. 
+In contrast, the density profile of \ce{Na^+} has a shoulder 
+at the region of the phosphate groups,
+which demonstrates that sodium cations adsorb weakly to the membrane interface compared to the aquaeous solution
+by yielding a small positive value of $\Gamma^{w}_{Na}$.
+
+\todo{JOE: Add a discussion about the still slightly overestimated affinity of \ce{Na^+}.}
 
@@
 as can be seen from its growing density at the interface.
+
+The relative surface excess of calcium, $\Gamma^{w}_{Ca}$, also reflects
+the higher affinity of the negatively charged PC:PS (5:1) membrane compared to a neutral POPC bilayer; 
+its value for the latter is       $\Gamma^{w}_{Ca} = 0.06\mathrm{nm^{-2}}$, 
+while for the PC:PS mixture it is $\Gamma^{w}_{Ca} = 0.24\mathrm{nm^{-2}}$
+at comparable concentrations $C _{Ca} = 260\pm 30 \mathrm{mM} $.
 

### 2019-03-05 Manuscript/manuscript.tex

@@
 
-\todo{EDIT: Taken from NMRlipids4 as an interesting information, to be edited.}
-The dehydrated complexes of PS headgroup and calcium ions can also lead to the
-phase separation \cite{hauser77,kurland79,hauser85,feigenson86,mattai89,roux90,roux91}.
-%
-CHARMM36 force field \cite{klauda10,venable13} combined with 2D infrared spectroscopy
-suggests that calcium ions interacts only with the carboxylate group of PS lipids \cite{valentine18}, while
-the same force field without the NBfix parameters together with the NMR chemical shifts and
-REDOR experiments suggests a significant binding affinity also to the phosphate region \cite{hallock18}.
-On the other hand, simulations with the Berger force field \cite{berger97,mukhopadhyay04}
-combined with fluorescent and vibrational sum frequency spectroscopy suggested a significant
-calcium binding also to the carbonyls in the acyl chains \cite{melcrova16}.
-
 \begin{figure}[tbp!] 
@@
 
-%%% Sum up what is to be presented briefly
-%The importance of treating polarizability for interactions of phospholipids with monovalent ions was demonstrated in the previous section. 
-Electronic polarization is a non-negligible contribution to the interactions of calcium even in simple aqueous solutions of \ce{CaCl2} \citep{martinek17, kohagen16, Pluharova2014}. 
-In this section,
-we will show the results from simulations of neutral and negatively charged phospholipid bilayers at varying \ce{CaCl2} concentrations
-using the recently developed models ECC-lipids and ECC-ions \citep{melcr18, martinek17}, 
-which implicitly include the effects of electronic polarization through electronic continuum correction \citep{leontyev11}. 
-Validated with the concept of a lipid electrometer in section \ref{section:electrometer_exp_sim},
-these implicitly polarizable models yield accurate description of 
-the interaction of \ce{Ca^{2+}} with both neutral and negatively charged phospholipids.
-
-The changes of the head group order parameters $S^\alpha$ and $S^\beta$ from simulations and experiments 
-are shown in Fig.~\ref{fig:delta_ordPar_CaCl} for a neutral POPC bilayer, 
-and in Fig.~\ref{fig:delta_ordPar_CaCl_PCPS} for a negatively charged bilayer with the composition 5\,PC:1\,PS. 
-For a direct comparison and a connection to our works \citep{catte16, nmrlipids_proj4},
+The interaction of various cations including calcium 
+with  neutral and also negatively charged lipid bilayers
+was thoroughly compared between simulations and experiments 
+from the point of the electrometer concept \cite{roux90, seelig87, altenbach84}
+in the references~\citenum{catte16, nmrlipids_proj4}. 
+It was concluded that the interaction of calcium with zwitterionic PC bilayers 
+is generally overestimated in all studied models \cite{catte16},
+while none of the models was capable of interpreting the NMR structural parameters
+with the negatively charged PC:PS (5:1) bilayers \cite{nmrlipids_proj4}. 
+Accounting for the electronic polarization was identified as 
+a non-negligible correction to the interaction between cations and POPC bilayers in Ref.~\cite{melcr18}. 
+Here, we demonstrate that the same approach 
+significantly improves the description of calcium binding 
+also to the negatively charged PC:PS (5:1) bilayers. 
+In correspondence with the Ref.~\citenum{nmrlipids_proj4},
+we also use the changes of the head group order parameters $S^\alpha$ and $S^\beta$ 
+in PC and PS lipids 
+to quantify the binding affinity of the calcium cations
+and the induced structural changes
+(Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}). 
+To directly connect to the previous works \citep{catte16, nmrlipids_proj4}
+and to mark the improvement after treating electronic polarizability, 
 we show simulation results from ECC-lipids and also from Lipid17 \citep{lipid17-future}. 
-This also highlights the improvements in ECC-lipids over Lipid17 arising from the electronic polarization. 
-%Although Lipid17 already belongs to the top-performing models in terms of the responses of the head group order parameters in those studies,  
-%including electronic polarizability to form ECC-lipids improves the results even further.
-The effect is probably the most striking for POPS, 
-for which also the structure of a pure POPS bilayer with only counterions is dramatically improved with the augmentation.
-
-%%% Discuss the changes of OPs and vectors from neutral and neg. membranes
-%While the changes induced by \ce{CaCl2} are \emph{qualitatively} correct for Lipid14/17, 
-Increasing concentrations of \ce{CaCl2} induce a systematic decrease of the order parameters $S^\alpha$ and $S^\beta$ in POPC. 
-Although the total magnitude of the response of the PC head group order parameters 
-is only slightly higher in the negatively charged bilayers than in the neutral bilayers, 
-the shape of the changes in the latter shows a steeper onset at low concentrations. 
-This is apparently due to the presence of POPS, 
-which has a higher affinity to \ce{Ca^{2+}} compared to POPC. 
-ECC-lipids is the first model,
-which achieves a \emph{quantitative} agreement with the changes induced by \ce{CaCl2} in experiments.
-
-\todo{SAMULI: The new results show that we cannot actually reproduce the rapid changes with low CaCl2
-  concentrations in POPS order parameters. This should be somehow discussed.}
+While the response of the order parameters of PC to \ce{CaCl2} concentrations
+is qualitatively correct in the Lipid14/17 model \cite{lipid17},
+it is quantitatively comparable to the experimental trends in ECC-lipids. 
+Moreover, the structure of PS head group is fundamentaly improved
+after accounting for the electronic polarization 
+(Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}).
+
+Increasing concentrations of \ce{CaCl2} 
+induce a systematic decrease of the order parameters $S^\alpha$ and $S^\beta$ 
+in both neutral POPC bilayers \cite{catte16, melcr18}
+as well as in the negatively charged PC:PS (5:1) bilayers \cite{nmrlipids_proj4, roux90}. 
+The change of the head group order parameters 
+are well reproduced only by ECC-lipids in Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}.
+
+The structural changes in POPS induced by increasing concentrations of \ce{CaCl2} 
+are very challenging to interpret from both experiments \cite{roux90} 
+and also from MD simulations \cite{nmrlipids_proj4}. 
+The changes of the order parameters $S^\alpha$ and $S^\beta$ in PS head groups from simulations 
+are not found to follow the experimental profiles, 
+nor any common trend,
+with the currently available models \cite{nmrlipids_proj4}. 
+In Fig.~\ref{fig:delta_ordPar_CaCl_PCPS},
+we plot these structural parameters for one of the models in that work, Lipid17, 
+and for ECC-lipids. 
+Although the agreement between ECC-lipids and experiments is far from optimal,
+the general trends and behaviour even at highly concentrated solutions of \ce{CaCl2} 
+are well captured by the model. 
+For example the order parameter $\beta$ in POPS 
+adopts a saturated value in experiments and ECC-lipids,
+but it is growing with the \ce{CaCl2} concentration in Lipid17.
+In addition, 
+the forking of the order parameters $\alpha$ 
+is almost constant in the experiment and ECC-lipids,
+but it varies greatly in th Lipid17.
+
+The changes of the order parameters $\alpha$ and $\beta$ at low \ce{CaCl2} concentrations
+have a steep onset at low concentrations in the negatively charged bilayers compared to neutral POPC bilayers. 
+However, such a detail is not captured by ECC-lipids
+indicating the necessity for further improvements.
+The source of this discrepancy may arise 
+from the interactions between PC and PS lipids and with \ce{Na^+} counterions,
+from the sub-optimal structure of the lipids (Fig.~\ref{simVSexpNOions_POPS},
+and also possibly from the approximate mean-field treatment of the electronic polarization.
 
@@
   Melcrova et al. Scientific Reports | 6:38035 | DOI: 10.1038/srep38035 }
+%
+\todo{EDIT: Taken from NMRlipids4 as an interesting information, to be edited.}
+The dehydrated complexes of PS headgroup and calcium ions can also lead to the
+phase separation \cite{hauser77,kurland79,hauser85,feigenson86,mattai89,roux90,roux91}.
+%
+CHARMM36 force field \cite{klauda10,venable13} combined with 2D infrared spectroscopy
+suggests that calcium ions interacts only with the carboxylate group of PS lipids \cite{valentine18}, while
+the same force field without the NBfix parameters together with the NMR chemical shifts and
+REDOR experiments suggests a significant binding affinity also to the phosphate region \cite{hallock18}.
+On the other hand, simulations with the Berger force field \cite{berger97,mukhopadhyay04}
+combined with fluorescent and vibrational sum frequency spectroscopy suggested a significant
+calcium binding also to the carbonyls in the acyl chains \cite{melcrova16}.
 

### 2019-03-06 Manuscript/manuscript.tex
Written/edited part dealing with P-N vector and COO- vector in the section with calcium. Next: Continue editing the part with the interacting moieties.

@@
 
+It has been pointed out in previous experimental works
+that the conformational flexibility of PS head groups 
+is different from other lipids \cite{browning80}. 
+While the static shielding tensor of PS is not substantially different from PC or PE, 
+the $^{31}$P~$T_1$ relaxation times are significantly shorter for PS. 
+It is suggested that the PS head group is more rigid compared to the other phospholipids, 
+which is also corroborated by a larger chemical shielding anisotropy of PS \cite{browning80}.
+
+We have investigated this effect using the mean orientations of P-N and \ce{COO^-} vectors 
+(defined and plotted in Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}). 
+While the mean angle between the P-N vector of PC head group and the membrane normal
+decreased by 11$^\circ$ from 64$^\circ$ to 55$^\circ$
+ at 620~mM buffer concentration of \ce{CaCl2},
+the mean orientation of the P-N vector in the PS head group has 
+decreased only by 3$^\circ$ from 67$^\circ$ to 64$^\circ$ 
+in the same concentration range.
+
+Such a largely diminished flexibility 
+may arise from the binding of calcium to the \ce{COO^-} group in PS. 
+Whe have measured the mean orientation of the \ce{COO^-} group 
+as the connector of the carbon atoms 
+forming the bond between the group and the $\beta$-carbon of the phospholipid. 
+The mean angle between the \ce{COO^-} vector and the membrane normal
+changes by $11^\circ$ from $62^\circ$ to $73^\circ$ (more towards the membrane interior)
+ at 620~mM buffer concentration of \ce{CaCl2}.
+The conformational flexibility of the PS head group is hence restrained 
+by the binding of calcium cations to the \ce{COO^-} group, 
+which reorients towards the membrane interior and to the phosphate region, 
+where calcium cations bind predominantly.
+
 \begin{figure}[htbp!] 
@@
 
-%%% Present the density profiles (different concentrations for both POPC and mixed)
-The increase in the amount of bound calcium cations from pure POPC \cite{melcr18}
-to the mixed negatively charged bilayer containing POPS
+The binding of calcium cations to the phospholipid bilayer is shown 
+in the density profiles of \ce{Ca^{2+}}, \ce{Na^+} counterions 
+and also \ce{Cl^-} ions in Fig.~\ref{fig:cacl-dens_PCPS}. 
+Clearly, after accounting for the electronic polarizability, 
+the population of the cations in the phosphate region is significantly diminished. 
+\todo{JOE: The density plot still needs to be changed to show both Lipid17 and ECC-lipids, and to support this argument.}
+Still, the dominant contribution to the binding of calcium comes from the phosphate groups. 
+Interestingly enough,
+although the \ce{COO^-} groups increase the affinity of calcium towards PS compared to PC,
+ there is no extra density at the outmost surface of the membrane. 
+This can be understood from the point of the structural rearangements of the \ce{COO^-} group, 
+which take place during the interactions with calcium cations
+described above and in Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}.
+
+The higher affinity of calcium to PS head groups compared to PC
 is well demonstrated using the relative surface excess, $\Gamma ^w _{Ca}$,
 summarized in Table~\ref{tab:binding}. 
-Distributions of \ce{Ca^{2+}}, \ce{Na^+} counterions and also \ce{Cl^-} 
-are plotted in Fig.~\ref{fig:cacl-dens_PCPS} for the negatively charged bilayer. 
-In contrast to \ce{KCl} or added concentrations of \ce{NaCl}, 
-\ce{Na^+} counterions are substituted with \ce{Ca^{2+}} even at low concentrations of \ce{CaCl2}. 
-The increasing concentration of \ce{CaCl2} and, hence, a higher amount of bound \ce{Ca^{2+}}
-also attracts \ce{Cl^-} anions to the bilayer 
-as can be seen from its growing density at the interface.
-
-The relative surface excess of calcium, $\Gamma^{w}_{Ca}$, also reflects
-the higher affinity of the negatively charged PC:PS (5:1) membrane compared to a neutral POPC bilayer; 
-its value for the latter is       $\Gamma^{w}_{Ca} = 0.06\mathrm{nm^{-2}}$, 
-while for the PC:PS mixture it is $\Gamma^{w}_{Ca} = 0.24\mathrm{nm^{-2}}$
-at comparable concentrations $C _{Ca} = 260\pm 30 \mathrm{mM} $.
-
-\todo{SAMULI: This paragraph in unclear. Should be clarified and the changes
-  of headgroup orientations in PC and PS clearly discussed.
-}
-%%% discuss the table with Gammas and follow up with molecular details, where do cations bind? Support with the spatial density figure
-The density profiles of ions suggest that
-the dominant contribution to the binding of \ce{Ca^{2+}} to phospholipid bilayers
-comes from the interactions with the phosphate groups in both POPC and POPS. 
-%The population analysis also suggests 
-%that the calcium cations prefer to reside in the phosphate region of the membrane. 
-%Such a finding is further validated with a more detailed analysis using Markov state modeling (MSM) \citep{Pande_MSM_paper} \todoi{Add papers reviewing MSM, e.g. recent Pande's paper.}. 
-%The set of states of a calcium cation 
-%encompassed all possible combinations of up to three surrounding lipids. 
-%In the case of PC, we used only the phosphate moitety, which forms the dominant contribution for calcium binding.
-%For PS we also distinguished configurations in which calcium interacts with the carboxylate moiety. 
-%All possible combinations of such states were used to build a Markov model at a lag time $25\,\mathrm{ns}$ using pyEMMA code by \citet{pyemma} \todoi{add citation for pyemma}. 
-%The resulting MSM was validated using Chapman-Kolmogorov test \citep{FrankNoe_papers_MSM} \todoi{Add citation for Noe's papers on MSM, especially Chapman Kolmogorov test.}. 
-%The stationary distribution of the states reveals a strong preference of the calcium cations to reside in the phosphate region of the mixed PC-PS membrane. 
-%When interacting with PS, configurations with the phosphate moiety from either PC or PS dominate the population.
-%Moreover, states containing interactions with the carboxylate group in PS 
-%bear higher probability when interacting also with the phosphate group of the same lipid or other lipids. 
-This is also reflected in the shift of the mean orientation 
-of the \ce{COO^-} group from $62^\circ$ to $73^\circ$ (420~mM \ce{CaCl2})
-which was measured as the connector of the carbon atoms 
-forming the bond between the group and the $\beta$-carbon of the phospholipid. 
-The interactions of the carboxylate group in PS with calcium and other phosphate groups
-shed light into the qualitatively different response of the head group order parameters $\alpha$ and $\beta$ in PS compared to PC 
-(see Figs.~\ref{fig:delta_ordPar_CaCl} and~\ref{fig:delta_ordPar_CaCl_PCPS}). 
-We find that the complex response of the head group order parameters of POPS 
-is affected by the conformational changes of the carboxylate group,
-which is attracted more towards the phosphate region, 
-where the calcium cations dominantly bind. 
-In line with the experimental work by \citet{browning80},
-this is also very likely the reason, 
-why the magnitude of the P-N vector change in POPS is diminished compared to POPC, 
-which is not restrained by an additional cation binding group like \ce{COO^-} in POPS.
-
-Further details about the interactions of \ce{Ca^{2+}} with various moieties in POPC or POPS
-were obtained by counting contacts between the cations and the oxygen atoms of the lipids
-similarly as was done in \cite{melcr18}. 
-The threshold for counting a contact was set to $0.3\,\mathrm{nm}$, 
-which encompasses the first peak of the radial distribution function between the cations and the oxygen atoms of the lipids.
+While $\Gamma^{w}_{Ca}$ at a neutral POPC bilayer
+is       $\Gamma^{w}_{Ca} = 0.06\mathrm{nm^{-2}}$, 
+its value for the PC:PS (5:1) bilayer is $\Gamma^{w}_{Ca} = 0.24\mathrm{nm^{-2}}$
+at comparable buffer concentrations  $350 \, \mathrm{mM}$ resp. $400\, \mathrm{mM}$.
 
@@
 \centering
-  \caption{Bulk concentrations, $C _{Ca}$, and molar fractions, $C' _{Ca}$, of Ca$^{2+}$;
+  \caption{Bulk concentrations, $C _{Ca}$, and buffer concentrations, $C' _{Ca}$, of Ca$^{2+}$;
            relative surface excess of calcium with respect to water ($\Gamma_{Ca}^{\rm water}$); 
@@
 \end{table}
+
+Further details about the interactions of \ce{Ca^{2+}} with various moieties in POPC or POPS
+were obtained by counting contacts between the cations and the oxygen atoms of the lipids
+similarly as was done in \cite{melcr18}. 
+The threshold for counting a contact was set to $0.3\,\mathrm{nm}$, 
+which encompasses the first peak of the radial distribution function between the cations and the oxygen atoms of the lipids.
 

### 2019-03-06 Manuscript/manuscript.tex
Update text

--- 2019-03-05 Manuscript/manuscript.tex
@@
 %% Generate PDF/A-2u
-\usepackage[a-2u]{pdfx}
+%\usepackage[a-2u]{pdfx}
 
@@
 \usepackage[english]{babel}
+\usepackage{xr}
+\externaldocument{manuscriptSUPPL}
 
@@
 %% Most settings are pre-set by the pdfx package.
-\hypersetup{unicode}
-\hypersetup{breaklinks=true}
-\hypersetup{colorlinks,allcolors=DodgerBlue4}
+%\hypersetup{unicode}
+%\hypersetup{breaklinks=true}
+%\hypersetup{colorlinks,allcolors=DodgerBlue4}
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
 %% The manuscript does not need to include \maketitle, which is 
@@
 \subsection{Validation of membrane properties}
-
-To validate the structures of lipid bilayers, the scattering form factors and C-H bond order parameters
-were calculated and compared against the experimental data as described previously~\cite{botan15,ollila16,melcr18,nmrlipids_proj4}.
-%in simulations without ions were validated against NMR 
-%by calculating the order parameters for the C-H bonds and against \mbox{X-ray} scattering experiments by evaluating 
-%as was done in Refs.~\citenum{catte16, melcr18, nmrlipids_proj4}.
-%NMR order parameters validate the structures sampled by the individual lipid molecules with atomic resolution. 
-%The simulated order parameters were calculated for all C-H bonds in lipid molecules, 
-%while all order paremeters in the head group and only some in the \ce{C-H} groups in tails were resolved experimentally
-%(see next section and Fig.~\ref{simVSexpNOions_POPS}). 
-The order parameters for each C-H bond of individual lipids in the trajectory were first averaged
-over the time. Then the average over lipids was taken and the standard error of the mean of this
-average was used as the error estimate of the mean \cite{botan15,ollila16,nmrlipids_proj4}.
 
@@
 
+The C--H bond order parameters were first averaged over the time for individual lipids and
+then the standard error of the mean of the average over lipids was used as the error estimate
+of the mean \cite{botan15,ollila16,nmrlipids_proj4}.
+
 \begin{table}[tbp]
@@
 
-We compared X-ray scattering form factors and NMR order parameters of bilayers
-in pure water without any ions (or only counter ions)
+The form factors from scattering and C-H bond order parameters from NMR experiments are good measures to validate
+the lipid bilayer structure in MD simulations against experiments because they
+are related to the bilayer dimensions (area per molecule and thickness) and
+conformational fluctuations of individual lipids, respectively, and 
+can be directly calculated from MD simulations~\cite{ollila16}.
+The X-ray scattering form factors and C-H bond order parameters in
+the headgroup and glycerol backbone region of POPS lipid bilayer are already available in the literature \cite{pan14,nmrlipids_proj4}.
+Here, we extend the available experimental dataset of POPS lipids by measuring the C--H bond
+order parameters also from the acyl chain region using R-PDLF experiments as described previously \cite{ferreira13}.
+We first determined the order parameters of each peak corresponding different C--H bonds detected
+in the first dimension (Fig. \ref{R-PDLF}) from the dipolar slices in the indirect dimension (Fig. \ref{R-PDLFslices}).
+Because the exact assignment of these peaks was not possible, we assigned the
+acyl chain order parameters to mimic the profile predicted by the MD simulations (Fig. \ref{simVSexpNOions_POPS}).
+\todo{Tiago has to check and update this and figures in SI.}
+
+To evaluate the effect of ECC on the structural properties of POPS lipid bilayer properties in MD simulations,
+we compared the results from a simulation with original Lipid17 \cite{??} combined with the Dang ion parameters \cite{??}
+with the results from ECC-POPS simulation and experiments (Fig. \ref{simVSexpNOions_POPS}).
+with sodium counterions are prevously publihed
+(or only counter ions)
 from simulations and experiments. 

### 2019-03-06 Manuscript/manuscript.tex
Update text

--- 2019-03-06 Manuscript/manuscript.tex
@@
 a non-negligible correction to the interaction between cations and POPC bilayers in Ref.~\cite{melcr18}.
+
 Here, we demonstrate that the same approach 
@@
 and to mark the improvement after treating electronic polarizability, 
-we show simulation results from ECC-lipids and also from Lipid17 \citep{lipid17-future}. 
+we show simulation results from the newly developed ECC-lipids and also from Lipid17 \citep{lipid17-future}. 
 While the response of the order parameters of PC to \ce{CaCl2} concentrations
@@
 and for ECC-lipids. 
-Although the agreement between ECC-lipids and experiments is far from optimal,
+Although the agreement between ECC-lipids and experiments is not optimal,
 the general trends and behaviour even at highly concentrated solutions of \ce{CaCl2} 
@@
 as the connector of the carbon atoms 
-forming the bond between the group and the $\beta$-carbon of the phospholipid. 
+forming the bond between the group and the $\beta$-carbon of the phospholipid 
+(Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}).
 The mean angle between the \ce{COO^-} vector and the membrane normal
@@
 Clearly, after accounting for the electronic polarizability, 
-the population of the cations in the phosphate region is significantly diminished. 
+the population of the cations at the membrane interface is significantly diminished. 
 \todo{JOE: The density plot still needs to be changed to show both Lipid17 and ECC-lipids, and to support this argument.}
@@
 
-The higher affinity of calcium to PS head groups compared to PC
-is well demonstrated using the relative surface excess, $\Gamma ^w _{Ca}$,
-summarized in Table~\ref{tab:binding}. 
-While $\Gamma^{w}_{Ca}$ at a neutral POPC bilayer
-is       $\Gamma^{w}_{Ca} = 0.06\mathrm{nm^{-2}}$, 
-its value for the PC:PS (5:1) bilayer is $\Gamma^{w}_{Ca} = 0.24\mathrm{nm^{-2}}$
-at comparable buffer concentrations  $350 \, \mathrm{mM}$ resp. $400\, \mathrm{mM}$.
+\todo{JOE: suggestion -- what is the charge at the membrane interface of memb+counterions -vs- memb+CaCl2 concnetration?
+                         this could relate well to the zeta potential measurements and the Fig 1 in \cite{melcrova16}. }
 
@@
 
-Further details about the interactions of \ce{Ca^{2+}} with various moieties in POPC or POPS
-were obtained by counting contacts between the cations and the oxygen atoms of the lipids
-similarly as was done in \cite{melcr18}. 
+% Gamma and higher affinity of PS to CA2+
+The higher affinity of calcium to PS head groups compared to PC
+is well demonstrated using the relative surface excess, $\Gamma ^w _{Ca}$,
+summarized in Table~\ref{tab:binding}. 
+While $\Gamma^{w}_{Ca}$ at a neutral POPC bilayer
+is       $\Gamma^{w}_{Ca} = 0.06\mathrm{nm^{-2}}$      \cite{melcr18}, 
+its value for the PC:PS (5:1) bilayer is $\Gamma^{w}_{Ca} = 0.24\mathrm{nm^{-2}}$
+at comparable buffer concentrations  $350 \, \mathrm{mM}$ resp. $400\, \mathrm{mM}$.
+
+% Description of no. conctacts counting
+Such a significantly increased binding also has an impact 
+on the structure and populations of the bound calcium cations. 
+We have counted contacts between the cations and the oxygen atoms of the lipids
+similarly as was done in \cite{melcr18} 
+to obtain such a detailed information 
+about the interactions of \ce{Ca^{2+}} with various moieties in POPC or POPS. 
 The threshold for counting a contact was set to $0.3\,\mathrm{nm}$, 
@@
 
+% Populations from contacts
 The percentages of the populations of membrane-bound calcium cations for various membrane moieties 
@@
 Even though the negatively charged membrane contains only 18\% of POPS, 
-approximately 41\% of the total population of bound calcium cations is in contact with PS lipids
+on average 41\% of the total population of bound calcium cations is in contact with PS lipids
 with 8\% bound only to them. 
-This corroborates the intrinsically higher affinity of PS lipids to calcium cations compared to neutral PC lipids. 
-POPC interacts with the calcium cations almost entirely through its phosphate group 
-in both neutral and negatively charged membranes. 
-Interactions of \ce{Ca^{2+}} with carbonyl groups are also present, 
+% Where Calcium interacts
+While POPC interacts with the calcium cations almost entirely through its phosphate group 
+in both neutral \cite{melcr18} and negatively charged membranes, 
+the inteaction with POPS lipids happens more through the \ce{COO^-} group than through the \ce{PO4} group. 
+Interactions of \ce{Ca^{2+}} with carbonyl groups are also present for both POPC and POPS,
 however, they are always accompanied by interactions with phosphate groups. 
-%but a very small negligable amount much below 1\%. 
-\todo{Also COO- vs. phosphate binding should be discussed and the results should be
-  discussed with respect to the literature: https://doi.org/10.1016/j.bpj.2018.08.044 and
-  Melcrova et al. Scientific Reports | 6:38035 | DOI: 10.1038/srep38035 }
-%
-\todo{EDIT: Taken from NMRlipids4 as an interesting information, to be edited.}
-The dehydrated complexes of PS headgroup and calcium ions can also lead to the
-phase separation \cite{hauser77,kurland79,hauser85,feigenson86,mattai89,roux90,roux91}.
-%
-CHARMM36 force field \cite{klauda10,venable13} combined with 2D infrared spectroscopy
-suggests that calcium ions interacts only with the carboxylate group of PS lipids \cite{valentine18}, while
-the same force field without the NBfix parameters together with the NMR chemical shifts and
-REDOR experiments suggests a significant binding affinity also to the phosphate region \cite{hallock18}.
-On the other hand, simulations with the Berger force field \cite{berger97,mukhopadhyay04}
-combined with fluorescent and vibrational sum frequency spectroscopy suggested a significant
-calcium binding also to the carbonyls in the acyl chains \cite{melcrova16}.
+This is in contrast to 
+simulations with the Berger force field \cite{berger97,mukhopadhyay04}
+combined with fluorescent and vibrational sum frequency spectroscopy,
+which suggest a significant
+calcium binding also to the carbonyls groups \cite{melcrova16}. 
+This feature arises likely from the inaccuracy of the Berger model 
+and its overestimated binding of cations to that region \cite{nmrlipids_proj4}.
+
+Measurements using infrared spectroscopy 
+combined with simulations with the CHARMM36 model 
+with a correction for overestimated binding of calcium \cite{valentine18}
+reveal similar trends for the populations at various PS moieties
+suggesting that POPS interacts mostly through its \ce{COO^-} group. 
+While this feature agrees nicely with our findings, 
+the balance of the interactions of PS moieties with calcium cations may be suboptimal
+and the interaction strength underestimated in total in those simulations \cite{nmrlipids_proj4}.
+
+In contrast, the combination of NMR experiments 
+and MD simulations with the same model, CHARMM36, but without the correction for calcium binding
+also reveal long-lived interactions of \ce{Ca^{2+}} with \ce{PO4} groups. \cite{hallock18}
+Although the ratios of the life times of coordination follow the same trends as in our work,
+the lengths of simulation trajectories are insufficient 
+to yield converged results in this aspect (see SI and Refs.~\cite{catte16, melcr18, nmrlipids_proj4}). 
+Moreover, the content of POPS in the bilayers in that work is
+much higher above the threshold after which 
+\ce{Ca^{2+}} cations induce lipid clustering. \cite{hauser77,kurland79,hauser85,feigenson86,mattai89,roux90,roux91}
+At such a high content of negatively charged lipids,
+PS head groups are found to form even four-membered clusters \cite{hallock18}. 
+Our simulations are, however, at different conditions,
+in which the clustering of POPS does not happen, 
+and PS lipids prefer 1:1 ratio with \ce{Ca^{2+}}
+as plotted in Fig.~\ref{fig:cacl_complexes}
+showing relative probabilities of \ce{Ca^{2+}} complexes with a certain number of PC or PS lipids. 
+Compared to a neutral POPC bilayer
+the negatively charged bilayer has its stoichiometry 
+shifted towards complexes with three phospholipids and one calcium. 
+This is obviously due to the presence of POPS, 
+which also contributes to the pre-formed one- and two-membered clusters of POPC. 
+Clusters of four or more lipids were not observed in either membrane.
+
+\todo{ THE REMAINING PIECES TO BE DISCUSSED  --  consider wheter the dynamics slowdown or the dihedral distributions could make good contributions to this paper. }
+
+% Ca2+ induces much slower dynamics of lipids when PS is present,
+% this does not happen with PC
+% uses NBFIX
+From Ref.~\cite{valentine18}://
+To summarize, calcium binding to PS lipids slows the dynamics at the ester position and introduces more heterogeneous molecular environments. 
+The slowdown cannot be attributed to complete ester dehydration or decreased hydrogen bonding to water, so we interpret these results as evidence that the lipid headgroups adopt more rigid but disordered conformations on the molecular level.
+The decrease in dynamics observed experimentally indicates that the network of hydrogen bonds between lipids and water at the polar-nonpolar interface becomes slower, as it was shown by Karathanou and Bondar (34).
+This is bolstered by recent NMR work showing that binding to Ca2+ induces two different rigid conformations in the glycerol backbone of lipid headgroups (18).
+We have investigated the molecular origin of the rigidification and decreased lateral diffusion, finding that calcium binding does not dehydrate the ester region, but even with only 10\% PS lipids, the local dynamics slow around the esters.
+The slower local dynamics provide a molecular explanation for the Ca2+-dependent decreases in fluidity observed on longer length and timescales.
+There is considerable room for further improvement in the force fields for ion-membrane interactions. Particularly, small differences in short-range interactions may determine the local geometries (34
+).
+Most of our conclusions are, however, based on long-range effects (i.e., electric field generated by the ion at significant distances from the carbonyl groups), which are less sensitive to the details of the force field.
+
+Their Figure 5 \cite{hallock18} shows the distributions of O-Ca-Cb-N dihedral angle
+revealing separate populations of the PS head groups. 
+These two structures are also compared in therm of P-Cg distances, 
+which can be measured from NMR REDOR exp. 
+It is concluded that since the two distinct populations in P-Cg distance can be represented with the 
+dihedral distributions, 
+they also likely represent the distinct populations. 
+ -> plotting the distribution of O-Ca-Cb-N dihedral angle may be also a good idea for us (SI?)
+     -> NMRlipids 4 in SI \cite{nmrlipids_proj4} plots this information: 
+           - C36 agress with the distribution in \cite{hallock18}
+           - Lipid17 (and Slipids) are almost identical to C36
+           - but the structure of the COO- is highly different (dih angle Ca-Cb-Cg-Og)
 
@@
 
-Relative probabilities of \ce{Ca^{2+}} complexes with a certain number of lipids are presented in Fig.~\ref{fig:cacl_complexes}. 
-Calcium cations that are bound only to PC in the mixed bilayer with PS 
-behave similarly as in the pure PC bilayer
-maintaining comparable probabilities for clustering one, two, or even three PC lipids together. 
-In contrast, PS lipids prefer 1:1 ratio with \ce{Ca^{2+}},
-which may also be due to their low molar fraction in the the mixed bilayer. 
-In total, however, the negatively charged membrane has its stoichiometry 
-shifted towards complexes with three phospholipids and one calcium. 
-This is obviously due to the presence of POPS, 
-which also contributes to the pre-formed one- and two-membered clusters of POPC. 
-Clusters of four or more lipids were not observed in either membrane.
-
-%This is also reflected in the increased probability 
-%of a single phospholipid interacting with two \ce{Ca^{2+}} cations 
-%compared to the neutral POPC bilayer, 
-%where such a probability is almost negligable. 
-% Not shown through numbers, however, I have the plots to back it.
-
-%\todo{Stationary distribution: Make a figure documenting the populations of bound \ce{Ca^{2+}} cations (like I have in the presentation) that would accompany Table \ref{tab:Ca_binding_PCPS}. 
-%This will roughly correspond to the PC stoichiometry plot \ref{fig:cacl_complexes}. }
+\section{Conclusions}
+
+\todo{SAMULI: I have putted here two preliminary sentences
+JOSEF: I can write conclusions, but I would prefer to wait with that after your comments and edits. }
+Applying ECC to POPS lipids improves cation binding affinity to membranes
+mixed with POPC and also the response of PS headgroup to the bound cations.
+However, further improvement of the parameters is needed to fully reproduce the
+details of calcium binding to PS headgroups.
+
+\listoftodos
+
+\section{SI}
+
+Timescales associated with the binding of calcium cations from solution to the membrane
+are plotted for each binding event as a histogram in Fig.~\ref{fig:hist_residence_times}. 
+Using these plots, we can estimate that 90\% of the residence times of any calcium cation 
+will be lower than $60\,\mathrm{ns}$ for pure POPC neutral bilayer 
+and shorter than $200\,\mathrm{ns}$ for the mixed 5\,PC:1\,PS negatively charged bilayer. 
+The longest observed residence times in the simulations were $141\,\mathrm{ns}$ for the neutral membrane 
+and $485\,\mathrm{ns}$ for the negatively charged membrane. 
+Both estimates of the residence times come from simulations with comparable concentrations of around $250\mathrm{mM}$;
+the simulation with the neutral membrane has a bulk concentration of calcium $C_{ion} = 280\mathrm{mM}$, 
+whereas the simulation with the negatively charged membrane has a bulk concentration of calcium $C_{ion} = 240\mathrm{mM}$.
 
@@
 
-Timescales associated with the binding of calcium cations from solution to the membrane
-are plotted for each binding event as a histogram in Fig.~\ref{fig:hist_residence_times}. 
-Using these plots, we can estimate that 90\% of the residence times of any calcium cation 
-will be lower than $60\,\mathrm{ns}$ for pure POPC neutral bilayer 
-and shorter than $200\,\mathrm{ns}$ for the mixed 5\,PC:1\,PS negatively charged bilayer. 
-The longest observed residence times in the simulations were $141\,\mathrm{ns}$ for the neutral membrane 
-and $485\,\mathrm{ns}$ for the negatively charged membrane. 
-Both estimates of the residence times come from simulations with comparable concentrations of around $250\mathrm{mM}$;
-the simulation with the neutral membrane has a bulk concentration of calcium $C_{ion} = 280\mathrm{mM}$, 
-whereas the simulation with the negatively charged membrane has a bulk concentration of calcium $C_{ion} = 240\mathrm{mM}$.
-
 In summary, the results from ECC-lipids suggest 
@@
 are required for the negatively charged bilayers.
-
-%In addition to such estimates of the time scales, we used the Markov model on top of the simulation with the negatviely charged mixed bilayer
-%to calculate the time of the mean first passage of calcium from solution to the membrane (and in reverse) resulting in $55\,\mathrm{ns}$  ($165\,\mathrm{ns}$).
-
-%\todo{Fluxes: committor analysis, dominant fluxes (table and figure)}
-%From the spectrum of the transition matrix, we also observe that 
-%the slowest transitions are associated with the binding to two or three phosphate moieties from either PC or PS,
-%which are also among the states with the highest probabilities.
-%The net fluxes of calcium cations from solution to such states also form a large contribution to the total flux ($\approx 45\%$). 
-%\todo{Make a table and a figure of the state probabilities and fluxes to support this statement.}. 
-%Analysis of the possible binding pathways of calcium cations to the negatively charged mixed bilayer 
-%reveal that the cations mostly enter the membrane bound states directly from solution 
-%without complicated transitions at the time scales of the order of the lag time of the Markov state model, $25\,\mathrm{ns}$.
-
-Provide density plots and show, where and how the cations bind to the membrane.
-
-\section{Conclusions}
-
-\todo{SAMULI: I have putted here two preliminary sentences}
-Applying ECC to POPS lipids improves cation binding affinity to membranes
-mixed with POPC and also the response of PS headgroup to the bound cations.
-However, further improvement of the parameters is needed to fully reproduce the
-details of calcium binding to PS headgroups.
-
-\listoftodos
 

### 2019-03-07 Manuscript/manuscript.tex
Added some missing literature and included exceprts from recent literature to be discussed in the paper in the section with Ca2+ which is still to be heavily edited.

--- 2019-03-06 Manuscript/manuscript.tex
@@
 
+% Description of no. conctacts counting
 Further details about the interactions of \ce{Ca^{2+}} with various moieties in POPC or POPS
@@
 This corroborates the intrinsically higher affinity of PS lipids to calcium cations compared to neutral PC lipids. 
+% Where Calcium interacts
 POPC interacts with the calcium cations almost entirely through its phosphate group 
-in both neutral and negatively charged membranes. 
+in both neutral \cite{melcr18} and negatively charged membranes. 
 Interactions of \ce{Ca^{2+}} with carbonyl groups are also present, 
@@
 %but a very small negligable amount much below 1\%.
-\todo{Also COO- vs. phosphate binding should be discussed and the results should be
-  discussed with respect to the literature: https://doi.org/10.1016/j.bpj.2018.08.044 and
-  Melcrova et al. Scientific Reports | 6:38035 | DOI: 10.1038/srep38035 }
-%
+
+Relative probabilities of \ce{Ca^{2+}} complexes with a certain number of lipids are presented in Fig.~\ref{fig:cacl_complexes}. 
+Calcium cations that are bound only to PC in the mixed bilayer with PS 
+behave similarly as in the pure PC bilayer
+maintaining comparable probabilities for clustering one, two, or even three PC lipids together. 
+In contrast, PS lipids prefer 1:1 ratio with \ce{Ca^{2+}},
+which may also be due to their low molar fraction in the the mixed bilayer. 
+In total, however, the negatively charged membrane has its stoichiometry 
+shifted towards complexes with three phospholipids and one calcium. 
+This is obviously due to the presence of POPS, 
+which also contributes to the pre-formed one- and two-membered clusters of POPC. 
+Clusters of four or more lipids were not observed in either membrane.
+
+%This is also reflected in the increased probability 
+%of a single phospholipid interacting with two \ce{Ca^{2+}} cations 
+%compared to the neutral POPC bilayer, 
+%where such a probability is almost negligable. 
+% Not shown through numbers, however, I have the plots to back it.
+
+Timescales associated with the binding of calcium cations from solution to the membrane
+are plotted for each binding event as a histogram in Fig.~\ref{fig:hist_residence_times}. 
+Using these plots, we can estimate that 90\% of the residence times of any calcium cation 
+will be lower than $60\,\mathrm{ns}$ for pure POPC neutral bilayer 
+and shorter than $200\,\mathrm{ns}$ for the mixed 5\,PC:1\,PS negatively charged bilayer. 
+The longest observed residence times in the simulations were $141\,\mathrm{ns}$ for the neutral membrane 
+and $485\,\mathrm{ns}$ for the negatively charged membrane. 
+Both estimates of the residence times come from simulations with comparable concentrations of around $250\mathrm{mM}$;
+the simulation with the neutral membrane has a bulk concentration of calcium $C_{ion} = 280\mathrm{mM}$, 
+whereas the simulation with the negatively charged membrane has a bulk concentration of calcium $C_{ion} = 240\mathrm{mM}$.
+
+In summary, the results from ECC-lipids suggest 
+that the exchange of calcium between the POPC bilayer and the solvent 
+occurs at the order of $\sim$10--100~ns, 
+which is significantly faster than observed in simulations 
+with other presently available non-polarizable models of lipids and ions~\citep{javanainen17, catte16}. 
+Our results suggest 
+that simulation trajectories with a characteristic length of several hundreds of nanoseconds 
+are necessary to capture the binding of calcium to neutral POPC bilayers 
+in equilibrium when more realistic \emph{polarizable} force fields are used. 
+Interestingly enough, almost an order of magnitude longer simulations
+are required for the negatively charged bilayers.
+
+%In addition to such estimates of the time scales, we used the Markov model on top of the simulation with the negatviely charged mixed bilayer
+%to calculate the time of the mean first passage of calcium from solution to the membrane (and in reverse) resulting in $55\,\mathrm{ns}$  ($165\,\mathrm{ns}$).
+
+%\todo{Fluxes: committor analysis, dominant fluxes (table and figure)}
+%From the spectrum of the transition matrix, we also observe that 
+%the slowest transitions are associated with the binding to two or three phosphate moieties from either PC or PS,
+%which are also among the states with the highest probabilities.
+%The net fluxes of calcium cations from solution to such states also form a large contribution to the total flux ($\approx 45\%$). 
+%\todo{Make a table and a figure of the state probabilities and fluxes to support this statement.}. 
+%Analysis of the possible binding pathways of calcium cations to the negatively charged mixed bilayer 
+%reveal that the cations mostly enter the membrane bound states directly from solution 
+%without complicated transitions at the time scales of the order of the lag time of the Markov state model, $25\,\mathrm{ns}$.
+
+FOLLOWING: lit. search on recent literature, which shall be discussed in this work.
+
 \todo{EDIT: Taken from NMRlipids4 as an interesting information, to be edited.}
@@
 calcium binding also to the carbonyls in the acyl chains \cite{melcrova16}.
+
+% Details of interactions between calcium and phospholipids (PC and PS) .
+% Major conclusions correct, only the model (Berger) over-exaggerates certain features. 
+From abstract of \cite{melcrova16}://
+Namely, time-resolved fluorescent spectroscopy of lipid vesicles 
+and vibrational sum frequency spectroscopy of lipid
+monolayers are used to characterize local binding sites of calcium in zwitterionic and anionic model
+lipid assemblies, while dynamic light scattering and zeta potential measurements are employed for
+macroscopic characterization of lipid vesicles in calcium-containing environments. To gain additional
+atomic-level information, the experiments are complemented by molecular simulations that utilize an
+accurate force field for calcium ions with scaled charges effectively accounting for electronic polarization
+effects. We demonstrate that lipid membranes have substantial calcium-binding capacity, with several
+types of binding sites present. Significantly, the binding mode depends on calcium concentration with
+important implications for calcium buffering, synaptic plasticity, and protein-membrane association.
+
+And From conclustions of \cite{melcrova16}://
+Calcium ion binding sites are heterogeneous both from the point of view of binding affinity and their positioning
+in the membrane. The character of the calcium binding varies with calcium concentration; this issue is of par-
+ticular importance as significant concentration spikes of calcium ions occur along calcium signaling pathways.
+The present results support the conjecture that lipid membranes, in particular the negatively charged inner leaflet
+of the plasma membrane, can act as effective calcium buffers upon calcium ions entering the cytosol. Of equal
+importance is the fact that the strong binding of this ion significantly alters the membranes by means of reduction
+of their hydration, lipid mobility, and lateral inter-lipid distance. Such local conformational membrane remod-
+eling may play a significant role in modulation of lipid-protein interactions as well as membrane-membrane
+interactions. Overall, the phenomena related to calcium ions-membrane interactions demonstrated here point to
+their diverse roles in biological systems.
+
+% Ca2+ induces much slower dynamics of lipids when PS is present,
+% this does not happen with PC
+From Ref.~\cite{valentine18}://
+To summarize, calcium binding to PS lipids slows the dynamics at the ester position and introduces more heterogeneous molecular environments. 
+The slowdown cannot be attributed to complete ester dehydration or decreased hydrogen bonding to water, so we interpret these results as evidence that the lipid headgroups adopt more rigid but disordered conformations on the molecular level.
+The decrease in dynamics observed experimentally indicates that the network of hydrogen bonds between lipids and water at the polar-nonpolar interface becomes slower, as it was shown by Karathanou and Bondar (34).
+This is bolstered by recent NMR work showing that binding to Ca2+ induces two different rigid conformations in the glycerol backbone of lipid headgroups (18).
+We have investigated the molecular origin of the rigidification and decreased lateral diffusion, finding that calcium binding does not dehydrate the ester region, but even with only 10\% PS lipids, the local dynamics slow around the esters.
+The slower local dynamics provide a molecular explanation for the Ca2+-dependent decreases in fluidity observed on longer length and timescales.
+There is considerable room for further improvement in the force fields for ion-membrane interactions. Particularly, small differences in short-range interactions may determine the local geometries (34
+).
+Most of our conclusions are, however, based on long-range effects (i.e., electric field generated by the ion at significant distances from the carbonyl groups), which are less sensitive to the details of the force field.
+
+% 2 major head group configurations of PS in 30-70 PC-PS mix -- paper 1
+From Ref.~\cite{boettcher11}://
+ In this study, combining magic angle spinning (MAS) solid-state NMR (SSNMR) measurements of isotopically labeled serine headgroups in mixed lipid bilayers with molecular dynamics (MD) simulations of PS lipid bilayers in the presence of different counterions, we provide site-resolved insights into the effects of Ca2+ on the structure and dynamics of lipid bilayers.
+Ca2+-induced conformational changes of PS in mixed bilayers are observed in both liposomes and Nanodiscs, a nanoscale membrane mimetic of bilayer patches.
+Site-resolved multidimensional correlation SSNMR spectra of bilayers containing 13C,15N-labeled PS demonstrate that Ca2+ ions promote two major PS headgroup conformations, which are well resolved in two-dimensional 13C-13C, 15N-13C, and 31P-13C spectra.
+The results of MD simulations performed on PS lipid bilayers in the presence or absence of Ca2+ provide an atomic view of the conformational effects underlying the observed spectra.
+but the NMR ...
+ Spectra were acquired of 70\% POPS* and 30\% POPC Nanodiscs with 2 mM Ca2+ using an MAS rate of 10 kHz on a 600 MHz (1H frequency) spectrometer
+and but the MD sim was done only for pure POPS bilayer and no other counterions but Ca2+ with CaCl2 concentration.
+
+% 2 major head group configurations of PS in 30-70 PC-PS mix -- paper 2
+% extended analysis and insight. 
+From Ref.~\cite{hallock18}://
+(conclusions) We investigated the atomistic details of how membranes containing anionic PS lipids are shaped into their functional state by Ca2+ ions.
+The lipid phosphate groups and the captured ions are made readily accessible to interact with proteins.
+Using extensive MD simulations with a lipid-accelerated membrane representation, we demonstrated that PS lipid nanoclusters are formed by two distinct PS headgroup conformations that are induced by Ca2+ ions, which bind to the lipids in a specific manner.
+Importantly, detailed characterization of MD-derived lipid nanoclusters and geometries of individual PS lipids agree with SSNMR measurements on two different spatial scales.
+(1) The chemical shifts show the accuracy of capturing the chemical environment around a single lipid in either of the two long-lived conformations, and (2) the Ca2+-coordinated PS lipid tetramer is a basic structural unit that is involved in membrane sculpturing.
+By studying the effects of ions on membranes containing anionic lipids, we have characterized a baseline behavior and structures of the lipids.
+These results will be important in interpreting the role of lipids in future protein–membrane studies, and they demonstrate that MD HMMM simulations combined and validated by SSNMR methodologies are effective at capturing high-resolution signature lipid structures that are stabilized by membrane-active agents in complex membrane environments.
+but ... again, the experiment with nanodiscs (and also the bilayer simulation with HMMM) 
+uses 70:30 PS:PC ratio, which induces PS clustering.
+ -> hence the markable difference between their figure 3 and our figure \ref{fig:cacl_complexes}
+ - they observe most probable Ca2+ clusters of 2, then 3, 1 and 4 PS together
+ - while we see rather only 1PS per ca2+ in Figure \ref{fig:cacl_complexes}
+On the other hand, I see correspondence with the observed longer life time ( stronger interaction?)
+with \ce{COO-} group in PS in \cite{hallock18} in Fig 4, for both Ca2+ and Na+. 
+This is followed by PO4 in both cases. 
+In the case of Ca2+, the ratio of the observed life time is also in line with our findings, that 
+the interaction with the PO4  and COO- together is of a smaller abundance compared to the interactions of the individual groups themselves, yet comparable to that of PO4. 
+Their Figure 5 shows the distributions of O-Ca-Cb-N dihedral angle
+revealing separate populations of the PS head groups. 
+These two structures are also compared in therm of P-Cg distances, 
+which can be measured from NMR REDOR exp. 
+It is concluded that since the two distinct populations in P-Cg distance can be represented with the 
+dihedral distributions, 
+they also likely represent the distinct populations. 
+ -> plotting the distribution of O-Ca-Cb-N dihedral angle may be also a good idea for us (SI?)
+     -> NMRlipids 4 in SI \cite{nmrlipids_proj4} plots this information: 
+           - C36 agress with the distribution in \cite{hallock18}
+           - Lipid17 (and Slipids) are almost identical to C36
+           - but the structure of the COO- is highly different (dih angle Ca-Cb-Cg-Og)
 
@@
 \end{figure}
-
-Relative probabilities of \ce{Ca^{2+}} complexes with a certain number of lipids are presented in Fig.~\ref{fig:cacl_complexes}. 
-Calcium cations that are bound only to PC in the mixed bilayer with PS 
-behave similarly as in the pure PC bilayer
-maintaining comparable probabilities for clustering one, two, or even three PC lipids together. 
-In contrast, PS lipids prefer 1:1 ratio with \ce{Ca^{2+}},
-which may also be due to their low molar fraction in the the mixed bilayer. 
-In total, however, the negatively charged membrane has its stoichiometry 
-shifted towards complexes with three phospholipids and one calcium. 
-This is obviously due to the presence of POPS, 
-which also contributes to the pre-formed one- and two-membered clusters of POPC. 
-Clusters of four or more lipids were not observed in either membrane.
-
-%This is also reflected in the increased probability 
-%of a single phospholipid interacting with two \ce{Ca^{2+}} cations 
-%compared to the neutral POPC bilayer, 
-%where such a probability is almost negligable. 
-% Not shown through numbers, however, I have the plots to back it.
 
@@
 
-Timescales associated with the binding of calcium cations from solution to the membrane
-are plotted for each binding event as a histogram in Fig.~\ref{fig:hist_residence_times}. 
-Using these plots, we can estimate that 90\% of the residence times of any calcium cation 
-will be lower than $60\,\mathrm{ns}$ for pure POPC neutral bilayer 
-and shorter than $200\,\mathrm{ns}$ for the mixed 5\,PC:1\,PS negatively charged bilayer. 
-The longest observed residence times in the simulations were $141\,\mathrm{ns}$ for the neutral membrane 
-and $485\,\mathrm{ns}$ for the negatively charged membrane. 
-Both estimates of the residence times come from simulations with comparable concentrations of around $250\mathrm{mM}$;
-the simulation with the neutral membrane has a bulk concentration of calcium $C_{ion} = 280\mathrm{mM}$, 
-whereas the simulation with the negatively charged membrane has a bulk concentration of calcium $C_{ion} = 240\mathrm{mM}$.
-
-In summary, the results from ECC-lipids suggest 
-that the exchange of calcium between the POPC bilayer and the solvent 
-occurs at the order of $\sim$10--100~ns, 
-which is significantly faster than observed in simulations 
-with other presently available non-polarizable models of lipids and ions~\citep{javanainen17, catte16}. 
-Our results suggest 
-that simulation trajectories with a characteristic length of several hundreds of nanoseconds 
-are necessary to capture the binding of calcium to neutral POPC bilayers 
-in equilibrium when more realistic \emph{polarizable} force fields are used. 
-Interestingly enough, almost an order of magnitude longer simulations
-are required for the negatively charged bilayers.
-
-%In addition to such estimates of the time scales, we used the Markov model on top of the simulation with the negatviely charged mixed bilayer
-%to calculate the time of the mean first passage of calcium from solution to the membrane (and in reverse) resulting in $55\,\mathrm{ns}$  ($165\,\mathrm{ns}$).
-
-%\todo{Fluxes: committor analysis, dominant fluxes (table and figure)}
-%From the spectrum of the transition matrix, we also observe that 
-%the slowest transitions are associated with the binding to two or three phosphate moieties from either PC or PS,
-%which are also among the states with the highest probabilities.
-%The net fluxes of calcium cations from solution to such states also form a large contribution to the total flux ($\approx 45\%$). 
-%\todo{Make a table and a figure of the state probabilities and fluxes to support this statement.}. 
-%Analysis of the possible binding pathways of calcium cations to the negatively charged mixed bilayer 
-%reveal that the cations mostly enter the membrane bound states directly from solution 
-%without complicated transitions at the time scales of the order of the lag time of the Markov state model, $25\,\mathrm{ns}$.
-
-Provide density plots and show, where and how the cations bind to the membrane.
-
 \section{Conclusions}

### 2019-03-07 Manuscript/manuscript.tex
Additional text from valentine18

@@
 
+ Ca2+ primarily binds carboxylates
+MD trajectories indicate that Ca2+ primarily binds to the carboxylates.
+These results are consistent with previous literature (19 , 22 , 32 , 33) but are surprising in the context of the IR results.
+Previous FTIR results show minimal changes to the carboxylate peaks and large changes to the phosphate and ester peaks (25
+).
+Even though the calcium ions primarily bind with the carboxylates in our MD trajectories, their effect on the electric field around the carboxylates is minimal because of screening by water molecules and averaging of orientations.
+In contrast, the less polar domain of the carbonyl experiences more significant electric field (Fig.
+6).
+This is consistent with recent 2D IR and ab initio MD simulations by Kuroda and co-workers in which no changes in the IR spectra or relaxation dynamics of the acetate carboxylate were observed in solutions up to 6 M NaCl (69
+).
+Recent results from our own group also point toward larger changes in carboxylate frequencies with a monodentate binding configuration (70).
+Likewise, in our study, carboxylates are fully hydrated and primarily experience electric field fluctuations from the surrounding water, and the ion does not appear to modify the local dynamics.
+Calcium ions exert a larger effect on the partially hydrated ester groups at the lipid-water interface, even though the esters rarely bind Ca2+.
+Further investigation is required to experimentally determine the location of Ca2+ ions.
+
+Ca2+ changes the dynamics and structure of the membrane-water interface
+ The ester region remains hydrated
+Previous experiments show that Ca2+ impacts the overall hydration of bilayers. Binder et al. found that dried lipid films hydrate differently in the presence of Ca2+ (71
+), calorimetric data suggests that PS-Ca2+ binding is driven entropically by both Ca2+ and lipid dehydration (21
+), and PS-Ca2+ binding changes the antisymmetric phosphate group’s IR absorption the same way that lipid dehydration does (7
+, 26
+). PS-rich membranes can even form a separate dehydrated phase, forming cylindrical aggregates, with excessive calcium ions (28
+, 44
+).
+The small changes in our measured ester line shapes, both FTIR and 2D IR, indicate that the carboxylate region remains hydrated.
+In fact, hydrogen-bond populations increase slightly with Ca2+, which points toward a slight increase in hydration.
+
 % 2 major head group configurations of PS in 30-70 PC-PS mix -- paper 1

### 2019-03-07 Manuscript/manuscript.tex
Further edits incorportating connection with the current literature. There are still a few point to be done.

--- 2019-03-06 Manuscript/manuscript.tex
@@
 %% Generate PDF/A-2u
-%\usepackage[a-2u]{pdfx}
+\usepackage[a-2u]{pdfx}
 
@@
 \usepackage[english]{babel}
-\usepackage{xr}
-\externaldocument{manuscriptSUPPL}
 
@@
 %% Most settings are pre-set by the pdfx package.
-%\hypersetup{unicode}
-%\hypersetup{breaklinks=true}
-%\hypersetup{colorlinks,allcolors=DodgerBlue4}
+\hypersetup{unicode}
+\hypersetup{breaklinks=true}
+\hypersetup{colorlinks,allcolors=DodgerBlue4}
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
 %% The manuscript does not need to include \maketitle, which is 
@@
 \subsection{Validation of membrane properties}
+
+To validate the structures of lipid bilayers, the scattering form factors and C-H bond order parameters
+were calculated and compared against the experimental data as described previously~\cite{botan15,ollila16,melcr18,nmrlipids_proj4}.
+%in simulations without ions were validated against NMR 
+%by calculating the order parameters for the C-H bonds and against \mbox{X-ray} scattering experiments by evaluating 
+%as was done in Refs.~\citenum{catte16, melcr18, nmrlipids_proj4}.
+%NMR order parameters validate the structures sampled by the individual lipid molecules with atomic resolution. 
+%The simulated order parameters were calculated for all C-H bonds in lipid molecules, 
+%while all order paremeters in the head group and only some in the \ce{C-H} groups in tails were resolved experimentally
+%(see next section and Fig.~\ref{simVSexpNOions_POPS}). 
+The order parameters for each C-H bond of individual lipids in the trajectory were first averaged
+over the time. Then the average over lipids was taken and the standard error of the mean of this
+average was used as the error estimate of the mean \cite{botan15,ollila16,nmrlipids_proj4}.
 
@@
 
-The C--H bond order parameters were first averaged over the time for individual lipids and
-then the standard error of the mean of the average over lipids was used as the error estimate
-of the mean \cite{botan15,ollila16,nmrlipids_proj4}.
-
 \begin{table}[tbp]
@@
 
-The form factors from scattering and C-H bond order parameters from NMR experiments are good measures to validate
-the lipid bilayer structure in MD simulations against experiments because they
-are related to the bilayer dimensions (area per molecule and thickness) and
-conformational fluctuations of individual lipids, respectively, and 
-can be directly calculated from MD simulations~\cite{ollila16}.
-The X-ray scattering form factors and C-H bond order parameters in
-the headgroup and glycerol backbone region of POPS lipid bilayer are already available in the literature \cite{pan14,nmrlipids_proj4}.
-Here, we extend the available experimental dataset of POPS lipids by measuring the C--H bond
-order parameters also from the acyl chain region using R-PDLF experiments as described previously \cite{ferreira13}.
-We first determined the order parameters of each peak corresponding different C--H bonds detected
-in the first dimension (Fig. \ref{R-PDLF}) from the dipolar slices in the indirect dimension (Fig. \ref{R-PDLFslices}).
-Because the exact assignment of these peaks was not possible, we assigned the
-acyl chain order parameters to mimic the profile predicted by the MD simulations (Fig. \ref{simVSexpNOions_POPS}).
-\todo{Tiago has to check and update this and figures in SI.}
-
-To evaluate the effect of ECC on the structural properties of POPS lipid bilayer properties in MD simulations,
-we compared the results from a simulation with original Lipid17 \cite{??} combined with the Dang ion parameters \cite{??}
-with the results from ECC-POPS simulation and experiments (Fig. \ref{simVSexpNOions_POPS}).
-with sodium counterions are prevously publihed
-(or only counter ions)
+We compared X-ray scattering form factors and NMR order parameters of bilayers
+in pure water without any ions (or only counter ions)
 from simulations and experiments. 
@@
 a non-negligible correction to the interaction between cations and POPC bilayers in Ref.~\cite{melcr18}. 
-
 Here, we demonstrate that the same approach 
@@
 and to mark the improvement after treating electronic polarizability, 
-we show simulation results from the newly developed ECC-lipids and also from Lipid17 \citep{lipid17-future}. 
+we show simulation results from ECC-lipids and also from Lipid17 \citep{lipid17-future}. 
 While the response of the order parameters of PC to \ce{CaCl2} concentrations
@@
 and for ECC-lipids. 
-Although the agreement between ECC-lipids and experiments is not optimal,
+Although the agreement between ECC-lipids and experiments is far from optimal,
 the general trends and behaviour even at highly concentrated solutions of \ce{CaCl2} 
@@
 as the connector of the carbon atoms 
-forming the bond between the group and the $\beta$-carbon of the phospholipid 
-(Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}).
+forming the bond between the group and the $\beta$-carbon of the phospholipid. 
 The mean angle between the \ce{COO^-} vector and the membrane normal
@@
 
-Measurements using infrared spectroscopy 
-combined with simulations with the CHARMM36 model 
-with a correction for overestimated binding of calcium \cite{valentine18}
-reveal similar trends for the populations at various PS moieties
-suggesting that POPS interacts mostly through its \ce{COO^-} group. 
-While this feature agrees nicely with our findings, 
-the balance of the interactions of PS moieties with calcium cations may be suboptimal
-and the interaction strength underestimated in total in those simulations \cite{nmrlipids_proj4}.
-
+Recent simulations reveal similar trends for the populations at various PS moieties. \cite{valentine18, hallock18}
+The results from infrared spectroscopy combined with 
+simulations with the CHARMM36 model 
+including a correction for overestimated binding of calcium \cite{valentine18}
+suggest that POPS interacts mostly through it \ce{COO^-} group. 
 In contrast, the combination of NMR experiments 
-and MD simulations with the same model, CHARMM36, but without the correction for calcium binding
+and MD simulations with the same model but without the correction 
 also reveal long-lived interactions of \ce{Ca^{2+}} with \ce{PO4} groups. \cite{hallock18}
-Although the ratios of the life times of coordination follow the same trends as in our work,
-the lengths of simulation trajectories are insufficient 
-to yield converged results in this aspect (see SI and Refs.~\cite{catte16, melcr18, nmrlipids_proj4}). 
-Moreover, the content of POPS in the bilayers in that work is
+The content of POPS in the bilayers in that work is, however, 
 much higher above the threshold after which 
 \ce{Ca^{2+}} cations induce lipid clustering. \cite{hauser77,kurland79,hauser85,feigenson86,mattai89,roux90,roux91}
-At such a high content of negatively charged lipids,
-PS head groups are found to form even four-membered clusters \cite{hallock18}. 
+At such highly charged membranes,
+POPS lipids are found to form even four-membered clusters \cite{hallock18}. 
 Our simulations are, however, at different conditions,
@@
 
-\todo{ THE REMAINING PIECES TO BE DISCUSSED  --  consider wheter the dynamics slowdown or the dihedral distributions could make good contributions to this paper. }
+THE REMAINING PIECES TO BE DISCUSSED:
 
@@
 
-\section{Conclusions}
-
-\todo{SAMULI: I have putted here two preliminary sentences
-JOSEF: I can write conclusions, but I would prefer to wait with that after your comments and edits. }
-Applying ECC to POPS lipids improves cation binding affinity to membranes
-mixed with POPC and also the response of PS headgroup to the bound cations.
-However, further improvement of the parameters is needed to fully reproduce the
-details of calcium binding to PS headgroups.
-
-\listoftodos
-
-\section{SI}
-
-Timescales associated with the binding of calcium cations from solution to the membrane
-are plotted for each binding event as a histogram in Fig.~\ref{fig:hist_residence_times}. 
-Using these plots, we can estimate that 90\% of the residence times of any calcium cation 
-will be lower than $60\,\mathrm{ns}$ for pure POPC neutral bilayer 
-and shorter than $200\,\mathrm{ns}$ for the mixed 5\,PC:1\,PS negatively charged bilayer. 
-The longest observed residence times in the simulations were $141\,\mathrm{ns}$ for the neutral membrane 
-and $485\,\mathrm{ns}$ for the negatively charged membrane. 
-Both estimates of the residence times come from simulations with comparable concentrations of around $250\mathrm{mM}$;
-the simulation with the neutral membrane has a bulk concentration of calcium $C_{ion} = 280\mathrm{mM}$, 
-whereas the simulation with the negatively charged membrane has a bulk concentration of calcium $C_{ion} = 240\mathrm{mM}$.
+%\todo{Stationary distribution: Make a figure documenting the populations of bound \ce{Ca^{2+}} cations (like I have in the presentation) that would accompany Table \ref{tab:Ca_binding_PCPS}. 
+%This will roughly correspond to the PC stoichiometry plot \ref{fig:cacl_complexes}. }
 
@@
 
+\section{Conclusions}
+
+\todo{SAMULI: I have putted here two preliminary sentences}
+Applying ECC to POPS lipids improves cation binding affinity to membranes
+mixed with POPC and also the response of PS headgroup to the bound cations.
+However, further improvement of the parameters is needed to fully reproduce the
+details of calcium binding to PS headgroups.
+
+\listoftodos
+
+\section{SI}
+
+Timescales associated with the binding of calcium cations from solution to the membrane
+are plotted for each binding event as a histogram in Fig.~\ref{fig:hist_residence_times}. 
+Using these plots, we can estimate that 90\% of the residence times of any calcium cation 
+will be lower than $60\,\mathrm{ns}$ for pure POPC neutral bilayer 
+and shorter than $200\,\mathrm{ns}$ for the mixed 5\,PC:1\,PS negatively charged bilayer. 
+The longest observed residence times in the simulations were $141\,\mathrm{ns}$ for the neutral membrane 
+and $485\,\mathrm{ns}$ for the negatively charged membrane. 
+Both estimates of the residence times come from simulations with comparable concentrations of around $250\mathrm{mM}$;
+the simulation with the neutral membrane has a bulk concentration of calcium $C_{ion} = 280\mathrm{mM}$, 
+whereas the simulation with the negatively charged membrane has a bulk concentration of calcium $C_{ion} = 240\mathrm{mM}$.
+
 In summary, the results from ECC-lipids suggest 

### 2019-03-08 Manuscript/manuscript.tex
small corrections to the comments on the works by valentine and hallock

@@
 
-Recent simulations reveal similar trends for the populations at various PS moieties. \cite{valentine18, hallock18}
-The results from infrared spectroscopy combined with 
-simulations with the CHARMM36 model 
-including a correction for overestimated binding of calcium \cite{valentine18}
-suggest that POPS interacts mostly through it \ce{COO^-} group. 
+Measurements using infrared spectroscopy 
+combined with simulations with the CHARMM36 model 
+with a correction for overestimated binding of calcium \cite{valentine18}
+reveal similar trends for the populations at various PS moieties
+suggesting that POPS interacts mostly through its \ce{COO^-} group. 
+While this feature agrees nicely with our findings, 
+the balance of the interactions of PS moieties with calcium cations may be suboptimal
+and the interaction strength underestimated in total \cite{nmrlipids_proj4}.
+
 In contrast, the combination of NMR experiments 
-and MD simulations with the same model but without the correction 
+and MD simulations with the same model, CHARMM36, but without the correction for calcium binding
 also reveal long-lived interactions of \ce{Ca^{2+}} with \ce{PO4} groups. \cite{hallock18}
-The content of POPS in the bilayers in that work is, however, 
+Although the ratios of the life times of coordination follow the same trends as in our work,
+the lengths of simulation trajectories are insufficient 
+to yield converged results in this aspect. \cite{catte16, melcr18, nmrlipids_proj4}
+Moreover, the content of POPS in the bilayers in that work is
 much higher above the threshold after which 
 \ce{Ca^{2+}} cations induce lipid clustering. \cite{hauser77,kurland79,hauser85,feigenson86,mattai89,roux90,roux91}
-At such highly charged membranes,
-POPS lipids are found to form even four-membered clusters \cite{hallock18}. 
+At such a high content of negatively charged lipids,
+PS head groups are found to form even four-membered clusters \cite{hallock18}. 
 Our simulations are, however, at different conditions,

### 2019-03-08 Manuscript/manuscript.tex
Update text

--- 2019-03-07 Manuscript/manuscript.tex
@@
 
+The form factors from scattering and C-H bond order parameters from NMR experiments are good measures to validate
+the lipid bilayer structure in MD simulations against experiments because they
+are related to the bilayer dimensions (area per molecule and thickness) and
+conformational fluctuations of individual lipids, respectively, and 
+can be directly calculated from MD simulations~\cite{ollila16}.
+The X-ray scattering form factors and C-H bond order parameters in
+the headgroup and glycerol backbone region of POPS lipid bilayer are already available in the literature \cite{pan14,nmrlipids_proj4}.
+Here, we extend the available experimental dataset of POPS lipids by measuring the C--H bond
+order parameters also from the acyl chain region using the R-PDLF experiments as described previously \cite{ferreira13}.
+We first determined the order parameters of each peak corresponding different C--H bonds detected
+in the first dimension (Fig. \ref{R-PDLF}) from the dipolar slices in the indirect dimension (Fig. \ref{R-PDLFslices}).
+Because the exact assignment of these peaks was not possible, we assigned the
+acyl chain order parameters to mimic the profile predicted by the MD simulations (Fig. \ref{simVSexpNOions_POPS}).
+\todo{Tiago has to check and update this and figures in SI.}
 \begin{figure}[tb!] 
@@
   }  
-\todo{J: flip the chemical structure so that it agrees with the left-to-right reading of the plot.}
+  \todo{J: flip the chemical structure so that it agrees with the left-to-right reading of the plot.} \\
+  \todo{The experimental order parameters from sn-1 chain are missing from the figure.} \\
+  \todo{We should use the same color coding for the models as in ECC-POPC paper, now it is opposite.} \\
+  \todo{Inverted axis and positive value as label would be more clear for the order parameter figure.}
 \end{figure}
+
+To evaluate the effect of ECC on the structural properties of POPS lipid bilayer,
+we compared the X-ray scattering form factors, C--H bond order parameters and area per molecules
+between simulation with original Lipid17 \cite{??} combined with the Dang ion parameters \cite{??},
+ECC-POPS simulation and experiments (Fig. \ref{simVSexpNOions_POPS} and table \ref{tab:apls}).
+The larger acyl chain order parameters, deviation in the form factor and smaller area per molecule 
+in the Lipid17 simulation with Dang ions when compared with experiments indicates that this simulation
+predicts too dense lipid bilayer. In previous study \cite{nmrlipids_proj4}, Lipid17 simulated with {\AA}qvist ions~\cite{aqvist90}
+gave larger area per molecule (57 \AA$^2$), probably due to the weaker counterion binding affinity.
+The Dang ions are used to here to avoid the artificial pairing of ions and condensation at larger salt
+concentrations observed with {\AA}qvist \cite{kohagen16, chen07, nmrlipids_proj4} when analysing the effect of ECC. 
+The scattering form factors and acyl chain order parameters from ECC-POPS are in good agreement with
+experiments (see Fig.~\ref{simVSexpNOions_POPS}), indicating the the bilayer dimensions and acyl chain
+conformations are well described by the force field. The area per lipid from ECC-POPS simulation is slightly
+smaller than the value reported from SDP model \cite{kucerka14} (table \ref{tab:apls}), but the values
+are within each others error bars. 
+%In order to obtain the area per lipid from experiments, modeling is used on top of the scattering factors \citep{kucerka14}. 
+%In ocntrast, this property is easily extracted directly from simulations with flat phospholipid bilayers. 
+%We compare the values from experiments and simulations in Table~\ref{tab:apls}. 
+%While the agreement between the scattering form factors from the simulation of a pure POPS bilayer and experiment 
+%are excellent (Fig.~\ref{simVSexpNOions_POPS}), there is a non-negligible difference between the values of the area per lipid in Table~\ref{tab:apls}. 
+%Since both values are derived from the scattering form factors through modeling of the electron density of the bilayer,
+%we cannot decide, which of the values is more reliable.
 
@@
     Lipid17/Dang              & 53.5$\pm$ 1.7   \\ 
-    Lipid17/ff99              & 57.9$\pm$ 1.7   \\ 
-    \hline 
+%    Lipid17/ff99              & 57.9$\pm$ 1.7   \\ 
+%    \hline 
     ECC-POPS/ECC-ions         & 59.8$\pm$ 1.8   \\ 
-    \hline 
-    experiment (SDP model) \citep{kucerka14} & 62.3  \\ 
-    \hline 
-  \end{tabular} 
+%    \hline 
+    experiment (SDP model) \citep{kucerka14} & 62.7$\pm$ 1.3  \\ 
+%    \hline 
+  \end{tabular} \\
+  \todo{I removed the value from Lipid17/ff99 simulation because this was not listed in Table I
+    and I think that it is more clear if we refer only to the NMRlipids IV result in the text.
+    I also changed the experimental value to correspond the value in the cited paper and added the error bar.}
 \end{table}
-
-We compared X-ray scattering form factors and NMR order parameters of bilayers
-in pure water without any ions (or only counter ions)
-from simulations and experiments. 
-The experimental X-ray scattering form factors 
-of a bilayer are well reproduced by ECC-POPS, 
-while for Lipid17/Dang they are in a slight disagreement
-(see Fig.~\ref{simVSexpNOions_POPS}). 
-The agreement of Lipid17 can be improved by using different parametes for ions \cite{aqvist90},
-which, however, bring other kinds of artifacts with them 
-(e.g., incorrect pairing of ions, condensation at larger salt concentrations \cite{kohagen16, chen07, nmrlipids_proj4}).
-
-%The area per lipid is often used as a relatively simple structural parameter reporting on the bilayer properties and the packing of lipids. 
-In order to obtain the area per lipid from experiments, modeling is used on top of the scattering factors \citep{kucerka14}. 
-In ocntrast, this property is easily extracted directly from simulations with flat phospholipid bilayers. 
-We compare the values from experiments and simulations in Table~\ref{tab:apls}.
 
@@
 
-While the agreement between the scattering form factors 
-from the simulation of a pure POPS bilayer and experiment 
-are excellent (Fig.~\ref{simVSexpNOions_POPS}),
-there is a non-negligible difference between the values of the area per lipid in Table~\ref{tab:apls}. 
-Since both values are derived from the scattering form factors through modeling of the electron density of the bilayer,
-we cannot decide, which of the values is more reliable. 
-In general, we can conclude that ECC-lipids
-%the presented lipid models with ECC-lipids 
-reproduce the experimental structural parameters of the lipid bilayers 
-with a comparable accuracy to existing state-of-the-art lipid models~\citep{botan15, ollila16, Pluhackova2016}.
-
 \begin{figure}[tb!] 
@@
 
-The head group and acyl chain order parameters within ECC-lipids
-are in general in a good agreement with the experimental values 
-as shown in Fig. \ref{simVSexpNOions_POPS}. 
-The acyl chain order parameters in particular agree with the experimental data within error bars.
-The order parameters of the head groups are at an accuracy comparable to 
-other currently available classical models of lipids \citep{botan15, catte16, Pluhackova2016, nmrlipids_proj4}. 
-In particular, ECC-POPS is closer to the experimental order parameters 
-than Lipid17 with either Dang \cite{smith94,chang1999,dang2006} 
-or ff99 ions \citet{aqvist90}.
+Despite the success in simulating bilayer dimensions and acyl chain conformations,
+lipid force fields have typically problems in capturing the correct glycerol backbone and
+headgroup order parameters and conformations \cite{botan15,ollila16,nmrlipids_proj4}.
+This is the case also for Lipid17 POPS simulations, where the headgroup and glycerol backbone
+order parameters are quite far from experimental values with both Dang (Fig. \ref{simVSexpNOions_POPS})
+and {\AA}qvist \cite{nmrlipids_proj4} ions. The values from ECC-POPS simulation are closer,
+but not in full agreement with experiments (Fig. \ref{simVSexpNOions_POPS}).
+Especially, the experimentally observed more negative values of the $\beta$-carbon
+and the larger forking (difference of order parameters between hydrogens attached to the
+same carbon) of the $\alpha$-carbon in POPS than in POPC \cite{nmrlipids_proj4}
+are not reproduced by neither of the models here. These differences are previously related
+to a more rigid structure of PS headgroup and the stiffness of the C$_\alpha$-C$_\beta$-C$_\gamma$-O$_\gamma$
+dihedral, which is \todo{the dihedral should be plotted and sentence finished.}
 
@@
 %For POPC in pure water, the order parameter $\beta$ agrees well with the experiment, while the order parameter $\alpha$ is somewhat lower.
-The head group order parameters of PS are a bit more complicated 
-compared to PC as the order parameter $\alpha$ exhibits a notable forking 
-(i.e., the hydrogen atoms in the group give rise to different order parameters, see Fig.~\ref{simVSexpNOions_POPS}).
-One of the order parameters of ECC-POPS, $\alpha_1$, agrees well with the experiment, 
-while the other, $\alpha_2$, adopts a higher value underestimating the experimentally reported forking. 
-There is only one order parameter $\beta$ in POPS, 
-which has a higher value closer to zero in the ECC-lipids model than in experiment. 
-Such a feature suggests that the model likely overestimates the orientational freedom of the PS head group.
+
+In conclusion, the ECC-POPS model reproduces the lipid bilayer dimensions and
+acyl chain conformations in good agreement with experiments, but has room for
+improvement in the glycerol backbone and headgroup region. Therefore,
+its structural accuracy is in the same level as in other state-of-the-art
+lipid force fields \cite{botan15, ollila16, Pluhackova2016, nmrlipids_proj4}.
 
@@
 
-It has been pointed out in previous experimental works
-that the conformational flexibility of PS head groups 
-is different from other lipids \cite{browning80}. 
-While the static shielding tensor of PS is not substantially different from PC or PE, 
-the $^{31}$P~$T_1$ relaxation times are significantly shorter for PS. 
-It is suggested that the PS head group is more rigid compared to the other phospholipids, 
-which is also corroborated by a larger chemical shielding anisotropy of PS \cite{browning80}.
-
-We have investigated this effect using the mean orientations of P-N and \ce{COO^-} vectors 
-(defined and plotted in Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}). 
-While the mean angle between the P-N vector of PC head group and the membrane normal
-decreased by 11$^\circ$ from 64$^\circ$ to 55$^\circ$
- at 620~mM buffer concentration of \ce{CaCl2},
-the mean orientation of the P-N vector in the PS head group has 
-decreased only by 3$^\circ$ from 67$^\circ$ to 64$^\circ$ 
-in the same concentration range.
-
-Such a largely diminished flexibility 
-may arise from the binding of calcium to the \ce{COO^-} group in PS. 
-Whe have measured the mean orientation of the \ce{COO^-} group 
-as the connector of the carbon atoms 
-forming the bond between the group and the $\beta$-carbon of the phospholipid. 
-The mean angle between the \ce{COO^-} vector and the membrane normal
-changes by $11^\circ$ from $62^\circ$ to $73^\circ$ (more towards the membrane interior)
- at 620~mM buffer concentration of \ce{CaCl2}.
-The conformational flexibility of the PS head group is hence restrained 
-by the binding of calcium cations to the \ce{COO^-} group, 
-which reorients towards the membrane interior and to the phosphate region, 
-where calcium cations bind predominantly.
-
 \begin{figure}[htbp!] 
@@
 
-The binding of calcium cations to the phospholipid bilayer is shown 
-in the density profiles of \ce{Ca^{2+}}, \ce{Na^+} counterions 
-and also \ce{Cl^-} ions in Fig.~\ref{fig:cacl-dens_PCPS}. 
-Clearly, after accounting for the electronic polarizability, 
-the population of the cations at the membrane interface is significantly diminished. 
-\todo{JOE: The density plot still needs to be changed to show both Lipid17 and ECC-lipids, and to support this argument.}
-Still, the dominant contribution to the binding of calcium comes from the phosphate groups. 
-Interestingly enough,
-although the \ce{COO^-} groups increase the affinity of calcium towards PS compared to PC,
- there is no extra density at the outmost surface of the membrane. 
-This can be understood from the point of the structural rearangements of the \ce{COO^-} group, 
-which take place during the interactions with calcium cations
-described above and in Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}.
-
-\todo{JOE: suggestion -- what is the charge at the membrane interface of memb+counterions -vs- memb+CaCl2 concnetration?
-                         this could relate well to the zeta potential measurements and the Fig 1 in \cite{melcrova16}. }
+%%% Present the density profiles (different concentrations for both POPC and mixed)
+The increase in the amount of bound calcium cations from pure POPC \cite{melcr18}
+to the mixed negatively charged bilayer containing POPS
+is well demonstrated using the relative surface excess, $\Gamma ^w _{Ca}$,
+summarized in Table~\ref{tab:binding}. 
+Distributions of \ce{Ca^{2+}}, \ce{Na^+} counterions and also \ce{Cl^-} 
+are plotted in Fig.~\ref{fig:cacl-dens_PCPS} for the negatively charged bilayer. 
+In contrast to \ce{KCl} or added concentrations of \ce{NaCl}, 
+\ce{Na^+} counterions are substituted with \ce{Ca^{2+}} even at low concentrations of \ce{CaCl2}. 
+The increasing concentration of \ce{CaCl2} and, hence, a higher amount of bound \ce{Ca^{2+}}
+also attracts \ce{Cl^-} anions to the bilayer 
+as can be seen from its growing density at the interface.
+
+The relative surface excess of calcium, $\Gamma^{w}_{Ca}$, also reflects
+the higher affinity of the negatively charged PC:PS (5:1) membrane compared to a neutral POPC bilayer; 
+its value for the latter is       $\Gamma^{w}_{Ca} = 0.06\mathrm{nm^{-2}}$, 
+while for the PC:PS mixture it is $\Gamma^{w}_{Ca} = 0.24\mathrm{nm^{-2}}$
+at comparable concentrations $C _{Ca} = 260\pm 30 \mathrm{mM} $.
+
+\todo{SAMULI: This paragraph in unclear. Should be clarified and the changes
+  of headgroup orientations in PC and PS clearly discussed.
+}
+%%% discuss the table with Gammas and follow up with molecular details, where do cations bind? Support with the spatial density figure
+The density profiles of ions suggest that
+the dominant contribution to the binding of \ce{Ca^{2+}} to phospholipid bilayers
+comes from the interactions with the phosphate groups in both POPC and POPS. 
+%The population analysis also suggests 
+%that the calcium cations prefer to reside in the phosphate region of the membrane. 
+%Such a finding is further validated with a more detailed analysis using Markov state modeling (MSM) \citep{Pande_MSM_paper} \todoi{Add papers reviewing MSM, e.g. recent Pande's paper.}. 
+%The set of states of a calcium cation 
+%encompassed all possible combinations of up to three surrounding lipids. 
+%In the case of PC, we used only the phosphate moitety, which forms the dominant contribution for calcium binding.
+%For PS we also distinguished configurations in which calcium interacts with the carboxylate moiety. 
+%All possible combinations of such states were used to build a Markov model at a lag time $25\,\mathrm{ns}$ using pyEMMA code by \citet{pyemma} \todoi{add citation for pyemma}. 
+%The resulting MSM was validated using Chapman-Kolmogorov test \citep{FrankNoe_papers_MSM} \todoi{Add citation for Noe's papers on MSM, especially Chapman Kolmogorov test.}. 
+%The stationary distribution of the states reveals a strong preference of the calcium cations to reside in the phosphate region of the mixed PC-PS membrane. 
+%When interacting with PS, configurations with the phosphate moiety from either PC or PS dominate the population.
+%Moreover, states containing interactions with the carboxylate group in PS 
+%bear higher probability when interacting also with the phosphate group of the same lipid or other lipids. 
+This is also reflected in the shift of the mean orientation 
+of the \ce{COO^-} group from $62^\circ$ to $73^\circ$ (420~mM \ce{CaCl2})
+which was measured as the connector of the carbon atoms 
+forming the bond between the group and the $\beta$-carbon of the phospholipid. 
+The interactions of the carboxylate group in PS with calcium and other phosphate groups
+shed light into the qualitatively different response of the head group order parameters $\alpha$ and $\beta$ in PS compared to PC 
+(see Figs.~\ref{fig:delta_ordPar_CaCl} and~\ref{fig:delta_ordPar_CaCl_PCPS}). 
+We find that the complex response of the head group order parameters of POPS 
+is affected by the conformational changes of the carboxylate group,
+which is attracted more towards the phosphate region, 
+where the calcium cations dominantly bind. 
+In line with the experimental work by \citet{browning80},
+this is also very likely the reason, 
+why the magnitude of the P-N vector change in POPS is diminished compared to POPC, 
+which is not restrained by an additional cation binding group like \ce{COO^-} in POPS.
+
+Further details about the interactions of \ce{Ca^{2+}} with various moieties in POPC or POPS
+were obtained by counting contacts between the cations and the oxygen atoms of the lipids
+similarly as was done in \cite{melcr18}. 
+The threshold for counting a contact was set to $0.3\,\mathrm{nm}$, 
+which encompasses the first peak of the radial distribution function between the cations and the oxygen atoms of the lipids.
 
@@
 \centering
-  \caption{Bulk concentrations, $C _{Ca}$, and buffer concentrations, $C' _{Ca}$, of Ca$^{2+}$;
+  \caption{Bulk concentrations, $C _{Ca}$, and molar fractions, $C' _{Ca}$, of Ca$^{2+}$;
            relative surface excess of calcium with respect to water ($\Gamma_{Ca}^{\rm water}$); 
@@
 
-% Gamma and higher affinity of PS to CA2+
-The higher affinity of calcium to PS head groups compared to PC
-is well demonstrated using the relative surface excess, $\Gamma ^w _{Ca}$,
-summarized in Table~\ref{tab:binding}. 
-While $\Gamma^{w}_{Ca}$ at a neutral POPC bilayer
-is       $\Gamma^{w}_{Ca} = 0.06\mathrm{nm^{-2}}$ \cite{melcr18}, 
-its value for the PC:PS (5:1) bilayer is $\Gamma^{w}_{Ca} = 0.24\mathrm{nm^{-2}}$
-at comparable buffer concentrations  $350 \, \mathrm{mM}$ resp. $400\, \mathrm{mM}$.
-
-% Description of no. conctacts counting
-Such a significantly increased binding also has an impact 
-on the structure and populations of the bound calcium cations. 
-We have counted contacts between the cations and the oxygen atoms of the lipids
-similarly as was done in \cite{melcr18} 
-to obtain such a detailed information 
-about the interactions of \ce{Ca^{2+}} with various moieties in POPC or POPS. 
-The threshold for counting a contact was set to $0.3\,\mathrm{nm}$, 
-which encompasses the first peak of the radial distribution function between the cations and the oxygen atoms of the lipids.
-
-% Populations from contacts
 The percentages of the populations of membrane-bound calcium cations for various membrane moieties 
@@
 Even though the negatively charged membrane contains only 18\% of POPS, 
-on average 41\% of the total population of bound calcium cations is in contact with PS lipids
+approximately 41\% of the total population of bound calcium cations is in contact with PS lipids
 with 8\% bound only to them. 
-% Where Calcium interacts
-While POPC interacts with the calcium cations almost entirely through its phosphate group 
-in both neutral \cite{melcr18} and negatively charged membranes, 
-the inteaction with POPS lipids happens more through the \ce{COO^-} group than through the \ce{PO4} group. 
-Interactions of \ce{Ca^{2+}} with carbonyl groups are also present for both POPC and POPS,
+This corroborates the intrinsically higher affinity of PS lipids to calcium cations compared to neutral PC lipids. 
+POPC interacts with the calcium cations almost entirely through its phosphate group 
+in both neutral and negatively charged membranes. 
+Interactions of \ce{Ca^{2+}} with carbonyl groups are also present, 
 however, they are always accompanied by interactions with phosphate groups. 
-This is in contrast to 
-simulations with the Berger force field \cite{berger97,mukhopadhyay04}
-combined with fluorescent and vibrational sum frequency spectroscopy,
-which suggest a significant
-calcium binding also to the carbonyls groups \cite{melcrova16}. 
-This feature arises likely from the inaccuracy of the Berger model 
-and its overestimated binding of cations to that region \cite{nmrlipids_proj4}.
-
-Recent simulations reveal similar trends for the populations at various PS moieties. \cite{valentine18, hallock18}
-The results from infrared spectroscopy combined with 
-simulations with the CHARMM36 model 
-including a correction for overestimated binding of calcium \cite{valentine18}
-suggest that POPS interacts mostly through it \ce{COO^-} group. 
-In contrast, the combination of NMR experiments 
-and MD simulations with the same model but without the correction 
-also reveal long-lived interactions of \ce{Ca^{2+}} with \ce{PO4} groups. \cite{hallock18}
-The content of POPS in the bilayers in that work is, however, 
-much higher above the threshold after which 
-\ce{Ca^{2+}} cations induce lipid clustering. \cite{hauser77,kurland79,hauser85,feigenson86,mattai89,roux90,roux91}
-At such highly charged membranes,
-POPS lipids are found to form even four-membered clusters \cite{hallock18}. 
-Our simulations are, however, at different conditions,
-in which the clustering of POPS does not happen, 
-and PS lipids prefer 1:1 ratio with \ce{Ca^{2+}}
-as plotted in Fig.~\ref{fig:cacl_complexes}
-showing relative probabilities of \ce{Ca^{2+}} complexes with a certain number of PC or PS lipids. 
-Compared to a neutral POPC bilayer
-the negatively charged bilayer has its stoichiometry 
-shifted towards complexes with three phospholipids and one calcium. 
-This is obviously due to the presence of POPS, 
-which also contributes to the pre-formed one- and two-membered clusters of POPC. 
-Clusters of four or more lipids were not observed in either membrane.
-
-THE REMAINING PIECES TO BE DISCUSSED:
-
-% Ca2+ induces much slower dynamics of lipids when PS is present,
-% this does not happen with PC
-% uses NBFIX
-From Ref.~\cite{valentine18}://
-To summarize, calcium binding to PS lipids slows the dynamics at the ester position and introduces more heterogeneous molecular environments. 
-The slowdown cannot be attributed to complete ester dehydration or decreased hydrogen bonding to water, so we interpret these results as evidence that the lipid headgroups adopt more rigid but disordered conformations on the molecular level.
-The decrease in dynamics observed experimentally indicates that the network of hydrogen bonds between lipids and water at the polar-nonpolar interface becomes slower, as it was shown by Karathanou and Bondar (34).
-This is bolstered by recent NMR work showing that binding to Ca2+ induces two different rigid conformations in the glycerol backbone of lipid headgroups (18).
-We have investigated the molecular origin of the rigidification and decreased lateral diffusion, finding that calcium binding does not dehydrate the ester region, but even with only 10\% PS lipids, the local dynamics slow around the esters.
-The slower local dynamics provide a molecular explanation for the Ca2+-dependent decreases in fluidity observed on longer length and timescales.
-There is considerable room for further improvement in the force fields for ion-membrane interactions. Particularly, small differences in short-range interactions may determine the local geometries (34
-).
-Most of our conclusions are, however, based on long-range effects (i.e., electric field generated by the ion at significant distances from the carbonyl groups), which are less sensitive to the details of the force field.
-
-Their Figure 5 \cite{hallock18} shows the distributions of O-Ca-Cb-N dihedral angle
-revealing separate populations of the PS head groups. 
-These two structures are also compared in therm of P-Cg distances, 
-which can be measured from NMR REDOR exp. 
-It is concluded that since the two distinct populations in P-Cg distance can be represented with the 
-dihedral distributions, 
-they also likely represent the distinct populations. 
- -> plotting the distribution of O-Ca-Cb-N dihedral angle may be also a good idea for us (SI?)
-     -> NMRlipids 4 in SI \cite{nmrlipids_proj4} plots this information: 
-           - C36 agress with the distribution in \cite{hallock18}
-           - Lipid17 (and Slipids) are almost identical to C36
-           - but the structure of the COO- is highly different (dih angle Ca-Cb-Cg-Og)
+%but a very small negligable amount much below 1\%. 
+\todo{Also COO- vs. phosphate binding should be discussed and the results should be
+  discussed with respect to the literature: https://doi.org/10.1016/j.bpj.2018.08.044 and
+  Melcrova et al. Scientific Reports | 6:38035 | DOI: 10.1038/srep38035 }
+%
+\todo{EDIT: Taken from NMRlipids4 as an interesting information, to be edited.}
+The dehydrated complexes of PS headgroup and calcium ions can also lead to the
+phase separation \cite{hauser77,kurland79,hauser85,feigenson86,mattai89,roux90,roux91}.
+%
+CHARMM36 force field \cite{klauda10,venable13} combined with 2D infrared spectroscopy
+suggests that calcium ions interacts only with the carboxylate group of PS lipids \cite{valentine18}, while
+the same force field without the NBfix parameters together with the NMR chemical shifts and
+REDOR experiments suggests a significant binding affinity also to the phosphate region \cite{hallock18}.
+On the other hand, simulations with the Berger force field \cite{berger97,mukhopadhyay04}
+combined with fluorescent and vibrational sum frequency spectroscopy suggested a significant
+calcium binding also to the carbonyls in the acyl chains \cite{melcrova16}.
 
@@
 \end{figure}
+
+Relative probabilities of \ce{Ca^{2+}} complexes with a certain number of lipids are presented in Fig.~\ref{fig:cacl_complexes}. 
+Calcium cations that are bound only to PC in the mixed bilayer with PS 
+behave similarly as in the pure PC bilayer
+maintaining comparable probabilities for clustering one, two, or even three PC lipids together. 
+In contrast, PS lipids prefer 1:1 ratio with \ce{Ca^{2+}},
+which may also be due to their low molar fraction in the the mixed bilayer. 
+In total, however, the negatively charged membrane has its stoichiometry 
+shifted towards complexes with three phospholipids and one calcium. 
+This is obviously due to the presence of POPS, 
+which also contributes to the pre-formed one- and two-membered clusters of POPC. 
+Clusters of four or more lipids were not observed in either membrane.
+
+%This is also reflected in the increased probability 
+%of a single phospholipid interacting with two \ce{Ca^{2+}} cations 
+%compared to the neutral POPC bilayer, 
+%where such a probability is almost negligable. 
+% Not shown through numbers, however, I have the plots to back it.
 
@@
 
-\section{Conclusions}
-
-\todo{SAMULI: I have putted here two preliminary sentences}
-Applying ECC to POPS lipids improves cation binding affinity to membranes
-mixed with POPC and also the response of PS headgroup to the bound cations.
-However, further improvement of the parameters is needed to fully reproduce the
-details of calcium binding to PS headgroups.
-
-\listoftodos
-
-\section{SI}
-
 Timescales associated with the binding of calcium cations from solution to the membrane
@@
 are required for the negatively charged bilayers.
+
+%In addition to such estimates of the time scales, we used the Markov model on top of the simulation with the negatviely charged mixed bilayer
+%to calculate the time of the mean first passage of calcium from solution to the membrane (and in reverse) resulting in $55\,\mathrm{ns}$  ($165\,\mathrm{ns}$).
+
+%\todo{Fluxes: committor analysis, dominant fluxes (table and figure)}
+%From the spectrum of the transition matrix, we also observe that 
+%the slowest transitions are associated with the binding to two or three phosphate moieties from either PC or PS,
+%which are also among the states with the highest probabilities.
+%The net fluxes of calcium cations from solution to such states also form a large contribution to the total flux ($\approx 45\%$). 
+%\todo{Make a table and a figure of the state probabilities and fluxes to support this statement.}. 
+%Analysis of the possible binding pathways of calcium cations to the negatively charged mixed bilayer 
+%reveal that the cations mostly enter the membrane bound states directly from solution 
+%without complicated transitions at the time scales of the order of the lag time of the Markov state model, $25\,\mathrm{ns}$.
+
+Provide density plots and show, where and how the cations bind to the membrane.
+
+\section{Conclusions}
+
+\todo{SAMULI: I have putted here two preliminary sentences}
+Applying ECC to POPS lipids improves cation binding affinity to membranes
+mixed with POPC and also the response of PS headgroup to the bound cations.
+However, further improvement of the parameters is needed to fully reproduce the
+details of calcium binding to PS headgroups.
+
+\listoftodos
 

### 2019-03-08 Manuscript/manuscript.tex
Update text

--- 2019-03-08 Manuscript/manuscript.tex
@@
 the balance of the interactions of PS moieties with calcium cations may be suboptimal
-and the interaction strength underestimated in total \cite{nmrlipids_proj4}.
+and the interaction strength underestimated in total in those simulations \cite{nmrlipids_proj4}.
 
@@
 the lengths of simulation trajectories are insufficient 
-to yield converged results in this aspect. \cite{catte16, melcr18, nmrlipids_proj4}
+to yield converged results in this aspect (see SI and Refs.~\cite{catte16, melcr18, nmrlipids_proj4}). 
 Moreover, the content of POPS in the bilayers in that work is
@@
 
-THE REMAINING PIECES TO BE DISCUSSED:
+\todo{ THE REMAINING PIECES TO BE DISCUSSED  --  consider wheter the dynamics slowdown or the dihedral distributions could make good contributions to this paper. }
 
@@
 
-%\todo{Stationary distribution: Make a figure documenting the populations of bound \ce{Ca^{2+}} cations (like I have in the presentation) that would accompany Table \ref{tab:Ca_binding_PCPS}. 
-%This will roughly correspond to the PC stoichiometry plot \ref{fig:cacl_complexes}. }
+\section{Conclusions}
+
+\todo{SAMULI: I have putted here two preliminary sentences
+JOSEF: I can write conclusions, but I would prefer to wait with that after your comments and edits. }
+Applying ECC to POPS lipids improves cation binding affinity to membranes
+mixed with POPC and also the response of PS headgroup to the bound cations.
+However, further improvement of the parameters is needed to fully reproduce the
+details of calcium binding to PS headgroups.
+
+\listoftodos
+
+\section{SI}
+
+Timescales associated with the binding of calcium cations from solution to the membrane
+are plotted for each binding event as a histogram in Fig.~\ref{fig:hist_residence_times}. 
+Using these plots, we can estimate that 90\% of the residence times of any calcium cation 
+will be lower than $60\,\mathrm{ns}$ for pure POPC neutral bilayer 
+and shorter than $200\,\mathrm{ns}$ for the mixed 5\,PC:1\,PS negatively charged bilayer. 
+The longest observed residence times in the simulations were $141\,\mathrm{ns}$ for the neutral membrane 
+and $485\,\mathrm{ns}$ for the negatively charged membrane. 
+Both estimates of the residence times come from simulations with comparable concentrations of around $250\mathrm{mM}$;
+the simulation with the neutral membrane has a bulk concentration of calcium $C_{ion} = 280\mathrm{mM}$, 
+whereas the simulation with the negatively charged membrane has a bulk concentration of calcium $C_{ion} = 240\mathrm{mM}$.
 
@@
 
-\section{Conclusions}
-
-\todo{SAMULI: I have putted here two preliminary sentences}
-Applying ECC to POPS lipids improves cation binding affinity to membranes
-mixed with POPC and also the response of PS headgroup to the bound cations.
-However, further improvement of the parameters is needed to fully reproduce the
-details of calcium binding to PS headgroups.
-
-\listoftodos
-
-\section{SI}
-
-Timescales associated with the binding of calcium cations from solution to the membrane
-are plotted for each binding event as a histogram in Fig.~\ref{fig:hist_residence_times}. 
-Using these plots, we can estimate that 90\% of the residence times of any calcium cation 
-will be lower than $60\,\mathrm{ns}$ for pure POPC neutral bilayer 
-and shorter than $200\,\mathrm{ns}$ for the mixed 5\,PC:1\,PS negatively charged bilayer. 
-The longest observed residence times in the simulations were $141\,\mathrm{ns}$ for the neutral membrane 
-and $485\,\mathrm{ns}$ for the negatively charged membrane. 
-Both estimates of the residence times come from simulations with comparable concentrations of around $250\mathrm{mM}$;
-the simulation with the neutral membrane has a bulk concentration of calcium $C_{ion} = 280\mathrm{mM}$, 
-whereas the simulation with the negatively charged membrane has a bulk concentration of calcium $C_{ion} = 240\mathrm{mM}$.
-
 In summary, the results from ECC-lipids suggest 

### 2019-03-08 Manuscript/manuscript.tex

@@
 the balance of the interactions of PS moieties with calcium cations may be suboptimal
-and the interaction strength underestimated in total in those simulations \cite{nmrlipids_proj4}.
+and the interaction strength underestimated in total \cite{nmrlipids_proj4}.
 
@@
 the lengths of simulation trajectories are insufficient 
-to yield converged results in this aspect (see SI and Refs.~\cite{catte16, melcr18, nmrlipids_proj4}). 
+to yield converged results in this aspect. \cite{catte16, melcr18, nmrlipids_proj4}
 Moreover, the content of POPS in the bilayers in that work is

### 2019-03-08 Manuscript/manuscript.tex
Update text

@@
 
-\subsection{Validation of membrane properties}
-
-Cation binding affinity to phosphatidylcholine (PC) lipid bilayers can be evaluated using 
-the order parameters of $\alpha$ and $\beta$ carbons in the lipid headgroup 
-(see Fig.~\ref{fig:chemstruct_pc_ps} for the labeling) \cite{catte16}.
-According to the ''electrometer concept'' \citep{seelig87}, these order parameters decrease
-proportionally to the bound charge and can therefore be used to quantify the cation binding
-affinity to lipid bilayers. The situation is more complicated for lipid bilayers containing
-PS lipids, for which there is no ion-free reference state due to the presence of counterions.
-Moreover, additional ions often lead to phase transitions in such bilayers \cite{nmrlipids_proj4,feigenson86,mattai89,roux91,roux90}
-Here, we overcome this complication by comparing simulations to the experimental data 
-measured from POPC:POPS (5:1) mixtures, 
-which are not as highly negatively charged, 
-and do not ondergo phase transitions with additional ions \cite{roux90} 
-as was also demonstrated in the recent work by the NMRlipids project \cite{nmrlipids_proj4}.
-
-%experimentally quantified
-%by measuring the changes of the head group order parameters in the lipids using solid state NMR.
-%In the case of 
-%such changes are known under the term 
-%It is experimentally observed that the \ce{C-H} bond
-%order parameters of $S^\alpha$ and $S^\beta$ carbons in a PC lipid head group
-%are proportional to the amount of charge, positive or negative, bound per lipid~\citep{seelig87, roux90}.
-%The change of the order parameters measured with varying concentrations of aqueous ions 
-%can be then related to the amount of bound ions.
+%\subsection{Validation of membrane properties}
+% SAMULI: I have dissolved this section into results and discussion
 
@@
 
-\begin{figure}[tbp!] 
-  \centering 
-  \includegraphics[height=\figheight]{../img/ecc_pops/density_profiles_na-k-counterions_wat_phos_compar_purePOPS_ecclipids-lipid17.pdf}
-  \includegraphics[height=\figheight]{../img/ecc_pops/density_profiles_na-k-counterions_wat_phos_compar_5PC-1PS_ecclipids-lipid17.pdf}
-  \caption{\label{fig:POPS-counterions-dens}
-    Number density profiles of \ce{K^{+}}, \ce{Na^{+}} counterions along the membrane normal axis 
-    for the negatively charged bilayers composed of only POPS (left) or POPC:POPS (5:1, right) 
-    using ECC-lipids/ECC-ions (solid lines) and Lipid17/Dang ions (dashed lines).  
-    The density profiles of phosphate groups and water are divided by 4 and 100, respectively.  
-}
-\todo{JOE: add labels with the relative surface excess $\Gamma$, which are promised to be present in the text.}
-\todo{JOE: split this figure into two: separate for POPS (to SI) and PC:PS 5:1 (main text)?}
-\end{figure}
-
-\begin{figure}[tb!] 
-  \centering 
-  \includegraphics[width=\figwidth]{../Fig/lipids_chemfig_POPC_POPS.pdf} 
-  \caption{ \label{fig:chemstruct_pc_ps} 
-            The chemical structure of POPC and POPS 
-            including the labeling of the carbon segments
-            used in the measurements of order parameteres. 
-  }  
-\end{figure}
-
 Despite the success in simulating bilayer dimensions and acyl chain conformations,
@@
 
-\subsection{Binding of counterions to negatively charged PC:PS bilayers and interactions between PC and PS head groups}
-
-According to the electrometer concept,
-the head group order parameters $\alpha$ and $\beta$ of PC lipids
-change proportionally to the amount of bound charge
-in a fixed ratio. \cite{seelig87,scherer87,roux90} 
-This was demonstrated for a bound positive charge in simulations with POPC and 
-cationic surfactants or aquaeous cations employing the ECC-POPC model \cite{melcr18}.
-In this work, we will use the electrometer concept also for a bound negative charge 
-using PS lipids as negatively charged surfactants. \cite{roux90}
+\subsection{Counterion binding affinity and interactions between PC and PS head groups}
+
+Binding affinity of monovalent ions to lipid bilayers depends strongly on
+force field parameters \cite{catte16,melcr18,nmrlipids_proj4},
+which is the case also here as the sodium binding affinity to
+POPS lipid bilayer is stronger in Lipid17 simulation with Dang ions than in
+ECC-POPS simulation (Fig. \ref{fig:POPS-counterions-dens}).
+The different cation binding affinities to lipid bilayers have been previously evaluated
+against experiments using the order parameters of $\alpha$ and $\beta$ carbons in the
+phosphatidylcholine (PC) lipid headgroup (see Fig.~\ref{fig:chemstruct_pc_ps} for the labeling) \cite{catte16,melcr18,nmrlipids_proj4}.
+According to the ''electrometer concept'' \citep{seelig87}, these order parameters decrease
+proportionally to the bound charge and can therefore be used to quantify the cation binding
+affinity to lipid bilayers. Because the response of PS lipid headgroup order parameters to the bound charge
+is less well undestood and addition of cation may cause phase transitions in negatively charged
+bilayers \cite{feigenson86,mattai89,roux91,roux90}, the cation binding affinity to bilayers with
+PS lipids has evaluated measuring the PC headgroup order parameters from POPC:POPS (5:1)
+mixtures\cite{roux90,nmrlipids_proj4}.
+
+The situation is more complicated for lipid bilayers containing
+PS lipids, for which there is no ion-free reference state due to the presence of counterions.
+Moreover, additional ions often lead to phase transitions in such bilayers \cite{nmrlipids_proj4,feigenson86,mattai89,roux91,roux90}
 
@@
 for both PS and mixed PC:PS (5:1) bilayers.
+
+%According to the electrometer concept,
+%the head group order parameters $\alpha$ and $\beta$ of PC lipids
+%change proportionally to the amount of bound charge
+%in a fixed ratio. \cite{seelig87,scherer87,roux90} 
+%This was demonstrated for a bound positive charge in simulations with POPC and 
+%cationic surfactants or aquaeous cations employing the ECC-POPC model \cite{melcr18}.
+%In this work, we will use the electrometer concept also for a bound negative charge 
+%using PS lipids as negatively charged surfactants. \cite{roux90}
+
+%is the
+%is not generally agreed on between simulations 
+%\citep{bockmann03,sachs04,berkowitz06,cordomi09}
+%and experiments 
+%\citep{cevc90,tocanne90,hauser76,herbette84,uhrikova08}.
+%We attempted to find a model, 
+%which would successfully interpret the experiments on binding of \ce{Na^+}
+%to neutral and negatively charged phospholipid bilayers \citep{akutsu81, roux90} 
+%in our works \citep{catte16, nmrlipids_proj4}.
+%After testing the currently available force fields for phospholipids,
+%we concluded
+%that the interactions are generally overestimated in magnitude in almost all models 
+%but Lipid14 (PC) \citep{dickson14}, resp. Lipid17 (PS) \citep{lipid17-future}. 
+%This model yields a semi-quantitative agreement with the experimentally measured small changes of the order parameters \cite{catte16}
+%when used with the model of ions by \citet{aqvist90}. 
+%However, when used with a more accurate model of ions by \citet{Pluharova2014, martinek17},
+%the model overestimates the binding affinity of \ce{Na^+}
+%measured with lipid electrometer concept. \citep{melcr18}
+%In total, these results suggest that improvements 
+%in the lipid parameters are required for more accurate interactions even with monovalent cations. 
+%\citep{catte16, melcr18, nmrlipids_proj4}
+
+\begin{figure}[tbp!] 
+  \centering 
+  \includegraphics[height=\figheight]{../img/ecc_pops/density_profiles_na-k-counterions_wat_phos_compar_purePOPS_ecclipids-lipid17.pdf}
+  \includegraphics[height=\figheight]{../img/ecc_pops/density_profiles_na-k-counterions_wat_phos_compar_5PC-1PS_ecclipids-lipid17.pdf}
+  \caption{\label{fig:POPS-counterions-dens}
+    Number density profiles of \ce{K^{+}}, \ce{Na^{+}} counterions along the membrane normal axis 
+    for the negatively charged bilayers composed of only POPS (left) or POPC:POPS (5:1, right) 
+    using ECC-lipids/ECC-ions (solid lines) and Lipid17/Dang ions (dashed lines).  
+    The density profiles of phosphate groups and water are divided by 4 and 100, respectively.  
+}
+\todo{JOE: add labels with the relative surface excess $\Gamma$, which are promised to be present in the text.} \\
+\todo{JOE: split this figure into two: separate for POPS (to SI) and PC:PS 5:1 (main text)? SAMULI: Should be POPS in main text, mixture in SI.}
+\end{figure}
+
+\begin{figure}[tb!] 
+  \centering 
+  \includegraphics[width=\figwidth]{../Fig/lipids_chemfig_POPC_POPS.pdf} 
+  \caption{ \label{fig:chemstruct_pc_ps} 
+            The chemical structure of POPC and POPS 
+            including the labeling of the carbon segments
+            used in the measurements of order parameteres. 
+          }
+          \todo{The POPS structure and labeling is already in figure \ref{simVSexpNOions_POPS}.
+            The POPC structure and labeling could be added on top of the new version of figure \ref{fig:delta_ordPar_NaCl_PC-PS_mix} with only PC,
+          where PC results occur in the first time, and this figure could be removed.}
+\end{figure}
+
+Generally improved behaviour 
+of the POPC and POPS head group order parameters 
+with \ce{NaCl} or \ce{KCl} concentrations 
+was achieved through the combination of models 
+ECC-lipids and ECC-ions \citep{martinek17, kohagen16, Pluharova2014}. 
+Simulations with these models 
+reveal a good agreement with the NMR experiments 
+for both neutral and negatively charged membranes.
+The results are plotted for \ce{NaCl} in Fig. \ref{fig:delta_ordPar_NaCl_PCPS}, 
+and for \ce{KCl} in Fig. \ref{fig:delta_ordPar_KCl_PCPS}.
 
@@
 
-\subsection{Molecular interaction and binding affinities of K$^+$ and Na$^+$ cations to mixed POPC:POPS (5:1) membrane} 
-\label{sec:K_Na_added}
+%\subsection{Molecular interaction and binding affinities of K$^+$ and Na$^+$ cations to mixed POPC:POPS (5:1) membrane} 
+%\label{sec:K_Na_added}
 
@@
 \end{figure}
-
-% This§: 
-% NA/K binding status, 
-% how we tried finding a model (NMRlipids4), and 
-% how it was futile
-% 
-The binding of \ce{Na^+} cations to phospholipid bilayers 
-is not generally agreed on between simulations 
-\citep{bockmann03,sachs04,berkowitz06,cordomi09}
-and experiments 
-\citep{cevc90,tocanne90,hauser76,herbette84,uhrikova08}.
-We attempted to find a model, 
-which would successfully interpret the experiments on binding of \ce{Na^+}
-to neutral and negatively charged phospholipid bilayers \citep{akutsu81, roux90} 
-in our works \citep{catte16, nmrlipids_proj4}.
-After testing the currently available force fields for phospholipids,
-we concluded
-that the interactions are generally overestimated in magnitude in almost all models 
-but Lipid14 (PC) \citep{dickson14}, resp. Lipid17 (PS) \citep{lipid17-future}. 
-This model yields a semi-quantitative agreement with the experimentally measured small changes of the order parameters \cite{catte16}
-when used with the model of ions by \citet{aqvist90}. 
-However, when used with a more accurate model of ions by \citet{Pluharova2014, martinek17},
-the model overestimates the binding affinity of \ce{Na^+}
-measured with lipid electrometer concept. \citep{melcr18}
-In total, these results suggest that improvements 
-in the lipid parameters are required for more accurate interactions even with monovalent cations. 
-\citep{catte16, melcr18, nmrlipids_proj4}
-
-% This§: 
-% we have developed a model that can describe the NMR exp. using ECC
-% 
-Generally improved behaviour 
-of the POPC and POPS head group order parameters 
-with \ce{NaCl} or \ce{KCl} concentrations 
-was achieved through the combination of models 
-ECC-lipids and ECC-ions \citep{martinek17, kohagen16, Pluharova2014}. 
-Simulations with these models 
-reveal a good agreement with the NMR experiments 
-for both neutral and negatively charged membranes.
-The results are plotted for \ce{NaCl} in Fig. \ref{fig:delta_ordPar_NaCl_PCPS}, 
-and for \ce{KCl} in Fig. \ref{fig:delta_ordPar_KCl_PCPS}.
 

### 2019-03-08 Manuscript/manuscript.tex
Went through the Ca2+ binding section I left a few things undone for Samuli to consider, wheter they would form a good contribution to the work. There still are Figures to be corrected.

--- 2019-03-08 Manuscript/manuscript.tex
@@
 
-The form factors from scattering and C-H bond order parameters from NMR experiments are good measures to validate
-the lipid bilayer structure in MD simulations against experiments because they
-are related to the bilayer dimensions (area per molecule and thickness) and
-conformational fluctuations of individual lipids, respectively, and 
-can be directly calculated from MD simulations~\cite{ollila16}.
-The X-ray scattering form factors and C-H bond order parameters in
-the headgroup and glycerol backbone region of POPS lipid bilayer are already available in the literature \cite{pan14,nmrlipids_proj4}.
-Here, we extend the available experimental dataset of POPS lipids by measuring the C--H bond
-order parameters also from the acyl chain region using the R-PDLF experiments as described previously \cite{ferreira13}.
-We first determined the order parameters of each peak corresponding different C--H bonds detected
-in the first dimension (Fig. \ref{R-PDLF}) from the dipolar slices in the indirect dimension (Fig. \ref{R-PDLFslices}).
-Because the exact assignment of these peaks was not possible, we assigned the
-acyl chain order parameters to mimic the profile predicted by the MD simulations (Fig. \ref{simVSexpNOions_POPS}).
-\todo{Tiago has to check and update this and figures in SI.}
 \begin{figure}[tb!] 
@@
   }  
-  \todo{J: flip the chemical structure so that it agrees with the left-to-right reading of the plot.} \\
-  \todo{The experimental order parameters from sn-1 chain are missing from the figure.} \\
-  \todo{We should use the same color coding for the models as in ECC-POPC paper, now it is opposite.} \\
-  \todo{Inverted axis and positive value as label would be more clear for the order parameter figure.}
+\todo{J: flip the chemical structure so that it agrees with the left-to-right reading of the plot.}
 \end{figure}
-
-To evaluate the effect of ECC on the structural properties of POPS lipid bilayer,
-we compared the X-ray scattering form factors, C--H bond order parameters and area per molecules
-between simulation with original Lipid17 \cite{??} combined with the Dang ion parameters \cite{??},
-ECC-POPS simulation and experiments (Fig. \ref{simVSexpNOions_POPS} and table \ref{tab:apls}).
-The larger acyl chain order parameters, deviation in the form factor and smaller area per molecule 
-in the Lipid17 simulation with Dang ions when compared with experiments indicates that this simulation
-predicts too dense lipid bilayer. In previous study \cite{nmrlipids_proj4}, Lipid17 simulated with {\AA}qvist ions~\cite{aqvist90}
-gave larger area per molecule (57 \AA$^2$), probably due to the weaker counterion binding affinity.
-The Dang ions are used to here to avoid the artificial pairing of ions and condensation at larger salt
-concentrations observed with {\AA}qvist \cite{kohagen16, chen07, nmrlipids_proj4} when analysing the effect of ECC. 
-The scattering form factors and acyl chain order parameters from ECC-POPS are in good agreement with
-experiments (see Fig.~\ref{simVSexpNOions_POPS}), indicating the the bilayer dimensions and acyl chain
-conformations are well described by the force field. The area per lipid from ECC-POPS simulation is slightly
-smaller than the value reported from SDP model \cite{kucerka14} (table \ref{tab:apls}), but the values
-are within each others error bars. 
-%In order to obtain the area per lipid from experiments, modeling is used on top of the scattering factors \citep{kucerka14}. 
-%In ocntrast, this property is easily extracted directly from simulations with flat phospholipid bilayers. 
-%We compare the values from experiments and simulations in Table~\ref{tab:apls}. 
-%While the agreement between the scattering form factors from the simulation of a pure POPS bilayer and experiment 
-%are excellent (Fig.~\ref{simVSexpNOions_POPS}), there is a non-negligible difference between the values of the area per lipid in Table~\ref{tab:apls}. 
-%Since both values are derived from the scattering form factors through modeling of the electron density of the bilayer,
-%we cannot decide, which of the values is more reliable.
 
@@
     Lipid17/Dang              & 53.5$\pm$ 1.7   \\ 
-%    Lipid17/ff99              & 57.9$\pm$ 1.7   \\ 
-%    \hline 
+    Lipid17/ff99              & 57.9$\pm$ 1.7   \\ 
+    \hline 
     ECC-POPS/ECC-ions         & 59.8$\pm$ 1.8   \\ 
-%    \hline 
-    experiment (SDP model) \citep{kucerka14} & 62.7$\pm$ 1.3  \\ 
-%    \hline 
-  \end{tabular} \\
-  \todo{I removed the value from Lipid17/ff99 simulation because this was not listed in Table I
-    and I think that it is more clear if we refer only to the NMRlipids IV result in the text.
-    I also changed the experimental value to correspond the value in the cited paper and added the error bar.}
+    \hline 
+    experiment (SDP model) \citep{kucerka14} & 62.3  \\ 
+    \hline 
+  \end{tabular} 
 \end{table}
+
+We compared X-ray scattering form factors and NMR order parameters of bilayers
+in pure water without any ions (or only counter ions)
+from simulations and experiments. 
+The experimental X-ray scattering form factors 
+of a bilayer are well reproduced by ECC-POPS, 
+while for Lipid17/Dang they are in a slight disagreement
+(see Fig.~\ref{simVSexpNOions_POPS}). 
+The agreement of Lipid17 can be improved by using different parametes for ions \cite{aqvist90},
+which, however, bring other kinds of artifacts with them 
+(e.g., incorrect pairing of ions, condensation at larger salt concentrations \cite{kohagen16, chen07, nmrlipids_proj4}).
+
+%The area per lipid is often used as a relatively simple structural parameter reporting on the bilayer properties and the packing of lipids. 
+In order to obtain the area per lipid from experiments, modeling is used on top of the scattering factors \citep{kucerka14}. 
+In ocntrast, this property is easily extracted directly from simulations with flat phospholipid bilayers. 
+We compare the values from experiments and simulations in Table~\ref{tab:apls}.
 
@@
 
+While the agreement between the scattering form factors 
+from the simulation of a pure POPS bilayer and experiment 
+are excellent (Fig.~\ref{simVSexpNOions_POPS}),
+there is a non-negligible difference between the values of the area per lipid in Table~\ref{tab:apls}. 
+Since both values are derived from the scattering form factors through modeling of the electron density of the bilayer,
+we cannot decide, which of the values is more reliable. 
+In general, we can conclude that ECC-lipids
+%the presented lipid models with ECC-lipids 
+reproduce the experimental structural parameters of the lipid bilayers 
+with a comparable accuracy to existing state-of-the-art lipid models~\citep{botan15, ollila16, Pluhackova2016}.
+
 \begin{figure}[tb!] 
@@
 
-Despite the success in simulating bilayer dimensions and acyl chain conformations,
-lipid force fields have typically problems in capturing the correct glycerol backbone and
-headgroup order parameters and conformations \cite{botan15,ollila16,nmrlipids_proj4}.
-This is the case also for Lipid17 POPS simulations, where the headgroup and glycerol backbone
-order parameters are quite far from experimental values with both Dang (Fig. \ref{simVSexpNOions_POPS})
-and {\AA}qvist \cite{nmrlipids_proj4} ions. The values from ECC-POPS simulation are closer,
-but not in full agreement with experiments (Fig. \ref{simVSexpNOions_POPS}).
-Especially, the experimentally observed more negative values of the $\beta$-carbon
-and the larger forking (difference of order parameters between hydrogens attached to the
-same carbon) of the $\alpha$-carbon in POPS than in POPC \cite{nmrlipids_proj4}
-are not reproduced by neither of the models here. These differences are previously related
-to a more rigid structure of PS headgroup and the stiffness of the C$_\alpha$-C$_\beta$-C$_\gamma$-O$_\gamma$
-dihedral, which is \todo{the dihedral should be plotted and sentence finished.}
+The head group and acyl chain order parameters within ECC-lipids
+are in general in a good agreement with the experimental values 
+as shown in Fig. \ref{simVSexpNOions_POPS}. 
+The acyl chain order parameters in particular agree with the experimental data within error bars.
+The order parameters of the head groups are at an accuracy comparable to 
+other currently available classical models of lipids \citep{botan15, catte16, Pluhackova2016, nmrlipids_proj4}. 
+In particular, ECC-POPS is closer to the experimental order parameters 
+than Lipid17 with either Dang \cite{smith94,chang1999,dang2006} 
+or ff99 ions \citet{aqvist90}.
 
@@
 %For POPC in pure water, the order parameter $\beta$ agrees well with the experiment, while the order parameter $\alpha$ is somewhat lower. 
-
-In conclusion, the ECC-POPS model reproduces the lipid bilayer dimensions and
-acyl chain conformations in good agreement with experiments, but has room for
-improvement in the glycerol backbone and headgroup region. Therefore,
-its structural accuracy is in the same level as in other state-of-the-art
-lipid force fields \cite{botan15, ollila16, Pluhackova2016, nmrlipids_proj4}.
+The head group order parameters of PS are a bit more complicated 
+compared to PC as the order parameter $\alpha$ exhibits a notable forking 
+(i.e., the hydrogen atoms in the group give rise to different order parameters, see Fig.~\ref{simVSexpNOions_POPS}).
+One of the order parameters of ECC-POPS, $\alpha_1$, agrees well with the experiment, 
+while the other, $\alpha_2$, adopts a higher value underestimating the experimentally reported forking. 
+There is only one order parameter $\beta$ in POPS, 
+which has a higher value closer to zero in the ECC-lipids model than in experiment. 
+Such a feature suggests that the model likely overestimates the orientational freedom of the PS head group.
 

### 2019-03-11 Manuscript/manuscript.tex
Update text

@@
 ECC-POPS simulation (Fig. \ref{fig:POPS-counterions-dens}).
-The different cation binding affinities to lipid bilayers have been previously evaluated
+The cation binding affinities to lipid bilayers in simulations have been previously evaluated
 against experiments using the order parameters of $\alpha$ and $\beta$ carbons in the
-phosphatidylcholine (PC) lipid headgroup (see Fig.~\ref{fig:chemstruct_pc_ps} for the labeling) \cite{catte16,melcr18,nmrlipids_proj4}.
-According to the ''electrometer concept'' \citep{seelig87}, these order parameters decrease
-proportionally to the bound charge and can therefore be used to quantify the cation binding
-affinity to lipid bilayers. Because the response of PS lipid headgroup order parameters to the bound charge
-is less well undestood and addition of cation may cause phase transitions in negatively charged
-bilayers \cite{feigenson86,mattai89,roux91,roux90}, the cation binding affinity to bilayers with
-PS lipids has evaluated measuring the PC headgroup order parameters from POPC:POPS (5:1)
+phosphatidylcholine (PC) lipid headgroup (see Fig.~\ref{fig:chemstruct_pc_ps} for the labeling) \cite{catte16,melcr18,nmrlipids_proj4},
+which decrease proportionally to the bound charge according to the ''electrometer concept'' \citep{seelig87}.
+Althought the response of PS lipid headgroup order parameters to the bound charge is also systematic,
+it is less well undestood and the addition of cations may cause phase transitions in negatively charged
+bilayers \cite{feigenson86,mattai89,roux91,roux90}. Therefore, the cation binding affinity to bilayers with
+PS lipids has been evaluated measuring the PC headgroup order parameters from POPC:POPS (5:1)
 mixtures\cite{roux90,nmrlipids_proj4}.
-
-The situation is more complicated for lipid bilayers containing
-PS lipids, for which there is no ion-free reference state due to the presence of counterions.
-Moreover, additional ions often lead to phase transitions in such bilayers \cite{nmrlipids_proj4,feigenson86,mattai89,roux91,roux90}
-
-Unlinke in pure PC bilayers,
-Interpreting the changes of the head group order parameters of 
-mixed PC:PS lipid bilayers 
-from both simulations and experiments
-is complicated by the presence of counterions. \cite{nmrlipids_proj4}
-In order to see the effects of different counterions, 
-we performed simulations with \ce{Na+} and also with \ce{K+} counterions, 
-which bind more weakly to phospholipids. \cite{nmrlipids_proj4, melcr18}
-This can also be seen in the density plots of the counterions in Fig.~\ref{fig:POPS-counterions-dens},
-which show a significantly reduced concentration of \ce{K+} ions 
-at the membrane head group region compared to \ce{Na+}
-for both PS and mixed PC:PS (5:1) bilayers.
-
-%According to the electrometer concept,
-%the head group order parameters $\alpha$ and $\beta$ of PC lipids
-%change proportionally to the amount of bound charge
-%in a fixed ratio. \cite{seelig87,scherer87,roux90} 
+Following our previous work \cite{nmrlipids_proj4}, we evaluate the counterion binding affinity
+to lipid bilayer in simulations containing PS using two different approaches.
+
+% The situation is more complicated for lipid bilayers containing
+%PS lipids, for which there is no ion-free reference state due to the presence of counterions.
+
+%Unlinke in pure PC bilayers,
+%Interpreting the changes of the head group order parameters of 
+%mixed PC:PS lipid bilayers 
+%from both simulations and experiments
+%is complicated by the presence of counterions. \cite{nmrlipids_proj4}
+
+\begin{figure}[tbp!] 
+  \centering 
+  \includegraphics[height=\figheight]{../img/ecc_pops/density_profiles_na-k-counterions_wat_phos_compar_purePOPS_ecclipids-lipid17.pdf}
+  \includegraphics[height=\figheight]{../img/ecc_pops/density_profiles_na-k-counterions_wat_phos_compar_5PC-1PS_ecclipids-lipid17.pdf}
+  \caption{\label{fig:POPS-counterions-dens}
+    Number density profiles of \ce{K^{+}}, \ce{Na^{+}} counterions along the membrane normal axis 
+    for the negatively charged bilayers composed of only POPS (left) or POPC:POPS (5:1, right) 
+    using ECC-lipids/ECC-ions (solid lines) and Lipid17/Dang ions (dashed lines).  
+    The density profiles of phosphate groups and water are divided by 4 and 100, respectively.  
+}
+\todo{JOE: add labels with the relative surface excess $\Gamma$, which are promised to be present in the text.} \\
+\todo{JOE: split this figure into two: separate for POPS (to SI) and PC:PS 5:1 (main text)? SAMULI: Should be POPS in main text, mixture in SI.}
+\end{figure}
+
+\begin{figure}[tb!] 
+  \centering 
+  \includegraphics[width=\figwidth]{../Fig/lipids_chemfig_POPC_POPS.pdf} 
+  \caption{ \label{fig:chemstruct_pc_ps} 
+            The chemical structure of POPC and POPS 
+            including the labeling of the carbon segments
+            used in the measurements of order parameteres. 
+          }
+          \todo{The POPS structure and labeling is already in figure \ref{simVSexpNOions_POPS}.
+            The POPC structure and labeling could be added on top of the new version of figure \ref{fig:delta_ordPar_NaCl_PC-PS_mix} with only PC,
+          where PC results occur in the first time, and this figure could be removed.}
+\end{figure}
+
 %This was demonstrated for a bound positive charge in simulations with POPC and 
@@
 
-\begin{figure}[tbp!] 
-  \centering 
-  \includegraphics[height=\figheight]{../img/ecc_pops/density_profiles_na-k-counterions_wat_phos_compar_purePOPS_ecclipids-lipid17.pdf}
-  \includegraphics[height=\figheight]{../img/ecc_pops/density_profiles_na-k-counterions_wat_phos_compar_5PC-1PS_ecclipids-lipid17.pdf}
-  \caption{\label{fig:POPS-counterions-dens}
-    Number density profiles of \ce{K^{+}}, \ce{Na^{+}} counterions along the membrane normal axis 
-    for the negatively charged bilayers composed of only POPS (left) or POPC:POPS (5:1, right) 
-    using ECC-lipids/ECC-ions (solid lines) and Lipid17/Dang ions (dashed lines).  
-    The density profiles of phosphate groups and water are divided by 4 and 100, respectively.  
-}
-\todo{JOE: add labels with the relative surface excess $\Gamma$, which are promised to be present in the text.} \\
-\todo{JOE: split this figure into two: separate for POPS (to SI) and PC:PS 5:1 (main text)? SAMULI: Should be POPS in main text, mixture in SI.}
-\end{figure}
-
-\begin{figure}[tb!] 
-  \centering 
-  \includegraphics[width=\figwidth]{../Fig/lipids_chemfig_POPC_POPS.pdf} 
-  \caption{ \label{fig:chemstruct_pc_ps} 
-            The chemical structure of POPC and POPS 
-            including the labeling of the carbon segments
-            used in the measurements of order parameteres. 
-          }
-          \todo{The POPS structure and labeling is already in figure \ref{simVSexpNOions_POPS}.
-            The POPC structure and labeling could be added on top of the new version of figure \ref{fig:delta_ordPar_NaCl_PC-PS_mix} with only PC,
-          where PC results occur in the first time, and this figure could be removed.}
-\end{figure}
-
-Generally improved behaviour 
-of the POPC and POPS head group order parameters 
-with \ce{NaCl} or \ce{KCl} concentrations 
-was achieved through the combination of models 
-ECC-lipids and ECC-ions \citep{martinek17, kohagen16, Pluharova2014}. 
-Simulations with these models 
-reveal a good agreement with the NMR experiments 
-for both neutral and negatively charged membranes.
-The results are plotted for \ce{NaCl} in Fig. \ref{fig:delta_ordPar_NaCl_PCPS}, 
-and for \ce{KCl} in Fig. \ref{fig:delta_ordPar_KCl_PCPS}.
-
-The slopes of the response of the PC head group order parameters $\alpha$ and $\beta$ 
-in the mixed PC:PS (5:1) bilayer with \ce{K+} counterions
-(Fig.~\ref{fig:delta_ordPar_NaCl_PC-PS_mix})
-agree qualitatively well with experiments for both employed models,
-i.e., for both Lipid14/17 and ECC-lipids. 
-The same property from the simualtions with \ce{Na+} counterions, 
-which were used in the experimental measurement,
-are not as good, however.
-In particular,
-ECC-lipids have a weaker response of their PC head group order parameters compared to the experiments, 
-while the simulations with Lipid17 model reveal a completely opposite trend.
-This effect likely arises from an overestimated binding of the sodium cations to the phospholipids \cite{nmrlipids_proj4},
-which is likely still present in a small amount also in the simulations with ECC-lipids. 
-The large difference between the affinities of the employed lipid models to the counterions 
-is visible in the density plots in Fig~\ref{fig:POPS-counterions-dens}.
-
-The order parameters of the PS head groups including their changes with the content of PC lipids
-are challenging to capture correctly in classical MD simulation models. \cite{nmrlipids_proj4}
-From such a perspective,
-the changes of the PS head group order parameters from ECC-lipids 
-with \ce{Na+}, and also \ce{K+}, counterions 
-agree comparably well to the experimental measurements
-(Fig.~\ref{fig:delta_ordPar_NaCl_PC-PS_mix}). 
-Interestingly enough, 
-although the Lipid17 model performed well with \ce{K+} counterions 
-for the PC head group order parameter changes,
-it does not agree in the equivalent changes in PS head groups
-pointing at non-negligible effects of even relatively weakly binding cations 
-on the structure of negatively charged lipids.
-
-Despite the great improvements in the overall structure of ECC-POPS 
-and its response to PC lipids and counterions,
-there is still a large room for improving it
-to agree with the experimental measurements.
-
+First, we monitor the response of PC headgroup order parameters to the increasing amount of PS (Fig.~\ref{fig:delta_ordPar_NaCl_PC-PS_mix}).
+According to the electrometer concept, the head group order parameters of PC lipids increase
+with the increasing amount of negatively charged PS lipids in experiments \cite{seelig87,scherer87}.
+The trend is observed with both force fields in simulations with \ce{K+} counterions and
+in ECC-POPS simulation with \ce{Na+} counterions, but not in Lipid17/Dang ion simulations with \ce{Na+} counterions.
+The result can be explained by the higher binding affinity of sodium in the
+Lipid17/Dang ion simulations (Fig~\ref{fig:POPS-counterions-dens}), which
+compensates the negative charge of PS lipids at the interface.
 \begin{figure*}[!tbp] 
@@
     and change the PS-plot x-axis to reflect the dilution of POPS by PC (now shows the content of PS). 
-The experimental lines in the bottom plot shall be different to make clear that they are not with K counterions (gray dashed?).}
+    The experimental lines in the bottom plot shall be different to make clear that they are not with K counterions (gray dashed?).} \\
+  \todo{Error bar of 0.02 should be shown for the experimental POPS results because they come from different experiments, see caption of figure 8 in NMRlipids IV publication.}
 \end{figure*}
+
+Second, we monitor the response of PC headgroup order parameters to the additional
+amount of sodium and potassium in POPC:POPS (5:1) mixtures.
+In experiments, the addition of K$^+$ ions led to a very modest decrease of the order parameters,
+while the decrease was more pronounced with Li$^+$ ions and strongest with divalent Ca$^{2+}$ and Mg$^{2+}$ ions \cite{roux90}.
+According to the electrometer concept, the decrease of PC lipid head group order parameters
+correlates with the amount of bound cations \cite{akutsu81,altenbach84,seelig87,roux90,catte16,nmrlipids_proj4},
+thus the binding affinity of the ions to POPC:POPS (5:1) mixtures increase in order K$^{+}$ $<$ Li$^{+}$  $<$ Mg$^{2+}$  $<$ Ca$^{2+}$.
+The response of PC headgroup order parameters to the increasing amount of KCl is close to experiments
+in both force fields (Fig. \ref{fig:delta_ordPar_KCl_PCPS}), athought the change of the $\alpha$-carbon
+order parameter may be slightly overestimated by the Lipid17/Dang simulations. This could be
+explained by the slightly deeper penetration of potassium into the bilayer (Fig. \ref{fig:POPS-counterions-dens}).
+Because experimental data with the additional amount of sodium in POPC:POPS (5:1) mixture was not available,
+we compared the simulation results to the experimental data with the additional amounts of
+lithium and potassium ions, which should give the lower and upper bounds to the sodium binding
+affinity (Fig. \ref{fig:delta_ordPar_KCl_PCPS}). The PC headgroup response ot the additional
+amount of sodium is clearly overestimated by the Lipid17/Dang simulation, which is in line
+the higher binding affinity (Fig. \ref{fig:POPS-counterions-dens}). Also, the response
+in ECC-POPS simulations is probably too close to the experimental results for lithium,
+indicating slight overbinding also in this simulation. In conclusion, the response of PC
+headgroups to added amount of PS and to the added amount of ions suggest that the Lipid17/Dang
+simulations sinficantly overestimate the sodium binding affinity to the PS containing lipid bilayers, while
+ECC-POPS simulations give results closer to the experiments, but may still slightly overestimate
+the binding affinity. The binding affinity of potassium is better described by both force fields.
+
+%Generally improved behaviour 
+%of the POPC and POPS head group order parameters 
+%with \ce{NaCl} or \ce{KCl} concentrations 
+%was achieved through the combination of models 
+%ECC-lipids and ECC-ions \citep{martinek17, kohagen16, Pluharova2014}. 
+%Simulations with these models 
+%reveal a good agreement with the NMR experiments 
+%for both neutral and negatively charged membranes.
+%The results are plotted for \ce{NaCl} in Fig. \ref{fig:delta_ordPar_NaCl_PCPS}, 
+%and for \ce{KCl} in Fig. \ref{fig:delta_ordPar_KCl_PCPS}.
+
+%The order parameters of the PS head groups including their changes with the content of PC lipids
+%are challenging to capture correctly in classical MD simulation models. \cite{nmrlipids_proj4}
+%From such a perspective,
+%the changes of the PS head group order parameters from ECC-lipids 
+%with \ce{Na+}, and also \ce{K+}, counterions 
+%agree comparably well to the experimental measurements
+%(Fig.~\ref{fig:delta_ordPar_NaCl_PC-PS_mix}). 
+%Interestingly enough, 
+%although the Lipid17 model performed well with \ce{K+} counterions 
+%for the PC head group order parameter changes,
+%it does not agree in the equivalent changes in PS head groups
+%pointing at non-negligible effects of even relatively weakly binding cations 
+%on the structure of negatively charged lipids.
+
+%Despite the great improvements in the overall structure of ECC-POPS 
+%and its response to PC lipids and counterions,
+%there is still a large room for improving it
+%to agree with the experimental measurements.
 
@@
   }
+  \todo{We should plot the experimental results with KCl also in this figure.} \\
+  \todo{Why there are no error bars for ECC-POPS in this and other figures?} \\
   \todo{SAMULI: Use the same y-axis scale in the bottom figures showing angles for POPC and POPS to ease the comparison.}
@@
   }
+  \todo{I am not sure if we need this figure at all.}
 \end{figure}

### 2019-03-12 Manuscript/manuscript.tex
Update text

@@
 in the intro.}
+
+The interaction of various cations including calcium 
+with  neutral and also negatively charged lipid bilayers
+was thoroughly compared between simulations and experiments 
+from the point of the electrometer concept \cite{roux90, seelig87, altenbach84}
+in the references~\citenum{catte16, nmrlipids_proj4}. 
+It was concluded that the interaction of calcium with zwitterionic PC bilayers 
+is generally overestimated in all studied models \cite{catte16},
+while none of the models was capable of interpreting the NMR structural parameters
+with the negatively charged PC:PS (5:1) bilayers \cite{nmrlipids_proj4}. 
+Accounting for the electronic polarization was identified as 
+a non-negligible correction to the interaction between cations and POPC bilayers in Ref.~\cite{melcr18}.
+
+Here, we demonstrate that the same approach 
+significantly improves the description of calcium binding 
+also to the negatively charged PC:PS (5:1) bilayers. 
+In correspondence with the Ref.~\citenum{nmrlipids_proj4},
+we also use the changes of the head group order parameters $S^\alpha$ and $S^\beta$ 
+in PC and PS lipids 
+to quantify the binding affinity of the calcium cations
+and the induced structural changes
+(Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}). 
+To directly connect to the previous works \citep{catte16, nmrlipids_proj4}
+and to mark the improvement after treating electronic polarizability, 
+we show simulation results from the newly developed ECC-lipids and also from Lipid17 \citep{lipid17-future}.
 
@@
 \todo{JOE: add labels with the relative surface excess $\Gamma$, which are promised to be present in the text.} \\
-\todo{JOE: split this figure into two: separate for POPS (to SI) and PC:PS 5:1 (main text)? SAMULI: Should be POPS in main text, mixture in SI.}
+\todo{JOE: split this figure into two: separate for POPS (to SI) and PC:PS 5:1 (main text)?} \\
+\todo{SAMULI: I feel now that we should keep the as it is, but if something is moved to SI, it should be POPS in main text, mixture in SI.}
 \end{figure}
@@
             The POPC structure and labeling could be added on top of the new version of figure \ref{fig:delta_ordPar_NaCl_PC-PS_mix} with only PC,
-          where PC results occur in the first time, and this figure could be removed.}
+            where PC results occur in the first time, and this figure could be removed.} \\
+          \todo{$\gamma$ carbon of POPS is the COO- minus carbon. The $\gamma$ label should be moved there.}
 \end{figure}
@@
 
-First, we monitor the response of PC headgroup order parameters to the increasing amount of PS (Fig.~\ref{fig:delta_ordPar_NaCl_PC-PS_mix}).
+First, we monitor the response of POPC headgroup order parameters to the increasing amount of POPS (Fig.~\ref{fig:delta_ordPar_NaCl_PC-PS_mix}).
 According to the electrometer concept, the head group order parameters of PC lipids increase
-with the increasing amount of negatively charged PS lipids in experiments \cite{seelig87,scherer87}.
+with the increasing amount of negatively charged PS lipids in experiments, because the headgroup P--N dipole
+tilts more parallel to the membrane interface \cite{seelig87,scherer87}.
 The trend is observed with both force fields in simulations with \ce{K+} counterions and
-in ECC-POPS simulation with \ce{Na+} counterions, but not in Lipid17/Dang ion simulations with \ce{Na+} counterions.
+in ECC-POPS simulation with \ce{Na+} counterions, but not in Lipid17/Dang ion simulations with \ce{Na+} counterions,
+where the order parameters are unchanged or decrease and the P--N vector tilts more perpendicular to the membrane interface.
 The result can be explained by the higher binding affinity of sodium in the
@@
 The response of PC headgroup order parameters to the increasing amount of KCl is close to experiments
-in both force fields (Fig. \ref{fig:delta_ordPar_KCl_PCPS}), athought the change of the $\alpha$-carbon
+and the change in P--N vector is negligible in both force fields (Fig. \ref{fig:delta_ordPar_KCl_PCPS}), athought the change of the $\alpha$-carbon
 order parameter may be slightly overestimated by the Lipid17/Dang simulations. This could be
@@
 lithium and potassium ions, which should give the lower and upper bounds to the sodium binding
-affinity (Fig. \ref{fig:delta_ordPar_KCl_PCPS}). The PC headgroup response ot the additional
-amount of sodium is clearly overestimated by the Lipid17/Dang simulation, which is in line
-the higher binding affinity (Fig. \ref{fig:POPS-counterions-dens}). Also, the response
-in ECC-POPS simulations is probably too close to the experimental results for lithium,
-indicating slight overbinding also in this simulation. In conclusion, the response of PC
-headgroups to added amount of PS and to the added amount of ions suggest that the Lipid17/Dang
-simulations sinficantly overestimate the sodium binding affinity to the PS containing lipid bilayers, while
-ECC-POPS simulations give results closer to the experiments, but may still slightly overestimate
-the binding affinity. The binding affinity of potassium is better described by both force fields.
+affinity (Fig. \ref{fig:delta_ordPar_NaCl_PCPS}). The PC headgroup response to the additional
+amount of sodium is clearly overestimated by the Lipid17/Dang simulation and
+the response in ECC-POPS simulations is probably too close to the experimental results for lithium.
+In conclusion, the results suggest that the sodium binding affinity (Fig. \ref{fig:POPS-counterions-dens}) and
+the tilting of P--N vector more perpendicular to the membrane surface (Fig. \ref{fig:delta_ordPar_NaCl_PCPS})
+are clearly overestimated in Lipid17/Dang simulations and slightly overestimated in ECC-POPS simulations,
+while 
+%In conclusion, the response of PC
+%headgroups to added amount of PS and to the added amount of ions suggest that the Lipid17/Dang
+%simulations sinficantly overestimate the sodium binding affinity to the PS containing lipid bilayers, while
+%ECC-POPS simulations give results closer to the experiments, but may still slightly overestimate
+%the binding affinity.
+the binding affinity of potassium is better described by both force fields.
 
@@
 % 
-The interaction with \ce{K^+}, which binds very weakly to both neutral and negatively charged membranes, 
-renders a qualitatively different response of the order parameter $S^\beta$ in POPS 
-in the mixed negatively charged bilayers compared to the neutral bilayers. 
-While the order parameter $S^\beta$ increases for both \ce{Na^+} and \ce{Ca^{2+}},
-it decreases in the presence of \ce{K^+}
-(\ce{KCl}:  Fig.~\ref{fig:delta_ordPar_KCl_PCPS}, 
- \ce{NaCl}: Fig.~\ref{fig:delta_ordPar_NaCl_PCPS}, and
-\ce{CaCl2}: Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}). 
-Despite the tremendous effort in the development of classical MD models of lipids, however,
-there is no such model, 
-which would agree with the experimental trends for these salt solutions \cite{nmrlipids_proj4}. 
-In contrast, ECC-lipids with ECC-ions capture the different response 
-of the order parameters $S^{\beta}$, $S^{\alpha _1}$ and $S^{\alpha _2}$ in POPS 
-to various concentrations of \ce{NaCl} and \ce{KCl} in a good agreement with the experiments
-(\ce{KCl}:  Fig.~\ref{fig:delta_ordPar_KCl_PCPS}, 
- \ce{NaCl}: Fig.~\ref{fig:delta_ordPar_NaCl_PCPS}). 
-Such a dramatic improvement in the structural parameters 
-demonstrates that including electronic polarization in phospholipids is crucial 
-even for interactions with soft cations like \ce{K^+}.
+The changes of POPS headgroup order parameters are almost negligible in experiments with increasing POPC fraction in bilayer
+or monovalent ion concentration in solvent
+(Figs. \ref{fig:delta_ordPar_NaCl_PCPS}, \ref{fig:delta_ordPar_KCl_PCPS} and \ref{fig:delta_ordPar_NaCl_PC-PS_mix}),
+while calcium leads to relatively large changes with low concentrations followed by a rapid saturation below 100mM (Fig. \ref{fig:delta_ordPar_CaCl_PCPS}) \cite{scherer87,roux90}.
+%The interaction with \ce{K^+}, which binds very weakly to both neutral and negatively charged membranes, 
+%renders a qualitatively different response of the order parameter $S^\beta$ in POPS 
+%in the mixed negatively charged bilayers compared to the neutral bilayers. 
+%While the order parameter $S^\beta$ increases for both \ce{Na^+} and \ce{Ca^{2+}},
+%it decreases in the presence of \ce{K^+}
+%(\ce{KCl}:  Fig.~\ref{fig:delta_ordPar_KCl_PCPS}, 
+% \ce{NaCl}: Fig.~\ref{fig:delta_ordPar_NaCl_PCPS}, and
+%\ce{CaCl2}: Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}). 
+In Lipid17/Dang simulations, the changes of POPS headgroup order parameters with increasing amount of POPC
+are overestimated, which is also related to the changes in the orientations of the P--N and C$_{\beta}$--C$_{\gamma}$
+vectors (Fig. \ref{fig:delta_ordPar_NaCl_PC-PS_mix}). The overestimation of the order parameter changes
+was also observed in previous simulations with other force fields \cite{nmrlipids_proj4}, but not in
+ECC-POPS simulation, where also the changes in the PS headgroup orientation are more modest (Fig. \ref{fig:delta_ordPar_NaCl_PC-PS_mix}).
+Also, the POPS headgroup orientation and order parameters are more sensitive to the additional amount
+of sodium and potassium in Lipid17/Dang simulations, while dependence is weaker and closer to experiments
+in ECC-POPS simulations (Figs. \ref{fig:delta_ordPar_NaCl_PCPS} and \ref{fig:delta_ordPar_KCl_PCPS}).
+%In contrast, ECC-lipids with ECC-ions capture the different response 
+%of the order parameters $S^{\beta}$, $S^{\alpha _1}$ and $S^{\alpha _2}$ in POPS 
+%to various concentrations of \ce{NaCl} and \ce{KCl} in a good agreement with the experiments
+%(\ce{KCl}:  Fig.~\ref{fig:delta_ordPar_KCl_PCPS}, 
+% \ce{NaCl}: Fig.~\ref{fig:delta_ordPar_NaCl_PCPS}). 
+%Such a dramatic improvement in the structural parameters 
+%demonstrates that including electronic polarization in phospholipids is crucial 
+%even for interactions with soft cations like \ce{K^+}. 
+In conclusion, using ECC to POPS lipid headgroup improves its
+interactions with monovalent ions and POPC headgroups when compared with NMR order parameter data,
+althought there is room for improvement in the sodium binding affinity.
 
@@
 
-% This§: 
-% adsorption and affinities of Na and K to PC:PS memb
-% from densities and rel.surf.excess
-% 
-%in  Fig.~\ref{fig:POPS-counterions-dens} (only counterions)
-%and Fig.~\ref{fig:nacl-dens_PCPS}        (additional salt concentrations), 
-%
-The affinities of \ce{Na^+} and \ce{K^+} to phospholipid membranes
-can be described by their relative surface excesses with respect to water, $\Gamma ^{w} _{ion}$. 
-Such a quantity compares the adsorption of ions to the adsorption of water molecules at an interface 
-without the necessity of defining a Gibbs dividing surface. \citep{melcr18, chattorajBOOK}
-Interestingly enough, \ce{K^+} maintains negative values of $\Gamma^{w}_{K}$ even for the negatively charged PC:PS~(5:1) bilayer.
-This is also evident from the density plots in  Fig.~\ref{fig:nacl-dens_PCPS}
-showing that the density profile of \ce{K^+} along the membrane normal 
-decays clearly faster towards the membrane centre compared to the density of water. 
-In contrast, the density profile of \ce{Na^+} has a shoulder 
-at the region of the phosphate groups,
-which demonstrates that sodium cations adsorb weakly to the membrane interface compared to the aquaeous solution
-by yielding a small positive value of $\Gamma^{w}_{Na}$.
-
-\todo{JOE: Add a discussion about the still slightly overestimated affinity of \ce{Na^+}.}
+%SAMULI: I WOULD LEAVE THIS PART OUT
+%% This§: 
+%% adsorption and affinities of Na and K to PC:PS memb
+%% from densities and rel.surf.excess
+%% 
+%%in  Fig.~\ref{fig:POPS-counterions-dens} (only counterions)
+%%and Fig.~\ref{fig:nacl-dens_PCPS}        (additional salt concentrations), 
+%%
+%The affinities of \ce{Na^+} and \ce{K^+} to phospholipid membranes
+%can be described by their relative surface excesses with respect to water, $\Gamma ^{w} _{ion}$. 
+%Such a quantity compares the adsorption of ions to the adsorption of water molecules at an interface 
+%without the necessity of defining a Gibbs dividing surface. \citep{melcr18, chattorajBOOK}
+%Interestingly enough, \ce{K^+} maintains negative values of $\Gamma^{w}_{K}$ even for the negatively charged PC:PS~(5:1) bilayer.
+%This is also evident from the density plots in  Fig.~\ref{fig:nacl-dens_PCPS}
+%showing that the density profile of \ce{K^+} along the membrane normal 
+%decays clearly faster towards the membrane centre compared to the density of water. 
+%In contrast, the density profile of \ce{Na^+} has a shoulder 
+%at the region of the phosphate groups,
+%which demonstrates that sodium cations adsorb weakly to the membrane interface compared to the aquaeous solution
+%by yielding a small positive value of $\Gamma^{w}_{Na}$.
 
@@
 \label{section:lip-ion_ca}
+
+The headgroup order parameters of PC lipids decrease more upon addition of \ce{CaCl2}
+in POPC:POPS (5:1) mixtures than in pure PC bilayers due to the higher calcium binding
+affinity to the negatively charged lipid bilayers \cite{akutsu81,altenbach84,roux90,nmrlipids_proj4}.
+Following our previous work \cite{nmrlipids_proj4}, we use the experimental
+headgroup order parameter data as a function of CaCl$_2$ concentration
+from POPC:POPS (5:1) mixture to evaluate the calcium binding affinity to lipid bilayers containing PS lipids.
+The decrease of the PC headgroup order parameters upon addition of \ce{CaCl2} 
+is overestimated in the Lipid17/Dang simulation, while ECC-POPS simulations give results
+closer to the experiments (Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}). 
+This indicates that the lower binding affinity of calcium in ECC-POPS
+simulation is more realistic (Fig.~\ref{fig:cacl-dens_PCPS}).
+\todo{JOE: The density plot still needs to be changed to show both Lipid17 and ECC-lipids, and to support this argument.}
+However, the steep decrease of the order parameters at low calcium concentrations
+observed especially in the negatively charged bilayers is not captured by the ECC-lipid simulations.
+
+%The binding of calcium cations to the phospholipid bilayer is shown 
+%in the density profiles of \ce{Ca^{2+}}, \ce{Na^+} counterions 
+%and also \ce{Cl^-} ions in Fig.~\ref{fig:cacl-dens_PCPS}. 
+%Clearly, after accounting for the electronic polarizability, 
+%the population of the cations at the membrane interface is significantly diminished.
 
@@
 \end{figure}
-
-The interaction of various cations including calcium 
-with  neutral and also negatively charged lipid bilayers
-was thoroughly compared between simulations and experiments 
-from the point of the electrometer concept \cite{roux90, seelig87, altenbach84}
-in the references~\citenum{catte16, nmrlipids_proj4}. 
-It was concluded that the interaction of calcium with zwitterionic PC bilayers 
-is generally overestimated in all studied models \cite{catte16},
-while none of the models was capable of interpreting the NMR structural parameters
-with the negatively charged PC:PS (5:1) bilayers \cite{nmrlipids_proj4}. 
-Accounting for the electronic polarization was identified as 
-a non-negligible correction to the interaction between cations and POPC bilayers in Ref.~\cite{melcr18}.
-
-Here, we demonstrate that the same approach 
-significantly improves the description of calcium binding 
-also to the negatively charged PC:PS (5:1) bilayers. 
-In correspondence with the Ref.~\citenum{nmrlipids_proj4},
-we also use the changes of the head group order parameters $S^\alpha$ and $S^\beta$ 
-in PC and PS lipids 
-to quantify the binding affinity of the calcium cations
-and the induced structural changes
-(Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}). 
-To directly connect to the previous works \citep{catte16, nmrlipids_proj4}
-and to mark the improvement after treating electronic polarizability, 
-we show simulation results from the newly developed ECC-lipids and also from Lipid17 \citep{lipid17-future}. 
-While the response of the order parameters of PC to \ce{CaCl2} concentrations
-is qualitatively correct in the Lipid14/17 model \cite{lipid17},
-it is quantitatively comparable to the experimental trends in ECC-lipids. 
-Moreover, the structure of PS head group is fundamentaly improved
-after accounting for the electronic polarization 
-(Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}).
-
-Increasing concentrations of \ce{CaCl2} 
-induce a systematic decrease of the order parameters $S^\alpha$ and $S^\beta$ 
-in both neutral POPC bilayers \cite{catte16, melcr18}
-as well as in the negatively charged PC:PS (5:1) bilayers \cite{nmrlipids_proj4, roux90}. 
-The change of the head group order parameters 
-are well reproduced only by ECC-lipids in Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}.
 
@@
 
-The changes of the order parameters $\alpha$ and $\beta$ at low \ce{CaCl2} concentrations
-have a steep onset at low concentrations in the negatively charged bilayers compared to neutral POPC bilayers. 
-However, such a detail is not captured by ECC-lipids
-indicating the necessity for further improvements.
-The source of this discrepancy may arise 
-from the interactions between PC and PS lipids and with \ce{Na^+} counterions,
+The source of this discrepancy may arise from the interactions between PC and PS lipids and with \ce{Na^+} counterions,
 from the sub-optimal structure of the lipids (Fig.~\ref{simVSexpNOions_POPS},
@@
 
-The binding of calcium cations to the phospholipid bilayer is shown 
-in the density profiles of \ce{Ca^{2+}}, \ce{Na^+} counterions 
-and also \ce{Cl^-} ions in Fig.~\ref{fig:cacl-dens_PCPS}. 
-Clearly, after accounting for the electronic polarizability, 
-the population of the cations at the membrane interface is significantly diminished. 
-\todo{JOE: The density plot still needs to be changed to show both Lipid17 and ECC-lipids, and to support this argument.}
 Still, the dominant contribution to the binding of calcium comes from the phosphate groups. 

### 2019-03-12 Manuscript/manuscript.tex
Update text

@@
   }
-  \todo{I am not sure if we need this figure at all.}
+  \todo{I am not sure if we need this figure at all or should we move it to SI.}
+  \todo{The concentrations in the figure do not match with the table.}
 \end{figure}
 
-%SAMULI: I WOULD LEAVE THIS PART OUT
-%% This§: 
-%% adsorption and affinities of Na and K to PC:PS memb
-%% from densities and rel.surf.excess
-%% 
-%%in  Fig.~\ref{fig:POPS-counterions-dens} (only counterions)
-%%and Fig.~\ref{fig:nacl-dens_PCPS}        (additional salt concentrations), 
-%%
-%The affinities of \ce{Na^+} and \ce{K^+} to phospholipid membranes
-%can be described by their relative surface excesses with respect to water, $\Gamma ^{w} _{ion}$. 
+% This§: 
+% adsorption and affinities of Na and K to PC:PS memb
+% from densities and rel.surf.excess
+% 
+%in  Fig.~\ref{fig:POPS-counterions-dens} (only counterions)
+%and Fig.~\ref{fig:nacl-dens_PCPS}        (additional salt concentrations), 
+%
+The influence of negatively charged PS on the counterion binding affinity can
+be estimated by comparing the relative surface excesses with respect to water, $\Gamma ^{w} _{\rm ion}$,
+between POPC:POPS (5:1) mixture and pure POPC \cite{melcr18}.
 %Such a quantity compares the adsorption of ions to the adsorption of water molecules at an interface 
 %without the necessity of defining a Gibbs dividing surface. \citep{melcr18, chattorajBOOK}
+The value of $\Gamma ^{w} _{\rm Na}=-0.11 \pm 0.01$ nm$^{-2}$ for 1 M sodium in pure POPC, calculated from
+ECC-POPC simulation, was reported in the previous work \cite{melcr18}. Here, the ECC-lipid simulations
+give
+%$\Gamma ^{w} _{\rm K}=-0.12 \pm ??$ nm$^{-2}$ and
+$\Gamma ^{w} _{\rm Na}=0.09 \pm ??$ nm$^{-2}$ for 0.621 M sodium in pure POPC:POPS (5:1) mixture.
+Therefore, the presence of $\sim$17\% POPS in POPC bilayer increases the relative surface
+excess of sodium with respect to water from negative to positive. As discussed above,
+the sodium binding to POPC:POPS (5:1) mixture may be slightly overestimated also in
+ECC-lipid simulations. However, the difference may not be affected because similar small
+overestimation seem to occur also in pure POPC simulation where the headgroup order parameter
+response to NaCl is also slightly overestimated \cite{meclr18}.
 %Interestingly enough, \ce{K^+} maintains negative values of $\Gamma^{w}_{K}$ even for the negatively charged PC:PS~(5:1) bilayer.
@@
 \todo{JOE: The density plot still needs to be changed to show both Lipid17 and ECC-lipids, and to support this argument.}
+The relative surface excess of calcium with respect to water, $\Gamma ^w _{Ca}$,
+calculated from ECC-lipid simulation gives $\Gamma^{w}_{Ca} = 0.24\mathrm{nm^{-2}}$ for the PC:PS (5:1) bilayer with 409 mM CaCl$_2$
+and $\Gamma^{w}_{Ca} = 0.06\mathrm{nm^{-2}}$ for pure POPC bilayer with 350 mM CaCl$_2$ \cite{melcr18},
+illustrating the higher affinity of calcium to bilayers containing PS lipids.
 However, the steep decrease of the order parameters at low calcium concentrations
@@
 
-The structural changes in POPS induced by increasing concentrations of \ce{CaCl2} 
-are very challenging to interpret from both experiments \cite{roux90} 
-and also from MD simulations \cite{nmrlipids_proj4}. 
-The changes of the order parameters $S^\alpha$ and $S^\beta$ in PS head groups from simulations 
-are not found to follow the experimental profiles, 
-nor any common trend,
-with the currently available models \cite{nmrlipids_proj4}. 
-In Fig.~\ref{fig:delta_ordPar_CaCl_PCPS},
-we plot these structural parameters for one of the models in that work, Lipid17, 
-and for ECC-lipids. 
-Although the agreement between ECC-lipids and experiments is not optimal,
-the general trends and behaviour even at highly concentrated solutions of \ce{CaCl2} 
-are well captured by the model. 
-For example the order parameter $\beta$ in POPS 
-adopts a saturated value in experiments and ECC-lipids,
-but it is growing with the \ce{CaCl2} concentration in Lipid17.
-In addition, 
-the forking of the order parameters $\alpha$ 
-is almost constant in the experiment and ECC-lipids,
-but it varies greatly in th Lipid17.
-
-The source of this discrepancy may arise from the interactions between PC and PS lipids and with \ce{Na^+} counterions,
-from the sub-optimal structure of the lipids (Fig.~\ref{simVSexpNOions_POPS},
-and also possibly from the approximate mean-field treatment of the electronic polarization.
-
-It has been pointed out in previous experimental works
-that the conformational flexibility of PS head groups 
-is different from other lipids \cite{browning80}. 
-While the static shielding tensor of PS is not substantially different from PC or PE, 
-the $^{31}$P~$T_1$ relaxation times are significantly shorter for PS. 
-It is suggested that the PS head group is more rigid compared to the other phospholipids, 
-which is also corroborated by a larger chemical shielding anisotropy of PS \cite{browning80}.
+Experimentally measured headgroup order parameters of POPS exhibit a strong dependence
+on low CaCl$_2$ concentrations with a rapid saturation below 100 mM \cite{roux90}.
+The changes are again overestimated by Lipid17/Dang simulations, which
+also exhibits non-systematic changes in both PS headroup orientation and order parameters
+upon addition of CaCl$_2$ (Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}).
+In ECC-lipid simulations,  the P--N and C$_{\beta}$--C$_{\gamma}$ vectors systematically
+tilt more perpendicular and parallel to the membrane interface, respectively,
+and the changes of the PS headgroup order parameters are again closer to the
+experiments (Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}).
+Despite significant improvements in the ECC-lipid against Lipid17/Dang
+and other force fields for the response of PS headgroup order parameters to
+the bound calcium \cite{nmrlipids_proj4}, the strong dependence of both PC and PS
+headgroup order parameters on low concetrations of CaCl$_2$ is not reproduced by
+the model (Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}). In addition to the inaccurate
+interactions between calcium ions and PS headgroup, the potential sources of this
+discrepancy include the above observed slight overestimation of \ce{Na^+} counterions and
+inaccurate structures of the lipid headgroup (Fig.~\ref{simVSexpNOions_POPS}).
+%and also possibly from the approximate mean-field treatment of the electronic polarization.
+
+%are again closer to the experiments is not optimal,
+%the general trends and behaviour even at highly concentrated solutions of \ce{CaCl2} 
+%are well captured by the model. 
+%For example the order parameter $\beta$ in POPS 
+%adopts a saturated value in experiments and ECC-lipids,
+%but it is growing with the \ce{CaCl2} concentration in Lipid17.
+%In addition, 
+%the forking of the order parameters $\alpha$ 
+%is almost constant in the experiment and ECC-lipids,
+%but it varies greatly in th Lipid17.
+
+%The structural changes in POPS induced by increasing concentrations of \ce{CaCl2} 
+%are very challenging to interpret from both experiments \cite{roux90} 
+%and also from MD simulations \cite{nmrlipids_proj4}. 
+%The changes of the order parameters $S^\alpha$ and $S^\beta$ in PS head groups from simulations 
+%are not found to follow the experimental profiles, 
+%nor any common trend, with the currently available models \cite{nmrlipids_proj4}.
+
+%SAMULI: We will not go to the spin relaxation time analysis in this work
+%and average angles do not give information about dynamics
+%
+%It has been pointed out in previous experimental works
+%that the conformational flexibility of PS head groups 
+%is different from other lipids \cite{browning80}. 
+%While the static shielding tensor of PS is not substantially different from PC or PE, 
+%the $^{31}$P~$T_1$ relaxation times are significantly shorter for PS. 
+%It is suggested that the PS head group is more rigid compared to the other phospholipids, 
+%which is also corroborated by a larger chemical shielding anisotropy of PS \cite{browning80}.
 
@@
 \centering
-  \caption{Bulk concentrations, $C _{Ca}$, and buffer concentrations, $C' _{Ca}$, of Ca$^{2+}$;
-           relative surface excess of calcium with respect to water ($\Gamma_{Ca}^{\rm water}$); 
-           and percentages of the population 
+  \caption{%Bulk concentrations, $C _{Ca}$, and buffer concentrations, $C' _{Ca}$, of Ca$^{2+}$;
+           %relative surface excess of calcium with respect to water ($\Gamma_{Ca}^{\rm water}$); 
+    %and
+    Percentages of the population 
            of bound Ca$^{2+}$ to various moieties 
@@
   \begin{tabular}{ l | c c } 
+        interacting moiety    &  \multicolumn{2}{c}{percentage of bound \ce{Ca^{2+}} } \\
+%	\hline
+%	$C _{Ca}\,/\,\mathrm{mM}$  &  $240\pm 10 $  &  $280\pm 10 $  \\
+%	$C'_{Ca}\,/\,\mathrm{mM}$  &  $400\pm 10 $  &  $350\pm 10 $  \\
+%	$\Gamma_{Ca}^{\rm water}\, / \,\mathrm{nm}^{-2}$  &  $0.24 \pm 0.01 $  &  $0.06 \pm 0.01 $  \\
+%	\hline
+%                             &  \multicolumn{2}{c}{ } \\
                              &  5\,POPC:1\,POPS &  POPC   \\
-	\hline
-	$C _{Ca}\,/\,\mathrm{mM}$  &  $240\pm 10 $  &  $280\pm 10 $  \\
-	$C'_{Ca}\,/\,\mathrm{mM}$  &  $400\pm 10 $  &  $350\pm 10 $  \\
-	$\Gamma_{Ca}^{\rm water}\, / \,\mathrm{nm}^{-2}$  &  $0.24 \pm 0.01 $  &  $0.06 \pm 0.01 $  \\
-	\hline
-                             &  \multicolumn{2}{c}{ } \\
-        interacting moiety   &  \multicolumn{2}{c}{percentage of bound \ce{Ca^{2+}} } \\
 	\hline
@@
 	both PC and PS       &   33   &        \\
-  \end{tabular} 
+  \end{tabular} \\
+  \todo{I have commented out the surface excess values (available in text) and concentrations.}
 \end{table}
-
-% Gamma and higher affinity of PS to CA2+
-The higher affinity of calcium to PS head groups compared to PC
-is well demonstrated using the relative surface excess, $\Gamma ^w _{Ca}$,
-summarized in Table~\ref{tab:binding}. 
-While $\Gamma^{w}_{Ca}$ at a neutral POPC bilayer
-is       $\Gamma^{w}_{Ca} = 0.06\mathrm{nm^{-2}}$      \cite{melcr18}, 
-its value for the PC:PS (5:1) bilayer is $\Gamma^{w}_{Ca} = 0.24\mathrm{nm^{-2}}$
-at comparable buffer concentrations  $350 \, \mathrm{mM}$ resp. $400\, \mathrm{mM}$.
 

### 2019-03-13 Manuscript/manuscript.tex
Update text

@@
 \centering
-  \caption{Values of the area per lipid (APL) of POPS bilayers 
-	   at 298~K with only \ce{Na^+} counterions and no additional ions. 
-           Standard deviations are given. \label{tab:apls} } 
+  \caption{Area per lipid (APL) of POPS bilayers from simulations with different force fields at 298~K with \ce{Na^+} counterions. \label{tab:apls} } 
   \begin{tabular}{l|c } 
@@
 }
-\todo{JOE: add labels with the relative surface excess $\Gamma$, which are promised to be present in the text.} \\
-\todo{JOE: split this figure into two: separate for POPS (to SI) and PC:PS 5:1 (main text)?} \\
-\todo{SAMULI: I feel now that we should keep the as it is, but if something is moved to SI, it should be POPS in main text, mixture in SI.}
+\todo{JOE: add labels with the relative surface excess $\Gamma$, which are promised to be present in the text.
+  SAMULI: I think that these are not promised in the current text and addition is not necessary.} \\
+\todo{JOE: split this figure into two: separate for POPS (to SI) and PC:PS 5:1 (main text)?
+  SAMULI: I feel now that we should keep the figure as it is. Also the POPC:POPS (5:1) results are now referred in the text.
+  However, if something is moved to SI, it should be POPS in main text, mixture in SI.}
 \end{figure}
@@
   \caption{\label{fig:delta_ordPar_NaCl_PC-PS_mix} 
-    The head group order parameters of a POPC:POPS (5:1) bilayer as a function of PS content
-    with \ce{Na^+} (top) and \ce{K^+} (bottom) counterions from simulations with ECC-lipids and Lipid17 \cite{lipid17-future} 
-    and experiments (only \ce{Na^+} counterions, but shown also in bottom plots). \cite{roux90}. 
-    The underestimated response of ECC-lipids with \ce{Na^+} counterions 
-    is likely due to a still slightly overestimated binding affinity of \ce{Na^+} to the phospholipids,
-    which is corroborated by the series with \ce{K^+} counterions (lower affinity than \ce{Na^+}),
-    where the response is closer to the experiment, which uses only \ce{Na^+} counterions. 
+    CAPTION FOR THE FIGURE WITH POPC: The POPC head group order parameters and the P--N vector angle
+    with respect to the membrane normal as a function of POPS content in a bilayer
+    from ECC-lipid and Lipid17/Dang simulations with \ce{Na^+} (top) and \ce{K^+} (bottom) counterions.
+    Experimental order parameter values are from Ref. \citenum{scherer87}
+    and the signs from Ref. \citenum{ferreira16} (only \ce{Na^+} counterions, but shown also in bottom plots).
+    CAPTION FOR THE FIGURE WITH POPS: The POPS head group order parameters, the P--N and C$_\beta$--C$_\gamma$ vector angles (stars)
+    with respect to the membrane normal as a function of POPC content in a bilayer
+    from ECC-lipid and Lipid17/Dang simulations with \ce{Na^+} (top) and \ce{K^+} (bottom) counterions.
+    Experimental order parameter values and signs for pure POPS are from Ref. \citenum{nmrlipids_proj4} and values
+    for POPC:POPS (5:1) mixture from Ref. \citenum{roux90} (only \ce{Na^+} counterions, but shown also in bottom plots).
+    %The underestimated response of ECC-lipids with \ce{Na^+} counterions 
+    %is likely due to a still slightly overestimated binding affinity of \ce{Na^+} to the phospholipids,
+    %which is corroborated by the series with \ce{K^+} counterions (lower affinity than \ce{Na^+}),
+    %where the response is closer to the experiment, which uses only \ce{Na^+} counterions. 
   }
@@
   \caption{\label{fig:delta_ordPar_NaCl_PCPS} 
-    Changes of the head group order parameters $\alpha$, $\beta$ and the orientations of the carboxylate group and the P-N vector  
-    of POPC (left) and POPS (right) phospholipids in a POPC:POPS 5:1 bilayer 
-    as a function of \ce{NaCl} concentration in the buffer 
-    from simulations with different force fields and experiments at 298 K. \citep{roux90}
+    Changes of the head group order parameters, the P--N and C$_\beta$--C$_\gamma$ vector angles (stars)
+    with respect to the membrane normal of POPC (left) and POPS (right) in a POPC:POPS (5:1) bilayer 
+    as a function of \ce{NaCl} concentration from ECC-lipid and Lipid17/Dang simulations at 298 K \cite{roux90}.
     Because data with \ce{NaCl} are not available for POPC, 
@@
     which has a lower affinity to phospholipid bilayers compared to \ce{LiCl} \citep{roux90}. 
-    The orientation of the \ce{COO^-} group is defined as 
-    the connector from the $\beta$ carbon to the carbon in \ce{COO^-} (stars, bottom right). 
+    %The orientation of the \ce{COO^-} group is defined as 
+    %the connector from the $\beta$ carbon to the carbon in \ce{COO^-} (stars, bottom right). 
   }
@@
 %$\Gamma ^{w} _{\rm K}=-0.12 \pm ??$ nm$^{-2}$ and
-$\Gamma ^{w} _{\rm Na}=0.09 \pm ??$ nm$^{-2}$ for 0.621 M sodium in pure POPC:POPS (5:1) mixture.
+$\Gamma ^{w} _{\rm Na}=0.09 \pm ??$ nm$^{-2}$ \todo{Error bar should be added.} for 0.621 M sodium in pure POPC:POPS (5:1) mixture.
 Therefore, the presence of $\sim$17\% POPS in POPC bilayer increases the relative surface
@@
 
-%SAMULI: We will not go to the spin relaxation time analysis in this work
-%and average angles do not give information about dynamics
-%
-%It has been pointed out in previous experimental works
-%that the conformational flexibility of PS head groups 
-%is different from other lipids \cite{browning80}. 
-%While the static shielding tensor of PS is not substantially different from PC or PE, 
-%the $^{31}$P~$T_1$ relaxation times are significantly shorter for PS. 
-%It is suggested that the PS head group is more rigid compared to the other phospholipids, 
-%which is also corroborated by a larger chemical shielding anisotropy of PS \cite{browning80}.
-
-We have investigated this effect using the mean orientations of P-N and \ce{COO^-} vectors 
-(defined and plotted in Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}). 
-While the mean angle between the P-N vector of PC head group and the membrane normal
-decreased by 11$^\circ$ from 64$^\circ$ to 55$^\circ$
- at 620~mM buffer concentration of \ce{CaCl2},
-the mean orientation of the P-N vector in the PS head group has 
-decreased only by 3$^\circ$ from 67$^\circ$ to 64$^\circ$ 
-in the same concentration range.
-
-Such a largely diminished flexibility 
-may arise from the binding of calcium to the \ce{COO^-} group in PS. 
-Whe have measured the mean orientation of the \ce{COO^-} group 
-as the connector of the carbon atoms 
-forming the bond between the group and the $\beta$-carbon of the phospholipid 
-(Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}).
-The mean angle between the \ce{COO^-} vector and the membrane normal
-changes by $11^\circ$ from $62^\circ$ to $73^\circ$ (more towards the membrane interior)
- at 620~mM buffer concentration of \ce{CaCl2}.
-The conformational flexibility of the PS head group is hence restrained 
-by the binding of calcium cations to the \ce{COO^-} group, 
-which reorients towards the membrane interior and to the phosphate region, 
-where calcium cations bind predominantly.
-
 \begin{figure}[htbp!] 
@@
 
-Still, the dominant contribution to the binding of calcium comes from the phosphate groups. 
-Interestingly enough,
-although the \ce{COO^-} groups increase the affinity of calcium towards PS compared to PC,
- there is no extra density at the outmost surface of the membrane. 
-This can be understood from the point of the structural rearangements of the \ce{COO^-} group, 
-which take place during the interactions with calcium cations
-described above and in Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}.
-
-\todo{JOE: suggestion -- what is the charge at the membrane interface of memb+counterions -vs- memb+CaCl2 concnetration?
-                         this could relate well to the zeta potential measurements and the Fig 1 in \cite{melcrova16}. }
-
+%SAMULI: We will not go to the spin relaxation time analysis in this work
+%and average angles do not give information about dynamics
+%
+%It has been pointed out in previous experimental works
+%that the conformational flexibility of PS head groups 
+%is different from other lipids \cite{browning80}. 
+%While the static shielding tensor of PS is not substantially different from PC or PE, 
+%the $^{31}$P~$T_1$ relaxation times are significantly shorter for PS. 
+%It is suggested that the PS head group is more rigid compared to the other phospholipids, 
+%which is also corroborated by a larger chemical shielding anisotropy of PS \cite{browning80}.
+
+% Description of no. conctacts counting
+%Such a significantly increased binding also has an impact 
+%on the structure and populations of the bound calcium cations. 
+%We have counted contacts between the cations and the oxygen atoms of the lipids
+%similarly as was done in \cite{melcr18} 
+%to obtain such a detailed information 
+%about the interactions of \ce{Ca^{2+}} with various moieties in POPC or POPS.
+
+% Populations from contacts
+Interaction sites between calcium and PS lipids \cite{??} have been recently analyzed using
+different force fields and experimental techniques, but the results have been controversial.
+Simulations using the NBfix parameters for calcium \cite{kim16} in CHARMM36 force field \cite{klauda10,venable13},
+combined with 2D infrared spectroscopy, suggest that calcium ions interact only with the carboxylate group of PS lipids \cite{valentine18}.
+However, the same force field without the NBfix parameters, combined with NMR chemical shifts and
+rotational-echo double-resonance (REDOR) experiments, suggests a significant binding affinity also to the phosphate region \cite{hallock18}.
+On the other hand, simulations with the Berger force field \cite{berger97,mukhopadhyay04},
+combined with fluorescent and vibrational sum frequency spectroscopy, suggest a significant
+calcium binding also to the carbonyls in the acyl chains \cite{melcrova16}.
+The advantage of the NMR order parameters used in this work is the direct connection between
+measured and simulation data, which minimizes the human interpretation when evaluating the
+force field quality \cite{ollila16}. Therefore, we believe that the ECC-lipid model, which is in the best
+agreement with experimental headgroup order parameters among the available models, gives the currently best available
+interpretation of the experimental data.
+
+In ECC-lipid simulation of POPC:POPS (5:1) mixture with 409 mM CaCl$_2$,
+the calcium ions bind twice more likely to the carboxylate than to phosphate moiety of POPS headgroup
+\todo{I am not sure if this is correct, because I do not understand the numbers in the table.} (Table \ref{tab:binding}),
+which is in line with the results with NBfix correction in CHARMM36 \cite{hallock18}.
+However, calcium ions coordinate with even four distinct PS lipids
+in CHARMM36 simulations without NBfix \cite{hallock18}, while coordination with more than two PS lipids
+was not observed in our ECC-lipid simulations (Fig.~\ref{fig:cacl_complexes}).
+This is not surprising because calcium significantly overbinds to PC lipids in
+CHARMM36 simulations without the NBfix correction for calcium \cite{catte16}.
+Furthermore, the POPS:POPC:Ca$^{2+}$ (7:3:7) mixture with relatively high
+PS lipid fraction and calcium concentration was used in previous work \cite{hallock18},
+which may lead to the complexation of \ce{Ca^{2+}} cations and PS lipids that is
+not observed with lower PS content used in ECC-lipid simulations in this work \cite{hauser77,kurland79,hauser85,feigenson86,mattai89,roux90,roux91}.
+%While POPC interacts with the calcium cations almost entirely through its phosphate group 
+%in both neutral \cite{melcr18} and negatively charged membranes, 
+%the inteaction with POPS lipids happens more through the \ce{COO^-} group than through the \ce{PO4} group. 
+%Interactions of \ce{Ca^{2+}} with carbonyl groups are also present for both POPC and POPS,
+%however, they are always accompanied by interactions with phosphate groups. 
+Significant coordination of calcium with carbonyls groups of PS and PC lipids observed in previous
+simulations with the Berger force field \cite{melcrova16} %\cite{berger97,mukhopadhyay04}
+was not observed in ECC-lipid simulations in here nor in the previous work with pure POPC \cite{melcr18}.
+%This is in contrast to 
+%combined with fluorescent and vibrational sum frequency spectroscopy,
+%which suggest a 
+Overestimated coordination of cations with carbonyl groups is a potential source for
+overestimated cation binding in Berger force fields \cite{catte16,nmrlipids_proj4}. 
+In CHARMM36 simulations with the NBfix, calcium mainly interacted with only
+carboxylate of PS lipids  \cite{valentine18}, which is a potential source for the
+underestimated calcium binding affinity to lipid bilayer with this force field \cite{nmrlipids_proj4}.
+
+%In contrast, the combination of NMR experiments 
+%and MD simulations with the same model, CHARMM36, but without the correction for calcium binding
+%also reveal long-lived interactions of \ce{Ca^{2+}} with \ce{PO4} groups. \cite{hallock18}
+%Although the ratios of the life times of coordination follow the same trends as in our work,
+%the lengths of simulation trajectories are insufficient 
+%to yield converged results in this aspect (see SI and Refs.~\cite{catte16, melcr18, nmrlipids_proj4}). 
 \begin{table}[tb!] 
@@
 \caption{%Bulk concentrations, $C _{Ca}$, and buffer concentrations, $C' _{Ca}$, of Ca$^{2+}$;
-           %relative surface excess of calcium with respect to water ($\Gamma_{Ca}^{\rm water}$); 
-    %and
-    Percentages of the population 
-           of bound Ca$^{2+}$ to various moieties 
-           in a neutral membrane composed of POPC
-           and in a negatively charged membrane with the composition 5\,PC:1\,PS.
+  % relative surface excess of calcium with respect to water ($\Gamma_{Ca}^{\rm water}$); 
+  % and
+  Percentages of the population of bound Ca$^{2+}$ to various lipid moieties in a pure POPC bilayer with 350 mM CaCl$_2$
+  and in a POPC:POPS (5:1) mixture with 409 mM CaCl$_2$. The threshold for counting a contact was set to $0.3\,\mathrm{nm}$, which encompasses the
+  first peak of the radial distribution function between the cations and the oxygen atoms of the lipids. 
   \label{tab:binding}} 
@@
   \end{tabular} \\
-  \todo{I have commented out the surface excess values (available in text) and concentrations.}
+  \todo{I have commented out the surface excess values (available in text) and concentrations (now in caption) from the table.} \\
+  \todo{I do not understand the numbers for different moieties. Are these probabilities that the calcium binds only to these moieties and
+    nothing else? In total, 41 \% binds to PS lipids, but only 2 \% and 4 \% to phosphates and COO$^-$, respectively? Where do the rest bind?
+    I think that we should show the total probabilities to bind to different moieties.}
 \end{table}
-
-% Description of no. conctacts counting
-Such a significantly increased binding also has an impact 
-on the structure and populations of the bound calcium cations. 
-We have counted contacts between the cations and the oxygen atoms of the lipids
-similarly as was done in \cite{melcr18} 
-to obtain such a detailed information 
-about the interactions of \ce{Ca^{2+}} with various moieties in POPC or POPS. 
-The threshold for counting a contact was set to $0.3\,\mathrm{nm}$, 
-which encompasses the first peak of the radial distribution function between the cations and the oxygen atoms of the lipids.
-
-% Populations from contacts
-The percentages of the populations of membrane-bound calcium cations for various membrane moieties 
-are summarized in Table~\ref{tab:binding}.
-Even though the negatively charged membrane contains only 18\% of POPS, 
-on average 41\% of the total population of bound calcium cations is in contact with PS lipids
-with 8\% bound only to them. 
-% Where Calcium interacts
-While POPC interacts with the calcium cations almost entirely through its phosphate group 
-in both neutral \cite{melcr18} and negatively charged membranes, 
-the inteaction with POPS lipids happens more through the \ce{COO^-} group than through the \ce{PO4} group. 
-Interactions of \ce{Ca^{2+}} with carbonyl groups are also present for both POPC and POPS,
-however, they are always accompanied by interactions with phosphate groups. 
-This is in contrast to 
-simulations with the Berger force field \cite{berger97,mukhopadhyay04}
-combined with fluorescent and vibrational sum frequency spectroscopy,
-which suggest a significant
-calcium binding also to the carbonyls groups \cite{melcrova16}. 
-This feature arises likely from the inaccuracy of the Berger model 
-and its overestimated binding of cations to that region \cite{nmrlipids_proj4}.
-
-Measurements using infrared spectroscopy 
-combined with simulations with the CHARMM36 model 
-with a correction for overestimated binding of calcium \cite{valentine18}
-reveal similar trends for the populations at various PS moieties
-suggesting that POPS interacts mostly through its \ce{COO^-} group. 
-While this feature agrees nicely with our findings, 
-the balance of the interactions of PS moieties with calcium cations may be suboptimal
-and the interaction strength underestimated in total in those simulations \cite{nmrlipids_proj4}.
-
-In contrast, the combination of NMR experiments 
-and MD simulations with the same model, CHARMM36, but without the correction for calcium binding
-also reveal long-lived interactions of \ce{Ca^{2+}} with \ce{PO4} groups. \cite{hallock18}
-Although the ratios of the life times of coordination follow the same trends as in our work,
-the lengths of simulation trajectories are insufficient 
-to yield converged results in this aspect (see SI and Refs.~\cite{catte16, melcr18, nmrlipids_proj4}). 
-Moreover, the content of POPS in the bilayers in that work is
-much higher above the threshold after which 
-\ce{Ca^{2+}} cations induce lipid clustering. \cite{hauser77,kurland79,hauser85,feigenson86,mattai89,roux90,roux91}
-At such a high content of negatively charged lipids,
-PS head groups are found to form even four-membered clusters \cite{hallock18}. 
-Our simulations are, however, at different conditions,
-in which the clustering of POPS does not happen, 
-and PS lipids prefer 1:1 ratio with \ce{Ca^{2+}}
-as plotted in Fig.~\ref{fig:cacl_complexes}
-showing relative probabilities of \ce{Ca^{2+}} complexes with a certain number of PC or PS lipids. 
-Compared to a neutral POPC bilayer
-the negatively charged bilayer has its stoichiometry 
-shifted towards complexes with three phospholipids and one calcium. 
-This is obviously due to the presence of POPS, 
-which also contributes to the pre-formed one- and two-membered clusters of POPC. 
-Clusters of four or more lipids were not observed in either membrane.
-
-\todo{ THE REMAINING PIECES TO BE DISCUSSED  --  consider wheter the dynamics slowdown or the dihedral distributions could make good contributions to this paper. }
-
-% Ca2+ induces much slower dynamics of lipids when PS is present,
-% this does not happen with PC
-% uses NBFIX
-From Ref.~\cite{valentine18}://
-To summarize, calcium binding to PS lipids slows the dynamics at the ester position and introduces more heterogeneous molecular environments. 
-The slowdown cannot be attributed to complete ester dehydration or decreased hydrogen bonding to water, so we interpret these results as evidence that the lipid headgroups adopt more rigid but disordered conformations on the molecular level.
-The decrease in dynamics observed experimentally indicates that the network of hydrogen bonds between lipids and water at the polar-nonpolar interface becomes slower, as it was shown by Karathanou and Bondar (34).
-This is bolstered by recent NMR work showing that binding to Ca2+ induces two different rigid conformations in the glycerol backbone of lipid headgroups (18).
-We have investigated the molecular origin of the rigidification and decreased lateral diffusion, finding that calcium binding does not dehydrate the ester region, but even with only 10\% PS lipids, the local dynamics slow around the esters.
-The slower local dynamics provide a molecular explanation for the Ca2+-dependent decreases in fluidity observed on longer length and timescales.
-There is considerable room for further improvement in the force fields for ion-membrane interactions. Particularly, small differences in short-range interactions may determine the local geometries (34
-).
-Most of our conclusions are, however, based on long-range effects (i.e., electric field generated by the ion at significant distances from the carbonyl groups), which are less sensitive to the details of the force field.
-
-Their Figure 5 \cite{hallock18} shows the distributions of O-Ca-Cb-N dihedral angle
-revealing separate populations of the PS head groups. 
-These two structures are also compared in therm of P-Cg distances, 
-which can be measured from NMR REDOR exp. 
-It is concluded that since the two distinct populations in P-Cg distance can be represented with the 
-dihedral distributions, 
-they also likely represent the distinct populations. 
- -> plotting the distribution of O-Ca-Cb-N dihedral angle may be also a good idea for us (SI?)
-     -> NMRlipids 4 in SI \cite{nmrlipids_proj4} plots this information: 
-           - C36 agress with the distribution in \cite{hallock18}
-           - Lipid17 (and Slipids) are almost identical to C36
-           - but the structure of the COO- is highly different (dih angle Ca-Cb-Cg-Og)
 
@@
   \caption{\label{fig:cacl_complexes} 
-      Relative probabilities of existence of \ce{Ca^{2+}} complexes 
-      with a certain number of lipids.  
-      All lipids were taken into account with the exception of the complexes in light green and orange, 
-      for which we counted only contacts with POPC resp POPS from the mixed 5\,PC:1\,PS negatively charged bilayer. 
-      The calculated probabilities of the calcium-lipid complexes also reflect only POPC (light green) resp POPS (orange). 
-      Probabilities were taken from simulations with comparable bulk concentrations of calcium around 250~mM. 
+    Relative probabilities of \ce{Ca^{2+}} ions to coordinate with a certain number of lipids
+    in pure POPC bilayer with 350 mM CaCl$_2$ and POPC:POPS (5:1) mixture with 400 mM CaCl$_2$.  
+    Analysis was done for both systems by considering all lipids (blue and violet) and
+    for POPC:POPS (5:1) mixture also by considering POPC and POPS lipids separately (green and orange).
+    %All lipids were taken into account with the exception of the complexes in light green and orange, 
+    %for which we counted only contacts with POPC resp POPS from the mixed 5\,PC:1\,PS negatively charged bilayer. 
+    %The calculated probabilities of the calcium-lipid complexes also reflect only POPC (light green) resp POPS (orange). 
+    % Probabilities were taken from simulations with comparable bulk concentrations of calcium around 250~mM. 
     Clusters of four or more lipids were not observed in either membrane.
-      Simulation data for POPC bilayers were taken directly from \cite{melcr18}. 
+    Previously published simulation data \cite{melcr18} for pure POPC bilayers were taken directly from ?? Add Zenodo reference, if available. 
   }
+  \todo{Add definition used for coordination and link to the Zenodo, if available.}
 \end{figure}
+
+In ECC-lipid simulations, on average 41\% of the total population of bound calcium cations is in contact with PS lipids
+with 8\% bound only to them, even though the negatively charged membrane contains only 18\% of POPS (Table \ref{tab:binding}).
+Therefore, calcium ions bind more likely to negatively charged PS lipids than neutral PC lipid in a mixed bilayer, as expected.
+In PS lipids, the calcium binds both phosphate and carboxylate, but probability for the latter is twice higher.
+In PC lipids, the calcium binds almost always to phosphate, but in some cases also to carbonyl, as also observed previously in pure POPC simulations \cite{melcr18}.
+%Still, the dominant contribution to the binding of calcium comes from the phosphate groups. 
+%Interestingly enough,
+%although the \ce{COO^-} groups increase the affinity of calcium towards PS compared to PC,
+% there is no extra density at the outmost surface of the membrane. 
+%This can be understood from the point of the structural rearangements of the \ce{COO^-} group, 
+%which take place during the interactions with calcium cations
+%described above and in Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}. 
+The average P--N vector orientation of both POPC and POPS headgroups tilts more perpendicular to the membrane surface
+upon addition of 620~mM CaCl$_2$  to POPC:POPS (5:1) mixture by  11$^\circ$ and  3$^\circ$, respectively (Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}).
+The average orientation of the PS headgroup C$_{\beta}$--C$_{\gamma}$ vector tilts to the opposite direction
+by $11^\circ$ in the same system.
+%changes by $11^\circ$ from $62^\circ$ to $73^\circ$ (more towards the membrane interior)
+% at 620~mM buffer concentration of \ce{CaCl2}.
+% We have investigated this effect using the mean orientations of P-N and \ce{COO^-} vectors 
+%(defined and plotted in Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}). 
+%While the mean angle between the P-N vector of PC head group and the membrane normal
+%decreased by 11$^\circ$ from 64$^\circ$ to 55$^\circ$
+% at 620~mM buffer concentration of \ce{CaCl2},
+%the mean orientation of the P-N vector in the PS head group has 
+%decreased only by 3$^\circ$ from 67$^\circ$ to 64$^\circ$ 
+%in the same concentration range. 
+%Such a largely diminished flexibility 
+%may arise from the binding of calcium to the \ce{COO^-} group in PS. 
+%Whe have measured the mean orientation of the \ce{COO^-} group 
+%as the connector of the carbon atoms 
+%forming the bond between the group and the $\beta$-carbon of the phospholipid 
+%(Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}).
+Therefore, we conclude that the PS headgroup orientation is less sensitive to the bound calcium than PC.
+%by the binding of calcium cations to the \ce{COO^-} group, 
+%which reorients towards the membrane interior and to the phosphate region, 
+%where calcium cations bind predominantly. 
+90\% of the calcium residence times are shorter than $60\,\mathrm{ns}$ for pure POPC bilayer
+and shorter than $200\,\mathrm{ns}$ for POPC:POPS (5:1) mixture, while the
+longest observed residence times are $141\,\mathrm{ns}$ and $485\,\mathrm{ns}$, respectively (Fig.~\ref{fig:hist_residence_times}),
+suggesting that the calcium residence times are 3-4 times longer in POPC:POPS (5:1) mixture than in
+pure POPC. Athough the ECC-lipid model gives the most realistic picture of calcium binding to
+lipid bilayers containing PS lipids with respect to NMR order parameters, it does not
+correctly capture the rapid changes with small concentrations. Therefore, further
+force field optimization is required for fully realistic MD simulations of cations
+in the vicinity of negatively charged lipid bilayers.
+
+%
+% SAMULI: We have problems to find time to finish this paper even with the current data, we should not add anything that is not necessary to finish this asap. 
+%
+%\todo{JOE: suggestion -- what is the charge at the membrane interface of memb+counterions -vs- memb+CaCl2 concnetration?
+% this could relate well to the zeta potential measurements and the Fig 1 in \cite{melcrova16}. }
+% 
+%\todo{ THE REMAINING PIECES TO BE DISCUSSED  --  consider wheter the dynamics slowdown or the dihedral distributions could make good contributions to this paper. }
+
+%% Ca2+ induces much slower dynamics of lipids when PS is present,
+%% this does not happen with PC
+%% uses NBFIX
+%From Ref.~\cite{valentine18}://
+%To summarize, calcium binding to PS lipids slows the dynamics at the ester position and introduces more heterogeneous molecular environments. 
+%The slowdown cannot be attributed to complete ester dehydration or decreased hydrogen bonding to water, so we interpret these results as evidence that the lipid headgroups adopt more rigid but %disordered conformations on the molecular level.
+%The decrease in dynamics observed experimentally indicates that the network of hydrogen bonds between lipids and water at the polar-nonpolar interface becomes slower, as it was shown by Karatha%nou and Bondar (34).
+%This is bolstered by recent NMR work showing that binding to Ca2+ induces two different rigid conformations in the glycerol backbone of lipid headgroups (18).
+%We have investigated the molecular origin of the rigidification and decreased lateral diffusion, finding that calcium binding does not dehydrate the ester region, but even with only 10\% PS lip%ids, the local dynamics slow around the esters.
+The slower local dynamics provide a molecular explanation for the Ca2+-dependent decreases in fluidity observed on longer length and timescales.
+%There is considerable room for further improvement in the force fields for ion-membrane interactions.
+Particularly, small differences in short-range interactions may determine the local geometr%ies (34
+%).
+Most of our conclusions are, however, based on long-range effects (i.e., electric field generated by the ion at significant distances from the carbonyl groups), which are less sensitive to t%he details of the force field.
+
+%Their Figure 5 \cite{hallock18} shows the distributions of O-Ca-Cb-N dihedral angle
+%revealing separate populations of the PS head groups. 
+%These two structures are also compared in therm of P-Cg distances, 
+%which can be measured from NMR REDOR exp. 
+%It is concluded that since the two distinct populations in P-Cg distance can be represented with the 
+%dihedral distributions, 
+%they also likely represent the distinct populations. 
+% -> plotting the distribution of O-Ca-Cb-N dihedral angle may be also a good idea for us (SI?)
+%     -> NMRlipids 4 in SI \cite{nmrlipids_proj4} plots this information: 
+%           - C36 agress with the distribution in \cite{hallock18}
+%           - Lipid17 (and Slipids) are almost identical to C36
+%           - but the structure of the COO- is highly different (dih angle Ca-Cb-Cg-Og)
 
@@
 \listoftodos
-
-\section{SI}
-
-Timescales associated with the binding of calcium cations from solution to the membrane
-are plotted for each binding event as a histogram in Fig.~\ref{fig:hist_residence_times}. 
-Using these plots, we can estimate that 90\% of the residence times of any calcium cation 
-will be lower than $60\,\mathrm{ns}$ for pure POPC neutral bilayer 
-and shorter than $200\,\mathrm{ns}$ for the mixed 5\,PC:1\,PS negatively charged bilayer. 
-The longest observed residence times in the simulations were $141\,\mathrm{ns}$ for the neutral membrane 
-and $485\,\mathrm{ns}$ for the negatively charged membrane. 
-Both estimates of the residence times come from simulations with comparable concentrations of around $250\mathrm{mM}$;
-the simulation with the neutral membrane has a bulk concentration of calcium $C_{ion} = 280\mathrm{mM}$, 
-whereas the simulation with the negatively charged membrane has a bulk concentration of calcium $C_{ion} = 240\mathrm{mM}$.
-
-\begin{figure}[tb!]
-  \centering
-  \includegraphics[width=\figwidth]{../img/histogram_bound_times_26CaCl2_comparison_PC-PCPS.pdf}
-  \caption{\label{fig:hist_residence_times}
-   Histograms of residence times of \ce{Ca^{2+}} 
-   in a neutral membrane composed of POPC (orange)
-   and in a negatively charged membrane with the composition 5\,PC:1\,PS (blue)
-   from simulations with ECC-lipids and ECC-ions.
-   The simulation with the neutral membrane has a bulk concentration of calcium $C_{ion} = 280\mathrm{mM}$, 
-   the simulation with the negatively charged membrane has a bulk concentration of calcium $C_{ion} = 240\mathrm{mM}$. 
-   In the simulation with the neutral membrane, 
-   90\% of the residence times of calcium cations are
-   shorter than $60\,\mathrm{ns}$, % exactly $53\,\mathrm{ns}$                                                                          
-   with the longest observed residence time being $141\,\mathrm{ns}$. 
-   In the simulation with the negatively charged membrane, 
-   90\% of the residence times of calcium cations are
-   shorter than $200\,\mathrm{ns}$, % exactly $53\,\mathrm{ns}$                                                                          
-   with the longest observed residence time being $485\,\mathrm{ns}$. 
-   Simulation data for POPC bilayers were taken directly from \cite{melcr18}. 
-   }
-\end{figure}
-
-In summary, the results from ECC-lipids suggest 
-that the exchange of calcium between the POPC bilayer and the solvent 
-occurs at the order of $\sim$10--100~ns, 
-which is significantly faster than observed in simulations 
-with other presently available non-polarizable models of lipids and ions~\citep{javanainen17, catte16}. 
-Our results suggest 
-that simulation trajectories with a characteristic length of several hundreds of nanoseconds 
-are necessary to capture the binding of calcium to neutral POPC bilayers 
-in equilibrium when more realistic \emph{polarizable} force fields are used. 
-Interestingly enough, almost an order of magnitude longer simulations
-are required for the negatively charged bilayers.
 

### 2019-03-13 Manuscript/manuscript.tex
Upodate manuscript

@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
 %% This is a (brief) model paper using the achemso class 
@@
   \caption{\label{fig:delta_ordPar_NaCl_PCPS} 
-    Changes of the head group order parameters, the P--N and C$_\beta$--C$_\gamma$ vector angles (stars)
+    Changes of the head group order parameters, and the angles of P--N and C$_\beta$--C$_\gamma$ (stars) vectors
     with respect to the membrane normal of POPC (left) and POPS (right) in a POPC:POPS (5:1) bilayer 
-    as a function of \ce{NaCl} concentration from ECC-lipid and Lipid17/Dang simulations at 298 K \cite{roux90}.
-    Because data with \ce{NaCl} are not available for POPC, 
-    we show experimental data for \ce{LiCl} (dashed line, left) 
-    as an upper bound for the magnitude of the response to \ce{NaCl}, 
-    which has a lower affinity to phospholipid bilayers compared to \ce{LiCl} \citep{roux90}. 
+    as a function of \ce{NaCl} concentration from ECC-lipid and Lipid17/Dang simulations at 298 K.
+    The y-axis for the $\alpha$-carbon results of POPS (middle right) is transferred
+    with the same value for both order parameters such that the lower order
+    parameter value from pure POPS is at zero to correctly illustrate the significant forking.
+    Because experimental data with \ce{NaCl} is not available for POPC, the data for KCl and \ce{LiCl} (dashed line, left)
+    are shown as lower and upper bounds for the response to \ce{NaCl}, respectively.
+    The magnitudes of experimental order parameters are from Ref. \citenum{roux90} and signs from Ref. \citenum{ferreira16}.
+    %which has a lower affinity to phospholipid bilayers compared to \ce{LiCl} \citep{roux90}. 
     %The orientation of the \ce{COO^-} group is defined as 
@@
   }
-  \todo{We should plot the experimental results with KCl also in this figure.} \\
+  \todo{We should plot also the experimental results for POPC with KCl in this figure for the lower bound.} \\
   \todo{Why there are no error bars for ECC-POPS in this and other figures?} \\
@@
   \caption{\label{fig:delta_ordPar_KCl_PCPS}
-    Changes of the head group order parameters $\alpha$, $\beta$ and the orientations of the carboxylate group and the P-N vector  
-    of POPC (left) and POPS (right) phospholipids in a POPC:POPS 5:1 bilayer 
-    as a function of \ce{KCl} concentration in the buffer 
-    from simulations with different force fields and experiments at 298 K. \citep{roux90}
-    The orientation of the \ce{COO^-} group is defined as 
-    the connector from the $\beta$ carbon to the carbon in \ce{COO^-} (stars, bottom right). 
+    Changes of the head group order parameters, and the angles of P--N and C$_\beta$--C$_\gamma$ (stars) vectors 
+    with respect to the membrane normal of POPC (left) and POPS (right) in a POPC:POPS (5:1) bilayer 
+    as a function of \ce{KCl} concentration from ECC-lipid and Lipid17/Dang simulations 
+    compared with experimental values from Ref. \citenum{roux90} (signs from Ref. \citenum{ferreira16}) at 298 K.
+    The y-axis for the $\alpha$-carbon results of POPS (middle right) is transferred
+    with the same value for both order parameters such that the lower order
+    parameter value from pure POPS is at zero to correctly illustrate the significant forking.
   }
@@
 
-\begin{figure}[tbp!] 
-  \centering 
-  \includegraphics[width=\figwidth]{../img/ecc_pops/density_profiles_na_k_cl_wat_phos_models-compar_5-6_NaCl-and-KCl-series.pdf}
-  \caption{\label{fig:nacl-dens_PCPS} 
-    Number density profiles of \ce{K^{+}}, \ce{Na^{+}} and \ce{Cl^-} along the membrane normal axis 
-    for the negatively charged membranes with the composition of 5\,PC:1\,PS using ECC-lipids and ECC-ions
-    from simulations with added concentrations of \ce{KCl} or \ce{NaCl}.  
-    The top    profile shows the simulation with an additional \ce{KCl} concentration and \ce{Na^+} counterions. 
-    The bottom profile shows the simulation with an additional \ce{NaCl} concentration and \ce{Na^+} counterions, which are not distinguished from the added salt. 
-    The density profiles of phosphate groups and water are divided by 4 and 100, respectively.  
-  }
-  \todo{I am not sure if we need this figure at all or should we move it to SI.}
-  \todo{The concentrations in the figure do not match with the table.}
-\end{figure}
-
 % This§: 
@@
 %without the necessity of defining a Gibbs dividing surface. \citep{melcr18, chattorajBOOK}
-The value of $\Gamma ^{w} _{\rm Na}=-0.11 \pm 0.01$ nm$^{-2}$ for 1 M sodium in pure POPC, calculated from
-ECC-POPC simulation, was reported in the previous work \cite{melcr18}. Here, the ECC-lipid simulations
-give
-%$\Gamma ^{w} _{\rm K}=-0.12 \pm ??$ nm$^{-2}$ and
-$\Gamma ^{w} _{\rm Na}=0.09 \pm ??$ nm$^{-2}$ \todo{Error bar should be added.} for 0.621 M sodium in pure POPC:POPS (5:1) mixture.
-Therefore, the presence of $\sim$17\% POPS in POPC bilayer increases the relative surface
-excess of sodium with respect to water from negative to positive. As discussed above,
+The value of $\Gamma ^{w} _{\rm Na}=-0.11 \pm 0.01$ nm$^{-2}$ for 1 M sodium in pure POPC
+was reported from a ECC-POPC simulation in the previous work \cite{melcr18}.
+The presence of $\sim$17\% POPS in POPC bilayer increases the
+%relative surface excess of sodium with respect to water from negative
+value to $\Gamma ^{w} _{\rm Na}=0.09 \pm ??$ nm$^{-2}$ \todo{Error bar should be added}
+for 0.621 M sodium, while value for potassium remains negative,
+$\Gamma ^{w} _{\rm K}=-0.12 \pm ??$ nm$^{-2}$ \todo{Error bar should be added} in POPC:POPS (5:1) mixture (Fig. \ref{fig:nacl-dens_PCPS}).
+As discussed above,
 the sodium binding to POPC:POPS (5:1) mixture may be slightly overestimated also in
@@
   \caption{\label{fig:delta_ordPar_CaCl_PCPS} 
-    Changes of the head group order parameters $\alpha$, $\beta$ and the orientations of the carboxylate group and the P-N vector  
-    of POPC (left) and POPS (right) phospholipids in a POPC:POPS 5:1 bilayer as a function of \ce{CaCl2} concentration 
-    in bulk ($C_{ion}$) from simulations with different force fields and experiments at 298~K. \citep{roux90}
-    The orientation of the \ce{COO^-} group is defined as 
-    the connector from the $\beta$ carbon to the carbon in \ce{COO^-} (stars, bottom right).
+    Changes of the head group order parameters, and the angles of P--N and C$_\beta$--C$_\gamma$ (stars) vectors 
+    with respect to the membrane normal of POPC (left) and POPS (right) in a POPC:POPS (5:1) bilayer 
+    as a function of \ce{CaCl} concentration from ECC-lipid and Lipid17/Dang simulations 
+    compared with experimental values from Ref. \citenum{roux90} (signs from Ref. \citenum{ferreira16}) at 298 K.
+    The y-axis for the $\alpha$-carbon results of POPS (middle right) is transferred
+    with the same value for both order parameters such that the lower order
+    parameter value from pure POPS is at zero to correctly illustrate the significant forking.
   }
@@
   }
-  \todo{SAMULI: We should compare to the density profiles from Lipid17 simulations. These are also needed for NMRlipids IV.}
+  \todo{SAMULI: We should compare to the density profiles from Lipid17 simulations.}
 \end{figure}
@@
     Relative probabilities of \ce{Ca^{2+}} ions to coordinate with a certain number of lipids
-    in pure POPC bilayer with 350 mM CaCl$_2$ and POPC:POPS (5:1) mixture with 400 mM CaCl$_2$.  
+    in pure POPC bilayer with 350 mM CaCl$_2$ and in POPC:POPS (5:1) mixture with 400 mM CaCl$_2$.  
     Analysis was done for both systems by considering all lipids (blue and violet) and
@@
   }
-  \todo{Add definition used for coordination and link to the Zenodo, if available.}
+  \todo{Add definition used for the coordination} \\
+  \todo{Add link to the Zenodo, if available.}
 \end{figure}

### 2019-03-14 Manuscript/manuscript.tex
Update text and references

@@
 
-Lipids with negatively charged headgroups, like phophatidylserine
-play important role in biomembranes. 
-Signaling molecules in cells interact with lipid membranes 
-due to specific, often electrostatically driven, interactions 
-with anionic lipids \cite{lemmon08,leventis10}.
-Anionic lipids and their interactions with ions play also role in membrane
-fusion \cite{??}. \todo{Pavel can probably improve this and give a citation}
-These interactions depend both on
-the chemical details of lipid headgroups as well as 
-on the binding of counter and co-ions to anionic membranes.
-
-Various interaction schemes have been proposed for the negatively charged
-lipids with ions and other lipids. 
+The most common negatively charged lipids in eukaryotic membranes bearing
+phophatidylserine (PS) headgroup are important signaling molecules \cite{lemmon08,leventis10,li14}.
+They interact with signaling proteins \cite{leventis10},
+regulate surface charge and protein localization \cite{yeung08}, 
+induce protein aggregation \cite{zhao04,gorbenko06} and membrane fusion \cite{??}.
+Because these functions are also often regulated by biological cations \cite{leventis10},
+the detailed interactions between PS lipids and biological cations, like calcium, are under great interest.
+\todo{Pavel can probably improve and complement this}
+
+%Various interaction schemes have been proposed for the negatively charged
+%lipids with ions and other lipids. 
 Some studies suggest that the
@@
 
-\todo{Joe: this paragraph requires editing.}
-The lack of electronic polarizability has been a concern since the
-early days of classical MD simulations of lipids \cite{jonsson86,egberts94, berendsen1996},
-where the atomic charges were scaled by a factor of 1/2 to effectively
-include the electrostatic screening from the missing electronic polarizability.
-The electronic continuum correction (ECC) is a more sophisticated version of the charge scaling.
-
-The polarization of electrons in ECC-lipids uses the implicit method termed electronic continuum correction (ECC). \cite{leontyev09, Pluharova2014, martinek17, melcr18} 
-In this method, a system of polarizable particles is represented 
-with an equivalent system of particles with fixed charges,
-which account for the electronic polarizability \emph{implicitly}. 
-This is practically achieved by rescaling the charges with a pre-determined scaling factor, $f_q$,
-which is derived from the high-frequency dielectric constant of electrons, $\epsilon_{el}$, as
-\begin{equation}
-  f_q = \frac{1}{\sqrt{\epsilon_{el}}} ,
-\end{equation}
-The value of $\epsilon_{el} = 1.78$ for water solutions gives the scaling
-factor of $f_q \approx 0.75$ for aquaeous environments \cite{leontyev09}.
-This has been recently applied to improve classical models for ions and
-other biomolecules \cite{Pluharova2014, martinek17, melcr18}
-\todo{Pavel can probably update this (insulin?)}.
-
-Recently we applied the ECC also to the Amber based lipid14 force field \cite{dickson14} to
-improve the cation binding affinity to zwitterionic POPC lipid bilayers \cite{melcr18}.
-In contrast to the total charge of ions, 
-the partial charges in force fields of zwitterionic molecules
-depend on the methods used to derive the force field, 
-which may already partially include the
-electronic polarizability effects. Therefore, the optimal scaling factor for charges in
-zwitterionic lipids may vary from (be even larger) than the theoretically derived $f_q \approx 0.75$.
-In the recent work, we found optimal results for cation binding affinity and
-bilayer structural properties when the partial charges and Lennard-Jones~$\sigma$
-parameters of all non-tail atoms of the lipid14 POPC model
-(i.e., headgroup, glycerol backbone, and carbonyl regions) 
-were scaled with the factors $f_q = 0.8$ and $f_\sigma = 0.89$  \cite{melcr18}.
-
-Here, we apply the ECC to the Amber based lipid17 parameters \cite{lipid17-future} of POPS lipid with
-a total monovalent negative charge. 
-In this case we treat the POPS lipids like ions
-applying to it the scaling factor of $f_q = 0.75$,
-as it carries a physical total charge similar to the ions in solution, 
-and the lipid charges has to neutralize the counterion charges in simulations.
-For the Lennard-Jones~$\sigma$ parameters, we used the same scaling factor of $f_\sigma = 0.89$
-as in the ECC-POPC model \cite{melcr18}.
-%we used exactly the pre-determined scaling factor 
+The electronic continuum correction (ECC) is a simple method to implicitly include
+the electronic polarizability into classical molecular dynamics (MD) simulations
+by scaling the electric charges defined in force fields.
+%The polarization of electrons in ECC-lipids
+%uses the implicit method termed electronic continuum correction (ECC). \cite{leontyev09, Pluharova2014, martinek17, melcr18} 
+%In this method, a system of polarizable particles is represented 
+%with an equivalent system of particles with fixed charges,
+%which account for the electronic polarizability \emph{implicitly}.
+For simple ions in water, the electronic polarizability
+can be taken into account by scaling the charge with a constant factor 
+%This is practically achieved by rescaling the charges with a pre-determined scaling factor, $f_q$,
+%\begin{equation}
+ $ f_q = \frac{1}{\sqrt{\epsilon_{el}}} \approx 0.75$,
+%\end{equation}
+where  $\epsilon_{el} = 1.78$ is the high-frequency dielectric constant of electrons in water \cite{leontyev09}.
+%,, as
+%The value of $\epsilon_{el} = 1.78$ for water solutions gives the scaling
+%factor of $f_q \approx 0.75$ for aquaeous environments \cite{leontyev09}.
+The scaling of charges with this factor has been recently applied to improve classical models for ions and
+other biomolecules \cite{Pluharova2014, martinek17} \todo{Pavel can probably update this (insulin?)}.
+In addition to the charge, the Lennard-Jones~$\sigma$ was scaled with a factor of $f_\sigma = ??$
+to improve the hydration properties of ions with respect to scattering data in these studies
+\todo{Pavel can probably deliver the state of the art information about these parameters and sharpen the discussion}.
+
+Derivation of the correct scaling factor for lipids is more complicated, because
+%In contrast to the total charge of ions,
+the partial charges depend on the methods used to derive the force field.
+For example, the partial charges may already include some electronic polarizability effects
+and the scaling factor should then be larger than the theoretically derived value for ions in water, $f_q \approx 0.75$.
+Before the ECC theory was rigorously derived, similar idea was employed in early classical 
+MD simulations of lipids and surfactants using the scaling factor of 0.5 for atomic charges  \cite{jonsson86,egberts94, berendsen1996}. 
+In our recent work, we applied the ECC to the Amber based lipid14 force field of zwitterionic POPC lipid \cite{dickson14}
+by scaling the partial charges and Lennard-Jones~$\sigma$ parameters of headgroup, glycerol backbone,
+and carbonyl regions \cite{melcr18}. The scaling factors were optimized to reproduce 
+the calcium binding affinity to POPC lipid bilayers, evaluated using the headgroup order parameters
+and the electrometer concept \cite{akutsu81,altenbach84,seelig87,catte16}, and the X-ray scattering form factor
+without additional ions \cite{kucerka11}, which resulted to the scaling factor values of $f_q = 0.8$ and $f_\sigma = 0.89$  \cite{melcr18}.
+
+Here, we apply the ECC to the Amber based lipid17 parameters \cite{lipid17-future} of POPS lipid with a monovalent negative charge.
+Because POPS carries a physical total charge as ions in water, we apply the theoretically derived scaling factor for
+ions $f_q = 0.75$ \cite{leontyev09} to the partial charges of POPS. The lipids with a total charge of -0.75 also neutralize
+the counterion charges of +0.75 in simulations with ECC-ions \cite{kohagen16}.
+Following our previous work for POPC, we use the scaling factor of $f_\sigma = 0.89$ for the Lennard-Jones~$\sigma$ parameters and
+scale the parameters of atoms only in the headgroup, glycerol backbone, and carbonyl regions \cite{melcr18}.
+Furhter optimization of parameters was not done in this work.
+% we used exactly the pre-determined scaling factor 
 %to represent the effects of electronic polarization
 %yielding a total charge of $-0.75$ in the ECC formalism. \cite{martinek17, leontyev09, melcr18}
-Similarly to the ECC-POPC model \cite{melcr18}, we applied the ECC only to the
-non-tail atoms, i.e., to atoms in the headgroup, glycerol backbone, and carbonyl regions. 
 %polar parts of the phospholipid,
@@
 A R-type Proton Detected Local Field (R-PDLF) experiment was performed to determine C--H bond order
-parameters, $S_{\rm{CH}}=1/2\langle3cos^2\theta-1\rangle$, where $\theta$ denotes the angle of the C--H bond
+parameters
+\begin{equation}\label{OPequation}
+  S_{\rm{CH}}=\frac{1}{2}\langle3\cos^2\theta-1\rangle,
+\end{equation}
+  where $\theta$ denotes the angle of the C--H bond
 with the bilayer normal and the angular brackets define a time average on a time scale of approximately 1 microsecond.
@@
 equipped with a standard 4 mm CP-MAS HXY probe. The set up of the R-PDLF experiment was the following
-(using the notation as in figures 1c and 2c of the original reference for the R-PDLF
-experiment [S. V. Dvinskikh, H. Zimmermann, A. Maliniak and D. Sandström, J. Mag. Res. 168, 194 (2004)]).
+(using the notation from figures 1c and 2c in the original publication describing the R-PDLF experiment \cite{dvinskikh04}).
 The magic angle spinning (MAS) frequency used was 5.15 kHz. The recoupling pulses in the R18 blocks had
@@
 resolution of +- 0.01. The C--H bond order parameter determined from a given dipolar
-splitting, $\Delta\nu$ (see e.g. Figure XX as SI) is equal to $\Delta\nu/(0.315\times21.5 kHz)$, where 0.315 is
-the scaling factor of the R18 recoupling sequence and 21.5 kHz is the maximum 1H-13C dipolar coupling for a C--H bond.
-The refocused-INEPT transfer [G. Morris and R. Freeman, J. Am. Chem. Soc. 101, 760 (1979); D. Burum, R. Ernst, J. Mag. Res. 39, 163 (1980)]
-was used for transferring the polarization from 1H nuclei to the covalently bond 13C nuclei with the delays $\tau_1$ = 1.94 ms and $\tau_2$ = 0.97 ms
+splitting, $\Delta\nu$ (see e.g. Fig. \ref{R-PDLFslices} \todo{Is this reference to the correct figure?})
+is equal to $\Delta\nu/(0.315\times21.5 kHz)$, where 0.315 is
+the scaling factor of the R18 recoupling sequence and 21.5 kHz is the maximum $^1$H-$^{13}$C dipolar coupling for a C--H bond.
+The refocused-INEPT transfer \cite{morris79,burum80} %[G. Morris and R. Freeman, J. Am. Chem. Soc. 101, 760 (1979);
+\todo{ I could not find this reference [D. Burum, R. Ernst, J. Mag. Res. 39, 163 (1980)]}
+was used for transferring the polarization from $^1$H nuclei to the covalently bond $^{13}$C nuclei with the delays $\tau_1$ = 1.94 ms and $\tau_2$ = 0.97 ms
 (set as multiples of the MAS rotation period). The pulses for the refocused-INEPT sequence had a nutation frequency equal to 63.45 kHz.
-For acquiring the spectra, a total number of 1024 transients were recorded for each point in the indirect dimension, using an acquisition time of 0.1 s under SPINAL64 1H decoupling [B.
-Fung, A.
-Khitrin, K.Ermolaev, J.
-Mag.
-Res.
-142, 97 (2000)] with a nutation frequency of 50 kHz, and with dwell time giving a spectral width of 200 ppm for the 13C chemical shift dimension.
-The set of free induction decays recorded were then processed as described e.g.
-in [T.
-M.
-Ferreira, B.
-Medronho, R.
-W.
-Martin, D.
-Topgaard, Phys.
-Chem.
-Chem.
-Phys.
-10, 6033 (2008)].
-For assigning the dipolar splittings in the crowded spectral region between 29 and 30 ppm both the assignments made for POPC in [ T.
-M.
-Ferreira, F.
-Coreta- Gomes, O.
-H.
-S.
-Ollila, M.
-J.
-Moreno, W.
-L.
-C.
-Vaz, D.
-Topgaard, Phys.
-Chem.
-Chem.
-Phys.
-15, 1976-1989 (2013)] and the POPS order parameters simulated and reported in this work were taken as a reference.
+For acquiring the spectra, a total number of 1024 transients were recorded for each point in the indirect dimension, using an acquisition
+time of 0.1 s under SPINAL64 $^1$H decoupling \cite{fung00} with a nutation frequency of 50 kHz, and with dwell time giving a spectral
+width of 200 ppm for the $^{13}$C chemical shift dimension. The set of free induction decays recorded were then processed as
+described e.g. in Ref. \citenum{ferreira08}. The dipolar splittings in the crowded spectral region
+between 29 and 30 ppm were assigned based on the previous assignments for POPC \cite{ferreira13} 
+and the POPS order parameters calculated from simulations in this work.
 
@@
 using the GROMACS 2018 \cite{Abraham15} simulation package. 
-%The simulated system consisted of lipids (POPC and POPS) forming a single lipid bilayer,
-%an amount of water molecules comparable to the hydration levels in experiments, 
-%\todo{Check this claim. SAMULI: Check discussin in the NMRlipids IV manuscript}
-The complete set of simulated systems is listed in Table~\ref{tbl:sim-list},
-and the used simulation parameters are in Table~\ref{tbl:mdpar}. 
+The simulated systems are listed in Table~\ref{tbl:sim-list}
+and the used simulation parameters in Table~\ref{tbl:mdpar}. 
 All simulations were ran for a minimum of $1 \, \mu$s,
@@
 
-\todo{EDIT: Taken from NMRlipids4 as an interesting information, to be edited.}
-In the experimental $^2$H NMR literature data used in this work~\cite{browning80,roux90},
-the lipids were first soluted to the buffer and then centrifuged to a pellet which
-was used in the measurements. Such samples have lower lipid concentration
-(approximately 10~wt~\% of lipids~\cite{browning80,roux88,roux90}) than 
-gravimetric samples (60~wt~\%) and simulations (approximately 50-60~wt~\%) in this work.
-Larger multilamellar repeat distances are expected in the samples with lower lipid
-concentrations due to the swelling caused by electrostatic repulsion in pure PS lipid systems~\cite{millman82}.
-However, the PS headgroup order parameters measured from gravimetric sample in this work
-are in good agreement with the results from centrifuged sample in the literature \cite{browning80} (Fig. \ref{HGorderParameters}).
-Furthermore, the equlibrium repeat distance rapidly decreases with the addtion of monovalent
-salts and is close to the simulation box sizes already above 500~mM concentrations~\cite{millman82,rand89}.
-Therefore, the hydration levels of multilamellae are expected to be sufficiently similar
-in the used simulations and reference experiments.
-
 In the reference simulations with the standard Amber force field,
-we used the Lipid14 parameters~\cite{dickson14} for POPC, the Lipid17 parameters for POPS \cite{lipid17-future},
-the TIP3P water model~\cite{jorgensen83}, and ion models by Dang and coworkers~\cite{smith94,chang1999,dang2006}.
-The Lipid14 parameters in Gromacs format were downloaded from Ref.~\citenum{lipid14files}. 
+the Lipid14 parameters~\cite{dickson14} for POPC, the Lipid17 parameters for POPS \cite{lipid17-future},
+the TIP3P water model~\cite{jorgensen83}, and ion models by Dang and coworkers~\cite{smith94,chang1999,dang2006} were used.
+The previously generated \cite{botan15} Lipid14 parameters in Gromacs format were downloaded from Ref.~\citenum{lipid14files}. 
 The Lipid17 parameters were obtained from AmberTools18 \cite{amber18} 
-and they were converted to Gromacs format using acpype \cite{acpype}.  
+and converted to Gromacs format using acpype \cite{acpype}.  
 The ion model by Dang and coworkers~\cite{smith94,chang1999,dang2006} was used with Amber lipids because
-analysis of simulations with the default Amber ion parameters is complicated by the artificial
-clustering of ions in solution \cite{nmrlipids_proj4}.
-In the simulations with ECC-lipids,
-we used the ECC-POPC parameters~\cite{melcr18} (available at \cite{ECC-POPC_nacl_cacl2_files}) and 
-ECC-POPS parameteres derived in this work (section \ref{section:ecc}) 
-together with the SPC/E~\cite{Berendsen1987} water model and 
-ECC-ions \cite{martinek17, kohagen16, Pluharova2014}.
-We used the SPC/E water model as its lower dielectric constant is consistent with the
+the default Amber ion parameters \cite{aqvist90} led to the artificial clustering of ions in solution \cite{nmrlipids_proj4}.
+In the ECC-lipid simulations,
+the ECC-POPC parameters~\cite{melcr18} (available from Ref. \citenum{ECC-POPC_nacl_cacl2_files}), 
+the ECC-POPS parameteres derived in this work (section \ref{section:ecc}), 
+the SPC/E~\cite{Berendsen1987} water model, and ECC-ions \cite{martinek17, kohagen16, Pluharova2014} were used.
+The SPC/E water model was selected because its lower dielectric constant is consistent with the
 ECC concept \cite{leontyev11,leontyev14}.
-
-The C--H bond order parameters were first averaged over the time for individual lipids and
-then the standard error of the mean of the average over lipids was used as the error estimate
-of the mean \cite{botan15,ollila16,nmrlipids_proj4}.
-
 \begin{table}[tbp]
@@
 
+The C--H bond order parameters were calculated directly from Eq. \ref{OPequation}.
+Average over the time for individual lipids was calculated first and
+the standard error of the mean of the average over lipids was then used as the
+error estimate \cite{botan15,ollila16,nmrlipids_proj4}.
+Python program ({\tt scripts/calcOrderParameters.py}) that uses the
+MDAnalysis library \cite{agrawal11,gowers16} is available in Ref. \citenum{MATCHgit}. 
+The ion number density profiles were calculated using the {\tt gmx density} tool
+of the Gromacs sofware package \cite{gromacsMANUAL}.
+The salt concentrations in buffer before solvating the lipids, reported in the used experimental data set \cite{roux90},
+were calculated as $[{\rm salt}]=N_{\rm c} \times[{\rm water}]\,/\,N_{\rm w}$,
+where $N_{\rm c}$ is the number of cations in simulation, $[{\rm water}]$\,=\,55.5~M and $N_{\rm w}$
+is the number of water molecules in the simulation.
+As discussed in our previous work \cite{nmrlipids_proj4}, the hydration levels of
+multilamellae are expected to be sufficiently similar in the used simulations and reference experiments \cite{roux90}.
+
+%The simulated system consisted of lipids (POPC and POPS) forming a single lipid bilayer,
+%an amount of water molecules comparable to the hydration levels in experiments, 
+%\todo{Check this claim. SAMULI: Check discussin in the NMRlipids IV manuscript}
+%\todo{EDIT: Taken from NMRlipids4 as an interesting information, to be edited.}
+%In the experimental $^2$H NMR literature data used in this work~\cite{browning80,roux90},
+%the lipids were first soluted to the buffer and then centrifuged to a pellet which
+%was used in the measurements. Such samples have lower lipid concentration
+%(approximately 10~wt~\% of lipids~\cite{browning80,roux88,roux90}) than 
+%gravimetric samples (60~wt~\%) and simulations (approximately 50-60~wt~\%) in this work.
+%Larger multilamellar repeat distances are expected in the samples with lower lipid
+%concentrations due to the swelling caused by electrostatic repulsion in pure PS lipid systems~\cite{millman82}.
+%However, the PS headgroup order parameters measured from gravimetric sample in this work
+%are in good agreement with the results from centrifuged sample in the literature \cite{browning80} (Fig. \ref{HGorderParameters}).
+%Furthermore, the equlibrium repeat distance rapidly decreases with the addtion of monovalent
+%salts and is close to the simulation box sizes already above 500~mM concentrations~\cite{millman82,rand89}.
+%Therefore, the hydration levels of multilamellae are expected to be sufficiently similar
+%in the used simulations and reference experiments.
+
 % ++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
@@
 The X-ray scattering form factors and C-H bond order parameters in
-the headgroup and glycerol backbone region of POPS lipid bilayer are already available in the literature \cite{pan14,nmrlipids_proj4}.
+the headgroup and glycerol backbone region of POPS lipid bilayer are already available in the literature \cite{kucerka14,nmrlipids_proj4}.
 Here, we extend the available experimental dataset of POPS lipids by measuring the C--H bond
@@
 overestimation seem to occur also in pure POPC simulation where the headgroup order parameter
-response to NaCl is also slightly overestimated \cite{meclr18}.
+response to NaCl is also slightly overestimated \cite{melcr18}.
 %Interestingly enough, \ce{K^+} maintains negative values of $\Gamma^{w}_{K}$ even for the negatively charged PC:PS~(5:1) bilayer.

### 2019-03-14 Manuscript/manuscript.tex
Update text

@@
 
-The most common negatively charged lipids in eukaryotic membranes bearing
-phophatidylserine (PS) headgroup are important signaling molecules \cite{lemmon08,leventis10,li14}.
+The phophatidylserine (PS) lipids are most common negatively charged lipids in eukaryotic membranes
+and important signaling molecules \cite{lemmon08,leventis10,li14}.
 They interact with signaling proteins \cite{leventis10},
@@
 %lipids with ions and other lipids. 
+Spectroscopic experiments give detailed information about the
+interactions between ions and PS lipids, but the data is often indirect and difficult to
+interpret \cite{hauser77,kurland79,eisenberg79,hauser83,dluhy83,hauser85,feigenson86,mattai89,roux90,roux91}.
 Some studies suggest that the
@@
 especially with physiological ion concentrations which are typically quite low.
-
-Spectroscopic experiments give detailed information about the
-interactions between ions and PS lipids, but the data is often indirect and difficult to
-interpret \cite{hauser77,kurland79,eisenberg79,hauser83,dluhy83,hauser85,feigenson86,mattai89,roux90,roux91,boettcher11}.
-Classical molecular dynamics (MD) simulations give
-access to all the details, but the current non-polarizable force fields are not
-sufficiently accurate to interpret the interactions between anionic
-lipid headgroups and other lipids or ions \cite{NMRlipidsIV}.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% On the simulation part .. %%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%From the perspective of phospholipid bilayers,
-There are force field parameters for neutral lipids PC and PE 
-with explicit polarizability using the Classical Durde model. \cite{chowdhary13, chowdhary17}
-While such a model shows a potential improvement over non-polarizable force fields from many perspectives -- 
-for example accounting for polarizability has a significant impact on the dipole potential of a membrane \cite{harder2009} -- 
-the non-polarizable version of the same force field 
-performes comparably well at a fraction of the computational cost 
-of the models with explicit polarization. \citep{lucas12,chowdhary13}
-
-Short paragraph describing previous MD-work on this topic (PS), i.e., \dots 
-\todo{SAMULI: Controversies in previous simulations (https://doi.org/10.1016/j.bpj.2018.08.044 and
-  Melcrova et al. Scientific Reports | 6:38035 | DOI: 10.1038/srep38035) should be mentioned also
-in the intro.}
-
-The interaction of various cations including calcium 
-with  neutral and also negatively charged lipid bilayers
-was thoroughly compared between simulations and experiments 
-from the point of the electrometer concept \cite{roux90, seelig87, altenbach84}
-in the references~\citenum{catte16, nmrlipids_proj4}. 
-It was concluded that the interaction of calcium with zwitterionic PC bilayers 
-is generally overestimated in all studied models \cite{catte16},
-while none of the models was capable of interpreting the NMR structural parameters
-with the negatively charged PC:PS (5:1) bilayers \cite{nmrlipids_proj4}. 
-Accounting for the electronic polarization was identified as 
-a non-negligible correction to the interaction between cations and POPC bilayers in Ref.~\cite{melcr18}.
-
-Here, we demonstrate that the same approach 
-significantly improves the description of calcium binding 
-also to the negatively charged PC:PS (5:1) bilayers. 
-In correspondence with the Ref.~\citenum{nmrlipids_proj4},
-we also use the changes of the head group order parameters $S^\alpha$ and $S^\beta$ 
-in PC and PS lipids 
-to quantify the binding affinity of the calcium cations
-and the induced structural changes
-(Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}). 
-To directly connect to the previous works \citep{catte16, nmrlipids_proj4}
-and to mark the improvement after treating electronic polarizability, 
-we show simulation results from the newly developed ECC-lipids and also from Lipid17 \citep{lipid17-future}.
-
-An empirical scaling factor, which reduces electrostatic forces between molecules, 
-is a well known concept in molecular modelling;
-such an approach was adopted in the early simulations of lipid bilayers \cite{egberts94, berendsen1996},
-in force fields employing explicit polarization \cite{lemkul2016empirical},
-or in the simulations of ionic liquids \cite{Bhargava2007}. 
-%An improvement of permeability of charged solutes in the gramicidin A channel 
-%is observed with a force field using charge equilibration to model polarizability. \cite{lucas2012} 
-At the same time, 
-the implicit inclusion of electronic polarizability to lipid models
-via charge re-scaling using electronic continuum correction (ECC) 
-substantially improved the interactions between cations and zwitterionic lipids \cite{melcr18}.
-
-Here, we extend the approach to anionic PS lipids.
-The newly developed model termed ECC-POPS 
+More recently, classical molecular dynamics (MD) simulations
+have been used to support the interpretation of spectroscopic experiments,
+but the results strongly depend on the force field parameters \cite{boettcher11,kucerka14,melcrova16,hallock18,valentine18}.
+In simulations with the CHARMM36 force field \cite{klauda10,venable13} using the NBfix parameters for calcium \cite{kim16}, 
+calcium ions interact only with the carboxylate group of PS lipids \cite{valentine18}.
+However, the same force field without the NBfix parameters, gives a significant binding affinity also to the phosphate region \cite{hallock18}.
+On the other hand, simulations with the Berger force field \cite{berger97,mukhopadhyay04}
+suggest a significant calcium binding also to the carbonyls in the acyl chains \cite{melcrova16}.
+
+The NMRlipids project \url{nmrlipids.blogspot.fi} has recently demonstrated
+that such controversies can be resolved by using the headgroup order parameters
+of phosphatidylcholine (PC) lipids \cite{catte16, nmrlipids_proj4}, which can
+be related to cation binding affinity to lipid bilayers using the electrometer
+concept \cite{akutsu81,altenbach84,seelig87}. The main advantage of the approach
+is the direct comparison between experimental and calculated order parameters,
+which reduced the ambiguity arising from the interpretation of the data.
+Unfortunately, none of the readily available force fields was sufficiently accurate to correctly reproduce the
+cation binding affinity to zwitterionic PC bilayers \cite{catte16} or to the
+mixtures with negatively charge PS lipids \cite{nmrlipids_proj4}.
+In our recent work~\cite{melcr18}, we were able to improve the cation binding affinity
+to zwitterionic 1-palmitoyl-2-oleoyl-sn-glycero-3-phosphocholine (POPC) lipid bilayer
+by implicitly including electronic polarizability using the electronic continuum
+correction (ECC) \cite{leontyev09}. The good agreement between the resulting ECC-POPC
+model and experiments enable detailed interpretation of calcium binding details to POPC lipid bilayer.
+
+%There are force field parameters for neutral lipids PC and PE 
+%with explicit polarizability using the Classical Durde model. \cite{chowdhary13, chowdhary17}
+%While such a model shows a potential improvement over non-polarizable force fields from many perspectives -- 
+%for example accounting for polarizability has a significant impact on the dipole potential of a membrane \cite{harder2009} -- 
+%the non-polarizable version of the same force field 
+%performes comparably well at a fraction of the computational cost 
+%of the models with explicit polarization. \citep{lucas12,chowdhary13}
+
+Here, we extend the ECC approach also to negatively charged 1-palmitoyl-2-oleoyl-sn-glycero-3-phospho-L-serine (POPS) lipid.
+The quality of the newly developed ECC-POPS force field parameters is
+evaluated against previously published X-ray scattering form factors\cite{kucerka14}
+and NMR \cite{roux90,nmrlipids_proj4} data for lipid headgroup C--H bond order parameters in various ionic conditions,
+as well as against acyl chain C--H bond order parameters reported in this work.
+as well as 
 not only describes the interactions of PS with ions more accurately, 
@@
 for a wide range of applications.
+
+%Here, we demonstrate that the same approach 
+%significantly improves the description of calcium binding 
+%also to the negatively charged PC:PS (5:1) bilayers. 
+%In correspondence with the Ref.~\citenum{nmrlipids_proj4},
+%we also use the changes of the head group order parameters $S^\alpha$ and $S^\beta$ 
+%in PC and PS lipids 
+%to quantify the binding affinity of the calcium cations
+%and the induced structural changes
+%(Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}). 
+%To directly connect to the previous works \citep{catte16, nmrlipids_proj4}
+%and to mark the improvement after treating electronic polarizability, 
+%we show simulation results from the newly developed ECC-lipids and also from Lipid17 \citep{lipid17-future}.
+
+%An empirical scaling factor, which reduces electrostatic forces between molecules, 
+%is a well known concept in molecular modelling;
+%such an approach was adopted in the early simulations of lipid bilayers \cite{egberts94, berendsen1996},
+%in force fields employing explicit polarization \cite{lemkul2016empirical},
+%or in the simulations of ionic liquids \cite{Bhargava2007}. 
+%%An improvement of permeability of charged solutes in the gramicidin A channel 
+%%is observed with a force field using charge equilibration to model polarizability. \cite{lucas2012} 
+%At the same time, 
+%the implicit inclusion of electronic polarizability to lipid models
+%via charge re-scaling using electronic continuum correction (ECC) 
+%substantially improved the interactions between cations and zwitterionic lipids \cite{melcr18}.
 

### 2019-03-15 Manuscript/manuscript.tex
Update text

@@
 Prague 6, Czech Republic} 
+\author{Tiago Ferreira}
+\affiliation{NMR group - Institut for Physics, Martin-Luther University Halle-Wittenberg} 
 \author{Pavel Jungwirth} 
@@
 %SAMULI: We need to put tiago because the experimental acyl chain order parameters
-\author{Tiago Ferreira}
-\affiliation{Halle} 
 \author{O. H. Samuli Ollila} 
@@
 that such controversies can be resolved by using the headgroup order parameters
-of phosphatidylcholine (PC) lipids \cite{catte16, nmrlipids_proj4}, which can
+of phosphatidylcholine (PC) lipids \cite{catte16, NMRlipidsIV}, which can
 be related to cation binding affinity to lipid bilayers using the electrometer
@@
 cation binding affinity to zwitterionic PC bilayers \cite{catte16} or to the
-mixtures with negatively charge PS lipids \cite{nmrlipids_proj4}.
+mixtures with negatively charge PS lipids \cite{NMRlipidsIV}.
 In our recent work~\cite{melcr18}, we were able to improve the cation binding affinity
@@
 Here, we extend the ECC approach also to negatively charged 1-palmitoyl-2-oleoyl-sn-glycero-3-phospho-L-serine (POPS) lipid.
-The quality of the newly developed ECC-POPS force field parameters is
-evaluated against previously published X-ray scattering form factors\cite{kucerka14}
-and NMR \cite{roux90,nmrlipids_proj4} data for lipid headgroup C--H bond order parameters in various ionic conditions,
-as well as against acyl chain C--H bond order parameters reported in this work.
-as well as 
-not only describes the interactions of PS with ions more accurately, 
-but also improves the structure of the PS head group 
-and its interactions with PC in mixed lipid bilayers. 
-The results pave the way to more realistic MD simulations 
+We also complement the available experimental data for the the force field quality
+evaluation by measuring the acyl chain C--H bond order parameters of POPS in bilayer using natural abundance $^{13}$C NMR.
+In addition to the acyl chain order parameters measured here, the quality of the newly developed ECC-POPS force field parameters is
+evaluated using previously published C--H bond order parameters of glycerol backbone and headgroup in various ionic
+conditions and lipid molar fractions \cite{roux90,NMRlipidsIV}, as well as X-ray scattering form factors\cite{kucerka14}.
+The overall improvement of the force field accuracy upon applying ECC
+%not only describes the interactions of PS with ions more accurately, 
+%but also improves the structure of the PS head group 
+%and its interactions with PC in mixed lipid bilayers. 
+%The results
+paves the way to more realistic MD simulations 
 of both neutral and negatively charged lipid bilayers 
@@
 %also to the negatively charged PC:PS (5:1) bilayers. 
-%In correspondence with the Ref.~\citenum{nmrlipids_proj4},
+%In correspondence with the Ref.~\citenum{NMRlipidsIV},
 %we also use the changes of the head group order parameters $S^\alpha$ and $S^\beta$ 
@@
 %(Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}). 
-%To directly connect to the previous works \citep{catte16, nmrlipids_proj4}
+%To directly connect to the previous works \citep{catte16, NMRlipidsIV}
 %and to mark the improvement after treating electronic polarizability, 
@@
 The ion model by Dang and coworkers~\cite{smith94,chang1999,dang2006} was used with Amber lipids because
-the default Amber ion parameters \cite{aqvist90} led to the artificial clustering of ions in solution \cite{nmrlipids_proj4}.
+the default Amber ion parameters \cite{aqvist90} led to the artificial clustering of ions in solution \cite{NMRlipidsIV}.
 In the ECC-lipid simulations,
@@
 the standard error of the mean of the average over lipids was then used as the
-error estimate \cite{botan15,ollila16,nmrlipids_proj4}.
+error estimate \cite{botan15,ollila16,NMRlipidsIV}.
 Python program ({\tt scripts/calcOrderParameters.py}) that uses the
@@
 is the number of water molecules in the simulation.
-As discussed in our previous work \cite{nmrlipids_proj4}, the hydration levels of
+As discussed in our previous work \cite{NMRlipidsIV}, the hydration levels of
 multilamellae are expected to be sufficiently similar in the used simulations and reference experiments \cite{roux90}.
@@
 
-The form factors from scattering and C-H bond order parameters from NMR experiments are good measures to validate
+The X-ray scattering form factors and C-H bond order parameters from NMR experiments are good measures to validate
 the lipid bilayer structure in MD simulations against experiments because they
@@
 The X-ray scattering form factors and C-H bond order parameters in
-the headgroup and glycerol backbone region of POPS lipid bilayer are already available in the literature \cite{kucerka14,nmrlipids_proj4}.
-Here, we extend the available experimental dataset of POPS lipids by measuring the C--H bond
-order parameters also from the acyl chain region using the R-PDLF experiments as described previously \cite{ferreira13}.
+the headgroup and glycerol backbone region of POPS lipid bilayer are already available in the literature \cite{kucerka14,NMRlipidsIV}.
+Here, we measure also the C--H bond order parameters of the acyl chain region using natural abundance $^{13}$C NMR.
 We first determined the order parameters of each peak corresponding different C--H bonds detected
@@
     from simulations with the Lipid17 \citep{lipid17-future} and the ECC-POPS models 
-    compared with experiments at 298~K. \citep{nmrlipids_proj4}
+    compared with experiments at 298~K. \citep{NMRlipidsIV}
     Open/closed symbols are used for palmitoyl/oleoyl chains of POPS. 
@@
 
-To evaluate the effect of ECC on the structural properties of POPS lipid bilayer,
-we compared the X-ray scattering form factors, C--H bond order parameters and area per molecules
-between simulation with original Lipid17 \cite{??} combined with the Dang ion parameters \cite{??},
-ECC-POPS simulation and experiments (Fig. \ref{simVSexpNOions_POPS} and table \ref{tab:apls}).
-The larger acyl chain order parameters, deviation in the form factor and smaller area per molecule 
-in the Lipid17 simulation with Dang ions when compared with experiments indicates that this simulation
-predicts too dense lipid bilayer. In previous study \cite{nmrlipids_proj4}, Lipid17 simulated with {\AA}qvist ions~\cite{aqvist90}
-gave larger area per molecule (57 \AA$^2$), probably due to the weaker counterion binding affinity.
-The Dang ions are used to here to avoid the artificial pairing of ions and condensation at larger salt
-concentrations observed with {\AA}qvist \cite{kohagen16, chen07, nmrlipids_proj4} when analysing the effect of ECC. 
-The scattering form factors and acyl chain order parameters from ECC-POPS are in good agreement with
-experiments (see Fig.~\ref{simVSexpNOions_POPS}), indicating the the bilayer dimensions and acyl chain
+The Lipid17/Dang simulation give larger acyl chain order parameters and smaller area per molecule than the
+experimental values (Fig. \ref{simVSexpNOions_POPS} and table \ref{tab:apls}),
+indicating that this simulation predicts too dense POPS lipid bilayer. This is also in line with
+the discrepancy observed in the X-ray form factor between this simulation and experiment (Fig. \ref{simVSexpNOions_POPS}).
+In previous study \cite{NMRlipidsIV}, the Lipid17 force field with {\AA}qvist ions~\cite{aqvist90}
+gave a larger area per molecule (57 \AA$^2$) for POPS being closer to the experimental value,
+which is probably due to the weaker counterion binding affinity of {\AA}qvist ions.
+However,  {\AA}qvist ions exhibit artificial aggregation of ions at larger salt concentrations \cite{kohagen16, chen07, NMRlipidsIV}.
+Therefore, the Dang ions with more realisitic bulk properties are used here to analyze the effect of ECC to the properties of POPS lipid bilayer. 
+The ECC-POPS simulation gives a good agreement with the X-ray scattering form factors and experimental
+acyl chain order parameters (see Fig.~\ref{simVSexpNOions_POPS}), indicating the the bilayer dimensions and acyl chain
 conformations are well described by the force field. The area per lipid from ECC-POPS simulation is slightly
@@
 are within each others error bars.
-%In order to obtain the area per lipid from experiments, modeling is used on top of the scattering factors \citep{kucerka14}. 
+
+% In order to obtain the area per lipid from experiments, modeling is used on top of the scattering factors \citep{kucerka14}. 
 %In ocntrast, this property is easily extracted directly from simulations with flat phospholipid bilayers. 
@@
 lipid force fields have typically problems in capturing the correct glycerol backbone and
-headgroup order parameters and conformations \cite{botan15,ollila16,nmrlipids_proj4}.
+headgroup order parameters and conformations \cite{botan15,ollila16,NMRlipidsIV}.
 This is the case also for Lipid17 POPS simulations, where the headgroup and glycerol backbone
 order parameters are quite far from experimental values with both Dang (Fig. \ref{simVSexpNOions_POPS})
-and {\AA}qvist \cite{nmrlipids_proj4} ions. The values from ECC-POPS simulation are closer,
+and {\AA}qvist \cite{NMRlipidsIV} ions. The values from ECC-POPS simulation are closer,
 but not in full agreement with experiments (Fig. \ref{simVSexpNOions_POPS}).
@@
 and the larger forking (difference of order parameters between hydrogens attached to the
-same carbon) of the $\alpha$-carbon in POPS than in POPC \cite{nmrlipids_proj4}
+same carbon \cite{ollila16}) of the $\alpha$-carbon in POPS than in POPC \cite{NMRlipidsIV}
 are not reproduced by neither of the models here. These differences are previously related
@@
 its structural accuracy is in the same level as in other state-of-the-art
-lipid force fields \cite{botan15, ollila16, Pluhackova2016, nmrlipids_proj4}.
+lipid force fields \cite{botan15, ollila16, Pluhackova2016, NMRlipidsIV}.
 
 \subsection{Counterion binding affinity and interactions between PC and PS head groups}
-
-Binding affinity of monovalent ions to lipid bilayers depends strongly on
-force field parameters \cite{catte16,melcr18,nmrlipids_proj4},
-which is the case also here as the sodium binding affinity to
-POPS lipid bilayer is stronger in Lipid17 simulation with Dang ions than in
-ECC-POPS simulation (Fig. \ref{fig:POPS-counterions-dens}).
-The cation binding affinities to lipid bilayers in simulations have been previously evaluated
-against experiments using the order parameters of $\alpha$ and $\beta$ carbons in the
-phosphatidylcholine (PC) lipid headgroup (see Fig.~\ref{fig:chemstruct_pc_ps} for the labeling) \cite{catte16,melcr18,nmrlipids_proj4},
-which decrease proportionally to the bound charge according to the ''electrometer concept'' \citep{seelig87}.
-Althought the response of PS lipid headgroup order parameters to the bound charge is also systematic,
-it is less well undestood and the addition of cations may cause phase transitions in negatively charged
-bilayers \cite{feigenson86,mattai89,roux91,roux90}. Therefore, the cation binding affinity to bilayers with
-PS lipids has been evaluated measuring the PC headgroup order parameters from POPC:POPS (5:1)
-mixtures\cite{roux90,nmrlipids_proj4}.
-Following our previous work \cite{nmrlipids_proj4}, we evaluate the counterion binding affinity
-to lipid bilayer in simulations containing PS using two different approaches.
-
-% The situation is more complicated for lipid bilayers containing
-%PS lipids, for which there is no ion-free reference state due to the presence of counterions.
-
-%Unlinke in pure PC bilayers,
-%Interpreting the changes of the head group order parameters of 
-%mixed PC:PS lipid bilayers 
-%from both simulations and experiments
-%is complicated by the presence of counterions. \cite{nmrlipids_proj4}
 
@@
   \caption{\label{fig:POPS-counterions-dens}
-    Number density profiles of \ce{K^{+}}, \ce{Na^{+}} counterions along the membrane normal axis 
-    for the negatively charged bilayers composed of only POPS (left) or POPC:POPS (5:1, right) 
-    using ECC-lipids/ECC-ions (solid lines) and Lipid17/Dang ions (dashed lines).  
+    Number density profiles of \ce{K^{+}} and \ce{Na^{+}} counterions along the membrane normal axis
+    in ECC-lipid (solid lines) and Lipid17/Dang (dashed lines) simulations of pure POPS (left) or POPC:POPS (5:1, right) bilayers.  
     The density profiles of phosphate groups and water are divided by 4 and 100, respectively.  
@@
 \end{figure} 
+Binding affinity of monovalent ions to lipid bilayers depends strongly on
+force field parameters \cite{catte16,NMRlipidsIV},
+which is the case also here as the sodium counterion binding affinity to
+POPS lipid bilayer is stronger in Lipid17/ Dang simulation than in
+ECC-POPS simulation (Fig. \ref{fig:POPS-counterions-dens}).
+The cation binding affinities to lipid bilayers in simulations have been previously evaluated
+against experiments using the order parameters of $\alpha$ and $\beta$ carbons in the
+phosphatidylcholine (PC) lipid headgroup (see Fig.~\ref{fig:chemstruct_pc_ps} for the labeling) \cite{catte16,melcr18,NMRlipidsIV},
+which decrease proportionally to the bound charge according to the ''electrometer concept'' \citep{seelig87}.
+Althought the response of PS lipid headgroup order parameters to the bound charge is also systematic,
+it is less well undestood and the addition of cations may cause phase transitions in negatively charged
+bilayers \cite{feigenson86,mattai89,roux91,roux90}. Therefore, the cation binding affinity to bilayers with
+PS lipids has been evaluated measuring the PC headgroup order parameters from POPC:POPS (5:1)
+mixtures\cite{roux90,NMRlipidsIV}.
+Following our previous work \cite{NMRlipidsIV}, we evaluate the counterion binding affinity
+to lipid bilayer in simulations containing PS using two different approaches.
+
+% The situation is more complicated for lipid bilayers containing
+%PS lipids, for which there is no ion-free reference state due to the presence of counterions.
+
+%Unlinke in pure PC bilayers,
+%Interpreting the changes of the head group order parameters of 
+%mixed PC:PS lipid bilayers 
+%from both simulations and experiments
+%is complicated by the presence of counterions. \cite{NMRlipidsIV}
 
@@
 %to neutral and negatively charged phospholipid bilayers \citep{akutsu81, roux90} 
-%in our works \citep{catte16, nmrlipids_proj4}.
+%in our works \citep{catte16, NMRlipidsIV}.
 %After testing the currently available force fields for phospholipids,
@@
 %in the lipid parameters are required for more accurate interactions even with monovalent cations. 
-%\citep{catte16, melcr18, nmrlipids_proj4}
-
-First, we monitor the response of POPC headgroup order parameters to the increasing amount of POPS (Fig.~\ref{fig:delta_ordPar_NaCl_PC-PS_mix}).
-According to the electrometer concept, the head group order parameters of PC lipids increase
-with the increasing amount of negatively charged PS lipids in experiments, because the headgroup P--N dipole
-tilts more parallel to the membrane interface \cite{seelig87,scherer87}.
-The trend is observed with both force fields in simulations with \ce{K+} counterions and
-in ECC-POPS simulation with \ce{Na+} counterions, but not in Lipid17/Dang ion simulations with \ce{Na+} counterions,
-where the order parameters are unchanged or decrease and the P--N vector tilts more perpendicular to the membrane interface.
-The result can be explained by the higher binding affinity of sodium in the
-Lipid17/Dang ion simulations (Fig~\ref{fig:POPS-counterions-dens}), which
-compensates the negative charge of PS lipids at the interface.
+%\citep{catte16, melcr18, NMRlipidsIV}
+
 \begin{figure*}[!tbp] 
@@
     from ECC-lipid and Lipid17/Dang simulations with \ce{Na^+} (top) and \ce{K^+} (bottom) counterions.
-    Experimental order parameter values and signs for pure POPS are from Ref. \citenum{nmrlipids_proj4} and values
+    Experimental order parameter values and signs for pure POPS are from Ref. \citenum{NMRlipidsIV} and values
     for POPC:POPS (5:1) mixture from Ref. \citenum{roux90} (only \ce{Na^+} counterions, but shown also in bottom plots).
@@
 \end{figure*} 
+First, we monitor the response of POPC headgroup order parameters to the increasing amount of POPS (Fig.~\ref{fig:delta_ordPar_NaCl_PC-PS_mix}).
+According to the electrometer concept, the head group order parameters of PC lipids increase
+with the increasing amount of negatively charged PS lipids in experiments, because the headgroup P--N dipole
+tilts more parallel to the membrane interface \cite{seelig87,scherer87}.
+The trend is observed with both force fields in simulations with \ce{K+} counterions and
+in ECC-POPS simulation with \ce{Na+} counterions, but not in Lipid17/Dang ion simulations with \ce{Na+} counterions,
+where the order parameters are unchanged or decrease and the P--N vector tilts more perpendicular to the membrane interface.
+The result can be explained by the higher binding affinity of sodium in the
+Lipid17/Dang ion simulations (Fig~\ref{fig:POPS-counterions-dens}), which
+compensates the negative charge of PS lipids at the interface.
 
@@
 According to the electrometer concept, the decrease of PC lipid head group order parameters
-correlates with the amount of bound cations \cite{akutsu81,altenbach84,seelig87,roux90,catte16,nmrlipids_proj4},
+correlates with the amount of bound cations \cite{akutsu81,altenbach84,seelig87,roux90,catte16,NMRlipidsIV},
 thus the binding affinity of the ions to POPC:POPS (5:1) mixtures increase in order K$^{+}$ $<$ Li$^{+}$  $<$ Mg$^{2+}$  $<$ Ca$^{2+}$.
@@
 are clearly overestimated in Lipid17/Dang simulations and slightly overestimated in ECC-POPS simulations,
-while 
-%In conclusion, the response of PC
-%headgroups to added amount of PS and to the added amount of ions suggest that the Lipid17/Dang
-%simulations sinficantly overestimate the sodium binding affinity to the PS containing lipid bilayers, while
-%ECC-POPS simulations give results closer to the experiments, but may still slightly overestimate
-%the binding affinity.
-the binding affinity of potassium is better described by both force fields.
+while the binding affinity of potassium is better described by both force fields.
 
@@
 %The order parameters of the PS head groups including their changes with the content of PC lipids
-%are challenging to capture correctly in classical MD simulation models. \cite{nmrlipids_proj4}
+%are challenging to capture correctly in classical MD simulation models. \cite{NMRlipidsIV}
 %From such a perspective,
@@
 or monovalent ion concentration in solvent
-(Figs. \ref{fig:delta_ordPar_NaCl_PCPS}, \ref{fig:delta_ordPar_KCl_PCPS} and \ref{fig:delta_ordPar_NaCl_PC-PS_mix}),
-while calcium leads to relatively large changes with low concentrations followed by a rapid saturation below 100mM (Fig. \ref{fig:delta_ordPar_CaCl_PCPS}) \cite{scherer87,roux90}.
+(Figs. \ref{fig:delta_ordPar_NaCl_PCPS}, \ref{fig:delta_ordPar_KCl_PCPS} and \ref{fig:delta_ordPar_NaCl_PC-PS_mix}).
+%while calcium leads to relatively large changes with low concentrations followed by a rapid saturation below 100mM (Fig. \ref{fig:delta_ordPar_CaCl_PCPS}) \cite{scherer87,roux90}.
 %The interaction with \ce{K^+}, which binds very weakly to both neutral and negatively charged membranes, 
@@
 vectors (Fig. \ref{fig:delta_ordPar_NaCl_PC-PS_mix}). The overestimation of the order parameter changes
-was also observed in previous simulations with other force fields \cite{nmrlipids_proj4}, but not in
+upon increasing amount of POPC in a bilayer 
+was also observed in previous simulations with other force fields \cite{NMRlipidsIV}, but not in
 ECC-POPS simulation, where also the changes in the PS headgroup orientation are more modest (Fig. \ref{fig:delta_ordPar_NaCl_PC-PS_mix}).
-Also, the POPS headgroup orientation and order parameters are more sensitive to the additional amount
+The POPS headgroup orientation and order parameters are more sensitive also to the additional amount
 of sodium and potassium in Lipid17/Dang simulations, while dependence is weaker and closer to experiments
@@
 %even for interactions with soft cations like \ce{K^+}.
-In conclusion, using ECC to POPS lipid headgroup improves its
-interactions with monovalent ions and POPC headgroups when compared with NMR order parameter data,
-althought there is room for improvement in the sodium binding affinity.
 
@@
 %without the necessity of defining a Gibbs dividing surface. \citep{melcr18, chattorajBOOK}
-The value of $\Gamma ^{w} _{\rm Na}=-0.11 \pm 0.01$ nm$^{-2}$ for 1 M sodium in pure POPC
-was reported from a ECC-POPC simulation in the previous work \cite{melcr18}.
+The value of $\Gamma ^{w} _{\rm Na}=-0.11 \pm 0.01$ nm$^{-2}$ for 1 M sodium in pure ECC-POPC
+simulation was reported in the previous work \cite{melcr18}.
 The presence of $\sim$17\% POPS in POPC bilayer increases the
@@
 $\Gamma ^{w} _{\rm K}=-0.12 \pm ??$ nm$^{-2}$ \todo{Error bar should be added} in POPC:POPS (5:1) mixture (Fig. \ref{fig:nacl-dens_PCPS}).
-As discussed above,
-the sodium binding to POPC:POPS (5:1) mixture may be slightly overestimated also in
-ECC-lipid simulations. However, the difference may not be affected because similar small
-overestimation seem to occur also in pure POPC simulation where the headgroup order parameter
-response to NaCl is also slightly overestimated \cite{melcr18}.
+As discussed above, the sodium binding to POPC:POPS (5:1) mixture may be slightly overestimated also in
+ECC-lipid simulations. The response of headgroup order parameters to the addition of NaCl are
+slightly overestimated also in ECC-POPC simulations (Fig 3 in Ref. \citenum{melcr18}),
+suggesting that similar small overestimation of sodium binding may also occur to a pure POPC bilayer.
+Therefore, the overestimation may arise from the sodium interactions with PC headgroup and
+increase in binding affinity due to PS lipids would not be affected.
 %Interestingly enough, \ce{K^+} maintains negative values of $\Gamma^{w}_{K}$ even for the negatively charged PC:PS~(5:1) bilayer.
@@
 
+In conclusion, applying ECC to Amber Lipid17 POPS force field parameters improves its
+interactions with both monovalent ions and POPC headgroups when compared with the NMR order parameter data.
+However, further optimization may be needed to fully reproduce the interactions between lipids and
+sodium ions.
+
 \subsection{Molecular interaction and binding affinities of \ce{Ca^{2+}} cations to mixed POPC:POPS (5:1) membrane} 
@@
 
-The headgroup order parameters of PC lipids decrease more upon addition of \ce{CaCl2}
-in POPC:POPS (5:1) mixtures than in pure PC bilayers due to the higher calcium binding
-affinity to the negatively charged lipid bilayers \cite{akutsu81,altenbach84,roux90,nmrlipids_proj4}.
-Following our previous work \cite{nmrlipids_proj4}, we use the experimental
-headgroup order parameter data as a function of CaCl$_2$ concentration
-from POPC:POPS (5:1) mixture to evaluate the calcium binding affinity to lipid bilayers containing PS lipids.
-The decrease of the PC headgroup order parameters upon addition of \ce{CaCl2} 
-is overestimated in the Lipid17/Dang simulation, while ECC-POPS simulations give results
-closer to the experiments (Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}). 
-This indicates that the lower binding affinity of calcium in ECC-POPS
-simulation is more realistic (Fig.~\ref{fig:cacl-dens_PCPS}).
+%The headgroup order parameters of PC lipids decrease more upon addition of \ce{CaCl2}
+%in POPC:POPS (5:1) mixtures than in pure PC bilayers due to the higher calcium binding
+%affinity to the negatively charged lipid bilayers \cite{akutsu81,altenbach84,roux90,NMRlipidsIV}.
+Our recent work demonstrates that the POPC headgroup order parameters measured
+from POPC:POPS (5:1) mixture as a function of CaCl$_2$ concentration \cite{roux90}
+can be used to evaluate the calcium binding affinity to lipid bilayers containing PS lipids \cite{NMRlipidsIV}.
+The decrease of the PC headgroup order parameters in this mixture upon addition of \ce{CaCl2} 
+were overestimated by the Lipid17/Dang simulations (also shown in Fig.~\ref{fig:delta_ordPar_CaCl_PCPS})
+and all the other tested force fields, except CHARMM36 with the recently introduced NBfix for calcium \cite{kim16},
+which underestimated the headgroup order parameter response.
+In ECC-lipid simulations, the headgroup responses are in better agreement with experiments (Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}),
+indicating that the lower binding affinity than in Lipid17/Dang simulations is more realistic (Fig.~\ref{fig:cacl-dens_PCPS}).
 \todo{JOE: The density plot still needs to be changed to show both Lipid17 and ECC-lipids, and to support this argument.}
-The relative surface excess of calcium with respect to water, $\Gamma ^w _{Ca}$,
-calculated from ECC-lipid simulation gives $\Gamma^{w}_{Ca} = 0.24\mathrm{nm^{-2}}$ for the PC:PS (5:1) bilayer with 409 mM CaCl$_2$
-and $\Gamma^{w}_{Ca} = 0.06\mathrm{nm^{-2}}$ for pure POPC bilayer with 350 mM CaCl$_2$ \cite{melcr18},
-illustrating the higher affinity of calcium to bilayers containing PS lipids.
-However, the steep decrease of the order parameters at low calcium concentrations
-observed especially in the negatively charged bilayers is not captured by the ECC-lipid simulations.
+%However, the steep decrease of the order parameters at low calcium concentrations,
+%observed especially in negatively charged bilayers \cite{borle85,macdonald87,roux90},
+%may not be fully captured by the ECC-lipid simulations.
+The increased calcium binding affinity to bilayers containing PS lipids is illustrated in
+ECC-lipid simulations by the increasing relative surface excess of calcium with respect to water from
+$\Gamma^{w}_{Ca} = 0.06\mathrm{nm^{-2}}$ for pure POPC bilayer with 350 mM CaCl$_2$ \cite{melcr18}
+to $\Gamma^{w}_{Ca} = 0.24\mathrm{nm^{-2}}$ for the POPC:POPS (5:1) bilayer with 409 mM CaCl$_2$.
 
@@
 on low CaCl$_2$ concentrations with a rapid saturation below 100 mM \cite{roux90}.
-The changes are again overestimated by Lipid17/Dang simulations, which
-also exhibits non-systematic changes in both PS headroup orientation and order parameters
-upon addition of CaCl$_2$ (Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}).
-In ECC-lipid simulations,  the P--N and C$_{\beta}$--C$_{\gamma}$ vectors systematically
-tilt more perpendicular and parallel to the membrane interface, respectively,
-and the changes of the PS headgroup order parameters are again closer to the
-experiments (Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}).
-Despite significant improvements in the ECC-lipid against Lipid17/Dang
-and other force fields for the response of PS headgroup order parameters to
-the bound calcium \cite{nmrlipids_proj4}, the strong dependence of both PC and PS
-headgroup order parameters on low concetrations of CaCl$_2$ is not reproduced by
-the model (Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}). In addition to the inaccurate
-interactions between calcium ions and PS headgroup, the potential sources of this
+Also these changes were overestimated by the Lipid17/Dang (also shown in Fig.~\ref{fig:delta_ordPar_CaCl_PCPS})
+and other tested force fields in our recent work \cite{NMRlipidsIV}, including CHARMM36 with the NBfix correction for calcium
+which underestimated the binding affinity.
+In ECC-lipid simulations, the changes of the PS headgroup order parameters are not overestimated,
+but the strong dependence on low concetrations of CaCl$_2$ is not fully reproduced (Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}).
+In addition to the inaccurate interactions between calcium ions and PS headgroup, the potential sources of this
 discrepancy include the above observed slight overestimation of \ce{Na^+} counterions and
 inaccurate structures of the lipid headgroup (Fig.~\ref{simVSexpNOions_POPS}).
-%and also possibly from the approximate mean-field treatment of the electronic polarization.
+
+In ECC-lipid simulations, the average orientation of P--N vectors of both POPC and POPS headgroups tilts more perpendicular to the membrane surface
+by 11$^\circ$ and  3$^\circ$, respectively, upon addition of 620~mM CaCl$_2$ to POPC:POPS (5:1) mixture (Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}),
+suggesting that the PS headgroup orientation is less sensitive to the bound calcium than PC.
+On the other hand, the average orientation of the C$_{\beta}$--C$_{\gamma}$ vector PS headgroup
+tilts to the opposite direction by $11^\circ$ in the same system.
+%the systematic tilting of P--N and C$_{\beta}$--C$_{\gamma}$ vectors 
+%to more perpendicular and parallel direction with respect to the membrane interface, respectively,
+%is observed upon addition of CaCl$_2$ (Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}).
+In Lipid17/Dang simulations, the changes in both headroup orientations and order parameters
+are larger and less systematic for the PS headgroup, especially with lower concentrations (Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}).
+Together with the smaller error bars and shorter residence times (Fig.~\ref{fig:hist_residence_times}) in
+ECC-lipid simulations than in typically observed for other force fields \cite{javanainen17,melcr18,NMRlipidsIV},
+this suggests that the ECC accelerates the equlibration of ions at lipid bilayer interface.
+%
+%changes by $11^\circ$ from $62^\circ$ to $73^\circ$ (more towards the membrane interior)
+% at 620~mM buffer concentration of \ce{CaCl2}.
+% We have investigated this effect using the mean orientations of P-N and \ce{COO^-} vectors 
+%(defined and plotted in Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}). 
+%While the mean angle between the P-N vector of PC head group and the membrane normal
+%decreased by 11$^\circ$ from 64$^\circ$ to 55$^\circ$
+% at 620~mM buffer concentration of \ce{CaCl2},
+%the mean orientation of the P-N vector in the PS head group has 
+%decreased only by 3$^\circ$ from 67$^\circ$ to 64$^\circ$ 
+%in the same concentration range. 
+%Such a largely diminished flexibility 
+%may arise from the binding of calcium to the \ce{COO^-} group in PS. 
+%Whe have measured the mean orientation of the \ce{COO^-} group 
+%as the connector of the carbon atoms 
+%forming the bond between the group and the $\beta$-carbon of the phospholipid 
+%(Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}).
+%by the binding of calcium cations to the \ce{COO^-} group, 
+%which reorients towards the membrane interior and to the phosphate region, 
+%where calcium cations bind predominantly. 
+%
+%
+Our 1 $\mu$s simulations seems to be sufficiently long for the ECC-lipid simulations, because
+90\% of the calcium residence times are shorter than $60\,\mathrm{ns}$ for pure POPC bilayer
+and shorter than $200\,\mathrm{ns}$ for POPC:POPS (5:1) mixture, while the
+longest observed residence times are $141\,\mathrm{ns}$ and $485\,\mathrm{ns}$, respectively (Fig.~\ref{fig:hist_residence_times}).
+Interestingly, the calcium residence times are 3-4 times longer in POPC:POPS (5:1) mixture than in pure POPC.
+In ECC-lipid simulations, on average 41\% of the total population of bound calcium cations is in contact with PS lipids
+with 8\% bound only to them, even though the negatively charged membrane contains only 18\% of POPS (Table \ref{tab:binding}).
+Therefore, calcium ions bind more likely to negatively charged PS lipids than neutral PC lipid in a mixed bilayer, as expected.
 
@@
 %are very challenging to interpret from both experiments \cite{roux90} 
-%and also from MD simulations \cite{nmrlipids_proj4}. 
+%and also from MD simulations \cite{NMRlipidsIV}. 
 %The changes of the order parameters $S^\alpha$ and $S^\beta$ in PS head groups from simulations 
 %are not found to follow the experimental profiles, 
-%nor any common trend, with the currently available models \cite{nmrlipids_proj4}.
+%nor any common trend, with the currently available models \cite{NMRlipidsIV}.
 
@@
 % Populations from contacts
-Interaction sites between calcium and PS lipids \cite{??} have been recently analyzed using
-different force fields and experimental techniques, but the results have been controversial.
-Simulations using the NBfix parameters for calcium \cite{kim16} in CHARMM36 force field \cite{klauda10,venable13},
-combined with 2D infrared spectroscopy, suggest that calcium ions interact only with the carboxylate group of PS lipids \cite{valentine18}.
-However, the same force field without the NBfix parameters, combined with NMR chemical shifts and
-rotational-echo double-resonance (REDOR) experiments, suggests a significant binding affinity also to the phosphate region \cite{hallock18}.
-On the other hand, simulations with the Berger force field \cite{berger97,mukhopadhyay04},
-combined with fluorescent and vibrational sum frequency spectroscopy, suggest a significant
-calcium binding also to the carbonyls in the acyl chains \cite{melcrova16}.
-The advantage of the NMR order parameters used in this work is the direct connection between
-measured and simulation data, which minimizes the human interpretation when evaluating the
-force field quality \cite{ollila16}. Therefore, we believe that the ECC-lipid model, which is in the best
-agreement with experimental headgroup order parameters among the available models, gives the currently best available
-interpretation of the experimental data.
-
+
+Recent analyzes of interaction sites between calcium and PS lipids combining
+MD simulations with various experimental techniques have been 
+controversial because the results strongly depend on force fields parameters \cite{melcrova16,valentine18,hallock18}.
 In ECC-lipid simulation of POPC:POPS (5:1) mixture with 409 mM CaCl$_2$,
@@
 \todo{I am not sure if this is correct, because I do not understand the numbers in the table.} (Table \ref{tab:binding}),
-which is in line with the results with NBfix correction in CHARMM36 \cite{hallock18}.
-However, calcium ions coordinate with even four distinct PS lipids
-in CHARMM36 simulations without NBfix \cite{hallock18}, while coordination with more than two PS lipids
+in line with the results without the NBfix corrections in CHARMM36 \cite{hallock18}.
+However, in CHARMM36 simulations without the NBfix correction for calcium,
+calcium ions significantly overbind to PC lipids  \cite{catte16} and calcium ions coordinate
+with even four distinct PS lipids \cite{hallock18}, while coordination with more than two PS lipids
 was not observed in our ECC-lipid simulations (Fig.~\ref{fig:cacl_complexes}).
-This is not surprising because calcium significantly overbinds to PC lipids in
-CHARMM36 simulations without the NBfix correction for calcium \cite{catte16}.
 Furthermore, the POPS:POPC:Ca$^{2+}$ (7:3:7) mixture with relatively high
@@
 Overestimated coordination of cations with carbonyl groups is a potential source for
-overestimated cation binding in Berger force fields \cite{catte16,nmrlipids_proj4}. 
+overestimated cation binding in Berger force fields \cite{catte16,NMRlipidsIV}. 
 In CHARMM36 simulations with the NBfix, calcium mainly interacted with only
 carboxylate of PS lipids  \cite{valentine18}, which is a potential source for the
-underestimated calcium binding affinity to lipid bilayer with this force field \cite{nmrlipids_proj4}.
+underestimated calcium binding affinity to lipid bilayer with this force field \cite{NMRlipidsIV}.
 
@@
 %the lengths of simulation trajectories are insufficient 
-%to yield converged results in this aspect (see SI and Refs.~\cite{catte16, melcr18, nmrlipids_proj4}). 
+%to yield converged results in this aspect (see SI and Refs.~\cite{catte16, melcr18, NMRlipidsIV}). 
 \begin{table}[tb!] 
@@
 
-In ECC-lipid simulations, on average 41\% of the total population of bound calcium cations is in contact with PS lipids
-with 8\% bound only to them, even though the negatively charged membrane contains only 18\% of POPS (Table \ref{tab:binding}).
-Therefore, calcium ions bind more likely to negatively charged PS lipids than neutral PC lipid in a mixed bilayer, as expected.
-In PS lipids, the calcium binds both phosphate and carboxylate, but probability for the latter is twice higher.
-In PC lipids, the calcium binds almost always to phosphate, but in some cases also to carbonyl, as also observed previously in pure POPC simulations \cite{melcr18}.
 %Still, the dominant contribution to the binding of calcium comes from the phosphate groups. 
@@
 %described above and in Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}.
-The average P--N vector orientation of both POPC and POPS headgroups tilts more perpendicular to the membrane surface
-upon addition of 620~mM CaCl$_2$  to POPC:POPS (5:1) mixture by  11$^\circ$ and  3$^\circ$, respectively (Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}).
-The average orientation of the PS headgroup C$_{\beta}$--C$_{\gamma}$ vector tilts to the opposite direction
-by $11^\circ$ in the same system.
-%changes by $11^\circ$ from $62^\circ$ to $73^\circ$ (more towards the membrane interior)
-% at 620~mM buffer concentration of \ce{CaCl2}.
-% We have investigated this effect using the mean orientations of P-N and \ce{COO^-} vectors 
-%(defined and plotted in Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}). 
-%While the mean angle between the P-N vector of PC head group and the membrane normal
-%decreased by 11$^\circ$ from 64$^\circ$ to 55$^\circ$
-% at 620~mM buffer concentration of \ce{CaCl2},
-%the mean orientation of the P-N vector in the PS head group has 
-%decreased only by 3$^\circ$ from 67$^\circ$ to 64$^\circ$ 
-%in the same concentration range. 
-%Such a largely diminished flexibility 
-%may arise from the binding of calcium to the \ce{COO^-} group in PS. 
-%Whe have measured the mean orientation of the \ce{COO^-} group 
-%as the connector of the carbon atoms 
-%forming the bond between the group and the $\beta$-carbon of the phospholipid 
-%(Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}).
-Therefore, we conclude that the PS headgroup orientation is less sensitive to the bound calcium than PC.
-%by the binding of calcium cations to the \ce{COO^-} group, 
-%which reorients towards the membrane interior and to the phosphate region, 
-%where calcium cations bind predominantly. 
-90\% of the calcium residence times are shorter than $60\,\mathrm{ns}$ for pure POPC bilayer
-and shorter than $200\,\mathrm{ns}$ for POPC:POPS (5:1) mixture, while the
-longest observed residence times are $141\,\mathrm{ns}$ and $485\,\mathrm{ns}$, respectively (Fig.~\ref{fig:hist_residence_times}),
-suggesting that the calcium residence times are 3-4 times longer in POPC:POPS (5:1) mixture than in
-pure POPC. Athough the ECC-lipid model gives the most realistic picture of calcium binding to
+
+%Simulations using CHARMM36 force field \cite{klauda10,venable13} with
+%the NBfix parameters for calcium \cite{kim16}, combined with 2D infrared spectroscopy,
+%suggest that calcium ions interact only with the carboxylate group of PS lipids \cite{valentine18}.
+%The same force field without the NBfix parameters, combined with NMR chemical shifts and
+%rotational-echo double-resonance (REDOR) experiments, suggests a significant binding affinity also to the phosphate region \cite{hallock18}.
+%Simulations with the Berger force field \cite{berger97,mukhopadhyay04},
+%combined with fluorescent and vibrational sum frequency spectroscopy, suggest a significant
+%calcium binding also to the carbonyls in the acyl chains \cite{melcrova16}.
+
+%In PS lipids, the calcium binds both phosphate and carboxylate, but probability for the latter is twice higher.
+%In PC lipids, the calcium binds almost always to phosphate, but in some cases also to carbonyl, as also observed previously in pure POPC simulations \cite{melcr18}.
+
+The advantage of the NMR order parameters used in this work is the direct connection between
+measured and simulation data, which enables more unambiguous evaluation of the force field quality \cite{catte16,ollila16}
+than more indirect comparison in previous studies using 2D infrared spectroscopy \cite{valentine18}, NMR chemical shifts and
+rotational-echo double-resonance (REDOR) experiments \cite{hallock18} or fluorescent and vibrational sum frequency spectroscopy \cite{melcrova16}.
+Because the ECC-lipid model gives the best agreement with the experimental headgroup order parameters
+with various ion conctenrations among the available models, we believe that it gives the currently best available
+interpretation of the experimental data. Athough the ECC-lipid model gives the most realistic picture of calcium binding to
 lipid bilayers containing PS lipids with respect to NMR order parameters, it does not
@@
 in the vicinity of negatively charged lipid bilayers.
+
+%Despite significant improvements in the ECC-lipid against Lipid17/Dang
+%and other force fields for the response of PS headgroup order parameters to
+%the bound calcium \cite{NMRlipidsIV}, 
+%the model (Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}).
+
+% and also possibly from the approximate mean-field treatment of the electronic polarization.
 
@@
 % -> plotting the distribution of O-Ca-Cb-N dihedral angle may be also a good idea for us (SI?)
-%     -> NMRlipids 4 in SI \cite{nmrlipids_proj4} plots this information: 
+%     -> NMRlipids 4 in SI \cite{NMRlipidsIV} plots this information: 
 %           - C36 agress with the distribution in \cite{hallock18}

### 2019-03-18 Manuscript/manuscript.tex
Small updates

@@
 scale the parameters of atoms only in the headgroup, glycerol backbone, and carbonyl regions \cite{melcr18}.
-Furhter optimization of parameters was not done in this work.
+Further optimization of parameters was not done in this work.
+\todo{Add here a reference to the parameters.}
 % we used exactly the pre-determined scaling factor 
@@
 the TIP3P water model~\cite{jorgensen83}, and ion models by Dang and coworkers~\cite{smith94,chang1999,dang2006} were used.
-The previously generated \cite{botan15} Lipid14 parameters in Gromacs format were downloaded from Ref.~\citenum{lipid14files}. 
-The Lipid17 parameters were obtained from AmberTools18 \cite{amber18} 
+The previously generated \cite{botan15} Lipid14 parameters for POPC in Gromacs format were downloaded from Ref.~\citenum{lipid14files}. 
+The Lipid17 parameters for POPS were obtained from AmberTools18 \cite{amber18} 
 and converted to Gromacs format using acpype \cite{acpype}.  
@@
 the ECC-POPC parameters~\cite{melcr18} (available from Ref. \citenum{ECC-POPC_nacl_cacl2_files}), 
-the ECC-POPS parameteres derived in this work (section \ref{section:ecc}), 
+the ECC-POPS parameteres derived in this work (see above), 
 the SPC/E~\cite{Berendsen1987} water model, and ECC-ions \cite{martinek17, kohagen16, Pluharova2014} were used.

### 2019-03-19 Manuscript/manuscript.tex

@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
 %% This is a (brief) model paper using the achemso class 
@@
 Tampere, Finland} 
-%SAMULI: We need to put tiago because the experimental acyl chain order parameters
 \author{O. H. Samuli Ollila} 
@@
 %yielding a total charge of $-0.75$ in the ECC formalism. \cite{martinek17, leontyev09, melcr18}
+Similarly to the ECC-POPC model \cite{melcr18}, we applied the ECC only to the
+non-tail atoms, i.e., to atoms in the headgroup, glycerol backbone, and carbonyl regions. 
 %polar parts of the phospholipid,
@@
 \end{figure}
+
+The head group and acyl chain order parameters within ECC-lipids
+are in general in a good agreement with the experimental values 
+as shown in Fig. \ref{simVSexpNOions_POPS}. 
+The acyl chain order parameters in particular agree with the experimental data within error bars.
+The order parameters of the head groups are at an accuracy comparable to 
+other currently available classical models of lipids \citep{botan15, catte16, Pluhackova2016, nmrlipids_proj4}. 
+In particular, ECC-POPS is closer to the experimental order parameters 
+than Lipid17 with either Dang \cite{smith94,chang1999,dang2006} 
+or ff99 ions \citet{aqvist90}.
 

### 2019-04-02 Manuscript/manuscript.tex
Removing TODOs -- updated figures PC-PS interaction and also with OP-response to ions.

@@
   }
-  \todo{JOE+SAMULI: split this figure into separate plots for PC and PS, 
-    and change the PS-plot x-axis to reflect the dilution of POPS by PC (now shows the content of PS). 
-    The experimental lines in the bottom plot shall be different to make clear that they are not with K counterions (gray dashed?).} \\
-  \todo{Error bar of 0.02 should be shown for the experimental POPS results because they come from different experiments, see caption of figure 8 in NMRlipids IV publication.}
+  \todo{JOE+SAMULI: split this figure into separate plots for PC and PS}
 \end{figure*} 
@@
   \todo{We should plot also the experimental results for POPC with KCl in this figure for the lower bound.} \\
-  \todo{Why there are no error bars for ECC-POPS in this and other figures?} \\
-  \todo{SAMULI: Use the same y-axis scale in the bottom figures showing angles for POPC and POPS to ease the comparison.}
   \end{figure}
@@
   }
-  \todo{SAMULI: Use the same y-axis scale in the bottom figures showing angles for POPC and POPS to ease the comparison.}
 \end{figure}

### 2019-04-03 Manuscript/manuscript.tex
Removing TODOs -- plotting also the experimental results for POPC with KCl for the lower bound.
Removing todos, density plot

@@
   \includegraphics[height=\figheightsmall]{../Fig/order_parameters_changes_A-B_PC-PS_mix_POPC_nacl.pdf} 
-  \includegraphics[height=\figheightsmall]{../Fig/order_parameters_changes_A-B_PC-PS_mix_POPS_nacl.pdf} 
   \includegraphics[height=\figheightsmall]{../Fig/order_parameters_changes_A-B_PC-PS_mix_POPC_kcl.pdf} 
-  \includegraphics[height=\figheightsmall]{../Fig/order_parameters_changes_A-B_PC-PS_mix_POPS_kcl.pdf} 
-  \caption{\label{fig:delta_ordPar_NaCl_PC-PS_mix} 
-    CAPTION FOR THE FIGURE WITH POPC: The POPC head group order parameters and the P--N vector angle
+  \caption{\label{fig:delta_ordPar_NaCl_PC-PS_mix_PC} 
+    The POPC head group order parameters and the P--N vector angle
     with respect to the membrane normal as a function of POPS content in a bilayer
@@
     and the signs from Ref. \citenum{ferreira16} (only \ce{Na^+} counterions, but shown also in bottom plots).
-    CAPTION FOR THE FIGURE WITH POPS: The POPS head group order parameters, the P--N and C$_\beta$--C$_\gamma$ vector angles (stars)
+  }
+\end{figure*}
+
+\begin{figure*}[!tbp] 
+  \centering 
+  \includegraphics[height=\figheightsmall]{../Fig/order_parameters_changes_A-B_PC-PS_mix_POPS_nacl.pdf} 
+  \includegraphics[height=\figheightsmall]{../Fig/order_parameters_changes_A-B_PC-PS_mix_POPS_kcl.pdf} 
+  \caption{\label{fig:delta_ordPar_NaCl_PC-PS_mix_PS} 
+    The POPS head group order parameters, the P--N and C$_\beta$--C$_\gamma$ vector angles (stars)
     with respect to the membrane normal as a function of POPC content in a bilayer
@@
   }
-  \todo{JOE+SAMULI: split this figure into separate plots for PC and PS}
 \end{figure*}
+
 First, we monitor the response of POPC headgroup order parameters to the increasing amount of POPS (Fig.~\ref{fig:delta_ordPar_NaCl_PC-PS_mix}).
@@
   }
-  \todo{We should plot also the experimental results for POPC with KCl in this figure for the lower bound.} \\
   \end{figure}
@@
 %relative surface excess of sodium with respect to water from negative
-value to $\Gamma ^{w} _{\rm Na}=0.09 \pm ??$ nm$^{-2}$ \todo{Error bar should be added}
+value to $\Gamma ^{w} _{\rm Na}=0.092 \pm 0.005 nm$^{-2}$ 
 for 0.621 M sodium, while value for potassium remains negative,
-$\Gamma ^{w} _{\rm K}=-0.12 \pm ??$ nm$^{-2}$ \todo{Error bar should be added} in POPC:POPS (5:1) mixture (Fig. \ref{fig:nacl-dens_PCPS}).
+$\Gamma ^{w} _{\rm K}=-0.123 \pm 0.005 nm$^{-2}$ in POPC:POPS (5:1) mixture (Fig. \ref{fig:nacl-dens_PCPS}).
 As discussed above, the sodium binding to POPC:POPS (5:1) mixture may be slightly overestimated also in
@@
 indicating that the lower binding affinity than in Lipid17/Dang simulations is more realistic (Fig.~\ref{fig:cacl-dens_PCPS}).
-\todo{JOE: The density plot still needs to be changed to show both Lipid17 and ECC-lipids, and to support this argument.}
 %However, the steep decrease of the order parameters at low calcium concentrations,
@@
   \centering 
-  \includegraphics[width=\figwidth]{../img/ecc_pops/density_profiles_ca_na_cl_wat_phos_models-compar_1-3_CaCl2-series.pdf}
+  \includegraphics[width=\figwidth]{../img/ecc_pops/density_profiles_ca_na_k_cl_wat_phos_ecclipids_lipid17_compar_80and200mMCaCl2.pdf}
   \caption{\label{fig:cacl-dens_PCPS} 
-    Number density profiles of \ce{Ca^{2+}}, \ce{Na^{+}} and \ce{Cl^-} along the normal of the membrane starting at the centre
+    Number density profiles of \ce{Ca^{2+}} and \ce{Cl^-} ions and \ce{Na^{+}} counterions 
+    along the normal of the membrane starting at the centre
     for the negatively charged membrane with the composition 5\,PC:1\,PS
-    at various bulk concentrations of \ce{CaCl2} from simulations with ECC-lipids. 
-    All profiles contain \ce{Na^+} counterions and an additional concentration of \ce{CaCl2}. 
-    In order to visualize the density profiles with a scale comparable to the profile of \ce{Ca^{2+}},  
-    the density profiles of~\ce{Cl^-} ions are divided by 2, and 
-    the density profiles of phosphate groups and water are divided by 5 and 200, respectively.  
+    at 80~mM (top) and 200~mM (bottom) added buffer concentrations of \ce{CaCl2} from simulations with ECC-lipids (solid) and Lipid17 (dashed). 
+    In order to visualize the density profiles with a comparable scale 
+    the density profile of~\ce{Ca^{2+}} ions are divided by 2, and 
+    the density profiles of phosphate groups and water are divided by 4 and 100, respectively.  
   }
-  \todo{SAMULI: We should compare to the density profiles from Lipid17 simulations.}
 \end{figure}

### 2019-04-03 Manuscript/manuscript.tex
Removing TODOs from MS and also from SI

@@
 scale the parameters of atoms only in the headgroup, glycerol backbone, and carbonyl regions \cite{melcr18}.
-Further optimization of parameters was not done in this work.
-\todo{Add here a reference to the parameters.}
+Further optimization of parameters was not done in this work. \cite{ECC-POPC_nacl_cacl2_files}
 % we used exactly the pre-determined scaling factor 
@@
 resolution of +- 0.01. The C--H bond order parameter determined from a given dipolar
-splitting, $\Delta\nu$ (see e.g. Fig. \ref{R-PDLFslices} \todo{Is this reference to the correct figure?})
+splitting, $\Delta\nu$ (see e.g. Fig. \ref{R-PDLFslices} in SI)
 is equal to $\Delta\nu/(0.315\times21.5 kHz)$, where 0.315 is
@@
 \todo{Tiago has to check and update this and figures in SI.}
+
 \begin{figure}[tb!] 
@@
   \todo{The experimental order parameters from sn-1 chain are missing from the figure.} \\
-  \todo{We should use the same color coding for the models as in ECC-POPC paper, now it is opposite.} \\
   \todo{Inverted axis and positive value as label would be more clear for the order parameter figure.}
@@
   \end{tabular} \\
-  \todo{I removed the value from Lipid17/ff99 simulation because this was not listed in Table I
-    and I think that it is more clear if we refer only to the NMRlipids IV result in the text.
-    I also changed the experimental value to correspond the value in the cited paper and added the error bar.}
 \end{table}
@@
 }
-\todo{JOE: add labels with the relative surface excess $\Gamma$, which are promised to be present in the text.
-  SAMULI: I think that these are not promised in the current text and addition is not necessary.} \\
-\todo{JOE: split this figure into two: separate for POPS (to SI) and PC:PS 5:1 (main text)?
-  SAMULI: I feel now that we should keep the figure as it is. Also the POPC:POPS (5:1) results are now referred in the text.
-  However, if something is moved to SI, it should be POPS in main text, mixture in SI.}
 \end{figure} 
@@
 %without the necessity of defining a Gibbs dividing surface. \citep{melcr18, chattorajBOOK}
-The value of $\Gamma ^{w} _{\rm Na}=-0.11 \pm 0.01$ nm$^{-2}$ for 1 M sodium in pure ECC-POPC
+The value of $\Gamma ^{w} _{\rm Na}=-0.11 \pm 0.01 \mathrm{nm}^{-2}$ for 1 M sodium in pure ECC-POPC
 simulation was reported in the previous work \cite{melcr18}.
-The presence of $\sim$17\% POPS in POPC bilayer increases the
+The presence of $\sim$17\$ POPS in POPC bilayer increases the
 %relative surface excess of sodium with respect to water from negative
-value to $\Gamma ^{w} _{\rm Na}=0.092 \pm 0.005 nm$^{-2}$ 
+value to $\Gamma ^{w} _{\rm Na}=0.092 \pm 0.005 \mathrm{nm}^{-2}$ 
 for 0.621 M sodium, while value for potassium remains negative,
-$\Gamma ^{w} _{\rm K}=-0.123 \pm 0.005 nm$^{-2}$ in POPC:POPS (5:1) mixture (Fig. \ref{fig:nacl-dens_PCPS}).
+$\Gamma ^{w} _{\rm K}=-0.123 \pm 0.005 \mathrm{nm}^{-2}$ in POPC:POPS (5:1) mixture (Fig. \ref{fig:nacl-dens_PCPS}).
 As discussed above, the sodium binding to POPC:POPS (5:1) mixture may be slightly overestimated also in
@@
     Clusters of four or more lipids were not observed in either membrane.
-    Previously published simulation data \cite{melcr18} for pure POPC bilayers were taken directly from ?? Add Zenodo reference, if available. 
+    The threshold for counting coordinated lipids in a complex with \ce{Ca^{2+}} was set to $0.3\,\mathrm{nm}$, 
+    as the distance between the cation and the oxygen atoms of the lipids. 
+    Previously published simulation data \cite{melcr18} for pure POPC bilayers were taken directly from \cite{ECC-POPC_nacl_cacl2_files}. 
   }
-  \todo{Add definition used for the coordination} \\
-  \todo{Add link to the Zenodo, if available.}
 \end{figure}

### 2019-04-05 Manuscript/manuscript.tex
Updated actual order parameters figure.

@@
   \todo{J: flip the chemical structure so that it agrees with the left-to-right reading of the plot.} \\
-  \todo{The experimental order parameters from sn-1 chain are missing from the figure.} \\
-  \todo{Inverted axis and positive value as label would be more clear for the order parameter figure.}
+  \todo{The experimental order parameters from sn-1 chain are missing from the figure -- Josef: they are not.
+the parameters are largely overlapping and also only a few of the parameters were resolved for palmitoyl.
+They are visible for C15 and C16 as open diamonds.
+} \\
 \end{figure}

### 2019-05-06 Manuscript/manuscript.tex
Added missing citation, fixed some latex related problems.

--- 2019-04-03 Manuscript/manuscript.tex
@@
 the indirect dimension were recorded. These settings enable to record dipolar slices with a C--H bond order parameter
-resolution of +- 0.01. The C--H bond order parameter determined from a given dipolar
-splitting, $\Delta\nu$ (see e.g. Fig. \ref{R-PDLFslices} in SI)
-is equal to $\Delta\nu/(0.315\times21.5 kHz)$, where 0.315 is
+resolution of $\pm 0.01$. The C--H bond order parameter determined from a given dipolar
+splitting, $\Delta \nu$ (see e.g. Fig. \ref{R-PDLFslices} in SI)
+is equal to $\Delta \nu / (0.315 \times 21.5 \mathrm{kHz})$, where 0.315 is
 the scaling factor of the R18 recoupling sequence and 21.5 kHz is the maximum $^1$H-$^{13}$C dipolar coupling for a C--H bond.
 The refocused-INEPT transfer \cite{morris79,burum80} %[G. Morris and R. Freeman, J. Am. Chem. Soc. 101, 760 (1979);
-\todo{ I could not find this reference [D. Burum, R. Ernst, J. Mag. Res. 39, 163 (1980)]}
 was used for transferring the polarization from $^1$H nuclei to the covalently bond $^{13}$C nuclei with the delays $\tau_1$ = 1.94 ms and $\tau_2$ = 0.97 ms
@@
   \todo{J: flip the chemical structure so that it agrees with the left-to-right reading of the plot.} \\
-  \todo{The experimental order parameters from sn-1 chain are missing from the figure.} \\
-  \todo{Inverted axis and positive value as label would be more clear for the order parameter figure.}
 \end{figure}

### 2019-05-07 Manuscript/manuscript.tex
Small edits in intro

@@
 \title[] 
-  {Effective Inclusion of Electronic Polarization to a Negatively Charged Membrane}
+  { Non-negligible Effects of Electronic Polarization at Negatively Charged Membranes and ions }
 
@@
 
-The phophatidylserine (PS) lipids are most common negatively charged lipids in eukaryotic membranes
+Phosphatidylserine (PS) lipids are the most common negatively charged lipids in eukaryotic membranes
 and important signaling molecules \cite{lemmon08,leventis10,li14}.
@@
 which cause phase separation \cite{hauser77,kurland79,hauser85,feigenson86,mattai89,roux90,roux91,boettcher11}.
-Different theories are difficult to evaluate experimentally,
-especially with physiological ion concentrations which are typically quite low.
+Theories about the interaction between calcium and PS 
+are difficult to evaluate experimentally,
+especially at physiological ion concentrations,
+which are typically too low to yield measurable effects.
+
 More recently, classical molecular dynamics (MD) simulations
@@
 but the results strongly depend on the force field parameters \cite{boettcher11,kucerka14,melcrova16,hallock18,valentine18}.
+For instance, the moieties of PS lipids that interact with calcium cations vary greatly.
 In simulations with the CHARMM36 force field \cite{klauda10,venable13} using the NBfix parameters for calcium \cite{kim16}, 
@@
 cation binding affinity to zwitterionic PC bilayers \cite{catte16} or to the
-mixtures with negatively charge PS lipids \cite{NMRlipidsIV}.
+mixtures with negatively charged PS lipids \cite{NMRlipidsIV}.
 In our recent work~\cite{melcr18}, we were able to improve the cation binding affinity
@@
 
-Here, we extend the ECC approach also to negatively charged 1-palmitoyl-2-oleoyl-sn-glycero-3-phospho-L-serine (POPS) lipid.
+Here, we extend the ECC approach also to the negatively charged 1-palmitoyl-2-oleoyl-sn-glycero-3-phospho-L-serine (POPS) lipid.
 We also complement the available experimental data for the the force field quality

### 2019-05-08 Manuscript/manuscript.tex
cosmetic modification to the title
small todo removal

--- 2019-05-06 Manuscript/manuscript.tex
@@
 controversial because the results strongly depend on force fields parameters \cite{melcrova16,valentine18,hallock18}.
-In ECC-lipid simulation of POPC:POPS (5:1) mixture with 409 mM CaCl$_2$,
-the calcium ions bind twice more likely to the carboxylate than to phosphate moiety of POPS headgroup
-\todo{I am not sure if this is correct, because I do not understand the numbers in the table.} (Table \ref{tab:binding}),
+In ECC-lipids simulation of POPC:POPS (5:1) mixture with 409 mM CaCl$_2$,
+the calcium ions bind more likely 
+to the carboxylate than to phosphate moiety of POPS headgroup (Table \ref{tab:binding}),
 in line with the results without the NBfix corrections in CHARMM36 \cite{hallock18}.
-However, in CHARMM36 simulations without the NBfix correction for calcium,
+However, in CHARMM36 simulations without the NBfix correction for calcium \cite{kim16},
 calcium ions significantly overbind to PC lipids  \cite{catte16} and calcium ions coordinate

### 2019-05-08 Manuscript/manuscript.tex

--- 2019-05-07 Manuscript/manuscript.tex
@@
 \title[] 
-  { Non-negligible Effects of Electronic Polarization at Negatively Charged Membranes and ions }
+  { Non-negligible Effects of Electronic Polarization at Negatively Charged Membranes with Ions }
 
@@
 but the results strongly depend on the force field parameters \cite{boettcher11,kucerka14,melcrova16,hallock18,valentine18}.
+% smoother connection between previous and following statements
 For instance, the moieties of PS lipids that interact with calcium cations vary greatly.

### 2019-05-08 Manuscript/manuscript.tex
Small edits

--- 2019-05-07 Manuscript/manuscript.tex
@@
 \begin{tabular}{ l | c c } 
-        interacting moiety    &  \multicolumn{2}{c}{percentage of bound \ce{Ca^{2+}} } \\
+ exclusive interacting moiety &  \multicolumn{2}{c}{percentage of bound \ce{Ca^{2+}} } \\
 %	\hline
@@
     \hline
-	     PC              &   59   &  100   \\
+    \textbf{PC}              &   59   &  100   \\
 	     PO$_4$    in PC &   41   &   67   \\
@@
     \hline
-	     PS              &    8   &        \\ 
+    \textbf{PS}              &    8   &        \\ 
 	     PO$_4$  in PS   &    2   &        \\
@@
     \hline
-	both PC and PS       &   33   &        \\
+    \textbf{both PC and PS}  &   33   &        \\
   \end{tabular} \\
-  \todo{I have commented out the surface excess values (available in text) and concentrations (now in caption) from the table.} \\
-  \todo{I do not understand the numbers for different moieties. Are these probabilities that the calcium binds only to these moieties and
-    nothing else? In total, 41 \% binds to PS lipids, but only 2 \% and 4 \% to phosphates and COO$^-$, respectively? Where do the rest bind?
-    I think that we should show the total probabilities to bind to different moieties.}
+  %\todo{I have commented out the surface excess values (available in text) and concentrations (now in caption) from the table.} \\
+  %\todo{I do not understand the numbers for different moieties. Are these probabilities that the calcium binds only to these moieties and
+  %  nothing else? In total, 41 \% binds to PS lipids, but only 2 \% and 4 \% to phosphates and COO$^-$, respectively? Where do the rest bind?
+  %  I think that we should show the total probabilities to bind to different moieties.}
 \end{table}

### 2019-05-09 Manuscript/manuscript.tex
Added the populations of calcium at PC&PS:various_moieties into table 4
Convert Table 4 with populations of bound Ca2+ in membrane to a new plot, which is a part of Fig 10, which also contains lipid stoichiometries.

@@
 %to yield converged results in this aspect (see SI and Refs.~\cite{catte16, melcr18, NMRlipidsIV}).
-\begin{table}[tb!] 
-\centering
-\caption{%Bulk concentrations, $C _{Ca}$, and buffer concentrations, $C' _{Ca}$, of Ca$^{2+}$;
-  % relative surface excess of calcium with respect to water ($\Gamma_{Ca}^{\rm water}$); 
-  % and
-  Percentages of the population of bound Ca$^{2+}$ to various lipid moieties in a pure POPC bilayer with 350 mM CaCl$_2$
-  and in a POPC:POPS (5:1) mixture with 409 mM CaCl$_2$. The threshold for counting a contact was set to $0.3\,\mathrm{nm}$, which encompasses the
-  first peak of the radial distribution function between the cations and the oxygen atoms of the lipids. 
-  \label{tab:binding}} 
-\begin{tabular}{ l | c c } 
- exclusive interacting moiety &  \multicolumn{2}{c}{percentage of bound \ce{Ca^{2+}} } \\
-%	\hline
-%	$C _{Ca}\,/\,\mathrm{mM}$  &  $240\pm 10 $  &  $280\pm 10 $  \\
-%	$C'_{Ca}\,/\,\mathrm{mM}$  &  $400\pm 10 $  &  $350\pm 10 $  \\
-%	$\Gamma_{Ca}^{\rm water}\, / \,\mathrm{nm}^{-2}$  &  $0.24 \pm 0.01 $  &  $0.06 \pm 0.01 $  \\
-%	\hline
-%                             &  \multicolumn{2}{c}{ } \\
-                             &  5\,POPC:1\,POPS &  POPC   \\
-    \hline
-    \textbf{PC}              &   59   &  100   \\
-	     PO$_4$    in PC &   41   &   67   \\
-	     carbonyls in PC &   <1   &   ~1   \\
-    \hline
-    \textbf{PS}              &    8   &        \\ 
-	     PO$_4$  in PS   &    2   &        \\
-	     COO$^-$ in PS   &    4   &        \\
-	     carbonyls in PS &   <1   &        \\
-    \hline
-    \textbf{both PC and PS}  &   33   &        \\
-  \end{tabular} \\
-  %\todo{I have commented out the surface excess values (available in text) and concentrations (now in caption) from the table.} \\
-  %\todo{I do not understand the numbers for different moieties. Are these probabilities that the calcium binds only to these moieties and
-  %  nothing else? In total, 41 \% binds to PS lipids, but only 2 \% and 4 \% to phosphates and COO$^-$, respectively? Where do the rest bind?
-  %  I think that we should show the total probabilities to bind to different moieties.}
-\end{table}
 
@@
   \centering 
-  \includegraphics[width=\figwidth]{../img/stoichiometry_CaCl2_comparison_Ecc-lipids_PC-vs-PCPS.pdf} \\ 
-  \caption{\label{fig:cacl_complexes} 
-    Relative probabilities of \ce{Ca^{2+}} ions to coordinate with a certain number of lipids
+  \includegraphics[width=\figwidthsmall]{../img/bound_ca_populations.pdf} 
+  \includegraphics[width=\figwidthsmall]{../img/stoichiometry_CaCl2_comparison_Ecc-lipids_PC-vs-PCPS.pdf} \\ 
+  \caption{ \label{fig:cacl_complexes} 
+  Left: Percentages of the population of bound Ca$^{2+}$ 
+in exclusive contact with various lipid moieties 
+in a POPC:POPS (5:1) bilayer with 409 mM CaCl$_2$. 
+For comparison, in a pure POPC bilayer with 350 mM CaCl$_2$, 
+67\% of bound \ce{Ca^{2+}} is exclusively in contact with \ce{PO_4} group. \cite{melcr18} \\
+    Right: Relative probabilities of \ce{Ca^{2+}} ions to coordinate with a certain number of lipids
     in pure POPC bilayer with 350 mM CaCl$_2$ and in POPC:POPS (5:1) mixture with 400 mM CaCl$_2$.  
     Analysis was done for both systems by considering all lipids (blue and violet) and
-    for POPC:POPS (5:1) mixture also by considering POPC and POPS lipids separately (green and orange).
+    for POPC:POPS (5:1) mixture also by considering POPC and POPS lipids separately (green and orange). \\
     %All lipids were taken into account with the exception of the complexes in light green and orange, 

### 2019-05-13 Manuscript/manuscript.tex
Improved Title of the paper

--- 2019-05-09 Manuscript/manuscript.tex
@@
 \title[] 
-  { Non-negligible Effects of Electronic Polarization at Negatively Charged Membranes with Ions }
+  { Detailed structure of phospholipid bilayers with negatively charged lipids and salt concentrations 
+    from molecular dynamics simulations including the effects of electronic polarization }
+ %{ Non-negligible Effects of Electronic Polarization at Negatively Charged Membranes with Ions }
+ %{ Effective Inclusion of Electronic Polarization to a Negatively Charged Membrane }
 

### 2019-05-13 Manuscript/manuscript.tex
Small edits in Methods

@@
 
-The electronic continuum correction (ECC) is a simple method to implicitly include
-the electronic polarizability into classical molecular dynamics (MD) simulations
-by scaling the electric charges defined in force fields.
-%The polarization of electrons in ECC-lipids
-%uses the implicit method termed electronic continuum correction (ECC). \cite{leontyev09, Pluharova2014, martinek17, melcr18} 
-%In this method, a system of polarizable particles is represented 
-%with an equivalent system of particles with fixed charges,
-%which account for the electronic polarizability \emph{implicitly}.
+Electronic continuum correction (ECC) is an implicit mean-field representation of 
+the electronic polarization in classical MD simulations. 
 For simple ions in water, the electronic polarizability
 can be taken into account by scaling the charge with a constant factor 
-%This is practically achieved by rescaling the charges with a pre-determined scaling factor, $f_q$,
 %\begin{equation}
- $ f_q = \frac{1}{\sqrt{\epsilon_{el}}} \approx 0.75$,
+ \mbox{$ f_q = \frac{1}{\sqrt{\epsilon_{el}}} \approx 0.75$,}
 %\end{equation}
 where  $\epsilon_{el} = 1.78$ is the high-frequency dielectric constant of electrons in water \cite{leontyev09}.
-%,, as
-%The value of $\epsilon_{el} = 1.78$ for water solutions gives the scaling
-%factor of $f_q \approx 0.75$ for aquaeous environments \cite{leontyev09}.
 The scaling of charges with this factor has been recently applied to improve classical models for ions and
-other biomolecules \cite{Pluharova2014, martinek17} \todo{Pavel can probably update this (insulin?)}.
-In addition to the charge, the Lennard-Jones~$\sigma$ was scaled with a factor of $f_\sigma = ??$
+other biomolecules \cite{Pluharova2014, martinek17, duboue2018insulin, Mason2019, Duboue2018MgZn} \todo{Pavel can probably update this (insulin?)}.
+In addition to the charge, the Lennard-Jones~$\sigma$ was scaled with a factor $f_\sigma \leq 1$
 to improve the hydration properties of ions with respect to scattering data in these studies
@@
 Derivation of the correct scaling factor for lipids is more complicated, because
-%In contrast to the total charge of ions,
-the partial charges depend on the methods used to derive the force field.
-For example, the partial charges may already include some electronic polarizability effects
-and the scaling factor should then be larger than the theoretically derived value for ions in water, $f_q \approx 0.75$.
-Before the ECC theory was rigorously derived, similar idea was employed in early classical 
+the partial charges depend on the methods used to derive the respective force fields.
+For example, the partial charges may already include the effects of electronic polarizability to some extent,
+and hence, the scaling factor can be larger than the theoretically derived value for ions in water, $f_q \approx 0.75$.
+Before the ECC theory was rigorously derived, similar idea was already employed in early classical 
 MD simulations of lipids and surfactants using the scaling factor of 0.5 for atomic charges  \cite{jonsson86,egberts94, berendsen1996}. 
@@
 
-Here, we apply the ECC to the Amber based lipid17 parameters \cite{lipid17-future} of POPS lipid with a monovalent negative charge.
-Because POPS carries a physical total charge as ions in water, we apply the theoretically derived scaling factor for
-ions $f_q = 0.75$ \cite{leontyev09} to the partial charges of POPS. The lipids with a total charge of -0.75 also neutralize
-the counterion charges of +0.75 in simulations with ECC-ions \cite{kohagen16}.
+Here, we apply the ECC to the Amber-based lipid17 parameters \cite{lipid17-future} of POPS lipid with a monovalent negative charge.
+Because POPS carries a physical total charge as aqueous ions, we apply the theoretically derived scaling factor for
+ions $f_q = 0.75$ \cite{leontyev09} to the partial charges of POPS. 
+The lipids with a total charge of -0.75 are also neutralized
+by the counterion charges of +0.75 in simulations with ECC-ions \cite{Pluharova2014, kohagen16, martinek17}.
 Following our previous work for POPC, we use the scaling factor of $f_\sigma = 0.89$ for the Lennard-Jones~$\sigma$ parameters and
-scale the parameters of atoms only in the headgroup, glycerol backbone, and carbonyl regions \cite{melcr18}.
-Further optimization of parameters was not done in this work. \cite{ECC-POPC_nacl_cacl2_files}
-% we used exactly the pre-determined scaling factor 
-%to represent the effects of electronic polarization
-%yielding a total charge of $-0.75$ in the ECC formalism. \cite{martinek17, leontyev09, melcr18}
-Similarly to the ECC-POPC model \cite{melcr18}, we applied the ECC only to the
-non-tail atoms, i.e., to atoms in the headgroup, glycerol backbone, and carbonyl regions. 
-%polar parts of the phospholipid,
-%i.e., charges on all atoms except on alkyl tails, 
-%which do not encounter direct contacts with ions.
-
-%\subsection{Validation of membrane properties}
-% SAMULI: I have dissolved this section into results and discussion
+use it to scale the parameters of atoms only in the headgroup, glycerol backbone, and carbonyl regions \cite{melcr18, ECC-POPC_nacl_cacl2_files}.
+Further optimization of parameters was not done in this work.
 

### 2019-05-13 Manuscript/manuscript.tex
Formal LateX related changes in the Experimental methods section

@@
 The experiment was done using a Bruker Avance III 400 spectrometer operating at a 1H Larmor frequency of 400.03 MHz
-equipped with a standard 4 mm CP-MAS HXY probe. The set up of the R-PDLF experiment was the following
+equipped with a standard 4~mm CP-MAS HXY probe. The set up of the R-PDLF experiment was the following
 (using the notation from figures 1c and 2c in the original publication describing the R-PDLF experiment \cite{dvinskikh04}).
 The magic angle spinning (MAS) frequency used was 5.15 kHz. The recoupling pulses in the R18 blocks had
-therefore a length of 10.79 us which correspond to a nutation frequency of 46.35 kHz. Increments of 18x2 recoupling
-pulses where used giving a spectral width of 2.6 kHz in the indirect dimension and a total number of 32 points in
+therefore a length of $10.79 \, \mu \mathrm{s}$, which correspond to a nutation frequency of 46.35~kHz. Increments of $18 \times 2$ recoupling
+pulses where used giving a spectral width of 2.6~kHz in the indirect dimension and a total number of 32 points in
 the indirect dimension were recorded. These settings enable to record dipolar slices with a C--H bond order parameter
@@
 splitting, $\Delta \nu$ (see e.g. Fig. \ref{R-PDLFslices} in SI)
-is equal to $\Delta \nu / (0.315 \times 21.5 \mathrm{kHz})$, where 0.315 is
+is equal to $\Delta \nu / (0.315 \times 21.5 \, \mathrm{kHz})$, where 0.315 is
 the scaling factor of the R18 recoupling sequence and 21.5 kHz is the maximum $^1$H-$^{13}$C dipolar coupling for a C--H bond.
 The refocused-INEPT transfer \cite{morris79,burum80} %[G. Morris and R. Freeman, J. Am. Chem. Soc. 101, 760 (1979);
-was used for transferring the polarization from $^1$H nuclei to the covalently bond $^{13}$C nuclei with the delays $\tau_1$ = 1.94 ms and $\tau_2$ = 0.97 ms
-(set as multiples of the MAS rotation period). The pulses for the refocused-INEPT sequence had a nutation frequency equal to 63.45 kHz.
+was used for transferring the polarization from $^1$H nuclei to the covalently bond $^{13}$C nuclei with the delays $\tau_1 = 1.94$~ms and $\tau_2 = 0.97$~ms
+(set as multiples of the MAS rotation period). The pulses for the refocused-INEPT sequence had a nutation frequency equal to 63.45~kHz.
 For acquiring the spectra, a total number of 1024 transients were recorded for each point in the indirect dimension, using an acquisition
-time of 0.1 s under SPINAL64 $^1$H decoupling \cite{fung00} with a nutation frequency of 50 kHz, and with dwell time giving a spectral
+time of 0.1~s under SPINAL64 $^1$H decoupling \cite{fung00} with a nutation frequency of 50~kHz, and with dwell time giving a spectral
 width of 200 ppm for the $^{13}$C chemical shift dimension. The set of free induction decays recorded were then processed as
-described e.g. in Ref. \citenum{ferreira08}. The dipolar splittings in the crowded spectral region
+described e.g. in Ref.~\citenum{ferreira08}. The dipolar splittings in the crowded spectral region
 between 29 and 30 ppm were assigned based on the previous assignments for POPC \cite{ferreira13} 

### 2019-05-13 Manuscript/manuscript.tex
Improved table with the list of simulations

@@
 \caption{  List of MD simulations reporting their respective
+           lipid composition,
+           used counterions,
            added buffer concentrations,
@@
          }\label{tbl:sim-list}
-\begin{tabular}{l | r r r r | r r | c c }
-                          &  \multicolumn{4}{c}{added buffer concentration / mM}    & \multicolumn{2}{c}{no. molecules} &  \multicolumn{2}{c}{simulated with model}  \\
-PC:PS (ratio)/counterions &  \ce{K^+}  &  \ce{Na^+} & \ce{Ca^{2+}} & \ce{Cl^-}      & \ce{H2O} &  PC:PS                 &  ECC-lipids  &  Lipid17    \\
+\begin{tabular}{l l | r r r r | r r | c c }
+   \multicolumn{2}{c}{ }  &  \multicolumn{4}{c}{added buffer conc. / mM}    & \multicolumn{2}{c}{no. molecules} &  \multicolumn{2}{c}{simulated with model}  \\
+PC:PS ratio & counterions &  \ce{K^+}  &  \ce{Na^+} & \ce{Ca^{2+}} & \ce{Cl^-}      & \ce{H2O} &  PC:PS                 &  ECC-lipids  &  Lipid17    \\
   \hline
-PC:PS (0:1)/K+   &      0  &      0  &      0  &      0  &  3600  &  0:72  &  X  &  X  \\ 
-PC:PS (0:1)/Na+  &      0  &      0  &      0  &      0  &  3600  &  0:72  &  X  &  X  \\ 
+0:1 & \ce{K^+}   &      0  &      0  &      0  &      0  &  3600  &  0:72  &  X  &  X  \\ 
+0:1 & \ce{Na^+}  &      0  &      0  &      0  &      0  &  3600  &  0:72  &  X  &  X  \\ 
   \hline
-PC:PS (10:1)/K+  &      0  &      0  &      0  &      0  &  3600  &  60:6  &  X  &  -  \\ 
-PC:PS (10:1)/Na+ &      0  &      0  &      0  &      0  &  3600  &  60:6  &  X  &  -  \\ 
+10:1 & \ce{K^+}  &      0  &      0  &      0  &      0  &  3600  &  60:6  &  X  &  -  \\ 
+10:1 & \ce{Na^+} &      0  &      0  &      0  &      0  &  3600  &  60:6  &  X  &  -  \\ 
   \hline
-PC:PS (1:1)/K+   &      0  &      0  &      0  &      0  &  5253  &  64:64  &  X  &  X  \\ 
-PC:PS (1:1)/Na+  &      0  &      0  &      0  &      0  &  5253  &  64:64  &  X  &  X  \\ 
+1:1 & \ce{K^+}   &      0  &      0  &      0  &      0  &  5253  &  64:64  &  X  &  X  \\ 
+1:1 & \ce{Na^+}  &      0  &      0  &      0  &      0  &  5253  &  64:64  &  X  &  X  \\ 
   \hline
-PC:PS (4:1)/K+   &      0  &      0  &      0  &      0  &  3600  &  48:12  &  X  &  -  \\ 
-PC:PS (4:1)/Na+  &      0  &      0  &      0  &      0  &  3600  &  48:12  &  X  &  -  \\ 
+4:1 & \ce{K^+}   &      0  &      0  &      0  &      0  &  3600  &  48:12  &  X  &  -  \\ 
+4:1 & \ce{Na^+}  &      0  &      0  &      0  &      0  &  3600  &  48:12  &  X  &  -  \\ 
   \hline
-PC:PS (5:1)/K+   &      0  &      0  &      0  &      0  &  3600  &  60:12  &  X  &  X  \\ 
-PC:PS (5:1)/Na+  &      0  &      0  &      0  &      0  &  3600  &  60:12  &  X  &  X  \\ 
-PC:PS (5:1)/Na+  &      0  &      0  &     78  &     78  &  3561  &  60:12  &  X  &  X  \\ 
-PC:PS (5:1)/Na+  &      0  &      0  &    125  &    125  &  3561  &  60:12  &  X  &  X  \\ 
-PC:PS (5:1)/Na+  &      0  &      0  &    202  &    202  &  3561  &  60:12  &  X  &  X  \\ 
-PC:PS (5:1)/Na+  &      0  &      0  &    409  &    409  &  3522  &  60:12  &  X  &  X  \\ 
-PC:PS (5:1)/Na+  &      0  &      0  &    621  &    621  &  3483  &  60:12  &  X  &  X  \\ 
-PC:PS (5:1)/Na+  &      0  &    621  &      0  &    621  &  3483  &  60:12  &  X  &  X  \\ 
-PC:PS (5:1)/Na+  &      0  &   1510  &      0  &   1510  &  3377  &  60:12  &  X  &  X  \\ 
-PC:PS (5:1)/Na+  &      0  &   3002  &      0  &   3002  &  3213  &  60:12  &  X  &  X  \\ 
-PC:PS (5:1)/Na+  &    621  &      0  &      0  &    621  &  3483  &  60:12  &  X  &  X  \\ 
-PC:PS (5:1)/Na+  &   1510  &      0  &      0  &   1510  &  3377  &  60:12  &  X  &  X  \\ 
-PC:PS (5:1)/Na+  &   3002  &      0  &      0  &   3002  &  3213  &  60:12  &  X  &  X  \\ 
+5:1 & \ce{K^+}   &      0  &      0  &      0  &      0  &  3600  &  60:12  &  X  &  X  \\ 
+5:1 & \ce{Na^+}  &      0  &      0  &      0  &      0  &  3600  &  60:12  &  X  &  X  \\ 
+5:1 & \ce{Na^+}  &      0  &      0  &     78  &     78  &  3561  &  60:12  &  X  &  X  \\ 
+5:1 & \ce{Na^+}  &      0  &      0  &    125  &    125  &  3561  &  60:12  &  X  &  X  \\ 
+5:1 & \ce{Na^+}  &      0  &      0  &    202  &    202  &  3561  &  60:12  &  X  &  X  \\ 
+5:1 & \ce{Na^+}  &      0  &      0  &    409  &    409  &  3522  &  60:12  &  X  &  X  \\ 
+5:1 & \ce{Na^+}  &      0  &      0  &    621  &    621  &  3483  &  60:12  &  X  &  X  \\ 
+5:1 & \ce{Na^+}  &      0  &    621  &      0  &    621  &  3483  &  60:12  &  X  &  X  \\ 
+5:1 & \ce{Na^+}  &      0  &   1510  &      0  &   1510  &  3377  &  60:12  &  X  &  X  \\ 
+5:1 & \ce{Na^+}  &      0  &   3002  &      0  &   3002  &  3213  &  60:12  &  X  &  X  \\ 
+5:1 & \ce{Na^+}  &    621  &      0  &      0  &    621  &  3483  &  60:12  &  X  &  X  \\ 
+5:1 & \ce{Na^+}  &   1510  &      0  &      0  &   1510  &  3377  &  60:12  &  X  &  X  \\ 
+5:1 & \ce{Na^+}  &   3002  &      0  &      0  &   3002  &  3213  &  60:12  &  X  &  X  \\ 
   \hline

### 2019-05-13 Manuscript/manuscript.tex
Improvements in the Methods section (small)

@@
   \hline
-0:1 & \ce{K^+}   &      0  &      0  &      0  &      0  &  3600  &  0:72  &  X  &  X  \\ 
-0:1 & \ce{Na^+}  &      0  &      0  &      0  &      0  &  3600  &  0:72  &  X  &  X  \\ 
-  \hline
-10:1 & \ce{K^+}  &      0  &      0  &      0  &      0  &  3600  &  60:6  &  X  &  -  \\ 
-10:1 & \ce{Na^+} &      0  &      0  &      0  &      0  &  3600  &  60:6  &  X  &  -  \\ 
+0:$\infty$ & \ce{K^+}   &      0  &      0  &      0  &      0  &  3600  &  0:72  &  X  &  X  \\ 
+0:$\infty$ & \ce{Na^+}  &      0  &      0  &      0  &      0  &  3600  &  0:72  &  X  &  X  \\ 
   \hline
@@
   \hline
+10:1 & \ce{K^+}  &      0  &      0  &      0  &      0  &  3600  &  60:6  &  X  &  -  \\ 
+10:1 & \ce{Na^+} &      0  &      0  &      0  &      0  &  3600  &  60:6  &  X  &  -  \\ 
 \end{tabular}
@@
 the default Amber ion parameters \cite{aqvist90} led to the artificial clustering of ions in solution \cite{NMRlipidsIV}.
-In the ECC-lipid simulations,
+In the ECC-lipids simulations,
 the ECC-POPC parameters~\cite{melcr18} (available from Ref. \citenum{ECC-POPC_nacl_cacl2_files}), 
@@
     equilibration time  & 50~ns  \\
-    total simulation time     & 1~$\mu$s  \\
+    total simulation time     & $\geq 1 \mu$s  \\
     temperature         & 298~K       \\
@@
 
-The C--H bond order parameters were calculated directly from Eq. \ref{OPequation}.
-Average over the time for individual lipids was calculated first and
-the standard error of the mean of the average over lipids was then used as the
+The C--H bond order parameters were calculated directly from Eq.~\ref{OPequation}.
+The time average was first calculated for each lipid, and
+the standard error of the mean from these individual values was then used as the
 error estimate \cite{botan15,ollila16,NMRlipidsIV}.
 Python program ({\tt scripts/calcOrderParameters.py}) that uses the
-MDAnalysis library \cite{agrawal11,gowers16} is available in Ref. \citenum{MATCHgit}. 
+MDAnalysis library \cite{agrawal11,gowers16} is available in Ref.~\citenum{MATCHgit}. 
 The ion number density profiles were calculated using the {\tt gmx density} tool
@@
 multilamellae are expected to be sufficiently similar in the used simulations and reference experiments \cite{roux90}.
-
-%The simulated system consisted of lipids (POPC and POPS) forming a single lipid bilayer,
-%an amount of water molecules comparable to the hydration levels in experiments, 
-%\todo{Check this claim. SAMULI: Check discussin in the NMRlipids IV manuscript}
-%\todo{EDIT: Taken from NMRlipids4 as an interesting information, to be edited.}
-%In the experimental $^2$H NMR literature data used in this work~\cite{browning80,roux90},
-%the lipids were first soluted to the buffer and then centrifuged to a pellet which
-%was used in the measurements. Such samples have lower lipid concentration
-%(approximately 10~wt~\% of lipids~\cite{browning80,roux88,roux90}) than 
-%gravimetric samples (60~wt~\%) and simulations (approximately 50-60~wt~\%) in this work.
-%Larger multilamellar repeat distances are expected in the samples with lower lipid
-%concentrations due to the swelling caused by electrostatic repulsion in pure PS lipid systems~\cite{millman82}.
-%However, the PS headgroup order parameters measured from gravimetric sample in this work
-%are in good agreement with the results from centrifuged sample in the literature \cite{browning80} (Fig. \ref{HGorderParameters}).
-%Furthermore, the equlibrium repeat distance rapidly decreases with the addtion of monovalent
-%salts and is close to the simulation box sizes already above 500~mM concentrations~\cite{millman82,rand89}.
-%Therefore, the hydration levels of multilamellae are expected to be sufficiently similar
-%in the used simulations and reference experiments.
 
@@
     Number density profiles of \ce{K^{+}} and \ce{Na^{+}} counterions along the membrane normal axis
-    in ECC-lipid (solid lines) and Lipid17/Dang (dashed lines) simulations of pure POPS (left) or POPC:POPS (5:1, right) bilayers.  
+    in ECC-lipids (solid lines) and Lipid17/Dang (dashed lines) simulations of pure POPS (left) or POPC:POPS (5:1, right) bilayers.  
     The density profiles of phosphate groups and water are divided by 4 and 100, respectively.  
@@
     with respect to the membrane normal as a function of POPS content in a bilayer
-    from ECC-lipid and Lipid17/Dang simulations with \ce{Na^+} (top) and \ce{K^+} (bottom) counterions.
+    from ECC-lipids and Lipid17/Dang simulations with \ce{Na^+} (top) and \ce{K^+} (bottom) counterions.
     Experimental order parameter values are from Ref. \citenum{scherer87}
@@
     with respect to the membrane normal as a function of POPC content in a bilayer
-    from ECC-lipid and Lipid17/Dang simulations with \ce{Na^+} (top) and \ce{K^+} (bottom) counterions.
+    from ECC-lipids and Lipid17/Dang simulations with \ce{Na^+} (top) and \ce{K^+} (bottom) counterions.
     Experimental order parameter values and signs for pure POPS are from Ref. \citenum{NMRlipidsIV} and values
@@
     with respect to the membrane normal of POPC (left) and POPS (right) in a POPC:POPS (5:1) bilayer 
-    as a function of \ce{NaCl} concentration from ECC-lipid and Lipid17/Dang simulations at 298 K.
+    as a function of \ce{NaCl} concentration from ECC-lipids and Lipid17/Dang simulations at 298 K.
     The y-axis for the $\alpha$-carbon results of POPS (middle right) is transferred
@@
     with respect to the membrane normal of POPC (left) and POPS (right) in a POPC:POPS (5:1) bilayer 
-    as a function of \ce{KCl} concentration from ECC-lipid and Lipid17/Dang simulations 
+    as a function of \ce{KCl} concentration from ECC-lipids and Lipid17/Dang simulations 
     compared with experimental values from Ref. \citenum{roux90} (signs from Ref. \citenum{ferreira16}) at 298 K.
@@
 As discussed above, the sodium binding to POPC:POPS (5:1) mixture may be slightly overestimated also in
-ECC-lipid simulations. The response of headgroup order parameters to the addition of NaCl are
+ECC-lipids simulations. The response of headgroup order parameters to the addition of NaCl are
 slightly overestimated also in ECC-POPC simulations (Fig 3 in Ref. \citenum{melcr18}),
@@
 which underestimated the headgroup order parameter response.
-In ECC-lipid simulations, the headgroup responses are in better agreement with experiments (Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}),
+In ECC-lipids simulations, the headgroup responses are in better agreement with experiments (Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}),
 indicating that the lower binding affinity than in Lipid17/Dang simulations is more realistic (Fig.~\ref{fig:cacl-dens_PCPS}).
@@
 %observed especially in negatively charged bilayers \cite{borle85,macdonald87,roux90},
-%may not be fully captured by the ECC-lipid simulations.
+%may not be fully captured by the ECC-lipids simulations.
 The increased calcium binding affinity to bilayers containing PS lipids is illustrated in
-ECC-lipid simulations by the increasing relative surface excess of calcium with respect to water from
+ECC-lipids simulations by the increasing relative surface excess of calcium with respect to water from
 $\Gamma^{w}_{Ca} = 0.06\mathrm{nm^{-2}}$ for pure POPC bilayer with 350 mM CaCl$_2$ \cite{melcr18}
@@
     with respect to the membrane normal of POPC (left) and POPS (right) in a POPC:POPS (5:1) bilayer 
-    as a function of \ce{CaCl} concentration from ECC-lipid and Lipid17/Dang simulations 
+    as a function of \ce{CaCl} concentration from ECC-lipids and Lipid17/Dang simulations 
     compared with experimental values from Ref. \citenum{roux90} (signs from Ref. \citenum{ferreira16}) at 298 K.
@@
 which underestimated the binding affinity.
-In ECC-lipid simulations, the changes of the PS headgroup order parameters are not overestimated,
+In ECC-lipids simulations, the changes of the PS headgroup order parameters are not overestimated,
 but the strong dependence on low concetrations of CaCl$_2$ is not fully reproduced (Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}).
@@
 
-In ECC-lipid simulations, the average orientation of P--N vectors of both POPC and POPS headgroups tilts more perpendicular to the membrane surface
+In ECC-lipids simulations, the average orientation of P--N vectors of both POPC and POPS headgroups tilts more perpendicular to the membrane surface
 by 11$^\circ$ and  3$^\circ$, respectively, upon addition of 620~mM CaCl$_2$ to POPC:POPS (5:1) mixture (Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}),
@@
 Together with the smaller error bars and shorter residence times (Fig.~\ref{fig:hist_residence_times}) in
-ECC-lipid simulations than in typically observed for other force fields \cite{javanainen17,melcr18,NMRlipidsIV},
+ECC-lipids simulations than in typically observed for other force fields \cite{javanainen17,melcr18,NMRlipidsIV},
 this suggests that the ECC accelerates the equlibration of ions at lipid bilayer interface.
@@
 %
-Our 1 $\mu$s simulations seems to be sufficiently long for the ECC-lipid simulations, because
+Our 1 $\mu$s simulations seems to be sufficiently long for the ECC-lipids simulations, because
 90\% of the calcium residence times are shorter than $60\,\mathrm{ns}$ for pure POPC bilayer
@@
 Interestingly, the calcium residence times are 3-4 times longer in POPC:POPS (5:1) mixture than in pure POPC.
-In ECC-lipid simulations, on average 41\% of the total population of bound calcium cations is in contact with PS lipids
+In ECC-lipids simulations, on average 41\% of the total population of bound calcium cations is in contact with PS lipids
 with 8\% bound only to them, even though the negatively charged membrane contains only 18\% of POPS (Table \ref{tab:binding}).
@@
 controversial because the results strongly depend on force fields parameters \cite{melcrova16,valentine18,hallock18}.
-In ECC-lipid simulation of POPC:POPS (5:1) mixture with 409 mM CaCl$_2$,
+In ECC-lipids simulation of POPC:POPS (5:1) mixture with 409 mM CaCl$_2$,
 the calcium ions bind twice more likely to the carboxylate than to phosphate moiety of POPS headgroup
@@
 with even four distinct PS lipids \cite{hallock18}, while coordination with more than two PS lipids
-was not observed in our ECC-lipid simulations (Fig.~\ref{fig:cacl_complexes}).
+was not observed in our ECC-lipids simulations (Fig.~\ref{fig:cacl_complexes}).
 Furthermore, the POPS:POPC:Ca$^{2+}$ (7:3:7) mixture with relatively high
@@
 which may lead to the complexation of \ce{Ca^{2+}} cations and PS lipids that is
-not observed with lower PS content used in ECC-lipid simulations in this work \cite{hauser77,kurland79,hauser85,feigenson86,mattai89,roux90,roux91}.
+not observed with lower PS content used in ECC-lipids simulations in this work \cite{hauser77,kurland79,hauser85,feigenson86,mattai89,roux90,roux91}.
 %While POPC interacts with the calcium cations almost entirely through its phosphate group 
@@
 simulations with the Berger force field \cite{melcrova16} %\cite{berger97,mukhopadhyay04}
-was not observed in ECC-lipid simulations in here nor in the previous work with pure POPC \cite{melcr18}.
+was not observed in ECC-lipids simulations in here nor in the previous work with pure POPC \cite{melcr18}.
 %This is in contrast to 
@@
 rotational-echo double-resonance (REDOR) experiments \cite{hallock18} or fluorescent and vibrational sum frequency spectroscopy \cite{melcrova16}.
-Because the ECC-lipid model gives the best agreement with the experimental headgroup order parameters
+Because the ECC-lipids model gives the best agreement with the experimental headgroup order parameters
 with various ion conctenrations among the available models, we believe that it gives the currently best available
-interpretation of the experimental data. Athough the ECC-lipid model gives the most realistic picture of calcium binding to
+interpretation of the experimental data. Athough the ECC-lipids model gives the most realistic picture of calcium binding to
 lipid bilayers containing PS lipids with respect to NMR order parameters, it does not

### 2019-05-13 Manuscript/manuscript.tex
Edited section Results: agreement w exp. and added dihedral angle data (associated plot TBA)

@@
 
-\subsection{Structural parameters of a pure ECC-POPS model membrane: Agreement with experiments}
+\subsection{ECC-POPS improves agreement in experimental structural parameters of a pure POPS bilayer }
 
@@
 Here, we measure also the C--H bond order parameters of the acyl chain region using natural abundance $^{13}$C NMR.
-We first determined the order parameters of each peak corresponding different C--H bonds detected
+We first determined the order parameters of each peak corresponding to different C--H bonds detected
 in the first dimension (Fig. \ref{R-PDLF}) from the dipolar slices in the indirect dimension (Fig. \ref{R-PDLFslices}).
@@
 In previous study \cite{NMRlipidsIV}, the Lipid17 force field with {\AA}qvist ions~\cite{aqvist90}
-gave a larger area per molecule (57 \AA$^2$) for POPS being closer to the experimental value,
+gave a larger area per molecule (57~\AA$^2$) for POPS being closer to the experimental value,
 which is probably due to the weaker counterion binding affinity of {\AA}qvist ions.
@@
 The ECC-POPS simulation gives a good agreement with the X-ray scattering form factors and experimental
-acyl chain order parameters (see Fig.~\ref{simVSexpNOions_POPS}), indicating the the bilayer dimensions and acyl chain
+acyl chain order parameters (see Fig.~\ref{simVSexpNOions_POPS}), indicating that the bilayer dimensions and acyl chain
 conformations are well described by the force field. The area per lipid from ECC-POPS simulation is slightly
 smaller than the value reported from SDP model \cite{kucerka14} (table \ref{tab:apls}), but the values
-are within each others error bars.
-
-% In order to obtain the area per lipid from experiments, modeling is used on top of the scattering factors \citep{kucerka14}. 
-%In ocntrast, this property is easily extracted directly from simulations with flat phospholipid bilayers. 
-%We compare the values from experiments and simulations in Table~\ref{tab:apls}. 
-%While the agreement between the scattering form factors from the simulation of a pure POPS bilayer and experiment 
-%are excellent (Fig.~\ref{simVSexpNOions_POPS}), there is a non-negligible difference between the values of the area per lipid in Table~\ref{tab:apls}. 
-%Since both values are derived from the scattering form factors through modeling of the electron density of the bilayer,
-%we cannot decide, which of the values is more reliable.
+agree within their error estimates.
 
@@
 \centering
-  \caption{Area per lipid (APL) of POPS bilayers from simulations with different force fields at 298~K with \ce{Na^+} counterions. \label{tab:apls} } 
+  \caption{Area per lipid of POPS bilayers, $A_L$, from simulations with different force fields at 298~K with \ce{Na^+} counterions. \label{tab:apls} } 
   \begin{tabular}{l|c } 
     \multicolumn{2}{c}{POPS} \\
-    model          & APL / Å$^2$    \\ 
+    model          & $A_L$ / Å$^2$    \\ 
     \hline 
@@
 and {\AA}qvist \cite{NMRlipidsIV} ions. The values from ECC-POPS simulation are closer,
-but not in full agreement with experiments (Fig. \ref{simVSexpNOions_POPS}).
+but not in full agreement with experiments (Fig.~\ref{simVSexpNOions_POPS}).
 Especially, the experimentally observed more negative values of the $\beta$-carbon
@@
 same carbon \cite{ollila16}) of the $\alpha$-carbon in POPS than in POPC \cite{NMRlipidsIV}
-are not reproduced by neither of the models here. These differences are previously related
-to a more rigid structure of PS headgroup and the stiffness of the C$_\alpha$-C$_\beta$-C$_\gamma$-O$_\gamma$
-dihedral, which is \todo{the dihedral should be plotted and sentence finished.}
-
-%The head group order parameters $\alpha$ and $\beta$ are highly relevant for this work,
-%as they are being used in the electrometer concept \cite{altenbach84, catte16, melcr18}.
-%For POPC in pure water, the order parameter $\beta$ agrees well with the experiment, while the order parameter $\alpha$ is somewhat lower.
-
-In conclusion, the ECC-POPS model reproduces the lipid bilayer dimensions and
-acyl chain conformations in good agreement with experiments, but has room for
-improvement in the glycerol backbone and headgroup region. Therefore,
-its structural accuracy is in the same level as in other state-of-the-art
-lipid force fields \cite{botan15, ollila16, Pluhackova2016, NMRlipidsIV}.
+are not reproduced by neither of the models here. 
+These differences are previously related
+to a more rigid structure of PS headgroup and 
+the stiffness of the C$_\alpha$-C$_\beta$-C$_\gamma$-O$_\gamma$ dihedral. 
+Interestingely, this dihedral angle is more flexible in ECC-POPS compared to Lipid17 
+suggesting that its rigidity is overestimated in Lipid17 due to the lack of electronic polarizability,
+while the same dihedral parameters yield too small forking in ECC-POPS.
+\todo{the dihedral to be plotted.}
+Although the ECC-POPS model reproduces the lipid bilayer dimensions and
+acyl chain conformations in a good agreement with experiments, 
+there is a room for improvements especially in the glycerol backbone and headgroup region,
+e.g. in the parameters of the associated dihedral angles. 
+\cite{botan15, ollila16, Pluhackova2016, NMRlipidsIV}.
 

### 2019-05-14 Manuscript/manuscript.tex
Included the dihedral distribution figure in the text.

@@
 to a more rigid structure of PS headgroup and 
-the stiffness of the C$_\alpha$-C$_\beta$-C$_\gamma$-O$_\gamma$ dihedral. 
+the stiffness of the C$_\alpha$-C$_\beta$-C$_\gamma$-O$_\gamma$ dihedral
+plotted in Fig.~\ref{fig:dihedral}. 
 Interestingely, this dihedral angle is more flexible in ECC-POPS compared to Lipid17 
@@
 while the same dihedral parameters yield too small forking in ECC-POPS.
-\todo{the dihedral to be plotted.}
 Although the ECC-POPS model reproduces the lipid bilayer dimensions and
@@
 \cite{botan15, ollila16, Pluhackova2016, NMRlipidsIV}.
+
+\begin{figure}[tbp!] 
+  \centering 
+  \includegraphics[width=\figwidth]{../img/dihedral_angle_distribution_Ca-Cb-Cg-Og_ECC-L17.pdf}
+  \caption{\label{fig:dihedral}
+    Distribution of dihedral angle \ce{C_\alpha-C_\beta-C_\gamma-O_\gamma} in POPS 
+    from simulations of a pure POPS bilayer with \ce{Na+} counterions using Lipid17 and ECC-lipids. 
+}
+\end{figure}
 

### 2019-05-14 Manuscript/manuscript.tex
Corrected figures with chemical structures of POPC and POPS and placed them in appropriate places.

@@
   \includegraphics[width=\figwidth]{../img/ecc_pops/Order-parameters_form-factors_exp-L17-ECC-lipids.pdf}
-  \includegraphics[width=\figwidth]{../img/ecc_pops/pops_chemfig.pdf} 
+  \includegraphics[width=\figwidth]{../img/POPSstructure.pdf} 
 \hfill
@@
   }  
-  \todo{J: flip the chemical structure so that it agrees with the left-to-right reading of the plot.} \\
 \end{figure}
@@
 
-\begin{figure}[tb!] 
-  \centering 
-  \includegraphics[width=\figwidth]{../Fig/lipids_chemfig_POPC_POPS.pdf} 
-  \caption{ \label{fig:chemstruct_pc_ps} 
-            The chemical structure of POPC and POPS 
-            including the labeling of the carbon segments
-            used in the measurements of order parameteres. 
-          }
-          \todo{The POPS structure and labeling is already in figure \ref{simVSexpNOions_POPS}.
-            The POPC structure and labeling could be added on top of the new version of figure \ref{fig:delta_ordPar_NaCl_PC-PS_mix} with only PC,
-            where PC results occur in the first time, and this figure could be removed.} \\
-          \todo{$\gamma$ carbon of POPS is the COO- minus carbon. The $\gamma$ label should be moved there.}
-\end{figure}
-
 The head group and acyl chain order parameters within ECC-lipids
@@
   \includegraphics[height=\figheightsmall]{../Fig/order_parameters_changes_A-B_PC-PS_mix_POPC_kcl.pdf} 
+  \includegraphics[height=\figheightsmall]{../img/POPCstructure.pdf} 
   \caption{\label{fig:delta_ordPar_NaCl_PC-PS_mix_PC} 

### 2019-05-17 Manuscript/manuscript.tex
Edits in Results section one (form fact. etc.)

@@
     \textbf{Top:} X-ray scattering form factors from simulations with Lipid17/Dang \citep{lipid17-future, dang2006} and 
-    the ECC-POPS/ECC-ions \cite{martinek17, Pluhackova2016} compared with experiments~\citep{kucerka14} at 298~K. 
+    ECC-POPS/ECC-ions \cite{martinek17, Pluhackova2016} compared with experiments~\citep{kucerka14} at 298~K. 
     \textbf{Middle:} Order parameters of POPS head group, glycerol backbone and acyl chains  
-    from simulations with the Lipid17 \citep{lipid17-future} and the ECC-POPS models 
+    from the same simulations 
     compared with experiments at 298~K. \citep{NMRlipidsIV}
@@
   }  
+\todo{Add label $\gamma$ to the figure -- it is used in Fig with the dihedral angle, 
+and it can be also used in the description of COO- vector. }
 \end{figure}
@@
     from simulations of a pure POPS bilayer with \ce{Na+} counterions using Lipid17 and ECC-lipids. 
+    Labelling of POPS segments is in Fig.~\ref{simVSexpNOions_POPS}. 
 }
@@
 In ECC-lipids simulation of POPC:POPS (5:1) mixture with 409 mM CaCl$_2$,
-the calcium ions bind twice more likely to the carboxylate than to phosphate moiety of POPS headgroup
-\todo{I am not sure if this is correct, because I do not understand the numbers in the table.} (Table \ref{tab:binding}),
+the calcium ions bind more likely 
+to the carboxylate than to phosphate moiety of POPS headgroup (Table \ref{tab:binding}),
 in line with the results without the NBfix corrections in CHARMM36 \cite{hallock18}.
-However, in CHARMM36 simulations without the NBfix correction for calcium,
+However, in CHARMM36 simulations without the NBfix correction for calcium \cite{kim16},
 calcium ions significantly overbind to PC lipids  \cite{catte16} and calcium ions coordinate
@@
   \caption{ \label{fig:cacl_complexes} 
-  Left: Percentages of the population of bound Ca$^{2+}$ 
+  \textbf{Left:} Percentages of the population of bound Ca$^{2+}$ 
 in exclusive contact with various lipid moieties 
 in a POPC:POPS (5:1) bilayer with 409 mM CaCl$_2$. 
+The bar representing total populations of bound \ce{Ca^{2+}} to PC, PS, or PC and PS (blue),
+is in part covered by areas representing contributions of individual moieties (shades of orange and red). 
+In the case of mixed simultaneous binding to PC and PS
+we distinguish different moieties only in PS.
+%i.e., PC can interact with any considered moiety. 
 For comparison, in a pure POPC bilayer with 350 mM CaCl$_2$, 
-67\% of bound \ce{Ca^{2+}} is exclusively in contact with \ce{PO_4} group. \cite{melcr18} \\
-    Right: Relative probabilities of \ce{Ca^{2+}} ions to coordinate with a certain number of lipids
+67\% of bound \ce{Ca^{2+}} is exclusively in contact with \ce{PO_4} group. \cite{melcr18} 
+    \textbf{Right:} Relative probabilities of \ce{Ca^{2+}} ions to coordinate with a certain number of lipids
     in pure POPC bilayer with 350 mM CaCl$_2$ and in POPC:POPS (5:1) mixture with 400 mM CaCl$_2$.  
     Analysis was done for both systems by considering all lipids (blue and violet) and
-    for POPC:POPS (5:1) mixture also by considering POPC and POPS lipids separately (green and orange). \\
+    for POPC:POPS (5:1) mixture also by considering POPC and POPS lipids separately (green and orange). 
     %All lipids were taken into account with the exception of the complexes in light green and orange, 

### 2019-05-17 Manuscript/manuscript.tex
Added gamma label to POPS chemfig and removed associated todo
Moved one paragraph from counterion affinity to structural agreement (previous) section

@@
   }  
-\todo{Add label $\gamma$ to the figure -- it is used in Fig with the dihedral angle, 
-and it can be also used in the description of COO- vector. }
 \end{figure}
@@
 while the same dihedral parameters yield too small forking in ECC-POPS.
+
+Overall,
+the head group and acyl chain order parameters within ECC-lipids
+are in general in a good agreement with the experimental values 
+as shown in Fig. \ref{simVSexpNOions_POPS}. 
+The acyl chain order parameters in particular agree with the experimental data within error bars.
+The order parameters of the head groups are at an accuracy comparable to 
+other currently available classical models of lipids \citep{botan15, catte16, Pluhackova2016, nmrlipids_proj4}. 
+In particular, ECC-POPS is closer to the experimental order parameters 
+than Lipid17 with either Dang \cite{smith94,chang1999,dang2006} 
+or ff99 ions \citet{aqvist90}. 
 Although the ECC-POPS model reproduces the lipid bilayer dimensions and
@@
 
-\subsection{Counterion binding affinity and interactions between PC and PS head groups}
+\subsection{Binding of counterions to POPS and POPC and interactions between their head groups}
 
@@
 \end{figure} 
-Binding affinity of monovalent ions to lipid bilayers depends strongly on
-force field parameters \cite{catte16,NMRlipidsIV},
-which is the case also here as the sodium counterion binding affinity to
-POPS lipid bilayer is stronger in Lipid17/ Dang simulation than in
-ECC-POPS simulation (Fig. \ref{fig:POPS-counterions-dens}).
+In simulations, binding affinity of monovalent ions to lipid bilayers depends strongly on
+force field parameters \cite{catte16,NMRlipidsIV}.
+For instance, the affinity of sodium counterions to a POPS bilayer 
+is stronger in Lipid17/Dang simulations 
+than in simulations with ECC-POPS (Fig.~\ref{fig:POPS-counterions-dens}).
 The cation binding affinities to lipid bilayers in simulations have been previously evaluated
 against experiments using the order parameters of $\alpha$ and $\beta$ carbons in the
-phosphatidylcholine (PC) lipid headgroup (see Fig.~\ref{fig:chemstruct_pc_ps} for the labeling) \cite{catte16,melcr18,NMRlipidsIV},
+phosphatidylcholine (PC) lipid headgroup (see Fig.~\ref{fig:delta_ordPar_NaCl_PC-PS_mix_PC} for the labeling) \cite{catte16,melcr18,NMRlipidsIV},
 which decrease proportionally to the bound charge according to the ''electrometer concept'' \citep{seelig87}.
-Althought the response of PS lipid headgroup order parameters to the bound charge is also systematic,
+Although the response of PS lipid headgroup order parameters to the bound charge is also systematic,
 it is less well undestood and the addition of cations may cause phase transitions in negatively charged
@@
 Following our previous work \cite{NMRlipidsIV}, we evaluate the counterion binding affinity
-to lipid bilayer in simulations containing PS using two different approaches.
-
-% The situation is more complicated for lipid bilayers containing
-%PS lipids, for which there is no ion-free reference state due to the presence of counterions.
-
-%Unlinke in pure PC bilayers,
-%Interpreting the changes of the head group order parameters of 
-%mixed PC:PS lipid bilayers 
-%from both simulations and experiments
-%is complicated by the presence of counterions. \cite{NMRlipidsIV}
-
-The head group and acyl chain order parameters within ECC-lipids
-are in general in a good agreement with the experimental values 
-as shown in Fig. \ref{simVSexpNOions_POPS}. 
-The acyl chain order parameters in particular agree with the experimental data within error bars.
-The order parameters of the head groups are at an accuracy comparable to 
-other currently available classical models of lipids \citep{botan15, catte16, Pluhackova2016, nmrlipids_proj4}. 
-In particular, ECC-POPS is closer to the experimental order parameters 
-than Lipid17 with either Dang \cite{smith94,chang1999,dang2006} 
-or ff99 ions \citet{aqvist90}.
-
-%This was demonstrated for a bound positive charge in simulations with POPC and 
-%cationic surfactants or aquaeous cations employing the ECC-POPC model \cite{melcr18}.
-%In this work, we will use the electrometer concept also for a bound negative charge 
-%using PS lipids as negatively charged surfactants. \cite{roux90}
-
-%is the
-%is not generally agreed on between simulations 
-%\citep{bockmann03,sachs04,berkowitz06,cordomi09}
-%and experiments 
-%\citep{cevc90,tocanne90,hauser76,herbette84,uhrikova08}.
-%We attempted to find a model, 
-%which would successfully interpret the experiments on binding of \ce{Na^+}
-%to neutral and negatively charged phospholipid bilayers \citep{akutsu81, roux90} 
-%in our works \citep{catte16, NMRlipidsIV}.
-%After testing the currently available force fields for phospholipids,
-%we concluded
-%that the interactions are generally overestimated in magnitude in almost all models 
-%but Lipid14 (PC) \citep{dickson14}, resp. Lipid17 (PS) \citep{lipid17-future}. 
-%This model yields a semi-quantitative agreement with the experimentally measured small changes of the order parameters \cite{catte16}
-%when used with the model of ions by \citet{aqvist90}. 
-%However, when used with a more accurate model of ions by \citet{Pluharova2014, martinek17},
-%the model overestimates the binding affinity of \ce{Na^+}
-%measured with lipid electrometer concept. \citep{melcr18}
-%In total, these results suggest that improvements 
-%in the lipid parameters are required for more accurate interactions even with monovalent cations. 
-%\citep{catte16, melcr18, NMRlipidsIV}
+to a lipid bilayer containing PS in simulations using two approaches.
 

### 2019-05-17 Manuscript/manuscript.tex
Edits in the PC-PS interaction section and counterions

@@
 \end{figure}
+
 In simulations, binding affinity of monovalent ions to lipid bilayers depends strongly on
@@
     with respect to the membrane normal as a function of POPS content in a bilayer
-    from ECC-lipids and Lipid17/Dang simulations with \ce{Na^+} (top) and \ce{K^+} (bottom) counterions.
+    from ECC-lipids and Lipid17/Dang simulations with \ce{Na^+} (left) and \ce{K^+} (right) counterions.
     Experimental order parameter values are from Ref. \citenum{scherer87}
-    and the signs from Ref. \citenum{ferreira16} (only \ce{Na^+} counterions, but shown also in bottom plots).
+    and the signs from Ref. \citenum{ferreira16} 
+    (only \ce{Na^+} counterions, but shown also in the right plots for \ce{K+}).
   }
@@
     with respect to the membrane normal as a function of POPC content in a bilayer
-    from ECC-lipids and Lipid17/Dang simulations with \ce{Na^+} (top) and \ce{K^+} (bottom) counterions.
+    from ECC-lipids and Lipid17/Dang simulations with \ce{Na^+} (left) and \ce{K^+} (right) counterions.
     Experimental order parameter values and signs for pure POPS are from Ref. \citenum{NMRlipidsIV} and values
-    for POPC:POPS (5:1) mixture from Ref. \citenum{roux90} (only \ce{Na^+} counterions, but shown also in bottom plots).
+    for POPC:POPS (5:1) mixture from Ref. \citenum{roux90} 
+    (only \ce{Na^+} counterions, but shown also in the right plots for \ce{K+}).
     %The underestimated response of ECC-lipids with \ce{Na^+} counterions 

### 2019-05-17 Manuscript/manuscript.tex
Edits in results section PC-PS and counterions and additional nacl, kcl

@@
 force field parameters \cite{catte16,NMRlipidsIV}.
-For instance, the affinity of sodium counterions to a POPS bilayer 
+For instance in the present work the affinity of sodium counterions to a POPS bilayer 
 is stronger in Lipid17/Dang simulations 
@@
 compensates the negative charge of PS lipids at the interface.
+In contrast, ECC-lipids represent this trend correctly
+with only slightly underestimated slopes of the order parameters 
+when \ce{Na+} counterions are used (Fig.~\ref{fig:delta_ordPar_NaCl_PC-PS_mix}).
 
@@
 In experiments, the addition of K$^+$ ions led to a very modest decrease of the order parameters,
-while the decrease was more pronounced with Li$^+$ ions and strongest with divalent Ca$^{2+}$ and Mg$^{2+}$ ions \cite{roux90}.
+while the decrease was more pronounced with Li$^+$ ions and was the strongest with divalent Ca$^{2+}$ and Mg$^{2+}$ ions \cite{roux90}.
 According to the electrometer concept, the decrease of PC lipid head group order parameters
@@
 The response of PC headgroup order parameters to the increasing amount of KCl is close to experiments
-and the change in P--N vector is negligible in both force fields (Fig. \ref{fig:delta_ordPar_KCl_PCPS}), athought the change of the $\alpha$-carbon
-order parameter may be slightly overestimated by the Lipid17/Dang simulations. This could be
-explained by the slightly deeper penetration of potassium into the bilayer (Fig. \ref{fig:POPS-counterions-dens}).
+and the change in P--N vector is negligible in both force fields (Fig. \ref{fig:delta_ordPar_KCl_PCPS}), athough the change of the $\alpha$-carbon
+order parameter may be slightly overestimated by the Lipid17/Dang simulations. 
+This could be explained by the slightly deeper penetration of potassium into the bilayer (Fig. \ref{fig:POPS-counterions-dens}).
+
 Because experimental data with the additional amount of sodium in POPC:POPS (5:1) mixture was not available,
@@
 amount of sodium is clearly overestimated by the Lipid17/Dang simulation and
-the response in ECC-POPS simulations is probably too close to the experimental results for lithium.
-In conclusion, the results suggest that the sodium binding affinity (Fig. \ref{fig:POPS-counterions-dens}) and
-the tilting of P--N vector more perpendicular to the membrane surface (Fig. \ref{fig:delta_ordPar_NaCl_PCPS})
-are clearly overestimated in Lipid17/Dang simulations and slightly overestimated in ECC-POPS simulations,
-while the binding affinity of potassium is better described by both force fields.
-
-%Generally improved behaviour 
-%of the POPC and POPS head group order parameters 
-%with \ce{NaCl} or \ce{KCl} concentrations 
-%was achieved through the combination of models 
-%ECC-lipids and ECC-ions \citep{martinek17, kohagen16, Pluharova2014}. 
-%Simulations with these models 
-%reveal a good agreement with the NMR experiments 
-%for both neutral and negatively charged membranes.
-%The results are plotted for \ce{NaCl} in Fig. \ref{fig:delta_ordPar_NaCl_PCPS}, 
-%and for \ce{KCl} in Fig. \ref{fig:delta_ordPar_KCl_PCPS}.
-
-%The order parameters of the PS head groups including their changes with the content of PC lipids
-%are challenging to capture correctly in classical MD simulation models. \cite{NMRlipidsIV}
-%From such a perspective,
-%the changes of the PS head group order parameters from ECC-lipids 
-%with \ce{Na+}, and also \ce{K+}, counterions 
-%agree comparably well to the experimental measurements
-%(Fig.~\ref{fig:delta_ordPar_NaCl_PC-PS_mix}). 
-%Interestingly enough, 
-%although the Lipid17 model performed well with \ce{K+} counterions 
-%for the PC head group order parameter changes,
-%it does not agree in the equivalent changes in PS head groups
-%pointing at non-negligible effects of even relatively weakly binding cations 
-%on the structure of negatively charged lipids.
-
-%Despite the great improvements in the overall structure of ECC-POPS 
-%and its response to PC lipids and counterions,
-%there is still a large room for improving it
-%to agree with the experimental measurements.
+the response in ECC-lipids simulations is probably too close to the experimental results for lithium.
+% The following conclusion is already in the previous sentence, so I commented it out
+%In conclusion, the results suggest that the sodium binding affinity (Fig. \ref{fig:POPS-counterions-dens}) and
+%the tilting of P--N vector more perpendicular to the membrane surface (Fig. \ref{fig:delta_ordPar_NaCl_PCPS})
+%are clearly overestimated in Lipid17/Dang simulations and slightly overestimated in ECC-POPS simulations,
+%while the binding affinity of potassium is better described by both force fields.
 
@@
     as a function of \ce{NaCl} concentration from ECC-lipids and Lipid17/Dang simulations at 298 K.
-    The y-axis for the $\alpha$-carbon results of POPS (middle right) is transferred
-    with the same value for both order parameters such that the lower order
-    parameter value from pure POPS is at zero to correctly illustrate the significant forking.
+    % No translation in y-axis done in this plot:
+    %The y-axis for the $\alpha$-carbon results of POPS (middle right) is transferred
+    %with the same value for both order parameters such that the lower order
+    %parameter value from pure POPS is at zero to correctly illustrate the significant forking.
     Because experimental data with \ce{NaCl} is not available for POPC, the data for KCl and \ce{LiCl} (dashed line, left)
@@
     The magnitudes of experimental order parameters are from Ref. \citenum{roux90} and signs from Ref. \citenum{ferreira16}.
-    %which has a lower affinity to phospholipid bilayers compared to \ce{LiCl} \citep{roux90}. 
-    %The orientation of the \ce{COO^-} group is defined as 
-    %the connector from the $\beta$ carbon to the carbon in \ce{COO^-} (stars, bottom right). 
   }
@@
     compared with experimental values from Ref. \citenum{roux90} (signs from Ref. \citenum{ferreira16}) at 298 K.
-    The y-axis for the $\alpha$-carbon results of POPS (middle right) is transferred
-    with the same value for both order parameters such that the lower order
-    parameter value from pure POPS is at zero to correctly illustrate the significant forking.
+    % No translation in y-axis done in this plot:
+    %The y-axis for the $\alpha$-carbon results of POPS (middle right) is transferred
+    %with the same value for both order parameters such that the lower order
+    %parameter value from pure POPS is at zero to correctly illustrate the significant forking.
   }
@@
 
-% This§: 
-% Other models do not yield good reponse, but
-% ECC-lipid describe the changes well
-% => Elect. polar. necessarry.
-% 
 The changes of POPS headgroup order parameters are almost negligible in experiments with increasing POPC fraction in bilayer
@@
 in ECC-POPS simulations (Figs. \ref{fig:delta_ordPar_NaCl_PCPS} and \ref{fig:delta_ordPar_KCl_PCPS}).
+These structural changes can relate to variations in the interactions between PC and PS head groups 
+and also to the changing amount of counterions and PS in the composition. 
 %In contrast, ECC-lipids with ECC-ions capture the different response 
@@
 %without the necessity of defining a Gibbs dividing surface. \citep{melcr18, chattorajBOOK}
-The value of $\Gamma ^{w} _{\rm Na}=-0.11 \pm 0.01 \mathrm{nm}^{-2}$ for 1 M sodium in pure ECC-POPC
+The value of $\Gamma ^{w} _{\rm Na}=-0.11 \pm 0.01 \mathrm{nm}^{-2}$ for 1~M sodium in pure ECC-POPC
 simulation was reported in the previous work \cite{melcr18}.

### 2019-05-17 Manuscript/manuscript.tex
Edited calcium binding section and cleaned up many commented lines

@@
     (only \ce{Na^+} counterions, but shown also in the right plots for \ce{K+}).
+    % No translation in y-axis done in this plot:
     %The underestimated response of ECC-lipids with \ce{Na^+} counterions 
@@
     as a function of \ce{NaCl} concentration from ECC-lipids and Lipid17/Dang simulations at 298 K.
-    % No translation in y-axis done in this plot:
-    %The y-axis for the $\alpha$-carbon results of POPS (middle right) is transferred
-    %with the same value for both order parameters such that the lower order
-    %parameter value from pure POPS is at zero to correctly illustrate the significant forking.
+    The y-axis for the $\alpha$-carbon results of POPS (middle right) is transferred
+    with the same value for both order parameters such that the lower order
+    parameter value from pure POPS is at zero to correctly illustrate the significant forking.
     Because experimental data with \ce{NaCl} is not available for POPC, the data for KCl and \ce{LiCl} (dashed line, left)
@@
     compared with experimental values from Ref. \citenum{roux90} (signs from Ref. \citenum{ferreira16}) at 298 K.
-    % No translation in y-axis done in this plot:
-    %The y-axis for the $\alpha$-carbon results of POPS (middle right) is transferred
-    %with the same value for both order parameters such that the lower order
-    %parameter value from pure POPS is at zero to correctly illustrate the significant forking.
+    The y-axis for the $\alpha$-carbon results of POPS (middle right) is transferred
+    with the same value for both order parameters such that the lower order
+    parameter value from pure POPS is at zero to correctly illustrate the significant forking.
   }
@@
 
-\subsection{Molecular interaction and binding affinities of \ce{Ca^{2+}} cations to mixed POPC:POPS (5:1) membrane} 
+\subsection{Molecular interaction and binding affinity of \ce{Ca^{2+}} cations to mixed POPC:POPS (5:1) membrane} 
 \label{section:lip-ion_ca}
@@
 Experimentally measured headgroup order parameters of POPS exhibit a strong dependence
-on low CaCl$_2$ concentrations with a rapid saturation below 100 mM \cite{roux90}.
+on low CaCl$_2$ concentrations with a rapid saturation below 100~mM \cite{roux90}.
 Also these changes were overestimated by the Lipid17/Dang (also shown in Fig.~\ref{fig:delta_ordPar_CaCl_PCPS})
-and other tested force fields in our recent work \cite{NMRlipidsIV}, including CHARMM36 with the NBfix correction for calcium
+and other tested force fields in our recent work \cite{NMRlipidsIV}, 
+but CHARMM36 with the NBfix correction for calcium
 which underestimated the binding affinity.
@@
 but the strong dependence on low concetrations of CaCl$_2$ is not fully reproduced (Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}).
-In addition to the inaccurate interactions between calcium ions and PS headgroup, the potential sources of this
+In addition to possibly suboptimal interactions between calcium ions and PS headgroup, the potential sources of this
 discrepancy include the above observed slight overestimation of \ce{Na^+} counterions and
@@
 tilts to the opposite direction by $11^\circ$ in the same system.
-%the systematic tilting of P--N and C$_{\beta}$--C$_{\gamma}$ vectors 
-%to more perpendicular and parallel direction with respect to the membrane interface, respectively,
-%is observed upon addition of CaCl$_2$ (Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}).
+Such a largely diminished flexibility of the PS head group can be explained 
+by the simultaneous binding of calcium to the \ce{COO^-} group in PS and phosphate group in PC. 
 In Lipid17/Dang simulations, the changes in both headroup orientations and order parameters
 are larger and less systematic for the PS headgroup, especially with lower concentrations (Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}).
-Together with the smaller error bars and shorter residence times (Fig.~\ref{fig:hist_residence_times}) in
-ECC-lipids simulations than in typically observed for other force fields \cite{javanainen17,melcr18,NMRlipidsIV},
-this suggests that the ECC accelerates the equlibration of ions at lipid bilayer interface.
-%
-%changes by $11^\circ$ from $62^\circ$ to $73^\circ$ (more towards the membrane interior)
-% at 620~mM buffer concentration of \ce{CaCl2}.
-% We have investigated this effect using the mean orientations of P-N and \ce{COO^-} vectors 
-%(defined and plotted in Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}). 
-%While the mean angle between the P-N vector of PC head group and the membrane normal
-%decreased by 11$^\circ$ from 64$^\circ$ to 55$^\circ$
-% at 620~mM buffer concentration of \ce{CaCl2},
-%the mean orientation of the P-N vector in the PS head group has 
-%decreased only by 3$^\circ$ from 67$^\circ$ to 64$^\circ$ 
-%in the same concentration range. 
-%Such a largely diminished flexibility 
-%may arise from the binding of calcium to the \ce{COO^-} group in PS. 
-%Whe have measured the mean orientation of the \ce{COO^-} group 
-%as the connector of the carbon atoms 
-%forming the bond between the group and the $\beta$-carbon of the phospholipid 
-%(Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}).
-%by the binding of calcium cations to the \ce{COO^-} group, 
-%which reorients towards the membrane interior and to the phosphate region, 
-%where calcium cations bind predominantly. 
-%
-%
+ECC-lipids simulations yield smaller error bars and shorter residence times (Fig.~\ref{fig:hist_residence_times}) in
+than those typically observed in simulations with other force fields \cite{javanainen17,melcr18,NMRlipidsIV}.
+This suggests that the ECC accelerates the equlibration of ions at lipid bilayer interface.
+
 Our 1 $\mu$s simulations seems to be sufficiently long for the ECC-lipids simulations, because
@@
 Therefore, calcium ions bind more likely to negatively charged PS lipids than neutral PC lipid in a mixed bilayer, as expected.
-
-%are again closer to the experiments is not optimal,
-%the general trends and behaviour even at highly concentrated solutions of \ce{CaCl2} 
-%are well captured by the model. 
-%For example the order parameter $\beta$ in POPS 
-%adopts a saturated value in experiments and ECC-lipids,
-%but it is growing with the \ce{CaCl2} concentration in Lipid17.
-%In addition, 
-%the forking of the order parameters $\alpha$ 
-%is almost constant in the experiment and ECC-lipids,
-%but it varies greatly in th Lipid17.
-
-%The structural changes in POPS induced by increasing concentrations of \ce{CaCl2} 
-%are very challenging to interpret from both experiments \cite{roux90} 
-%and also from MD simulations \cite{NMRlipidsIV}. 
-%The changes of the order parameters $S^\alpha$ and $S^\beta$ in PS head groups from simulations 
-%are not found to follow the experimental profiles, 
-%nor any common trend, with the currently available models \cite{NMRlipidsIV}.
 
@@
 the calcium ions bind more likely 
-to the carboxylate than to phosphate moiety of POPS headgroup (Table \ref{tab:binding}),
+to the carboxylate than to phosphate moiety of POPS headgroup (Fig.~\ref{fig:cacl_complexes}),
 in line with the results without the NBfix corrections in CHARMM36 \cite{hallock18}.
@@
 Furthermore, the POPS:POPC:Ca$^{2+}$ (7:3:7) mixture with relatively high
-PS lipid fraction and calcium concentration was used in previous work \cite{hallock18},
+PS lipid fraction and calcium concentration was used in the previous work \cite{hallock18},
 which may lead to the complexation of \ce{Ca^{2+}} cations and PS lipids that is
@@
 %however, they are always accompanied by interactions with phosphate groups.
+
 Significant coordination of calcium with carbonyls groups of PS and PC lipids observed in previous
@@
 was not observed in ECC-lipids simulations in here nor in the previous work with pure POPC \cite{melcr18}.
-%This is in contrast to 
-%combined with fluorescent and vibrational sum frequency spectroscopy,
-%which suggest a 
 Overestimated coordination of cations with carbonyl groups is a potential source for
@@
 The bar representing total populations of bound \ce{Ca^{2+}} to PC, PS, or PC and PS (blue),
-is in part covered by areas representing contributions of individual moieties (shades of orange and red). 
+is in part covered by areas representing contributions of individual moieties exclusively (shades of orange and red). 
 In the case of mixed simultaneous binding to PC and PS
@@
 
-%Still, the dominant contribution to the binding of calcium comes from the phosphate groups. 
-%Interestingly enough,
-%although the \ce{COO^-} groups increase the affinity of calcium towards PS compared to PC,
-% there is no extra density at the outmost surface of the membrane. 
-%This can be understood from the point of the structural rearangements of the \ce{COO^-} group, 
-%which take place during the interactions with calcium cations
-%described above and in Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}.
-
-%Simulations using CHARMM36 force field \cite{klauda10,venable13} with
-%the NBfix parameters for calcium \cite{kim16}, combined with 2D infrared spectroscopy,
-%suggest that calcium ions interact only with the carboxylate group of PS lipids \cite{valentine18}.
-%The same force field without the NBfix parameters, combined with NMR chemical shifts and
-%rotational-echo double-resonance (REDOR) experiments, suggests a significant binding affinity also to the phosphate region \cite{hallock18}.
-%Simulations with the Berger force field \cite{berger97,mukhopadhyay04},
-%combined with fluorescent and vibrational sum frequency spectroscopy, suggest a significant
-%calcium binding also to the carbonyls in the acyl chains \cite{melcrova16}.
-
-%In PS lipids, the calcium binds both phosphate and carboxylate, but probability for the latter is twice higher.
-%In PC lipids, the calcium binds almost always to phosphate, but in some cases also to carbonyl, as also observed previously in pure POPC simulations \cite{melcr18}.
-
 The advantage of the NMR order parameters used in this work is the direct connection between
@@
 correctly capture the rapid changes with small concentrations. Therefore, further
-force field optimization is required for fully realistic MD simulations of cations
+force field optimization is required for even more accurate MD simulations of cations
 in the vicinity of negatively charged lipid bilayers.
-
-%Despite significant improvements in the ECC-lipid against Lipid17/Dang
-%and other force fields for the response of PS headgroup order parameters to
-%the bound calcium \cite{NMRlipidsIV}, 
-%the model (Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}).
-
 % and also possibly from the approximate mean-field treatment of the electronic polarization.
-
-%
-% SAMULI: We have problems to find time to finish this paper even with the current data, we should not add anything that is not necessary to finish this asap. 
-%
-%\todo{JOE: suggestion -- what is the charge at the membrane interface of memb+counterions -vs- memb+CaCl2 concnetration?
-% this could relate well to the zeta potential measurements and the Fig 1 in \cite{melcrova16}. }
-% 
-%\todo{ THE REMAINING PIECES TO BE DISCUSSED  --  consider wheter the dynamics slowdown or the dihedral distributions could make good contributions to this paper. }
-
-%% Ca2+ induces much slower dynamics of lipids when PS is present,
-%% this does not happen with PC
-%% uses NBFIX
-%From Ref.~\cite{valentine18}://
-%To summarize, calcium binding to PS lipids slows the dynamics at the ester position and introduces more heterogeneous molecular environments. 
-%The slowdown cannot be attributed to complete ester dehydration or decreased hydrogen bonding to water, so we interpret these results as evidence that the lipid headgroups adopt more rigid but %disordered conformations on the molecular level.
-%The decrease in dynamics observed experimentally indicates that the network of hydrogen bonds between lipids and water at the polar-nonpolar interface becomes slower, as it was shown by Karatha%nou and Bondar (34).
-%This is bolstered by recent NMR work showing that binding to Ca2+ induces two different rigid conformations in the glycerol backbone of lipid headgroups (18).
-%We have investigated the molecular origin of the rigidification and decreased lateral diffusion, finding that calcium binding does not dehydrate the ester region, but even with only 10\% PS lip%ids, the local dynamics slow around the esters.
-The slower local dynamics provide a molecular explanation for the Ca2+-dependent decreases in fluidity observed on longer length and timescales.
-%There is considerable room for further improvement in the force fields for ion-membrane interactions.
-Particularly, small differences in short-range interactions may determine the local geometr%ies (34
-%).
-Most of our conclusions are, however, based on long-range effects (i.e., electric field generated by the ion at significant distances from the carbonyl groups), which are less sensitive to t%he details of the force field.
-
-%Their Figure 5 \cite{hallock18} shows the distributions of O-Ca-Cb-N dihedral angle
-%revealing separate populations of the PS head groups. 
-%These two structures are also compared in therm of P-Cg distances, 
-%which can be measured from NMR REDOR exp. 
-%It is concluded that since the two distinct populations in P-Cg distance can be represented with the 
-%dihedral distributions, 
-%they also likely represent the distinct populations. 
-% -> plotting the distribution of O-Ca-Cb-N dihedral angle may be also a good idea for us (SI?)
-%     -> NMRlipids 4 in SI \cite{NMRlipidsIV} plots this information: 
-%           - C36 agress with the distribution in \cite{hallock18}
-%           - Lipid17 (and Slipids) are almost identical to C36
-%           - but the structure of the COO- is highly different (dih angle Ca-Cb-Cg-Og)
 

### 2019-05-20 Manuscript/manuscript.tex
Improved title (and gave an alternative title in comment)

@@
 \author{Josef Melcr} 
-\email{melcr@uochb.cas.cz} 
+\email{melcr@marge.uochb.cas.cz} 
 %%\homepage[]{https://jmelcr.github.io/}
@@
 \title[] 
-  { Detailed structure of phospholipid bilayers with negatively charged lipids and salt concentrations 
-    from molecular dynamics simulations including the effects of electronic polarization }
- %{ Non-negligible Effects of Electronic Polarization at Negatively Charged Membranes with Ions }
- %{ Effective Inclusion of Electronic Polarization to a Negatively Charged Membrane }
+  { Detailed structure of negatively charged membranes 
+    of phosphatidylserine and phosphatidylcholine 
+    at concentrations of calcium, sodium and potassium salts
+    from molecular dynamics simulations with electronic polarization }
+  % JOSEF: I feel that the above title is a bit better, because it focuses on PS,
+  %        but the following one is an alternative, which focuses more on the cations as in roux90
+  %{ Detailed structure of calcium, sodium and potassium cations
+  %  at negatively charged membranes of phosphatidylserine and phosphatidylcholine 
+  %  from molecular dynamics simulations with electronic polarization }
 

### 2019-05-20 Manuscript/manuscript.tex
Improved table 1 - list of simulations

@@
            and the deployed models.
-           The fatty acid chains for all lipids are plamitoyl (sn-1) and oleoyl (sn-2).
+           The fatty acid chains for all lipids are palmitoyl (sn-1) and oleoyl (sn-2).
          }\label{tbl:sim-list}
-\begin{tabular}{l l | r r r r | r r | c c }
-   \multicolumn{2}{c}{ }  &  \multicolumn{4}{c}{added buffer conc. / mM}    & \multicolumn{2}{c}{no. molecules} &  \multicolumn{2}{c}{simulated with model}  \\
-PC:PS ratio & counterions &  \ce{K^+}  &  \ce{Na^+} & \ce{Ca^{2+}} & \ce{Cl^-}      & \ce{H2O} &  PC:PS                 &  ECC-lipids  &  Lipid17    \\
+\begin{tabular}{r  l | r r r r | r r | c c }
+   \multicolumn{1}{r}{PC:PS} & \multicolumn{1}{l}{ }  &  \multicolumn{4}{c}{added buffer conc. / mM}    & \multicolumn{2}{c}{no. molecules} &  \multicolumn{2}{c}{simulated with}  \\
+   ratio & counterion &  \ce{K^+}  &  \ce{Na^+} & \ce{Ca^{2+}} & \ce{Cl^-}      & \ce{H2O} &  PC:PS                 &  ECC-lipids  &  Lipid17    \\
   \hline
-0:$\infty$ & \ce{K^+}   &      0  &      0  &      0  &      0  &  3600  &  0:72  &  X  &  X  \\ 
-0:$\infty$ & \ce{Na^+}  &      0  &      0  &      0  &      0  &  3600  &  0:72  &  X  &  X  \\ 
+only PS & \ce{K^+}   &      0  &      0  &      0  &      0  &  3600  &  0:72  &  \textbullet &  \textbullet \\ 
+only PS & \ce{Na^+}  &      0  &      0  &      0  &      0  &  3600  &  0:72  &  \textbullet &  \textbullet \\ 
   \hline
-1:1 & \ce{K^+}   &      0  &      0  &      0  &      0  &  5253  &  64:64  &  X  &  X  \\ 
-1:1 & \ce{Na^+}  &      0  &      0  &      0  &      0  &  5253  &  64:64  &  X  &  X  \\ 
+1:1 & \ce{K^+}   &      0  &      0  &      0  &      0  &  5253  &  64:64  &  \textbullet &  \textbullet \\ 
+1:1 & \ce{Na^+}  &      0  &      0  &      0  &      0  &  5253  &  64:64  &  \textbullet &  \textbullet \\ 
   \hline
-4:1 & \ce{K^+}   &      0  &      0  &      0  &      0  &  3600  &  48:12  &  X  &  -  \\ 
-4:1 & \ce{Na^+}  &      0  &      0  &      0  &      0  &  3600  &  48:12  &  X  &  -  \\ 
+4:1 & \ce{K^+}   &      0  &      0  &      0  &      0  &  3600  &  48:12  &  \textbullet &  -  \\ 
+4:1 & \ce{Na^+}  &      0  &      0  &      0  &      0  &  3600  &  48:12  &  \textbullet &  -  \\ 
   \hline
-5:1 & \ce{K^+}   &      0  &      0  &      0  &      0  &  3600  &  60:12  &  X  &  X  \\ 
-5:1 & \ce{Na^+}  &      0  &      0  &      0  &      0  &  3600  &  60:12  &  X  &  X  \\ 
-5:1 & \ce{Na^+}  &      0  &      0  &     78  &     78  &  3561  &  60:12  &  X  &  X  \\ 
-5:1 & \ce{Na^+}  &      0  &      0  &    125  &    125  &  3561  &  60:12  &  X  &  X  \\ 
-5:1 & \ce{Na^+}  &      0  &      0  &    202  &    202  &  3561  &  60:12  &  X  &  X  \\ 
-5:1 & \ce{Na^+}  &      0  &      0  &    409  &    409  &  3522  &  60:12  &  X  &  X  \\ 
-5:1 & \ce{Na^+}  &      0  &      0  &    621  &    621  &  3483  &  60:12  &  X  &  X  \\ 
-5:1 & \ce{Na^+}  &      0  &    621  &      0  &    621  &  3483  &  60:12  &  X  &  X  \\ 
-5:1 & \ce{Na^+}  &      0  &   1510  &      0  &   1510  &  3377  &  60:12  &  X  &  X  \\ 
-5:1 & \ce{Na^+}  &      0  &   3002  &      0  &   3002  &  3213  &  60:12  &  X  &  X  \\ 
-5:1 & \ce{Na^+}  &    621  &      0  &      0  &    621  &  3483  &  60:12  &  X  &  X  \\ 
-5:1 & \ce{Na^+}  &   1510  &      0  &      0  &   1510  &  3377  &  60:12  &  X  &  X  \\ 
-5:1 & \ce{Na^+}  &   3002  &      0  &      0  &   3002  &  3213  &  60:12  &  X  &  X  \\ 
+5:1 & \ce{K^+}   &      0  &      0  &      0  &      0  &  3600  &  60:12  &  \textbullet &  \textbullet \\ 
+5:1 & \ce{Na^+}  &      0  &      0  &      0  &      0  &  3600  &  60:12  &  \textbullet &  \textbullet \\ 
+5:1 & \ce{Na^+}  &      0  &      0  &     78  &     78  &  3561  &  60:12  &  \textbullet &  \textbullet \\ 
+5:1 & \ce{Na^+}  &      0  &      0  &    125  &    125  &  3561  &  60:12  &  \textbullet &  \textbullet \\ 
+5:1 & \ce{Na^+}  &      0  &      0  &    202  &    202  &  3561  &  60:12  &  \textbullet &  \textbullet \\ 
+5:1 & \ce{Na^+}  &      0  &      0  &    409  &    409  &  3522  &  60:12  &  \textbullet &  \textbullet \\ 
+5:1 & \ce{Na^+}  &      0  &      0  &    621  &    621  &  3483  &  60:12  &  \textbullet &  \textbullet \\ 
+5:1 & \ce{Na^+}  &      0  &    621  &      0  &    621  &  3483  &  60:12  &  \textbullet &  \textbullet \\ 
+5:1 & \ce{Na^+}  &      0  &   1510  &      0  &   1510  &  3377  &  60:12  &  \textbullet &  \textbullet \\ 
+5:1 & \ce{Na^+}  &      0  &   3002  &      0  &   3002  &  3213  &  60:12  &  \textbullet &  \textbullet \\ 
+5:1 & \ce{Na^+}  &    621  &      0  &      0  &    621  &  3483  &  60:12  &  \textbullet &  \textbullet \\ 
+5:1 & \ce{Na^+}  &   1510  &      0  &      0  &   1510  &  3377  &  60:12  &  \textbullet &  \textbullet \\ 
+5:1 & \ce{Na^+}  &   3002  &      0  &      0  &   3002  &  3213  &  60:12  &  \textbullet &  \textbullet \\ 
   \hline
-10:1 & \ce{K^+}  &      0  &      0  &      0  &      0  &  3600  &  60:6  &  X  &  -  \\ 
-10:1 & \ce{Na^+} &      0  &      0  &      0  &      0  &  3600  &  60:6  &  X  &  -  \\ 
+10:1 & \ce{K^+}  &      0  &      0  &      0  &      0  &  3600  &  60:6  &  \textbullet &  -  \\ 
+10:1 & \ce{Na^+} &      0  &      0  &      0  &      0  &  3600  &  60:6  &  \textbullet &  -  \\ 
 \end{tabular}

### 2019-05-20 Manuscript/manuscript.tex
Edits in reults-counterions moved dihedrtal distribution to SI

@@
 The X-ray scattering form factors and C-H bond order parameters in
-the headgroup and glycerol backbone region of POPS lipid bilayer are already available in the literature \cite{kucerka14,NMRlipidsIV}.
+the headgroup and glycerol backbone region of a POPS lipid bilayer are already available in the literature \cite{kucerka14,NMRlipidsIV}.
 Here, we measure also the C--H bond order parameters of the acyl chain region using natural abundance $^{13}$C NMR.
@@
 
-The Lipid17/Dang simulation give larger acyl chain order parameters and smaller area per molecule than the
+The Lipid17/Dang simulation gives larger acyl chain order parameters and smaller area per molecule than the
 experimental values (Fig. \ref{simVSexpNOions_POPS} and table \ref{tab:apls}),
-indicating that this simulation predicts too dense POPS lipid bilayer. This is also in line with
-the discrepancy observed in the X-ray form factor between this simulation and experiment (Fig. \ref{simVSexpNOions_POPS}).
+indicating that this simulation predicts a too compact POPS lipid bilayer. This is also in line with
+the discrepancy observed in the X-ray form factor between the simulation and experiment (Fig.~\ref{simVSexpNOions_POPS}).
 In previous study \cite{NMRlipidsIV}, the Lipid17 force field with {\AA}qvist ions~\cite{aqvist90}
 gave a larger area per molecule (57~\AA$^2$) for POPS being closer to the experimental value,
-which is probably due to the weaker counterion binding affinity of {\AA}qvist ions.
-However,  {\AA}qvist ions exhibit artificial aggregation of ions at larger salt concentrations \cite{kohagen16, chen07, NMRlipidsIV}.
-Therefore, the Dang ions with more realisitic bulk properties are used here to analyze the effect of ECC to the properties of POPS lipid bilayer. 
+which is probably due to the weaker counterion binding affinity of {\AA}qvist ions to phospholipids. \cite{catte16, NMRlipidsIV}
+However,  {\AA}qvist ions produce known artifacts, 
+e.g. they exhibit artificial aggregation of ions at larger salt concentrations \cite{kohagen16, chen07, NMRlipidsIV}.
+Therefore, the Dang ions with more realisitic bulk properties are used here to analyze the effect of ECC to the properties of a POPS lipid bilayer. 
 The ECC-POPS simulation gives a good agreement with the X-ray scattering form factors and experimental
-acyl chain order parameters (see Fig.~\ref{simVSexpNOions_POPS}), indicating that the bilayer dimensions and acyl chain
+acyl chain order parameters (see Fig.~\ref{simVSexpNOions_POPS}) indicating that the bilayer dimensions and acyl chain
 conformations are well described by the force field. The area per lipid from ECC-POPS simulation is slightly
-smaller than the value reported from SDP model \cite{kucerka14} (table \ref{tab:apls}), but the values
+smaller than the value reported from SDP model \cite{kucerka14} (table~\ref{tab:apls}), but the values
 agree within their error estimates.
@@
   \end{tabular} \\
+\todo{J: This table will become a small inset plot in the top of Fig.~\ref{simVSexpNOions_POPS} }
 \end{table}
@@
 order parameters are quite far from experimental values with both Dang (Fig. \ref{simVSexpNOions_POPS})
-and {\AA}qvist \cite{NMRlipidsIV} ions. The values from ECC-POPS simulation are closer,
+and {\AA}qvist ions. \cite{NMRlipidsIV} The values from ECC-POPS simulation are closer,
 but not in full agreement with experiments (Fig.~\ref{simVSexpNOions_POPS}).
@@
 These differences are previously related
-to a more rigid structure of PS headgroup and 
-the stiffness of the C$_\alpha$-C$_\beta$-C$_\gamma$-O$_\gamma$ dihedral
+to a more rigid structure of PS headgroup \cite{browning80,buldt81} and 
+the stiffness of the C$_\alpha$-C$_\beta$-C$_\gamma$-O$_\gamma$ dihedral \cite{NMRlipidsIV}
 plotted in Fig.~\ref{fig:dihedral}. 
@@
 the head group and acyl chain order parameters within ECC-lipids
-are in general in a good agreement with the experimental values 
+are in a good agreement with the experimental values 
 as shown in Fig. \ref{simVSexpNOions_POPS}. 
@@
 than Lipid17 with either Dang \cite{smith94,chang1999,dang2006} 
-or ff99 ions \citet{aqvist90}. 
+or {\AA}qvist ions \cite{aqvist90}. 
 Although the ECC-POPS model reproduces the lipid bilayer dimensions and
@@
 e.g. in the parameters of the associated dihedral angles. 
-\cite{botan15, ollila16, Pluhackova2016, NMRlipidsIV}.
-
-\begin{figure}[tbp!] 
-  \centering 
-  \includegraphics[width=\figwidth]{../img/dihedral_angle_distribution_Ca-Cb-Cg-Og_ECC-L17.pdf}
-  \caption{\label{fig:dihedral}
-    Distribution of dihedral angle \ce{C_\alpha-C_\beta-C_\gamma-O_\gamma} in POPS 
-    from simulations of a pure POPS bilayer with \ce{Na+} counterions using Lipid17 and ECC-lipids. 
-    Labelling of POPS segments is in Fig.~\ref{simVSexpNOions_POPS}. 
-}
-\end{figure}
+\cite{botan15, ollila16, Pluhackova2016, NMRlipidsIV}
 

### 2019-05-20 Manuscript/manuscript.tex
Rearranged and united figures in the section with monovalent ions. Final retouche of the Figs yet to be done manually.

@@
 \newlength{\figheightsmall}
-\setlength{\figheightsmall}{8 cm} 
+\setlength{\figheightsmall}{6 cm} 
 \newlength{\figheight}
@@
 
-\begin{figure}[tbp!] 
-  \centering 
-  \includegraphics[height=\figheight]{../img/ecc_pops/density_profiles_na-k-counterions_wat_phos_compar_purePOPS_ecclipids-lipid17.pdf}
-  \includegraphics[height=\figheight]{../img/ecc_pops/density_profiles_na-k-counterions_wat_phos_compar_5PC-1PS_ecclipids-lipid17.pdf}
-  \caption{\label{fig:POPS-counterions-dens}
-    Number density profiles of \ce{K^{+}} and \ce{Na^{+}} counterions along the membrane normal axis
-    in ECC-lipids (solid lines) and Lipid17/Dang (dashed lines) simulations of pure POPS (left) or POPC:POPS (5:1, right) bilayers.  
-    The density profiles of phosphate groups and water are divided by 4 and 100, respectively.  
-}
-\end{figure}
-
 In simulations, binding affinity of monovalent ions to lipid bilayers depends strongly on
 force field parameters \cite{catte16,NMRlipidsIV}.
-For instance in the present work the affinity of sodium counterions to a POPS bilayer 
-is stronger in Lipid17/Dang simulations 
-than in simulations with ECC-POPS (Fig.~\ref{fig:POPS-counterions-dens}).
+% J: I think the following sentence can be omitted
+%For instance in the present work the affinity of sodium counterions to a POPS bilayer 
+%is stronger in Lipid17/Dang simulations 
+%than in simulations with ECC-POPS (Fig.~\ref{fig:POPS-counterions-dens} and~\ref{fig:delta_ordPar_NaCl_PC-PS_mix_PC}).
 The cation binding affinities to lipid bilayers in simulations have been previously evaluated
@@
   \centering 
+  \includegraphics[height=\figheightsmall]{../img/ecc_pops/density_profiles_na-k-counterions_wat_phos_compar_5PC-1PS_ecclipids-lipid17.pdf}
   \includegraphics[height=\figheightsmall]{../Fig/order_parameters_changes_A-B_PC-PS_mix_POPC_nacl.pdf} 
@@
   \caption{\label{fig:delta_ordPar_NaCl_PC-PS_mix_PC} 
-    The POPC head group order parameters and the P--N vector angle
+    \textbf{Left:} Number density profiles of \ce{K^{+}} and \ce{Na^{+}} counterions along the membrane normal axis
+    in ECC-lipids (solid lines) and Lipid17/Dang (dashed lines) simulations of POPC:POPS (5:1) bilayers.  
+    The density profiles of phosphate groups and water are divided by 4 and 100, respectively.  
+    \textbf{Middle:} The POPC head group order parameters and the P--N vector angle
     with respect to the membrane normal as a function of POPS content in a bilayer
@@
     (only \ce{Na^+} counterions, but shown also in the right plots for \ce{K+}).
+    \textbf{Right: Chemical structure and labelling of carbon segments of POPC. }
   }
+\todo{J: Will manually improve this figure with panels A, B, C and D 
+and will erase the duplicate y-axis description 
+and will add ticks to the right side of the plotting area. }
 \end{figure*}
+
+First, we monitor the response of POPC headgroup order parameters to the increasing amount of POPS (Fig.~\ref{fig:delta_ordPar_NaCl_PC-PS_mix}).
+According to the electrometer concept, the head group order parameters of PC lipids increase
+with the increasing amount of negatively charged PS lipids in experiments, because the headgroup P--N dipole
+tilts more parallel to the membrane interface \cite{seelig87,scherer87}.
+The trend is observed with both force fields in simulations with \ce{K+} counterions and
+in ECC-POPS simulation with \ce{Na+} counterions, but not in Lipid17/Dang ion simulations with \ce{Na+} counterions,
+where the order parameters are unchanged or decrease and the P--N vector tilts more perpendicular to the membrane interface.
+The result can be explained by the higher binding affinity of sodium in the
+Lipid17/Dang ion simulations (Fig~\ref{fig:delta_ordPar_NaCl_PC-PS_mix_PC}), which
+compensates the negative charge of PS lipids at the interface.
+In contrast, ECC-lipids represent this trend correctly
+with only slightly underestimated slopes of the order parameters 
+when \ce{Na+} counterions are used (Fig.~\ref{fig:delta_ordPar_NaCl_PC-PS_mix}).
+
+Second, we monitor the response of PC headgroup order parameters to the additional
+amount of sodium and potassium in POPC:POPS (5:1) mixtures.
+In experiments, the addition of K$^+$ ions led to a very modest decrease of the order parameters,
+while the decrease was more pronounced with Li$^+$ ions and was the strongest with divalent Ca$^{2+}$ and Mg$^{2+}$ ions \cite{roux90}.
+According to the electrometer concept, the decrease of PC lipid head group order parameters
+correlates with the amount of bound cations \cite{akutsu81,altenbach84,seelig87,roux90,catte16,NMRlipidsIV},
+thus the binding affinity of the ions to POPC:POPS (5:1) mixtures increase in order K$^{+}$ $<$ Li$^{+}$  $<$ Mg$^{2+}$  $<$ Ca$^{2+}$.
+The response of PC headgroup order parameters to the increasing amount of KCl is close to experiments
+and the change in P--N vector is negligible in both force fields (Fig. \ref{fig:delta_ordPar_monoval_PCPS}), athough the change of the $\alpha$-carbon
+order parameter may be slightly overestimated by the Lipid17/Dang simulations. 
+This could be explained by the slightly deeper penetration of potassium into the bilayer (Fig. \ref{fig:delta_ordPar_NaCl_PC-PS_mix_PC}).
+
+Because experimental data with the additional amount of sodium in POPC:POPS (5:1) mixture was not available,
+we compared the simulation results to the experimental data with the additional amounts of
+lithium and potassium ions, which should give the lower and upper bounds to the sodium binding
+affinity (Fig. \ref{fig:delta_ordPar_monoval_PCPS}). The PC headgroup response to the additional
+amount of sodium is clearly overestimated by the Lipid17/Dang simulation and
+the response in ECC-lipids simulations is probably too close to the experimental results for lithium.
+% The following conclusion is already in the previous sentence, so I commented it out
+%In conclusion, the results suggest that the sodium binding affinity (Fig. \ref{fig:POPS-counterions-dens}) and
+%the tilting of P--N vector more perpendicular to the membrane surface (Fig. \ref{fig:delta_ordPar_NaCl_PCPS})
+%are clearly overestimated in Lipid17/Dang simulations and slightly overestimated in ECC-POPS simulations,
+%while the binding affinity of potassium is better described by both force fields.
+
+%\subsection{Molecular interaction and binding affinities of K$^+$ and Na$^+$ cations to mixed POPC:POPS (5:1) membrane} 
+%\label{sec:K_Na_added}
+
+\begin{figure}[tbp!] 
+  \centering 
+  \includegraphics[height=\figheightsmall]{../img/ecc_pops/order_parameters_changes_ecc-lip_L14_A-B-PN-COO_POPC_kcl.pdf} 
+  \includegraphics[height=\figheightsmall]{../img/ecc_pops/order_parameters_changes_ecc-lip_L14_A-B-PN-COO_POPC_nacl.pdf} 
+  \includegraphics[height=\figheightsmall]{../img/ecc_pops/order_parameters_changes_ecc-lip_L14_A-B-PN-COO_POPS_kcl.pdf} 
+  \includegraphics[height=\figheightsmall]{../img/ecc_pops/order_parameters_changes_ecc-lip_L14_A-B-PN-COO_POPS_nacl.pdf} 
+  \caption{\label{fig:delta_ordPar_monoval_PCPS}
+    Changes of the head group order parameters, and the angles of P--N and \ce{C_\beta-C_\gamma} (stars) vectors 
+    with respect to the membrane normal of POPC (left) and POPS (right) in a POPC:POPS (5:1) bilayer 
+    as a function of \ce{KCl} and \ce{NaCl} concentration from ECC-lipids and Lipid17/Dang simulations 
+    compared with experimental values from Ref. \citenum{roux90} (signs from Ref. \citenum{ferreira16}) at 298 K.
+    The y-axis for the $\alpha$-carbon results of POPS (middle right) is transferred
+    with the same value for both order parameters such that the lower order
+    parameter value from pure POPS is at zero to correctly illustrate the significant forking.
+    Because experimental data with \ce{NaCl} is not available for POPC, the data for KCl and \ce{LiCl} (dashed line, left)
+    are shown as lower and upper bounds for the response to \ce{NaCl}, respectively.
+  }
+\todo{J: Will improve this figure with panels A, B, C and D 
+and put POPC together, POPS together 
+and remove the duplicate y-axis titles 
+and make a common legend. }
+\end{figure}
 
@@
   }
+\todo{J: I suggest to put this to SI. }
 \end{figure*}
-
-First, we monitor the response of POPC headgroup order parameters to the increasing amount of POPS (Fig.~\ref{fig:delta_ordPar_NaCl_PC-PS_mix}).
-According to the electrometer concept, the head group order parameters of PC lipids increase
-with the increasing amount of negatively charged PS lipids in experiments, because the headgroup P--N dipole
-tilts more parallel to the membrane interface \cite{seelig87,scherer87}.
-The trend is observed with both force fields in simulations with \ce{K+} counterions and
-in ECC-POPS simulation with \ce{Na+} counterions, but not in Lipid17/Dang ion simulations with \ce{Na+} counterions,
-where the order parameters are unchanged or decrease and the P--N vector tilts more perpendicular to the membrane interface.
-The result can be explained by the higher binding affinity of sodium in the
-Lipid17/Dang ion simulations (Fig~\ref{fig:POPS-counterions-dens}), which
-compensates the negative charge of PS lipids at the interface.
-In contrast, ECC-lipids represent this trend correctly
-with only slightly underestimated slopes of the order parameters 
-when \ce{Na+} counterions are used (Fig.~\ref{fig:delta_ordPar_NaCl_PC-PS_mix}).
-
-Second, we monitor the response of PC headgroup order parameters to the additional
-amount of sodium and potassium in POPC:POPS (5:1) mixtures.
-In experiments, the addition of K$^+$ ions led to a very modest decrease of the order parameters,
-while the decrease was more pronounced with Li$^+$ ions and was the strongest with divalent Ca$^{2+}$ and Mg$^{2+}$ ions \cite{roux90}.
-According to the electrometer concept, the decrease of PC lipid head group order parameters
-correlates with the amount of bound cations \cite{akutsu81,altenbach84,seelig87,roux90,catte16,NMRlipidsIV},
-thus the binding affinity of the ions to POPC:POPS (5:1) mixtures increase in order K$^{+}$ $<$ Li$^{+}$  $<$ Mg$^{2+}$  $<$ Ca$^{2+}$.
-The response of PC headgroup order parameters to the increasing amount of KCl is close to experiments
-and the change in P--N vector is negligible in both force fields (Fig. \ref{fig:delta_ordPar_KCl_PCPS}), athough the change of the $\alpha$-carbon
-order parameter may be slightly overestimated by the Lipid17/Dang simulations. 
-This could be explained by the slightly deeper penetration of potassium into the bilayer (Fig. \ref{fig:POPS-counterions-dens}).
-
-Because experimental data with the additional amount of sodium in POPC:POPS (5:1) mixture was not available,
-we compared the simulation results to the experimental data with the additional amounts of
-lithium and potassium ions, which should give the lower and upper bounds to the sodium binding
-affinity (Fig. \ref{fig:delta_ordPar_NaCl_PCPS}). The PC headgroup response to the additional
-amount of sodium is clearly overestimated by the Lipid17/Dang simulation and
-the response in ECC-lipids simulations is probably too close to the experimental results for lithium.
-% The following conclusion is already in the previous sentence, so I commented it out
-%In conclusion, the results suggest that the sodium binding affinity (Fig. \ref{fig:POPS-counterions-dens}) and
-%the tilting of P--N vector more perpendicular to the membrane surface (Fig. \ref{fig:delta_ordPar_NaCl_PCPS})
-%are clearly overestimated in Lipid17/Dang simulations and slightly overestimated in ECC-POPS simulations,
-%while the binding affinity of potassium is better described by both force fields.
-
-%\subsection{Molecular interaction and binding affinities of K$^+$ and Na$^+$ cations to mixed POPC:POPS (5:1) membrane} 
-%\label{sec:K_Na_added}
-
-\begin{figure}[tbp!] 
-  \centering 
-  \includegraphics[height=\figheight]{../img/ecc_pops/order_parameters_changes_ecc-lip_L14_A-B-PN-COO_POPC_nacl.pdf} 
-  \includegraphics[height=\figheight]{../img/ecc_pops/order_parameters_changes_ecc-lip_L14_A-B-PN-COO_POPS_nacl.pdf} 
-  \caption{\label{fig:delta_ordPar_NaCl_PCPS} 
-    Changes of the head group order parameters, and the angles of P--N and C$_\beta$--C$_\gamma$ (stars) vectors
-    with respect to the membrane normal of POPC (left) and POPS (right) in a POPC:POPS (5:1) bilayer 
-    as a function of \ce{NaCl} concentration from ECC-lipids and Lipid17/Dang simulations at 298 K.
-    The y-axis for the $\alpha$-carbon results of POPS (middle right) is transferred
-    with the same value for both order parameters such that the lower order
-    parameter value from pure POPS is at zero to correctly illustrate the significant forking.
-    Because experimental data with \ce{NaCl} is not available for POPC, the data for KCl and \ce{LiCl} (dashed line, left)
-    are shown as lower and upper bounds for the response to \ce{NaCl}, respectively.
-    The magnitudes of experimental order parameters are from Ref. \citenum{roux90} and signs from Ref. \citenum{ferreira16}.
-  }
-  \end{figure}
-
-\begin{figure}[tbp!] 
-  \centering 
-  \includegraphics[height=\figheight]{../img/ecc_pops/order_parameters_changes_ecc-lip_L14_A-B-PN-COO_POPC_kcl.pdf} 
-  \includegraphics[height=\figheight]{../img/ecc_pops/order_parameters_changes_ecc-lip_L14_A-B-PN-COO_POPS_kcl.pdf} 
-  \caption{\label{fig:delta_ordPar_KCl_PCPS}
-    Changes of the head group order parameters, and the angles of P--N and C$_\beta$--C$_\gamma$ (stars) vectors 
-    with respect to the membrane normal of POPC (left) and POPS (right) in a POPC:POPS (5:1) bilayer 
-    as a function of \ce{KCl} concentration from ECC-lipids and Lipid17/Dang simulations 
-    compared with experimental values from Ref. \citenum{roux90} (signs from Ref. \citenum{ferreira16}) at 298 K.
-    The y-axis for the $\alpha$-carbon results of POPS (middle right) is transferred
-    with the same value for both order parameters such that the lower order
-    parameter value from pure POPS is at zero to correctly illustrate the significant forking.
-  }
-\end{figure}
 
@@
 or monovalent ion concentration in solvent
-(Figs. \ref{fig:delta_ordPar_NaCl_PCPS}, \ref{fig:delta_ordPar_KCl_PCPS} and \ref{fig:delta_ordPar_NaCl_PC-PS_mix}).
+(Figs. \ref{fig:delta_ordPar_monoval_PCPS}, \ref{fig:delta_ordPar_monoval_PCPS} and \ref{fig:delta_ordPar_NaCl_PC-PS_mix}).
 %while calcium leads to relatively large changes with low concentrations followed by a rapid saturation below 100mM (Fig. \ref{fig:delta_ordPar_CaCl_PCPS}) \cite{scherer87,roux90}.
@@
 of sodium and potassium in Lipid17/Dang simulations, while dependence is weaker and closer to experiments
-in ECC-POPS simulations (Figs. \ref{fig:delta_ordPar_NaCl_PCPS} and \ref{fig:delta_ordPar_KCl_PCPS}).
+in ECC-POPS simulations (Figs. \ref{fig:delta_ordPar_monoval_PCPS} and \ref{fig:delta_ordPar_monoval_PCPS}).
 These structural changes can relate to variations in the interactions between PC and PS head groups 

### 2019-05-21 Manuscript/manuscript.tex
Edited monovalent ions section

@@
 resolution of $\pm 0.01$. The C--H bond order parameter determined from a given dipolar
-splitting, $\Delta \nu$ (see e.g. Fig. \ref{R-PDLFslices} in SI)
+splitting, $\Delta \nu$ (see e.g. Fig.~\ref{R-PDLFslices} in SI)
 is equal to $\Delta \nu / (0.315 \times 21.5 \, \mathrm{kHz})$, where 0.315 is
@@
 We first determined the order parameters of each peak corresponding to different C--H bonds detected
-in the first dimension (Fig. \ref{R-PDLF}) from the dipolar slices in the indirect dimension (Fig. \ref{R-PDLFslices}).
+in the first dimension (Fig.~\ref{R-PDLF}) from the dipolar slices in the indirect dimension (Fig.~\ref{R-PDLFslices}).
 Because the exact assignment of these peaks was not possible, we assigned the
-acyl chain order parameters to mimic the profile predicted by the MD simulations (Fig. \ref{simVSexpNOions_POPS}).
+acyl chain order parameters to mimic the profile predicted by the MD simulations (Fig.~\ref{simVSexpNOions_POPS}).
 \todo{Tiago has to check and update this and figures in SI.}
@@
 The Lipid17/Dang simulation gives larger acyl chain order parameters and smaller area per molecule than the
-experimental values (Fig. \ref{simVSexpNOions_POPS} and table \ref{tab:apls}),
+experimental values (Fig.~\ref{simVSexpNOions_POPS} and table~\ref{tab:apls}),
 indicating that this simulation predicts a too compact POPS lipid bilayer. This is also in line with
@@
 This is the case also for Lipid17 POPS simulations, where the headgroup and glycerol backbone
-order parameters are quite far from experimental values with both Dang (Fig. \ref{simVSexpNOions_POPS})
+order parameters are quite far from experimental values with both Dang (Fig.~\ref{simVSexpNOions_POPS})
 and {\AA}qvist ions. \cite{NMRlipidsIV} The values from ECC-POPS simulation are closer,
@@
 are in a good agreement with the experimental values 
-as shown in Fig. \ref{simVSexpNOions_POPS}. 
+as shown in Fig.~\ref{simVSexpNOions_POPS}. 
 The acyl chain order parameters in particular agree with the experimental data within error bars.
@@
     (only \ce{Na^+} counterions, but shown also in the right plots for \ce{K+}).
-    \textbf{Right: Chemical structure and labelling of carbon segments of POPC. }
+    \textbf{Right:} Chemical structure and labelling of carbon segments of POPC. 
   }
@@
 
-First, we monitor the response of POPC headgroup order parameters to the increasing amount of POPS (Fig.~\ref{fig:delta_ordPar_NaCl_PC-PS_mix}).
+First, we monitor the response of POPC headgroup order parameters to the increasing amount of POPS (Fig.~\ref{fig:delta_ordPar_NaCl_PC-PS_mix_PC}).
 According to the electrometer concept, the head group order parameters of PC lipids increase
@@
 The trend is observed with both force fields in simulations with \ce{K+} counterions and
-in ECC-POPS simulation with \ce{Na+} counterions, but not in Lipid17/Dang ion simulations with \ce{Na+} counterions,
-where the order parameters are unchanged or decrease and the P--N vector tilts more perpendicular to the membrane interface.
+in ECC-lipids simulation with \ce{Na+} counterions. 
+In contrast, the order parameters are unchanged or decrease 
+and the P--N vector tilts more perpendicular to the membrane interface.
+in Lipid17/Dang ion simulations with \ce{Na+} counterions.
 The result can be explained by the higher binding affinity of sodium in the
@@
 compensates the negative charge of PS lipids at the interface.
-In contrast, ECC-lipids represent this trend correctly
-with only slightly underestimated slopes of the order parameters 
-when \ce{Na+} counterions are used (Fig.~\ref{fig:delta_ordPar_NaCl_PC-PS_mix}).
+This is corrected in ECC-lipids, 
+which only slightly underestimate the slopes of the order parameters 
+when \ce{Na+} counterions are used (Fig.~\ref{fig:delta_ordPar_NaCl_PC-PS_mix_PC}).
 
 Second, we monitor the response of PC headgroup order parameters to the additional
-amount of sodium and potassium in POPC:POPS (5:1) mixtures.
+amount of \ce{NaCl} and \ce{KCl} in POPC:POPS (5:1) mixtures.
 In experiments, the addition of K$^+$ ions led to a very modest decrease of the order parameters,
@@
 The response of PC headgroup order parameters to the increasing amount of KCl is close to experiments
-and the change in P--N vector is negligible in both force fields (Fig. \ref{fig:delta_ordPar_monoval_PCPS}), athough the change of the $\alpha$-carbon
+and the change in P--N vector is negligible in both force fields (Fig.~\ref{fig:delta_ordPar_monoval_PCPS}), athough the change of the $\alpha$-carbon
 order parameter may be slightly overestimated by the Lipid17/Dang simulations. 
-This could be explained by the slightly deeper penetration of potassium into the bilayer (Fig. \ref{fig:delta_ordPar_NaCl_PC-PS_mix_PC}).
-
-Because experimental data with the additional amount of sodium in POPC:POPS (5:1) mixture was not available,
+This could be explained by the deeper penetration of \ce{K+} into the bilayer (Fig.~\ref{fig:delta_ordPar_NaCl_PC-PS_mix_PC}).
+
+Because experimental data with the additional amount of \ce{Na+} in POPC:POPS (5:1) mixture was not available,
 we compared the simulation results to the experimental data with the additional amounts of
-lithium and potassium ions, which should give the lower and upper bounds to the sodium binding
-affinity (Fig. \ref{fig:delta_ordPar_monoval_PCPS}). The PC headgroup response to the additional
-amount of sodium is clearly overestimated by the Lipid17/Dang simulation and
+\ce{Li+} and \ce{K+} ions, which should give the lower and the upper bounds to the sodium binding
+affinity (Fig.~\ref{fig:delta_ordPar_monoval_PCPS}). The PC headgroup response to the additional
+amount of sodium is clearly overestimated by the Lipid17/Dang simulation, and
 the response in ECC-lipids simulations is probably too close to the experimental results for lithium.
@@
 %the tilting of P--N vector more perpendicular to the membrane surface (Fig. \ref{fig:delta_ordPar_NaCl_PCPS})
-%are clearly overestimated in Lipid17/Dang simulations and slightly overestimated in ECC-POPS simulations,
+%are clearly overestimated in Lipid17/Dang simulations and slightly overestimated in ECC-lipids simulations,
 %while the binding affinity of potassium is better described by both force fields.
@@
 or monovalent ion concentration in solvent
-(Figs. \ref{fig:delta_ordPar_monoval_PCPS}, \ref{fig:delta_ordPar_monoval_PCPS} and \ref{fig:delta_ordPar_NaCl_PC-PS_mix}).
+(Figs.~\ref{fig:delta_ordPar_monoval_PCPS},~\ref{fig:delta_ordPar_monoval_PCPS} and~\ref{fig:delta_ordPar_NaCl_PC-PS_mix}).
 %while calcium leads to relatively large changes with low concentrations followed by a rapid saturation below 100mM (Fig. \ref{fig:delta_ordPar_CaCl_PCPS}) \cite{scherer87,roux90}.
@@
 are overestimated, which is also related to the changes in the orientations of the P--N and C$_{\beta}$--C$_{\gamma}$
-vectors (Fig. \ref{fig:delta_ordPar_NaCl_PC-PS_mix}). The overestimation of the order parameter changes
-upon increasing amount of POPC in a bilayer 
-was also observed in previous simulations with other force fields \cite{NMRlipidsIV}, but not in
-ECC-POPS simulation, where also the changes in the PS headgroup orientation are more modest (Fig. \ref{fig:delta_ordPar_NaCl_PC-PS_mix}).
-The POPS headgroup orientation and order parameters are more sensitive also to the additional amount
-of sodium and potassium in Lipid17/Dang simulations, while dependence is weaker and closer to experiments
-in ECC-POPS simulations (Figs. \ref{fig:delta_ordPar_monoval_PCPS} and \ref{fig:delta_ordPar_monoval_PCPS}).
-These structural changes can relate to variations in the interactions between PC and PS head groups 
-and also to the changing amount of counterions and PS in the composition. 
+vectors (Fig.~\ref{fig:delta_ordPar_NaCl_PC-PS_mix_PC}). 
+%These structural changes can relate to variations in the interactions between PC and PS head groups 
+%and also to the changing amount of counterions and PS in the composition. 
+Also, the PS headgroup order parameters from simulations with Lipid17/Dang model
+are too sensitive to an additional amount of \ce{KCl} and \ce{NaCl} 
+(Fig.~\ref{fig:delta_ordPar_monoval_PCPS}).
+The overestimation of the order parameter changes upon increasing amount of POPC in a bilayer 
+was also observed in previous simulations with other force fields \cite{NMRlipidsIV}, 
+but not in ECC-lipids simulation, 
+where the changes in the PS headgroup orientation are more modest (Fig.~\ref{fig:delta_ordPar_NaCl_PC-PS_mix_PC}).
+In particular, 
+the dependence of the PS headgroup order parameters from ECC-lipids simulations
+to increasing amount of POPC, \ce{KCl} or \ce{NaCl}
+is weaker and closer to experiments 
+(Fig.~\ref{fig:delta_ordPar_monoval_PCPS}).
 %In contrast, ECC-lipids with ECC-ions capture the different response 
@@
 %
-The influence of negatively charged PS on the counterion binding affinity can
-be estimated by comparing the relative surface excesses with respect to water, $\Gamma ^{w} _{\rm ion}$,
+The influence of negatively charged PS on the counterion binding affinity can be estimated 
+%by calculating the amounts of phospholipids per bound \ce{Ca^{2+}} cation, $\xi _{Ca^{2+}}$,
+by comparing the relative surface excesses with respect to water, $\Gamma ^{w} _{\rm ion}$,
 between POPC:POPS (5:1) mixture and pure POPC \cite{melcr18}.
+%Threshold for counting bound \ce{Ca^{2+}} cations to a membrane
+%was set to $0.3\,\mathrm{nm}$ from oxygen atoms in lipids as in Ref.~\citenum{melcr18}.  
 %Such a quantity compares the adsorption of ions to the adsorption of water molecules at an interface 
@@
 simulation was reported in the previous work \cite{melcr18}.
-The presence of $\sim$17\$ POPS in POPC bilayer increases the
+The presence of $\sim$17\% POPS in a POPC bilayer increases the
 %relative surface excess of sodium with respect to water from negative
@@
 for 0.621 M sodium, while value for potassium remains negative,
-$\Gamma ^{w} _{\rm K}=-0.123 \pm 0.005 \mathrm{nm}^{-2}$ in POPC:POPS (5:1) mixture (Fig. \ref{fig:nacl-dens_PCPS}).
+$\Gamma ^{w} _{\rm K}=-0.123 \pm 0.005 \mathrm{nm}^{-2}$ in POPC:POPS (5:1) mixture (Fig.~\ref{fig:nacl-dens_PCPS}).
 As discussed above, the sodium binding to POPC:POPS (5:1) mixture may be slightly overestimated also in
@@
 %by yielding a small positive value of $\Gamma^{w}_{Na}$.
-
-In conclusion, applying ECC to Amber Lipid17 POPS force field parameters improves its
-interactions with both monovalent ions and POPC headgroups when compared with the NMR order parameter data.
-However, further optimization may be needed to fully reproduce the interactions between lipids and
-sodium ions.
 
@@
 In ECC-lipids simulations, on average 41\% of the total population of bound calcium cations is in contact with PS lipids
-with 8\% bound only to them, even though the negatively charged membrane contains only 18\% of POPS (Table \ref{tab:binding}).
+with 8\% bound only to them, even though the negatively charged membrane contains only 17\% of POPS (Table~\ref{tab:binding}).
 Therefore, calcium ions bind more likely to negatively charged PS lipids than neutral PC lipid in a mixed bilayer, as expected.
@@
 
+In conclusion, applying ECC to Amber Lipid17 POPS force field parameters improves its
+interactions with both monovalent ions and POPC headgroups when compared with the NMR order parameter data.
+However, further optimization may be needed to fully reproduce the interactions between lipids and
+sodium ions.
+
 \listoftodos

### 2019-05-21 Manuscript/manuscript.tex
Edited section about calcium interactions

@@
 
-\begin{figure*}[!tbp] 
-  \centering 
-  \includegraphics[height=\figheightsmall]{../Fig/order_parameters_changes_A-B_PC-PS_mix_POPS_nacl.pdf} 
-  \includegraphics[height=\figheightsmall]{../Fig/order_parameters_changes_A-B_PC-PS_mix_POPS_kcl.pdf} 
-  \caption{\label{fig:delta_ordPar_NaCl_PC-PS_mix_PS} 
-    The POPS head group order parameters, the P--N and C$_\beta$--C$_\gamma$ vector angles (stars)
-    with respect to the membrane normal as a function of POPC content in a bilayer
-    from ECC-lipids and Lipid17/Dang simulations with \ce{Na^+} (left) and \ce{K^+} (right) counterions.
-    Experimental order parameter values and signs for pure POPS are from Ref. \citenum{NMRlipidsIV} and values
-    for POPC:POPS (5:1) mixture from Ref. \citenum{roux90} 
-    (only \ce{Na^+} counterions, but shown also in the right plots for \ce{K+}).
-    % No translation in y-axis done in this plot:
-    %The underestimated response of ECC-lipids with \ce{Na^+} counterions 
-    %is likely due to a still slightly overestimated binding affinity of \ce{Na^+} to the phospholipids,
-    %which is corroborated by the series with \ce{K^+} counterions (lower affinity than \ce{Na^+}),
-    %where the response is closer to the experiment, which uses only \ce{Na^+} counterions. 
-  }
-\todo{J: I suggest to put this to SI. }
-\end{figure*}
-
 The changes of POPS headgroup order parameters are almost negligible in experiments with increasing POPC fraction in bilayer
@@
 Our recent work demonstrates that the POPC headgroup order parameters measured
-from POPC:POPS (5:1) mixture as a function of CaCl$_2$ concentration \cite{roux90}
-can be used to evaluate the calcium binding affinity to lipid bilayers containing PS lipids \cite{NMRlipidsIV}.
+from POPC:POPS (5:1) mixture as a function of CaCl$_2$ concentration 
+can be used to evaluate the calcium binding affinity to lipid bilayers containing PS lipids \cite{roux90, NMRlipidsIV}.
 The decrease of the PC headgroup order parameters in this mixture upon addition of \ce{CaCl2} 
 were overestimated by the Lipid17/Dang simulations (also shown in Fig.~\ref{fig:delta_ordPar_CaCl_PCPS})
-and all the other tested force fields, except CHARMM36 with the recently introduced NBfix for calcium \cite{kim16},
+and all other tested force fields except CHARMM36 with the recently introduced NBfix for calcium \cite{kim16},
 which underestimated the headgroup order parameter response.
@@
 ECC-lipids simulations by the increasing relative surface excess of calcium with respect to water from
-$\Gamma^{w}_{Ca} = 0.06\mathrm{nm^{-2}}$ for pure POPC bilayer with 350 mM CaCl$_2$ \cite{melcr18}
-to $\Gamma^{w}_{Ca} = 0.24\mathrm{nm^{-2}}$ for the POPC:POPS (5:1) bilayer with 409 mM CaCl$_2$.
+%  previously we used 350mM conc, but I think it is better to compare to a higher concentration to support the whole argument bullet-proof
+%  $\Gamma^{w}_{Ca} = 0.06\mathrm{nm^{-2}}$ ($\sim ?.?$ phospholipids per bound \ce{Ca^{2+}}) for pure POPC bilayer with 346~mM CaCl$_2$ \cite{melcr18, ECC-POPC_nacl_cacl2_files} 
+   $\Gamma^{w}_{Ca} = 0.09\mathrm{nm^{-2}}$ ($\sim 5.4$ phospholipids per bound \ce{Ca^{2+}}) for pure POPC bilayer with 467~mM CaCl$_2$ \cite{ECC-POPC_nacl_cacl2_files} 
+to $\Gamma^{w}_{Ca} = 0.24\mathrm{nm^{-2}}$ ($\sim 4.8$ phospholipids per bound \ce{Ca^{2+}}) for the POPC:POPS (5:1) bilayer with 409 mM CaCl$_2$.
 
@@
 discrepancy include the above observed slight overestimation of \ce{Na^+} counterions and
-inaccurate structures of the lipid headgroup (Fig.~\ref{simVSexpNOions_POPS}).
+imperfect structures of the lipid headgroup (Fig.~\ref{simVSexpNOions_POPS}).
 
 In ECC-lipids simulations, the average orientation of P--N vectors of both POPC and POPS headgroups tilts more perpendicular to the membrane surface
-by 11$^\circ$ and  3$^\circ$, respectively, upon addition of 620~mM CaCl$_2$ to POPC:POPS (5:1) mixture (Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}),
+by 11$^\circ$ and  3$^\circ$, respectively, upon addition of 620~mM CaCl$_2$ to POPC:POPS (5:1) mixture (Fig.~\ref{fig:delta_ordPar_CaCl_PCPS})
 suggesting that the PS headgroup orientation is less sensitive to the bound calcium than PC.
-On the other hand, the average orientation of the C$_{\beta}$--C$_{\gamma}$ vector PS headgroup
+On the other hand, the average orientation of the C$_{\beta}$--C$_{\gamma}$ vector in PS headgroup
 tilts to the opposite direction by $11^\circ$ in the same system.
-Such a largely diminished flexibility of the PS head group can be explained 
-by the simultaneous binding of calcium to the \ce{COO^-} group in PS and phosphate group in PC. 
 In Lipid17/Dang simulations, the changes in both headroup orientations and order parameters
 are larger and less systematic for the PS headgroup, especially with lower concentrations (Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}).
-ECC-lipids simulations yield smaller error bars and shorter residence times (Fig.~\ref{fig:hist_residence_times}) in
+
+ECC-lipids simulations yield smaller error bars and shorter residence times (Fig.~\ref{fig:hist_residence_times}) 
 than those typically observed in simulations with other force fields \cite{javanainen17,melcr18,NMRlipidsIV}.
 This suggests that the ECC accelerates the equlibration of ions at lipid bilayer interface.
-
-Our 1 $\mu$s simulations seems to be sufficiently long for the ECC-lipids simulations, because
+Our 1 $\mu$s simulations seem to be sufficiently long for the ECC-lipids simulations, because
 90\% of the calcium residence times are shorter than $60\,\mathrm{ns}$ for pure POPC bilayer
@@
 Interestingly, the calcium residence times are 3-4 times longer in POPC:POPS (5:1) mixture than in pure POPC.
-In ECC-lipids simulations, on average 41\% of the total population of bound calcium cations is in contact with PS lipids
-with 8\% bound only to them, even though the negatively charged membrane contains only 17\% of POPS (Table~\ref{tab:binding}).
-Therefore, calcium ions bind more likely to negatively charged PS lipids than neutral PC lipid in a mixed bilayer, as expected.
 
@@
 In ECC-lipids simulation of POPC:POPS (5:1) mixture with 409 mM CaCl$_2$,
-the calcium ions bind more likely 
-to the carboxylate than to phosphate moiety of POPS headgroup (Fig.~\ref{fig:cacl_complexes}),
+on average 41\% of the total population of bound calcium cations is in contact with PS lipids
+with 8\% bound exclusively only to them even though the negatively charged membrane contains only 17\% of POPS (Fig.~\ref{fig:cacl_complexes}).
+%Therefore, calcium ions bind more likely to negatively charged PS lipids than neutral PC lipid in a mixed bilayer, as expected.
+Also, calcium ions bind more likely 
+to the carboxylate than to the phosphate moiety of POPS headgroup (Fig.~\ref{fig:cacl_complexes}),
 in line with the results without the NBfix corrections in CHARMM36 \cite{hallock18}.
+This can explain the diminished flexibility of the PS head group 
+from simultaneous binding of \ce{Ca^{2+}} to the \ce{COO^-} group in PS and a phosphate group in PC 
+as the most probable configuration (Fig.~\ref{fig:cacl_complexes}).
+
 However, in CHARMM36 simulations without the NBfix correction for calcium \cite{kim16},
@@
 with even four distinct PS lipids \cite{hallock18}, while coordination with more than two PS lipids
-was not observed in our ECC-lipids simulations (Fig.~\ref{fig:cacl_complexes}).
+was very rare in our ECC-lipids simulations (Fig.~\ref{fig:cacl_complexes}).
 Furthermore, the POPS:POPC:Ca$^{2+}$ (7:3:7) mixture with relatively high
@@
 
-Significant coordination of calcium with carbonyls groups of PS and PC lipids observed in previous
+Significant coordination of calcium with carbonyl groups of PS and PC lipids observed in previous
 simulations with the Berger force field \cite{melcrova16} %\cite{berger97,mukhopadhyay04}
@@
 Overestimated coordination of cations with carbonyl groups is a potential source for
-overestimated cation binding in Berger force fields \cite{catte16,NMRlipidsIV}. 
+the overestimated cation binding in Berger force field \cite{catte16,NMRlipidsIV}. 
 In CHARMM36 simulations with the NBfix, calcium mainly interacted with only
@@
   \textbf{Left:} Percentages of the population of bound Ca$^{2+}$ 
-in exclusive contact with various lipid moieties 
+in exclusive contact with the given oxygen moiety
 in a POPC:POPS (5:1) bilayer with 409 mM CaCl$_2$. 
-The bar representing total populations of bound \ce{Ca^{2+}} to PC, PS, or PC and PS (blue),
-is in part covered by areas representing contributions of individual moieties exclusively (shades of orange and red). 
+The areas represent populations of bound \ce{Ca^{2+}} 
+only to the given moieties exclusively (shades of orange and red)
+or their combinations (CHANGE:blue)
+separately for PC, PS, or PC and PS simultaneously. 
+The areas sum up to total populations at the respective lipids. 
 In the case of mixed simultaneous binding to PC and PS
@@
   }
+\todo{J: ad bar plots -- change total to combinations as described in caption, 
+add simultaneously to "PC and PS", 
+add ticks to the right side of the plot and
+change the colour of "combinations" from blue to any other so that it is different from the legend in the other plot. }
 \end{figure}
+
+\section{Conclusions}
+
+\todo{SAMULI: I have putted here two preliminary sentences
+JOSEF: I can write conclusions, but I would prefer to wait with that after your comments and edits. }
+Applying ECC to POPS lipids improves cation binding affinity to membranes
+mixed with POPC and also the response of PS headgroup to the bound cations.
+However, further improvement of the parameters is needed to fully reproduce the
+details of calcium binding to PS headgroups.
 
@@
 Because the ECC-lipids model gives the best agreement with the experimental headgroup order parameters
-with various ion conctenrations among the available models, we believe that it gives the currently best available
+with various ion conctenrations among the available models, we believe that it also gives the currently best available
 interpretation of the experimental data. Athough the ECC-lipids model gives the most realistic picture of calcium binding to
@@
 % and also possibly from the approximate mean-field treatment of the electronic polarization.
-
-\section{Conclusions}
-
-\todo{SAMULI: I have putted here two preliminary sentences
-JOSEF: I can write conclusions, but I would prefer to wait with that after your comments and edits. }
-Applying ECC to POPS lipids improves cation binding affinity to membranes
-mixed with POPC and also the response of PS headgroup to the bound cations.
-However, further improvement of the parameters is needed to fully reproduce the
-details of calcium binding to PS headgroups.
 

### 2019-05-21 Manuscript/manuscript.tex
Grouped the figures in the section with calcium into one panel and glued the paragraphs in conclusion.

@@
 and remove the duplicate y-axis titles 
-and make a common legend. }
+and make a common legend distinguishing different experiments better. }
 \end{figure}
@@
   \centering 
-  \includegraphics[height=\figheight]{../img/ecc_pops/order_parameters_changes_ecc-lip_L14_A-B-PN-COO_POPC_cacl.pdf} 
-  \includegraphics[height=\figheight]{../img/ecc_pops/order_parameters_changes_ecc-lip_L14_A-B-PN-COO_POPS_cacl.pdf} 
+  \begin{tabular}{ c }
+  \includegraphics[height=\figheightsmall]{../img/ecc_pops/density_profiles_ca_na_k_cl_wat_phos_ecclipids_lipid17_compar_80and200mMCaCl2.pdf}
+  \includegraphics[height=\figheightsmall]{../img/ecc_pops/order_parameters_changes_ecc-lip_L14_A-B-PN-COO_POPC_cacl.pdf} 
+  \includegraphics[height=\figheightsmall]{../img/ecc_pops/order_parameters_changes_ecc-lip_L14_A-B-PN-COO_POPS_cacl.pdf} 
+  \end{tabular}
+  \begin{tabular}{ c }
+  \includegraphics[width=3cm]{../img/bound_ca_populations.pdf} \\
+  \includegraphics[width=3cm]{../img/stoichiometry_CaCl2_comparison_Ecc-lipids_PC-vs-PCPS.pdf} 
+  \end{tabular}
+  \caption{\label{fig:cacl-dens_PCPS} 
+    \textbf{Left panel:}
+    Number density profiles of \ce{Ca^{2+}} and \ce{Cl^-} ions and \ce{Na^{+}} counterions 
+    along the normal of the membrane starting at the centre
+    for the negatively charged membrane with the composition 5\,PC:1\,PS
+    at 80~mM (top) and 200~mM (bottom) added buffer concentrations of \ce{CaCl2} from simulations with ECC-lipids (solid) and Lipid17 (dashed). 
+    In order to visualize the density profiles with a comparable scale 
+    the density profile of~\ce{Ca^{2+}} ions are divided by 2, and 
+    the density profiles of phosphate groups and water are divided by 4 and 100, respectively.  
+  }
   \caption{\label{fig:delta_ordPar_CaCl_PCPS} 
+    \textbf{Middle panels:}
     Changes of the head group order parameters, and the angles of P--N and C$_\beta$--C$_\gamma$ (stars) vectors 
@@
   }
+  \caption{ \label{fig:cacl_complexes} 
+  \textbf{Right top:} Percentages of the population of bound Ca$^{2+}$ 
+	in exclusive contact with the given oxygen moiety
+	in a POPC:POPS (5:1) bilayer with 409 mM CaCl$_2$. 
+	The areas represent populations of bound \ce{Ca^{2+}} 
+	only to the given moieties exclusively (shades of orange and red)
+	or their combinations (CHANGE:blue)
+	separately for PC, PS, or PC and PS simultaneously. 
+	The areas sum up to total populations at the respective lipids. 
+	In the case of mixed simultaneous binding to PC and PS
+	we distinguish different moieties only in PS.
+	%i.e., PC can interact with any considered moiety. 
+	For comparison, in a pure POPC bilayer with 350 mM CaCl$_2$, 
+	67\% of bound \ce{Ca^{2+}} is exclusively in contact with \ce{PO_4} group. \cite{melcr18} 
+    \textbf{Right bottom:} Relative probabilities of \ce{Ca^{2+}} ions to coordinate with a certain number of lipids
+    in pure POPC bilayer with 350 mM CaCl$_2$ and in POPC:POPS (5:1) mixture with 400 mM CaCl$_2$.  
+    Analysis was done for both systems by considering all lipids (blue and violet) and
+    for POPC:POPS (5:1) mixture also by considering POPC and POPS lipids separately (green and orange). 
+    %All lipids were taken into account with the exception of the complexes in light green and orange, 
+    %for which we counted only contacts with POPC resp POPS from the mixed 5\,PC:1\,PS negatively charged bilayer. 
+    %The calculated probabilities of the calcium-lipid complexes also reflect only POPC (light green) resp POPS (orange). 
+    % Probabilities were taken from simulations with comparable bulk concentrations of calcium around 250~mM. 
+    Clusters of four or more lipids were not observed in either membrane.
+    The threshold for counting coordinated lipids in a complex with \ce{Ca^{2+}} was set to $0.3\,\mathrm{nm}$, 
+    as the distance between the cation and the oxygen atoms of the lipids. 
+    Previously published simulation data \cite{melcr18} for pure POPC bilayers were taken directly from \cite{ECC-POPC_nacl_cacl2_files}. 
+  }
+\todo{J: Will adjust the Fig into a nice panel and change the captions accordingly. 
+Ad bar plots -- change total to combinations as described in caption, 
+add simultaneously to "PC and PS", 
+add ticks to the right side of the plot and
+change the colour of "combinations" from blue to any other so that it is different from the legend in the other plot. }
 \end{figure}
@@
 Interestingly, the calcium residence times are 3-4 times longer in POPC:POPS (5:1) mixture than in pure POPC.
-
-\begin{figure}[htbp!] 
-  \centering 
-  \includegraphics[width=\figwidth]{../img/ecc_pops/density_profiles_ca_na_k_cl_wat_phos_ecclipids_lipid17_compar_80and200mMCaCl2.pdf}
-  \caption{\label{fig:cacl-dens_PCPS} 
-    Number density profiles of \ce{Ca^{2+}} and \ce{Cl^-} ions and \ce{Na^{+}} counterions 
-    along the normal of the membrane starting at the centre
-    for the negatively charged membrane with the composition 5\,PC:1\,PS
-    at 80~mM (top) and 200~mM (bottom) added buffer concentrations of \ce{CaCl2} from simulations with ECC-lipids (solid) and Lipid17 (dashed). 
-    In order to visualize the density profiles with a comparable scale 
-    the density profile of~\ce{Ca^{2+}} ions are divided by 2, and 
-    the density profiles of phosphate groups and water are divided by 4 and 100, respectively.  
-  }
-\end{figure}
 
@@
 
-\begin{figure}[tb!] 
-  \centering 
-  \includegraphics[width=\figwidthsmall]{../img/bound_ca_populations.pdf} 
-  \includegraphics[width=\figwidthsmall]{../img/stoichiometry_CaCl2_comparison_Ecc-lipids_PC-vs-PCPS.pdf} \\ 
-  \caption{ \label{fig:cacl_complexes} 
-  \textbf{Left:} Percentages of the population of bound Ca$^{2+}$ 
-in exclusive contact with the given oxygen moiety
-in a POPC:POPS (5:1) bilayer with 409 mM CaCl$_2$. 
-The areas represent populations of bound \ce{Ca^{2+}} 
-only to the given moieties exclusively (shades of orange and red)
-or their combinations (CHANGE:blue)
-separately for PC, PS, or PC and PS simultaneously. 
-The areas sum up to total populations at the respective lipids. 
-In the case of mixed simultaneous binding to PC and PS
-we distinguish different moieties only in PS.
-%i.e., PC can interact with any considered moiety. 
-For comparison, in a pure POPC bilayer with 350 mM CaCl$_2$, 
-67\% of bound \ce{Ca^{2+}} is exclusively in contact with \ce{PO_4} group. \cite{melcr18} 
-    \textbf{Right:} Relative probabilities of \ce{Ca^{2+}} ions to coordinate with a certain number of lipids
-    in pure POPC bilayer with 350 mM CaCl$_2$ and in POPC:POPS (5:1) mixture with 400 mM CaCl$_2$.  
-    Analysis was done for both systems by considering all lipids (blue and violet) and
-    for POPC:POPS (5:1) mixture also by considering POPC and POPS lipids separately (green and orange). 
-    %All lipids were taken into account with the exception of the complexes in light green and orange, 
-    %for which we counted only contacts with POPC resp POPS from the mixed 5\,PC:1\,PS negatively charged bilayer. 
-    %The calculated probabilities of the calcium-lipid complexes also reflect only POPC (light green) resp POPS (orange). 
-    % Probabilities were taken from simulations with comparable bulk concentrations of calcium around 250~mM. 
-    Clusters of four or more lipids were not observed in either membrane.
-    The threshold for counting coordinated lipids in a complex with \ce{Ca^{2+}} was set to $0.3\,\mathrm{nm}$, 
-    as the distance between the cation and the oxygen atoms of the lipids. 
-    Previously published simulation data \cite{melcr18} for pure POPC bilayers were taken directly from \cite{ECC-POPC_nacl_cacl2_files}. 
-  }
-\todo{J: ad bar plots -- change total to combinations as described in caption, 
-add simultaneously to "PC and PS", 
-add ticks to the right side of the plot and
-change the colour of "combinations" from blue to any other so that it is different from the legend in the other plot. }
-\end{figure}
-
 \section{Conclusions}
 
-\todo{SAMULI: I have putted here two preliminary sentences
-JOSEF: I can write conclusions, but I would prefer to wait with that after your comments and edits. }
+\todo{J: I have moved several paragraphs from the body to here and added a bit of glue. This section still requires to be written properly.  }
+
+We have developed a classical MD model of POPS, ECC-POPS, 
+which accurately describes features of both pure POPS membrane and its mixes with POPC. 
+This was achieved by applying ECC which accounts for electronic polarization. 
 Applying ECC to POPS lipids improves cation binding affinity to membranes
 mixed with POPC and also the response of PS headgroup to the bound cations.
-However, further improvement of the parameters is needed to fully reproduce the
-details of calcium binding to PS headgroups.
-
+This was compared to accurate NMR measurements of the changes of headgroup order parameter within the ``electrometer concept''. 
 The advantage of the NMR order parameters used in this work is the direct connection between
@@
 with various ion conctenrations among the available models, we believe that it also gives the currently best available
-interpretation of the experimental data. Athough the ECC-lipids model gives the most realistic picture of calcium binding to
-lipid bilayers containing PS lipids with respect to NMR order parameters, it does not
-correctly capture the rapid changes with small concentrations. Therefore, further
-force field optimization is required for even more accurate MD simulations of cations
-in the vicinity of negatively charged lipid bilayers.
-% and also possibly from the approximate mean-field treatment of the electronic polarization.
+interpretation of the experimental data.
 
@@
 interactions with both monovalent ions and POPC headgroups when compared with the NMR order parameter data.
-However, further optimization may be needed to fully reproduce the interactions between lipids and
-sodium ions.
+Although the ECC-lipids model gives the most realistic picture of calcium binding to
+lipid bilayers containing PS lipids with respect to NMR order parameters, it does not
+correctly capture the rapid changes with small concentrations. 
+Therefore, further force field optimization is required for even more accurate MD simulations of cations
+in the vicinity of negatively charged lipid bilayers.
 

### 2019-05-22 Manuscript/manuscript.tex
Some changes

--- 2019-05-21 Manuscript/manuscript.tex
@@
 \newlength{\figheightsmall}
-\setlength{\figheightsmall}{6 cm} 
+\setlength{\figheightsmall}{8 cm} 
 \newlength{\figheight}
@@
 \author{Josef Melcr} 
-\email{melcr@marge.uochb.cas.cz} 
+\email{melcr@uochb.cas.cz} 
 %%\homepage[]{https://jmelcr.github.io/}
@@
 \title[] 
-  { Detailed structure of negatively charged membranes 
-    of phosphatidylserine and phosphatidylcholine 
-    at concentrations of calcium, sodium and potassium salts
-    from molecular dynamics simulations with electronic polarization }
-  % JOSEF: I feel that the above title is a bit better, because it focuses on PS,
-  %        but the following one is an alternative, which focuses more on the cations as in roux90
-  %{ Detailed structure of calcium, sodium and potassium cations
-  %  at negatively charged membranes of phosphatidylserine and phosphatidylcholine 
-  %  from molecular dynamics simulations with electronic polarization }
+  {Effective Inclusion of Electronic Polarization to a Negatively Charged Membrane}
 
@@
 
-Phosphatidylserine (PS) lipids are the most common negatively charged lipids in eukaryotic membranes
+The phophatidylserine (PS) lipids are most common negatively charged lipids in eukaryotic membranes
 and important signaling molecules \cite{lemmon08,leventis10,li14}.
@@
 which cause phase separation \cite{hauser77,kurland79,hauser85,feigenson86,mattai89,roux90,roux91,boettcher11}.
-Theories about the interaction between calcium and PS 
-are difficult to evaluate experimentally,
-especially at physiological ion concentrations,
-which are typically too low to yield measurable effects.
-
+Different theories are difficult to evaluate experimentally,
+especially with physiological ion concentrations which are typically quite low.
 More recently, classical molecular dynamics (MD) simulations
@@
 but the results strongly depend on the force field parameters \cite{boettcher11,kucerka14,melcrova16,hallock18,valentine18}.
-% smoother connection between previous and following statements
-For instance, the moieties of PS lipids that interact with calcium cations vary greatly.
 In simulations with the CHARMM36 force field \cite{klauda10,venable13} using the NBfix parameters for calcium \cite{kim16}, 
@@
 cation binding affinity to zwitterionic PC bilayers \cite{catte16} or to the
-mixtures with negatively charged PS lipids \cite{NMRlipidsIV}.
+mixtures with negatively charge PS lipids \cite{NMRlipidsIV}.
 In our recent work~\cite{melcr18}, we were able to improve the cation binding affinity
@@
 
-Here, we extend the ECC approach also to the negatively charged 1-palmitoyl-2-oleoyl-sn-glycero-3-phospho-L-serine (POPS) lipid.
+Here, we extend the ECC approach also to negatively charged 1-palmitoyl-2-oleoyl-sn-glycero-3-phospho-L-serine (POPS) lipid.
 We also complement the available experimental data for the the force field quality
@@
 
-Electronic continuum correction (ECC) is an implicit mean-field representation of 
-the electronic polarization in classical MD simulations. 
+The electronic continuum correction (ECC) is a simple method to implicitly include
+the electronic polarizability into classical molecular dynamics (MD) simulations
+by scaling the electric charges defined in force fields.
+%The polarization of electrons in ECC-lipids
+%uses the implicit method termed electronic continuum correction (ECC). \cite{leontyev09, Pluharova2014, martinek17, melcr18} 
+%In this method, a system of polarizable particles is represented 
+%with an equivalent system of particles with fixed charges,
+%which account for the electronic polarizability \emph{implicitly}.
 For simple ions in water, the electronic polarizability
 can be taken into account by scaling the charge with a constant factor 
+%This is practically achieved by rescaling the charges with a pre-determined scaling factor, $f_q$,
 %\begin{equation}
- \mbox{$ f_q = \frac{1}{\sqrt{\epsilon_{el}}} \approx 0.75$,}
+ $ f_q = \frac{1}{\sqrt{\epsilon_{el}}} \approx 0.75$,
 %\end{equation}
 where  $\epsilon_{el} = 1.78$ is the high-frequency dielectric constant of electrons in water \cite{leontyev09}.
+%,, as
+%The value of $\epsilon_{el} = 1.78$ for water solutions gives the scaling
+%factor of $f_q \approx 0.75$ for aquaeous environments \cite{leontyev09}.
 The scaling of charges with this factor has been recently applied to improve classical models for ions and
-other biomolecules \cite{Pluharova2014, martinek17, duboue2018insulin, Mason2019, Duboue2018MgZn} \todo{Pavel can probably update this (insulin?)}.
-In addition to the charge, the Lennard-Jones~$\sigma$ was scaled with a factor $f_\sigma \leq 1$
+other biomolecules \cite{Pluharova2014, martinek17} \todo{Pavel can probably update this (insulin?)}.
+In addition to the charge, the Lennard-Jones~$\sigma$ was scaled with a factor of $f_\sigma = ??$
 to improve the hydration properties of ions with respect to scattering data in these studies
@@
 Derivation of the correct scaling factor for lipids is more complicated, because
-the partial charges depend on the methods used to derive the respective force fields.
-For example, the partial charges may already include the effects of electronic polarizability to some extent,
-and hence, the scaling factor can be larger than the theoretically derived value for ions in water, $f_q \approx 0.75$.
-Before the ECC theory was rigorously derived, similar idea was already employed in early classical 
+%In contrast to the total charge of ions,
+the partial charges depend on the methods used to derive the force field.
+For example, the partial charges may already include some electronic polarizability effects
+and the scaling factor should then be larger than the theoretically derived value for ions in water, $f_q \approx 0.75$.
+Before the ECC theory was rigorously derived, similar idea was employed in early classical 
 MD simulations of lipids and surfactants using the scaling factor of 0.5 for atomic charges  \cite{jonsson86,egberts94, berendsen1996}. 
@@
 
-Here, we apply the ECC to the Amber-based lipid17 parameters \cite{lipid17-future} of POPS lipid with a monovalent negative charge.
-Because POPS carries a physical total charge as aqueous ions, we apply the theoretically derived scaling factor for
-ions $f_q = 0.75$ \cite{leontyev09} to the partial charges of POPS. 
-The lipids with a total charge of -0.75 are also neutralized
-by the counterion charges of +0.75 in simulations with ECC-ions \cite{Pluharova2014, kohagen16, martinek17}.
+Here, we apply the ECC to the Amber based lipid17 parameters \cite{lipid17-future} of POPS lipid with a monovalent negative charge.
+Because POPS carries a physical total charge as ions in water, we apply the theoretically derived scaling factor for
+ions $f_q = 0.75$ \cite{leontyev09} to the partial charges of POPS. The lipids with a total charge of -0.75 also neutralize
+the counterion charges of +0.75 in simulations with ECC-ions \cite{kohagen16}.
 Following our previous work for POPC, we use the scaling factor of $f_\sigma = 0.89$ for the Lennard-Jones~$\sigma$ parameters and
-use it to scale the parameters of atoms only in the headgroup, glycerol backbone, and carbonyl regions \cite{melcr18, ECC-POPC_nacl_cacl2_files}.
-Further optimization of parameters was not done in this work.
+scale the parameters of atoms only in the headgroup, glycerol backbone, and carbonyl regions \cite{melcr18}.
+Further optimization of parameters was not done in this work. \cite{ECC-POPC_nacl_cacl2_files}
+% we used exactly the pre-determined scaling factor 
+%to represent the effects of electronic polarization
+%yielding a total charge of $-0.75$ in the ECC formalism. \cite{martinek17, leontyev09, melcr18}
+Similarly to the ECC-POPC model \cite{melcr18}, we applied the ECC only to the
+non-tail atoms, i.e., to atoms in the headgroup, glycerol backbone, and carbonyl regions. 
+%polar parts of the phospholipid,
+%i.e., charges on all atoms except on alkyl tails, 
+%which do not encounter direct contacts with ions.
+
+%\subsection{Validation of membrane properties}
+% SAMULI: I have dissolved this section into results and discussion
 
@@
 The experiment was done using a Bruker Avance III 400 spectrometer operating at a 1H Larmor frequency of 400.03 MHz
-equipped with a standard 4~mm CP-MAS HXY probe. The set up of the R-PDLF experiment was the following
+equipped with a standard 4 mm CP-MAS HXY probe. The set up of the R-PDLF experiment was the following
 (using the notation from figures 1c and 2c in the original publication describing the R-PDLF experiment \cite{dvinskikh04}).
 The magic angle spinning (MAS) frequency used was 5.15 kHz. The recoupling pulses in the R18 blocks had
-therefore a length of $10.79 \, \mu \mathrm{s}$, which correspond to a nutation frequency of 46.35~kHz. Increments of $18 \times 2$ recoupling
-pulses where used giving a spectral width of 2.6~kHz in the indirect dimension and a total number of 32 points in
+therefore a length of 10.79 us which correspond to a nutation frequency of 46.35 kHz. Increments of 18x2 recoupling
+pulses where used giving a spectral width of 2.6 kHz in the indirect dimension and a total number of 32 points in
 the indirect dimension were recorded. These settings enable to record dipolar slices with a C--H bond order parameter
-resolution of $\pm 0.01$. The C--H bond order parameter determined from a given dipolar
-splitting, $\Delta \nu$ (see e.g. Fig.~\ref{R-PDLFslices} in SI)
-is equal to $\Delta \nu / (0.315 \times 21.5 \, \mathrm{kHz})$, where 0.315 is
+resolution of $\pm$0.01. The C--H bond order parameter determined from a given dipolar
+splitting, $\Delta\nu$ (see e.g. Fig. \ref{R-PDLFslices} in \todo{Is this the correct figure?}) is equal to $\Delta\nu/(0.315\times21.5 kHz)$, where 0.315 is
 the scaling factor of the R18 recoupling sequence and 21.5 kHz is the maximum $^1$H-$^{13}$C dipolar coupling for a C--H bond.
 The refocused-INEPT transfer \cite{morris79,burum80} %[G. Morris and R. Freeman, J. Am. Chem. Soc. 101, 760 (1979);
-was used for transferring the polarization from $^1$H nuclei to the covalently bond $^{13}$C nuclei with the delays $\tau_1 = 1.94$~ms and $\tau_2 = 0.97$~ms
-(set as multiples of the MAS rotation period). The pulses for the refocused-INEPT sequence had a nutation frequency equal to 63.45~kHz.
+\todo{ I could not find this reference [D. Burum, R. Ernst, J. Mag. Res. 39, 163 (1980)]}
+was used for transferring the polarization from $^1$H nuclei to the covalently bond $^{13}$C nuclei with the delays $\tau_1$ = 1.94 ms and $\tau_2$ = 0.97 ms
+(set as multiples of the MAS rotation period). The pulses for the refocused-INEPT sequence had a nutation frequency equal to 63.45 kHz.
 For acquiring the spectra, a total number of 1024 transients were recorded for each point in the indirect dimension, using an acquisition
-time of 0.1~s under SPINAL64 $^1$H decoupling \cite{fung00} with a nutation frequency of 50~kHz, and with dwell time giving a spectral
+time of 0.1 s under SPINAL64 $^1$H decoupling \cite{fung00} with a nutation frequency of 50 kHz, and with dwell time giving a spectral
 width of 200 ppm for the $^{13}$C chemical shift dimension. The set of free induction decays recorded were then processed as
-described e.g. in Ref.~\citenum{ferreira08}. The dipolar splittings in the crowded spectral region
+described e.g. in Ref. \citenum{ferreira08}. The dipolar splittings in the crowded spectral region
 between 29 and 30 ppm were assigned based on the previous assignments for POPC \cite{ferreira13} 
@@
 \caption{  List of MD simulations reporting their respective
-           lipid composition,
-           used counterions,
            added buffer concentrations,
@@
            and the deployed models.
-           The fatty acid chains for all lipids are palmitoyl (sn-1) and oleoyl (sn-2).
+           The fatty acid chains for all lipids are plamitoyl (sn-1) and oleoyl (sn-2).
          }\label{tbl:sim-list}
-\begin{tabular}{r  l | r r r r | r r | c c }
-   \multicolumn{1}{r}{PC:PS} & \multicolumn{1}{l}{ }  &  \multicolumn{4}{c}{added buffer conc. / mM}    & \multicolumn{2}{c}{no. molecules} &  \multicolumn{2}{c}{simulated with}  \\
-   ratio & counterion &  \ce{K^+}  &  \ce{Na^+} & \ce{Ca^{2+}} & \ce{Cl^-}      & \ce{H2O} &  PC:PS                 &  ECC-lipids  &  Lipid17    \\
+\begin{tabular}{l | r r r r | r r | c c }
+                          &  \multicolumn{4}{c}{added buffer concentration / mM}    & \multicolumn{2}{c}{no. molecules} &  \multicolumn{2}{c}{simulated with model}  \\
+PC:PS (ratio)/counterions &  \ce{K^+}  &  \ce{Na^+} & \ce{Ca^{2+}} & \ce{Cl^-}      & \ce{H2O} &  PC:PS                 &  ECC-lipids  &  Lipid17    \\
   \hline
-only PS & \ce{K^+}   &      0  &      0  &      0  &      0  &  3600  &  0:72  &  \textbullet &  \textbullet \\ 
-only PS & \ce{Na^+}  &      0  &      0  &      0  &      0  &  3600  &  0:72  &  \textbullet &  \textbullet \\ 
+PC:PS (0:1)/K+   &      0  &      0  &      0  &      0  &  3600  &  0:72  &  X  &  X  \\ 
+PC:PS (0:1)/Na+  &      0  &      0  &      0  &      0  &  3600  &  0:72  &  X  &  X  \\ 
   \hline
-1:1 & \ce{K^+}   &      0  &      0  &      0  &      0  &  5253  &  64:64  &  \textbullet &  \textbullet \\ 
-1:1 & \ce{Na^+}  &      0  &      0  &      0  &      0  &  5253  &  64:64  &  \textbullet &  \textbullet \\ 
+PC:PS (10:1)/K+  &      0  &      0  &      0  &      0  &  3600  &  60:6  &  X  &  -  \\ 
+PC:PS (10:1)/Na+ &      0  &      0  &      0  &      0  &  3600  &  60:6  &  X  &  -  \\ 
   \hline
-4:1 & \ce{K^+}   &      0  &      0  &      0  &      0  &  3600  &  48:12  &  \textbullet &  -  \\ 
-4:1 & \ce{Na^+}  &      0  &      0  &      0  &      0  &  3600  &  48:12  &  \textbullet &  -  \\ 
+PC:PS (1:1)/K+   &      0  &      0  &      0  &      0  &  5253  &  64:64  &  X  &  X  \\ 
+PC:PS (1:1)/Na+  &      0  &      0  &      0  &      0  &  5253  &  64:64  &  X  &  X  \\ 
   \hline
-5:1 & \ce{K^+}   &      0  &      0  &      0  &      0  &  3600  &  60:12  &  \textbullet &  \textbullet \\ 
-5:1 & \ce{Na^+}  &      0  &      0  &      0  &      0  &  3600  &  60:12  &  \textbullet &  \textbullet \\ 
-5:1 & \ce{Na^+}  &      0  &      0  &     78  &     78  &  3561  &  60:12  &  \textbullet &  \textbullet \\ 
-5:1 & \ce{Na^+}  &      0  &      0  &    125  &    125  &  3561  &  60:12  &  \textbullet &  \textbullet \\ 
-5:1 & \ce{Na^+}  &      0  &      0  &    202  &    202  &  3561  &  60:12  &  \textbullet &  \textbullet \\ 
-5:1 & \ce{Na^+}  &      0  &      0  &    409  &    409  &  3522  &  60:12  &  \textbullet &  \textbullet \\ 
-5:1 & \ce{Na^+}  &      0  &      0  &    621  &    621  &  3483  &  60:12  &  \textbullet &  \textbullet \\ 
-5:1 & \ce{Na^+}  &      0  &    621  &      0  &    621  &  3483  &  60:12  &  \textbullet &  \textbullet \\ 
-5:1 & \ce{Na^+}  &      0  &   1510  &      0  &   1510  &  3377  &  60:12  &  \textbullet &  \textbullet \\ 
-5:1 & \ce{Na^+}  &      0  &   3002  &      0  &   3002  &  3213  &  60:12  &  \textbullet &  \textbullet \\ 
-5:1 & \ce{Na^+}  &    621  &      0  &      0  &    621  &  3483  &  60:12  &  \textbullet &  \textbullet \\ 
-5:1 & \ce{Na^+}  &   1510  &      0  &      0  &   1510  &  3377  &  60:12  &  \textbullet &  \textbullet \\ 
-5:1 & \ce{Na^+}  &   3002  &      0  &      0  &   3002  &  3213  &  60:12  &  \textbullet &  \textbullet \\ 
+PC:PS (4:1)/K+   &      0  &      0  &      0  &      0  &  3600  &  48:12  &  X  &  -  \\ 
+PC:PS (4:1)/Na+  &      0  &      0  &      0  &      0  &  3600  &  48:12  &  X  &  -  \\ 
   \hline
-10:1 & \ce{K^+}  &      0  &      0  &      0  &      0  &  3600  &  60:6  &  \textbullet &  -  \\ 
-10:1 & \ce{Na^+} &      0  &      0  &      0  &      0  &  3600  &  60:6  &  \textbullet &  -  \\ 
+PC:PS (5:1)/K+   &      0  &      0  &      0  &      0  &  3600  &  60:12  &  X  &  X  \\ 
+PC:PS (5:1)/Na+  &      0  &      0  &      0  &      0  &  3600  &  60:12  &  X  &  X  \\ 
+PC:PS (5:1)/Na+  &      0  &      0  &     78  &     78  &  3561  &  60:12  &  X  &  X  \\ 
+PC:PS (5:1)/Na+  &      0  &      0  &    125  &    125  &  3561  &  60:12  &  X  &  X  \\ 
+PC:PS (5:1)/Na+  &      0  &      0  &    202  &    202  &  3561  &  60:12  &  X  &  X  \\ 
+PC:PS (5:1)/Na+  &      0  &      0  &    409  &    409  &  3522  &  60:12  &  X  &  X  \\ 
+PC:PS (5:1)/Na+  &      0  &      0  &    621  &    621  &  3483  &  60:12  &  X  &  X  \\ 
+PC:PS (5:1)/Na+  &      0  &    621  &      0  &    621  &  3483  &  60:12  &  X  &  X  \\ 
+PC:PS (5:1)/Na+  &      0  &   1510  &      0  &   1510  &  3377  &  60:12  &  X  &  X  \\ 
+PC:PS (5:1)/Na+  &      0  &   3002  &      0  &   3002  &  3213  &  60:12  &  X  &  X  \\ 
+PC:PS (5:1)/Na+  &    621  &      0  &      0  &    621  &  3483  &  60:12  &  X  &  X  \\ 
+PC:PS (5:1)/Na+  &   1510  &      0  &      0  &   1510  &  3377  &  60:12  &  X  &  X  \\ 
+PC:PS (5:1)/Na+  &   3002  &      0  &      0  &   3002  &  3213  &  60:12  &  X  &  X  \\ 
+  \hline
 \end{tabular}
@@
 the default Amber ion parameters \cite{aqvist90} led to the artificial clustering of ions in solution \cite{NMRlipidsIV}.
-In the ECC-lipids simulations,
+In the ECC-lipid simulations,
 the ECC-POPC parameters~\cite{melcr18} (available from Ref. \citenum{ECC-POPC_nacl_cacl2_files}), 
@@
     equilibration time  & 50~ns  \\
-    total simulation time     & $\geq 1 \mu$s  \\
+    total simulation time     & 1~$\mu$s  \\
     temperature         & 298~K       \\
@@
 
-The C--H bond order parameters were calculated directly from Eq.~\ref{OPequation}.
-The time average was first calculated for each lipid, and
-the standard error of the mean from these individual values was then used as the
+The C--H bond order parameters were calculated directly from Eq. \ref{OPequation}.
+Average over the time for individual lipids was calculated first and
+the standard error of the mean of the average over lipids was then used as the
 error estimate \cite{botan15,ollila16,NMRlipidsIV}.
 Python program ({\tt scripts/calcOrderParameters.py}) that uses the
-MDAnalysis library \cite{agrawal11,gowers16} is available in Ref.~\citenum{MATCHgit}. 
+MDAnalysis library \cite{agrawal11,gowers16} is available in Ref. \citenum{MATCHgit}. 
 The ion number density profiles were calculated using the {\tt gmx density} tool
@@
 
+%The simulated system consisted of lipids (POPC and POPS) forming a single lipid bilayer,
+%an amount of water molecules comparable to the hydration levels in experiments, 
+%\todo{Check this claim. SAMULI: Check discussin in the NMRlipids IV manuscript}
+%\todo{EDIT: Taken from NMRlipids4 as an interesting information, to be edited.}
+%In the experimental $^2$H NMR literature data used in this work~\cite{browning80,roux90},
+%the lipids were first soluted to the buffer and then centrifuged to a pellet which
+%was used in the measurements. Such samples have lower lipid concentration
+%(approximately 10~wt~\% of lipids~\cite{browning80,roux88,roux90}) than 
+%gravimetric samples (60~wt~\%) and simulations (approximately 50-60~wt~\%) in this work.
+%Larger multilamellar repeat distances are expected in the samples with lower lipid
+%concentrations due to the swelling caused by electrostatic repulsion in pure PS lipid systems~\cite{millman82}.
+%However, the PS headgroup order parameters measured from gravimetric sample in this work
+%are in good agreement with the results from centrifuged sample in the literature \cite{browning80} (Fig. \ref{HGorderParameters}).
+%Furthermore, the equlibrium repeat distance rapidly decreases with the addtion of monovalent
+%salts and is close to the simulation box sizes already above 500~mM concentrations~\cite{millman82,rand89}.
+%Therefore, the hydration levels of multilamellae are expected to be sufficiently similar
+%in the used simulations and reference experiments.
+
 % ++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
@@
 
-\subsection{ECC-POPS improves agreement in experimental structural parameters of a pure POPS bilayer }
+\subsection{Structural parameters of a pure ECC-POPS model membrane: Agreement with experiments}
 
@@
 The X-ray scattering form factors and C-H bond order parameters in
-the headgroup and glycerol backbone region of a POPS lipid bilayer are already available in the literature \cite{kucerka14,NMRlipidsIV}.
+the headgroup and glycerol backbone region of POPS lipid bilayer are already available in the literature \cite{kucerka14,NMRlipidsIV}.
 Here, we measure also the C--H bond order parameters of the acyl chain region using natural abundance $^{13}$C NMR.
-We first determined the order parameters of each peak corresponding to different C--H bonds detected
-in the first dimension (Fig.~\ref{R-PDLF}) from the dipolar slices in the indirect dimension (Fig.~\ref{R-PDLFslices}).
+We first determined the order parameters of each peak corresponding different C--H bonds detected
+in the first dimension (Fig. \ref{R-PDLF}) from the dipolar slices in the indirect dimension (Fig. \ref{R-PDLFslices}).
 Because the exact assignment of these peaks was not possible, we assigned the
-acyl chain order parameters to mimic the profile predicted by the MD simulations (Fig.~\ref{simVSexpNOions_POPS}).
+acyl chain order parameters to mimic the profile predicted by the MD simulations (Fig. \ref{simVSexpNOions_POPS}).
 \todo{Tiago has to check and update this and figures in SI.}
@@
   \includegraphics[width=\figwidth]{../img/ecc_pops/Order-parameters_form-factors_exp-L17-ECC-lipids.pdf}
-  \includegraphics[width=\figwidth]{../img/POPSstructure.pdf} 
+  \includegraphics[width=\figwidth]{../img/ecc_pops/pops_chemfig.pdf} 
 \hfill
@@
     \textbf{Top:} X-ray scattering form factors from simulations with Lipid17/Dang \citep{lipid17-future, dang2006} and 
-    ECC-POPS/ECC-ions \cite{martinek17, Pluhackova2016} compared with experiments~\citep{kucerka14} at 298~K. 
+    the ECC-POPS/ECC-ions \cite{martinek17, Pluhackova2016} compared with experiments~\citep{kucerka14} at 298~K. 
     \textbf{Middle:} Order parameters of POPS head group, glycerol backbone and acyl chains  
-    from the same simulations 
+    from simulations with the Lipid17 \citep{lipid17-future} and the ECC-POPS models 
     compared with experiments at 298~K. \citep{NMRlipidsIV}
@@
   }  
+  \todo{J: flip the chemical structure so that it agrees with the left-to-right reading of the plot.} \\
 \end{figure}
 
-The Lipid17/Dang simulation gives larger acyl chain order parameters and smaller area per molecule than the
-experimental values (Fig.~\ref{simVSexpNOions_POPS} and table~\ref{tab:apls}),
-indicating that this simulation predicts a too compact POPS lipid bilayer. This is also in line with
-the discrepancy observed in the X-ray form factor between the simulation and experiment (Fig.~\ref{simVSexpNOions_POPS}).
+The Lipid17/Dang simulation give larger acyl chain order parameters and smaller area per molecule than the
+experimental values (Fig. \ref{simVSexpNOions_POPS} and table \ref{tab:apls}),
+indicating that this simulation predicts too dense POPS lipid bilayer. This is also in line with
+the discrepancy observed in the X-ray form factor between this simulation and experiment (Fig. \ref{simVSexpNOions_POPS}).
 In previous study \cite{NMRlipidsIV}, the Lipid17 force field with {\AA}qvist ions~\cite{aqvist90}
-gave a larger area per molecule (57~\AA$^2$) for POPS being closer to the experimental value,
-which is probably due to the weaker counterion binding affinity of {\AA}qvist ions to phospholipids. \cite{catte16, NMRlipidsIV}
-However,  {\AA}qvist ions produce known artifacts, 
-e.g. they exhibit artificial aggregation of ions at larger salt concentrations \cite{kohagen16, chen07, NMRlipidsIV}.
-Therefore, the Dang ions with more realisitic bulk properties are used here to analyze the effect of ECC to the properties of a POPS lipid bilayer. 
+gave a larger area per molecule (57 \AA$^2$) for POPS being closer to the experimental value,
+which is probably due to the weaker counterion binding affinity of {\AA}qvist ions.
+However,  {\AA}qvist ions exhibit artificial aggregation of ions at larger salt concentrations \cite{kohagen16, chen07, NMRlipidsIV}.
+Therefore, the Dang ions with more realisitic bulk properties are used here to analyze the effect of ECC to the properties of POPS lipid bilayer. 
 The ECC-POPS simulation gives a good agreement with the X-ray scattering form factors and experimental
-acyl chain order parameters (see Fig.~\ref{simVSexpNOions_POPS}) indicating that the bilayer dimensions and acyl chain
+acyl chain order parameters (see Fig.~\ref{simVSexpNOions_POPS}), indicating the the bilayer dimensions and acyl chain
 conformations are well described by the force field. The area per lipid from ECC-POPS simulation is slightly
-smaller than the value reported from SDP model \cite{kucerka14} (table~\ref{tab:apls}), but the values
-agree within their error estimates.
+smaller than the value reported from SDP model \cite{kucerka14} (table \ref{tab:apls}), but the values
+are within each others error bars.
+
+% In order to obtain the area per lipid from experiments, modeling is used on top of the scattering factors \citep{kucerka14}. 
+%In ocntrast, this property is easily extracted directly from simulations with flat phospholipid bilayers. 
+%We compare the values from experiments and simulations in Table~\ref{tab:apls}. 
+%While the agreement between the scattering form factors from the simulation of a pure POPS bilayer and experiment 
+%are excellent (Fig.~\ref{simVSexpNOions_POPS}), there is a non-negligible difference between the values of the area per lipid in Table~\ref{tab:apls}. 
+%Since both values are derived from the scattering form factors through modeling of the electron density of the bilayer,
+%we cannot decide, which of the values is more reliable.
 
@@
 \centering
-  \caption{Area per lipid of POPS bilayers, $A_L$, from simulations with different force fields at 298~K with \ce{Na^+} counterions. \label{tab:apls} } 
+  \caption{Area per lipid (APL) of POPS bilayers from simulations with different force fields at 298~K with \ce{Na^+} counterions. \label{tab:apls} } 
   \begin{tabular}{l|c } 
     \multicolumn{2}{c}{POPS} \\
-    model          & $A_L$ / Å$^2$    \\ 
+    model          & APL / Å$^2$    \\ 
     \hline 
@@
   \end{tabular} \\
-\todo{J: This table will become a small inset plot in the top of Fig.~\ref{simVSexpNOions_POPS} }
 \end{table}
@@
 This is the case also for Lipid17 POPS simulations, where the headgroup and glycerol backbone
-order parameters are quite far from experimental values with both Dang (Fig.~\ref{simVSexpNOions_POPS})
-and {\AA}qvist ions. \cite{NMRlipidsIV} The values from ECC-POPS simulation are closer,
-but not in full agreement with experiments (Fig.~\ref{simVSexpNOions_POPS}).
+order parameters are quite far from experimental values with both Dang (Fig. \ref{simVSexpNOions_POPS})
+and {\AA}qvist \cite{NMRlipidsIV} ions. The values from ECC-POPS simulation are closer,
+but not in full agreement with experiments (Fig. \ref{simVSexpNOions_POPS}).
 Especially, the experimentally observed more negative values of the $\beta$-carbon
@@
 same carbon \cite{ollila16}) of the $\alpha$-carbon in POPS than in POPC \cite{NMRlipidsIV}
-are not reproduced by neither of the models here. 
-These differences are previously related
-to a more rigid structure of PS headgroup \cite{browning80,buldt81} and 
-the stiffness of the C$_\alpha$-C$_\beta$-C$_\gamma$-O$_\gamma$ dihedral \cite{NMRlipidsIV}
-plotted in Fig.~\ref{fig:dihedral}. 
-Interestingely, this dihedral angle is more flexible in ECC-POPS compared to Lipid17 
-suggesting that its rigidity is overestimated in Lipid17 due to the lack of electronic polarizability,
-while the same dihedral parameters yield too small forking in ECC-POPS.
-
-Overall,
-the head group and acyl chain order parameters within ECC-lipids
-are in a good agreement with the experimental values 
-as shown in Fig.~\ref{simVSexpNOions_POPS}. 
+are not reproduced by neither of the models here. These differences are previously related
+to a more rigid structure of PS headgroup and the stiffness of the C$_\alpha$-C$_\beta$-C$_\gamma$-O$_\gamma$
+dihedral, which is \todo{the dihedral should be plotted and sentence finished.}
+
+%The head group order parameters $\alpha$ and $\beta$ are highly relevant for this work,
+%as they are being used in the electrometer concept \cite{altenbach84, catte16, melcr18}.
+%For POPC in pure water, the order parameter $\beta$ agrees well with the experiment, while the order parameter $\alpha$ is somewhat lower.
+
+In conclusion, the ECC-POPS model reproduces the lipid bilayer dimensions and
+acyl chain conformations in good agreement with experiments, but has room for
+improvement in the glycerol backbone and headgroup region. Therefore,
+its structural accuracy is in the same level as in other state-of-the-art
+lipid force fields \cite{botan15, ollila16, Pluhackova2016, NMRlipidsIV}.
+
+\subsection{Counterion binding affinity and interactions between PC and PS head groups}
+
+\begin{figure}[tbp!] 
+  \centering 
+  \includegraphics[height=\figheight]{../img/ecc_pops/density_profiles_na-k-counterions_wat_phos_compar_purePOPS_ecclipids-lipid17.pdf}
+  \includegraphics[height=\figheight]{../img/ecc_pops/density_profiles_na-k-counterions_wat_phos_compar_5PC-1PS_ecclipids-lipid17.pdf}
+  \caption{\label{fig:POPS-counterions-dens}
+    Number density profiles of \ce{K^{+}} and \ce{Na^{+}} counterions along the membrane normal axis
+    in ECC-lipid (solid lines) and Lipid17/Dang (dashed lines) simulations of pure POPS (left) or POPC:POPS (5:1, right) bilayers.  
+    The density profiles of phosphate groups and water are divided by 4 and 100, respectively.  
+}
+\end{figure} 
+Binding affinity of monovalent ions to lipid bilayers depends strongly on
+force field parameters \cite{catte16,NMRlipidsIV},
+which is the case also here as the sodium counterion binding affinity to
+POPS lipid bilayer is stronger in Lipid17/ Dang simulation than in
+ECC-POPS simulation (Fig. \ref{fig:POPS-counterions-dens}).
+The cation binding affinities to lipid bilayers in simulations have been previously evaluated
+against experiments using the order parameters of $\alpha$ and $\beta$ carbons in the
+phosphatidylcholine (PC) lipid headgroup (see Fig.~\ref{fig:chemstruct_pc_ps} for the labeling) \cite{catte16,melcr18,NMRlipidsIV},
+which decrease proportionally to the bound charge according to the ''electrometer concept'' \citep{seelig87}.
+Althought the response of PS lipid headgroup order parameters to the bound charge is also systematic,
+it is less well undestood and the addition of cations may cause phase transitions in negatively charged
+bilayers \cite{feigenson86,mattai89,roux91,roux90}. Therefore, the cation binding affinity to bilayers with
+PS lipids has been evaluated measuring the PC headgroup order parameters from POPC:POPS (5:1)
+mixtures\cite{roux90,NMRlipidsIV}.
+Following our previous work \cite{NMRlipidsIV}, we evaluate the counterion binding affinity
+to lipid bilayer in simulations containing PS using two different approaches.
+
+% The situation is more complicated for lipid bilayers containing
+%PS lipids, for which there is no ion-free reference state due to the presence of counterions.
+
+%Unlinke in pure PC bilayers,
+%Interpreting the changes of the head group order parameters of 
+%mixed PC:PS lipid bilayers 
+%from both simulations and experiments
+%is complicated by the presence of counterions. \cite{NMRlipidsIV}
+
+\begin{figure}[tb!] 
+  \centering 
+  \includegraphics[width=\figwidth]{../Fig/lipids_chemfig_POPC_POPS.pdf} 
+  \caption{ \label{fig:chemstruct_pc_ps} 
+            The chemical structure of POPC and POPS 
+            including the labeling of the carbon segments
+            used in the measurements of order parameteres. 
+          }
+          \todo{The POPS structure and labeling is already in figure \ref{simVSexpNOions_POPS}.
+            The POPC structure and labeling could be added on top of the new version of figure \ref{fig:delta_ordPar_NaCl_PC-PS_mix} with only PC,
+            where PC results occur in the first time, and this figure could be removed.} \\
+          \todo{$\gamma$ carbon of POPS is the COO- minus carbon. The $\gamma$ label should be moved there.}
+\end{figure}
+
+The head group and acyl chain order parameters within ECC-lipids
+are in general in a good agreement with the experimental values 
+as shown in Fig. \ref{simVSexpNOions_POPS}. 
 The acyl chain order parameters in particular agree with the experimental data within error bars.
@@
 than Lipid17 with either Dang \cite{smith94,chang1999,dang2006} 
-or {\AA}qvist ions \cite{aqvist90}. 
-Although the ECC-POPS model reproduces the lipid bilayer dimensions and
-acyl chain conformations in a good agreement with experiments, 
-there is a room for improvements especially in the glycerol backbone and headgroup region,
-e.g. in the parameters of the associated dihedral angles. 
-\cite{botan15, ollila16, Pluhackova2016, NMRlipidsIV}
-
-\subsection{Binding of counterions to POPS and POPC and interactions between their head groups}
-
-In simulations, binding affinity of monovalent ions to lipid bilayers depends strongly on
-force field parameters \cite{catte16,NMRlipidsIV}.
-% J: I think the following sentence can be omitted
-%For instance in the present work the affinity of sodium counterions to a POPS bilayer 
-%is stronger in Lipid17/Dang simulations 
-%than in simulations with ECC-POPS (Fig.~\ref{fig:POPS-counterions-dens} and~\ref{fig:delta_ordPar_NaCl_PC-PS_mix_PC}).
-The cation binding affinities to lipid bilayers in simulations have been previously evaluated
-against experiments using the order parameters of $\alpha$ and $\beta$ carbons in the
-phosphatidylcholine (PC) lipid headgroup (see Fig.~\ref{fig:delta_ordPar_NaCl_PC-PS_mix_PC} for the labeling) \cite{catte16,melcr18,NMRlipidsIV},
-which decrease proportionally to the bound charge according to the ''electrometer concept'' \citep{seelig87}.
-Although the response of PS lipid headgroup order parameters to the bound charge is also systematic,
-it is less well undestood and the addition of cations may cause phase transitions in negatively charged
-bilayers \cite{feigenson86,mattai89,roux91,roux90}. Therefore, the cation binding affinity to bilayers with
-PS lipids has been evaluated measuring the PC headgroup order parameters from POPC:POPS (5:1)
-mixtures\cite{roux90,NMRlipidsIV}.
-Following our previous work \cite{NMRlipidsIV}, we evaluate the counterion binding affinity
-to a lipid bilayer containing PS in simulations using two approaches.
+or ff99 ions \citet{aqvist90}.
+
+%This was demonstrated for a bound positive charge in simulations with POPC and 
+%cationic surfactants or aquaeous cations employing the ECC-POPC model \cite{melcr18}.
+%In this work, we will use the electrometer concept also for a bound negative charge 
+%using PS lipids as negatively charged surfactants. \cite{roux90}
+
+%is the
+%is not generally agreed on between simulations 
+%\citep{bockmann03,sachs04,berkowitz06,cordomi09}
+%and experiments 
+%\citep{cevc90,tocanne90,hauser76,herbette84,uhrikova08}.
+%We attempted to find a model, 
+%which would successfully interpret the experiments on binding of \ce{Na^+}
+%to neutral and negatively charged phospholipid bilayers \citep{akutsu81, roux90} 
+%in our works \citep{catte16, NMRlipidsIV}.
+%After testing the currently available force fields for phospholipids,
+%we concluded
+%that the interactions are generally overestimated in magnitude in almost all models 
+%but Lipid14 (PC) \citep{dickson14}, resp. Lipid17 (PS) \citep{lipid17-future}. 
+%This model yields a semi-quantitative agreement with the experimentally measured small changes of the order parameters \cite{catte16}
+%when used with the model of ions by \citet{aqvist90}. 
+%However, when used with a more accurate model of ions by \citet{Pluharova2014, martinek17},
+%the model overestimates the binding affinity of \ce{Na^+}
+%measured with lipid electrometer concept. \citep{melcr18}
+%In total, these results suggest that improvements 
+%in the lipid parameters are required for more accurate interactions even with monovalent cations. 
+%\citep{catte16, melcr18, NMRlipidsIV}
 
@@
   \centering 
-  \includegraphics[height=\figheightsmall]{../img/ecc_pops/density_profiles_na-k-counterions_wat_phos_compar_5PC-1PS_ecclipids-lipid17.pdf}
   \includegraphics[height=\figheightsmall]{../Fig/order_parameters_changes_A-B_PC-PS_mix_POPC_nacl.pdf} 
   \includegraphics[height=\figheightsmall]{../Fig/order_parameters_changes_A-B_PC-PS_mix_POPC_kcl.pdf} 
-  \includegraphics[height=\figheightsmall]{../img/POPCstructure.pdf} 
   \caption{\label{fig:delta_ordPar_NaCl_PC-PS_mix_PC} 
-    \textbf{Left:} Number density profiles of \ce{K^{+}} and \ce{Na^{+}} counterions along the membrane normal axis
-    in ECC-lipids (solid lines) and Lipid17/Dang (dashed lines) simulations of POPC:POPS (5:1) bilayers.  
-    The density profiles of phosphate groups and water are divided by 4 and 100, respectively.  
-    \textbf{Middle:} The POPC head group order parameters and the P--N vector angle
+    The POPC head group order parameters and the P--N vector angle
     with respect to the membrane normal as a function of POPS content in a bilayer
-    from ECC-lipids and Lipid17/Dang simulations with \ce{Na^+} (left) and \ce{K^+} (right) counterions.
+    from ECC-lipid and Lipid17/Dang simulations with \ce{Na^+} (top) and \ce{K^+} (bottom) counterions.
     Experimental order parameter values are from Ref. \citenum{scherer87}
-    and the signs from Ref. \citenum{ferreira16} 
-    (only \ce{Na^+} counterions, but shown also in the right plots for \ce{K+}).
-    \textbf{Right:} Chemical structure and labelling of carbon segments of POPC. 
+    and the signs from Ref. \citenum{ferreira16} (only \ce{Na^+} counterions, but shown also in bottom plots).
   }
-\todo{J: Will manually improve this figure with panels A, B, C and D 
-and will erase the duplicate y-axis description 
-and will add ticks to the right side of the plotting area. }
 \end{figure*}
-
-First, we monitor the response of POPC headgroup order parameters to the increasing amount of POPS (Fig.~\ref{fig:delta_ordPar_NaCl_PC-PS_mix_PC}).
-According to the electrometer concept, the head group order parameters of PC lipids increase
-with the increasing amount of negatively charged PS lipids in experiments, because the headgroup P--N dipole
-tilts more parallel to the membrane interface \cite{seelig87,scherer87}.
-The trend is observed with both force fields in simulations with \ce{K+} counterions and
-in ECC-lipids simulation with \ce{Na+} counterions. 
-In contrast, the order parameters are unchanged or decrease 
-and the P--N vector tilts more perpendicular to the membrane interface.
-in Lipid17/Dang ion simulations with \ce{Na+} counterions.
-The result can be explained by the higher binding affinity of sodium in the
-Lipid17/Dang ion simulations (Fig~\ref{fig:delta_ordPar_NaCl_PC-PS_mix_PC}), which
-compensates the negative charge of PS lipids at the interface.
-This is corrected in ECC-lipids, 
-which only slightly underestimate the slopes of the order parameters 
-when \ce{Na+} counterions are used (Fig.~\ref{fig:delta_ordPar_NaCl_PC-PS_mix_PC}).
-
-Second, we monitor the response of PC headgroup order parameters to the additional
-amount of \ce{NaCl} and \ce{KCl} in POPC:POPS (5:1) mixtures.
-In experiments, the addition of K$^+$ ions led to a very modest decrease of the order parameters,
-while the decrease was more pronounced with Li$^+$ ions and was the strongest with divalent Ca$^{2+}$ and Mg$^{2+}$ ions \cite{roux90}.
-According to the electrometer concept, the decrease of PC lipid head group order parameters
-correlates with the amount of bound cations \cite{akutsu81,altenbach84,seelig87,roux90,catte16,NMRlipidsIV},
-thus the binding affinity of the ions to POPC:POPS (5:1) mixtures increase in order K$^{+}$ $<$ Li$^{+}$  $<$ Mg$^{2+}$  $<$ Ca$^{2+}$.
-The response of PC headgroup order parameters to the increasing amount of KCl is close to experiments
-and the change in P--N vector is negligible in both force fields (Fig.~\ref{fig:delta_ordPar_monoval_PCPS}), athough the change of the $\alpha$-carbon
-order parameter may be slightly overestimated by the Lipid17/Dang simulations. 
-This could be explained by the deeper penetration of \ce{K+} into the bilayer (Fig.~\ref{fig:delta_ordPar_NaCl_PC-PS_mix_PC}).
-
-Because experimental data with the additional amount of \ce{Na+} in POPC:POPS (5:1) mixture was not available,
-we compared the simulation results to the experimental data with the additional amounts of
-\ce{Li+} and \ce{K+} ions, which should give the lower and the upper bounds to the sodium binding
-affinity (Fig.~\ref{fig:delta_ordPar_monoval_PCPS}). The PC headgroup response to the additional
-amount of sodium is clearly overestimated by the Lipid17/Dang simulation, and
-the response in ECC-lipids simulations is probably too close to the experimental results for lithium.
-% The following conclusion is already in the previous sentence, so I commented it out
-%In conclusion, the results suggest that the sodium binding affinity (Fig. \ref{fig:POPS-counterions-dens}) and
-%the tilting of P--N vector more perpendicular to the membrane surface (Fig. \ref{fig:delta_ordPar_NaCl_PCPS})
-%are clearly overestimated in Lipid17/Dang simulations and slightly overestimated in ECC-lipids simulations,
-%while the binding affinity of potassium is better described by both force fields.
-
-%\subsection{Molecular interaction and binding affinities of K$^+$ and Na$^+$ cations to mixed POPC:POPS (5:1) membrane} 
-%\label{sec:K_Na_added}
-
-\begin{figure}[tbp!] 
-  \centering 
-  \includegraphics[height=\figheightsmall]{../img/ecc_pops/order_parameters_changes_ecc-lip_L14_A-B-PN-COO_POPC_kcl.pdf} 
-  \includegraphics[height=\figheightsmall]{../img/ecc_pops/order_parameters_changes_ecc-lip_L14_A-B-PN-COO_POPC_nacl.pdf} 
-  \includegraphics[height=\figheightsmall]{../img/ecc_pops/order_parameters_changes_ecc-lip_L14_A-B-PN-COO_POPS_kcl.pdf} 
-  \includegraphics[height=\figheightsmall]{../img/ecc_pops/order_parameters_changes_ecc-lip_L14_A-B-PN-COO_POPS_nacl.pdf} 
-  \caption{\label{fig:delta_ordPar_monoval_PCPS}
-    Changes of the head group order parameters, and the angles of P--N and \ce{C_\beta-C_\gamma} (stars) vectors 
-    with respect to the membrane normal of POPC (left) and POPS (right) in a POPC:POPS (5:1) bilayer 
-    as a function of \ce{KCl} and \ce{NaCl} concentration from ECC-lipids and Lipid17/Dang simulations 
-    compared with experimental values from Ref. \citenum{roux90} (signs from Ref. \citenum{ferreira16}) at 298 K.
-    The y-axis for the $\alpha$-carbon results of POPS (middle right) is transferred
-    with the same value for both order parameters such that the lower order
-    parameter value from pure POPS is at zero to correctly illustrate the significant forking.
-    Because experimental data with \ce{NaCl} is not available for POPC, the data for KCl and \ce{LiCl} (dashed line, left)
-    are shown as lower and upper bounds for the response to \ce{NaCl}, respectively.
-  }
-\todo{J: Will improve this figure with panels A, B, C and D 
-and put POPC together, POPS together 
-and remove the duplicate y-axis titles 
-and make a common legend. }
-\end{figure}
 
@@
     with respect to the membrane normal as a function of POPC content in a bilayer
-    from ECC-lipids and Lipid17/Dang simulations with \ce{Na^+} (left) and \ce{K^+} (right) counterions.
+    from ECC-lipid and Lipid17/Dang simulations with \ce{Na^+} (top) and \ce{K^+} (bottom) counterions.
     Experimental order parameter values and signs for pure POPS are from Ref. \citenum{NMRlipidsIV} and values
-    for POPC:POPS (5:1) mixture from Ref. \citenum{roux90} 
-    (only \ce{Na^+} counterions, but shown also in the right plots for \ce{K+}).
-    % No translation in y-axis done in this plot:
+    for POPC:POPS (5:1) mixture from Ref. \citenum{roux90} (only \ce{Na^+} counterions, but shown also in bottom plots).
     %The underestimated response of ECC-lipids with \ce{Na^+} counterions 
@@
   }
-\todo{J: I suggest to put this to SI. }
 \end{figure*}
 
+First, we monitor the response of POPC headgroup order parameters to the increasing amount of POPS (Fig.~\ref{fig:delta_ordPar_NaCl_PC-PS_mix}).
+According to the electrometer concept, the head group order parameters of PC lipids increase
+with the increasing amount of negatively charged PS lipids in experiments, because the headgroup P--N dipole
+tilts more parallel to the membrane interface \cite{seelig87,scherer87}.
+The trend is observed with both force fields in simulations with \ce{K+} counterions and
+in ECC-POPS simulation with \ce{Na+} counterions, but not in Lipid17/Dang ion simulations with \ce{Na+} counterions,
+where the order parameters are unchanged or decrease and the P--N vector tilts more perpendicular to the membrane interface.
+The result can be explained by the higher binding affinity of sodium in the
+Lipid17/Dang ion simulations (Fig~\ref{fig:POPS-counterions-dens}), which
+compensates the negative charge of PS lipids at the interface.
+
+Second, we monitor the response of PC headgroup order parameters to the additional
+amount of sodium and potassium in POPC:POPS (5:1) mixtures.
+In experiments, the addition of K$^+$ ions led to a very modest decrease of the order parameters,
+while the decrease was more pronounced with Li$^+$ ions and strongest with divalent Ca$^{2+}$ and Mg$^{2+}$ ions \cite{roux90}.
+According to the electrometer concept, the decrease of PC lipid head group order parameters
+correlates with the amount of bound cations \cite{akutsu81,altenbach84,seelig87,roux90,catte16,NMRlipidsIV},
+thus the binding affinity of the ions to POPC:POPS (5:1) mixtures increase in order K$^{+}$ $<$ Li$^{+}$  $<$ Mg$^{2+}$  $<$ Ca$^{2+}$.
+The response of PC headgroup order parameters to the increasing amount of KCl is close to experiments
+and the change in P--N vector is negligible in both force fields (Fig. \ref{fig:delta_ordPar_KCl_PCPS}), athought the change of the $\alpha$-carbon
+order parameter may be slightly overestimated by the Lipid17/Dang simulations. This could be
+explained by the slightly deeper penetration of potassium into the bilayer (Fig. \ref{fig:POPS-counterions-dens}).
+Because experimental data with the additional amount of sodium in POPC:POPS (5:1) mixture was not available,
+we compared the simulation results to the experimental data with the additional amounts of
+lithium and potassium ions, which should give the lower and upper bounds to the sodium binding
+affinity (Fig. \ref{fig:delta_ordPar_NaCl_PCPS}). The PC headgroup response to the additional
+amount of sodium is clearly overestimated by the Lipid17/Dang simulation and
+the response in ECC-POPS simulations is probably too close to the experimental results for lithium.
+In conclusion, the results suggest that the sodium binding affinity (Fig. \ref{fig:POPS-counterions-dens}) and
+the tilting of P--N vector more perpendicular to the membrane surface (Fig. \ref{fig:delta_ordPar_NaCl_PCPS})
+are clearly overestimated in Lipid17/Dang simulations and slightly overestimated in ECC-POPS simulations,
+while the binding affinity of potassium is better described by both force fields.
+
+%Generally improved behaviour 
+%of the POPC and POPS head group order parameters 
+%with \ce{NaCl} or \ce{KCl} concentrations 
+%was achieved through the combination of models 
+%ECC-lipids and ECC-ions \citep{martinek17, kohagen16, Pluharova2014}. 
+%Simulations with these models 
+%reveal a good agreement with the NMR experiments 
+%for both neutral and negatively charged membranes.
+%The results are plotted for \ce{NaCl} in Fig. \ref{fig:delta_ordPar_NaCl_PCPS}, 
+%and for \ce{KCl} in Fig. \ref{fig:delta_ordPar_KCl_PCPS}.
+
+%The order parameters of the PS head groups including their changes with the content of PC lipids
+%are challenging to capture correctly in classical MD simulation models. \cite{NMRlipidsIV}
+%From such a perspective,
+%the changes of the PS head group order parameters from ECC-lipids 
+%with \ce{Na+}, and also \ce{K+}, counterions 
+%agree comparably well to the experimental measurements
+%(Fig.~\ref{fig:delta_ordPar_NaCl_PC-PS_mix}). 
+%Interestingly enough, 
+%although the Lipid17 model performed well with \ce{K+} counterions 
+%for the PC head group order parameter changes,
+%it does not agree in the equivalent changes in PS head groups
+%pointing at non-negligible effects of even relatively weakly binding cations 
+%on the structure of negatively charged lipids.
+
+%Despite the great improvements in the overall structure of ECC-POPS 
+%and its response to PC lipids and counterions,
+%there is still a large room for improving it
+%to agree with the experimental measurements.
+
+%\subsection{Molecular interaction and binding affinities of K$^+$ and Na$^+$ cations to mixed POPC:POPS (5:1) membrane} 
+%\label{sec:K_Na_added}
+
+\begin{figure}[tbp!] 
+  \centering 
+  \includegraphics[height=\figheight]{../img/ecc_pops/order_parameters_changes_ecc-lip_L14_A-B-PN-COO_POPC_nacl.pdf} 
+  \includegraphics[height=\figheight]{../img/ecc_pops/order_parameters_changes_ecc-lip_L14_A-B-PN-COO_POPS_nacl.pdf} 
+  \caption{\label{fig:delta_ordPar_NaCl_PCPS} 
+    Changes of the head group order parameters, and the angles of P--N and C$_\beta$--C$_\gamma$ (stars) vectors
+    with respect to the membrane normal of POPC (left) and POPS (right) in a POPC:POPS (5:1) bilayer 
+    as a function of \ce{NaCl} concentration from ECC-lipid and Lipid17/Dang simulations at 298 K.
+    The y-axis for the $\alpha$-carbon results of POPS (middle right) is transferred
+    with the same value for both order parameters such that the lower order
+    parameter value from pure POPS is at zero to correctly illustrate the significant forking.
+    Because experimental data with \ce{NaCl} is not available for POPC, the data for KCl and \ce{LiCl} (dashed line, left)
+    are shown as lower and upper bounds for the response to \ce{NaCl}, respectively.
+    The magnitudes of experimental order parameters are from Ref. \citenum{roux90} and signs from Ref. \citenum{ferreira16}.
+    %which has a lower affinity to phospholipid bilayers compared to \ce{LiCl} \citep{roux90}. 
+    %The orientation of the \ce{COO^-} group is defined as 
+    %the connector from the $\beta$ carbon to the carbon in \ce{COO^-} (stars, bottom right). 
+  }
+  \end{figure}
+
+\begin{figure}[tbp!] 
+  \centering 
+  \includegraphics[height=\figheight]{../img/ecc_pops/order_parameters_changes_ecc-lip_L14_A-B-PN-COO_POPC_kcl.pdf} 
+  \includegraphics[height=\figheight]{../img/ecc_pops/order_parameters_changes_ecc-lip_L14_A-B-PN-COO_POPS_kcl.pdf} 
+  \caption{\label{fig:delta_ordPar_KCl_PCPS}
+    Changes of the head group order parameters, and the angles of P--N and C$_\beta$--C$_\gamma$ (stars) vectors 
+    with respect to the membrane normal of POPC (left) and POPS (right) in a POPC:POPS (5:1) bilayer 
+    as a function of \ce{KCl} concentration from ECC-lipid and Lipid17/Dang simulations 
+    compared with experimental values from Ref. \citenum{roux90} (signs from Ref. \citenum{ferreira16}) at 298 K.
+    The y-axis for the $\alpha$-carbon results of POPS (middle right) is transferred
+    with the same value for both order parameters such that the lower order
+    parameter value from pure POPS is at zero to correctly illustrate the significant forking.
+  }
+\end{figure}
+
+% This§: 
+% Other models do not yield good reponse, but
+% ECC-lipid describe the changes well
+% => Elect. polar. necessarry.
+% 
 The changes of POPS headgroup order parameters are almost negligible in experiments with increasing POPC fraction in bilayer
 or monovalent ion concentration in solvent
-(Figs.~\ref{fig:delta_ordPar_monoval_PCPS},~\ref{fig:delta_ordPar_monoval_PCPS} and~\ref{fig:delta_ordPar_NaCl_PC-PS_mix}).
+(Figs. \ref{fig:delta_ordPar_NaCl_PCPS}, \ref{fig:delta_ordPar_KCl_PCPS} and \ref{fig:delta_ordPar_NaCl_PC-PS_mix}).
 %while calcium leads to relatively large changes with low concentrations followed by a rapid saturation below 100mM (Fig. \ref{fig:delta_ordPar_CaCl_PCPS}) \cite{scherer87,roux90}.
@@
 are overestimated, which is also related to the changes in the orientations of the P--N and C$_{\beta}$--C$_{\gamma}$
-vectors (Fig.~\ref{fig:delta_ordPar_NaCl_PC-PS_mix_PC}). 
-%These structural changes can relate to variations in the interactions between PC and PS head groups 
-%and also to the changing amount of counterions and PS in the composition. 
-Also, the PS headgroup order parameters from simulations with Lipid17/Dang model
-are too sensitive to an additional amount of \ce{KCl} and \ce{NaCl} 
-(Fig.~\ref{fig:delta_ordPar_monoval_PCPS}).
-The overestimation of the order parameter changes upon increasing amount of POPC in a bilayer 
-was also observed in previous simulations with other force fields \cite{NMRlipidsIV}, 
-but not in ECC-lipids simulation, 
-where the changes in the PS headgroup orientation are more modest (Fig.~\ref{fig:delta_ordPar_NaCl_PC-PS_mix_PC}).
-In particular, 
-the dependence of the PS headgroup order parameters from ECC-lipids simulations
-to increasing amount of POPC, \ce{KCl} or \ce{NaCl}
-is weaker and closer to experiments 
-(Fig.~\ref{fig:delta_ordPar_monoval_PCPS}).
+vectors (Fig. \ref{fig:delta_ordPar_NaCl_PC-PS_mix}). The overestimation of the order parameter changes
+upon increasing amount of POPC in a bilayer 
+was also observed in previous simulations with other force fields \cite{NMRlipidsIV}, but not in
+ECC-POPS simulation, where also the changes in the PS headgroup orientation are more modest (Fig. \ref{fig:delta_ordPar_NaCl_PC-PS_mix}).
+The POPS headgroup orientation and order parameters are more sensitive also to the additional amount
+of sodium and potassium in Lipid17/Dang simulations, while dependence is weaker and closer to experiments
+in ECC-POPS simulations (Figs. \ref{fig:delta_ordPar_NaCl_PCPS} and \ref{fig:delta_ordPar_KCl_PCPS}).
 %In contrast, ECC-lipids with ECC-ions capture the different response 
@@
 %
-The influence of negatively charged PS on the counterion binding affinity can be estimated 
-%by calculating the amounts of phospholipids per bound \ce{Ca^{2+}} cation, $\xi _{Ca^{2+}}$,
-by comparing the relative surface excesses with respect to water, $\Gamma ^{w} _{\rm ion}$,
+The influence of negatively charged PS on the counterion binding affinity can
+be estimated by comparing the relative surface excesses with respect to water, $\Gamma ^{w} _{\rm ion}$,
 between POPC:POPS (5:1) mixture and pure POPC \cite{melcr18}.
-%Threshold for counting bound \ce{Ca^{2+}} cations to a membrane
-%was set to $0.3\,\mathrm{nm}$ from oxygen atoms in lipids as in Ref.~\citenum{melcr18}.  
 %Such a quantity compares the adsorption of ions to the adsorption of water molecules at an interface 
 %without the necessity of defining a Gibbs dividing surface. \citep{melcr18, chattorajBOOK}
-The value of $\Gamma ^{w} _{\rm Na}=-0.11 \pm 0.01 \mathrm{nm}^{-2}$ for 1~M sodium in pure ECC-POPC
+The value of $\Gamma ^{w} _{\rm Na}=-0.11 \pm 0.01 \mathrm{nm}^{-2}$ for 1 M sodium in pure ECC-POPC
 simulation was reported in the previous work \cite{melcr18}.
-The presence of $\sim$17\% POPS in a POPC bilayer increases the
+The presence of $\sim$17\$ POPS in POPC bilayer increases the
 %relative surface excess of sodium with respect to water from negative
@@
 for 0.621 M sodium, while value for potassium remains negative,
-$\Gamma ^{w} _{\rm K}=-0.123 \pm 0.005 \mathrm{nm}^{-2}$ in POPC:POPS (5:1) mixture (Fig.~\ref{fig:nacl-dens_PCPS}).
+$\Gamma ^{w} _{\rm K}=-0.123 \pm 0.005 \mathrm{nm}^{-2}$ in POPC:POPS (5:1) mixture (Fig. \ref{fig:nacl-dens_PCPS}).
 As discussed above, the sodium binding to POPC:POPS (5:1) mixture may be slightly overestimated also in
-ECC-lipids simulations. The response of headgroup order parameters to the addition of NaCl are
+ECC-lipid simulations. The response of headgroup order parameters to the addition of NaCl are
 slightly overestimated also in ECC-POPC simulations (Fig 3 in Ref. \citenum{melcr18}),
@@
 
-\subsection{Molecular interaction and binding affinity of \ce{Ca^{2+}} cations to mixed POPC:POPS (5:1) membrane} 
+In conclusion, applying ECC to Amber Lipid17 POPS force field parameters improves its
+interactions with both monovalent ions and POPC headgroups when compared with the NMR order parameter data.
+However, further optimization may be needed to fully reproduce the interactions between lipids and
+sodium ions.
+
+\subsection{Molecular interaction and binding affinities of \ce{Ca^{2+}} cations to mixed POPC:POPS (5:1) membrane} 
 \label{section:lip-ion_ca}
@@
 which underestimated the headgroup order parameter response.
-In ECC-lipids simulations, the headgroup responses are in better agreement with experiments (Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}),
+In ECC-lipid simulations, the headgroup responses are in better agreement with experiments (Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}),
 indicating that the lower binding affinity than in Lipid17/Dang simulations is more realistic (Fig.~\ref{fig:cacl-dens_PCPS}).
@@
 %observed especially in negatively charged bilayers \cite{borle85,macdonald87,roux90},
-%may not be fully captured by the ECC-lipids simulations.
+%may not be fully captured by the ECC-lipid simulations.
 The increased calcium binding affinity to bilayers containing PS lipids is illustrated in
-ECC-lipids simulations by the increasing relative surface excess of calcium with respect to water from
+ECC-lipid simulations by the increasing relative surface excess of calcium with respect to water from
 $\Gamma^{w}_{Ca} = 0.06\mathrm{nm^{-2}}$ for pure POPC bilayer with 350 mM CaCl$_2$ \cite{melcr18}
@@
     with respect to the membrane normal of POPC (left) and POPS (right) in a POPC:POPS (5:1) bilayer 
-    as a function of \ce{CaCl} concentration from ECC-lipids and Lipid17/Dang simulations 
+    as a function of \ce{CaCl} concentration from ECC-lipid and Lipid17/Dang simulations 
     compared with experimental values from Ref. \citenum{roux90} (signs from Ref. \citenum{ferreira16}) at 298 K.
@@
 Experimentally measured headgroup order parameters of POPS exhibit a strong dependence
-on low CaCl$_2$ concentrations with a rapid saturation below 100~mM \cite{roux90}.
+on low CaCl$_2$ concentrations with a rapid saturation below 100 mM \cite{roux90}.
 Also these changes were overestimated by the Lipid17/Dang (also shown in Fig.~\ref{fig:delta_ordPar_CaCl_PCPS})
-and other tested force fields in our recent work \cite{NMRlipidsIV}, 
-but CHARMM36 with the NBfix correction for calcium
+and other tested force fields in our recent work \cite{NMRlipidsIV}, including CHARMM36 with the NBfix correction for calcium
 which underestimated the binding affinity.
-In ECC-lipids simulations, the changes of the PS headgroup order parameters are not overestimated,
+In ECC-lipid simulations, the changes of the PS headgroup order parameters are not overestimated,
 but the strong dependence on low concetrations of CaCl$_2$ is not fully reproduced (Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}).
-In addition to possibly suboptimal interactions between calcium ions and PS headgroup, the potential sources of this
+In addition to the inaccurate interactions between calcium ions and PS headgroup, the potential sources of this
 discrepancy include the above observed slight overestimation of \ce{Na^+} counterions and
@@
 
-In ECC-lipids simulations, the average orientation of P--N vectors of both POPC and POPS headgroups tilts more perpendicular to the membrane surface
+In ECC-lipid simulations, the average orientation of P--N vectors of both POPC and POPS headgroups tilts more perpendicular to the membrane surface
 by 11$^\circ$ and  3$^\circ$, respectively, upon addition of 620~mM CaCl$_2$ to POPC:POPS (5:1) mixture (Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}),
@@
 tilts to the opposite direction by $11^\circ$ in the same system.
-Such a largely diminished flexibility of the PS head group can be explained 
-by the simultaneous binding of calcium to the \ce{COO^-} group in PS and phosphate group in PC. 
+%the systematic tilting of P--N and C$_{\beta}$--C$_{\gamma}$ vectors 
+%to more perpendicular and parallel direction with respect to the membrane interface, respectively,
+%is observed upon addition of CaCl$_2$ (Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}).
 In Lipid17/Dang simulations, the changes in both headroup orientations and order parameters
 are larger and less systematic for the PS headgroup, especially with lower concentrations (Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}).
-ECC-lipids simulations yield smaller error bars and shorter residence times (Fig.~\ref{fig:hist_residence_times}) in
-than those typically observed in simulations with other force fields \cite{javanainen17,melcr18,NMRlipidsIV}.
-This suggests that the ECC accelerates the equlibration of ions at lipid bilayer interface.
-
-Our 1 $\mu$s simulations seems to be sufficiently long for the ECC-lipids simulations, because
+Together with the smaller error bars and shorter residence times (Fig.~\ref{fig:hist_residence_times}) in
+ECC-lipid simulations than in typically observed for other force fields \cite{javanainen17,melcr18,NMRlipidsIV},
+this suggests that the ECC accelerates the equlibration of ions at lipid bilayer interface.
+%
+%changes by $11^\circ$ from $62^\circ$ to $73^\circ$ (more towards the membrane interior)
+% at 620~mM buffer concentration of \ce{CaCl2}.
+% We have investigated this effect using the mean orientations of P-N and \ce{COO^-} vectors 
+%(defined and plotted in Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}). 
+%While the mean angle between the P-N vector of PC head group and the membrane normal
+%decreased by 11$^\circ$ from 64$^\circ$ to 55$^\circ$
+% at 620~mM buffer concentration of \ce{CaCl2},
+%the mean orientation of the P-N vector in the PS head group has 
+%decreased only by 3$^\circ$ from 67$^\circ$ to 64$^\circ$ 
+%in the same concentration range. 
+%Such a largely diminished flexibility 
+%may arise from the binding of calcium to the \ce{COO^-} group in PS. 
+%Whe have measured the mean orientation of the \ce{COO^-} group 
+%as the connector of the carbon atoms 
+%forming the bond between the group and the $\beta$-carbon of the phospholipid 
+%(Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}).
+%by the binding of calcium cations to the \ce{COO^-} group, 
+%which reorients towards the membrane interior and to the phosphate region, 
+%where calcium cations bind predominantly. 
+%
+%
+Our 1 $\mu$s simulations seems to be sufficiently long for the ECC-lipid simulations, because
 90\% of the calcium residence times are shorter than $60\,\mathrm{ns}$ for pure POPC bilayer
@@
 Interestingly, the calcium residence times are 3-4 times longer in POPC:POPS (5:1) mixture than in pure POPC.
-In ECC-lipids simulations, on average 41\% of the total population of bound calcium cations is in contact with PS lipids
-with 8\% bound only to them, even though the negatively charged membrane contains only 17\% of POPS (Table~\ref{tab:binding}).
+In ECC-lipid simulations, on average 41\% of the total population of bound calcium cations is in contact with PS lipids
+with 8\% bound only to them, even though the negatively charged membrane contains only 18\% of POPS (Table \ref{tab:binding}).
 Therefore, calcium ions bind more likely to negatively charged PS lipids than neutral PC lipid in a mixed bilayer, as expected.
+
+%are again closer to the experiments is not optimal,
+%the general trends and behaviour even at highly concentrated solutions of \ce{CaCl2} 
+%are well captured by the model. 
+%For example the order parameter $\beta$ in POPS 
+%adopts a saturated value in experiments and ECC-lipids,
+%but it is growing with the \ce{CaCl2} concentration in Lipid17.
+%In addition, 
+%the forking of the order parameters $\alpha$ 
+%is almost constant in the experiment and ECC-lipids,
+%but it varies greatly in th Lipid17.
+
+%The structural changes in POPS induced by increasing concentrations of \ce{CaCl2} 
+%are very challenging to interpret from both experiments \cite{roux90} 
+%and also from MD simulations \cite{NMRlipidsIV}. 
+%The changes of the order parameters $S^\alpha$ and $S^\beta$ in PS head groups from simulations 
+%are not found to follow the experimental profiles, 
+%nor any common trend, with the currently available models \cite{NMRlipidsIV}.
 
@@
 controversial because the results strongly depend on force fields parameters \cite{melcrova16,valentine18,hallock18}.
-In ECC-lipids simulation of POPC:POPS (5:1) mixture with 409 mM CaCl$_2$,
-the calcium ions bind more likely 
-to the carboxylate than to phosphate moiety of POPS headgroup (Fig.~\ref{fig:cacl_complexes}),
+In ECC-lipid simulation of POPC:POPS (5:1) mixture with 409 mM CaCl$_2$,
+the calcium ions bind twice more likely to the carboxylate than to phosphate moiety of POPS headgroup
+\todo{I am not sure if this is correct, because I do not understand the numbers in the table.} (Table \ref{tab:binding}),
 in line with the results without the NBfix corrections in CHARMM36 \cite{hallock18}.
-However, in CHARMM36 simulations without the NBfix correction for calcium \cite{kim16},
+However, in CHARMM36 simulations without the NBfix correction for calcium,
 calcium ions significantly overbind to PC lipids  \cite{catte16} and calcium ions coordinate
 with even four distinct PS lipids \cite{hallock18}, while coordination with more than two PS lipids
-was not observed in our ECC-lipids simulations (Fig.~\ref{fig:cacl_complexes}).
+was not observed in our ECC-lipid simulations (Fig.~\ref{fig:cacl_complexes}).
 Furthermore, the POPS:POPC:Ca$^{2+}$ (7:3:7) mixture with relatively high
-PS lipid fraction and calcium concentration was used in the previous work \cite{hallock18},
+PS lipid fraction and calcium concentration was used in previous work \cite{hallock18},
 which may lead to the complexation of \ce{Ca^{2+}} cations and PS lipids that is
-not observed with lower PS content used in ECC-lipids simulations in this work \cite{hauser77,kurland79,hauser85,feigenson86,mattai89,roux90,roux91}.
+not observed with lower PS content used in ECC-lipid simulations in this work \cite{hauser77,kurland79,hauser85,feigenson86,mattai89,roux90,roux91}.
 %While POPC interacts with the calcium cations almost entirely through its phosphate group 
@@
 %however, they are always accompanied by interactions with phosphate groups. 
-
 Significant coordination of calcium with carbonyls groups of PS and PC lipids observed in previous
 simulations with the Berger force field \cite{melcrova16} %\cite{berger97,mukhopadhyay04}
-was not observed in ECC-lipids simulations in here nor in the previous work with pure POPC \cite{melcr18}.
+was not observed in ECC-lipid simulations in here nor in the previous work with pure POPC \cite{melcr18}.
+%This is in contrast to 
+%combined with fluorescent and vibrational sum frequency spectroscopy,
+%which suggest a 
 Overestimated coordination of cations with carbonyl groups is a potential source for
@@
 %to yield converged results in this aspect (see SI and Refs.~\cite{catte16, melcr18, NMRlipidsIV}). 
+\begin{table}[tb!] 
+\centering
+\caption{%Bulk concentrations, $C _{Ca}$, and buffer concentrations, $C' _{Ca}$, of Ca$^{2+}$;
+  % relative surface excess of calcium with respect to water ($\Gamma_{Ca}^{\rm water}$); 
+  % and
+  Percentages of the population of bound Ca$^{2+}$ to various lipid moieties in a pure POPC bilayer with 350 mM CaCl$_2$
+  and in a POPC:POPS (5:1) mixture with 409 mM CaCl$_2$. The threshold for counting a contact was set to $0.3\,\mathrm{nm}$, which encompasses the
+  first peak of the radial distribution function between the cations and the oxygen atoms of the lipids. 
+  \label{tab:binding}} 
+\begin{tabular}{ l | c c } 
+        interacting moiety    &  \multicolumn{2}{c}{percentage of bound \ce{Ca^{2+}} } \\
+%	\hline
+%	$C _{Ca}\,/\,\mathrm{mM}$  &  $240\pm 10 $  &  $280\pm 10 $  \\
+%	$C'_{Ca}\,/\,\mathrm{mM}$  &  $400\pm 10 $  &  $350\pm 10 $  \\
+%	$\Gamma_{Ca}^{\rm water}\, / \,\mathrm{nm}^{-2}$  &  $0.24 \pm 0.01 $  &  $0.06 \pm 0.01 $  \\
+%	\hline
+%                             &  \multicolumn{2}{c}{ } \\
+                             &  5\,POPC:1\,POPS &  POPC   \\
+	\hline
+	     PC              &   59   &  100   \\
+	     PO$_4$    in PC &   41   &   67   \\
+	     carbonyls in PC &   <1   &   ~1   \\
+	\hline
+	     PS              &    8   &        \\ 
+	     PO$_4$  in PS   &    2   &        \\
+	     COO$^-$ in PS   &    4   &        \\
+	     carbonyls in PS &   <1   &        \\
+	\hline
+	both PC and PS       &   33   &        \\
+  \end{tabular} \\
+  \todo{I have commented out the surface excess values (available in text) and concentrations (now in caption) from the table.} \\
+  \todo{I do not understand the numbers for different moieties. Are these probabilities that the calcium binds only to these moieties and
+    nothing else? In total, 41 \% binds to PS lipids, but only 2 \% and 4 \% to phosphates and COO$^-$, respectively? Where do the rest bind?
+    I think that we should show the total probabilities to bind to different moieties.}
+\end{table}
 
@@
   \centering 
-  \includegraphics[width=\figwidthsmall]{../img/bound_ca_populations.pdf} 
-  \includegraphics[width=\figwidthsmall]{../img/stoichiometry_CaCl2_comparison_Ecc-lipids_PC-vs-PCPS.pdf} \\ 
-  \caption{ \label{fig:cacl_complexes} 
-  \textbf{Left:} Percentages of the population of bound Ca$^{2+}$ 
-in exclusive contact with various lipid moieties 
-in a POPC:POPS (5:1) bilayer with 409 mM CaCl$_2$. 
-The bar representing total populations of bound \ce{Ca^{2+}} to PC, PS, or PC and PS (blue),
-is in part covered by areas representing contributions of individual moieties exclusively (shades of orange and red). 
-In the case of mixed simultaneous binding to PC and PS
-we distinguish different moieties only in PS.
-%i.e., PC can interact with any considered moiety. 
-For comparison, in a pure POPC bilayer with 350 mM CaCl$_2$, 
-67\% of bound \ce{Ca^{2+}} is exclusively in contact with \ce{PO_4} group. \cite{melcr18} 
-    \textbf{Right:} Relative probabilities of \ce{Ca^{2+}} ions to coordinate with a certain number of lipids
+  \includegraphics[width=\figwidth]{../img/stoichiometry_CaCl2_comparison_Ecc-lipids_PC-vs-PCPS.pdf} \\ 
+  \caption{\label{fig:cacl_complexes} 
+    Relative probabilities of \ce{Ca^{2+}} ions to coordinate with a certain number of lipids
     in pure POPC bilayer with 350 mM CaCl$_2$ and in POPC:POPS (5:1) mixture with 400 mM CaCl$_2$.  
@@
 
+%Still, the dominant contribution to the binding of calcium comes from the phosphate groups. 
+%Interestingly enough,
+%although the \ce{COO^-} groups increase the affinity of calcium towards PS compared to PC,
+% there is no extra density at the outmost surface of the membrane. 
+%This can be understood from the point of the structural rearangements of the \ce{COO^-} group, 
+%which take place during the interactions with calcium cations
+%described above and in Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}.
+
+%Simulations using CHARMM36 force field \cite{klauda10,venable13} with
+%the NBfix parameters for calcium \cite{kim16}, combined with 2D infrared spectroscopy,
+%suggest that calcium ions interact only with the carboxylate group of PS lipids \cite{valentine18}.
+%The same force field without the NBfix parameters, combined with NMR chemical shifts and
+%rotational-echo double-resonance (REDOR) experiments, suggests a significant binding affinity also to the phosphate region \cite{hallock18}.
+%Simulations with the Berger force field \cite{berger97,mukhopadhyay04},
+%combined with fluorescent and vibrational sum frequency spectroscopy, suggest a significant
+%calcium binding also to the carbonyls in the acyl chains \cite{melcrova16}.
+
+%In PS lipids, the calcium binds both phosphate and carboxylate, but probability for the latter is twice higher.
+%In PC lipids, the calcium binds almost always to phosphate, but in some cases also to carbonyl, as also observed previously in pure POPC simulations \cite{melcr18}.
+
 The advantage of the NMR order parameters used in this work is the direct connection between
@@
 rotational-echo double-resonance (REDOR) experiments \cite{hallock18} or fluorescent and vibrational sum frequency spectroscopy \cite{melcrova16}.
-Because the ECC-lipids model gives the best agreement with the experimental headgroup order parameters
+Because the ECC-lipid model gives the best agreement with the experimental headgroup order parameters
 with various ion conctenrations among the available models, we believe that it gives the currently best available
-interpretation of the experimental data. Athough the ECC-lipids model gives the most realistic picture of calcium binding to
+interpretation of the experimental data. Athough the ECC-lipid model gives the most realistic picture of calcium binding to
 lipid bilayers containing PS lipids with respect to NMR order parameters, it does not
 correctly capture the rapid changes with small concentrations. Therefore, further
-force field optimization is required for even more accurate MD simulations of cations
+force field optimization is required for fully realistic MD simulations of cations
 in the vicinity of negatively charged lipid bilayers.
+
+%Despite significant improvements in the ECC-lipid against Lipid17/Dang
+%and other force fields for the response of PS headgroup order parameters to
+%the bound calcium \cite{NMRlipidsIV}, 
+%the model (Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}).
+
 % and also possibly from the approximate mean-field treatment of the electronic polarization.
+
+%
+% SAMULI: We have problems to find time to finish this paper even with the current data, we should not add anything that is not necessary to finish this asap. 
+%
+%\todo{JOE: suggestion -- what is the charge at the membrane interface of memb+counterions -vs- memb+CaCl2 concnetration?
+% this could relate well to the zeta potential measurements and the Fig 1 in \cite{melcrova16}. }
+% 
+%\todo{ THE REMAINING PIECES TO BE DISCUSSED  --  consider wheter the dynamics slowdown or the dihedral distributions could make good contributions to this paper. }
+
+%% Ca2+ induces much slower dynamics of lipids when PS is present,
+%% this does not happen with PC
+%% uses NBFIX
+%From Ref.~\cite{valentine18}://
+%To summarize, calcium binding to PS lipids slows the dynamics at the ester position and introduces more heterogeneous molecular environments. 
+%The slowdown cannot be attributed to complete ester dehydration or decreased hydrogen bonding to water, so we interpret these results as evidence that the lipid headgroups adopt more rigid but %disordered conformations on the molecular level.
+%The decrease in dynamics observed experimentally indicates that the network of hydrogen bonds between lipids and water at the polar-nonpolar interface becomes slower, as it was shown by Karatha%nou and Bondar (34).
+%This is bolstered by recent NMR work showing that binding to Ca2+ induces two different rigid conformations in the glycerol backbone of lipid headgroups (18).
+%We have investigated the molecular origin of the rigidification and decreased lateral diffusion, finding that calcium binding does not dehydrate the ester region, but even with only 10\% PS lip%ids, the local dynamics slow around the esters.
+The slower local dynamics provide a molecular explanation for the Ca2+-dependent decreases in fluidity observed on longer length and timescales.
+%There is considerable room for further improvement in the force fields for ion-membrane interactions.
+Particularly, small differences in short-range interactions may determine the local geometr%ies (34
+%).
+Most of our conclusions are, however, based on long-range effects (i.e., electric field generated by the ion at significant distances from the carbonyl groups), which are less sensitive to t%he details of the force field.
+
+%Their Figure 5 \cite{hallock18} shows the distributions of O-Ca-Cb-N dihedral angle
+%revealing separate populations of the PS head groups. 
+%These two structures are also compared in therm of P-Cg distances, 
+%which can be measured from NMR REDOR exp. 
+%It is concluded that since the two distinct populations in P-Cg distance can be represented with the 
+%dihedral distributions, 
+%they also likely represent the distinct populations. 
+% -> plotting the distribution of O-Ca-Cb-N dihedral angle may be also a good idea for us (SI?)
+%     -> NMRlipids 4 in SI \cite{NMRlipidsIV} plots this information: 
+%           - C36 agress with the distribution in \cite{hallock18}
+%           - Lipid17 (and Slipids) are almost identical to C36
+%           - but the structure of the COO- is highly different (dih angle Ca-Cb-Cg-Og)
 
@@
 details of calcium binding to PS headgroups.
-
-In conclusion, applying ECC to Amber Lipid17 POPS force field parameters improves its
-interactions with both monovalent ions and POPC headgroups when compared with the NMR order parameter data.
-However, further optimization may be needed to fully reproduce the interactions between lipids and
-sodium ions.
 

### 2019-05-22 Manuscript/manuscript.tex

--- 2019-05-21 Manuscript/manuscript.tex
@@
 the indirect dimension were recorded. These settings enable to record dipolar slices with a C--H bond order parameter
-resolution of $\pm 0.01$. The C--H bond order parameter determined from a given dipolar
-splitting, $\Delta \nu$ (see e.g. Fig.~\ref{R-PDLFslices} in SI)
-is equal to $\Delta \nu / (0.315 \times 21.5 \, \mathrm{kHz})$, where 0.315 is
+resolution of $\pm$0.01. The C--H bond order parameter determined from a given dipolar
+splitting, $\Delta\nu$ (see e.g. Fig. \ref{R-PDLFslices} in \todo{Is this the correct figure?}) is equal to $\Delta\nu/(0.315\times21.5 kHz)$, where 0.315 is
 the scaling factor of the R18 recoupling sequence and 21.5 kHz is the maximum $^1$H-$^{13}$C dipolar coupling for a C--H bond.
@@
   }
-\todo{J: Will adjust the Fig into a nice panel and change the captions accordingly. 
-Ad bar plots -- change total to combinations as described in caption, 
-add simultaneously to "PC and PS", 
-add ticks to the right side of the plot and
-change the colour of "combinations" from blue to any other so that it is different from the legend in the other plot. }
+\todo{J: Will adjust the Fig into a nice panel and change the captions accordingly. }
 \end{figure}

### 2019-05-22 Manuscript/manuscript.tex
Some update

@@
 Further optimization of parameters was not done in this work. 
+The generated ECC-POPS parameters are available from Ref.~\citenum{??}
+\todo{We need here a permanent citation with DOI.
+  Easiest would be to add the parameters into a one of the existing Zenodo repositories containing other simulation parameters.
+  We could also put a release of Git into Zenodo, cite that and specify the exact directory where parameters are found,
+  but this is more complicated for users to find.
+}
 

### 2019-05-23 Manuscript/manuscript.tex
update text

@@
 Here, we extend the ECC approach also to the negatively charged 1-palmitoyl-2-oleoyl-sn-glycero-3-phospho-L-serine (POPS) lipid.
-We also complement the available experimental data for the the force field quality
+We also complement the available experimental data for the force field quality
 evaluation by measuring the acyl chain C--H bond order parameters of POPS in bilayer using natural abundance $^{13}$C NMR.
@@
 Following our previous work for POPC, we use the scaling factor of $f_\sigma = 0.89$ for the Lennard-Jones~$\sigma$ parameters and
-use it to scale the parameters of atoms only in the headgroup, glycerol backbone, and carbonyl regions \cite{melcr18, ECC-POPC_nacl_cacl2_files}.
+use it to scale the parameters of atoms only in the headgroup, glycerol backbone, and carbonyl regions \cite{melcr18}.
 Further optimization of parameters was not done in this work. 
@@
 error estimate \cite{botan15,ollila16,NMRlipidsIV}.
-Python program ({\tt scripts/calcOrderParameters.py}) that uses the
-MDAnalysis library \cite{agrawal11,gowers16} is available in Ref.~\citenum{MATCHgit}. 
+Python program that uses the
+MDAnalysis library \cite{agrawal11,gowers16} is available in Ref.~\citenum{MATCHgit} ({\tt scripts/calcOrderParameters.py}). 
 The ion number density profiles were calculated using the {\tt gmx density} tool
@@
 
-The Lipid17/Dang simulation gives larger acyl chain order parameters and smaller area per molecule than the
-experimental values (Fig.~\ref{simVSexpNOions_POPS} and table~\ref{tab:apls}),
-indicating that this simulation predicts a too compact POPS lipid bilayer. This is also in line with
-the discrepancy observed in the X-ray form factor between the simulation and experiment (Fig.~\ref{simVSexpNOions_POPS}).
-In previous study \cite{NMRlipidsIV}, the Lipid17 force field with {\AA}qvist ions~\cite{aqvist90}
-gave a larger area per molecule (57~\AA$^2$) for POPS being closer to the experimental value,
-which is probably due to the weaker counterion binding affinity of {\AA}qvist ions to phospholipids. \cite{catte16, NMRlipidsIV}
-However,  {\AA}qvist ions produce known artifacts, 
-e.g. they exhibit artificial aggregation of ions at larger salt concentrations \cite{kohagen16, chen07, NMRlipidsIV}.
-Therefore, the Dang ions with more realisitic bulk properties are used here to analyze the effect of ECC to the properties of a POPS lipid bilayer. 
-The ECC-POPS simulation gives a good agreement with the X-ray scattering form factors and experimental
-acyl chain order parameters (see Fig.~\ref{simVSexpNOions_POPS}) indicating that the bilayer dimensions and acyl chain
+The Lipid17/Dang simulation gives 
+discrepancies with the experimental X-ray form factor,
+larger acyl chain order parameters,
+and smaller area per molecule than the experimental values (Fig.~\ref{simVSexpNOions_POPS} and table~\ref{tab:apls}),
+indicating that this simulation predicts a too compact POPS lipid bilayer. 
+The ECC-POPS simulation gives better agreement with experiments for the X-ray scattering form factors and 
+acyl chain order parameters (Fig.~\ref{simVSexpNOions_POPS}),
+indicating that the bilayer dimensions and acyl chain
 conformations are well described by the force field. The area per lipid from ECC-POPS simulation is slightly
 smaller than the value reported from SDP model \cite{kucerka14} (table~\ref{tab:apls}), but the values
-agree within their error estimates.
+agree within their error estimates. The larger area in ECC-POPS simulation can be
+explained by increased headgroup repulsion due to lower counterion binding affinity (Fig. \ref{??}).
+This explains also the larger area per molecule (57~\AA$^2$) reported for Lipid17 POPS simulations
+%being closer to the experimental value, 
+%In previous study \cite{NMRlipidsIV}, the Lipid17 force field
+with {\AA}qvist ions~\cite{NMRlipidsIV}.
+%gave a 
+%which is probably due to the weaker counterion binding affinity of {\AA}qvist ions to phospholipids. \cite{catte16, NMRlipidsIV}
+However, the Dang ions %with more realisitic bulk properties
+are used here to analyze the effect of ECC to the properties of a POPS lipid bilayer because
+{\AA}qvist ions produce known artifacts, like artificial aggregation of ions at larger salt concentrations \cite{kohagen16, chen07, NMRlipidsIV}.
 

### 2019-05-28 Manuscript/manuscript.tex
Update text

@@
 agree within their error estimates. The larger area in ECC-POPS simulation can be
-explained by increased headgroup repulsion due to lower counterion binding affinity (Fig. \ref{??}).
+explained by increased headgroup repulsion due to lower counterion binding affinity (Fig. \ref{fig:POPS-counterions-dens}).
 This explains also the larger area per molecule (57~\AA$^2$) reported for Lipid17 POPS simulations
@@
 {\AA}qvist ions produce known artifacts, like artificial aggregation of ions at larger salt concentrations \cite{kohagen16, chen07, NMRlipidsIV}.
+
+%SAMULI: I have moved this figure back to the main text
+\begin{figure}[hbp!] 
+  \centering 
+  \includegraphics[width=8cm]{../img/ecc_pops/density_profiles_na-k-counterions_wat_phos_compar_purePOPS_ecclipids-lipid17.pdf}
+  \caption{\label{fig:POPS-counterions-dens}
+    Number density profiles of \ce{K^{+}} and \ce{Na^{+}} counterions along the membrane normal axis
+    in ECC-lipids (solid lines) and Lipid17/Dang (dashed lines) simulations of pure POPS bilayers.  
+    The density profiles of phosphate groups and water are divided by 4 and 100, respectively.  
+  }
+  \todo{This figure might be more clear without the lipids in background.}
+\end{figure}
 
@@
 This is the case also for Lipid17 POPS simulations, where the headgroup and glycerol backbone
-order parameters are quite far from experimental values with both Dang (Fig.~\ref{simVSexpNOions_POPS})
-and {\AA}qvist ions. \cite{NMRlipidsIV} The values from ECC-POPS simulation are closer,
+order parameters are quite far from experimental values with Dang (Fig.~\ref{simVSexpNOions_POPS}),
+{\AA}qvist and Joung-Cheatham ions~\cite{NMRlipidsIV}. The values from ECC-POPS simulation are closer,
 but not in full agreement with experiments (Fig.~\ref{simVSexpNOions_POPS}).
-Especially, the experimentally observed more negative values of the $\beta$-carbon
-and the larger forking (difference of order parameters between hydrogens attached to the
-same carbon \cite{ollila16}) of the $\alpha$-carbon in POPS than in POPC \cite{NMRlipidsIV}
-are not reproduced by neither of the models here. 
-These differences are previously related
-to a more rigid structure of PS headgroup \cite{browning80,buldt81} and 
-the stiffness of the C$_\alpha$-C$_\beta$-C$_\gamma$-O$_\gamma$ dihedral \cite{NMRlipidsIV}
-plotted in Fig.~\ref{fig:dihedral}. 
-Interestingely, this dihedral angle is more flexible in ECC-POPS compared to Lipid17 
-suggesting that its rigidity is overestimated in Lipid17 due to the lack of electronic polarizability,
-while the same dihedral parameters yield too small forking in ECC-POPS.
-
-Overall,
-the head group and acyl chain order parameters within ECC-lipids
-are in a good agreement with the experimental values 
-as shown in Fig.~\ref{simVSexpNOions_POPS}. 
-The acyl chain order parameters in particular agree with the experimental data within error bars.
-The order parameters of the head groups are at an accuracy comparable to 
-other currently available classical models of lipids \citep{botan15, catte16, Pluhackova2016, nmrlipids_proj4}. 
-In particular, ECC-POPS is closer to the experimental order parameters 
-than Lipid17 with either Dang \cite{smith94,chang1999,dang2006} 
-or {\AA}qvist ions \cite{aqvist90}. 
-Although the ECC-POPS model reproduces the lipid bilayer dimensions and
-acyl chain conformations in a good agreement with experiments, 
-there is a room for improvements especially in the glycerol backbone and headgroup region,
-e.g. in the parameters of the associated dihedral angles. 
-\cite{botan15, ollila16, Pluhackova2016, NMRlipidsIV}
+
+% SAMULI: I commented this out to simplify the manuscript
+%
+%Especially, the experimentally observed more negative values of the $\beta$-carbon
+%and the larger forking (difference of order parameters between hydrogens attached to the
+%same carbon \cite{ollila16}) of the $\alpha$-carbon in POPS than in POPC \cite{NMRlipidsIV}
+%are not reproduced by neither of the models here. 
+%These differences are previously related
+%to a more rigid structure of PS headgroup \cite{browning80,buldt81} and 
+%the stiffness of the C$_\alpha$-C$_\beta$-C$_\gamma$-O$_\gamma$ dihedral \cite{NMRlipidsIV}
+%plotted in Fig.~\ref{fig:dihedral}. 
+%Interestingely, this dihedral angle is more flexible in ECC-POPS compared to Lipid17 
+%suggesting that its rigidity is overestimated in Lipid17 due to the lack of electronic polarizability,
+%while the same dihedral parameters yield too small forking in ECC-POPS.
+
+%Overall,
+%the head group and acyl chain order parameters within ECC-lipids
+%are in a good agreement with the experimental values 
+%as shown in Fig.~\ref{simVSexpNOions_POPS}. 
+In conclusion,
+%the acyl chain order parameters in particular agree with the experimental data within error bars.
+%The order parameters of the head groups are at an accuracy comparable to 
+the stuctural quality of ECC-POPS is similar to the other currently available
+lipid force fields \cite{botan15, catte16, Pluhackova2016, nmrlipids_proj4},
+%In particular, ECC-POPS is closer to the experimental order parameters 
+%than Lipid17 with either Dang \cite{smith94,chang1999,dang2006} 
+%or {\AA}qvist ions \cite{aqvist90}. 
+%Although the ECC-POPS model reproduces the lipid bilayer dimensions and
+giving good agreement with experiments for acyl chain conformations and bilayer dimensions,
+while there is a room for improvement in the glycerol backbone and headgroup region
+conformations.
+%e.g. in the parameters of the associated dihedral angles. 
+%\cite{botan15, ollila16, Pluhackova2016, NMRlipidsIV}
 
@@
 
-In simulations, binding affinity of monovalent ions to lipid bilayers depends strongly on
-force field parameters \cite{catte16,NMRlipidsIV}.
-% J: I think the following sentence can be omitted
-%For instance in the present work the affinity of sodium counterions to a POPS bilayer 
-%is stronger in Lipid17/Dang simulations 
-%than in simulations with ECC-POPS (Fig.~\ref{fig:POPS-counterions-dens} and~\ref{fig:delta_ordPar_NaCl_PC-PS_mix_PC}).
+Binding affinity of monovalent ions to lipid bilayers depends strongly on
+force field parameters in simulations \cite{catte16,NMRlipidsIV}.
 The cation binding affinities to lipid bilayers in simulations have been previously evaluated
@@
 PS lipids has been evaluated measuring the PC headgroup order parameters from POPC:POPS (5:1)
-mixtures\cite{roux90,NMRlipidsIV}.
-Following our previous work \cite{NMRlipidsIV}, we evaluate the counterion binding affinity
-to a lipid bilayer containing PS in simulations using two approaches.
+mixtures\cite{roux90,NMRlipidsIV}. The evaluation of monovalent ion binding affinity to
+lipid bilayers containing PS lipids is complicated also by the lack of ion-free state due to the 
+counterions, which are always present with negatively charged lipids~\cite{NMRlipidsIV}.
+Therefore, we evaluate the counterion binding affinity to POPC:POPS mixtures by monitoring
+the response of POPC headgroup order parameters to both increasing mole fraction of PS lipids and
+to the increasing monovalent ion concetrations. %, as 
+%to a lipid bilayer containing PS in simulations using two approaches.
+%Following our previous work \cite{NMRlipidsIV},
 
@@
 
-First, we monitor the response of POPC headgroup order parameters to the increasing amount of POPS (Fig.~\ref{fig:delta_ordPar_NaCl_PC-PS_mix_PC}).
+%First, we monitor the response of POPC headgroup order parameters to the increasing amount of POPS (Fig.~\ref{fig:delta_ordPar_NaCl_PC-PS_mix_PC}).
 According to the electrometer concept, the head group order parameters of PC lipids increase
@@
 tilts more parallel to the membrane interface \cite{seelig87,scherer87}.
-The trend is observed with both force fields in simulations with \ce{K+} counterions and
-in ECC-lipids simulation with \ce{Na+} counterions. 
+However, in simulations with strong counterion binding affinity, the charge of bound
+counterions cancel the influence of negatively charged lipids and the increase in PC headgroup
+order parameters is not observed \cite{NMRlipidsIV}. This is observed in 
+Lipid17/Dang ion simulations with \ce{Na+} counterions (Fig~\ref{fig:delta_ordPar_NaCl_PC-PS_mix_PC}),
+while in ECC-lipids simulations with \ce{Na+} counterions and in Lipid17/Dang
+%The trend is observed with both force fields in simulations
+with \ce{K+} counterions the order parameters increase
+in 
 In contrast, the order parameters are unchanged or decrease 
 and the P--N vector tilts more perpendicular to the membrane interface.
-in Lipid17/Dang ion simulations with \ce{Na+} counterions.
+in 
 The result can be explained by the higher binding affinity of sodium in the

### 2019-05-29 Manuscript/manuscript.tex
Update text

@@
     (only \ce{Na^+} counterions, but shown also in the right plots for \ce{K+}).
+    Error bars are not visible for most of the simulation points because they are smaller than the point size.
     \textbf{Right:} Chemical structure and labelling of carbon segments of POPC. 
@@
 and will add ticks to the right side of the plotting area. }
+\todo{The scales should be the same in the left and right panel.}
 \end{figure*}
@@
 counterions cancel the influence of negatively charged lipids and the increase in PC headgroup
-order parameters is not observed \cite{NMRlipidsIV}. This is observed in 
-Lipid17/Dang ion simulations with \ce{Na+} counterions (Fig~\ref{fig:delta_ordPar_NaCl_PC-PS_mix_PC}),
-while in ECC-lipids simulations with \ce{Na+} counterions and in Lipid17/Dang
+order parameters is not observed \cite{NMRlipidsIV}. This is the case in 
+Lipid17/Dang ion simulations with \ce{Na+} counterions,
+while in simulations with \ce{K+} counterions and in
+ECC-lipids simulations both counterions,
 %The trend is observed with both force fields in simulations
-with \ce{K+} counterions the order parameters increase
-in 
-In contrast, the order parameters are unchanged or decrease 
-and the P--N vector tilts more perpendicular to the membrane interface.
-in 
-The result can be explained by the higher binding affinity of sodium in the
-Lipid17/Dang ion simulations (Fig~\ref{fig:delta_ordPar_NaCl_PC-PS_mix_PC}), which
-compensates the negative charge of PS lipids at the interface.
-This is corrected in ECC-lipids, 
-which only slightly underestimate the slopes of the order parameters 
-when \ce{Na+} counterions are used (Fig.~\ref{fig:delta_ordPar_NaCl_PC-PS_mix_PC}).
-
-Second, we monitor the response of PC headgroup order parameters to the additional
-amount of \ce{NaCl} and \ce{KCl} in POPC:POPS (5:1) mixtures.
-In experiments, the addition of K$^+$ ions led to a very modest decrease of the order parameters,
-while the decrease was more pronounced with Li$^+$ ions and was the strongest with divalent Ca$^{2+}$ and Mg$^{2+}$ ions \cite{roux90}.
-According to the electrometer concept, the decrease of PC lipid head group order parameters
-correlates with the amount of bound cations \cite{akutsu81,altenbach84,seelig87,roux90,catte16,NMRlipidsIV},
-thus the binding affinity of the ions to POPC:POPS (5:1) mixtures increase in order K$^{+}$ $<$ Li$^{+}$  $<$ Mg$^{2+}$  $<$ Ca$^{2+}$.
-The response of PC headgroup order parameters to the increasing amount of KCl is close to experiments
-and the change in P--N vector is negligible in both force fields (Fig.~\ref{fig:delta_ordPar_monoval_PCPS}), athough the change of the $\alpha$-carbon
-order parameter may be slightly overestimated by the Lipid17/Dang simulations. 
-This could be explained by the deeper penetration of \ce{K+} into the bilayer (Fig.~\ref{fig:delta_ordPar_NaCl_PC-PS_mix_PC}).
-
-Because experimental data with the additional amount of \ce{Na+} in POPC:POPS (5:1) mixture was not available,
-we compared the simulation results to the experimental data with the additional amounts of
-\ce{Li+} and \ce{K+} ions, which should give the lower and the upper bounds to the sodium binding
-affinity (Fig.~\ref{fig:delta_ordPar_monoval_PCPS}). The PC headgroup response to the additional
-amount of sodium is clearly overestimated by the Lipid17/Dang simulation, and
-the response in ECC-lipids simulations is probably too close to the experimental results for lithium.
+the counterion binding affinity is weaker and the
+order parameters increase as in experiments (Fig~\ref{fig:delta_ordPar_NaCl_PC-PS_mix_PC}).  
+%In contrast, the order parameters are unchanged or decrease 
+%and the P--N vector tilts more perpendicular to the membrane interface.
+%in 
+%The result can be explained by the higher binding affinity of sodium in the
+%Lipid17/Dang ion simulations (Fig~\ref{fig:delta_ordPar_NaCl_PC-PS_mix_PC}), which
+%compensates the negative charge of PS lipids at the interface.
+%This is corrected in ECC-lipids, 
+%which only slightly underestimate the slopes of the order parameters 
+%when \ce{Na+} counterions are used (Fig.~\ref{fig:delta_ordPar_NaCl_PC-PS_mix_PC}). 
+According to the electrometer concept, the decrease of PC head group order parameters
+correlates with the amount of bound cations to lipid bilayer \cite{akutsu81,altenbach84,seelig87,roux90,catte16,NMRlipidsIV}.
+%Second, we monitor the response of PC headgroup order parameters to the additional
+%amount of \ce{NaCl} and \ce{KCl} in POPC:POPS (5:1) mixtures.
+In experiments, the addition of K$^+$ ions led to a very modest decrease of POPC headgroup order parameters in POPC:POPS (5:1) mixture,
+while the decrease was more pronounced with Li$^+$ ions and was strongest with divalent Mg$^{2+}$ and Ca$^{2+}$ ions~\cite{roux90},
+suggesting that the binding affinity increases in order K$^{+}$ $<$ Li$^{+}$  $<$ Mg$^{2+}$  $<$ Ca$^{2+}$.
+The small changes of POPC headgroup order parameters with increasing amount of KCl are close to experiments
+%and the change in P--N vector is negligible
+in both force fields (Fig.~\ref{fig:delta_ordPar_monoval_PCPS}).
+The slightly overestimated change of the $\alpha$-carbon
+order parameter in the Lipid17/Dang simulations may be due to the
+deeper penetration of \ce{K+} into the bilayer (Fig.~\ref{fig:delta_ordPar_monoval_PCPS}).
+Experimental data with the additional amount of \ce{Na+} in POPC:POPS (5:1) mixture is not available,
+%we compared the simulation results to the experimental data with the additional amounts of
+but \ce{Li+} and \ce{K+} results give the lower and the upper bounds, respectively, to the sodium binding
+affinity (Fig.~\ref{fig:delta_ordPar_monoval_PCPS}). In ECC-lipids simulations,
+the response of POPC headgroup order parameters to the additional sodium is close to the
+experimental results for lithium, while in Lipid17/Dang simulations the response is larger.
 % The following conclusion is already in the previous sentence, so I commented it out
-%In conclusion, the results suggest that the sodium binding affinity (Fig. \ref{fig:POPS-counterions-dens}) and
-%the tilting of P--N vector more perpendicular to the membrane surface (Fig. \ref{fig:delta_ordPar_NaCl_PCPS})
-%are clearly overestimated in Lipid17/Dang simulations and slightly overestimated in ECC-lipids simulations,
-%while the binding affinity of potassium is better described by both force fields.
+Overall, the results suggest that the \ce{Na+} binding affinity is
+% (Fig. \ref{fig:POPS-counterions-dens})
+%and the tilting of P--N vector more perpendicular to the membrane surface (Fig. \ref{fig:delta_ordPar_NaCl_PCPS})
+clearly overestimated in Lipid17/Dang simulations and slightly overestimated in ECC-lipids simulations,
+while the binding affinity of potassium is better described by both force fields.
 
@@
     Because experimental data with \ce{NaCl} is not available for POPC, the data for KCl and \ce{LiCl} (dashed line, left)
-    are shown as lower and upper bounds for the response to \ce{NaCl}, respectively.
+    are shown as lower and upper bounds, respectively, for the response to \ce{NaCl}.
+    Error bars are not visible for ECC-lipids simulations because they are smaller than the point size.
   }
@@
 and make a common legend distinguishing different experiments better. }
+\todo{The scale in left and right columns (response to KCl and NaCl) should be the same (not necessarily the same scale for POPC and POPS though).}
 \end{figure}
 
-The changes of POPS headgroup order parameters are almost negligible in experiments with increasing POPC fraction in bilayer
-or monovalent ion concentration in solvent
-(Figs.~\ref{fig:delta_ordPar_monoval_PCPS},~\ref{fig:delta_ordPar_monoval_PCPS} and~\ref{fig:delta_ordPar_NaCl_PC-PS_mix}).
+The behaviour of POPS headgroup can be further characterized by monitoring its order parameters in different
+lipid mixtures and ionic conditions~\cite{NMRlipidsIV,roux90}.
+Upon increasing the POPC content or monovalent ion concetration in experiments, 
+the headgroup order parameters of POPS are almost unchanged
+(\ref{fig:delta_ordPar_NaCl_PC-PS_mix_PS} and Figs.~\ref{fig:delta_ordPar_monoval_PCPS}).
 %while calcium leads to relatively large changes with low concentrations followed by a rapid saturation below 100mM (Fig. \ref{fig:delta_ordPar_CaCl_PCPS}) \cite{scherer87,roux90}.
@@
 %\ce{CaCl2}: Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}). 
-In Lipid17/Dang simulations, the changes of POPS headgroup order parameters with increasing amount of POPC
-are overestimated, which is also related to the changes in the orientations of the P--N and C$_{\beta}$--C$_{\gamma}$
-vectors (Fig.~\ref{fig:delta_ordPar_NaCl_PC-PS_mix_PC}). 
+In Lipid17/Dang simulations, the changes of POPS headgroup order parameters with increasing amount of
+both POPC or monovalent salts (\ce{KCl} and \ce{NaCl}) are overestimated (\ref{fig:delta_ordPar_NaCl_PC-PS_mix_PS} and Figs.~\ref{fig:delta_ordPar_monoval_PCPS}).
 %These structural changes can relate to variations in the interactions between PC and PS head groups 
 %and also to the changing amount of counterions and PS in the composition. 
-Also, the PS headgroup order parameters from simulations with Lipid17/Dang model
-are too sensitive to an additional amount of \ce{KCl} and \ce{NaCl} 
-(Fig.~\ref{fig:delta_ordPar_monoval_PCPS}).
-The overestimation of the order parameter changes upon increasing amount of POPC in a bilayer 
-was also observed in previous simulations with other force fields \cite{NMRlipidsIV}, 
-but not in ECC-lipids simulation, 
-where the changes in the PS headgroup orientation are more modest (Fig.~\ref{fig:delta_ordPar_NaCl_PC-PS_mix_PC}).
-In particular, 
-the dependence of the PS headgroup order parameters from ECC-lipids simulations
-to increasing amount of POPC, \ce{KCl} or \ce{NaCl}
-is weaker and closer to experiments 
-(Fig.~\ref{fig:delta_ordPar_monoval_PCPS}).
+%Also, the PS headgroup order parameters from simulations with Lipid17/Dang model
+%are too sensitive to an additional amount of
+%(Fig.~\ref{fig:delta_ordPar_monoval_PCPS}).
+Similar results %The overestimation of the order parameter changes upon increasing amount of POPC in a bilayer 
+also for other force fields are reported elsewhere~\cite{NMRlipidsIV}.
+In ECC-lipids simulations, the response of POPS headgroup order parameters to both POPC and monovalent salt concentration (\ce{KCl} and \ce{NaCl})
+are more modest and closer to experiments (\ref{fig:delta_ordPar_NaCl_PC-PS_mix_PS} and Figs.~\ref{fig:delta_ordPar_monoval_PCPS}).
+The changes in order parameters can be related to the orientations of the P--N and C$_{\beta}$--C$_{\gamma}$
+vectors in POPC and POPS, which change much less and more systematically in ECC-lipids
+than in Lipid17/Dang simulations (\ref{fig:delta_ordPar_NaCl_PC-PS_mix_PS} and Figs.~\ref{fig:delta_ordPar_monoval_PCPS}).
+
+% where the changes in the PS headgroup orientation are more modest (Fig.~\ref{fig:delta_ordPar_NaCl_PC-PS_mix_PC}).
+%In particular, 
+%the dependence of the PS headgroup order parameters from ECC-lipids simulations
+%to increasing amount of POPC, \ce{KCl} or \ce{NaCl}
+%is weaker and closer to experiments 
+%(Fig.~\ref{fig:delta_ordPar_monoval_PCPS}).
 %In contrast, ECC-lipids with ECC-ions capture the different response 
@@
 %
-The influence of negatively charged PS on the counterion binding affinity can be estimated 
+We estimate the influence of negatively charged POPS on the counterion binding affinity
+% can be estimated 
 %by calculating the amounts of phospholipids per bound \ce{Ca^{2+}} cation, $\xi _{Ca^{2+}}$,
 by comparing the relative surface excesses with respect to water, $\Gamma ^{w} _{\rm ion}$,
-between POPC:POPS (5:1) mixture and pure POPC \cite{melcr18}.
+between POPC:POPS (5:1) mixture and pure POPC in ECC-lipids simulations with the most realistic
+cation binding affinities~\cite{melcr18}.
 %Threshold for counting bound \ce{Ca^{2+}} cations to a membrane
@@
 %may not be fully captured by the ECC-lipids simulations.
-The increased calcium binding affinity to bilayers containing PS lipids is illustrated in
-ECC-lipids simulations by the increasing relative surface excess of calcium with respect to water from
+Therefore, we use ECC-lipids simulations to estimate the influence of negative charged POPS on
+calcium binding affinity to lipid bilayers.
+%containing PS lipids is illustrated in
+%ECC-lipids simulations by the increasing
+The relative surface excess of calcium with respect to water 
 %  previously we used 350mM conc, but I think it is better to compare to a higher concentration to support the whole argument bullet-proof
@@
 $\Gamma^{w}_{Ca} = 0.09\mathrm{nm^{-2}}$ ($\sim 5.4$ phospholipids per bound \ce{Ca^{2+}}) for pure POPC bilayer with 467~mM CaCl$_2$ \cite{ECC-POPC_nacl_cacl2_files}
-to $\Gamma^{w}_{Ca} = 0.24\mathrm{nm^{-2}}$ ($\sim 4.8$ phospholipids per bound \ce{Ca^{2+}}) for the POPC:POPS (5:1) bilayer with 409 mM CaCl$_2$.
-
-%The binding of calcium cations to the phospholipid bilayer is shown 
-%in the density profiles of \ce{Ca^{2+}}, \ce{Na^+} counterions 
-%and also \ce{Cl^-} ions in Fig.~\ref{fig:cacl-dens_PCPS}. 
-%Clearly, after accounting for the electronic polarizability, 
-%the population of the cations at the membrane interface is significantly diminished.
+is increased to
+$\Gamma^{w}_{Ca} = 0.24\mathrm{nm^{-2}}$ ($\sim 4.8$ phospholipids per bound \ce{Ca^{2+}}) for the POPC:POPS (5:1) bilayer with 409 mM CaCl$_2$.
+Besides lower binding affinity, ECC-lipids simulations yield smaller error bars for order parameters and shorter residence times (Fig.~\ref{fig:hist_residence_times}) 
+than those typically observed in simulations with other force fields \cite{javanainen17,melcr18,NMRlipidsIV},
+suggesting that the ECC accelerates the equlibration of ions at lipid bilayer interface.
+Therefore, our 1 $\mu$s simulations seem to be sufficiently long for the ECC-lipids simulations, because
+90\% of the calcium residence times are shorter than $60\,\mathrm{ns}$ for pure POPC bilayer
+and shorter than $200\,\mathrm{ns}$ for POPC:POPS (5:1) mixture, while the
+longest observed residence times are $141\,\mathrm{ns}$ and $485\,\mathrm{ns}$, respectively (Fig.~\ref{fig:hist_residence_times}).
+Interestingly, the calcium residence times are 3-4 times longer in POPC:POPS (5:1) mixture than in pure POPC.
 
@@
     \textbf{Middle panels:}
-    Changes of the head group order parameters, and the angles of P--N and C$_\beta$--C$_\gamma$ (stars) vectors 
+    Changes of the head group order parameters, and the angles of P--N (circles) and C$_\beta$--C$_\gamma$ (stars) vectors 
     with respect to the membrane normal of POPC (left) and POPS (right) in a POPC:POPS (5:1) bilayer 
-    as a function of \ce{CaCl} concentration from ECC-lipids and Lipid17/Dang simulations 
-    compared with experimental values from Ref. \citenum{roux90} (signs from Ref. \citenum{ferreira16}) at 298 K.
+    as a function of \ce{CaCl$_2$} concentration from ECC-lipids and Lipid17/Dang simulations 
+    compared with experimental values from Ref. \citenum{roux90} (signs from Refs. \citenum{ferreira16} and \citenum{NMRlipidsIV}) at 298 K.
     The y-axis for the $\alpha$-carbon results of POPS (middle right) is transferred
@@
     parameter value from pure POPS is at zero to correctly illustrate the significant forking.
+    Error bars are not visible for ECC-lipids simulations because they are smaller than the point size.
   }
   \caption{ \label{fig:cacl_complexes} 
-  \textbf{Right top:} Percentages of the population of bound Ca$^{2+}$ 
+    \textbf{Right top:} Percentages of the bound Ca$^{2+}$ 
     in exclusive contact with the given oxygen moiety
-	in a POPC:POPS (5:1) bilayer with 409 mM CaCl$_2$. 
-	The areas represent populations of bound \ce{Ca^{2+}} 
-	only to the given moieties exclusively (shades of orange and red)
-	or their combinations (CHANGE:blue)
-	separately for PC, PS, or PC and PS simultaneously. 
-	The areas sum up to total populations at the respective lipids. 
-	In the case of mixed simultaneous binding to PC and PS
-	we distinguish different moieties only in PS.
-	%i.e., PC can interact with any considered moiety. 
-	For comparison, in a pure POPC bilayer with 350 mM CaCl$_2$, 
-	67\% of bound \ce{Ca^{2+}} is exclusively in contact with \ce{PO_4} group. \cite{melcr18} 
+    when bound to PC only, PS only or simultaneously to both
+    calculated from ECC-lipids simulation of POPC:POPS (5:1) bilayer with 409 mM CaCl$_2$. 
+%	The areas represent populations of bound \ce{Ca^{2+}} 
+%	only to the given moieties exclusively (shades of orange and red)
+%	or their combinations (CHANGE:blue)
+%	separately for PC, PS, or PC and PS simultaneously. 
+%	The areas sum up to total populations at the respective lipids. 
+	In the case of simultaneous binding to both PC and PS,
+	the percentages refer to the moieties in PS.
+	% i.e., PC can interact with any considered moiety.
+        For numerical values, see table~\ref{tab:binding}.
     \textbf{Right bottom:} Relative probabilities of \ce{Ca^{2+}} ions to coordinate with a certain number of lipids
@@
   }
-\todo{J: Will adjust the Fig into a nice panel and change the captions accordingly. }
+  \todo{J: Will adjust the Fig into a nice panel and change the captions accordingly. } \\
+  \todo{All these in the same figure might be too much, maybe the most right panel into a separate figure?} \\
+  \todo{In the most right panel, the x-axis label should be changed to ''number of lipids directly bound to Ca$^{2+}$''.
+    The current label has currently another meaning in the text.}
 \end{figure}
 
-Experimentally measured headgroup order parameters of POPS exhibit a strong dependence
-on low CaCl$_2$ concentrations with a rapid saturation below 100~mM \cite{roux90}.
-Also these changes were overestimated by the Lipid17/Dang (also shown in Fig.~\ref{fig:delta_ordPar_CaCl_PCPS})
+Interactions between Ca$^{2+}$ ions and PS lipids can be further characterized by monitoring
+the POPS headgroup order parameters from POPC:POPS (5:1) mixture with increasing  CaCl$_2$ concentration~\cite{roux90}.
+In experiments, the headgroup order parameters of POPS exhibit a strong dependence
+on low CaCl$_2$ concentrations with a rapid saturation below 100~mM  (Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}) \cite{roux90}.
+These changes are overestimated in simulations with Lipid17/Dang (Fig.~\ref{fig:delta_ordPar_CaCl_PCPS})
 and other tested force fields in our recent work \cite{NMRlipidsIV}, 
-but CHARMM36 with the NBfix correction for calcium
+including CHARMM36 with the NBfix correction for calcium
 which underestimated the binding affinity.
@@
 In addition to possibly suboptimal interactions between calcium ions and PS headgroup, the potential sources of this
-discrepancy include the above observed slight overestimation of \ce{Na^+} counterions and
+discrepancy include the above observed sligth overestimation of \ce{Na^+} counterions and
 imperfect structures of the lipid headgroup (Fig.~\ref{simVSexpNOions_POPS}).
-
-In ECC-lipids simulations, the average orientation of P--N vectors of both POPC and POPS headgroups tilts more perpendicular to the membrane surface
-by 11$^\circ$ and  3$^\circ$, respectively, upon addition of 620~mM CaCl$_2$ to POPC:POPS (5:1) mixture (Fig.~\ref{fig:delta_ordPar_CaCl_PCPS})
+The POPS headgroup order parameters are also related to the changes of average orientation of P--N and
+C$_{\beta}$--C$_{\gamma}$ vectors in the PS headgroup, which are smaller and more systematic
+in ECC-lipids than in Lipid7/Dang simulations (Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}).
+Upon addition of 620~mM CaCl$_2$ to POPC:POPS (5:1) mixture,
+the average orientation of P--N vectors in both POPC and POPS headgroups tilts more perpendicular to the membrane surface
+by 11$^\circ$ and  3$^\circ$, respectively, in ECC-lipids simulations (Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}),
 suggesting that the PS headgroup orientation is less sensitive to the bound calcium than PC.
@@
 tilts to the opposite direction by $11^\circ$ in the same system.
-In Lipid17/Dang simulations, the changes in both headroup orientations and order parameters
-are larger and less systematic for the PS headgroup, especially with lower concentrations (Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}).
-
-ECC-lipids simulations yield smaller error bars and shorter residence times (Fig.~\ref{fig:hist_residence_times}) 
-than those typically observed in simulations with other force fields \cite{javanainen17,melcr18,NMRlipidsIV}.
-This suggests that the ECC accelerates the equlibration of ions at lipid bilayer interface.
-Our 1 $\mu$s simulations seem to be sufficiently long for the ECC-lipids simulations, because
-90\% of the calcium residence times are shorter than $60\,\mathrm{ns}$ for pure POPC bilayer
-and shorter than $200\,\mathrm{ns}$ for POPC:POPS (5:1) mixture, while the
-longest observed residence times are $141\,\mathrm{ns}$ and $485\,\mathrm{ns}$, respectively (Fig.~\ref{fig:hist_residence_times}).
-Interestingly, the calcium residence times are 3-4 times longer in POPC:POPS (5:1) mixture than in pure POPC.
+%In Lipid17/Dang simulations, the changes in both headroup orientations and order parameters
+%are larger and less systematic for the PS headgroup, especially with lower concentrations (Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}).
 
@@
 %Therefore, calcium ions bind more likely to negatively charged PS lipids than neutral PC lipid in a mixed bilayer, as expected.
+	For comparison, in a pure POPC bilayer with 350 mM CaCl$_2$, 
+	67\% of bound \ce{Ca^{2+}} is exclusively in contact with \ce{PO_4} group. \cite{melcr18} 
 Also, calcium ions bind more likely 

### 2019-05-29 Manuscript/manuscript.tex
Update text

@@
   \begin{tabular}{ c }
-  \includegraphics[height=\figheightsmall]{../img/ecc_pops/density_profiles_ca_na_k_cl_wat_phos_ecclipids_lipid17_compar_80and200mMCaCl2.pdf}
-  \includegraphics[height=\figheightsmall]{../img/ecc_pops/order_parameters_changes_ecc-lip_L14_A-B-PN-COO_POPC_cacl.pdf} 
-  \includegraphics[height=\figheightsmall]{../img/ecc_pops/order_parameters_changes_ecc-lip_L14_A-B-PN-COO_POPS_cacl.pdf} 
+  \includegraphics[height=1.5\figheightsmall]{../img/ecc_pops/density_profiles_ca_na_k_cl_wat_phos_ecclipids_lipid17_compar_80and200mMCaCl2.pdf}
+  \includegraphics[height=1.5\figheightsmall]{../img/ecc_pops/order_parameters_changes_ecc-lip_L14_A-B-PN-COO_POPC_cacl.pdf} 
+  \includegraphics[height=1.5\figheightsmall]{../img/ecc_pops/order_parameters_changes_ecc-lip_L14_A-B-PN-COO_POPS_cacl.pdf} 
   \end{tabular}
-  \begin{tabular}{ c }
-  \includegraphics[width=3cm]{../img/bound_ca_populations.pdf} \\
-  \includegraphics[width=3cm]{../img/stoichiometry_CaCl2_comparison_Ecc-lipids_PC-vs-PCPS.pdf} 
-  \end{tabular}
+%  \begin{tabular}{ c }
+%  \includegraphics[width=3cm]{../img/bound_ca_populations.pdf} \\
+%  \includegraphics[width=3cm]{../img/stoichiometry_CaCl2_comparison_Ecc-lipids_PC-vs-PCPS.pdf} 
+%  \end{tabular}
   \caption{\label{fig:cacl-dens_PCPS} 
@@
   }
+%  \caption{ \label{fig:cacl_complexes} 
+%    \textbf{Right top:} Percentages of the bound Ca$^{2+}$ 
+%    in exclusive contact with the given oxygen moiety
+%    when bound to PC only, PS only or simultaneously to both
+%    calculated from ECC-lipids simulation of POPC:POPS (5:1) bilayer with 409 mM CaCl$_2$. 
+%%	The areas represent populations of bound \ce{Ca^{2+}} 
+%%	only to the given moieties exclusively (shades of orange and red)
+%%	or their combinations (CHANGE:blue)
+%%	separately for PC, PS, or PC and PS simultaneously. 
+%%	The areas sum up to total populations at the respective lipids. %
+%	In the case of simultaneous binding to both PC and PS%,
+%	the percentages refer to the moieties in PS.
+%	% i.e., PC can interact with any considered moiety.
+%        For numerical values, see table~\ref{tab:binding}.
+%    \textbf{Right bottom:} Relative probabilities of \ce{Ca^{2+}} ions to coordinate with a certain number of lipids
+%    in pure POPC bilayer with 350 mM CaCl$_2$ and in POPC:POPS (5:1) mixture with 400 mM CaCl$_2$.  
+%    Analysis was done for both systems by considering all lipids (blue and violet) and
+%    for POPC:POPS (5:1) mixture also by considering POPC and POPS lipids separately (green and orange). 
+%    %All lipids were taken into account with the exception of the complexes in light green and orange, 
+%    %for which we counted only contacts with POPC resp POPS from the mixed 5\,PC:1\,PS negatively charged bilayer. 
+%    %The calculated probabilities of the calcium-lipid complexes also reflect only POPC (light green) resp POPS (orange). 
+%    % Probabilities were taken from simulations with comparable bulk concentrations of calcium around 250~mM. 
+%    Clusters of four or more lipids were not observed in either membrane.
+%    The threshold for counting coordinated lipids in a complex with \ce{Ca^{2+}} was set to $0.3\,\mathrm{nm}$, 
+%    as the distance between the cation and the oxygen atoms of the lipids. 
+%    Previously published simulation data \cite{melcr18} for pure POPC bilayers were taken directly from \cite{ECC-POPC_nacl_cacl2_files}. 
+%  }
+  \todo{J: Will adjust the Fig into a nice panel and change the captions accordingly. } \\
+\end{figure}
+
+% SAMULI: I separated the figure below, because it was too small to be read when combined with others.
+% It might be too much to all of these into the same.
+
+\begin{figure}[tbp!] 
+  \centering 
+  \begin{tabular}{ c }
+  \includegraphics[width=8cm]{../img/bound_ca_populations.pdf} \\
+  \includegraphics[width=8cm]{../img/stoichiometry_CaCl2_comparison_Ecc-lipids_PC-vs-PCPS.pdf} 
+  \end{tabular}
   \caption{ \label{fig:cacl_complexes} 
@@
   }
-  \todo{J: Will adjust the Fig into a nice panel and change the captions accordingly. } \\
-  \todo{All these in the same figure might be too much, maybe the most right panel into a separate figure?} \\
-  \todo{In the most right panel, the x-axis label should be changed to ''number of lipids directly bound to Ca$^{2+}$''.
+  \todo{The x-axis label should be changed to ''number of lipids directly bound to Ca$^{2+}$''.
     The current label has currently another meaning in the text.}
@@
 
-% Description of no. conctacts counting
-%Such a significantly increased binding also has an impact 
-%on the structure and populations of the bound calcium cations. 
-%We have counted contacts between the cations and the oxygen atoms of the lipids
-%similarly as was done in \cite{melcr18} 
-%to obtain such a detailed information 
-%about the interactions of \ce{Ca^{2+}} with various moieties in POPC or POPS.
-
-% Populations from contacts
-
 Recent analyzes of interaction sites between calcium and PS lipids combining
@@
 controversial because the results strongly depend on force fields parameters \cite{melcrova16,valentine18,hallock18}.
-In ECC-lipids simulation of POPC:POPS (5:1) mixture with 409 mM CaCl$_2$,
-on average 41\% of the total population of bound calcium cations is in contact with PS lipids
-with 8\% bound exclusively only to them even though the negatively charged membrane contains only 17\% of POPS (Fig.~\ref{fig:cacl_complexes}).
-%Therefore, calcium ions bind more likely to negatively charged PS lipids than neutral PC lipid in a mixed bilayer, as expected.
-	For comparison, in a pure POPC bilayer with 350 mM CaCl$_2$, 
-	67\% of bound \ce{Ca^{2+}} is exclusively in contact with \ce{PO_4} group. \cite{melcr18} 
-Also, calcium ions bind more likely 
-to the carboxylate than to the phosphate moiety of POPS headgroup (Fig.~\ref{fig:cacl_complexes}),
-in line with the results without the NBfix corrections in CHARMM36 \cite{hallock18}.
-This can explain the diminished flexibility of the PS head group 
-from simultaneous binding of \ce{Ca^{2+}} to the \ce{COO^-} group in PS and a phosphate group in PC 
-as the most probable configuration (Fig.~\ref{fig:cacl_complexes}).
-
-However, in CHARMM36 simulations without the NBfix correction for calcium \cite{kim16},
-calcium ions significantly overbind to PC lipids  \cite{catte16} and calcium ions coordinate
-with even four distinct PS lipids \cite{hallock18}, while coordination with more than two PS lipids
+Here, we analyze the calcium binding details from ECC-lipids simulation of POPC:POPS (5:1) mixture with 409 mM CaCl$_2$,
+because it surpasses the quality of other force fields in direct comparison with experimental order parameter data.
+In the ECC-lipid simulation, calcium ions binds approximately twice more likely 
+to the carboxylate than to the phosphate moiety of POPS headgroup,
+while binding to acyl chain carbonyls is almost negligible (Fig.~\ref{fig:cacl_complexes} and table~\ref{tab:binding}).
+The result is consistent with CHARMM36 simulations without the NBfix correction \cite{hallock18},
+but CHARMM36 simulations with the NBfix correction \cite{kim16} predict almost exclusive binding on carboxylate group \cite{valentine18}
+and Berger simulations signifincant binding affinity also on acyl chain carbonyls \cite{melcrova16}.
+On the other hand, calcium ions bind too strongly on bilayers simulated using CHARMM36 without the NBfix correction
+or Berger models, and too weakly on CHARMM36 bilayers with the NBfix correction \cite{catte16,NMRlipidsIV}.
+Furthermore, in CHARMM36 simulations without the NBfix,
+calcium ions coordinate with even four distinct PS lipids \cite{hallock18}, while coordination with more than two PS lipids
 was very rare in our ECC-lipids simulations (Fig.~\ref{fig:cacl_complexes}).
-Furthermore, the POPS:POPC:Ca$^{2+}$ (7:3:7) mixture with relatively high
-PS lipid fraction and calcium concentration was used in the previous work \cite{hallock18},
-which may lead to the complexation of \ce{Ca^{2+}} cations and PS lipids that is
-not observed with lower PS content used in ECC-lipids simulations in this work \cite{hauser77,kurland79,hauser85,feigenson86,mattai89,roux90,roux91}.
+%Furthermore, the POPS:POPC:Ca$^{2+}$ (7:3:7) mixture with relatively high
+%PS lipid fraction and calcium concentration was used in the previous work \cite{hallock18},
+%which may lead to the complexation of \ce{Ca^{2+}} cations and PS lipids that is
+%not observed with lower PS content used in ECC-lipids simulations in this work \cite{hauser77,kurland79,hauser85,feigenson86,mattai89,roux90,roux91}.
 %While POPC interacts with the calcium cations almost entirely through its phosphate group 
@@
 
-Significant coordination of calcium with carbonyl groups of PS and PC lipids observed in previous
-simulations with the Berger force field \cite{melcrova16} %\cite{berger97,mukhopadhyay04}
-was not observed in ECC-lipids simulations in here nor in the previous work with pure POPC \cite{melcr18}.
-Overestimated coordination of cations with carbonyl groups is a potential source for
-the overestimated cation binding in Berger force field \cite{catte16,NMRlipidsIV}. 
-In CHARMM36 simulations with the NBfix, calcium mainly interacted with only
-carboxylate of PS lipids  \cite{valentine18}, which is a potential source for the
-underestimated calcium binding affinity to lipid bilayer with this force field \cite{NMRlipidsIV}.
+%Significant coordination of calcium with carbonyl groups of PS and PC lipids observed in previous
+%simulations with the Berger force field \cite{melcrova16} %\cite{berger97,mukhopadhyay04}
+%was not observed in ECC-lipids simulations in here nor in the previous work with pure POPC \cite{melcr18}.
+%Overestimated coordination of cations with carbonyl groups is a potential source for
+%the overestimated cation binding in Berger force field \cite{catte16,NMRlipidsIV}. 
+%In CHARMM36 simulations with the NBfix, calcium mainly interacted with only
+%carboxylate of PS lipids  \cite{valentine18}, which is a potential source for the
+%underestimated calcium binding affinity to lipid bilayer with this force field \cite{NMRlipidsIV}.
+
+%on average 41\% of the total population of bound calcium cations is in contact with PS lipids
+%with 8\% bound exclusively only to them even though the negatively charged membrane contains only 17\% of POPS (Fig.~\ref{fig:cacl_complexes}).
+%%Therefore, calcium ions bind more likely to negatively charged PS lipids than neutral PC lipid in a mixed bilayer, as expected.
+%	For comparison, in a pure POPC bilayer with 350 mM CaCl$_2$, 
+%	67\% of bound \ce{Ca^{2+}} is exclusively in contact with \ce{PO_4} group. \cite{melcr18}
+
+%This can explain the diminished flexibility of the PS head group 
+%from simultaneous binding of \ce{Ca^{2+}} to the \ce{COO^-} group in PS and a phosphate group in PC 
+%as the most probable configuration (Fig.~\ref{fig:cacl_complexes}).
 

### 2019-05-31 Manuscript/manuscript.tex
Update text

@@
 
-%%SAMULI: new suggestion for the preliminary title
 \title[] 
-  { Detailed structure of negatively charged membranes 
-    of phosphatidylserine and phosphatidylcholine 
-    at concentrations of calcium, sodium and potassium salts
-    from molecular dynamics simulations with electronic polarization }
+{Improved Cation Binding to Lipid Bilayer with Negatively Charged POPS by Effective Inclusion of Electronic Polarization}
+  %Detailed structure of negatively charged membranes 
+  %  of phosphatidylserine and phosphatidylcholine 
+  %  at concentrations of calcium, sodium and potassium salts
+  %  from molecular dynamics simulations with electronic polarization }
   % JOSEF: I feel that the above title is a bit better, because it focuses on PS,
@@
 \begin{abstract} 
-
-\textbf{Main message: 
-ECC generally improves interaction and structure of both neutral and charged molecules. 
-Polarizability is necessarry for accurate simulations of cellular membranes at physiological ionic conditions.}
-
-%%% SAMULI: I commented this out because it way too general. We can write abstract later when the final content is clear.
-
-%Binding affinities and stoichiometries of sodium and calcium ions to phospholipid bilayers are of paramount significance in the properties and functionality of cellular membranes.
-Current estimates of binding affinities and stoichiometries of cations are, however, controversial due to limitations in the available experimental and theoretical methods.
-Experimentally one can assess these parameters by titrating several membrane properties as a function of the ion concentration.
-However, the interpretation of the experiments relies on theoretical models, as direct molecular information is not available.
-
-%Classical molecular dynamics (MD) simulations provide details of the ion binding process with atomistic resolution, therefore offering all the necessary information to interpret experimental data without the need to resort to over-simplified models.
-However, the accuracy of the currently available lipid models when interacting with ions is not sufficient for such interpretation \cite{catte16}.
-Recently, we have improved the binding details of sodium and calcium ions to the 1-Palmitoyl-2-oleoyl-phosphatidylcholine (POPC) bilayer by implicitly including electronic polarization as a mean field correction, known as the electronic continuum correction (ECC).
-Our improved ECC-POPC model reproduces not only the experimentally measured structural parameters for the ion-free membrane, but also improves significantly the response of lipid head group to a bound positive charge, and the binding affinities of sodium and calcium ions.
-\cite{melcr18}
-
-%We are now extending our work to other lipid membranes composed of larger varieties of different biologically relevant lipids. In particular, we focus our attention on membranes with 
-%	%phosphatidylethanolamine (PE) and 
-%	negatively charged phosphatidylserine (PS) lipids. 
-%	We compare available NMR and x-ray experiments with MD simulations to obtain atomistic-resolution picture of the interactions between cations and phospholipid membranes.
-Our results indicate that electronic polarization is a non-negligable interaction, which dramatically improves the description of cation binding to phospholipid membranes.
+  Phosphatidylserine (PS) lipids are important signaling molecules and the most common negatively charged lipids in eukaryotic membranes.
+  The signaling can be often regulated with calcium, but its interactions with PS headgroups are not fully understood.
+  Classical molecular dynamics (MD) simulations can potentially give detailed description of lipid-ion interactions,
+  but the results strongly depend on the used force field.
+  Here, we apply the electronic continuum correction (ECC) to the Amber Lipid17 parameters of POPS lipid to improve its
+  interactions with Na$^+$ and Ca$^{2+}$ ions.
+  The partial charges of headgroup, glycerol backbone and carbonyls of POPS, bearing a unit negative charge, were scaled with a
+  factor of 0.75, derived for monovalent ions and the Lennard-Jones $\sigma$ parameters of the same segments were scaled with a factor of 0.89.
+  The resulting ECC-POPS models, gives more realistic interactions with Na$^+$ and Ca$^{2+}$ cations than the original Amber Lipid17 parameters,
+  when validated using the headgroup order parameters and ''electrometer concept''.
+  In ECC-lipids simulations, the Ca$^{2+}$ cations do not simultaneously interact with more than two PS lipids, and interactions 
+  carboxylate groups is twice more likely than with phosphate group, while interaction with carbonyls is almost negligible.
+  Our results pave the way for more realistic MD simulations of anionic biological membranes and demonstrate the ECC
+  approach is useful also for charged lipids.
 \end{abstract}
@@
 acyl chain order parameters to mimic the profile predicted by the MD simulations (Fig.~\ref{simVSexpNOions_POPS}).
-\todo{Tiago has to check and update this and figures in SI.}
+\todo{Tiago should check and update this and figures in SI.}
 
@@
 it is less well undestood and the addition of cations may cause phase transitions in negatively charged
-bilayers \cite{feigenson86,mattai89,roux91,roux90}. Therefore, the cation binding affinity to bilayers with
-PS lipids has been evaluated measuring the PC headgroup order parameters from POPC:POPS (5:1)
-mixtures\cite{roux90,NMRlipidsIV}. The evaluation of monovalent ion binding affinity to
-lipid bilayers containing PS lipids is complicated also by the lack of ion-free state due to the 
-counterions, which are always present with negatively charged lipids~\cite{NMRlipidsIV}.
-Therefore, we evaluate the counterion binding affinity to POPC:POPS mixtures by monitoring
-the response of POPC headgroup order parameters to both increasing mole fraction of PS lipids and
-to the increasing monovalent ion concetrations. %, as 
+bilayers \cite{feigenson86,mattai89,roux91,roux90}.
+Therefore, the cation binding affinity to bilayers with PS lipids has been evaluated by measuring the decrease
+of PC headgroup order parameters from POPC:POPS (5:1) mixtures upon addition of salt concentration,
+which correlate with the amount of bound cations to lipid bilayer according to the
+electrometer concept~\cite{akutsu81,altenbach84,seelig87,roux90,catte16,NMRlipidsIV}.
+Because the evaluation of monovalent ion binding affinity to bilayers with
+%with negatively charged
+PS lipids is complicated by the lack of ion-free state (counterions are always present with negatively charged lipids),
+%Therefore,
+we evaluate the counterion binding affinity to POPC:POPS mixtures also by monitoring
+the response of POPC headgroup order parameters to the increasing mole fraction of PS lipids~\cite{NMRlipidsIV}.
+%and
+%to the increasing monovalent ion concetrations. %, as 
 %to a lipid bilayer containing PS in simulations using two approaches.
@@
 
+\begin{figure}[tbp!] 
+  \centering 
+  \includegraphics[height=\figheightsmall]{../img/ecc_pops/order_parameters_changes_ecc-lip_L14_A-B-PN-COO_POPC_kcl.pdf} 
+  \includegraphics[height=\figheightsmall]{../img/ecc_pops/order_parameters_changes_ecc-lip_L14_A-B-PN-COO_POPC_nacl.pdf} 
+  \includegraphics[height=\figheightsmall]{../img/ecc_pops/order_parameters_changes_ecc-lip_L14_A-B-PN-COO_POPS_kcl.pdf} 
+  \includegraphics[height=\figheightsmall]{../img/ecc_pops/order_parameters_changes_ecc-lip_L14_A-B-PN-COO_POPS_nacl.pdf} 
+  \caption{\label{fig:delta_ordPar_monoval_PCPS}
+    Changes of the head group order parameters, and the angles of P--N and \ce{C_\beta-C_\gamma} (stars) vectors 
+    with respect to the membrane normal of POPC (left) and POPS (right) in a POPC:POPS (5:1) bilayer 
+    as a function of \ce{KCl} and \ce{NaCl} concentration from ECC-lipids and Lipid17/Dang simulations 
+    compared with experimental values from Ref. \citenum{roux90} (signs from Ref. \citenum{ferreira16}) at 298 K.
+    The y-axis for the $\alpha$-carbon results of POPS (middle right) is transferred
+    with the same value for both order parameters such that the lower order
+    parameter value from pure POPS is at zero to correctly illustrate the significant forking.
+    Because experimental data with \ce{NaCl} is not available for POPC, the data for KCl and \ce{LiCl} (dashed line, left)
+    are shown as lower and upper bounds, respectively, for the response to \ce{NaCl}.
+    Error bars are not visible for ECC-lipids simulations because they are smaller than the point size.
+  }
+\todo{J: Will improve this figure with panels A, B, C and D 
+and put POPC together, POPS together 
+and remove the duplicate y-axis titles 
+and make a common legend distinguishing different experiments better. }
+\todo{The scale in left and right columns (response to KCl and NaCl) should be the same (not necessarily the same scale for POPC and POPS though).}
+\end{figure}
+
+In experiments, the addition of K$^+$ ions led to a very modest decrease of POPC headgroup order parameters in POPC:POPS (5:1) mixture,
+while the decrease was more pronounced with Li$^+$ ions and was strongest with divalent Mg$^{2+}$ and Ca$^{2+}$ ions~\cite{roux90},
+suggesting that the binding affinity increases in order K$^{+}$ $<$ Li$^{+}$  $<$ Mg$^{2+}$  $<$ Ca$^{2+}$.
+The small changes of POPC headgroup order parameters with increasing amount of KCl are close to experiments
+%and the change in P--N vector is negligible
+in both force fields (Fig.~\ref{fig:delta_ordPar_monoval_PCPS}).
+The slightly overestimated change of the $\alpha$-carbon
+order parameter in the Lipid17/Dang simulations may be due to the
+deeper penetration of \ce{K+} into the bilayer (Fig.~\ref{fig:delta_ordPar_NaCl_PC-PS_mix_PC}).
+Experimental data with the additional amount of \ce{Na+} in POPC:POPS (5:1) mixture is not available,
+%we compared the simulation results to the experimental data with the additional amounts of
+but \ce{Li+} and \ce{K+} results give the lower and the upper bounds, respectively, to the sodium binding
+affinity (Fig.~\ref{fig:delta_ordPar_monoval_PCPS}). In ECC-lipids simulations,
+the response of POPC headgroup order parameters to the additional sodium is close to the
+experimental results for lithium, while in Lipid17/Dang simulations the response is larger.
+% The following conclusion is already in the previous sentence, so I commented it out
+Overall, the results suggest that the \ce{Na+} binding affinity is
+% (Fig. \ref{fig:POPS-counterions-dens})
+%and the tilting of P--N vector more perpendicular to the membrane surface (Fig. \ref{fig:delta_ordPar_NaCl_PCPS})
+clearly overestimated in Lipid17/Dang simulations and slightly overestimated in ECC-lipids simulations,
+while the binding affinity of potassium is better described by both force fields.
+%First, we monitor the response of POPC headgroup order parameters to the increasing amount of POPS (Fig.~\ref{fig:delta_ordPar_NaCl_PC-PS_mix_PC}).
+This conclusion is also supported by the POPC headgroup responses to the
+increasing amount of POPS in different simulations (Fig~\ref{fig:delta_ordPar_NaCl_PC-PS_mix_PC}).
+In experiments,
+the head group order parameters of PC lipids increase upon addition of
+% with the increasing amount of
+negatively charged PS lipids, as predicted by the electrometer concept~\cite{seelig87,scherer87}.
+%, because the headgroup P--N dipole
+%tilts more parallel to the membrane interface \cite{seelig87,scherer87}.
+This is the case also in simulations with the exception of
+Lipid17/Dang with \ce{Na+} counterions, where the
+%but not with \ce{Na+} counterions,
+%However, in simulations with
+stronger counterion binding affinity
+%, the charge of bound
+%counterions
+cancels the influence of negatively charged lipids and the increase in PC headgroup
+order parameters is not observed.% \cite{NMRlipidsIV}.
+
+%Lipid17/Dang ion simulations 
+%while in simulations with \ce{K+} counterions and in
+
+%The trend is observed with both force fields in simulations
+%the counterion binding affinity is weaker and the
+%order parameters increase as in experiments (Fig~\ref{fig:delta_ordPar_NaCl_PC-PS_mix_PC}).  
+%In contrast, the order parameters are unchanged or decrease 
+%and the P--N vector tilts more perpendicular to the membrane interface.
+%in 
+%The result can be explained by the higher binding affinity of sodium in the
+%Lipid17/Dang ion simulations (Fig~\ref{fig:delta_ordPar_NaCl_PC-PS_mix_PC}), which
+%compensates the negative charge of PS lipids at the interface.
+%This is corrected in ECC-lipids, 
+%which only slightly underestimate the slopes of the order parameters 
+%when \ce{Na+} counterions are used (Fig.~\ref{fig:delta_ordPar_NaCl_PC-PS_mix_PC}). 
+%According to the electrometer concept, the decrease of PC head group order parameters
+%correlates with the amount of bound cations to lipid bilayer \cite{akutsu81,altenbach84,seelig87,roux90,catte16,NMRlipidsIV}.
+%Second, we monitor the response of PC headgroup order parameters to the additional
+%amount of \ce{NaCl} and \ce{KCl} in POPC:POPS (5:1) mixtures.
+
+%\subsection{Molecular interaction and binding affinities of K$^+$ and Na$^+$ cations to mixed POPC:POPS (5:1) membrane} 
+%\label{sec:K_Na_added} 
 \begin{figure*}[!tbp] 
@@
 \end{figure*}
-
-%First, we monitor the response of POPC headgroup order parameters to the increasing amount of POPS (Fig.~\ref{fig:delta_ordPar_NaCl_PC-PS_mix_PC}).
-According to the electrometer concept, the head group order parameters of PC lipids increase
-with the increasing amount of negatively charged PS lipids in experiments, because the headgroup P--N dipole
-tilts more parallel to the membrane interface \cite{seelig87,scherer87}.
-However, in simulations with strong counterion binding affinity, the charge of bound
-counterions cancel the influence of negatively charged lipids and the increase in PC headgroup
-order parameters is not observed \cite{NMRlipidsIV}. This is the case in 
-Lipid17/Dang ion simulations with \ce{Na+} counterions,
-while in simulations with \ce{K+} counterions and in
-ECC-lipids simulations both counterions,
-%The trend is observed with both force fields in simulations
-the counterion binding affinity is weaker and the
-order parameters increase as in experiments (Fig~\ref{fig:delta_ordPar_NaCl_PC-PS_mix_PC}).  
-%In contrast, the order parameters are unchanged or decrease 
-%and the P--N vector tilts more perpendicular to the membrane interface.
-%in 
-%The result can be explained by the higher binding affinity of sodium in the
-%Lipid17/Dang ion simulations (Fig~\ref{fig:delta_ordPar_NaCl_PC-PS_mix_PC}), which
-%compensates the negative charge of PS lipids at the interface.
-%This is corrected in ECC-lipids, 
-%which only slightly underestimate the slopes of the order parameters 
-%when \ce{Na+} counterions are used (Fig.~\ref{fig:delta_ordPar_NaCl_PC-PS_mix_PC}). 
-According to the electrometer concept, the decrease of PC head group order parameters
-correlates with the amount of bound cations to lipid bilayer \cite{akutsu81,altenbach84,seelig87,roux90,catte16,NMRlipidsIV}.
-%Second, we monitor the response of PC headgroup order parameters to the additional
-%amount of \ce{NaCl} and \ce{KCl} in POPC:POPS (5:1) mixtures.
-In experiments, the addition of K$^+$ ions led to a very modest decrease of POPC headgroup order parameters in POPC:POPS (5:1) mixture,
-while the decrease was more pronounced with Li$^+$ ions and was strongest with divalent Mg$^{2+}$ and Ca$^{2+}$ ions~\cite{roux90},
-suggesting that the binding affinity increases in order K$^{+}$ $<$ Li$^{+}$  $<$ Mg$^{2+}$  $<$ Ca$^{2+}$.
-The small changes of POPC headgroup order parameters with increasing amount of KCl are close to experiments
-%and the change in P--N vector is negligible
-in both force fields (Fig.~\ref{fig:delta_ordPar_monoval_PCPS}).
-The slightly overestimated change of the $\alpha$-carbon
-order parameter in the Lipid17/Dang simulations may be due to the
-deeper penetration of \ce{K+} into the bilayer (Fig.~\ref{fig:delta_ordPar_monoval_PCPS}).
-Experimental data with the additional amount of \ce{Na+} in POPC:POPS (5:1) mixture is not available,
-%we compared the simulation results to the experimental data with the additional amounts of
-but \ce{Li+} and \ce{K+} results give the lower and the upper bounds, respectively, to the sodium binding
-affinity (Fig.~\ref{fig:delta_ordPar_monoval_PCPS}). In ECC-lipids simulations,
-the response of POPC headgroup order parameters to the additional sodium is close to the
-experimental results for lithium, while in Lipid17/Dang simulations the response is larger.
-% The following conclusion is already in the previous sentence, so I commented it out
-Overall, the results suggest that the \ce{Na+} binding affinity is
-% (Fig. \ref{fig:POPS-counterions-dens})
-%and the tilting of P--N vector more perpendicular to the membrane surface (Fig. \ref{fig:delta_ordPar_NaCl_PCPS})
-clearly overestimated in Lipid17/Dang simulations and slightly overestimated in ECC-lipids simulations,
-while the binding affinity of potassium is better described by both force fields.
-
-%\subsection{Molecular interaction and binding affinities of K$^+$ and Na$^+$ cations to mixed POPC:POPS (5:1) membrane} 
-%\label{sec:K_Na_added}
-
-\begin{figure}[tbp!] 
-  \centering 
-  \includegraphics[height=\figheightsmall]{../img/ecc_pops/order_parameters_changes_ecc-lip_L14_A-B-PN-COO_POPC_kcl.pdf} 
-  \includegraphics[height=\figheightsmall]{../img/ecc_pops/order_parameters_changes_ecc-lip_L14_A-B-PN-COO_POPC_nacl.pdf} 
-  \includegraphics[height=\figheightsmall]{../img/ecc_pops/order_parameters_changes_ecc-lip_L14_A-B-PN-COO_POPS_kcl.pdf} 
-  \includegraphics[height=\figheightsmall]{../img/ecc_pops/order_parameters_changes_ecc-lip_L14_A-B-PN-COO_POPS_nacl.pdf} 
-  \caption{\label{fig:delta_ordPar_monoval_PCPS}
-    Changes of the head group order parameters, and the angles of P--N and \ce{C_\beta-C_\gamma} (stars) vectors 
-    with respect to the membrane normal of POPC (left) and POPS (right) in a POPC:POPS (5:1) bilayer 
-    as a function of \ce{KCl} and \ce{NaCl} concentration from ECC-lipids and Lipid17/Dang simulations 
-    compared with experimental values from Ref. \citenum{roux90} (signs from Ref. \citenum{ferreira16}) at 298 K.
-    The y-axis for the $\alpha$-carbon results of POPS (middle right) is transferred
-    with the same value for both order parameters such that the lower order
-    parameter value from pure POPS is at zero to correctly illustrate the significant forking.
-    Because experimental data with \ce{NaCl} is not available for POPC, the data for KCl and \ce{LiCl} (dashed line, left)
-    are shown as lower and upper bounds, respectively, for the response to \ce{NaCl}.
-    Error bars are not visible for ECC-lipids simulations because they are smaller than the point size.
-  }
-\todo{J: Will improve this figure with panels A, B, C and D 
-and put POPC together, POPS together 
-and remove the duplicate y-axis titles 
-and make a common legend distinguishing different experiments better. }
-\todo{The scale in left and right columns (response to KCl and NaCl) should be the same (not necessarily the same scale for POPC and POPS though).}
-\end{figure}
 
@@
   \caption{ \label{fig:cacl_complexes} 
-    \textbf{Right top:} Percentages of the bound Ca$^{2+}$ 
+    \textbf{Top:} Percentages of the bound Ca$^{2+}$ 
     in exclusive contact with the given oxygen moiety
@@
         For numerical values, see table~\ref{tab:binding}.
-    \textbf{Right bottom:} Relative probabilities of \ce{Ca^{2+}} ions to coordinate with a certain number of lipids
+    \textbf{Bottom:} Relative probabilities of \ce{Ca^{2+}} ions to coordinate with a certain number of lipids
     in pure POPC bilayer with 350 mM CaCl$_2$ and in POPC:POPS (5:1) mixture with 400 mM CaCl$_2$.  
@@
   }
-  \todo{The x-axis label should be changed to ''number of lipids directly bound to Ca$^{2+}$''.
-    The current label has currently another meaning in the text.}
+  \todo{Bottom: The x-axis label should be changed to ''number of lipids directly bound to Ca$^{2+}$''.
+    The current label has another meaning in the text now.}
+  \todo{Bottom: The labels could be more clear. Maybe ''PC in POPC:POPS(5:1)'' ...}
 \end{figure}
@@
 \section{Conclusions} 
-
-\todo{J: I have moved several paragraphs from the body to here and added a bit of glue. This section still requires to be written properly.  }
-
-We have developed a classical MD model of POPS, ECC-POPS, 
-which accurately describes features of both pure POPS membrane and its mixes with POPC. 
-This was achieved by applying ECC which accounts for electronic polarization. 
-Applying ECC to POPS lipids improves cation binding affinity to membranes
-mixed with POPC and also the response of PS headgroup to the bound cations.
-This was compared to accurate NMR measurements of the changes of headgroup order parameter within the ``electrometer concept''. 
-The advantage of the NMR order parameters used in this work is the direct connection between
-measured and simulation data, which enables more unambiguous evaluation of the force field quality \cite{catte16,ollila16}
-than more indirect comparison in previous studies using 2D infrared spectroscopy \cite{valentine18}, NMR chemical shifts and
-rotational-echo double-resonance (REDOR) experiments \cite{hallock18} or fluorescent and vibrational sum frequency spectroscopy \cite{melcrova16}.
-Because the ECC-lipids model gives the best agreement with the experimental headgroup order parameters
-with various ion conctenrations among the available models, we believe that it also gives the currently best available
-interpretation of the experimental data.
-
-In conclusion, applying ECC to Amber Lipid17 POPS force field parameters improves its
-interactions with both monovalent ions and POPC headgroups when compared with the NMR order parameter data.
-Although the ECC-lipids model gives the most realistic picture of calcium binding to
-lipid bilayers containing PS lipids with respect to NMR order parameters, it does not
-correctly capture the rapid changes with small concentrations. 
-Therefore, further force field optimization is required for even more accurate MD simulations of cations
-in the vicinity of negatively charged lipid bilayers.
+We have applied ECC to implicitly include electronic polarization to the Amber Lipid17 force field
+parameters of negatively charged POPS lipid.
+Because PS headgroup bears a unit negative charge, we apply the scaling factor of 0.75,
+derived for monovalent ions~\cite{leontyev09}, to the partial charges
+in the headgroup, glycerol backbone and carbonyl regions of POPS.
+Following our previous work for zwitterionic POPC lipids, the 
+Lennard-Jones $\sigma$ parameters of the same POPS segments were
+scaled by a factor of 0.89 to optimize the hydration properties of the bilayer~\cite{melcr18}.
+Similarly to other state of the art lipid models, the created ECC-POPS parameters give
+lipid bilayer dimensions and acyl chain structures in agremeement with experiments,
+but leaves room for improvement for the headgroup structure when validated against
+NMR and scattering experiments~\cite{botan15,ollila16}.
+Nevertheless, ECC-lipid parameters describe cation (Na$^+$ and Ca$^{2+}$) binding affinities
+and their interactions with PS headgroups better than other available lipid force fields,
+when validated using the headgroup order parameters and ``electrometer concept''~\cite{NMRlipidsIV}.
+There is, however, room for improvement in capturing the sensitive headgroup responses to the
+small concentrations of CaCl$_2$.
+
+%Although the ECC-lipids model gives the most realistic picture of calcium binding to
+%lipid bilayers containing PS lipids with respect to NMR order parameters,
+
+%The scaling factor for the charge
+%We have developed a classical MD model of POPS, ECC-POPS, 
+%which accurately describes features of both pure POPS membrane and its mixes with POPC. 
+%This was achieved by applying ECC which accounts for electronic polarization. 
+%Applying ECC to POPS lipids improves cation binding affinity to membranes
+%mixed with POPC and also the response of PS headgroup to the bound cations.
+
+In our ECC-lipids simulation, Ca$^{2+}$ ions bind twice more likely to carboxylate groups
+of PS headgroups than to phosphate groups, while binding to carbonyls is almost negligible.
+Binding of Ca$^{2+}$ ions to more than two POPS lipids simultaneously is very rare in ECC-lipids simulations.
+Because ECC-lipids parameters give the best results in direct comparison with NMR order parameters, %used in this work is the direct connection between
+we believe that our results help in resoving controversial interpretations of more indirect experiments
+using different force fields~\cite{melcrova16,valentine18,hallock18}.
+%measured and simulation data, which enables more unambiguous evaluation of the force field quality \cite{catte16,ollila16}
+%than more indirect comparison in previous studies using 2D infrared spectroscopy \cite{valentine18}, NMR chemical shifts and
+%rotational-echo double-resonance (REDOR) experiments \cite{hallock18} or fluorescent and vibrational sum frequency spectroscopy \cite{melcrova16}.
+%Because the ECC-lipids model gives the best agreement with the experimental headgroup order parameters
+%with various ion conctenrations among the available models, we believe that it also gives the currently best available
+%interpretation of the experimental data. 
+Furthermore, our results pave the way for more realistic MD simulations of anionic biological membranes,
+and demonstrate the usefulness of ECC also for charged lipids.
+\todo{Maybe Pavel could extend this by mentioning something that this is yet another case where ECC works and/or
+  mentioning some concrete biological examples where anionic membrane/calcium interactions are important?}
+
+%Therefore, further force field optimization is required for even more accurate MD simulations of cations
+%in the vicinity of negatively charged lipid bilayers.
 

### 2019-05-31 Manuscript/manuscript.tex
Update text

@@
 and remove the duplicate y-axis titles 
-and make a common legend distinguishing different experiments better. }
+and make a common legend distinguishing different experiments better. } \\
 \todo{The scale in left and right columns (response to KCl and NaCl) should be the same (not necessarily the same scale for POPC and POPS though).}
@@
 and will erase the duplicate y-axis description 
-and will add ticks to the right side of the plotting area. }
+and will add ticks to the right side of the plotting area. } \\
 \todo{The scales should be the same in the left and right panel.}
@@
 lipid mixtures and ionic conditions~\cite{NMRlipidsIV,roux90}.
-Upon increasing the POPC content or monovalent ion concetration in experiments, 
-the headgroup order parameters of POPS are almost unchanged
+In experiments, the headgroup order parameters of POPS are almost unchanged upon increasing the POPC content or monovalent ion concetration
 (\ref{fig:delta_ordPar_NaCl_PC-PS_mix_PS} and Figs.~\ref{fig:delta_ordPar_monoval_PCPS}).
@@
 The decrease of the PC headgroup order parameters in this mixture upon addition of \ce{CaCl2} 
-were overestimated by the Lipid17/Dang simulations (also shown in Fig.~\ref{fig:delta_ordPar_CaCl_PCPS})
+were overestimated by the Lipid17/Dang simulations (Fig.~\ref{fig:delta_ordPar_CaCl_PCPS})
 and all other tested force fields except CHARMM36 with the recently introduced NBfix for calcium \cite{kim16},
@@
 than those typically observed in simulations with other force fields \cite{javanainen17,melcr18,NMRlipidsIV},
-suggesting that the ECC accelerates the equlibration of ions at lipid bilayer interface.
+suggesting that the ECC accelerates the equilibration of ions at lipid bilayer interface.
 Therefore, our 1 $\mu$s simulations seem to be sufficiently long for the ECC-lipids simulations, because
@@
   \todo{Bottom: The x-axis label should be changed to ''number of lipids directly bound to Ca$^{2+}$''.
-    The current label has another meaning in the text now.}
+    The current label has another meaning in the text now.} \\
   \todo{Bottom: The labels could be more clear. Maybe ''PC in POPC:POPS(5:1)'' ...}
@@
 Similarly to other state of the art lipid models, the created ECC-POPS parameters give
-lipid bilayer dimensions and acyl chain structures in agremeement with experiments,
+lipid bilayer dimensions and acyl chain structures in agreement with experiments,
 but leaves room for improvement for the headgroup structure when validated against

### 2019-05-31 Manuscript/manuscript.tex
Remove comments from tex files

@@
 
-%Various interaction schemes have been proposed for the negatively charged
-%lipids with ions and other lipids. 
 Spectroscopic experiments give detailed information about the
@@
 but the results strongly depend on the force field parameters \cite{boettcher11,kucerka14,melcrova16,hallock18,valentine18}.
-% smoother connection between previous and following statements
 For instance, the moieties of PS lipids that interact with calcium cations vary greatly.
@@
 The overall improvement of the force field accuracy upon applying ECC
-%not only describes the interactions of PS with ions more accurately, 
-%but also improves the structure of the PS head group 
-%and its interactions with PC in mixed lipid bilayers. 
-%The results
 paves the way to more realistic MD simulations 
@@
 for a wide range of applications.
-
-%Here, we demonstrate that the same approach 
-%significantly improves the description of calcium binding 
-%also to the negatively charged PC:PS (5:1) bilayers. 
-%In correspondence with the Ref.~\citenum{NMRlipidsIV},
-%we also use the changes of the head group order parameters $S^\alpha$ and $S^\beta$ 
-%in PC and PS lipids 
-%to quantify the binding affinity of the calcium cations
-%and the induced structural changes
-%(Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}). 
-%To directly connect to the previous works \citep{catte16, NMRlipidsIV}
-%and to mark the improvement after treating electronic polarizability, 
-%we show simulation results from the newly developed ECC-lipids and also from Lipid17 \citep{lipid17-future}.
-
-%An empirical scaling factor, which reduces electrostatic forces between molecules, 
-%is a well known concept in molecular modelling;
-%such an approach was adopted in the early simulations of lipid bilayers \cite{egberts94, berendsen1996},
-%in force fields employing explicit polarization \cite{lemkul2016empirical},
-%or in the simulations of ionic liquids \cite{Bhargava2007}. 
-%%An improvement of permeability of charged solutes in the gramicidin A channel 
-%%is observed with a force field using charge equilibration to model polarizability. \cite{lucas2012} 
-%At the same time, 
-%the implicit inclusion of electronic polarizability to lipid models
-%via charge re-scaling using electronic continuum correction (ECC) 
-%substantially improved the interactions between cations and zwitterionic lipids \cite{melcr18}.
 
@@
 %being closer to the experimental value, 
-%In previous study \cite{NMRlipidsIV}, the Lipid17 force field
 with {\AA}qvist ions~\cite{NMRlipidsIV}.
-%gave a 
-%which is probably due to the weaker counterion binding affinity of {\AA}qvist ions to phospholipids. \cite{catte16, NMRlipidsIV}
-However, the Dang ions %with more realisitic bulk properties
-are used here to analyze the effect of ECC to the properties of a POPS lipid bilayer because
+However, the Dang ions are used here to analyze the effect of ECC to the properties of a POPS lipid bilayer because
 {\AA}qvist ions produce known artifacts, like artificial aggregation of ions at larger salt concentrations \cite{kohagen16, chen07, NMRlipidsIV}.
@@
 
-%Overall,
-%the head group and acyl chain order parameters within ECC-lipids
-%are in a good agreement with the experimental values 
-%as shown in Fig.~\ref{simVSexpNOions_POPS}. 
-In conclusion,
-%the acyl chain order parameters in particular agree with the experimental data within error bars.
-%The order parameters of the head groups are at an accuracy comparable to 
-the stuctural quality of ECC-POPS is similar to the other currently available
+In conclusion, the stuctural quality of ECC-POPS is similar to the other currently available
 lipid force fields \cite{botan15, catte16, Pluhackova2016, nmrlipids_proj4},
-%In particular, ECC-POPS is closer to the experimental order parameters 
-%than Lipid17 with either Dang \cite{smith94,chang1999,dang2006} 
-%or {\AA}qvist ions \cite{aqvist90}. 
-%Although the ECC-POPS model reproduces the lipid bilayer dimensions and
 giving good agreement with experiments for acyl chain conformations and bilayer dimensions,
@@
 conformations.
-%e.g. in the parameters of the associated dihedral angles. 
-%\cite{botan15, ollila16, Pluhackova2016, NMRlipidsIV}
 
 \subsection{Binding of counterions to POPS and POPC and interactions between their head groups}
-
 Binding affinity of monovalent ions to lipid bilayers depends strongly on
@@
 Because the evaluation of monovalent ion binding affinity to bilayers with
-%with negatively charged
 PS lipids is complicated by the lack of ion-free state (counterions are always present with negatively charged lipids),
-%Therefore,
 we evaluate the counterion binding affinity to POPC:POPS mixtures also by monitoring
 the response of POPC headgroup order parameters to the increasing mole fraction of PS lipids~\cite{NMRlipidsIV}.
-%and
-%to the increasing monovalent ion concetrations. %, as 
-%to a lipid bilayer containing PS in simulations using two approaches.
-%Following our previous work \cite{NMRlipidsIV},
 
@@
 Experimental data with the additional amount of \ce{Na+} in POPC:POPS (5:1) mixture is not available,
-%we compared the simulation results to the experimental data with the additional amounts of
 but \ce{Li+} and \ce{K+} results give the lower and the upper bounds, respectively, to the sodium binding
@@
 experimental results for lithium, while in Lipid17/Dang simulations the response is larger.
-% The following conclusion is already in the previous sentence, so I commented it out
 Overall, the results suggest that the \ce{Na+} binding affinity is
-% (Fig. \ref{fig:POPS-counterions-dens})
-%and the tilting of P--N vector more perpendicular to the membrane surface (Fig. \ref{fig:delta_ordPar_NaCl_PCPS})
 clearly overestimated in Lipid17/Dang simulations and slightly overestimated in ECC-lipids simulations,
 while the binding affinity of potassium is better described by both force fields.
-%First, we monitor the response of POPC headgroup order parameters to the increasing amount of POPS (Fig.~\ref{fig:delta_ordPar_NaCl_PC-PS_mix_PC}).
 This conclusion is also supported by the POPC headgroup responses to the
 increasing amount of POPS in different simulations (Fig~\ref{fig:delta_ordPar_NaCl_PC-PS_mix_PC}).
-In experiments,
-the head group order parameters of PC lipids increase upon addition of
-% with the increasing amount of
+In experiments, the head group order parameters of PC lipids increase upon addition of
 negatively charged PS lipids, as predicted by the electrometer concept~\cite{seelig87,scherer87}.
-%, because the headgroup P--N dipole
-%tilts more parallel to the membrane interface \cite{seelig87,scherer87}.
 This is the case also in simulations with the exception of
-Lipid17/Dang with \ce{Na+} counterions, where the
-%but not with \ce{Na+} counterions,
-%However, in simulations with
-stronger counterion binding affinity
-%, the charge of bound
-%counterions
+Lipid17/Dang with \ce{Na+} counterions, where the stronger counterion binding affinity
 cancels the influence of negatively charged lipids and the increase in PC headgroup
@@
 
-%Lipid17/Dang ion simulations 
-%while in simulations with \ce{K+} counterions and in
-
-%The trend is observed with both force fields in simulations
-%the counterion binding affinity is weaker and the
-%order parameters increase as in experiments (Fig~\ref{fig:delta_ordPar_NaCl_PC-PS_mix_PC}).  
-%In contrast, the order parameters are unchanged or decrease 
-%and the P--N vector tilts more perpendicular to the membrane interface.
-%in 
-%The result can be explained by the higher binding affinity of sodium in the
-%Lipid17/Dang ion simulations (Fig~\ref{fig:delta_ordPar_NaCl_PC-PS_mix_PC}), which
-%compensates the negative charge of PS lipids at the interface.
-%This is corrected in ECC-lipids, 
-%which only slightly underestimate the slopes of the order parameters 
-%when \ce{Na+} counterions are used (Fig.~\ref{fig:delta_ordPar_NaCl_PC-PS_mix_PC}). 
-%According to the electrometer concept, the decrease of PC head group order parameters
-%correlates with the amount of bound cations to lipid bilayer \cite{akutsu81,altenbach84,seelig87,roux90,catte16,NMRlipidsIV}.
-%Second, we monitor the response of PC headgroup order parameters to the additional
-%amount of \ce{NaCl} and \ce{KCl} in POPC:POPS (5:1) mixtures.
-
-%\subsection{Molecular interaction and binding affinities of K$^+$ and Na$^+$ cations to mixed POPC:POPS (5:1) membrane} 
-%\label{sec:K_Na_added} 
 \begin{figure*}[!tbp] 
@@
 (\ref{fig:delta_ordPar_NaCl_PC-PS_mix_PS} and Figs.~\ref{fig:delta_ordPar_monoval_PCPS}).
-%while calcium leads to relatively large changes with low concentrations followed by a rapid saturation below 100mM (Fig. \ref{fig:delta_ordPar_CaCl_PCPS}) \cite{scherer87,roux90}.
-%The interaction with \ce{K^+}, which binds very weakly to both neutral and negatively charged membranes, 
-%renders a qualitatively different response of the order parameter $S^\beta$ in POPS 
-%in the mixed negatively charged bilayers compared to the neutral bilayers. 
-%While the order parameter $S^\beta$ increases for both \ce{Na^+} and \ce{Ca^{2+}},
-%it decreases in the presence of \ce{K^+}
-%(\ce{KCl}:  Fig.~\ref{fig:delta_ordPar_KCl_PCPS}, 
-% \ce{NaCl}: Fig.~\ref{fig:delta_ordPar_NaCl_PCPS}, and
-%\ce{CaCl2}: Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}). 
 In Lipid17/Dang simulations, the changes of POPS headgroup order parameters with increasing amount of
 both POPC or monovalent salts (\ce{KCl} and \ce{NaCl}) are overestimated (\ref{fig:delta_ordPar_NaCl_PC-PS_mix_PS} and Figs.~\ref{fig:delta_ordPar_monoval_PCPS}).
-%These structural changes can relate to variations in the interactions between PC and PS head groups 
-%and also to the changing amount of counterions and PS in the composition. 
-%Also, the PS headgroup order parameters from simulations with Lipid17/Dang model
-%are too sensitive to an additional amount of
-%(Fig.~\ref{fig:delta_ordPar_monoval_PCPS}).
-Similar results %The overestimation of the order parameter changes upon increasing amount of POPC in a bilayer 
-also for other force fields are reported elsewhere~\cite{NMRlipidsIV}.
+Similar results also for other force fields are reported elsewhere~\cite{NMRlipidsIV}.
 In ECC-lipids simulations, the response of POPS headgroup order parameters to both POPC and monovalent salt concentration (\ce{KCl} and \ce{NaCl})
@@
 
-% where the changes in the PS headgroup orientation are more modest (Fig.~\ref{fig:delta_ordPar_NaCl_PC-PS_mix_PC}).
-%In particular, 
-%the dependence of the PS headgroup order parameters from ECC-lipids simulations
-%to increasing amount of POPC, \ce{KCl} or \ce{NaCl}
-%is weaker and closer to experiments 
-%(Fig.~\ref{fig:delta_ordPar_monoval_PCPS}).
-%In contrast, ECC-lipids with ECC-ions capture the different response 
-%of the order parameters $S^{\beta}$, $S^{\alpha _1}$ and $S^{\alpha _2}$ in POPS 
-%to various concentrations of \ce{NaCl} and \ce{KCl} in a good agreement with the experiments
-%(\ce{KCl}:  Fig.~\ref{fig:delta_ordPar_KCl_PCPS}, 
-% \ce{NaCl}: Fig.~\ref{fig:delta_ordPar_NaCl_PCPS}). 
-%Such a dramatic improvement in the structural parameters 
-%demonstrates that including electronic polarization in phospholipids is crucial 
-%even for interactions with soft cations like \ce{K^+}.
-
-% This§: 
-% adsorption and affinities of Na and K to PC:PS memb
-% from densities and rel.surf.excess
-% 
-%in  Fig.~\ref{fig:POPS-counterions-dens} (only counterions)
-%and Fig.~\ref{fig:nacl-dens_PCPS}        (additional salt concentrations), 
-%
 We estimate the influence of negatively charged POPS on the counterion binding affinity
-% can be estimated 
-%by calculating the amounts of phospholipids per bound \ce{Ca^{2+}} cation, $\xi _{Ca^{2+}}$,
 by comparing the relative surface excesses with respect to water, $\Gamma ^{w} _{\rm ion}$,
@@
 The presence of $\sim$17\% POPS in a POPC bilayer increases the
-%relative surface excess of sodium with respect to water from negative
 value to $\Gamma ^{w} _{\rm Na}=0.092 \pm 0.005 \mathrm{nm}^{-2}$ 
@@
 increase in binding affinity due to PS lipids would not be affected.
-%Interestingly enough, \ce{K^+} maintains negative values of $\Gamma^{w}_{K}$ even for the negatively charged PC:PS~(5:1) bilayer.
-%This is also evident from the density plots in  Fig.~\ref{fig:nacl-dens_PCPS}
-%showing that the density profile of \ce{K^+} along the membrane normal 
-%decays clearly faster towards the membrane centre compared to the density of water. 
-%In contrast, the density profile of \ce{Na^+} has a shoulder 
-%at the region of the phosphate groups,
-%which demonstrates that sodium cations adsorb weakly to the membrane interface compared to the aquaeous solution
-%by yielding a small positive value of $\Gamma^{w}_{Na}$.
 
@@
 \label{section:lip-ion_ca}
-
-%The headgroup order parameters of PC lipids decrease more upon addition of \ce{CaCl2}
-%in POPC:POPS (5:1) mixtures than in pure PC bilayers due to the higher calcium binding
-%affinity to the negatively charged lipid bilayers \cite{akutsu81,altenbach84,roux90,NMRlipidsIV}.
 Our recent work demonstrates that the POPC headgroup order parameters measured
@@
 indicating that the lower binding affinity than in Lipid17/Dang simulations is more realistic (Fig.~\ref{fig:cacl-dens_PCPS}).
-%However, the steep decrease of the order parameters at low calcium concentrations,
-%observed especially in negatively charged bilayers \cite{borle85,macdonald87,roux90},
-%may not be fully captured by the ECC-lipids simulations.
 Therefore, we use ECC-lipids simulations to estimate the influence of negative charged POPS on
 calcium binding affinity to lipid bilayers.
-%containing PS lipids is illustrated in
-%ECC-lipids simulations by the increasing
 The relative surface excess of calcium with respect to water 
@@
 % SAMULI: I separated the figure below, because it was too small to be read when combined with others.
-% It might be too much to all of these into the same.
 
@@
     calculated from ECC-lipids simulation of POPC:POPS (5:1) bilayer with 409 mM CaCl$_2$. 
-%	The areas represent populations of bound \ce{Ca^{2+}} 
-%	only to the given moieties exclusively (shades of orange and red)
-%	or their combinations (CHANGE:blue)
-%	separately for PC, PS, or PC and PS simultaneously. 
-%	The areas sum up to total populations at the respective lipids. 
 	In the case of simultaneous binding to both PC and PS,
 	the percentages refer to the moieties in PS.
-	% i.e., PC can interact with any considered moiety.
         For numerical values, see table~\ref{tab:binding}.
@@
 tilts to the opposite direction by $11^\circ$ in the same system.
-%In Lipid17/Dang simulations, the changes in both headroup orientations and order parameters
-%are larger and less systematic for the PS headgroup, especially with lower concentrations (Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}).
-
-%SAMULI: We will not go to the spin relaxation time analysis in this work
-%and average angles do not give information about dynamics
-%
-%It has been pointed out in previous experimental works
-%that the conformational flexibility of PS head groups 
-%is different from other lipids \cite{browning80}. 
-%While the static shielding tensor of PS is not substantially different from PC or PE, 
-%the $^{31}$P~$T_1$ relaxation times are significantly shorter for PS. 
-%It is suggested that the PS head group is more rigid compared to the other phospholipids, 
-%which is also corroborated by a larger chemical shielding anisotropy of PS \cite{browning80}.
 
@@
 
-%Although the ECC-lipids model gives the most realistic picture of calcium binding to
-%lipid bilayers containing PS lipids with respect to NMR order parameters,
-
-%The scaling factor for the charge
-%We have developed a classical MD model of POPS, ECC-POPS, 
-%which accurately describes features of both pure POPS membrane and its mixes with POPC. 
-%This was achieved by applying ECC which accounts for electronic polarization. 
-%Applying ECC to POPS lipids improves cation binding affinity to membranes
-%mixed with POPC and also the response of PS headgroup to the bound cations.
-
 In our ECC-lipids simulation, Ca$^{2+}$ ions bind twice more likely to carboxylate groups
@@
 Binding of Ca$^{2+}$ ions to more than two POPS lipids simultaneously is very rare in ECC-lipids simulations.
-Because ECC-lipids parameters give the best results in direct comparison with NMR order parameters, %used in this work is the direct connection between
+Because ECC-lipids parameters give the best results in direct comparison with NMR order parameters, 
 we believe that our results help in resoving controversial interpretations of more indirect experiments

### 2019-06-15 Manuscript/manuscript.tex
Tiago, to do list

@@
 Here, we measure also the C--H bond order parameters of the acyl chain region using natural abundance $^{13}$C NMR.
-We first determined the order parameters of each peak corresponding to different C--H bonds detected
-in the first dimension (Fig.~\ref{R-PDLF}) from the dipolar slices in the indirect dimension (Fig.~\ref{R-PDLFslices}).
-Because the exact assignment of these peaks was not possible, we assigned the
-acyl chain order parameters to mimic the profile predicted by the MD simulations (Fig.~\ref{simVSexpNOions_POPS}).
-\todo{Tiago should check and update this and figures in SI.}
+The assignment of peaks to acyl chain carbon sites was done based on previous work {\cite{ferreira13}} and is shown in Figs.~\ref{INEPT}~and~\ref{R-PDLF}.
+The corresponding dipolar splittings for each of the assigned peaks are shown in Fig.~\ref{R-PDLFslices}.
+Because the complete assignment of the peaks in the crowded spectral region between 29-31 ppm was not possible due to the overlapp of different carbons, we partially assigned the acyl chain order parameters of this region to mimic the profile predicted by the MD simulations (Fig.~\ref{simVSexpNOions_POPS}).
 
@@
   \caption{\label{fig:delta_ordPar_monoval_PCPS}
-    Changes of the head group order parameters, and the angles of P--N and \ce{C_\beta-C_\gamma} (stars) vectors 
+    Changes of the head group order parameters, and the angles of P--N and ${\rm{C}}_\beta-{\rm{C}}_\gamma$ (stars) vectors
     with respect to the membrane normal of POPC (left) and POPS (right) in a POPC:POPS (5:1) bilayer 
@@
     Because experimental data with \ce{NaCl} is not available for POPC, the data for KCl and \ce{LiCl} (dashed line, left)
-    are shown as lower and upper bounds, respectively, for the response to \ce{NaCl}.
-    Error bars are not visible for ECC-lipids simulations because they are smaller than the point size.
-  }
+    are shown as lower and upper bounds, respectively, for the response to \ce{NaCl}. Error bars are not visible for ECC-lipids simulations because they are smaller than the point size.}
 \todo{J: Will improve this figure with panels A, B, C and D 

### 2019-06-15 Manuscript/manuscript.tex
Tiago, to do list

@@
 the headgroup and glycerol backbone region of a POPS lipid bilayer are already available in the literature \cite{kucerka14,NMRlipidsIV}.
-Here, we measure also the C--H bond order parameters of the acyl chain region using natural abundance $^{13}$C NMR.
-The assignment of peaks to acyl chain carbon sites was done based on previous work {\cite{ferreira13}} and is shown in Figs.~\ref{INEPT}~and~\ref{R-PDLF}.
-The corresponding dipolar splittings for each of the assigned peaks are shown in Fig.~\ref{R-PDLFslices}.
-Because the complete assignment of the peaks in the crowded spectral region between 29-31 ppm was not possible due to the overlapp of different carbons, we partially assigned the acyl chain order parameters of this region to mimic the profile predicted by the MD simulations (Fig.~\ref{simVSexpNOions_POPS}).
+Here, we measure also the C--H bond order parameters of the acyl chain region from a sample of POPS multi-lamellar vesicles by means of a $^1$H--$^{13}$C dipolar recoupling 2D NMR experiment.
+More specifically, as described in detail in the Methods section, we performed R-PDLF 2D spectroscopy~\cite{dvinskikh04} to obtain spectra correlating chemical shifts (direct dimension) and C--H bond dipolar couplings (indirect dimension) from which the C--H bond order parameters defined in eq.~\ref{OPequation} can be determined.
+The assignment of the peaks in the $^{13}$C chemical shift dimension to acyl chain carbon sites was done based on previous work {\cite{ferreira13}} and is shown in Figs.~\ref{INEPT}~and~\ref{R-PDLF}.
+The dipolar splittings for the assigned carbons that were used to calculate their C--H bond order parameter magnitudes are shown in Fig.~\ref{R-PDLFslices}.
+Because the complete assignment of the peaks in the crowded spectral region between 29-31 ppm was not possible due to the chemical shift overlap of different carbons (~\ref{R-PDLF}), we partially assigned the acyl chain order parameters of this region to mimic the profile predicted by the MD simulations (Fig.~\ref{simVSexpNOions_POPS}).
 

### 2019-06-20 Manuscript/manuscript.tex
Updates made by Pavel

--- 2019-05-31 Manuscript/manuscript.tex
@@
 \affiliation{Institute of Organic Chemistry and Biochemistry, 
-Academy of Sciences of the Czech Republic,  
+Czech Academy of Sciences,  
 Prague 6, Czech Republic} 
@@
 \affiliation{Institute of Organic Chemistry and Biochemistry, 
-Academy of Sciences of the Czech Republic,  
+Czech Academy of Sciences,  
 Prague 6, Czech Republic} 
-\alsoaffiliation{Department of Physics, Tampere University of Technology, P.O. Box 692, FI-33101 
-Tampere, Finland} 
 \author{O. H. Samuli Ollila} 
@@
 \title[] 
-{Improved Cation Binding to Lipid Bilayer with Negatively Charged POPS by Effective Inclusion of Electronic Polarization}
+{Improved Description of Cation Binding to Lipid Bilayer with Negatively Charged POPS by Effective Inclusion of Electronic Polarization}
   %Detailed structure of negatively charged membranes 
@@
 induce protein aggregation \cite{zhao04,gorbenko06} and membrane fusion \cite{??}.
-Because these functions are also often regulated by biological cations \cite{leventis10},
-the detailed interactions between PS lipids and biological cations, like calcium, are under great interest.
-\todo{Pavel can probably improve and complement this}
+As these functions are also often regulated by ions \cite{leventis10},
+a detailed understanding of interactions between negatively charged lipids and biologically relevant cations such as calcium is of a great importance.
+As a recent example, we have shown by a combination of spectroscopic experiments, biochemical essays, and molecular simulations, that binding of
+calcium to PIP2 lipids inhibits their recognition by a phospolipase C, which demonstrates that not only proteins but also lipids can get involved in calcium 
+signaling \cite{DOI: 10.1021/jacs.6b11760}.
 
@@
 where  $\epsilon_{el} = 1.78$ is the high-frequency dielectric constant of electrons in water \cite{leontyev09}.
-The scaling of charges with this factor has been recently applied to improve classical models for ions and
-other biomolecules \cite{Pluharova2014, martinek17, duboue2018insulin, Mason2019, Duboue2018MgZn} \todo{Pavel can probably update this (insulin?)}.
-In addition to the charge, the Lennard-Jones~$\sigma$ was scaled with a factor $f_\sigma \leq 1$
-to improve the hydration properties of ions with respect to scattering data in these studies
-\todo{Pavel can probably deliver the state of the art information about these parameters and sharpen the discussion}.
+The scaling of charges with this factor has been recently applied to improve empirical models for ions and
+ biomolecules containing charged groups such as proteins\cite{Pluharova2014, martinek17, duboue2018insulin, Mason2019, Duboue2018MgZn}.
+In addition to the charge, the Lennard-Jones~$\sigma$ was scaled with a factor $ 0.9 < f_\sigma \leq 1$
+to improve the description of hydration properties of ions with respect to scattering data in these studies.
+The reason for the small readjustment of the ionic radii is that the original force fields were often developed such that the first peak on the experimental ion-water oxygen radial distribution function was reproduced.
+Upon charge scaling this peak typically moved to slightly larger distances and the agreement with experiment was then restored by a small decrease of the Lennard-Jones~$\sigma$ parameter.
 
@@
 \centering
-\caption{  List of MD simulations reporting their respective
+\caption{  MOVE THE WHOLE TABLE TO SI!!!
+           List of MD simulations reporting their respective
            lipid composition,
@@
 \begin{table}[tbp]
-  \caption{Simulation parameters}
+  \caption{MOVE TABLE TO SI AND MENTION ONLY THE KEY INFO IN 1 SENTENCE IN THE MAIN TEXT
+Simulation parameters}
   \label{tbl:mdpar}
@@
 imperfect structures of the lipid headgroup (Fig.~\ref{simVSexpNOions_POPS}).
+
+MOVE THIS WHOLE PARAGRAPH TO SI
 The POPS headgroup order parameters are also related to the changes of average orientation of P--N and
@@
 and demonstrate the usefulness of ECC also for charged lipids.
-\todo{Maybe Pavel could extend this by mentioning something that this is yet another case where ECC works and/or
-  mentioning some concrete biological examples where anionic membrane/calcium interactions are important?}
+The general philosophy behind our approach is to build upon existing non-polarizable force fields making the minimum of necessary changes by scaling only groups bearing a sizable (partial) charge, followed by slight readjustments of van der Waals radii and possibly also other parameters such as those concerning dihedral angle terms.
+The present work thus represents another piece in the mosaic that leads to development of an accurate empirical force field for biomolecular simulations involving aqueous salt solutions, proteins, peptides, lipid membranes, and/or nucleic acids, which include electronic polarization effects in a mean field way via charge scaling without extra computational costs compared to standard non-polarizable simulations.
 
@@
 % Put your acknowledgments here. 
-P.J. acknowledges support from the Czech Science Foundation (grant no. 16-01074S)  
-and from the Academy of Finland via the FiDiPro award. 
+P.J. acknowledges support from the Czech Science Foundation via an EXPRO grant no. 19-26854X. 
 Computational resources were supplied by the Ministry of Education, Youth and Sports 

### 2019-06-21 Manuscript/manuscript.tex
Updates made by Pavel

@@
 the headgroup and glycerol backbone region of a POPS lipid bilayer are already available in the literature \cite{kucerka14,NMRlipidsIV}.
-Here, we measure also the C--H bond order parameters of the acyl chain region using natural abundance $^{13}$C NMR.
-We first determined the order parameters of each peak corresponding to different C--H bonds detected
-in the first dimension (Fig.~\ref{R-PDLF}) from the dipolar slices in the indirect dimension (Fig.~\ref{R-PDLFslices}).
-Because the exact assignment of these peaks was not possible, we assigned the
-acyl chain order parameters to mimic the profile predicted by the MD simulations (Fig.~\ref{simVSexpNOions_POPS}).
-\todo{Tiago should check and update this and figures in SI.}
+Here, we measure also the C--H bond order parameters of the acyl chain region from a sample of POPS multi-lamellar vesicles by means of a $^1$H--$^{13}$C dipolar recoupling 2D NMR experiment.
+More specifically, as described in detail in the Methods section, we performed R-PDLF 2D spectroscopy~\cite{dvinskikh04} to obtain spectra correlating chemical shifts (direct dimension) and C--H bond dipolar couplings (indirect dimension) from which the C--H bond order parameters defined in eq.~\ref{OPequation} can be determined.
+The assignment of the peaks in the $^{13}$C chemical shift dimension to acyl chain carbon sites was done based on previous work {\cite{ferreira13}} and is shown in Figs.~\ref{INEPT}~and~\ref{R-PDLF}.
+The dipolar splittings for the assigned carbons that were used to calculate their C--H bond order parameter magnitudes are shown in Fig.~\ref{R-PDLFslices}.
+Because the complete assignment of the peaks in the crowded spectral region between 29-31 ppm was not possible due to the chemical shift overlap of different carbons (~\ref{R-PDLF}), we partially assigned the acyl chain order parameters of this region to mimic the profile predicted by the MD simulations (Fig.~\ref{simVSexpNOions_POPS}).
 

### 2019-07-10 Manuscript/manuscript.tex
Turned Figure 1 "form-f, NMR OPs and POPS chem struct" into a panel with an inset commented out table with APLs and updated captions and ref{} commands accordingly

@@
   \centering 
-  \includegraphics[width=\figwidth]{../img/ecc_pops/Order-parameters_form-factors_exp-L17-ECC-lipids.pdf}
-  \includegraphics[width=\figwidth]{../img/POPSstructure.pdf} 
+  %\includegraphics[width=\figwidth]{../img/ecc_pops/Order-parameters_form-factors_exp-L17-ECC-lipids.pdf}
+  %\includegraphics[width=\figwidth]{../img/POPSstructure.pdf} 
+  \includegraphics[width=\figwidth]{../img/formF_OPs_APLinset_POPSchemfig.pdf} 
 \hfill
   \caption{\label{simVSexpNOions_POPS} 
-    \textbf{Top:} X-ray scattering form factors from simulations with Lipid17/Dang \citep{lipid17-future, dang2006} and 
+    \textbf{(A)} X-ray scattering form factors and area per phospholipid of a POPS bilayer
+    from simulations with Lipid17/Dang \citep{lipid17-future, dang2006} and 
     ECC-POPS/ECC-ions \cite{martinek17, Pluhackova2016} compared with experiments~\citep{kucerka14} at 298~K. 
-    \textbf{Middle:} Order parameters of POPS head group, glycerol backbone and acyl chains  
+    \textbf{(B)} Order parameters of POPS head group, glycerol backbone and acyl chains  
     from the same simulations 
@@
     Open/closed symbols are used for palmitoyl/oleoyl chains of POPS. 
-    \textbf{Bottom:} The chemical structure of POPS and the labeling of the carbon segments. 
+    \textbf{(C)} Chemical structure of POPS and labeling of carbon segments. 
   }  
@@
 larger acyl chain order parameters,
-and smaller area per molecule than the experimental values (Fig.~\ref{simVSexpNOions_POPS} and table~\ref{tab:apls}),
+and smaller area per molecule than the experimental values (Fig.~\ref{simVSexpNOions_POPS}),
 indicating that this simulation predicts a too compact POPS lipid bilayer. 
@@
 conformations are well described by the force field. The area per lipid from ECC-POPS simulation is slightly
-smaller than the value reported from SDP model \cite{kucerka14} (table~\ref{tab:apls}), but the values
+smaller than the value reported from SDP model \cite{kucerka14} (Fig.~\ref{simVSexpNOions_POPS}), but the values
 agree within their error estimates. The larger area in ECC-POPS simulation can be
@@
   }
-  \todo{This figure might be more clear without the lipids in background.}
 \end{figure}
 
-\begin{table}[tb!] 
-\centering
-  \caption{Area per lipid of POPS bilayers, $A_L$, from simulations with different force fields at 298~K with \ce{Na^+} counterions. \label{tab:apls} } 
-  \begin{tabular}{l|c } 
-    \multicolumn{2}{c}{POPS} \\
-    model          & $A_L$ / Å$^2$    \\ 
-    \hline 
-    Lipid17/Dang              & 53.5$\pm$ 1.7   \\ 
-%    Lipid17/ff99              & 57.9$\pm$ 1.7   \\ 
-%    \hline 
-    ECC-POPS/ECC-ions         & 59.8$\pm$ 1.8   \\ 
-%    \hline 
-    experiment (SDP model) \citep{kucerka14} & 62.7$\pm$ 1.3  \\ 
-%    \hline 
-  \end{tabular} \\
-\todo{J: This table will become a small inset plot in the top of Fig.~\ref{simVSexpNOions_POPS} }
-\end{table}
+%% moved into the figure 1 as an inset
+%\begin{table}[tb!] 
+%\centering
+  %\caption{Area per lipid of POPS bilayers, $A_L$, from simulations with different force fields at 298~K with \ce{Na^+} counterions. \label{tab:apls} } 
+  %\begin{tabular}{l|c } 
+    %\multicolumn{2}{c}{POPS} \\
+    %model          & $A_L$ / Å$^2$    \\ 
+    %\hline 
+    %Lipid17/Dang              & 53.5$\pm$ 1.7   \\ 
+%%    Lipid17/ff99              & 57.9$\pm$ 1.7   \\ 
+%%    \hline 
+    %ECC-POPS/ECC-ions         & 59.8$\pm$ 1.8   \\ 
+%%    \hline 
+    %experiment (SDP model) \citep{kucerka14} & 62.7$\pm$ 1.3  \\ 
+%%    \hline 
+  %\end{tabular} \\
+%\todo{J: This table will become a small inset plot in the top of Fig.~\ref{simVSexpNOions_POPS} }
+%%\end{table} 
+%
 
@@
   \includegraphics[height=\figheightsmall]{../img/ecc_pops/density_profiles_na-k-counterions_wat_phos_compar_5PC-1PS_ecclipids-lipid17.pdf}
-  \includegraphics[height=\figheightsmall]{../Fig/order_parameters_changes_A-B_PC-PS_mix_POPC_nacl.pdf} 
-  \includegraphics[height=\figheightsmall]{../Fig/order_parameters_changes_A-B_PC-PS_mix_POPC_kcl.pdf} 
+  \includegraphics[height=\figheightsmall]{../img/ecc_pops/order_parameters_changes_A-B_PC-PS_mix_POPC_nacl.pdf} 
+  \includegraphics[height=\figheightsmall]{../img/ecc_pops/order_parameters_changes_A-B_PC-PS_mix_POPC_kcl.pdf} 
   \includegraphics[height=\figheightsmall]{../img/POPCstructure.pdf} 

### 2019-07-10 Manuscript/manuscript.tex
Updated Fig 3 "changes of OPs w.r.t. nacl and kcl" made a common legend, grouped together plots for PC and PS and placed ABCD to make it a nice panel

@@
   \centering 
-  \includegraphics[height=\figheightsmall]{../img/ecc_pops/order_parameters_changes_ecc-lip_L14_A-B-PN-COO_POPC_kcl.pdf} 
-  \includegraphics[height=\figheightsmall]{../img/ecc_pops/order_parameters_changes_ecc-lip_L14_A-B-PN-COO_POPC_nacl.pdf} 
-  \includegraphics[height=\figheightsmall]{../img/ecc_pops/order_parameters_changes_ecc-lip_L14_A-B-PN-COO_POPS_kcl.pdf} 
-  \includegraphics[height=\figheightsmall]{../img/ecc_pops/order_parameters_changes_ecc-lip_L14_A-B-PN-COO_POPS_nacl.pdf} 
+  %\includegraphics[height=\figheightsmall]{../img/ecc_pops/order_parameters_changes_ecc-lip_L14_A-B-PN-COO_POPC_kcl.pdf} 
+  %\includegraphics[height=\figheightsmall]{../img/ecc_pops/order_parameters_changes_ecc-lip_L14_A-B-PN-COO_POPC_nacl.pdf} 
+  %\includegraphics[height=\figheightsmall]{../img/ecc_pops/order_parameters_changes_ecc-lip_L14_A-B-PN-COO_POPS_kcl.pdf} 
+  %\includegraphics[height=\figheightsmall]{../img/ecc_pops/order_parameters_changes_ecc-lip_L14_A-B-PN-COO_POPS_nacl.pdf} 
+  \includegraphics[width=\figwidthfull]{../img/deltaOP_nacl_kcl_PC-PS.pdf} 
   \caption{\label{fig:delta_ordPar_monoval_PCPS}
     Changes of the head group order parameters, and the angles of P--N and ${\rm{C}}_\beta-{\rm{C}}_\gamma$ (stars) vectors
-    with respect to the membrane normal of POPC (left) and POPS (right) in a POPC:POPS (5:1) bilayer 
+    with respect to the membrane normal of POPC (A, B) and POPS (C, D) in a POPC:POPS (5:1) bilayer 
     as a function of \ce{KCl} and \ce{NaCl} concentration from ECC-lipids and Lipid17/Dang simulations 
@@
     Because experimental data with \ce{NaCl} is not available for POPC, the data for KCl and \ce{LiCl} (dashed line, left)
-    are shown as lower and upper bounds, respectively, for the response to \ce{NaCl}. Error bars are not visible for ECC-lipids simulations because they are smaller than the point size.}
-\todo{J: Will improve this figure with panels A, B, C and D 
-and put POPC together, POPS together 
-and remove the duplicate y-axis titles 
-and make a common legend distinguishing different experiments better. } \\
-\todo{The scale in left and right columns (response to KCl and NaCl) should be the same (not necessarily the same scale for POPC and POPS though).}
+    are shown as lower and upper bounds, respectively, for the response to \ce{NaCl}. 
+    Error bars are not visible for ECC-lipids simulations because they are smaller than the point size.}
 \end{figure}

### 2019-07-10 Manuscript/manuscript.tex
Updated Fig 3 "changes of OPs e.r.t kcl, nacl ..." to include the chem struct of POPC head group with labeling as an inset in the panel A
Updated Fig.4 "PC response to PS fraction and density plot" turned it into an ABC panel made a common legend and tweaked the caption accordingly

@@
     are shown as lower and upper bounds, respectively, for the response to \ce{NaCl}. 
-    Error bars are not visible for ECC-lipids simulations because they are smaller than the point size.}
+    Error bars are not visible for ECC-lipids simulations because they are smaller than the point size.
+    Inset shows the labeling of carbon segments in a POPC head group.}
 \end{figure}
@@
   \centering 
-  \includegraphics[height=\figheightsmall]{../img/ecc_pops/density_profiles_na-k-counterions_wat_phos_compar_5PC-1PS_ecclipids-lipid17.pdf}
-  \includegraphics[height=\figheightsmall]{../img/ecc_pops/order_parameters_changes_A-B_PC-PS_mix_POPC_nacl.pdf} 
-  \includegraphics[height=\figheightsmall]{../img/ecc_pops/order_parameters_changes_A-B_PC-PS_mix_POPC_kcl.pdf} 
-  \includegraphics[height=\figheightsmall]{../img/POPCstructure.pdf} 
+  %\includegraphics[height=\figheightsmall]{../img/ecc_pops/density_profiles_na-k-counterions_wat_phos_compar_5PC-1PS_ecclipids-lipid17.pdf}
+  %\includegraphics[height=\figheightsmall]{../img/ecc_pops/order_parameters_changes_A-B_PC-PS_mix_POPC_nacl.pdf} 
+  %\includegraphics[height=\figheightsmall]{../img/ecc_pops/order_parameters_changes_A-B_PC-PS_mix_POPC_kcl.pdf} 
+  %\includegraphics[height=\figheightsmall]{../img/POPCstructure.pdf} 
+  \includegraphics[width=\figwidthfull]{../img/deltaOP_mix_PC-PS.pdf} 
   \caption{\label{fig:delta_ordPar_NaCl_PC-PS_mix_PC} 
-    \textbf{Left:} Number density profiles of \ce{K^{+}} and \ce{Na^{+}} counterions along the membrane normal axis
+    \textbf{(A)} Number density profiles of \ce{K^{+}} and \ce{Na^{+}} counterions along the membrane normal axis
     in ECC-lipids (solid lines) and Lipid17/Dang (dashed lines) simulations of POPC:POPS (5:1) bilayers.  
     The density profiles of phosphate groups and water are divided by 4 and 100, respectively.  
-    \textbf{Middle:} The POPC head group order parameters and the P--N vector angle
+    \textbf{(B, C)} The POPC head group order parameters and the P--N vector angle
     with respect to the membrane normal as a function of POPS content in a bilayer
-    from ECC-lipids and Lipid17/Dang simulations with \ce{Na^+} (left) and \ce{K^+} (right) counterions.
+    from ECC-lipids and Lipid17/Dang simulations with \ce{Na^+} (C) and \ce{K^+} (B) counterions.
     Experimental order parameter values are from Ref. \citenum{scherer87}
-    and the signs from Ref. \citenum{ferreira16} 
-    (only \ce{Na^+} counterions, but shown also in the right plots for \ce{K+}).
+    and the signs from Ref.~\citenum{ferreira16} 
+    (only \ce{Na^+} counterions, but shown also in the left plots (B) for \ce{K+}).
     Error bars are not visible for most of the simulation points because they are smaller than the point size.
-    \textbf{Right:} Chemical structure and labelling of carbon segments of POPC. 
+    Chemical structure and labelling of carbon segments of POPC is shown in Fig.~\ref{fig:delta_ordPar_monoval_PCPS}. 
   }
-\todo{J: Will manually improve this figure with panels A, B, C and D 
-and will erase the duplicate y-axis description 
-and will add ticks to the right side of the plotting area. } \\
-\todo{The scales should be the same in the left and right panel.}
 \end{figure*}

### 2019-07-10 Manuscript/manuscript.tex
Updated Fig. 6 " populations and stoichiometry of Ca2+ binding" added AB and made it into a panel, put the plots horizontally

@@
   \begin{tabular}{ c }
-  \includegraphics[height=1.5\figheightsmall]{../img/ecc_pops/density_profiles_ca_na_k_cl_wat_phos_ecclipids_lipid17_compar_80and200mMCaCl2.pdf}
-  \includegraphics[height=1.5\figheightsmall]{../img/ecc_pops/order_parameters_changes_ecc-lip_L14_A-B-PN-COO_POPC_cacl.pdf} 
-  \includegraphics[height=1.5\figheightsmall]{../img/ecc_pops/order_parameters_changes_ecc-lip_L14_A-B-PN-COO_POPS_cacl.pdf} 
+  %\includegraphics[height=1.5\figheightsmall]{../img/ecc_pops/density_profiles_ca_na_k_cl_wat_phos_ecclipids_lipid17_compar_80and200mMCaCl2.pdf}
+  %\includegraphics[height=1.5\figheightsmall]{../img/ecc_pops/order_parameters_changes_ecc-lip_L14_A-B-PN-COO_POPC_cacl.pdf} 
+  %\includegraphics[height=1.5\figheightsmall]{../img/ecc_pops/order_parameters_changes_ecc-lip_L14_A-B-PN-COO_POPS_cacl.pdf} 
+  \includegraphics[width=\figwidthfull]{../img/deltaOP_cacl_PC-PS.pdf} 
   \end{tabular}
@@
 %  \end{tabular}
-  \caption{\label{fig:cacl-dens_PCPS} 
-    \textbf{Left panel:}
+  \caption{
+    \label{fig:cacl-dens_PCPS} 
+    \label{fig:delta_ordPar_CaCl_PCPS} 
+    \textbf{(A)}
     Number density profiles of \ce{Ca^{2+}} and \ce{Cl^-} ions and \ce{Na^{+}} counterions 
@@
     the density profiles of phosphate groups and water are divided by 4 and 100, respectively.  
-  }
-  \caption{\label{fig:delta_ordPar_CaCl_PCPS} 
-    \textbf{Middle panels:}
+    \textbf{(B, C)}
     Changes of the head group order parameters, and the angles of P--N (circles) and C$_\beta$--C$_\gamma$ (stars) vectors 
-    with respect to the membrane normal of POPC (left) and POPS (right) in a POPC:POPS (5:1) bilayer 
+    with respect to the membrane normal of POPC (B) and POPS (C) in a POPC:POPS (5:1) bilayer 
     as a function of \ce{CaCl$_2$} concentration from ECC-lipids and Lipid17/Dang simulations 
@@
 %  }
-  \todo{J: Will adjust the Fig into a nice panel and change the captions accordingly. } \\
 \end{figure}
@@
   \begin{tabular}{ c }
-  \includegraphics[width=8cm]{../img/bound_ca_populations.pdf} \\
-  \includegraphics[width=8cm]{../img/stoichiometry_CaCl2_comparison_Ecc-lipids_PC-vs-PCPS.pdf} 
+  %\includegraphics[width=8cm]{../img/bound_ca_populations.pdf} \\
+  %\includegraphics[width=8cm]{../img/stoichiometry_CaCl2_comparison_Ecc-lipids_PC-vs-PCPS.pdf} 
+  \includegraphics[width=\figwidthfull]{../img/populations_stoichiometry.pdf} 
   \end{tabular}
   \caption{ \label{fig:cacl_complexes} 
-    \textbf{Top:} Percentages of the bound Ca$^{2+}$ 
+    \textbf{(A)} Percentages of the bound Ca$^{2+}$ 
     in exclusive contact with the given oxygen moiety
@@
         For numerical values, see table~\ref{tab:binding}.
-    \textbf{Bottom:} Relative probabilities of \ce{Ca^{2+}} ions to coordinate with a certain number of lipids
+    \textbf{(B)} Relative probabilities of \ce{Ca^{2+}} ions to coordinate with a certain number of lipids
     in pure POPC bilayer with 350 mM CaCl$_2$ and in POPC:POPS (5:1) mixture with 400 mM CaCl$_2$.  
@@
   }
-  \todo{Bottom: The x-axis label should be changed to ''number of lipids directly bound to Ca$^{2+}$''.
-    The current label has another meaning in the text now.} \\
-  \todo{Bottom: The labels could be more clear. Maybe ''PC in POPC:POPS(5:1)'' ...}
 \end{figure}

### 2019-07-11 Manuscript/manuscript.tex
Fixed Zenodo deposit reference and added reference to TO BE DEPOSITED ECC-POPS paprameters The reference has to be updated again (change of DOI) after that.
Updated reference with nacl/kcl simulations to include also simulation parameters and topologies, updated text (removed associated todo)

@@
 Further optimization of parameters was not done in this work. 
-The generated ECC-POPS parameters are available from Ref.~\citenum{??}
-\todo{We need here a permanent citation with DOI.
-  Easiest would be to add the parameters into a one of the existing Zenodo repositories containing other simulation parameters.
-  We could also put a release of Git into Zenodo, cite that and specify the exact directory where parameters are found,
-  but this is more complicated for users to find.
-}
+The generated ECC-POPS parameters are available from Ref.~\citenum{ecclipids_pcps_nacl_kcl_series}.
 

### 2019-07-11 Manuscript/manuscript.tex

@@
 \title[] 
-{Improved Description of Cation Binding to Lipid Bilayer with Negatively Charged POPS by Effective Inclusion of Electronic Polarization}
+{Accurate Description of 
+Cation Binding to Lipid Bilayers 
+with Negatively Charged Phosphatidylserine 
+Requires Accounting for Electronic Polarization}
   %Detailed structure of negatively charged membranes 
@@
 regulate surface charge and protein localization \cite{yeung08}, 
-induce protein aggregation \cite{zhao04,gorbenko06} and membrane fusion \cite{??}.
+induce protein aggregation \cite{zhao04,gorbenko06} and membrane fusion \cite{wilschut1981calcium, papahadjopoulos90, verma2018cell}.
 As these functions are also often regulated by ions \cite{leventis10},
@@
 As a recent example, we have shown by a combination of spectroscopic experiments, biochemical essays, and molecular simulations, that binding of
-calcium to PIP2 lipids inhibits their recognition by a phospolipase C, which demonstrates that not only proteins but also lipids can get involved in calcium 
-signaling \cite{DOI: 10.1021/jacs.6b11760}.
+calcium to \ce{PIP_2} lipids inhibits their recognition by a phospolipase C, which demonstrates that not only proteins but also lipids can get involved in calcium 
+signaling \cite{Bilkova2017Calcium}.
 

### 2019-07-12 Manuscript/manuscript.tex
Improvements in abstract +1 todo

@@
   Phosphatidylserine (PS) lipids are important signaling molecules and the most common negatively charged lipids in eukaryotic membranes.
-  The signaling can be often regulated with calcium, but its interactions with PS headgroups are not fully understood.
+  The signaling can be often regulated by calcium, but its interactions with PS headgroups are not fully understood.
+  \todo{The way we write "head groups" resp. "headgroups" has to be united. 
+        I prefer "head groups" as it is gramatically correct, but I am fine with either of the options.}
   Classical molecular dynamics (MD) simulations can potentially give detailed description of lipid-ion interactions,
@@
   The resulting ECC-POPS models, gives more realistic interactions with Na$^+$ and Ca$^{2+}$ cations than the original Amber Lipid17 parameters,
-  when validated using the headgroup order parameters and ''electrometer concept''.
-  In ECC-lipids simulations, the Ca$^{2+}$ cations do not simultaneously interact with more than two PS lipids, and interactions 
-  carboxylate groups is twice more likely than with phosphate group, while interaction with carbonyls is almost negligible.
-  Our results pave the way for more realistic MD simulations of anionic biological membranes and demonstrate the ECC
-  approach is useful also for charged lipids.
+  when validated using headgroup order parameters and ''electrometer concept''.
+  In ECC-lipids simulations, Ca$^{2+}$ cations do not simultaneously interact with more than two PS lipids, and interactions 
+  with carboxylate groups is twice more likely than with phosphate groups, while interaction with carbonyls is almost negligible.
+  Our results pave the way for more realistic MD simulations of anionic biological membranes and demonstrate the 
+  usefulness of ECC also to charged lipids.
 \end{abstract}

### 2019-07-12 Manuscript/manuscript.tex
Improvements and comments in Introduction +2 todos

@@
 induce protein aggregation \cite{zhao04,gorbenko06} and membrane fusion \cite{wilschut1981calcium, papahadjopoulos90, verma2018cell}.
-As these functions are also often regulated by ions \cite{leventis10},
+Recently, we have shown by a combination of spectroscopic experiments, biochemical essays, and molecular simulations, that binding of
+calcium to \ce{PIP_2} lipids inhibits their recognition by a phospolipase~C, 
+which demonstrates that not only proteins but also lipids can get involved in calcium signaling \cite{Bilkova2017Calcium}.
+As such \todo{lipid-related functions resp. signalling (choose more suitable)} are also often regulated by ions \cite{leventis10},
 a detailed understanding of interactions between negatively charged lipids and biologically relevant cations such as calcium is of a great importance.
-As a recent example, we have shown by a combination of spectroscopic experiments, biochemical essays, and molecular simulations, that binding of
-calcium to \ce{PIP_2} lipids inhibits their recognition by a phospolipase C, which demonstrates that not only proteins but also lipids can get involved in calcium 
-signaling \cite{Bilkova2017Calcium}.
-
-Spectroscopic experiments give detailed information about the
+
+Spectroscopic experiments give \todo{detailed .vs. accurate (better?). 
+Reason: I did not like that the sentence says that experiments give detailed information and at the same time such information is indirect and difficult to interpret.} 
+information about the
 interactions between ions and PS lipids, but the data is often indirect and difficult to
@@
 is the direct comparison between experimental and calculated order parameters,
-which reduced the ambiguity arising from the interpretation of the data.
+which reduce the ambiguity arising from the interpretation of the data.
 Unfortunately, none of the readily available force fields was sufficiently accurate to correctly reproduce the
@@
 correction (ECC) \cite{leontyev09}. The good agreement between the resulting ECC-POPC
-model and experiments enable detailed interpretation of calcium binding details to POPC lipid bilayer.
+model and experiments enable detailed interpretation of calcium binding details to a POPC lipid bilayer.
 

### 2019-07-12 Manuscript/manuscript.tex
Improved MD Methods section +1 todo

@@
 In addition to the charge, the Lennard-Jones~$\sigma$ was scaled with a factor $ 0.9 < f_\sigma \leq 1$
+\todo{The given range is not correct, e.g. scaling of POPC (0.89), Mg2+ (0.72 and 0.82). 
+Extending the lower bound to 0.8 fits the cases I know of with the exception of the small-Mg2+. }
 to improve the description of hydration properties of ions with respect to scattering data in these studies.
@@
 the partial charges depend on the methods used to derive the respective force fields.
-For example, the partial charges may already include the effects of electronic polarizability to some extent,
+For example, the partial charges of overall neutral molecules may already include the effects of electronic polarizability to some extent,
 and hence, the scaling factor can be larger than the theoretically derived value for ions in water, $f_q \approx 0.75$.  
+% above does not apply to molecules and moieties with non-zero total charge (hence the edit). 
 Before the ECC theory was rigorously derived, similar idea was already employed in early classical 
@@
 Here, we apply the ECC to the Amber-based lipid17 parameters \cite{lipid17-future} of POPS lipid with a monovalent negative charge.
-Because POPS carries a physical total charge as aqueous ions, we apply the theoretically derived scaling factor for
-ions $f_q = 0.75$ \cite{leontyev09} to the partial charges of POPS. 
+Because POPS is anionic at physiological pH$\approx 7$ carrying a physical total charge as aqueous ions, 
+we apply the theoretically derived scaling factor\cite{leontyev09} for
+ions $f_q = 0.75$ to the partial charges of POPS 
+only in the headgroup, glycerol backbone, and carbonyl regions. % added because the information where we scale charges was missing before.
 The lipids with a total charge of -0.75 are also neutralized
 by the counterion charges of +0.75 in simulations with ECC-ions \cite{Pluharova2014, kohagen16, martinek17}.
-Following our previous work for POPC, we use the scaling factor of $f_\sigma = 0.89$ for the Lennard-Jones~$\sigma$ parameters and
-use it to scale the parameters of atoms only in the headgroup, glycerol backbone, and carbonyl regions \cite{melcr18}.
+Following our previous work for POPC, we use the scaling factor $f_\sigma = 0.89$ for the Lennard-Jones~$\sigma$ parameters 
+to scale the corresponding parameters of atoms 
+only in the headgroup, glycerol backbone, and carbonyl regions \cite{melcr18}. % Can be shortened by saying it is the same set as that for scaling charges.
 Further optimization of parameters was not done in this work. 

### 2019-07-12 Manuscript/manuscript.tex
Improvements in Simulation details See also commented lines in the tex file

@@
 \centering
-\caption{  MOVE THE WHOLE TABLE TO SI!!!
+\caption{  MOVE THE WHOLE TABLE TO SI!!! % J: I do not have a strong opinion on this. I think it can stay within the main text in a JCTC paper.
            List of MD simulations reporting their respective
@@
 and the first  $50 \,$ns were omitted as an equilibration period. 
-For deposited trajectories and parameter files see 
+Generated trajectories and parameter files are available from 
 Refs.~\citenum{ecclipids_pcps_nacl_kcl_series, ecclipids_pcps_cacl2_series, ecclipids_pcps_mixtures_counterions, 
-               lipid17_nacl_kcl_series, lipid17_ff99_ions, lipid17_cacl_series}
+               lipid17_nacl_kcl_series, lipid17_ff99_ions, lipid17_cacl_series}.
 
@@
 the default Amber ion parameters \cite{aqvist90} led to the artificial clustering of ions in solution \cite{NMRlipidsIV}.
+
 In the ECC-lipids simulations,
@@
 \begin{table}[tbp]
-  \caption{MOVE TABLE TO SI AND MENTION ONLY THE KEY INFO IN 1 SENTENCE IN THE MAIN TEXT
+  \caption{MOVE TABLE TO SI AND MENTION ONLY THE KEY INFO IN 1 SENTENCE IN THE MAIN TEXT  % J: I am fine with this, the simulation parameters are rather standard.
+There is nothing to be done in the text, the reference to this table will only obtain prefix "S".
 Simulation parameters}

### 2019-07-12 Manuscript/manuscript.tex
Improvements in section Results-monovalent ions

@@
 smaller than the value reported from SDP model \cite{kucerka14} (Fig.~\ref{simVSexpNOions_POPS}), but the values
-agree within their error estimates. The larger area in ECC-POPS simulation can be
+agree within their error estimates. The larger area in ECC-POPS compared to Lipid17/Dang simulation can be
 explained by increased headgroup repulsion due to lower counterion binding affinity (Fig. \ref{fig:POPS-counterions-dens}).
@@
 %SAMULI: I have moved this figure back to the main text
-\begin{figure}[hbp!] 
+\begin{figure}[tbp!] 
   \centering 
-  \includegraphics[width=8cm]{../img/ecc_pops/density_profiles_na-k-counterions_wat_phos_compar_purePOPS_ecclipids-lipid17.pdf}
+  \includegraphics[width=\figwidthsmall]{../img/ecc_pops/density_profiles_na-k-counterions_wat_phos_compar_purePOPS_ecclipids-lipid17.pdf}
   \caption{\label{fig:POPS-counterions-dens}
@@
 against experiments using the order parameters of $\alpha$ and $\beta$ carbons in the
-phosphatidylcholine (PC) lipid headgroup (see Fig.~\ref{fig:delta_ordPar_NaCl_PC-PS_mix_PC} for the labeling) \cite{catte16,melcr18,NMRlipidsIV},
+phosphatidylcholine (PC) lipid headgroup (see Fig.~\ref{fig:delta_ordPar_monoval_PCPS} for the labeling) \cite{catte16,melcr18,NMRlipidsIV},
 which decrease proportionally to the bound charge according to the ''electrometer concept'' \citep{seelig87}.
@@
 we evaluate the counterion binding affinity to POPC:POPS mixtures also by monitoring
-the response of POPC headgroup order parameters to the increasing mole fraction of PS lipids~\cite{NMRlipidsIV}.
+the response of POPC headgroup order parameters to the increasing molar fraction of PS lipids~\cite{NMRlipidsIV} 
+(Fig.~\ref{fig:delta_ordPar_NaCl_PC-PS_mix_PC}).
 
@@
     compared with experimental values from Ref. \citenum{roux90} (signs from Ref. \citenum{ferreira16}) at 298 K.
-    The y-axis for the $\alpha$-carbon results of POPS (middle right) is transferred
+    The y-axis for the $\alpha$-carbon results of POPS (C and D, middle) is shifted
     with the same value for both order parameters such that the lower order
     parameter value from pure POPS is at zero to correctly illustrate the significant forking.
-    Because experimental data with \ce{NaCl} is not available for POPC, the data for KCl and \ce{LiCl} (dashed line, left)
+    Because experimental data with \ce{NaCl} is not available for POPC, the data for KCl and \ce{LiCl} (B, dashed lines)
     are shown as lower and upper bounds, respectively, for the response to \ce{NaCl}. 
     Error bars are not visible for ECC-lipids simulations because they are smaller than the point size.
-    Inset shows the labeling of carbon segments in a POPC head group.}
+    Inset in A shows the labeling of carbon segments in a POPC head group.}
 \end{figure}
@@
 lipid mixtures and ionic conditions~\cite{NMRlipidsIV,roux90}.
-In experiments, the headgroup order parameters of POPS are almost unchanged upon increasing the POPC content or monovalent ion concetration
-(\ref{fig:delta_ordPar_NaCl_PC-PS_mix_PS} and Figs.~\ref{fig:delta_ordPar_monoval_PCPS}).
-In Lipid17/Dang simulations, the changes of POPS headgroup order parameters with increasing amount of
-both POPC or monovalent salts (\ce{KCl} and \ce{NaCl}) are overestimated (\ref{fig:delta_ordPar_NaCl_PC-PS_mix_PS} and Figs.~\ref{fig:delta_ordPar_monoval_PCPS}).
+In experiments, the headgroup order parameters of POPS are almost unchanged upon increasing the POPC content or monovalent ion concentration, 
+while in Lipid17/Dang simulations the changes of POPS headgroup order parameters with increasing amount of
+both POPC or monovalent salts (\ce{KCl} and \ce{NaCl}) are overestimated 
+(Figs.~\ref{fig:delta_ordPar_monoval_PCPS} and \ref{fig:delta_ordPar_NaCl_PC-PS_mix_PS}).
 Similar results also for other force fields are reported elsewhere~\cite{NMRlipidsIV}.
 In ECC-lipids simulations, the response of POPS headgroup order parameters to both POPC and monovalent salt concentration (\ce{KCl} and \ce{NaCl})
-are more modest and closer to experiments (\ref{fig:delta_ordPar_NaCl_PC-PS_mix_PS} and Figs.~\ref{fig:delta_ordPar_monoval_PCPS}).
+are more modest and closer to experiments. 
 The changes in order parameters can be related to the orientations of the P--N and C$_{\beta}$--C$_{\gamma}$
 vectors in POPC and POPS, which change much less and more systematically in ECC-lipids
-than in Lipid17/Dang simulations (\ref{fig:delta_ordPar_NaCl_PC-PS_mix_PS} and Figs.~\ref{fig:delta_ordPar_monoval_PCPS}).
+than in Lipid17/Dang simulations 
+(Figs.~\ref{fig:delta_ordPar_monoval_PCPS} and \ref{fig:delta_ordPar_NaCl_PC-PS_mix_PS}).
 
@@
 $\Gamma ^{w} _{\rm K}=-0.123 \pm 0.005 \mathrm{nm}^{-2}$ in POPC:POPS (5:1) mixture (Fig.~\ref{fig:nacl-dens_PCPS}).
-As discussed above, the sodium binding to POPC:POPS (5:1) mixture may be slightly overestimated also in
-ECC-lipids simulations. The response of headgroup order parameters to the addition of NaCl are
-slightly overestimated also in ECC-POPC simulations (Fig 3 in Ref. \citenum{melcr18}),
-suggesting that similar small overestimation of sodium binding may also occur to a pure POPC bilayer.
-Therefore, the overestimation may arise from the sodium interactions with PC headgroup and
-increase in binding affinity due to PS lipids would not be affected.
+%
+% J: I deemed the following sentences need improvement, my suggestion is below.
+%
+%As discussed above, the sodium binding to POPC:POPS (5:1) mixture may be slightly overestimated also in
+%ECC-lipids simulations. The response of headgroup order parameters to the addition of NaCl are
+%slightly overestimated also in ECC-POPC simulations (Fig 3 in Ref.~\citenum{melcr18}),
+%suggesting that similar small overestimation of sodium binding may also occur to a pure POPC bilayer.
+%Therefore, the overestimation may arise from the sodium interactions with PC headgroup and
+%increase in binding affinity due to PS lipids would not be affected.
+Slightly overestimated binding of \ce{Na+} to a POPC bilayer can be deduced from
+the response of headgroup order parameters to the addition of NaCl (Fig~3 in Ref.~\citenum{melcr18}) 
+and similarly also that to a POPC:POPS (5:1) mixture from the discussion above. 
+In spite of this small inaccuracy
+the increase in binding affinity of \ce{Na+} due to PS lipids remains unaffected.
 
@@
     compared with experimental values from Ref. \citenum{roux90} (signs from Refs. \citenum{ferreira16} and \citenum{NMRlipidsIV}) at 298 K.
-    The y-axis for the $\alpha$-carbon results of POPS (middle right) is transferred
+    The y-axis for the $\alpha$-carbon results of POPS (middle right) is shifted
     with the same value for both order parameters such that the lower order
@@
 In ECC-lipids simulations, the changes of the PS headgroup order parameters are not overestimated,
-but the strong dependence on low concetrations of CaCl$_2$ is not fully reproduced (Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}).
+but the strong dependence on low concentrations of CaCl$_2$ is not fully reproduced (Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}).
 In addition to possibly suboptimal interactions between calcium ions and PS headgroup, the potential sources of this

### 2019-07-12 Manuscript/manuscript.tex
Improvements in the Ca2+ section in results

@@
 which underestimated the headgroup order parameter response.
-In ECC-lipids simulations, the headgroup responses are in better agreement with experiments (Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}),
+In ECC-lipids simulations, the headgroup responses are in better agreement with experiments,
 indicating that the lower binding affinity than in Lipid17/Dang simulations is more realistic (Fig.~\ref{fig:cacl-dens_PCPS}).
@@
 suggesting that the ECC accelerates the equilibration of ions at lipid bilayer interface.
-Therefore, our 1 $\mu$s simulations seem to be sufficiently long for the ECC-lipids simulations, because
+Therefore, our 1 $\mu$s simulations are sufficiently long for the ECC-lipids simulations, because
 90\% of the calcium residence times are shorter than $60\,\mathrm{ns}$ for pure POPC bilayer
@@
 longest observed residence times are $141\,\mathrm{ns}$ and $485\,\mathrm{ns}$, respectively (Fig.~\ref{fig:hist_residence_times}).
-Interestingly, the calcium residence times are 3-4 times longer in POPC:POPS (5:1) mixture than in pure POPC.
+Interestingly, the calcium residence times are 3-4 times longer in POPC:POPS (5:1) mixture than in pure POPC
+in line with increased relaxation times of fluorescence probes from experiments. \cite{melcrova16}
 
@@
     Number density profiles of \ce{Ca^{2+}} and \ce{Cl^-} ions and \ce{Na^{+}} counterions 
-    along the normal of the membrane starting at the centre
-    for the negatively charged membrane with the composition 5\,PC:1\,PS
-    at 80~mM (top) and 200~mM (bottom) added buffer concentrations of \ce{CaCl2} from simulations with ECC-lipids (solid) and Lipid17 (dashed). 
+    along the normal of the negatively charged membrane with the composition 5\,PC:1\,PS
+    at 80~mM (A, top) and 200~mM (A, bottom) added buffer concentrations of \ce{CaCl2} from simulations with ECC-lipids (solid) and Lipid17 (dashed). 
     In order to visualize the density profiles with a comparable scale 
@@
     with respect to the membrane normal of POPC (B) and POPS (C) in a POPC:POPS (5:1) bilayer 
-    as a function of \ce{CaCl$_2$} concentration from ECC-lipids and Lipid17/Dang simulations 
+    as a function of \ce{CaCl$_2$} added buffer concentration from ECC-lipids and Lipid17/Dang simulations 
     compared with experimental values from Ref. \citenum{roux90} (signs from Refs. \citenum{ferreira16} and \citenum{NMRlipidsIV}) at 298 K.
-    The y-axis for the $\alpha$-carbon results of POPS (middle right) is shifted
+    The y-axis for the $\alpha$-carbon results of POPS (C, right) is shifted
     with the same value for both order parameters such that the lower order
@@
     when bound to PC only, PS only or simultaneously to both
-    calculated from ECC-lipids simulation of POPC:POPS (5:1) bilayer with 409 mM CaCl$_2$. 
+    calculated from ECC-lipids simulation of POPC:POPS (5:1) bilayer with 409 mM CaCl$_2$ added buffer concentration. 
 	In the case of simultaneous binding to both PC and PS,
@@
     Analysis was done for both systems by considering all lipids (blue and violet) and
-    for POPC:POPS (5:1) mixture also by considering POPC and POPS lipids separately (green and orange). 
+    for POPC:POPS (5:1) mixture also by considering only POPC and POPS lipids separately (green and orange). 
     %All lipids were taken into account with the exception of the complexes in light green and orange, 
@@
     as the distance between the cation and the oxygen atoms of the lipids. 
-    Previously published simulation data \cite{melcr18} for pure POPC bilayers were taken directly from \cite{ECC-POPC_nacl_cacl2_files}. 
+    Previously published simulation data \cite{melcr18} for pure POPC bilayers were taken directly from Ref.~\citenum{ECC-POPC_nacl_cacl2_files}. 
   }
@@
 In addition to possibly suboptimal interactions between calcium ions and PS headgroup, the potential sources of this
-discrepancy include the above observed sligth overestimation of \ce{Na^+} counterions and
+discrepancy include the above observed slight overestimation of \ce{Na^+} counterions and
 imperfect structures of the lipid headgroup (Fig.~\ref{simVSexpNOions_POPS}).
 
-MOVE THIS WHOLE PARAGRAPH TO SI
+\todo{MOVE THIS WHOLE PARAGRAPH TO SI}
 The POPS headgroup order parameters are also related to the changes of average orientation of P--N and
@@
 On the other hand, the average orientation of the C$_{\beta}$--C$_{\gamma}$ vector in PS headgroup
-tilts to the opposite direction by $11^\circ$ in the same system.
+tilts to the opposite direction by $11^\circ$ in the same system
+due to the attraction between \ce{COO^-} groups in PS and the cations adsorbed at the phosphate region.
 
@@
 to the carboxylate than to the phosphate moiety of POPS headgroup,
-while binding to acyl chain carbonyls is almost negligible (Fig.~\ref{fig:cacl_complexes} and table~\ref{tab:binding}).
+while binding to acyl chain carbonyls is almost negligible (Fig.~\ref{fig:cacl_complexes} and Table~\ref{tab:binding}).
 The result is consistent with CHARMM36 simulations without the NBfix correction \cite{hallock18},
-but CHARMM36 simulations with the NBfix correction \cite{kim16} predict almost exclusive binding on carboxylate group \cite{valentine18}
-and Berger simulations signifincant binding affinity also on acyl chain carbonyls \cite{melcrova16}.
+but CHARMM36 simulations with the NBfix correction \cite{kim16} predict almost exclusive binding to carboxylate group \cite{valentine18}
+and Berger simulations show signifincant binding affinity also to acyl chain carbonyls \cite{melcrova16}.
 On the other hand, calcium ions bind too strongly on bilayers simulated using CHARMM36 without the NBfix correction

### 2019-07-12 Manuscript/manuscript.tex
Improvements in Conclusions

@@
 of PS headgroups than to phosphate groups, while binding to carbonyls is almost negligible.
-Binding of Ca$^{2+}$ ions to more than two POPS lipids simultaneously is very rare in ECC-lipids simulations.
-Because ECC-lipids parameters give the best results in direct comparison with NMR order parameters, 
-we believe that our results help in resoving controversial interpretations of more indirect experiments
+Binding of Ca$^{2+}$ ions to more than two POPS lipids simultaneously is very rare in \mbox{ECC-lipids} simulations.
+Because \mbox{ECC-lipids} parameters give the best results in direct comparison with NMR order parameters, 
+we believe that our results help in resolving controversial interpretations of more indirect experiments
 using different force fields~\cite{melcrova16,valentine18,hallock18}.
@@
 and demonstrate the usefulness of ECC also for charged lipids.
-The general philosophy behind our approach is to build upon existing non-polarizable force fields making the minimum of necessary changes by scaling only groups bearing a sizable (partial) charge, followed by slight readjustments of van der Waals radii and possibly also other parameters such as those concerning dihedral angle terms.
+The general philosophy behind our approach is to build upon existing non-polarizable force fields making the minimum of necessary changes by scaling only groups bearing a sizeable (partial) charge, followed by slight readjustments of van der Waals radii and possibly also other parameters such as those concerning dihedral angle terms.
 The present work thus represents another piece in the mosaic that leads to development of an accurate empirical force field for biomolecular simulations involving aqueous salt solutions, proteins, peptides, lipid membranes, and/or nucleic acids, which include electronic polarization effects in a mean field way via charge scaling without extra computational costs compared to standard non-polarizable simulations.

### 2019-08-06 Manuscript/manuscript.tex
Fixes in bibliography and SI. Please, also read commented lines!

@@
     and the signs from Ref.~\citenum{ferreira16} 
-    (only \ce{Na^+} counterions, but shown also in the left plots (B) for \ce{K+}).
+    (only \ce{Na^+} counterions, but shown also in the left plots (B) for \ce{K+} with dashed lines).
     Error bars are not visible for most of the simulation points because they are smaller than the point size.
@@
 
+% I do not think it is necessary to
 \todo{MOVE THIS WHOLE PARAGRAPH TO SI}

### 2019-08-13 Manuscript/manuscript.tex

--- 2019-07-12 Manuscript/manuscript.tex
@@
 \title[] 
-{Accurate Description of 
-Cation Binding to Lipid Bilayers 
-with Negatively Charged Phosphatidylserine 
-Requires Accounting for Electronic Polarization}
-  %Detailed structure of negatively charged membranes 
-  %  of phosphatidylserine and phosphatidylcholine 
-  %  at concentrations of calcium, sodium and potassium salts
-  %  from molecular dynamics simulations with electronic polarization }
-  % JOSEF: I feel that the above title is a bit better, because it focuses on PS,
-  %        but the following one is an alternative, which focuses more on the cations as in roux90
-  %{ Detailed structure of calcium, sodium and potassium cations
-  %  at negatively charged membranes of phosphatidylserine and phosphatidylcholine 
-  %  from molecular dynamics simulations with electronic polarization }
+      {Improved Cation Binding to Lipid Bilayer with
+Negatively Charged POPS by Effective
+Inclusion of Electronic Polarization}
 
@@
   The signaling can be often regulated by calcium, but its interactions with PS headgroups are not fully understood.
-  \todo{The way we write "head groups" resp. "headgroups" has to be united. 
-        I prefer "head groups" as it is gramatically correct, but I am fine with either of the options.}
   Classical molecular dynamics (MD) simulations can potentially give detailed description of lipid-ion interactions,
   but the results strongly depend on the used force field.
-  Here, we apply the electronic continuum correction (ECC) to the Amber Lipid17 parameters of POPS lipid to improve its
+  Here, we apply the electronic continuum correction (ECC) to the Amber Lipid17 parameters of 1-palmitoyl-2-oleoyl-{\it sn}-glycero-3-phospho-L-serine (POPS) lipid to improve its
   interactions with Na$^+$ and Ca$^{2+}$ ions.
@@
   factor of 0.75, derived for monovalent ions and the Lennard-Jones $\sigma$ parameters of the same segments were scaled with a factor of 0.89.
-  The resulting ECC-POPS models, gives more realistic interactions with Na$^+$ and Ca$^{2+}$ cations than the original Amber Lipid17 parameters,
+  The resulting ECC-POPS model gives more realistic interactions with Na$^+$ and Ca$^{2+}$ cations than the original Amber Lipid17 parameters,
   when validated using headgroup order parameters and ''electrometer concept''.
@@
 induce protein aggregation \cite{zhao04,gorbenko06} and membrane fusion \cite{wilschut1981calcium, papahadjopoulos90, verma2018cell}.
-Recently, we have shown by a combination of spectroscopic experiments, biochemical essays, and molecular simulations, that binding of
+As such lipid-related functions are also often regulated by ions \cite{leventis10},
+a detailed understanding of interactions between negatively charged lipids and biologically relevant cations such as calcium is of a great importance.
+Recent combination of spectroscopic experiments, biochemical essays, and molecular simulations showed that binding of
 calcium to \ce{PIP_2} lipids inhibits their recognition by a phospolipase~C, 
 which demonstrates that not only proteins but also lipids can get involved in calcium signaling \cite{Bilkova2017Calcium}.
-As such \todo{lipid-related functions resp. signalling (choose more suitable)} are also often regulated by ions \cite{leventis10},
-a detailed understanding of interactions between negatively charged lipids and biologically relevant cations such as calcium is of a great importance.
-
-Spectroscopic experiments give \todo{detailed .vs. accurate (better?). 
-Reason: I did not like that the sentence says that experiments give detailed information and at the same time such information is indirect and difficult to interpret.} 
-information about the
+
+Spectroscopic experiments give accurate information about the
 interactions between ions and PS lipids, but the data is often indirect and difficult to
@@
 Some studies suggest that the
-binding constant of ions to negatively charged lipids is similar to that of zwitterionic lipids,
+binding affinity of ions to negatively charged lipids is similar to that of zwitterionic lipids,
 and the binding affinity is increased only due to the increased cation
@@
 
-The NMRlipids project \url{nmrlipids.blogspot.fi} has recently demonstrated
+The NMRlipids project (\url{nmrlipids.blogspot.fi}) has recently demonstrated
 that such controversies can be resolved by using the headgroup order parameters
@@
 be related to cation binding affinity to lipid bilayers using the electrometer
-concept \cite{akutsu81,altenbach84,seelig87}. The main advantage of the approach
+concept \cite{akutsu81,altenbach84,seelig87}. The main advantage of this approach
 is the direct comparison between experimental and calculated order parameters,
@@
 
-Here, we extend the ECC approach also to the negatively charged 1-palmitoyl-2-oleoyl-sn-glycero-3-phospho-L-serine (POPS) lipid.
+Here, we extend the ECC approach also to the negatively charged 1-palmitoyl-2-oleoyl-{\it sn}-glycero-3-phospho-L-serine (POPS) lipid.
 We also complement the available experimental data for the force field quality
@@
 In addition to the charge, the Lennard-Jones~$\sigma$ was scaled with a factor $ 0.9 < f_\sigma \leq 1$
-\todo{The given range is not correct, e.g. scaling of POPC (0.89), Mg2+ (0.72 and 0.82). 
+\todo{JOE: The given range is not correct, e.g. scaling of POPC (0.89), Mg2+ (0.72 and 0.82). 
 Extending the lower bound to 0.8 fits the cases I know of with the exception of the small-Mg2+. }
@@
 
-Here, we apply the ECC to the Amber-based lipid17 parameters \cite{lipid17-future} of POPS lipid with a monovalent negative charge.
-Because POPS is anionic at physiological pH$\approx 7$ carrying a physical total charge as aqueous ions, 
+Here, we apply the ECC to the Amber-based lipid17 parameters \cite{lipid17-future} (available in AmberTools18 \cite{amber18})
+of POPS lipid with a monovalent negative charge.
+Because POPS is anionic at physiological pH $\approx 7$ carrying a physical total charge as aqueous ions, 
 we apply the theoretically derived scaling factor\cite{leontyev09} for
 ions $f_q = 0.75$ to the partial charges of POPS 
-only in the headgroup, glycerol backbone, and carbonyl regions. % added because the information where we scale charges was missing before.
+in the headgroup, glycerol backbone, and carbonyl regions. % added because the information where we scale charges was missing before.
 The lipids with a total charge of -0.75 are also neutralized
@@
 to scale the corresponding parameters of atoms 
-only in the headgroup, glycerol backbone, and carbonyl regions \cite{melcr18}. % Can be shortened by saying it is the same set as that for scaling charges.
+in the headgroup, glycerol backbone, and carbonyl regions \cite{melcr18}. % Can be shortened by saying it is the same set as that for scaling charges.
 Further optimization of parameters was not done in this work. 
@@
 with the bilayer normal and the angular brackets define a time average on a time scale of approximately 1 microsecond.
-The experiment was done using a Bruker Avance III 400 spectrometer operating at a 1H Larmor frequency of 400.03 MHz
+The experiment was done using a Bruker Avance III 400 spectrometer operating at a $^1$H Larmor frequency of 400.03 MHz
 equipped with a standard 4~mm CP-MAS HXY probe. The set up of the R-PDLF experiment was the following
@@
 resolution of $\pm$0.01. The C--H bond order parameter determined from a given dipolar
-splitting, $\Delta\nu$ (see e.g. Fig. \ref{R-PDLFslices} in \todo{Is this the correct figure?}) is equal to $\Delta\nu/(0.315\times21.5 kHz)$, where 0.315 is
+splitting, $\Delta\nu$ (see e.g. Fig. \ref{R-PDLFslices} in supplementary information) is equal to $\Delta\nu/(0.315\times21.5 kHz)$, where 0.315 is
 the scaling factor of the R18 recoupling sequence and 21.5 kHz is the maximum $^1$H-$^{13}$C dipolar coupling for a C--H bond.
 The refocused-INEPT transfer \cite{morris79,burum80} %[G. Morris and R. Freeman, J. Am. Chem. Soc. 101, 760 (1979);
-was used for transferring the polarization from $^1$H nuclei to the covalently bond $^{13}$C nuclei with the delays $\tau_1 = 1.94$~ms and $\tau_2 = 0.97$~ms
+was used for transferring the polarization from $^1$H nuclei to the covalently bond $^{13}$C nuclei with the delays of $\tau_1 = 1.94$~ms and $\tau_2 = 0.97$~ms
 (set as multiples of the MAS rotation period). The pulses for the refocused-INEPT sequence had a nutation frequency equal to 63.45~kHz.
@@
 
+\todo{We need sample preparation, temperature and other potentially missing details.}
+
 \subsection{Simulation details}
@@
 \centering
-\caption{  MOVE THE WHOLE TABLE TO SI!!! % J: I do not have a strong opinion on this. I think it can stay within the main text in a JCTC paper.
+\caption{  %MOVE THE WHOLE TABLE TO SI!!! 
            List of MD simulations reporting their respective
@@
            and the deployed models.
-           The fatty acid chains for all lipids are palmitoyl (sn-1) and oleoyl (sn-2).
+           The fatty acid chains for all lipids are palmitoyl ({\it sn}-1) and oleoyl ({\it sn}-2).
          }\label{tbl:sim-list}
@@
 The simulated systems are listed in Table~\ref{tbl:sim-list}
-and the used simulation parameters in Table~\ref{tbl:mdpar}. 
-All simulations were ran for a minimum of $1 \, \mu$s,
+and the used simulation parameters in Table~\ref{tbl:mdpar} in supplementary information. 
+All simulations were ran for a minimum of $1 \, \mu$s at 298 K,
 and the first  $50 \,$ns were omitted as an equilibration period. 
@@
 the ECC-POPC parameters~\cite{melcr18} (available from Ref. \citenum{ECC-POPC_nacl_cacl2_files}), 
-the ECC-POPS parameteres derived in this work (see above), 
+the ECC-POPS parameteres derived in this work (see above for details and Ref. \citenum{ecclipids_pcps_nacl_kcl_series} for parameters), 
 the SPC/E~\cite{Berendsen1987} water model, and ECC-ions \cite{martinek17, kohagen16, Pluharova2014} were used.
@@
 ECC concept \cite{leontyev11,leontyev14}.
-\begin{table}[tbp]
-  \caption{MOVE TABLE TO SI AND MENTION ONLY THE KEY INFO IN 1 SENTENCE IN THE MAIN TEXT  % J: I am fine with this, the simulation parameters are rather standard.
-There is nothing to be done in the text, the reference to this table will only obtain prefix "S".
-Simulation parameters}
-  \label{tbl:mdpar}
-  \begin{tabular}{ll}
-    simulation property & parameter   \\
-    \hline
-    time-step           & 2~fs         \\
-    equilibration time  & 50~ns  \\
-    total simulation time     & $\geq 1 \mu$s  \\
-    temperature         & 298~K       \\
-    thermostat          & v-rescale  \cite{bussi07}   \\
-    barostat            & Parrinello-Rahman, semi-isotropic \cite{parrinello81} \\
-    long-range electrostatics & PME  \cite{darden93}  \\
-    cut-off scheme      & Verlet \cite{Pall13}      \\
-    Coulomb and VdW cut-off & 1.0~nm \\
-    constraints         & LINCS, only hydrogen atoms \cite{hess97} \\
-    constraints for water & SETTLE  \cite{miyamoto92} \\
-    \hline
-  \end{tabular}
-\end{table}
 
@@
 the headgroup and glycerol backbone region of a POPS lipid bilayer are already available in the literature \cite{kucerka14,NMRlipidsIV}.
-Here, we measure also the C--H bond order parameters of the acyl chain region from a sample of POPS multi-lamellar vesicles by means of a $^1$H--$^{13}$C dipolar recoupling 2D NMR experiment.
-More specifically, as described in detail in the Methods section, we performed R-PDLF 2D spectroscopy~\cite{dvinskikh04} to obtain spectra correlating chemical shifts (direct dimension) and C--H bond dipolar couplings (indirect dimension) from which the C--H bond order parameters defined in eq.~\ref{OPequation} can be determined.
+Here, we measure also the C--H bond order parameters of the acyl chain region from multi-lamellar POPS vesicles using $^1$H--$^{13}$C
+dipolar recoupling 2D NMR experiment. As described in the Methods section, we performed R-PDLF 2D spectroscopy~\cite{dvinskikh04} to
+obtain a spectra correlating chemical shifts (direct dimension) and C--H bond dipolar couplings (indirect dimension) from which the C--H bond order parameters defined in eq.~\ref{OPequation} can be determined.
 The assignment of the peaks in the $^{13}$C chemical shift dimension to acyl chain carbon sites was done based on previous work {\cite{ferreira13}} and is shown in Figs.~\ref{INEPT}~and~\ref{R-PDLF}.
 The dipolar splittings for the assigned carbons that were used to calculate their C--H bond order parameter magnitudes are shown in Fig.~\ref{R-PDLFslices}.
-Because the complete assignment of the peaks in the crowded spectral region between 29-31 ppm was not possible due to the chemical shift overlap of different carbons (~\ref{R-PDLF}), we partially assigned the acyl chain order parameters of this region to mimic the profile predicted by the MD simulations (Fig.~\ref{simVSexpNOions_POPS}).
+Because the complete assignment of the peaks in the crowded spectral region between 29-31 ppm was not possible due to the chemical shift overlap of different carbons (Fig.~\ref{R-PDLF}), we partially assigned the acyl chain order parameters of this region to mimic the profile predicted by the MD simulations (Fig.~\ref{simVSexpNOions_POPS}).
 
@@
   %\includegraphics[width=\figwidth]{../img/POPSstructure.pdf} 
-  \includegraphics[width=\figwidth]{../img/formF_OPs_APLinset_POPSchemfig.pdf} 
+  \includegraphics[width=3.33in]{../img/formF_OPs_APLinset_POPSchemfig.pdf} 
 \hfill
@@
     ECC-POPS/ECC-ions \cite{martinek17, Pluhackova2016} compared with experiments~\citep{kucerka14} at 298~K. 
-    \textbf{(B)} Order parameters of POPS head group, glycerol backbone and acyl chains  
+    \textbf{(B)} Order parameters of POPS headgroup, glycerol backbone and acyl chains  
     from the same simulations 
@@
   \centering 
-  \includegraphics[width=\figwidthsmall]{../img/ecc_pops/density_profiles_na-k-counterions_wat_phos_compar_purePOPS_ecclipids-lipid17.pdf}
+  \includegraphics[width=3.33in]{../img/ecc_pops/density_profiles_na-k-counterions_wat_phos_compar_purePOPS_ecclipids-lipid17.pdf}
   \caption{\label{fig:POPS-counterions-dens}
@@
 In conclusion, the stuctural quality of ECC-POPS is similar to the other currently available
-lipid force fields \cite{botan15, catte16, Pluhackova2016, nmrlipids_proj4},
+lipid force fields \cite{botan15, catte16, Pluhackova2016,NMRlipidsIV},
 giving good agreement with experiments for acyl chain conformations and bilayer dimensions,
@@
 
-\subsection{Binding of counterions to POPS and POPC and interactions between their head groups}
+\subsection{Binding of counterions to POPS and POPC and interactions between their headgroups}
 Binding affinity of monovalent ions to lipid bilayers depends strongly on
@@
   %\includegraphics[height=\figheightsmall]{../img/ecc_pops/order_parameters_changes_ecc-lip_L14_A-B-PN-COO_POPS_nacl.pdf} 
-  \includegraphics[width=\figwidthfull]{../img/deltaOP_nacl_kcl_PC-PS.pdf} 
+  \includegraphics[width=7in]{../img/deltaOP_nacl_kcl_PC-PS.pdf} 
   \caption{\label{fig:delta_ordPar_monoval_PCPS}
-    Changes of the head group order parameters, and the angles of P--N and ${\rm{C}}_\beta-{\rm{C}}_\gamma$ (stars) vectors
+    Changes of the headgroup order parameters, and the angles of P--N and ${\rm{C}}_\beta-{\rm{C}}_\gamma$ (stars) vectors
     with respect to the membrane normal of POPC (A, B) and POPS (C, D) in a POPC:POPS (5:1) bilayer 
     as a function of \ce{KCl} and \ce{NaCl} concentration from ECC-lipids and Lipid17/Dang simulations 
-    compared with experimental values from Ref. \citenum{roux90} (signs from Ref. \citenum{ferreira16}) at 298 K.
+    compared with experimental values from Ref. \citenum{roux90} (signs from Refs.~\citenum{ferreira16} and~\citenum{NMRlipidsIV}) at 298 K.
+    Because experimental data with \ce{NaCl} is not available for POPC, the data for KCl and \ce{LiCl} (B, dashed lines)
+    are shown as lower and upper bounds, respectively, for the response to \ce{NaCl}. 
     The y-axis for the $\alpha$-carbon results of POPS (C and D, middle) is shifted
@@
     parameter value from pure POPS is at zero to correctly illustrate the significant forking.
-    Because experimental data with \ce{NaCl} is not available for POPC, the data for KCl and \ce{LiCl} (B, dashed lines)
-    are shown as lower and upper bounds, respectively, for the response to \ce{NaCl}. 
     Error bars are not visible for ECC-lipids simulations because they are smaller than the point size.
-    Inset in A shows the labeling of carbon segments in a POPC head group.}
+    Inset in A shows the labeling of carbon segments in a POPC headgroup.
+    For the ion density profiles, see Fig. \ref{fig:nacl-dens_PCPS} in supplementary information.
+  }
 \end{figure}
@@
 %and the change in P--N vector is negligible
-in both force fields (Fig.~\ref{fig:delta_ordPar_monoval_PCPS}).
+in both force fields (Fig.~\ref{fig:delta_ordPar_monoval_PCPS} A).
 The slightly overestimated change of the $\alpha$-carbon
 order parameter in the Lipid17/Dang simulations may be due to the
-deeper penetration of \ce{K+} into the bilayer (Fig.~\ref{fig:delta_ordPar_NaCl_PC-PS_mix_PC}).
+deeper penetration of \ce{K+} into the bilayer (Fig.~\ref{fig:delta_ordPar_NaCl_PC-PS_mix_PC} A).
 Experimental data with the additional amount of \ce{Na+} in POPC:POPS (5:1) mixture is not available,
 but \ce{Li+} and \ce{K+} results give the lower and the upper bounds, respectively, to the sodium binding
-affinity (Fig.~\ref{fig:delta_ordPar_monoval_PCPS}). In ECC-lipids simulations,
+affinity (Fig.~\ref{fig:delta_ordPar_monoval_PCPS} B). In ECC-lipids simulations,
 the response of POPC headgroup order parameters to the additional sodium is close to the
@@
 increasing amount of POPS in different simulations (Fig~\ref{fig:delta_ordPar_NaCl_PC-PS_mix_PC}).
-In experiments, the head group order parameters of PC lipids increase upon addition of
+In experiments, the headgroup order parameters of PC lipids increase upon addition of
 negatively charged PS lipids, as predicted by the electrometer concept~\cite{seelig87,scherer87}.
@@
   %\includegraphics[height=\figheightsmall]{../img/POPCstructure.pdf} 
-  \includegraphics[width=\figwidthfull]{../img/deltaOP_mix_PC-PS.pdf} 
+  \includegraphics[width=7in]{../img/deltaOP_mix_PC-PS.pdf} 
   \caption{\label{fig:delta_ordPar_NaCl_PC-PS_mix_PC} 
@@
     The density profiles of phosphate groups and water are divided by 4 and 100, respectively.  
-    \textbf{(B, C)} The POPC head group order parameters and the P--N vector angle
+    \textbf{(B, C)} The POPC headgroup order parameters and the P--N vector angle
     with respect to the membrane normal as a function of POPS content in a bilayer
-    from ECC-lipids and Lipid17/Dang simulations with \ce{Na^+} (C) and \ce{K^+} (B) counterions.
-    Experimental order parameter values are from Ref. \citenum{scherer87}
-    and the signs from Ref.~\citenum{ferreira16} 
-    (only \ce{Na^+} counterions, but shown also in the left plots (B) for \ce{K+}).
+    from ECC-lipids and Lipid17/Dang simulations with \ce{K^+} (B) and \ce{Na^+} (C) counterions.
+    Experimental order parameter values with \ce{Na^+} are from Ref. \citenum{scherer87}
+    and the signs from Ref.~\citenum{ferreira16}. Because experimental values with \ce{K^+} are
+    not available, the data with \ce{Na^+} is shown with dashed lines in (B).
     Error bars are not visible for most of the simulation points because they are smaller than the point size.
@@
 The behaviour of POPS headgroup can be further characterized by monitoring its order parameters in different
-lipid mixtures and ionic conditions~\cite{NMRlipidsIV,roux90}.
+lipid mixtures and ionic conditions (Figs.~\ref{fig:delta_ordPar_monoval_PCPS} and \ref{fig:delta_ordPar_NaCl_PC-PS_mix_PS})~\cite{NMRlipidsIV,roux90}.
 In experiments, the headgroup order parameters of POPS are almost unchanged upon increasing the POPC content or monovalent ion concentration, 
@@
 
-We estimate the influence of negatively charged POPS on the counterion binding affinity
+We quantify the influence of negatively charged POPS on the counterion binding affinity
 by comparing the relative surface excesses with respect to water, $\Gamma ^{w} _{\rm ion}$,
@@
 $\Gamma ^{w} _{\rm K}=-0.123 \pm 0.005 \mathrm{nm}^{-2}$ in POPC:POPS (5:1) mixture (Fig.~\ref{fig:nacl-dens_PCPS}).
-%
-% J: I deemed the following sentences need improvement, my suggestion is below.
-%
-%As discussed above, the sodium binding to POPC:POPS (5:1) mixture may be slightly overestimated also in
-%ECC-lipids simulations. The response of headgroup order parameters to the addition of NaCl are
-%slightly overestimated also in ECC-POPC simulations (Fig 3 in Ref.~\citenum{melcr18}),
-%suggesting that similar small overestimation of sodium binding may also occur to a pure POPC bilayer.
-%Therefore, the overestimation may arise from the sodium interactions with PC headgroup and
-%increase in binding affinity due to PS lipids would not be affected.
-Slightly overestimated binding of \ce{Na+} to a POPC bilayer can be deduced from
-the response of headgroup order parameters to the addition of NaCl (Fig~3 in Ref.~\citenum{melcr18}) 
-and similarly also that to a POPC:POPS (5:1) mixture from the discussion above. 
-In spite of this small inaccuracy
-the increase in binding affinity of \ce{Na+} due to PS lipids remains unaffected.
+The increase in sodium binding affinity upon addition of anionic POPS is not expected to depend on
+the slightly overestimated \ce{Na+} binding in ECC-lipid simulations, because the
+inaccuracy is similar for pure POPC (Fig~3 in Ref.~\citenum{melcr18}) and POPC:POPS (5:1) mixture (Fig. \ref{fig:delta_ordPar_monoval_PCPS}).
 
@@
 from POPC:POPS (5:1) mixture as a function of CaCl$_2$ concentration 
-can be used to evaluate the calcium binding affinity to lipid bilayers containing PS lipids \cite{roux90, NMRlipidsIV}.
+can be used to evaluate the calcium binding affinity to lipid bilayers containing PS lipids \cite{roux90,NMRlipidsIV}.
 The decrease of the PC headgroup order parameters in this mixture upon addition of \ce{CaCl2} 
@@
 indicating that the lower binding affinity than in Lipid17/Dang simulations is more realistic (Fig.~\ref{fig:cacl-dens_PCPS}).
-Therefore, we use ECC-lipids simulations to estimate the influence of negative charged POPS on
+Therefore, we use ECC-lipids simulations to quantify the influence of negative charged POPS on
 calcium binding affinity to lipid bilayers.
@@
 $\Gamma^{w}_{Ca} = 0.24\mathrm{nm^{-2}}$ ($\sim 4.8$ phospholipids per bound \ce{Ca^{2+}}) for the POPC:POPS (5:1) bilayer with 409 mM CaCl$_2$.
+Interestingly, the calcium residence times are 3-4 times longer in POPC:POPS (5:1) mixture (Fig.~\ref{fig:hist_residence_times}) than in pure POPC \cite{melcr18}.
+%in line with increased relaxation times of fluorescence probes from experiments \cite{melcrova16}.
+%SAMULI: I commented this out because the cited reference does not discuss residence times.
 Besides lower binding affinity, ECC-lipids simulations yield smaller error bars for order parameters and shorter residence times (Fig.~\ref{fig:hist_residence_times}) 
@@
 longest observed residence times are $141\,\mathrm{ns}$ and $485\,\mathrm{ns}$, respectively (Fig.~\ref{fig:hist_residence_times}).
-Interestingly, the calcium residence times are 3-4 times longer in POPC:POPS (5:1) mixture than in pure POPC
-in line with increased relaxation times of fluorescence probes from experiments. \cite{melcrova16}
 
@@
   %\includegraphics[height=1.5\figheightsmall]{../img/ecc_pops/order_parameters_changes_ecc-lip_L14_A-B-PN-COO_POPS_cacl.pdf} 
-  \includegraphics[width=\figwidthfull]{../img/deltaOP_cacl_PC-PS.pdf} 
+  \includegraphics[width=7in]{../img/deltaOP_cacl_PC-PS.pdf} 
   \end{tabular}
@@
     Number density profiles of \ce{Ca^{2+}} and \ce{Cl^-} ions and \ce{Na^{+}} counterions 
-    along the normal of the negatively charged membrane with the composition 5\,PC:1\,PS
-    at 80~mM (A, top) and 200~mM (A, bottom) added buffer concentrations of \ce{CaCl2} from simulations with ECC-lipids (solid) and Lipid17 (dashed). 
-    In order to visualize the density profiles with a comparable scale 
-    the density profile of~\ce{Ca^{2+}} ions are divided by 2, and 
-    the density profiles of phosphate groups and water are divided by 4 and 100, respectively.  
+    along the normal of POPC:POPS (5:1) bilayer with 80~mM (A, top) and 200~mM (A, bottom)
+    of \ce{CaCl2} from simulations with ECC-lipids (solid) and Lipid17 (dashed). 
+    The~\ce{Ca^{2+}}, phosphate and water densities are divided by 2, 4 and 100, respectively.  
     \textbf{(B, C)}
-    Changes of the head group order parameters, and the angles of P--N (circles) and C$_\beta$--C$_\gamma$ (stars) vectors 
+    Changes of the headgroup order parameters, and the angles of P--N (circles) and C$_\beta$--C$_\gamma$ (stars) vectors 
     with respect to the membrane normal of POPC (B) and POPS (C) in a POPC:POPS (5:1) bilayer 
-    as a function of \ce{CaCl$_2$} added buffer concentration from ECC-lipids and Lipid17/Dang simulations 
+    as a function of \ce{CaCl$_2$} from ECC-lipids and Lipid17/Dang simulations 
     compared with experimental values from Ref. \citenum{roux90} (signs from Refs. \citenum{ferreira16} and \citenum{NMRlipidsIV}) at 298 K.
@@
   %\includegraphics[width=8cm]{../img/stoichiometry_CaCl2_comparison_Ecc-lipids_PC-vs-PCPS.pdf} 
-  \includegraphics[width=\figwidthfull]{../img/populations_stoichiometry.pdf} 
+  \includegraphics[width=7in]{../img/populations_stoichiometry.pdf} 
   \end{tabular}
@@
     when bound to PC only, PS only or simultaneously to both
-    calculated from ECC-lipids simulation of POPC:POPS (5:1) bilayer with 409 mM CaCl$_2$ added buffer concentration. 
+    calculated from ECC-lipids simulation of POPC:POPS (5:1) bilayer with 409 mM CaCl$_2$. 
 	In the case of simultaneous binding to both PC and PS,
@@
     The threshold for counting coordinated lipids in a complex with \ce{Ca^{2+}} was set to $0.3\,\mathrm{nm}$, 
-    as the distance between the cation and the oxygen atoms of the lipids. 
+    for the distance between the cation and the oxygen atoms of the lipids. 
     Previously published simulation data \cite{melcr18} for pure POPC bilayers were taken directly from Ref.~\citenum{ECC-POPC_nacl_cacl2_files}. 
@@
 imperfect structures of the lipid headgroup (Fig.~\ref{simVSexpNOions_POPS}).
-
-\todo{MOVE THIS WHOLE PARAGRAPH TO SI}
-The POPS headgroup order parameters are also related to the changes of average orientation of P--N and
-C$_{\beta}$--C$_{\gamma}$ vectors in the PS headgroup, which are smaller and more systematic
-in ECC-lipids than in Lipid7/Dang simulations (Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}).
-Upon addition of 620~mM CaCl$_2$ to POPC:POPS (5:1) mixture,
-the average orientation of P--N vectors in both POPC and POPS headgroups tilts more perpendicular to the membrane surface
-by 11$^\circ$ and  3$^\circ$, respectively, in ECC-lipids simulations (Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}),
-suggesting that the PS headgroup orientation is less sensitive to the bound calcium than PC.
-On the other hand, the average orientation of the C$_{\beta}$--C$_{\gamma}$ vector in PS headgroup
-tilts to the opposite direction by $11^\circ$ in the same system
-due to the attraction between \ce{COO^-} groups in PS and the cations adsorbed at the phosphate region.
 
@@
 calcium ions coordinate with even four distinct PS lipids \cite{hallock18}, while coordination with more than two PS lipids
-was very rare in our ECC-lipids simulations (Fig.~\ref{fig:cacl_complexes}).
+is very rare in our ECC-lipids simulations (Fig.~\ref{fig:cacl_complexes}).
 %Furthermore, the POPS:POPC:Ca$^{2+}$ (7:3:7) mixture with relatively high
@@
 %not observed with lower PS content used in ECC-lipids simulations in this work \cite{hauser77,kurland79,hauser85,feigenson86,mattai89,roux90,roux91}.
-%While POPC interacts with the calcium cations almost entirely through its phosphate group 
-%in both neutral \cite{melcr18} and negatively charged membranes, 
-%the inteaction with POPS lipids happens more through the \ce{COO^-} group than through the \ce{PO4} group. 
-%Interactions of \ce{Ca^{2+}} with carbonyl groups are also present for both POPC and POPS,
-%however, they are always accompanied by interactions with phosphate groups.
-
-%Significant coordination of calcium with carbonyl groups of PS and PC lipids observed in previous
-%simulations with the Berger force field \cite{melcrova16} %\cite{berger97,mukhopadhyay04}
-%was not observed in ECC-lipids simulations in here nor in the previous work with pure POPC \cite{melcr18}.
-%Overestimated coordination of cations with carbonyl groups is a potential source for
-%the overestimated cation binding in Berger force field \cite{catte16,NMRlipidsIV}. 
-%In CHARMM36 simulations with the NBfix, calcium mainly interacted with only
-%carboxylate of PS lipids  \cite{valentine18}, which is a potential source for the
-%underestimated calcium binding affinity to lipid bilayer with this force field \cite{NMRlipidsIV}.
-
-%on average 41\% of the total population of bound calcium cations is in contact with PS lipids
-%with 8\% bound exclusively only to them even though the negatively charged membrane contains only 17\% of POPS (Fig.~\ref{fig:cacl_complexes}).
-%%Therefore, calcium ions bind more likely to negatively charged PS lipids than neutral PC lipid in a mixed bilayer, as expected.
-%	For comparison, in a pure POPC bilayer with 350 mM CaCl$_2$, 
-%	67\% of bound \ce{Ca^{2+}} is exclusively in contact with \ce{PO_4} group. \cite{melcr18}
-
-%This can explain the diminished flexibility of the PS head group 
-%from simultaneous binding of \ce{Ca^{2+}} to the \ce{COO^-} group in PS and a phosphate group in PC 
-%as the most probable configuration (Fig.~\ref{fig:cacl_complexes}).
-
-%In contrast, the combination of NMR experiments 
-%and MD simulations with the same model, CHARMM36, but without the correction for calcium binding
-%also reveal long-lived interactions of \ce{Ca^{2+}} with \ce{PO4} groups. \cite{hallock18}
-%Although the ratios of the life times of coordination follow the same trends as in our work,
-%the lengths of simulation trajectories are insufficient 
-%to yield converged results in this aspect (see SI and Refs.~\cite{catte16, melcr18, NMRlipidsIV}).
+
+%\todo{MOVE THIS WHOLE PARAGRAPH TO SI}
+The improved accuracy in ECC-lipids simulations motivates also the more detailed analysis of
+POPS headgroup response to the bound calcium ions, which is potentially important in calcium regulated
+lipid-protein interactions \cite{leventis10}.
+%The POPS headgroup order parameters are also related to the changes of average orientation of P--N and
+%C$_{\beta}$--C$_{\gamma}$ vectors in the PS headgroup, which are smaller and more systematic
+%in ECC-lipids than in Lipid7/Dang simulations (Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}).
+Upon addition of 620~mM CaCl$_2$ to POPC:POPS (5:1) mixture,
+the average orientation of P--N vectors in both POPC and POPS headgroups tilts more perpendicular to the membrane surface
+by 11$^\circ$ and  3$^\circ$, respectively, in ECC-lipids simulations (Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}),
+suggesting that the PS headgroup orientation is less sensitive to the bound calcium than PC.
+On the other hand, the average orientation of the C$_{\beta}$--C$_{\gamma}$ vector in PS headgroup
+tilts to the opposite direction by $11^\circ$ in the same system, probably
+due to the attraction between \ce{COO^-} groups and cations adsorbed at the phosphate region.
 
@@
 Development and Innovations Infrastructures. 
-O.H.S.O. acknowledges financial support from 
+O.H.S.O. acknowledges financial support from Academy of Finland (315596),
 Integrated Structural Biology Research Infrastructure of
-Helsinki Institute of Life Science (Instruct-HiLIFE). 
+Helsinki Institute of Life Science (Instruct-HiLIFE), and
+CSC-IT center for science for computational resources.
 \end{acknowledgement}

### 2019-08-13 Manuscript/manuscript.tex

@@
 \newlength{\figwidthfull}
-\setlength{\figwidthfull}{14 cm} 
+\setlength{\figwidthfull}{16 cm} 
 \newlength{\figheightsmall}
@@
   %\includegraphics[height=\figheightsmall]{../img/ecc_pops/order_parameters_changes_ecc-lip_L14_A-B-PN-COO_POPS_nacl.pdf} 
-  \includegraphics[width=7in]{../img/deltaOP_nacl_kcl_PC-PS.pdf} 
+  \includegraphics[width=\figwidthfull]{../img/deltaOP_nacl_kcl_PC-PS.pdf} 
   \caption{\label{fig:delta_ordPar_monoval_PCPS}
@@
   %\includegraphics[height=\figheightsmall]{../img/POPCstructure.pdf} 
-  \includegraphics[width=7in]{../img/deltaOP_mix_PC-PS.pdf} 
+  \includegraphics[width=\figwidthfull]{../img/deltaOP_mix_PC-PS.pdf} 
   \caption{\label{fig:delta_ordPar_NaCl_PC-PS_mix_PC} 
@@
   %\includegraphics[height=1.5\figheightsmall]{../img/ecc_pops/order_parameters_changes_ecc-lip_L14_A-B-PN-COO_POPS_cacl.pdf} 
-  \includegraphics[width=7in]{../img/deltaOP_cacl_PC-PS.pdf} 
+  \includegraphics[width=\figwidthfull]{../img/deltaOP_cacl_PC-PS.pdf} 
   \end{tabular}
@@
   %\includegraphics[width=8cm]{../img/stoichiometry_CaCl2_comparison_Ecc-lipids_PC-vs-PCPS.pdf} 
-  \includegraphics[width=7in]{../img/populations_stoichiometry.pdf} 
+  \includegraphics[width=\figwidthfull]{../img/populations_stoichiometry.pdf} 
   \end{tabular}

### 2019-08-14 Manuscript/manuscript.tex

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
-\documentclass[journal=jpcbfk,manuscript=article]{achemso}
+\documentclass[journal=jctcce,manuscript=article]{achemso}
 
@@
 
-%\author{Andrew N. Other} 
-%\altaffiliation{A shared footnote} 
-%\author{Fred T. Secondauthor} 
-%\altaffiliation{Current address: Some other place, Othert\"own, 
-%Germany} 
-%\author{I. Ken Groupleader} 
-%\altaffiliation{A shared footnote} 
-%\email{i.k.groupleader@unknown.uu} 
-%\phone{+123 (0)123 4445556} 
-%\fax{+123 (0)123 4445557} 
-%\affiliation[Unknown University] 
-%{Department of Chemistry, Unknown University, Unknown Town} 
-%\alsoaffiliation[Second University] 
-%{Department of Chemistry, Second University, Nearby Town} 
-%\author{Susanne K. Laborator} 
-%\email{s.k.laborator@bigpharma.co} 
-%\affiliation[BigPharma] 
-%{Lead Discovery, BigPharma, Big Town, USA} 
-%\author{Kay T. Finally} 
-%\affiliation[Unknown University] 
-%{Department of Chemistry, Unknown University, Unknown Town} 
-%\alsoaffiliation[Second University] 
-%{Department of Chemistry, Second University, Nearby Town}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
@@
   Classical molecular dynamics (MD) simulations can potentially give detailed description of lipid-ion interactions,
-  but the results strongly depend on the used force field.
-  Here, we apply the electronic continuum correction (ECC) to the Amber Lipid17 parameters of 1-palmitoyl-2-oleoyl-{\it sn}-glycero-3-phospho-L-serine (POPS) lipid to improve its
+  but the results strongly depend on the used force field. Here, we apply the electronic continuum correction (ECC) to
+  the Amber Lipid17 parameters of 1-palmitoyl-2-oleoyl-{\it sn}-glycero-3-phospho-L-serine (POPS) lipid to improve its
   interactions with Na$^+$ and Ca$^{2+}$ ions.
@@
 
-%There are force field parameters for neutral lipids PC and PE 
-%with explicit polarizability using the Classical Durde model. \cite{chowdhary13, chowdhary17}
-%While such a model shows a potential improvement over non-polarizable force fields from many perspectives -- 
-%for example accounting for polarizability has a significant impact on the dipole potential of a membrane \cite{harder2009} -- 
-%the non-polarizable version of the same force field 
-%performes comparably well at a fraction of the computational cost 
-%of the models with explicit polarization. \citep{lucas12,chowdhary13}
-
 Here, we extend the ECC approach also to the negatively charged 1-palmitoyl-2-oleoyl-{\it sn}-glycero-3-phospho-L-serine (POPS) lipid.
@@
  biomolecules containing charged groups such as proteins\cite{Pluharova2014, martinek17, duboue2018insulin, Mason2019, Duboue2018MgZn}.
-In addition to the charge, the Lennard-Jones~$\sigma$ was scaled with a factor $ 0.9 < f_\sigma \leq 1$
-\todo{JOE: The given range is not correct, e.g. scaling of POPC (0.89), Mg2+ (0.72 and 0.82). 
-Extending the lower bound to 0.8 fits the cases I know of with the exception of the small-Mg2+. }
+In addition to the charge, the Lennard-Jones~$\sigma$ was scaled with a factor $ 0.72 < f_\sigma \leq 1$
 to improve the description of hydration properties of ions with respect to scattering data in these studies.
@@
 and hence, the scaling factor can be larger than the theoretically derived value for ions in water, $f_q \approx 0.75$.  
-% above does not apply to molecules and moieties with non-zero total charge (hence the edit). 
 Before the ECC theory was rigorously derived, similar idea was already employed in early classical 
@@
 ions $f_q = 0.75$ to the partial charges of POPS 
-in the headgroup, glycerol backbone, and carbonyl regions. % added because the information where we scale charges was missing before.
-The lipids with a total charge of -0.75 are also neutralized
+in the headgroup, glycerol backbone, and carbonyl regions. The lipids with a total charge of -0.75 are also neutralized
 by the counterion charges of +0.75 in simulations with ECC-ions \cite{Pluharova2014, kohagen16, martinek17}.
@@
 to scale the corresponding parameters of atoms 
-in the headgroup, glycerol backbone, and carbonyl regions \cite{melcr18}. % Can be shortened by saying it is the same set as that for scaling charges.
+in the headgroup, glycerol backbone, and carbonyl regions \cite{melcr18}. 
 Further optimization of parameters was not done in this work. 
@@
 the scaling factor of the R18 recoupling sequence and 21.5 kHz is the maximum $^1$H-$^{13}$C dipolar coupling for a C--H bond.
-The refocused-INEPT transfer \cite{morris79,burum80} %[G. Morris and R. Freeman, J. Am. Chem. Soc. 101, 760 (1979);
+The refocused-INEPT transfer \cite{morris79,burum80} 
 was used for transferring the polarization from $^1$H nuclei to the covalently bond $^{13}$C nuclei with the delays of $\tau_1 = 1.94$~ms and $\tau_2 = 0.97$~ms
@@
 
-\todo{We need sample preparation, temperature and other potentially missing details.}
+As in our previous work~\cite{NMRlipidsIV}, the sample was prepared by first mixing POPS powder
+(1-palmitoyl-2-oleoyl-sn-glycero-3-phospho-L-serine, purchased from Avanti Polar Lipids as sodium salt)
+with water (lipid:water 60:40 wt-\%) in an Eppendorf tube. The mixture was repeatedly centrifuged and
+stirred (approximately 5 to 6 times) until a homogeneous viscous fluid was visually observed.
+Then 20 mg of the sample was transferred to an NMR insert suitable for 4 mm NMR rotors.
+Experiments were done at 298~K.
 
@@
 \centering
-\caption{  %MOVE THE WHOLE TABLE TO SI!!! 
-           List of MD simulations reporting their respective
+\caption{  List of MD simulations reporting their respective
            lipid composition,
@@
   \centering 
-  %\includegraphics[width=\figwidth]{../img/ecc_pops/Order-parameters_form-factors_exp-L17-ECC-lipids.pdf}
-  %\includegraphics[width=\figwidth]{../img/POPSstructure.pdf} 
   \includegraphics[width=3.33in]{../img/formF_OPs_APLinset_POPSchemfig.pdf} 
@@
 This explains also the larger area per molecule (57~\AA$^2$) reported for Lipid17 POPS simulations
-%being closer to the experimental value, 
 with {\AA}qvist ions~\cite{NMRlipidsIV}.
@@
 
-%SAMULI: I have moved this figure back to the main text
 \begin{figure}[tbp!] 
@@
 
-%% moved into the figure 1 as an inset
-%\begin{table}[tb!] 
-%\centering
-  %\caption{Area per lipid of POPS bilayers, $A_L$, from simulations with different force fields at 298~K with \ce{Na^+} counterions. \label{tab:apls} } 
-  %\begin{tabular}{l|c } 
-    %\multicolumn{2}{c}{POPS} \\
-    %model          & $A_L$ / Å$^2$    \\ 
-    %\hline 
-    %Lipid17/Dang              & 53.5$\pm$ 1.7   \\ 
-%%    Lipid17/ff99              & 57.9$\pm$ 1.7   \\ 
-%%    \hline 
-    %ECC-POPS/ECC-ions         & 59.8$\pm$ 1.8   \\ 
-%%    \hline 
-    %experiment (SDP model) \citep{kucerka14} & 62.7$\pm$ 1.3  \\ 
-%%    \hline 
-  %\end{tabular} \\
-%\todo{J: This table will become a small inset plot in the top of Fig.~\ref{simVSexpNOions_POPS} }
-%%\end{table} 
-%
-
 Despite the success in simulating bilayer dimensions and acyl chain conformations,
@@
 but not in full agreement with experiments (Fig.~\ref{simVSexpNOions_POPS}).
-
-% SAMULI: I commented this out to simplify the manuscript
-%
-%Especially, the experimentally observed more negative values of the $\beta$-carbon
-%and the larger forking (difference of order parameters between hydrogens attached to the
-%same carbon \cite{ollila16}) of the $\alpha$-carbon in POPS than in POPC \cite{NMRlipidsIV}
-%are not reproduced by neither of the models here. 
-%These differences are previously related
-%to a more rigid structure of PS headgroup \cite{browning80,buldt81} and 
-%the stiffness of the C$_\alpha$-C$_\beta$-C$_\gamma$-O$_\gamma$ dihedral \cite{NMRlipidsIV}
-%plotted in Fig.~\ref{fig:dihedral}. 
-%Interestingely, this dihedral angle is more flexible in ECC-POPS compared to Lipid17 
-%suggesting that its rigidity is overestimated in Lipid17 due to the lack of electronic polarizability,
-%while the same dihedral parameters yield too small forking in ECC-POPS.
 
@@
   \centering 
-  %\includegraphics[height=\figheightsmall]{../img/ecc_pops/order_parameters_changes_ecc-lip_L14_A-B-PN-COO_POPC_kcl.pdf} 
-  %\includegraphics[height=\figheightsmall]{../img/ecc_pops/order_parameters_changes_ecc-lip_L14_A-B-PN-COO_POPC_nacl.pdf} 
-  %\includegraphics[height=\figheightsmall]{../img/ecc_pops/order_parameters_changes_ecc-lip_L14_A-B-PN-COO_POPS_kcl.pdf} 
-  %\includegraphics[height=\figheightsmall]{../img/ecc_pops/order_parameters_changes_ecc-lip_L14_A-B-PN-COO_POPS_nacl.pdf} 
   \includegraphics[width=\figwidthfull]{../img/deltaOP_nacl_kcl_PC-PS.pdf} 
@@
 The small changes of POPC headgroup order parameters with increasing amount of KCl are close to experiments
-%and the change in P--N vector is negligible
 in both force fields (Fig.~\ref{fig:delta_ordPar_monoval_PCPS} A).
@@
 cancels the influence of negatively charged lipids and the increase in PC headgroup
-order parameters is not observed.% \cite{NMRlipidsIV}.
+order parameters is not observed.
 
@@
   \centering 
-  %\includegraphics[height=\figheightsmall]{../img/ecc_pops/density_profiles_na-k-counterions_wat_phos_compar_5PC-1PS_ecclipids-lipid17.pdf}
-  %\includegraphics[height=\figheightsmall]{../img/ecc_pops/order_parameters_changes_A-B_PC-PS_mix_POPC_nacl.pdf} 
-  %\includegraphics[height=\figheightsmall]{../img/ecc_pops/order_parameters_changes_A-B_PC-PS_mix_POPC_kcl.pdf} 
-  %\includegraphics[height=\figheightsmall]{../img/POPCstructure.pdf} 
   \includegraphics[width=\figwidthfull]{../img/deltaOP_mix_PC-PS.pdf} 
@@
 cation binding affinities~\cite{melcr18}.
-%Threshold for counting bound \ce{Ca^{2+}} cations to a membrane
-%was set to $0.3\,\mathrm{nm}$ from oxygen atoms in lipids as in Ref.~\citenum{melcr18}.  
-%Such a quantity compares the adsorption of ions to the adsorption of water molecules at an interface 
-%without the necessity of defining a Gibbs dividing surface. \citep{melcr18, chattorajBOOK}
 The value of $\Gamma ^{w} _{\rm Na}=-0.11 \pm 0.01 \mathrm{nm}^{-2}$ for 1~M sodium in pure ECC-POPC
@@
 The relative surface excess of calcium with respect to water 
-%  previously we used 350mM conc, but I think it is better to compare to a higher concentration to support the whole argument bullet-proof
-%  $\Gamma^{w}_{Ca} = 0.06\mathrm{nm^{-2}}$ ($\sim ?.?$ phospholipids per bound \ce{Ca^{2+}}) for pure POPC bilayer with 346~mM CaCl$_2$ \cite{melcr18, ECC-POPC_nacl_cacl2_files} 
 $\Gamma^{w}_{Ca} = 0.09\mathrm{nm^{-2}}$ ($\sim 5.4$ phospholipids per bound \ce{Ca^{2+}}) for pure POPC bilayer with 467~mM CaCl$_2$ \cite{ECC-POPC_nacl_cacl2_files}
@@
 Interestingly, the calcium residence times are 3-4 times longer in POPC:POPS (5:1) mixture (Fig.~\ref{fig:hist_residence_times}) than in pure POPC \cite{melcr18}.
-%in line with increased relaxation times of fluorescence probes from experiments \cite{melcrova16}.
-%SAMULI: I commented this out because the cited reference does not discuss residence times.
 Besides lower binding affinity, ECC-lipids simulations yield smaller error bars for order parameters and shorter residence times (Fig.~\ref{fig:hist_residence_times}) 
@@
   \begin{tabular}{ c }
-  %\includegraphics[height=1.5\figheightsmall]{../img/ecc_pops/density_profiles_ca_na_k_cl_wat_phos_ecclipids_lipid17_compar_80and200mMCaCl2.pdf}
-  %\includegraphics[height=1.5\figheightsmall]{../img/ecc_pops/order_parameters_changes_ecc-lip_L14_A-B-PN-COO_POPC_cacl.pdf} 
-  %\includegraphics[height=1.5\figheightsmall]{../img/ecc_pops/order_parameters_changes_ecc-lip_L14_A-B-PN-COO_POPS_cacl.pdf} 
   \includegraphics[width=\figwidthfull]{../img/deltaOP_cacl_PC-PS.pdf} 
   \end{tabular}
-%  \begin{tabular}{ c }
-%  \includegraphics[width=3cm]{../img/bound_ca_populations.pdf} \\
-%  \includegraphics[width=3cm]{../img/stoichiometry_CaCl2_comparison_Ecc-lipids_PC-vs-PCPS.pdf} 
-%  \end{tabular}
   \caption{
@@
   }
-%  \caption{ \label{fig:cacl_complexes} 
-%    \textbf{Right top:} Percentages of the bound Ca$^{2+}$ 
-%    in exclusive contact with the given oxygen moiety
-%    when bound to PC only, PS only or simultaneously to both
-%    calculated from ECC-lipids simulation of POPC:POPS (5:1) bilayer with 409 mM CaCl$_2$. 
-%%	The areas represent populations of bound \ce{Ca^{2+}} 
-%%	only to the given moieties exclusively (shades of orange and red)
-%%	or their combinations (CHANGE:blue)
-%%	separately for PC, PS, or PC and PS simultaneously. 
-%%	The areas sum up to total populations at the respective lipids. %
-%	In the case of simultaneous binding to both PC and PS%,
-%	the percentages refer to the moieties in PS.
-%	% i.e., PC can interact with any considered moiety.
-%        For numerical values, see table~\ref{tab:binding}.
-%    \textbf{Right bottom:} Relative probabilities of \ce{Ca^{2+}} ions to coordinate with a certain number of lipids
-%    in pure POPC bilayer with 350 mM CaCl$_2$ and in POPC:POPS (5:1) mixture with 400 mM CaCl$_2$.  
-%    Analysis was done for both systems by considering all lipids (blue and violet) and
-%    for POPC:POPS (5:1) mixture also by considering POPC and POPS lipids separately (green and orange). 
-%    %All lipids were taken into account with the exception of the complexes in light green and orange, 
-%    %for which we counted only contacts with POPC resp POPS from the mixed 5\,PC:1\,PS negatively charged bilayer. 
-%    %The calculated probabilities of the calcium-lipid complexes also reflect only POPC (light green) resp POPS (orange). 
-%    % Probabilities were taken from simulations with comparable bulk concentrations of calcium around 250~mM. 
-%    Clusters of four or more lipids were not observed in either membrane.
-%    The threshold for counting coordinated lipids in a complex with \ce{Ca^{2+}} was set to $0.3\,\mathrm{nm}$, 
-%    as the distance between the cation and the oxygen atoms of the lipids. 
-%    Previously published simulation data \cite{melcr18} for pure POPC bilayers were taken directly from \cite{ECC-POPC_nacl_cacl2_files}. 
-%  }
 \end{figure}
-
-% SAMULI: I separated the figure below, because it was too small to be read when combined with others.
 
@@
   \begin{tabular}{ c }
-  %\includegraphics[width=8cm]{../img/bound_ca_populations.pdf} \\
-  %\includegraphics[width=8cm]{../img/stoichiometry_CaCl2_comparison_Ecc-lipids_PC-vs-PCPS.pdf} 
   \includegraphics[width=\figwidthfull]{../img/populations_stoichiometry.pdf} 
@@
     for POPC:POPS (5:1) mixture also by considering only POPC and POPS lipids separately (green and orange). 
-    %All lipids were taken into account with the exception of the complexes in light green and orange, 
-    %for which we counted only contacts with POPC resp POPS from the mixed 5\,PC:1\,PS negatively charged bilayer. 
-    %The calculated probabilities of the calcium-lipid complexes also reflect only POPC (light green) resp POPS (orange). 
-    % Probabilities were taken from simulations with comparable bulk concentrations of calcium around 250~mM. 
     Clusters of four or more lipids were not observed in either membrane.
@@
 is very rare in our ECC-lipids simulations (Fig.~\ref{fig:cacl_complexes}).
-%Furthermore, the POPS:POPC:Ca$^{2+}$ (7:3:7) mixture with relatively high
-%PS lipid fraction and calcium concentration was used in the previous work \cite{hallock18},
-%which may lead to the complexation of \ce{Ca^{2+}} cations and PS lipids that is
-%not observed with lower PS content used in ECC-lipids simulations in this work \cite{hauser77,kurland79,hauser85,feigenson86,mattai89,roux90,roux91}.
 
-%\todo{MOVE THIS WHOLE PARAGRAPH TO SI}
 The improved accuracy in ECC-lipids simulations motivates also the more detailed analysis of
@@
 lipid-protein interactions \cite{leventis10}.
-%The POPS headgroup order parameters are also related to the changes of average orientation of P--N and
-%C$_{\beta}$--C$_{\gamma}$ vectors in the PS headgroup, which are smaller and more systematic
-%in ECC-lipids than in Lipid7/Dang simulations (Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}).
 Upon addition of 620~mM CaCl$_2$ to POPC:POPS (5:1) mixture,
@@
 using different force fields~\cite{melcrova16,valentine18,hallock18}.
-%measured and simulation data, which enables more unambiguous evaluation of the force field quality \cite{catte16,ollila16}
-%than more indirect comparison in previous studies using 2D infrared spectroscopy \cite{valentine18}, NMR chemical shifts and
-%rotational-echo double-resonance (REDOR) experiments \cite{hallock18} or fluorescent and vibrational sum frequency spectroscopy \cite{melcrova16}.
-%Because the ECC-lipids model gives the best agreement with the experimental headgroup order parameters
-%with various ion conctenrations among the available models, we believe that it also gives the currently best available
-%interpretation of the experimental data. 
 Furthermore, our results pave the way for more realistic MD simulations of anionic biological membranes,
@@
 The present work thus represents another piece in the mosaic that leads to development of an accurate empirical force field for biomolecular simulations involving aqueous salt solutions, proteins, peptides, lipid membranes, and/or nucleic acids, which include electronic polarization effects in a mean field way via charge scaling without extra computational costs compared to standard non-polarizable simulations.
-
-%Therefore, further force field optimization is required for even more accurate MD simulations of cations
-%in the vicinity of negatively charged lipid bilayers.
-
-\listoftodos
 
@@
 
-%The following files are available free of charge. 
-%\begin{itemize} 
-%  \item Filename: brief description 
-%\end{itemize}
+The following files are available free of charge. 
+\begin{itemize} 
+  \item SI.pdf: Additional simulation and experimental details, figures and tables. 
+\end{itemize}
 

### 2019-10-01 Manuscript/manuscript.tex

@@
 The ion model by Dang and coworkers~\cite{smith94,chang1999,dang2006} was used with Amber lipids because
-the default Amber ion parameters \cite{aqvist90} led to the artificial clustering of ions in solution \cite{NMRlipidsIV}.
+the default Amber ion parameters, derived by {\AA}qvist \cite{aqvist90}, led to the artificial clustering of ions in solution \cite{NMRlipidsIV}.
 
@@
 This explains also the larger area per molecule (57~\AA$^2$) reported for Lipid17 POPS simulations
-with {\AA}qvist ions~\cite{NMRlipidsIV}.
-However, the Dang ions are used here to analyze the effect of ECC to the properties of a POPS lipid bilayer because
-{\AA}qvist ions produce known artifacts, like artificial aggregation of ions at larger salt concentrations \cite{kohagen16, chen07, NMRlipidsIV}.
+with {\AA}qvist ion parameters~\cite{NMRlipidsIV}.
+However, the Dang ion parameters are used here to analyze the effect of ECC to the properties of a POPS lipid bilayer because
+{\AA}qvist ion parameters produce known artifacts, like artificial aggregation of ions at larger salt concentrations \cite{kohagen16, chen07, NMRlipidsIV}.
 
@@
 order parameters are quite far from experimental values with Dang (Fig.~\ref{simVSexpNOions_POPS}),
-{\AA}qvist and Joung-Cheatham ions~\cite{NMRlipidsIV}. The values from ECC-POPS simulation are closer,
+{\AA}qvist and Joung-Cheatham ion parameters~\cite{NMRlipidsIV}. The values from ECC-POPS simulation are closer,
 but not in full agreement with experiments (Fig.~\ref{simVSexpNOions_POPS}).
@@
 against experiments using the order parameters of $\alpha$ and $\beta$ carbons in the
-phosphatidylcholine (PC) lipid headgroup (see Fig.~\ref{fig:delta_ordPar_monoval_PCPS} for the labeling) \cite{catte16,melcr18,NMRlipidsIV},
-which decrease proportionally to the bound charge according to the ''electrometer concept'' \citep{seelig87}.
+phosphatidylcholine (PC) lipid headgroup (see Fig.~\ref{fig:delta_ordPar_monoval_PCPS} for the labeling) \cite{catte16,melcr18,NMRlipidsIV}.
+According to the ''electrometer concept'' \citep{seelig87}, these order parameters
+decrease proportionally to the amount of bound positive charge because the headgroup dipole tilts more
+parallel to the mebrane normal.
 Although the response of PS lipid headgroup order parameters to the bound charge is also systematic,
@@
 were overestimated by the Lipid17/Dang simulations (Fig.~\ref{fig:delta_ordPar_CaCl_PCPS})
-and all other tested force fields except CHARMM36 with the recently introduced NBfix for calcium \cite{kim16},
+and all other tested force fields except CHARMM36 with the recently introduced NBfix parameters for calcium \cite{kim16,han2018graph},
 which underestimated the headgroup order parameter response.
@@
 The relative surface excess of calcium with respect to water 
-$\Gamma^{w}_{Ca} = 0.09\mathrm{nm^{-2}}$ ($\sim 5.4$ phospholipids per bound \ce{Ca^{2+}}) for pure POPC bilayer with 467~mM CaCl$_2$ \cite{ECC-POPC_nacl_cacl2_files}
+$\Gamma^{w}_{Ca} = 0.09\mathrm{nm^{-2}}$
+for pure POPC bilayer with 467~mM CaCl$_2$ \cite{ECC-POPC_nacl_cacl2_files}
 is increased to
-$\Gamma^{w}_{Ca} = 0.24\mathrm{nm^{-2}}$ ($\sim 4.8$ phospholipids per bound \ce{Ca^{2+}}) for the POPC:POPS (5:1) bilayer with 409 mM CaCl$_2$.
+$\Gamma^{w}_{Ca} = 0.24\mathrm{nm^{-2}}$
+for the POPC:POPS (5:1) bilayer with 409 mM CaCl$_2$.
+With these concentrations, the total amount of lipids per bound \ce{Ca^{2+}} is
+$\sim 5.4$ in the pure POPC system and $\sim 4.8$ in the POPC:POPS (5:1) mixture.
 Interestingly, the calcium residence times are 3-4 times longer in POPC:POPS (5:1) mixture (Fig.~\ref{fig:hist_residence_times}) than in pure POPC \cite{melcr18}.
@@
 and other tested force fields in our recent work \cite{NMRlipidsIV}, 
-including CHARMM36 with the NBfix correction for calcium
+including CHARMM36 with the NBfix corrections for calcium
 which underestimated the binding affinity.
@@
 and demonstrate the usefulness of ECC also for charged lipids.
+
 The general philosophy behind our approach is to build upon existing non-polarizable force fields making the minimum of necessary changes by scaling only groups bearing a sizeable (partial) charge, followed by slight readjustments of van der Waals radii and possibly also other parameters such as those concerning dihedral angle terms.
 The present work thus represents another piece in the mosaic that leads to development of an accurate empirical force field for biomolecular simulations involving aqueous salt solutions, proteins, peptides, lipid membranes, and/or nucleic acids, which include electronic polarization effects in a mean field way via charge scaling without extra computational costs compared to standard non-polarizable simulations.
+We believe that ECC corrected lipid force fields will be useful in modelling complex biomembranes when 
+explicitly polarizable force fields for lipids \cite{li17,chu18} are not necessary.
 

### 2019-10-16 Manuscript/manuscript.tex
Added example PC:PS:ca2+ structures into Fig6 (main) and FigS6 (SI) added captions to them. Fixed typo in the figure definition "popul...._SI.sla (wrongly stated "PC", shall be "PS" instead),   this is yet to be re-plotted also into the PDF.

@@
   \begin{tabular}{ c }
-  \includegraphics[width=\figwidthfull]{../img/populations_stoichiometry.pdf} 
+  \includegraphics[width=\figwidthfull]{../img/populations_stoichiometry_structures.pdf} 
   \end{tabular}
@@
         For numerical values, see table~\ref{tab:binding}.
-    \textbf{(B)} Relative probabilities of \ce{Ca^{2+}} ions to coordinate with a certain number of lipids
-    in pure POPC bilayer with 350 mM CaCl$_2$ and in POPC:POPS (5:1) mixture with 400 mM CaCl$_2$.  
+    \textbf{(B)} Relative probabilities of \ce{Ca^{2+}} ions to coordinate with the given number of lipids
+    in a pure POPC bilayer with 350~mM~CaCl$_2$ and in POPC:POPS (5:1) mixture with 400~mM~CaCl$_2$.  
     Analysis was done for both systems by considering all lipids (blue and violet) and
@@
     for the distance between the cation and the oxygen atoms of the lipids. 
+    \textbf{(C and D)} Representative example configurations of \ce{Ca^{2+}} coordinated complexes 
+    of one POPS with one POPC (C) or two POPC lipids (D). 
+    Extra configurations are given in Fig.~\ref{fig:strcutures_SI} in the SI. 
     Previously published simulation data \cite{melcr18} for pure POPC bilayers were taken directly from Ref.~\citenum{ECC-POPC_nacl_cacl2_files}. 

### 2019-10-16 Manuscript/manuscript.tex
Updated affiliation of JM

@@
 \email{melcr@marge.uochb.cas.cz} 
-%%\homepage[]{https://jmelcr.github.io/}
-\affiliation{Institute of Organic Chemistry and Biochemistry, 
-Czech Academy of Sciences,  
-Prague 6, Czech Republic} 
+\affiliation[Czech Academy of Sciences]{Institute of Organic Chemistry and Biochemistry of the 
+Czech Academy of Sciences, Flemingovo n\'{a}m. 542/2, CZ-16610 Prague 6, Czech Republic}
+\alsoaffiliation{Groningen Biomolecular Sciences and Biotechnology Institute 
+and The Zernike Institute for Advanced Materials, 
+University of Groningen, 9747 AG Groningen, The Netherlands}
+
 \author{Tiago Ferreira}

### 2019-10-16 Manuscript/manuscript.tex
Improved first paragraph in sec. Methods according to Rev1 comments.

@@
 Electronic continuum correction (ECC) is an implicit mean-field representation of 
-the electronic polarization in classical MD simulations. 
+the electronic polarization in classical MD simulations, 
+which has been recently applied to improve empirical models for ions and
+ biomolecules containing charged groups such as proteins and lipids. 
+\cite{melcr18, Pluharova2014, martinek17, duboue2018insulin, Mason2019, Duboue2018MgZn}
 For simple ions in water, the electronic polarizability
@@
 where  $\epsilon_{el} = 1.78$ is the high-frequency dielectric constant of electrons in water \cite{leontyev09}.
-The scaling of charges with this factor has been recently applied to improve empirical models for ions and
- biomolecules containing charged groups such as proteins\cite{Pluharova2014, martinek17, duboue2018insulin, Mason2019, Duboue2018MgZn}.
-In addition to the charge, the Lennard-Jones~$\sigma$ was scaled with a factor $ 0.72 < f_\sigma \leq 1$
-to improve the description of hydration properties of ions with respect to scattering data in these studies.
+In addition to the charge, the Lennard-Jones parameter~$\sigma$ of the ions needed to be modified 
+to improve the description of hydration properties of ions with respect to scattering data in these studies 
+by scaling with a factor $ 0.72 < f_\sigma \leq 1$.
 The reason for the small readjustment of the ionic radii is that the original force fields were often developed such that the first peak on the experimental ion-water oxygen radial distribution function was reproduced.
 Upon charge scaling this peak typically moved to slightly larger distances and the agreement with experiment was then restored by a small decrease of the Lennard-Jones~$\sigma$ parameter.
+More details about ECC are contained in references \citenum{melcr18, martinek17, leontyev09, leontyev11, leontyev14}.
 

### 2019-10-18 Manuscript/manuscript.tex

@@
 to the carboxylate than to the phosphate moiety of POPS headgroup,
-while binding to acyl chain carbonyls is almost negligible (Fig.~\ref{fig:cacl_complexes} and Table~\ref{tab:binding}).
+while binding only to acyl chain carbonyls is almost negligible (Fig.~\ref{fig:cacl_complexes} and Table~\ref{tab:binding}).
 The result is consistent with CHARMM36 simulations without the NBfix correction \cite{hallock18},
@@
 In our ECC-lipids simulation, Ca$^{2+}$ ions bind twice more likely to carboxylate groups
-of PS headgroups than to phosphate groups, while binding to carbonyls is almost negligible.
+of PS headgroups than to phosphate groups, while binding only to carbonyls is almost negligible.
 Binding of Ca$^{2+}$ ions to more than two POPS lipids simultaneously is very rare in \mbox{ECC-lipids} simulations.

### 2019-10-18 Manuscript/manuscript.tex
Merging text by Pavel to the methods

@@
 
-Electronic continuum correction (ECC) is an implicit mean-field representation of 
-the electronic polarization in classical MD simulations, 
-which has been recently applied to improve empirical models for ions and
- biomolecules containing charged groups such as proteins and lipids. 
-\cite{melcr18, Pluharova2014, martinek17, duboue2018insulin, Mason2019, Duboue2018MgZn}
-For simple ions in water, the electronic polarizability
-can be taken into account by scaling the charge with a constant factor 
-%\begin{equation}
- \mbox{$ f_q = \frac{1}{\sqrt{\epsilon_{el}}} \approx 0.75$,}
-%\end{equation}
-where  $\epsilon_{el} = 1.78$ is the high-frequency dielectric constant of electrons in water \cite{leontyev09}.
+%Electronic continuum correction (ECC) is an implicit mean-field representation of 
+%the electronic polarization in classical MD simulations, 
+%which has been recently applied to improve empirical models for ions and
+% biomolecules containing charged groups such as proteins and lipids. 
+%\cite{melcr18, Pluharova2014, martinek17, duboue2018insulin, Mason2019, Duboue2018MgZn}
+%For simple ions in water, the electronic polarizability
+%can be taken into account by scaling the charge with a constant factor 
+%%\begin{equation}
+% \mbox{$ f_q = \frac{1}{\sqrt{\epsilon_{el}}} \approx 0.75$,}
+%%\end{equation}
+%where  $\epsilon_{el} = 1.78$ is the high-frequency dielectric constant of electrons in water \cite{leontyev09}.
+
+Electrostatic interactions between charges in a dielectric medium are attenuated by a factor equal to the
+corresponding relative permittivity (dielectric constant). Standard non-polarizable force fields are capable to
+account for the dielectric response from the orientation of molecules (low-frequency), but fail to account for the electronic (high frequency) polarization.
+In most biologically relevant environments this high frequency dielectric constant is ~2 (for water the value is 1.78).
+This means that for a non-polarizable force field that would strictly account only for the low-frequency dielectric response
+interactions between charges in aqueous environments should be further attenuated by a factor of 1.78, which is mathematically
+equivalent by scaling the charges by a factor of ~0.75 \cite{leontyev09,leontyev14}. 
+This electronic continuum correction (ECC), which can be view as a mean-field continuum way for accounting for electronic polarization,
+has been shown to significantly improve the description of interactions between aqueous ions and biologically relevant charged groups \cite{pluharova13,kohagen14,kohagen14b,melcr18,xie16,kroutil17,bruce18}
+%[dx.doi.org/10.1021/jp402532e, dx.doi.org/10.1021/jp5005693, dx.doi.org/10.1021/jz502099g, DOI: 10.1021/acs.jpcb.7b12510, https://doi.org/10.1021/acs.jpcb.5b10755, DOI:10.1007/s00894-017-3490-x, https://doi.org/10.1063/1.5017101].
+This correction is particularly important for high charge density species where it helps to remove artifacts such as excessive ion pairing in water [dx.doi.org/10.1021/jp402532e, dx.doi.org/10.1021/jp310719g].
+One should, however, bear in mind that existing non-polarizable force fields have been parameterized also using experimental data, which means that they may to a certain extent include implicitly electronic polarization effects.
+Therefore, in such cases a charge scaling factor between 0.75 and 1 should carefully be adjusted such that electronic polarization effect are not overaccounted for.
+
 In addition to the charge, the Lennard-Jones parameter~$\sigma$ of the ions needed to be modified 

### 2019-10-18 Manuscript/manuscript.tex
Improvements in abstract

--- 2019-10-18 Manuscript/manuscript.tex
@@
   the Amber Lipid17 parameters of 1-palmitoyl-2-oleoyl-{\it sn}-glycero-3-phospho-L-serine (POPS) lipid to improve its
-  interactions with Na$^+$ and Ca$^{2+}$ ions.
+  interactions with \ce{K^+}, Na$^+$ and Ca$^{2+}$ ions.
   The partial charges of headgroup, glycerol backbone and carbonyls of POPS, bearing a unit negative charge, were scaled with a
@@
   when validated using headgroup order parameters and ''electrometer concept''.
-  In ECC-lipids simulations, Ca$^{2+}$ cations do not simultaneously interact with more than two PS lipids, and interactions 
-  with carboxylate groups is twice more likely than with phosphate groups, while interaction with carbonyls is almost negligible.
+  In ECC-lipids simulations, populations of complexes of Ca$^{2+}$ cations with more than two PS lipids are negligible, 
+and interactions of Ca$^{2+}$ cations with only carboxylate groups is twice more likely than with only phosphate groups, 
+while interactions with carbonyls almost entirely involve also other groups. 
   Our results pave the way for more realistic MD simulations of anionic biological membranes and demonstrate the 

### 2019-10-18 Manuscript/manuscript.tex
small edits in Methods-ECC

@@
 binding affinity of ions to negatively charged lipids is similar to that of zwitterionic lipids,
-and the binding affinity is increased only due to the increased cation
+and it increases only due to the increased cation
 concentration in the vicinity of the membrane \cite{seelig90,sinn06}.
@@
 Upon charge scaling this peak typically moved to slightly larger distances and the agreement with experiment was then restored by a small decrease of the Lennard-Jones~$\sigma$ parameter.
-More details about ECC are contained in references \citenum{melcr18, martinek17, leontyev09, leontyev11, leontyev14}.
+More detailed description of ECC and its applications are contained in references \citenum{melcr18, martinek17, Pluharova2014, leontyev09, leontyev11, leontyev14}.
 
@@
 
-Here, we apply the ECC to the Amber-based lipid17 parameters \cite{lipid17-future} (available in AmberTools18 \cite{amber18})
-of POPS lipid with a monovalent negative charge.
+Here, we apply the ECC to the Amber Lipid17 parameters of POPS \cite{lipid17-future} (available in AmberTools18 \cite{amber18}). 
 Because POPS is anionic at physiological pH $\approx 7$ carrying a physical total charge as aqueous ions, 
-we apply the theoretically derived scaling factor\cite{leontyev09} for
-ions $f_q = 0.75$ to the partial charges of POPS 
-in the headgroup, glycerol backbone, and carbonyl regions. The lipids with a total charge of -0.75 are also neutralized
+we apply the scaling factor derived for
+ions\cite{leontyev09} $f_q = 0.75$ to the partial charges of POPS 
+in the headgroup, glycerol backbone, and carbonyl regions. The POPS lipids with a total charge of -0.75 are also neutralized
 by the counterion charges of +0.75 in simulations with ECC-ions \cite{Pluharova2014, kohagen16, martinek17}.
@@
 Further optimization of parameters was not done in this work. 
-The generated ECC-POPS parameters are available from Ref.~\citenum{ecclipids_pcps_nacl_kcl_series}.
+The generated ECC-POPS parameters are openly available for download from Ref.~\citenum{ecclipids_pcps_nacl_kcl_series}.
 

### 2019-10-18 Manuscript/manuscript.tex
Small edits in simulation details

@@
 resolution of $\pm$0.01. The C--H bond order parameter determined from a given dipolar
-splitting, $\Delta\nu$ (see e.g. Fig. \ref{R-PDLFslices} in supplementary information) is equal to $\Delta\nu/(0.315\times21.5 kHz)$, where 0.315 is
+splitting, $\Delta\nu$ (see e.g. Fig.~\ref{R-PDLFslices} in supplementary information) is equal to $\Delta\nu/(0.315\times21.5 kHz)$, where 0.315 is
 the scaling factor of the R18 recoupling sequence and 21.5 kHz is the maximum $^1$H-$^{13}$C dipolar coupling for a C--H bond.
@@
 
-In the ECC-lipids simulations,
+In the ECC-lipids simulations, we used
 the ECC-POPC parameters~\cite{melcr18} (available from Ref. \citenum{ECC-POPC_nacl_cacl2_files}), 
 the ECC-POPS parameteres derived in this work (see above for details and Ref. \citenum{ecclipids_pcps_nacl_kcl_series} for parameters), 
-the SPC/E~\cite{Berendsen1987} water model, and ECC-ions \cite{martinek17, kohagen16, Pluharova2014} were used.
+the SPC/E~\cite{Berendsen1987} water model, and ECC-ions \cite{martinek17, kohagen16, Pluharova2014}. 
 The SPC/E water model was selected because its lower dielectric constant is consistent with the
@@
 
-The C--H bond order parameters were calculated directly from Eq.~\ref{OPequation}.
+The C--H bond order parameters were calculated directly from Eq.~\ref{OPequation}
+using a Python code openly available at Ref.~\citenum{MATCHgit} ({\tt scripts/calcOrderParameters.py}) 
+that uses the MDAnalysis library \cite{agrawal11,gowers16}.  
 The time average was first calculated for each lipid, and
@@
 error estimate \cite{botan15,ollila16,NMRlipidsIV}.
-Python program that uses the
-MDAnalysis library \cite{agrawal11,gowers16} is available in Ref.~\citenum{MATCHgit} ({\tt scripts/calcOrderParameters.py}). 
 The ion number density profiles were calculated using the {\tt gmx density} tool
 of the Gromacs sofware package \cite{gromacsMANUAL}.
-The salt concentrations in buffer before solvating the lipids, reported in the used experimental data set \cite{roux90},
+In simulations, the salt concentrations in a buffer before solvating the lipids, 
+as reported in the used experimental data set \cite{roux90},
 were calculated as $[{\rm salt}]=N_{\rm c} \times[{\rm water}]\,/\,N_{\rm w}$,
-where $N_{\rm c}$ is the number of cations in simulation, $[{\rm water}]$\,=\,55.5~M and $N_{\rm w}$
-is the number of water molecules in the simulation.
+where $N_{\rm c}$ is the number of additional cations, 
+$N_{\rm w}$ is the number of water molecules,
+and $[{\rm water}]$\,=\,55.5~M. 
 As discussed in our previous work \cite{NMRlipidsIV}, the hydration levels of

### 2019-10-18 Manuscript/manuscript.tex
Edits in results section

@@
 (Figs.~\ref{fig:delta_ordPar_monoval_PCPS} and \ref{fig:delta_ordPar_NaCl_PC-PS_mix_PS}).
-Similar results also for other force fields are reported elsewhere~\cite{NMRlipidsIV}.
+Similar results are reported also for other force fields. \cite{NMRlipidsIV}
 In ECC-lipids simulations, the response of POPS headgroup order parameters to both POPC and monovalent salt concentration (\ce{KCl} and \ce{NaCl})
@@
 by comparing the relative surface excesses with respect to water, $\Gamma ^{w} _{\rm ion}$,
-between POPC:POPS (5:1) mixture and pure POPC in ECC-lipids simulations with the most realistic
+between POPC:POPS (5:1) mixture and pure POPC in ECC-lipids simulations with currently the most realistic
 cation binding affinities~\cite{melcr18}.
@@
 indicating that the lower binding affinity than in Lipid17/Dang simulations is more realistic (Fig.~\ref{fig:cacl-dens_PCPS}).
-Therefore, we use ECC-lipids simulations to quantify the influence of negative charged POPS on
+Therefore, we use ECC-lipids simulations to quantify the influence of negatively charged POPS on
 calcium binding affinity to lipid bilayers.
@@
 for the POPC:POPS (5:1) bilayer with 409 mM CaCl$_2$.
-With these concentrations, the total amount of lipids per bound \ce{Ca^{2+}} is
+At these concentrations, the total amount of lipids per bound \ce{Ca^{2+}} is
 $\sim 5.4$ in the pure POPC system and $\sim 4.8$ in the POPC:POPS (5:1) mixture.
-Interestingly, the calcium residence times are 3-4 times longer in POPC:POPS (5:1) mixture (Fig.~\ref{fig:hist_residence_times}) than in pure POPC \cite{melcr18}.
+Also, the residence times of \ce{Ca^{2+}} at the membrane 
+are 3-4 times longer in POPC:POPS (5:1) mixture (Fig.~\ref{fig:hist_residence_times}) than in pure POPC \cite{melcr18}.
 Besides lower binding affinity, ECC-lipids simulations yield smaller error bars for order parameters and shorter residence times (Fig.~\ref{fig:hist_residence_times}) 
@@
     Extra configurations are given in Fig.~\ref{fig:strcutures_SI} in the SI. 
-    Previously published simulation data \cite{melcr18} for pure POPC bilayers were taken directly from Ref.~\citenum{ECC-POPC_nacl_cacl2_files}. 
+    Previously published simulation data for pure POPC bilayers \cite{melcr18} were taken directly from Ref.~\citenum{ECC-POPC_nacl_cacl2_files}. 
   }
@@
 Recent analyzes of interaction sites between calcium and PS lipids combining
-MD simulations with various experimental techniques have been 
-controversial because the results strongly depend on force fields parameters \cite{melcrova16,valentine18,hallock18}.
-Here, we analyze the calcium binding details from ECC-lipids simulation of POPC:POPS (5:1) mixture with 409 mM CaCl$_2$,
-because it surpasses the quality of other force fields in direct comparison with experimental order parameter data.
+MD simulations with various experimental techniques remain controversial 
+as the results strongly depend on force fields parameters \cite{melcrova16,valentine18,hallock18, NMRlipidsIV}.
+Here, we analyze the calcium binding details of POPC:POPS (5:1) mixture with 409 mM CaCl$_2$ from ECC-lipids simulations,
+which in this respect surpass the quality of other force fields in direct comparison with experimental order parameters.
 In the ECC-lipid simulation, calcium ions binds approximately twice more likely 

### 2019-10-18 Manuscript/manuscript.tex
edits in conclustions

@@
 when validated using the headgroup order parameters and ``electrometer concept''~\cite{NMRlipidsIV}.
-There is, however, room for improvement in capturing the sensitive headgroup responses to the
+There is, however, still room for improvement in capturing the sensitive headgroup responses to 
 small concentrations of CaCl$_2$.
@@
 In our ECC-lipids simulation, Ca$^{2+}$ ions bind twice more likely to carboxylate groups
-of PS headgroups than to phosphate groups, while binding only to carbonyls is almost negligible.
+of PS headgroups than to phosphate groups, 
+while binding only to carbonyls is negligible and almost entirely involves also other groups. 
 Binding of Ca$^{2+}$ ions to more than two POPS lipids simultaneously is very rare in \mbox{ECC-lipids} simulations.
@@
 we believe that our results help in resolving controversial interpretations of more indirect experiments
-using different force fields~\cite{melcrova16,valentine18,hallock18}.
-Furthermore, our results pave the way for more realistic MD simulations of anionic biological membranes,
-and demonstrate the usefulness of ECC also for charged lipids.
+using different force fields~\cite{melcrova16,valentine18,hallock18}
+and demonstrate the usefulness of ECC also for membranes with charged lipids.
+Furthermore, the present simulations pave the way for more realistic MD simulations of anionic biological membranes
+allowing to elucidate biological signalling processes involving PS and \ce{Ca^{2+}}.
 
 The general philosophy behind our approach is to build upon existing non-polarizable force fields making the minimum of necessary changes by scaling only groups bearing a sizeable (partial) charge, followed by slight readjustments of van der Waals radii and possibly also other parameters such as those concerning dihedral angle terms.
-The present work thus represents another piece in the mosaic that leads to development of an accurate empirical force field for biomolecular simulations involving aqueous salt solutions, proteins, peptides, lipid membranes, and/or nucleic acids, which include electronic polarization effects in a mean field way via charge scaling without extra computational costs compared to standard non-polarizable simulations.
-We believe that ECC corrected lipid force fields will be useful in modelling complex biomembranes when 
-explicitly polarizable force fields for lipids \cite{li17,chu18} are not necessary.
+The present work thus represents another piece in the mosaic that leads to development of an improved empirical force field for biomolecular simulations involving aqueous salt solutions, proteins, peptides, lipid membranes, and/or nucleic acids, which include electronic polarization effects in a mean field way via charge scaling without extra computational costs compared to standard non-polarizable simulations.
+We believe that such ECC corrected force fields will be useful in modeling complex biomembranes when 
+explicitly polarizable force fields for biomolecules \cite{amoeba13,li17,chu18} are not necessary.
 

### 2019-10-18 Manuscript/manuscript.tex
Improved formulation in conclusion and absract and making them consistent

@@
 while interactions with carbonyls almost entirely involve also other groups. 
-  Our results pave the way for more realistic MD simulations of anionic biological membranes and demonstrate the 
-  usefulness of ECC also to charged lipids.
+  Our results pave the way for more realistic MD simulations of biomolecular systems with anionic membranes
+allowing to elucidate signaling processes involving PS and \ce{Ca^{2+}}. 
 \end{abstract}
@@
 and demonstrate the usefulness of ECC also for membranes with charged lipids.
-Furthermore, the present simulations pave the way for more realistic MD simulations of anionic biological membranes
-allowing to elucidate biological signalling processes involving PS and \ce{Ca^{2+}}.
+Furthermore, the present simulations pave the way
+for more realistic MD simulations of biomolecular systems with anionic membranes
+allowing to elucidate signaling processes involving PS and \ce{Ca^{2+}}.
 

### 2019-10-20 Manuscript/manuscript.tex
Update text

@@
 This correction is particularly important for high charge density species where it helps to remove artifacts such as excessive ion pairing in water [dx.doi.org/10.1021/jp402532e, dx.doi.org/10.1021/jp310719g].
-One should, however, bear in mind that existing non-polarizable force fields have been parameterized also using experimental data, which means that they may to a certain extent include implicitly electronic polarization effects.
-Therefore, in such cases a charge scaling factor between 0.75 and 1 should carefully be adjusted such that electronic polarization effect are not overaccounted for.
-
+One should, however, bear in mind that existing non-polarizable force fields have been parameterized also using experimental data, which means that they may to a certain extent include implicitly electronic polaization effects.
+Therefore, in such cases a charge scaling factor between 0.75 and 1 should carefully be adjusted such that electronic polarization effect are not overaccounted.
 In addition to the charge, the Lennard-Jones parameter~$\sigma$ of the ions needed to be modified 
@@
 Upon charge scaling this peak typically moved to slightly larger distances and the agreement with experiment was then restored by a small decrease of the Lennard-Jones~$\sigma$ parameter.
-More detailed description of ECC and its applications are contained in references \citenum{melcr18, martinek17, Pluharova2014, leontyev09, leontyev11, leontyev14}.
+%More detailed description of ECC and its applications are contained in references \citenum{melcr18, martinek17, Pluharova2014, leontyev09, leontyev11, leontyev14}.
 
-Derivation of the correct scaling factor for lipids is more complicated, because
-the partial charges depend on the methods used to derive the respective force fields.
-For example, the partial charges of overall neutral molecules may already include the effects of electronic polarizability to some extent,
-and hence, the scaling factor can be larger than the theoretically derived value for ions in water, $f_q \approx 0.75$.  
+%Derivation of the correct scaling factor for lipids is more complicated, because
+%the partial charges depend on the methods used to derive the respective force fields.
+%For example, the partial charges of overall neutral molecules may already include the effects of electronic polarizability to some extent,
+%and hence, the scaling factor can be larger than the theoretically derived value for ions in water, $f_q \approx 0.75$.  
 Before the ECC theory was rigorously derived, similar idea was already employed in early classical 
-MD simulations of lipids and surfactants using the scaling factor of 0.5 for atomic charges  \cite{jonsson86,egberts94, berendsen1996}. 
+MD simulations of lipids and surfactants using the ad hoc scaling factor of 0.5 for atomic charges  \cite{jonsson86,egberts94, berendsen1996}. 
 In our recent work, we applied the ECC to the Amber based lipid14 force field of zwitterionic POPC lipid \cite{dickson14}

### 2019-10-21 Manuscript/manuscript.tex
Update text

@@
 %[dx.doi.org/10.1021/jp402532e, dx.doi.org/10.1021/jp5005693, dx.doi.org/10.1021/jz502099g, DOI: 10.1021/acs.jpcb.7b12510, https://doi.org/10.1021/acs.jpcb.5b10755, DOI:10.1007/s00894-017-3490-x, https://doi.org/10.1063/1.5017101].
-This correction is particularly important for high charge density species where it helps to remove artifacts such as excessive ion pairing in water [dx.doi.org/10.1021/jp402532e, dx.doi.org/10.1021/jp310719g].
+This correction is particularly important for high charge density species where it helps to remove artifacts such as excessive ion pairing in water~\cite{pluharova13,vazdar13}.
+%[dx.doi.org/10.1021/jp402532e, dx.doi.org/10.1021/jp310719g].
 One should, however, bear in mind that existing non-polarizable force fields have been parameterized also using experimental data, which means that they may to a certain extent include implicitly electronic polaization effects.
@@
 decrease proportionally to the amount of bound positive charge because the headgroup dipole tilts more
-parallel to the mebrane normal.
+parallel to the membrane normal.
 Although the response of PS lipid headgroup order parameters to the bound charge is also systematic,

### 2019-10-21 Manuscript/manuscript.tex
Update text

@@
 binding affinity of ions to negatively charged lipids is similar to that of zwitterionic lipids,
-and it increases only due to the increased cation
+and the amount of bound ions increases only due to the increased cation
 concentration in the vicinity of the membrane \cite{seelig90,sinn06}.
@@
 corresponding relative permittivity (dielectric constant). Standard non-polarizable force fields are capable to
-account for the dielectric response from the orientation of molecules (low-frequency), but fail to account for the electronic (high frequency) polarization.
+account for the dielectric response from the orientation of the solvent molecules (low-frequency nuclear response), but fail to account for the electronic (high frequency) polarization.
 In most biologically relevant environments this high frequency dielectric constant is ~2 (for water the value is 1.78).
@@
 of the Gromacs sofware package \cite{gromacsMANUAL}.
-In simulations, the salt concentrations in a buffer before solvating the lipids, 
+The salt concentrations in a buffer before solvating the lipids, 
 as reported in the used experimental data set \cite{roux90},
+%is simulations
 were calculated as $[{\rm salt}]=N_{\rm c} \times[{\rm water}]\,/\,N_{\rm w}$,
@@
     \textbf{(C and D)} Representative example configurations of \ce{Ca^{2+}} coordinated complexes 
-    of one POPS with one POPC (C) or two POPC lipids (D). 
-    Extra configurations are given in Fig.~\ref{fig:strcutures_SI} in the SI. 
+    of one POPS with one (C) or two POPC lipids (D). 
+    More configurations are shown in Fig.~\ref{fig:strcutures_SI} in the SI. 
     Previously published simulation data for pure POPC bilayers \cite{melcr18} were taken directly from Ref.~\citenum{ECC-POPC_nacl_cacl2_files}. 
@@
 Here, we analyze the calcium binding details of POPC:POPS (5:1) mixture with 409 mM CaCl$_2$ from ECC-lipids simulations,
-which in this respect surpass the quality of other force fields in direct comparison with experimental order parameters.
+which surpass the quality of other force fields in direct comparison with experimental order parameters.
 In the ECC-lipid simulation, calcium ions binds approximately twice more likely 
@@
 In our ECC-lipids simulation, Ca$^{2+}$ ions bind twice more likely to carboxylate groups
-of PS headgroups than to phosphate groups, 
-while binding only to carbonyls is negligible and almost entirely involves also other groups. 
+of PS headgroups than to phosphate groups, while binding only to carbonyls is negligible.
+However, calcium can bind also to carbonyls when simultaneously bound also to carboxylate (Fig. \ref{fig:strcutures_SI}) or phosphate group.
 Binding of Ca$^{2+}$ ions to more than two POPS lipids simultaneously is very rare in \mbox{ECC-lipids} simulations.

### 2019-10-28 Manuscript/manuscript.tex
Updated manuscript PDFs including also TOC graphic

--- 2019-10-20 Manuscript/manuscript.tex
@@
 
-Some journals require a graphical entry for the Table of Contents. 
-This should be laid out ``print ready'' so that the sizing of the 
-text is correct.
+\includegraphics[height=1.4 in]{../img/TOC_graphic.pdf}
 
-Inside the \texttt{tocentry} environment, the font used is Helvetica 
-8\,pt, as required by \emph{Journal of the American Chemical 
-Society}.
-
-The surrounding frame is 9\,cm by 3.5\,cm, which is the maximum 
-permitted for  \emph{Journal of the American Chemical Society} 
-graphical table of content entries. The box will not resize if the 
-content is too big: instead it will overflow the edge of the box.
-
-This box and the associated title will always be printed on a 
-separate page at the end of the document.
+For table of contents only.
 
@@
 
-%Electronic continuum correction (ECC) is an implicit mean-field representation of 
-%the electronic polarization in classical MD simulations, 
-%which has been recently applied to improve empirical models for ions and
-% biomolecules containing charged groups such as proteins and lipids. 
-%\cite{melcr18, Pluharova2014, martinek17, duboue2018insulin, Mason2019, Duboue2018MgZn}
-%For simple ions in water, the electronic polarizability
-%can be taken into account by scaling the charge with a constant factor 
-%%\begin{equation}
-% \mbox{$ f_q = \frac{1}{\sqrt{\epsilon_{el}}} \approx 0.75$,}
-%%\end{equation}
-%where  $\epsilon_{el} = 1.78$ is the high-frequency dielectric constant of electrons in water \cite{leontyev09}.
-
-Electrostatic interactions between charges in a dielectric medium are attenuated by a factor equal to the
-corresponding relative permittivity (dielectric constant). Standard non-polarizable force fields are capable to
-account for the dielectric response from the orientation of molecules (low-frequency), but fail to account for the electronic (high frequency) polarization.
-In most biologically relevant environments this high frequency dielectric constant is ~2 (for water the value is 1.78).
-This means that for a non-polarizable force field that would strictly account only for the low-frequency dielectric response
-interactions between charges in aqueous environments should be further attenuated by a factor of 1.78, which is mathematically
-equivalent by scaling the charges by a factor of ~0.75 \cite{leontyev09,leontyev14}. 
-This electronic continuum correction (ECC), which can be view as a mean-field continuum way for accounting for electronic polarization,
-has been shown to significantly improve the description of interactions between aqueous ions and biologically relevant charged groups \cite{pluharova13,kohagen14,kohagen14b,melcr18,xie16,kroutil17,bruce18}
-%[dx.doi.org/10.1021/jp402532e, dx.doi.org/10.1021/jp5005693, dx.doi.org/10.1021/jz502099g, DOI: 10.1021/acs.jpcb.7b12510, https://doi.org/10.1021/acs.jpcb.5b10755, DOI:10.1007/s00894-017-3490-x, https://doi.org/10.1063/1.5017101].
-This correction is particularly important for high charge density species where it helps to remove artifacts such as excessive ion pairing in water [dx.doi.org/10.1021/jp402532e, dx.doi.org/10.1021/jp310719g].
-One should, however, bear in mind that existing non-polarizable force fields have been parameterized also using experimental data, which means that they may to a certain extent include implicitly electronic polaization effects.
-Therefore, in such cases a charge scaling factor between 0.75 and 1 should carefully be adjusted such that electronic polarization effect are not overaccounted.
+Electronic continuum correction (ECC) is an implicit mean-field representation of 
+the electronic polarization in classical MD simulations, 
+which has been recently applied to improve empirical models for ions and
+ biomolecules containing charged groups such as proteins and lipids. 
+\cite{melcr18, Pluharova2014, martinek17, duboue2018insulin, Mason2019, Duboue2018MgZn}
+For simple ions in water, the electronic polarizability
+can be taken into account by scaling the charge with a constant factor 
+%\begin{equation}
+ \mbox{$ f_q = \frac{1}{\sqrt{\epsilon_{el}}} \approx 0.75$,}
+%\end{equation}
+where  $\epsilon_{el} = 1.78$ is the high-frequency dielectric constant of electrons in water \cite{leontyev09}.
 In addition to the charge, the Lennard-Jones parameter~$\sigma$ of the ions needed to be modified 
@@
 Upon charge scaling this peak typically moved to slightly larger distances and the agreement with experiment was then restored by a small decrease of the Lennard-Jones~$\sigma$ parameter.
-%More detailed description of ECC and its applications are contained in references \citenum{melcr18, martinek17, Pluharova2014, leontyev09, leontyev11, leontyev14}.
+More detailed description of ECC and its applications are contained in references \citenum{melcr18, martinek17, Pluharova2014, leontyev09, leontyev11, leontyev14}.
 
-%Derivation of the correct scaling factor for lipids is more complicated, because
-%the partial charges depend on the methods used to derive the respective force fields.
-%For example, the partial charges of overall neutral molecules may already include the effects of electronic polarizability to some extent,
-%and hence, the scaling factor can be larger than the theoretically derived value for ions in water, $f_q \approx 0.75$.  
+Derivation of the correct scaling factor for lipids is more complicated, because
+the partial charges depend on the methods used to derive the respective force fields.
+For example, the partial charges of overall neutral molecules may already include the effects of electronic polarizability to some extent,
+and hence, the scaling factor can be larger than the theoretically derived value for ions in water, $f_q \approx 0.75$.  
 Before the ECC theory was rigorously derived, similar idea was already employed in early classical 
-MD simulations of lipids and surfactants using the ad hoc scaling factor of 0.5 for atomic charges  \cite{jonsson86,egberts94, berendsen1996}. 
+MD simulations of lipids and surfactants using the scaling factor of 0.5 for atomic charges  \cite{jonsson86,egberts94, berendsen1996}. 
 In our recent work, we applied the ECC to the Amber based lipid14 force field of zwitterionic POPC lipid \cite{dickson14}

### 2019-10-31 Manuscript/manuscript.tex
Update text

@@
 
-\includegraphics[height=1.4 in]{../img/TOC_graphic.pdf}
-
-For table of contents only.
+\includegraphics[height=1.375 in]{../img/TOC_graphic.pdf}
 
@@
 Recent analyzes of interaction sites between calcium and PS lipids combining
-MD simulations with various experimental techniques remain controversial 
-as the results strongly depend on force fields parameters \cite{melcrova16,valentine18,hallock18, NMRlipidsIV}.
+MD simulations with various experimental techniques have been controversial 
+because the results strongly depend on force fields parameters \cite{melcrova16,valentine18,hallock18, NMRlipidsIV}.
 Here, we analyze the calcium binding details of POPC:POPS (5:1) mixture with 409 mM CaCl$_2$ from ECC-lipids simulations,
@@
 \begin{itemize} 
-  \item SI.pdf: Additional simulation and experimental details, figures and tables. 
+\item SI.pdf: Measured 1D and 2D NMR spectra, additional details and results from simulations
+  about monovalent ion binding, configurations and populations of bound calcium to POPS lipids, and
+  residence times of bound calcium.
 \end{itemize}

### 2019-11-06 Manuscript/manuscript.tex
Update to the submitted revised version

@@
   The resulting ECC-POPS model gives more realistic interactions with Na$^+$ and Ca$^{2+}$ cations than the original Amber Lipid17 parameters,
-  when validated using headgroup order parameters and ''electrometer concept''.
+  when validated using headgroup order parameters and the ''electrometer concept''.
   In ECC-lipids simulations, populations of complexes of Ca$^{2+}$ cations with more than two PS lipids are negligible, 
@@
 binding affinity of ions to negatively charged lipids is similar to that of zwitterionic lipids,
-and the amount of bound ions increases only due to the increased cation
+and that the amount of bound ions increases only due to the increased cation
 concentration in the vicinity of the membrane \cite{seelig90,sinn06}.
@@
 correction (ECC) \cite{leontyev09}. The good agreement between the resulting ECC-POPC
-model and experiments enable detailed interpretation of calcium binding details to a POPC lipid bilayer.
+model and experiments enables a more detailed molecular interpretation of the calcium binding to a POPC lipid bilayer.
 
@@
 
-%Electronic continuum correction (ECC) is an implicit mean-field representation of 
-%the electronic polarization in classical MD simulations, 
-%which has been recently applied to improve empirical models for ions and
-% biomolecules containing charged groups such as proteins and lipids. 
-%\cite{melcr18, Pluharova2014, martinek17, duboue2018insulin, Mason2019, Duboue2018MgZn}
-%For simple ions in water, the electronic polarizability
-%can be taken into account by scaling the charge with a constant factor 
-%%\begin{equation}
-% \mbox{$ f_q = \frac{1}{\sqrt{\epsilon_{el}}} \approx 0.75$,}
-%%\end{equation}
-%where  $\epsilon_{el} = 1.78$ is the high-frequency dielectric constant of electrons in water \cite{leontyev09}.
-
 Electrostatic interactions between charges in a dielectric medium are attenuated by a factor equal to the
@@
 account for the dielectric response from the orientation of the solvent molecules (low-frequency nuclear response), but fail to account for the electronic (high frequency) polarization.
-In most biologically relevant environments this high frequency dielectric constant is ~2 (for water the value is 1.78).
+In most biologically relevant environments this high frequency dielectric constant is approximately 2 (for water the value is 1.78).
 This means that for a non-polarizable force field that would strictly account only for the low-frequency dielectric response
@@
 has been shown to significantly improve the description of interactions between aqueous ions and biologically relevant charged groups \cite{pluharova13,kohagen14,kohagen14b,melcr18,xie16,kroutil17,bruce18}
-%[dx.doi.org/10.1021/jp402532e, dx.doi.org/10.1021/jp5005693, dx.doi.org/10.1021/jz502099g, DOI: 10.1021/acs.jpcb.7b12510, https://doi.org/10.1021/acs.jpcb.5b10755, DOI:10.1007/s00894-017-3490-x, https://doi.org/10.1063/1.5017101].
 This correction is particularly important for high charge density species where it helps to remove artifacts such as excessive ion pairing in water~\cite{pluharova13,vazdar13}.
-%[dx.doi.org/10.1021/jp402532e, dx.doi.org/10.1021/jp310719g].
-One should, however, bear in mind that existing non-polarizable force fields have been parameterized also using experimental data, which means that they may to a certain extent include implicitly electronic polaization effects.
-Therefore, in such cases a charge scaling factor between 0.75 and 1 should carefully be adjusted such that electronic polarization effect are not overaccounted.
+One should, however, bear in mind that existing non-polarizable force fields have been parameterized also using experimental data, which means that they may to a certain extent include implicitly electronic polarization effects.
+Therefore, in such cases a charge scaling factor between 0.75 and 1 should carefully be adjusted such that the electronic polarization effects are not overaccounted.
 In addition to the charge, the Lennard-Jones parameter~$\sigma$ of the ions needed to be modified 
@@
 Upon charge scaling this peak typically moved to slightly larger distances and the agreement with experiment was then restored by a small decrease of the Lennard-Jones~$\sigma$ parameter.
-%More detailed description of ECC and its applications are contained in references \citenum{melcr18, martinek17, Pluharova2014, leontyev09, leontyev11, leontyev14}.
 
-%Derivation of the correct scaling factor for lipids is more complicated, because
-%the partial charges depend on the methods used to derive the respective force fields.
-%For example, the partial charges of overall neutral molecules may already include the effects of electronic polarizability to some extent,
-%and hence, the scaling factor can be larger than the theoretically derived value for ions in water, $f_q \approx 0.75$.  
 Before the ECC theory was rigorously derived, similar idea was already employed in early classical 
@@
 Here, we measure also the C--H bond order parameters of the acyl chain region from multi-lamellar POPS vesicles using $^1$H--$^{13}$C
-dipolar recoupling 2D NMR experiment. As described in the Methods section, we performed R-PDLF 2D spectroscopy~\cite{dvinskikh04} to
+dipolar recoupling 2D NMR spectroscopy. As described in the Methods section, we performed R-PDLF 2D spectroscopy~\cite{dvinskikh04} to
 obtain a spectra correlating chemical shifts (direct dimension) and C--H bond dipolar couplings (indirect dimension) from which the C--H bond order parameters defined in eq.~\ref{OPequation} can be determined.
@@
 smaller than the value reported from SDP model \cite{kucerka14} (Fig.~\ref{simVSexpNOions_POPS}), but the values
-agree within their error estimates. The larger area in ECC-POPS compared to Lipid17/Dang simulation can be
+agree within their error estimates. Both the smaller order parameters and larger area in ECC-POPS compared to Lipid17/Dang simulation can be
 explained by increased headgroup repulsion due to lower counterion binding affinity (Fig. \ref{fig:POPS-counterions-dens}).
@@
 The present work thus represents another piece in the mosaic that leads to development of an improved empirical force field for biomolecular simulations involving aqueous salt solutions, proteins, peptides, lipid membranes, and/or nucleic acids, which include electronic polarization effects in a mean field way via charge scaling without extra computational costs compared to standard non-polarizable simulations.
-We believe that such ECC corrected force fields will be useful in modeling complex biomembranes when 
-explicitly polarizable force fields for biomolecules \cite{amoeba13,li17,chu18} are not necessary.
+We believe that such ECC force fields will be useful in modeling complex biomembranes when simulations with
+explicitly polarizable force fields for biomolecules \cite{amoeba13,li17,chu18} are not practical.
 
